\section{Christmas Message (2007)}

What message should I give for Him Who is the Eternal One?

He as God alone exists. There is only One Who exists eternally; not many, but One. No one else exists. Whatever appears as existing exists in non-existence. Manyness comes because of the dream, that is, ignorance. It is the shadow of Truth.

So what is this Creation?

Creation is that which was created in order for God to know the answer to the question, ``Who am I?''

When God was in the Beyond-Beyond state, He had neither consciousness nor unconsciousness. Suddenly, the Whim came: ``Who am I?''

And what happened because of that?

As soon as that Whim, ``Who am I?'' came in the Infinite Ocean, innumerable drops emerged from it to find out who they were. First, they entered into the mental sphere, then the subtle sphere and finally the gross sphere. So we have come down with the gross, subtle and mental worlds within us. But the drops had no consciousness. The Whim was in every drop, but actually, the drops were not there. It was the Ocean.

But for the Ocean to know itself as the Ocean, the drop has to experience the dream. In this dream, whatever you see yourself as being, you believe that you are that.

Mind was created, and energy was created, to play their roles in Creation in order to give the reply to the question, ``Who am I?''

And because of that, mind (which is illusory and within the drop) starts working.

In the gross sphere, there are seven kingdoms: gas-stone-metal, vegetable, worm and insect, fish; bird, animal and human being.

Mind and energy do not develop in the stone kingdom. But in the vegetable kingdom, we find energy starting to develop, because a vegetable needs water, manure, light and air. Then comes the kingdom of worms and insects. Worms have movement. If you try to catch one, it runs away.

Why?

Because mind is developing. In this way, energy and mind go on developing and developing in the fish, bird, animal and human kingdom.

Actually, the Ocean is every form in each kingdom, but to the drop, all forms look like drops. The Ocean asks, ``Who am I?''

And each drop says, ``I am a stone, I am a vegetable, I am a worm, I am a fish, I am a bird, I am an animal. I am a human being.''

If you try to count the species and varieties in every kingdom, you will find that they are innumerable. But this is all a dream.

So you may say, ``God has a dream in order to find out the Real Answer to the Real Question, ``Who am I?''

After ages and ages and ages of passing through all the kingdoms of the gross sphere, you come into human form. The impressions of the last gorilla form come with you, and in human form, reincarnation as a human being starts. Mind and energy are fully developed. You have passed through all the lower kingdoms, and now you experience gross consciousness in the gross sphere. You take the form of a man, the form of a woman. Sometimes you are born poor, sometimes rich. Sometimes you are a king, sometimes beggar. Sometimes you are in this country, sometimes in another. It goes on and on.

You say, ``I am this, or I am that.'' But you realize that these are not the Real Answer to the Whim, the question, ``Who am I?''

Earth is very close to the Creation Point. Though the gross sphere is the seventh shadow of the Truth, God, this seventh shadow is infinite. In it, there are different planets. On some, there is only stone; on others, there are stones and vegetables and on others, stones, vegetables and insects.

And on 18,000 of these planets, human beings live. However, on these planets, there is only mental development, no development of the heart. The earth is the only planet upon which human beings can get the Real Answer to the Whim, ``Who am I?''

After you achieve human form, it takes ages to enter in the subtle sphere. You pass through eight billion four hundred thousand births and deaths before your progress enough to enter into the subtle sphere. When you enter into the subtle sphere, you forget the gross sphere and experience energy.

Every night when we sleep, we dream. Sometimes we dream that we are kings and have big kingdoms. When we dream this, we enjoy! Sometimes we dream that we are beggars. And then we suffer.

But happiness and suffering and misery are all just a dream. They are not real.

The dream world that we experience when we sleep is the first part of the first plane of the subtle sphere. We pass through it whenever we sleep. At the time, we don't know that we are dreaming. But when we wake up, then we know that it was a dream.

And what are we passing through for ages and ages?

Nothing but the dream, Illusion. God alone exists. Nothing exists besides God. But we have to experience Illusion in order to find out the Real Answer to the Whim, ``Who am I?''

In the second plane of the subtle sphere are the hell and heaven states. They are but dreams into a dream. When we collect good impressions, we feel happy. When we drop our body, we go into either heaven or hell. If we have more good impressions than bad, then we enjoy the heaven state, in which the body is not there, but the soul, mind and energy are.

Say, for example, you enjoy liquor. When the gross medium of the body is no longer there, the impressions of liquor are still with you. When you are in heaven and a thought of liquor comes, you have thousands of times more enjoyment than when you are in a gross form.

In this way, in the heaven state, you spend good impressions quickly. But what happens?

Suppose you have one billion good impressions and have spent all but sixty million of them. You also have forty million bad impressions. You are nearing equality, or a balance of good and bad impressions. If good impressions and bad impressions equalize, you get liberation. But this never happens. Before you come to the point of equality, you become free and come into the body again. You take birth, and now you bring fifty million good and forty million bad impressions with you.

So in this life, we are spending our past impressions, and while we are, we create new ones. But we don't spend the impressions that we create in this birth. They are for the next life. This goes on and on. We spend the grossness of these good and bad impressions until equality comes.

Then we enter into the subtle sphere. Because of gross consciousness, you see the world and time and space. But as soon as you enter into the subtle world, there is no time and no space. We enter into subtle sphere when we don't create new impressions and are only spending past ones. Still, we see the dream. The dream continues, but it becomes subtle, and in the subtle world, we experience energy.

Anyone in the subtle world is not attached to the gross sphere. In the subtle world you have only subtle impressions. In this world, there are only subtle thoughts and subtle desires. Any action you perform is not gross, but subtle.

From the subtle sphere, we enter into the mental sphere, where there are only mental thoughts, not subtle or gross ones. If you are in the mental planes, you remain mentally conscious. You have no attachment to the gross or subtle. You are completely detached.

Those in the subtle and mental spheres may eat, but they cannot create gross thoughts. They remain detached from the physical body. Someone in the mental world knows the thoughts of everyone, even in different countries. When they sit here in India, they see the thoughts of everyone in the world. They do not have to remember them; they just know.

There was a mast here at Meherabad, Mohammed. Baba had taken him to France, and when he returned, Mohammed would say, ``I want to go to France! Take me to France!''

He went on saying this until the Second World War took place. Nobody can understand what work Baba had given to him during this war.

After his return from France with Baba, Mohammed would often say about a servant who took care of him, ``I want to marry that one. I want to marry her.''

Every day, he would catch hold of her hand, saying, ``I will marry you.''

Sometimes he would not let go of her hand for hours together.

People would call Mohammed ``Deva,'' and at the time, Sidhu was looking after him. Sidhu would say, ``Deva, leave her, because her children are hungry. She has to cook food for them.''

With great difficulty, Mohammed would release her hand and say, ``Go and cook food for your children.''

This would go on every day. When Baba would go to see him, Mohammed would still hold on to her hand, saying, ``I want to marry her.''

Baba would smile and laugh. ``Allow her to go,'' Baba would tell Mohammed. ``Her children are very small, and they must be waiting for food.''

Then Mohammed would release her.

Mohammed was a mast of the fifth plane, the mental plane, so he did not have gross thoughts or gross desires. Those on the mental planes do not have such thoughts or desires, and they cannot do any physical action. Their actions are non-action.

The sixth plane of the mental sphere is the plane where you see only the effulgence, the Light, of God. When you sees see this, you want to become One with the Light. Your longing increases infinitely.

As soon as you cross the mental plane into the seventh plane, mind is annihilated. Then you know Who you are. The ``I am God'' state is experienced. You experience All-Knowledge, All-Power and All-Bliss. That is the goal of life.

So Creation is there in order to for each drop of the Infinite Ocean to experience God.

There are always five Perfect Masters who have-God consciousness as well as consciousness of Creation. This is under the Divine Plan and follows Divine Law. But they cannot give a Universal Push to the world or anyone in it. So when consciousness in the gross sphere remains suspended, the Ancient One has to come down upon earth. Age after age, He comes down, and He gives a Universal Push, a push to all the seven kingdoms from stone to human being.

You must have read the story about Adam and Eve. Both were in Paradise, and there was a particular tree, the fruit of which was prohibited. No one should eat it. But one day, what happened?

The serpent persuaded Eve to pluck the fruit from the prohibited tree, and she gave that fruit to Adam. Adam ate it, and when he did, God got very angry, and threw them out of Paradise.

What a beautiful story it is. It is about the beginning of Creation and how it was created. Who was Adam?

Adam was the first one you read about in GOD SPEAKS: God was neither conscious nor unconscious. Then He got the Whim, ``Who am I?''

That is why God in that state started searching, why Creation took place and why evolution and involution began.

Adam and Eve were God. Eve was the shadow of God, and Adam was God Himself. People say, ``He was a very bad man. He ate the fruit from a prohibited tree!''

But no one knows the essence of this story. Adam is the Ancient One who comes down on earth, age after age, to help the universe. In every religion His Name is different, His face is different, and the place where He comes is different, but He is One and the same Adam Who comes down on earth age after age. He is the Saviour of Creation.

Another story given by Baba was entitled, ``The Mischievous Chicken.''

He also told different stories about the Creation. You can read them in THE NOTHING AND THE EVERYTHING.

God is not different. God is infinitely One, indivisible. God cannot be divided -- it is mind that divides the Ocean into every drop, that is, everyone and everything in the Creation.

Because the Whim, ``Who am I?,'' is always there, each drop is searching for God unconsciously. One day, after crossing the seven kingdoms of the evolution of consciousness and six kingdoms of the involution of consciousness, you reach the seventh plane. Mind is annihilated, and you know Who you are.

The Divine Plan and Divine Law are under Him, and He always remains in Paradise, having Infinite Knowledge, Infinite Power and Infinite Bliss. He lives on the seventh plane, and though He does not have Creation consciousness except when He comes down upon the earth, He remains eternally connected with each individual eternally. He is duty bound to take everyone and everything towards God.

When Jesus said, ``My father and I are One,'' nobody understood Him because the situation then was quite different. During His crucifixion, He said, ``O Father, forgive them all, who know not what they are doing.''

He would always tell people, ``I am the Son of My Father.'' At the time people, were so engrossed in rituals and ceremonies that even the Ancient One would not utter before their ignorance, ``I am God.''

The same thing happened in the time of Prophet Mohammed. He said, ``I am the Messenger of God.''

But God is infinite and indivisible. He is All-Eternal. It is God Who alone exists, and nothing exists besides God. Whatever appears as existing exists in non-existence.

And now again, I say about Lord Jesus:

He took birth, and He was crucified for the world. He is birthless and He is deathless, but when He comes down on earth, His birth is for the universe. When His work is completed in the gross sphere, He drops the body. He is the only Sustainer, and He lives in everyone and everything. But when ignorance goes on increasing and increasing, He has to come down on earth in order to lead human beings this from ignorance and push the progress of consciousness toward Reality.

What message should I give to you about Lord Jesus?

We need His help all the time, and we always must seek His help to find Him in His true form. He is always the same: Zoroaster, Ram, Krishna, Buddha, Jesus, Prophet Mohammed and now Meher Baba. This is the end of the cycle of cycles, and Meher Baba is the last One of this cycle. Now a new cycle will begin after His Universal Manifestation. The world will be awakened, and a new cycle will begin.

For Lord Jesus, whose birthday and New Year we are celebrating, I say:

May Beloved Lord Jesus help us all, so that we forget the beginning and end of Creation and know our Real state.

May He guide us at every step of our life towards Him.

May He keep our path clear so that we proceed towards Him without any hindrance by the unwanted impressions that we create every day.

May the Lord give us inner strength so that we become blind to Creation and see only Him and Him alone as infinite.

May He awaken us so that we depend upon Him fully at our every step and take any step only according to His Divine Wish.

May the Lord give us that longing to awaken us towards Him. May we come to know that we never took birth, and we never died, and the births and deaths which we have passed through were all mere illusion, the dream.

May Lord Jesus make us His slave so that we become free from anger, lust, temptation, greed, selfishness and all desires and wants in order to lead our life for Him and remain awakened towards Him.

O Lord, listen to our prayers. We are ignorant and you are the only One who can give us awakening. Only You can make us free from ignorance. You are the only One who can make us free from all rituals, ceremonies, customs and religious limits. Only You can give us the awakening that we are all One and that some day, we become One and realize that from the beginning until the end, we were One, but because of the play of Illusion, we found divisions and were surrounded by hatred and all low desires.

May the Lord give us that awakening in order that we do not become the prey of ignorance so that He may not suffer for us. May the Lord be with us in such a way that we know that the Lord is with us and that we follow His pleasure.

I pray to the Lord for you all --- my mothers, fathers, sisters and brothers. Our Lord is One, and we should not make divisions because of ignorance. Everything happens according to the Divine Plan, but everything does not happen according to His Divine Wish. And therefore, we must follow His Divine Wish with all love, not in a ritualistic way.

O Lord, I bow down to you in such a way that I should not lift my head from Your feet, but remain there until this mind goes away and I forget all the differences and remain One with Beloved God.

*****

In my 2006 Christmas message to all you dear ones, I wrote about Ravana and Rama. Below is a ``Letter of Ravana to Rama'' which was given to me by dear Falu Mistry from Meherazad. The source is unknown (perhaps not from the Ramayana itself). But the letter from the dead Demon King to the Avatar who killed him, Rama, also beautifully illustrates that God alone exists.

THE LETTER OF RAVANA TO RAMA

Rama stood alone over the corpse of the Demon King, Ravana. There was no one else about, nothing stirred.

Sukha, the Rakhsa Minister walked straight to Rama and handed Him a sealed letter. ``For you from the dead King..

Rama cracked open the stone envelope with his strong hands. He looked at the letter awhile, then turned to Suka. ``Ah my friend,'' said Rama fondly. ``Listen:

``Lord Naryana, you are the witness, you make the moon walk in brightness and the stars vanish in the daylight.

``Dear Rama, Lord of the Worlds, think and remember, how you promised Indra to kill me forever. Nothing is forever except for Yourself. Except by dying at Your hand, how else could I make You take me into Your Own Self?

``I was only a Rakshasa, and You were very hard to approach. Yet seeking wisdom, I learned many things. You do not know Who You are again. I knew it all along, but even still You do not know. Nothing You do ever fails. One glance of Yours, and people sing again the good old songs.

``I took no protection against men. You go everywhere and know everything that ever has or ever will be done. How was I careless? I was nowhere careless! O Narayana, lo, I looked, I marveled. Men are mines. Men are precious mines. O Rama, did you think that dark was bad?

``You see whatever happens, and You support all creatures. I saw that Heaven was impermanent, and Hell itself did not endure. I discovered that the time of every life is one day full, and I found how all creatures that are separate from You are ever and again reborn, over and over, always changing. I do not love things that come and go and slip away with Time, and Time himself I hate. I warned him when we first met, that I took him for my enemy. I told him so.

``Best of men, there are many kinds of Love. I never hurt her. I kept Lakshmi to lure you here. I offered you my life and you accepted it.

``You are Narayana who moves on the waters. You flow through us all. You are Rama and Sita born out of Earth and Ravana the Demon King. You are Hanuman like the wind. You are Lakshmana like a mirror; you are Indrajit and Indra. You are the Poet and the Players and the Play. Born as a man, You forget this. You lose the memory, and take on man's ignorance again, as You will, every time.

``Therefore welcome back your Sita. The war is done, and so we close our letter.''

\begin{center} ***** \end{center}

Dear ones of Beloved Avatar Meher Baba, every day is Christmas. During this season we call Christmas, we eat, drink and make merry, but we forget Him. There is nothing wrong in enjoying the food, drink and merry-making, but never forget Him. Remember Him. Remember Beloved Meher Baba, Who was once born as Jesus Christ and Who comes again and again as the same Ancient One.

AVATAR MEHER BABA KI JAI!

With all love and Jai Baba to you,

In His love and service

Bhau Kalchuri


\flrightit{Posted on 2008-01-02 by Isabeau}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{SoldaTHeero on 2010-10-22 at 23:27 said:}

What an excellent synopsis of Meher's teachings and what an awesome Christmas message I will be sure to read this again come Christmas. Jai Baba

BTW I stumbled upon this site a year or so ago while researching something besides Meher, I belive anyhow. \& Now I see you also are aware of Meher Baba. : )

\hfill

\texttt{Cologero on 2010-10-23 at 12:57 said:}

``Isabeau'' has her own blog now.

Please visit \textit{Contessa Isabella Vacani}\footnote{\url{http://isabellavacani.blogspot.com}} for more of her writings.


\hfill

\end{sffamily}
\end{footnotesize}

