\section{Life’s a Bitch, but she’s my bitch so I love her}

\begin{quotex}
Death is certain, the time and manner is not.
\flright{\textsc{Buddha}}\end{quotex}

In \emph{Twilight of the Idols}, Nietzsche wrote:

\begin{quotex}
The invalid is a parasite on society. In a certain state it is indecent to go on living... .The highest interest of life, of ascending life, demands the most ruthless suppression and sequestration of degenerating life --- for example in determining the right to reproduce, the right to be born, the right to live... . To die proudly when it is no longer possible to live proudly. Death of one's own free choice, death at the proper time, with a clear head and with joyfulness, consummated in the midst of children and witnesses.
\end{quotex}

Nietzsche himself died totally insane, staring out the window in his sister's flat.


\flrightit{Posted on 2009-08-13 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{S on 2011-05-11 at 11:32 said:}

Only because he knew nobody with any mercy.

Leaving a man like that is torture.

\hfill

\texttt{santiago on 2020-01-23 at 01:49 said:}

I thought the same thing, recently reading through The Genealogy of Morals. In the third essay, section 14, Nietzsche rails against the sick and the invalids, and all who would take pity on them. He likens the sick and the frail to abhorrent malformed and grotesque ``physiological distortions and worm-riddled objects!''

The parallels between Nietzsche and the most degraded aspects of the Fascist ideal are worth considering. One can see many parallels between his thought and what would become the artistic and inspirational motifs behind the creation of the National Socialist party (especially his fetishism of unjustified temporal power and the--distanced--respect for the `Blonde beast of prey'). I think too often among the far right, the train stops here, and everyday compassion makes way for jaded cut-throat idealism.

Recently I had a death in the family. For a while towards the end, I was entrusted to be his caretaker. The weeks and months leading to this mans death were slow and drawn out to their utmost extreme; he really had strength, that much I can say. But--the confused looks, the constant forgetting of faces and names, the paranoia, the childlike and happy moments, the odd spurts of incredible energy, the anger, the inability to eat or maintain even the slightest trace of equilibrium, the decrepitude and absolute weakness--I will never forget it.

This is nothing new: death has always been an influence in my life. This is not special, everyone experiences death second- and first-hand at some point. How ironic it was that the death of this man was almost immediately succeeded by the considering of ``retirement.''

I hope death comes quick, if I'll never see the other side sooner than later.


\hfill

\end{sffamily}
\end{footnotesize}

