\section{Life is the Absence of Sleep}

\begin{quotationx}
The universality of sin necessitates the universality of purification, which in its turn requires pain to be universal, that the whole human race may be purified in its mysterious waters. This explains why all men born suffer from their birth to their death. Pain is the inseparable companion of life in this obscure valley, filled with our sighs, deafened with our lamentations, and moistened with our tears. Every man is a suffering being, and everything not painful is strange to him. If he fixes his eyes on the past, he grieves to see it vanished; if on the present, he bewails the past as bitter; if on the future, he feels perturbation, because the future is full of shadows and mysteries.

How little soever he considers, he discovers that the past and present and future are all, and all is nothing --- the past is gone, the present is rapidly going, and the future has not come. The poor are loaded with fatigue, the rich with indigestion, the powerful with pride, the lazy with weariness, the lowly with envy, and the mighty with disdain. The conquerors who drive the nations, are themselves driven by furies, and only stumble on others because they are flying from themselves. Lust consumes the flesh of the youth with its impure flames; ambition takes the youth, made man, from the hands of lust, and burns him in other flames, and drives him into other conflagrations; avarice seizes him when lust rejects and ambition abandons him; she gives him an artificial life called sleepless; old misers only live because they do not sleep --- their life is nothing but the absence of sleep.

\flright{\textsc{Juan Donoso Cortés} (marqués de Valdegamas), \textit{Essays on Catholicism, Liberalism and Socialism}\footnote{\url{https://gornahoor.net/library/CortesEssays.pdf}}}

\end{quotationx}



\flrightit{Posted on 2009-08-29 by Cologero}