\subsection{Julius Evola and Lourdes}

\textbf{Letter to Fr. Clemente Rebora (May 14, 1949)}

I thank you sincerely both for the trouble that you took in coming to visit me and for the concern that your offer of a visit to Lourdes shows me. About the visit, I don't believe that it makes any sense, not because there are prejudices on my part, but for reasons, let us say, so technical that whoever takes the point of view not only of faith but moreover of knowledge could easily recognize. I will point out that at Lourdes I would have to go to ask, in the state of grace, that the physical disability be removed. Now, I have already told you how little this thing means to me, and even if the illness were more serious, a man worthy of this name should not turn to the supernatural for things of this type.

First of all, the basic premise, which is that of a burning desire for a healing, is lacking. If you have to ask for a favour, it should instead be that of understanding the \emph{meaning} of what happened in the spiritual realm -- the thing remains or it does not; furthermore, there is the favour of understanding the reason for my continuing to live. I have already noted that the accident was like an enigmatic answer to my asking -- through exposing myself to danger -- if my earthly life could come to an end.

In the second place, I must say that among the symbols of Christianity even that of the ``Mother of God'' is among those that speak the least to me, that are the most strange to me; finally, that all my psychic constitution leaves precious few possibilities to processes that involve emotional and affective factors, starting with those that stand at the base of the very element of ``believing'' in general. Given these circumstances, at Lourdes what would have to happen is something similar to a mechanical action, completely external, that is barely conceivable in itself -- and moreover -- how much of the spiritual design that always accompanies every true ``miracle''. \emph{Being ``cured'' in such a manner would change nothing in me; the true problem, that is the increase of interior light, would not result.} In my gift, of being absolutely sincere with myself, Father, this is what I must tell you. In every way, for the disposition that your proposal reveals, I am grateful to you.

Sincerely,

J. Evola

\begin{quotex}
Subsequently Fr. Rivora wrote to Pittoni: A supernatural tenderness for him grows in my heart. He told me of an inner fact necessary for him in the bombardment at Vienna and he added that his ongoing test still remained mysterious to him. I believe instead I notice a decisive and providential act for his soul. He thinks that this morning I wrote to him suggesting that he come to Lourdes with the hospital train ...
\end{quotex}

\flrightit{Posted on 2009-03-19 by Cologero}
