\section{The Third Testament}

\begin{quotationx}
In order of succession, the \emph{Gnosis} that Jesus revealed to John, James, and Peter after His resurrection reached Clement of Alexandria (about 160-215) and his direct disciples. Due to the persecutions of the third century, and the troubles that arose in the heart of Christianity after it became the State religion, it became imperative to make it `hermetic' if it were to survive. Hidden like a treasure buried in the earth, it slienty made its way and, like a subterranean stream, flowed from master to disciple and from generation to generation until the present, when it rose to the surface again. Stripped of its occult character, it reappears with it original significance as an esoteric projection into the future taking the form of a \emph{New Covenant}, or, in other words, a \emph{Third Testament}.

The Law of the Old Testament, dictated to Moses on Mount Sinai amid thunder and lightning, took the form of a command. On the other hand the New Testament was not imposed on human beings, it came to them as \emph{Good News}, and each was free to welcome or reject it. Though of great significance, this difference goes unnoticed. We will try to understand this different attitude of the divine Will in the two cases. This will enable us to penetrate more deeply the true meaning of the \emph{Third Testamnet}, as well as its message.


\flright{\textsc{Boris Mouravieff}, \emph{Gnosis}, Book Three, VII, IV (1)}
\end{quotationx}



\flrightit{Posted on 2010-05-01 by Cologero}

