\section{In pain there is something fortifying, manly, and profound}

\begin{quotex}
Pain establishes a certain equality between all those who suffer, which is to establish it between all men, for all suffer: by pleasure we are separated, by pain united in fraternal bonds. Pain removes the superfluous, and gives us what we want, and establishes a most perfect equilibrium in man: the proud man does not suffer without losing some of his pride, nor the ambitious man some of his ambition, nor the passionate man some of his anger, nor the impure man some of his impurity.

Pain is sovereign in extinguishing the fires of the passions. At the same time that it removes what injures us, it gives us what ennobles us---the hard-hearted do not suffer without feeling themselves more inclined to compassion, nor the disdainful without feeling more humble, nor the voluptuous without feeling more chaste. The violent become tamed, the weak fortified. No one comes out of that furnace of pain worse than he entered: the greater number come out with sublime virtues they knew not of.

One goes in impious, and comes out religious; another avaricious, and comes out an almsgiver; another without ever having wept, and comes out with the gift of tears; another heart-hardened, and comes out merciful. In pain there is something fortifying, manly, and profound, which is the origin of all heroism and of all greatness; no one has felt its mysterious contact without improving: the child acquires by pain the vitality of youth, youth the maturity and gravity of men, men the bravery of heroes, heroes the sanctity of saints.

\flright{\textsc{Donoso Cortés}}
\end{quotex}


\flrightit{Posted on 2010-03-22 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{J. R. R. Tolkien on 2020-09-29 at 09:41 said:}

I want more of this. Recommendations?

\hfill

\texttt{Cologero on 2020-09-29 at 22:53 said:}

@Tolkein:\\ It is a passage from \textit{Essays on Catholicism, Liberalism, and Socialism by Juan Donoso Cortes}\footnote{\url{https://www.gornahoor.net/library/CortesEssays.pdf}}.

\hfill

\texttt{Taivasvaeltaja on 2024-01-27 at 15:20 said:}

``Cried the young, cried the old, cried the unmarried men,

the males and females wept, the half-aged boys wept, as well as the boys to them.''

-- Kalevala, 49th poem

Let's say it right at the beginning so that there is no doubt about it: real men cry when the burden of the world becomes too heavy to bear. Gone are the days, fortunately, when a man was not allowed to show his feelings and crying was considered a sign of weakness; real strength and power is to also recognize one's apparent weakness, because if anything it makes us genuine and whole people. Of course, we immediately hear the Old Testament cries that only sentimental and weak mourners cry, while a real man is some kind of emotionless automaton, and that crying is the job of batteries.

But is this really in any way a new thing, as we might imagine if we were ignorant of history or myths; no, but the matter is ancient. In the mysticism of Christianity, there has been a lot of talk about the gift of tears and how this contributes to our spiritual growth, but we also notice from our very own Kalevala that all its characters sometimes cry from behind their shoulders. The European story, on the other hand, tells that a man met Christ during his travels and asked him what he should do in order to be saved. With his keen spiritual eye, Jesus saw what the man's problem was and said to him: You must give up your anger, for those who carry anger within themselves will turn into devouring dragons. The best way to tame the raging dragon inside us can sometimes be crying, because it releases the heavy energy that has been pent up inside us, so that our whole being feels like it has been cleansed after crying; energy flows freely again and we can feel as if we have been relieved of a heavy burden -- even for a moment.

``For Satan gives them heavy burdens to carry, so that they would feel the relief of bearing the liberating burden of salvation, even for a moment''.

We could only speculate about the mystical dimensions of crying and the gift of tears and, for example, its positive effect on our emotional body, but anyone who knows the subject knows, even without these technical considerations, how liberating and cleansing a good whistle can be.

We can even say that in the raging tears of a true warrior, the world is cleansed of its own part, and even though crying is the culmination of emotions, in the waters of purification there is not a drop of sentimentality left in the end, but pure Joy.

Let's cry.

\end{sffamily}
\end{footnotesize}

