\section{To Love the Truth}

\begin{quotex}
The coming of the lawless one is by the activity of Satan with all power and false signs and wonders, and with all wicked deception for those who are perishing, because they refused to love the truth and so be saved. Therefore God sends them a strong delusion, so that they may believe a lie, in order that all may be condemned who did not believe the truth , but had pleasure in unrighteousness.
\flright{\textsc{2 Thessalonians 2:9-12}}
Some of us have exhausted ourselves trying to distribute factual information and convince relatives and friends of the truth, to no avail. We ought rather to first teach them to love the truth; only then will God lift His curse from them; only then will they be receptive to the truth.

(In speaking of possessing truth, we say this with humility, vigilant against our own pride, and cognizant that we ourselves still have much to learn, and are as potentially fallible as any other human being).


\flright{\textsc{Michael Hoffman}}
\end{quotex}

The above passage is from the current issue of Revisionist History\footnote{\url{http://revisionisthistorystore.blogspot.com/2010/03/hoffmans-revisionist-store-newsletters\_31.html}}, titled \textbf{``Judaizing Protestantism'' and Neo-Platonic Catholicism in Legend and Reality}. Mr. Hoffman offers up a perspective in opposition to the Hermetic viewpoint of this project. He traces its roots to the Renaissance and the influence of the Talmud on both Catholics and Protestants.

Although the article is valuable for its historical references, we prefer to think that the Hermetic and Platonic traditions are much more ancient than the Renaissance, are free from anti-Western influences, and represent a valuable gateway to authentic Tradition in the West.


\flrightit{Posted on 2010-06-05 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{GFJF on 2010-06-06 at 01:23 said:}

Yes, but how? A man must reach a certain level autonomously before he can be brought to love the truth. ``If any man come to me, and hate not his father, and mother, and wife, and children, and brethren, and sisters, yea, and his own life also, he cannot be my disciple. And whosoever doth not bear his cross, and come after me, cannot be my disciple . . . whosoever he be of you that forsaketh not all that he hath, he cannot be my disciple.''

\hfill

\texttt{Cologero on 2010-06-06 at 23:12 said:}

In \textit{Stages of Man}\footnote{\url{http://www.gornahoor.net/?p=214}}, I show the three stages of consciousness described by Evola in the \textbf{Individual and the Becoming of the World}. In the first stage, the question of epistemology does not even arise. People believe any old thing and remain unconcerned about the question of truth. If the question engages the consciousness, a man reaches the next stage and looks for truth in various religious and scientific systems. This is a good start, although such a person usually gets bogged down in some ideological system and never progresses to the next stage.

\end{sffamily}
\end{footnotesize}

