\section{A Way Out, Further In}

One of Evola's most prescient suggestions is that the modern aspirant towards transcendence should ``ride the tiger'' -- the individual (or the person, but not the personality) should use the negative energies of the modern world to advance himself, until they are exhausted, for opposing them directly would invite destruction. There is an important qualification:

\begin{quotex}
This restriction must be kept in mind. What I am about to say does \emph{not} concern the ordinary man of our day. On the contrary, I have in mind the man who finds himself involved in today's world, even at its most problematic and paroxysimal points; yet he does not belong inwardly to such a world, nor will he give in to it. He feels himself, in essence, as belonging to a different race from that of the overwhelming majority of his contemporaries.
\end{quotex}


\emph{When one cannot find a way out, find one further in}. At first glance, this is a seeming difficulty for the Christian, who is attuned to essentially reject the possibility of ``utilizing'' something lower to actualize something ``higher''. It smacks of Shelley-like Satanism. Gornahoor has addressed this more precisely here\footnote{\url{http://www.meditationsonthetarot.com/axioms-of-the-will-i}}:

\begin{quotex}
One can and one should, accept evil as the means of good; but one must never will it or do it, otherwise one would destroy with one hand what one builds with the other. Good faith never justifies bad means; it corrects them when one undergoes them, and condemns them when one takes them. \textbf{COMMENTARY.} Evil increases the virtue of the wise but corrupts the weak.
\end{quotex}

One way to re-speak this (to the Christian) is that though ends don't justify means, yet means do not pardon ends either; so that what is called ``moralizing'' is often nothing but abdication of the entire moral sense of man at its root, whether it is ``Christian'' or not. Although might does not make right (since the gross is ruled by the subtle), neither does it follow that weakness makes right (Nietzsche's objection to the Faith, as expressed). ``Force and justice rule the world; force, till right is ready'' (M. Arnold, quoting Joubert). To think that weakness and victim-hood make right is to succumb to the lowest form of mental degradation possible (currently). Right makes right.

\begin{quotex}
For all doing is shrouded in guilt, like the blaze of the flame in smoke.
\flright{\emph{Bhagavad-Gita} (XVIII, 48b)}
\end{quotex}


In this sense, all action in the world seems to fall afoul of evil's tincture. This is the reason for the ethical asceticism advocated by the \emph{Bhagavad-Gita}: the true man (compelled to action) must do what is right, regardless of consequences\footnote{\url{http://www.youtube.com/watch?v=lq3zjTmVLbM}}, but not regardless of results.

Results are carefully calculated, but not for personal gain (or, one might add, even for personal ``salvation''). It is here that Evola's remarks on Calvin\footnote{\url{http://www.gornahoor.net/?p=2194}} are relevant. Being willing to be damned for God's glory means being committed to ethical action in a total way that re-unites the ``afterlife'' to this life. One begins to operate with reference to transcendence, of ways and means and also end. This was a way of exalting the moral sense in man without subjecting that sense to the ego. ``Let God be true, though every man a liar''.

Or, ``By patient continuance in well-doing, seeking for glory, honor, immortality... ''

There is ultimately no conflict between the warrior \& the saint. The cross, which is an inverted sword, once stood as the symbol of this.

One possible re-unifying thinker which could re-mediate between Christian ethics \& perennial metaphysics is the Japanese Buddhist, Kukai\footnote{\url{http://plato.stanford.edu/entries/kukai/}}. Kukai's diagram for gnosis was

\begin{quotex}
1. Freedom from bias

2. Insight into particulars

3. Improvisation towards enlightenment.
\end{quotex}


Like most perennialists, he placed the religious worldview below the metaphysical, but superior to the secular. Kukai's work is a naturally magic bit of either one, \& stands as an exemplar (for example, he argued that a ``Triad'' was at the root of reality, yet he maintains the hierarchy of metaphysic over ``religion'').

Recently, venturing on such ground, the pope himself has directed Christian attention\footnote{\url{http://www.zenit.org/article-22588?l=english}} towards P-Dionysius' reworking of Proclus as a bridge towards Asia. There are some things to object to in the speech, but it is important because its aim is to conserve Proclus through Dionysius.

\begin{quotationx}
Apparently what Plato says and what great philosophy says about God is much more elevated, much more true; the Bible seems very ``bárbara,'' simple, precritical, we would say today. But he observes that precisely this is necessary so that we can thus understand that the most elevated concepts of God never reach his true greatness. They are always beneath him.

These images bring us to understand, in reality, that God is above every concept; in the simplicity of the images, we find more truth than in the great concepts. The face of God is our incapacity to truly express what he is. In this way he speaks --- Pseudo-Dionysius himself says --- of a ``negative theology.'' It is easier to say what God is not than to express what he really is. Only through these images can we grasp at his true face and, on the other hand, this face of God is very concrete: It is Jesus Christ. If Dionysius shows us, following Proclo, the harmony of the celestial choirs, in such a way that it seems that all of them depend on each other, it is true that our path toward God remains very far from him. Pseudo-Dionysius shows that in the end, the path to God is God himself, who makes himself close to us in Jesus Christ.
\end{quotationx}


The pope's essay touches upon the bitter wound or nub between the two -- it is precisely this idea of reunion in God (or reabsorption into God) which is contested by traditionalism. I do not, however, see why perennialism could not accept the idea of exoteric, simple language as a manifestation of that which is beyond words.

It would certainly improve upon the Kantian idea that men are universally not merely equal, but \emph{fungible}, which is to say, absolutely interchangeable. For the medieval imagination and the \emph{prisca theologica,} as well as Evola, this is a grave and serious delusion, perhaps a prime one. The ``Three Orders''\footnote{\url{http://www.wwnorton.com/college/english/nael/middleages/topic\_1/aelfric.htm}} were once of those who fight, those who work, and those who pray. Evola speaks of ``types'' and Ortega y Gasset (a classic liberal) speaks of ``qualitative man''. Many other cultured Europeans noted the rise of ``American man''\footnote{\url{http://www.scribd.com/doc/24545690/American-Civilisation-Julius-Evola}}. As Gornahoor has argued\footnote{\url{https://www.gornahoor.net/?p=2244}}, the dividing line is the Middle Ages because after they were destroyed, the chain of being and the ``orders'' were obliterated.

Christianity's temptation (and even tendency) is reduce everything to the lowest common denominator in the name of compassion. If we take a great yet flawed interpreter of Western history like Rosenstock-Huessy\footnote{\url{http://www.argobooks.org/english/the\_christian\_future.html}} in the modern era, we find him proclaiming the era of the individual:

\begin{quotex}
Today we are living through the agonies of transition to the third epoch. We have yet to establish Man, the great singular of humanity, in one household, over the plurality of races, classes, and and age groups. This will be the center of struggle in the future. They pose the questions the Third Millennium will have to answer... the State is on the defensive because it is inadequate for the needs of the coming age. The theme of future history will be not territorial nor political, but social...
\end{quotex}


Huessy's sequence ran

\begin{enumerate}


\item The Christianization of Europe/Romanization of the Church

\item The Papal Revolution 11th-13\textsuperscript{th} century, where the Church destroys the yoke of State

\item Protestant Reformation

\item The Enlightenment
\end{enumerate}


Huessy again profoundly incarnates the basic tension between perennialism \& Christianity. To the perennialist, Christianity must always seem in danger of simply being turned inside out, \& consequently, self-destroyed. Huessy fell into the trap of not perceiving that the ``Zeitgeist'' might not necessarily be the breath of the highest Being nor the pattern which was beyond Being. Another way of putting this is that being absorbed into the \emph{Anima Mundi} \emph{\textbf{or even into what is perceived as ``God''}} may not be in accord with the true possibilities of the inner man.

Again, we are back to the argument over ``absorption'' into God. That, and that Perennialists are looking for something other than this ``We'' that is always thrown about in secular democracies.

\begin{quotex}
Most important of all through this physical rebirth you are able to achieve the state of Vajradhara within one short lifetime in this degenerate age; otherwise it would take thee countless great aeons to achieve. Thus this rebirth is worth more than one thousand billion precious jewels.
\flright{\textsc{Pabongka Rimpoche}}
\end{quotex}


For the sake of this (the kingdom of God), Christians must therefore part ways with the likes of Augustine at times, though all manner of evil is said against us falsely for this (even by the Church):

\begin{quotex}
Please notice carefully that in speaking of the good morals of some atheists, I have not attributed any real virtues to them. Their sobriety, their chastity, their probity, their contempt for riches, their zeal for the public good, their inclination to be helpful to their neighbor were not the effect of the love of God and tended neither to honor nor to glorify him. They themselves were the source and end of all this. Self-love was the basis, the boundaries, and the cause of it. These were all glittering sins, \emph{\textbf{splendida peccata}}, as St. Augustine has said of all the fine actio of the pagans.
\flright{\textsc{Pierre Bayle}}
\end{quotex}


And thus, we are back to the moralizing ``we''. It would seem that the entire debate as it has ossified between Christianity \& traditionalism could potentially be a ``we''-argument dead end.

Given the historical examples of middle ground, and indeed either in the other's camp, it should prove relatively easy to simply will the seemingly impossible reconciliation. If it is chosen to do so.

Continuing the argument at its highest levels and lowest roots (focusing specifically on the idea of ``reabsorption'', with all the attendant assumptions about man's ultimate fungibility or equation with all other men and indeed things) would seem to be an excellent place to begin.

The Christian should begin the exercise by meditating on what the possible meaning of this riddle is: Why does Christ never speak of Yahweh/Jehovah, but only of His ``Father''?

And why did the ``religion'' of Christ not begin until after his triumph over death?

Finally, why does Christ come at the end of time, not ``re-absorbed'' into God, but giving all back, \& riding a horse of war, with a sword coming out of his mouth?

The servants of the secret fire\footnote{\url{http://en.wikipedia.org/wiki/Secret\_Fire}} should stand together.


\flrightit{Posted on 2011-06-24 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Perennial on 2011-06-25 at 01:38 said:}

Please elaborate on ``absorption'' into God. By this you surely do not mean Theosis, which it would seem is a seperate concept.

\hfill

\texttt{Ismo on 2011-06-25 at 11:45 said:}

There is a small error in the first link: it works only if one removes the \_One can -part from it. If the text is open to correction it would be a good idea to remove the flawed part from the end of the link.

As it comes to `Shelley-like Satanism', in some sense this could be true, if one identifies `Ha-Satan' with manifestation itself, with the `bright face' of Shekinah. But this sort of `Satanism' would be very far from the usual popular hedonistic, self-centred and materialistic satanism or from the descending path of counter-tradition of the `awliya-ash-Shaitán'; it would be instead very close to the traditional `left hand path', exalted and elaborated by Evola also in some connections, and suitable only for `regal souls', or virya and kaula, in eastern terms. There are a growing number of people in todays world who identify themselves as `satanists' or as `left-hand-pathers', but shun away from both of the first-mentioned blind alleys. In this sense the term could and may refer to western form of the left hand path and maybe even one of the signs of east and west coming together. Of course it is understandable that there's alot of confusion and lower possibilities that need to be cleared away before this sort of satanism can be seen `in traditional terms'. Still, let us not forget that Saturn is the god of the Golden Age.

Good article, by the way!

\hfill

\texttt{logres on 2011-06-25 at 11:48 said:}

Yes, as I see it, perhaps the biggest sticking point between Christianity and traditionalism is precisely whether or not theosis is absorption or reabsorption into God. If it is not, then much of the difficulty between the two goes away, on the pagan side. The difficulty is that these are experiential concepts -- it's not possible for me to arbitrate them (as I've not experienced it). However, in a tradition where there is less of a conflict between the nation and the Church (eg., Russia) figures existed/may exist who can arbitrate and tell one whether compassion is compatible with supreme individuality and survival. My understanding of traditionalism is that they will avoid (at all costs) compromising on the issue of absorption into God and forfeiting the ``being beyond being''. Isn't this what caused the difficulty for Evola?

\hfill

\texttt{Logres on 2011-06-25 at 11:56 said:}

Thank you, it's fixed. ``Satanism'' is a misleading term (from my perspective) for the left-hand path. The ``royal path'' or alchemy (?) is a much friendlier Western phrasing. I came across a term for Crowley recently, which was ``vulnerable master'' -- that is, someone who had managed to integrate to higher reality by bypassing the heart chakra entirely (being almost entirely Service-to-Self, like Genghis Khan for instance). This term notes that he has something to say, but has also lost something in the process. To Christianity, in general, this is the danger of the ``left-hand'' approach, although such characters as Ehud in the Old Testament embody it.

\hfill

\texttt{Ismo on 2011-06-25 at 15:40 said:}

Yes, `Satanism' has a bad ring to it as a term, because of numerous mis-interpretations and confusions, but I'm convinced that it cannot be completely disregarded the more further we approach the end of the cycle and the beginning of the new one, that many of those who are able and will `ride the tiger' will identify themselves as satanists. As in basic kabbalistic terms Satan (or Saturn, Aima) represents `the left hand of God', in this way satanism is a very correct term for the left-hand-path. Destructive? Yes. Evil? Not necessarily, if one is able to traverse the tree without giving in to the lowest possibilities and \emph{transform the destructive tendencies into something positive and constructive} = to ride the tiger.

There are also many different sides to this issue that I cannot go into in this short post, but by delving into the nature of Metatron's cube and into the cubic form of `the new Jerusalem', and comparing these symbolic issues into the saying of Jesus to Peter (Petrus = stone) ``Upon you I will build my church'', many things about the last two thousand years and about the history of the Catholic Church (``The Synagogue of Satan'', as some have said about it) start to open up.

Shortly, I see this issue about the coming age of Fire and our transitional times in the following way: Nietzsche got a hunch of it but failed to understand it; Crowley got a big glimpse of it but failed to formulate it without errors, intentionally or not; Lavey (or Levy) brought -- typically for a judaic soul -- out the lowest possibilities of it; Evola formulated it into a coherent form. The rest is yet to come.

\hfill

\texttt{Logres on 2011-06-25 at 19:46 said:}

I certainly accept that that which is Good in the Kali Yuga era (in spite of the era) will look ``bad'' to virtually everyone. The confusion is still so great on the ground that it's difficult to say much more. But the line of formulation laid out there through to Evola I basically accept, except Crowley troubles me somewhat. I think it does have something to do with the bypassing the heart chakra, which Evola does not do, as is evident in his essay on Crowley (he accepts what is worthwhile, ignores the rest). Crowley was also a great showman, as well as (in my opinion) definitely a ``lower'' type of man, even in his exaltation beyond the ego. I don't think that the Kali Yuga necessarily entails the arrival or dominance of this ``type'', although it certainly may entail the necessary understanding \& use of them. These are just sketchy opinions.

\hfill

\texttt{Ismo on 2011-06-26 at 06:54 said:}

Logres, I completely agree with you on Crowley. In many ways he had great potential and great theroetical knowledge at his disposal, but his personal charlatanism was quite repulsive and his writings leave a lot of room for very negative impulses in regard of spiritual development and not mere psychism. In this sense he can perhaps be counted among `the grey ones', and if the knowledge of his secret political actions are true it only strengthens this.

The reason I draw a line from Nietzsche to Evola and included Crowley (and even Lavey, sic!) is that all these gentlemen seem like precursors -- and warningsigns! -- of the coming age. Evola also made his own mistakes (his injury in the third chakra is quite telling) and left some room for very negative impulses, although in every other sense he understood the most clearly and with Arya -like nobility the way of the differentiated man in the Kali Yuga. By taking what is worthwile even from the most modern and anti-esoteric writers and thinkers, and dumping the rest, I see him as an archetypical LHP -practicioner.

\hfill

\texttt{Logres on 2011-06-26 at 10:36 said:}

Ismo, I know nothing of Crowley's politics. And I agree about the warning -- signs, that's a good way to put it. Trying to steer between residual sentimentalism \& inversionism, here... 

\hfill

\texttt{Ismo on 2011-06-29 at 15:42 said:}

Basing my judgment about Crowley and Politics on this article, his views seem to be actually somewhat enlightened:
\url{http://www.arcane-archive.org/religion/thelema/aleister-crowley-on-politics-1.php}


... but as I understand and have read from his biography, he worked as some sort of a double-agent during the first or second world war. I cannot say this with an authority though, since I read it quite a while ago from a good friends book, and can't say if I remember absolutely correctly. Still, I think it is clear there was some shady businesses that he was into and politics was part of it.

Fighting with residual sentimentalism and inversionism... ? Join the Camp!

\hfill

\end{sffamily}
\end{footnotesize}

