\section{The Interior Lifelessness of the Nordics}

This is a section from \textit{Il Fascismo visto dalla destra}, published by Julius Evola shortly after WWII. As is evident from this paragraph, the process of decay, or progress, depending on your perspective, in the Scandinavian countries, has been going on for quite some time. Obviously, this cannot be attributed to biological racial characteristics. I would like to translate the entire chapter which is a commentary on the Volkisch policies of the third Reich. However, that project is contingent on intelligent discussion of this topic, since biological racists have hijacked a large part of the right, even claiming to represent the values of Tradition, when they are actually examples of the mental deformation Evola criticizes.

\begin{quotex}
That biological aspect had a relevant part and created many illusiond because of a mental ``scientistic'' deformation when they thought that it sufficed to to proceed to a prophylaxis and erect barriers against crossing and hybridism, along with more eugenic measures, so that by virtue of it, man, the creator of a superior civilization, vanishes, reappears, and rises again almost automatically. Today, there are populations, such as the Norwegians, Swedes, or Dutch, who are of a high level of racial and even ``Nordic'' purity, but who interiorally are more or less lifeless \emph{spente} , are spiritually bastardized, deprived of the virtues that characterized them in other times.
\end{quotex}


\flrightit{Posted on 2011-08-06 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{I on 2011-08-09 at 15:14 said:}

Nothing could be more true than this. Sentimental chickens danzing to jazz, tango \& r'n b combined with arch-materialistic `pure-blooded nordics' listening to and making rap music, treating their ladies like whores and going around cities and drive-ins with their new cars, that's what it in large parts basically amounts to nowadays. Can't really blame it upon anyone else than themselves.

\hfill

\texttt{Spir on 2011-08-10 at 20:18 said:}

An examination of the ideas of the various rightist groups that have made active reference to the writings of Evola, and to a lesser extent Guenon, can be disturbing.

Not only a perverted biological racism, loosely defined and sentimental compared to 19th and 20th century Europe's `scientific' attempts. In addition, there is a demented longing not for Tradition, which is truly unknown to most, but for some kind of ideal 1950s or Victorian era style of society where white men work hard and raise big families and everyone is polite and dresses well, with some folk dancing and improvised ancestor worship thrown in.

This criticism is not offered gratuitously -- it should be obvious that we are witnessing the confused attempts of underqualified and sometimes suspect individuals attempting to endorse a restoration `from the outside'; vain to say the least. These groups are ripe pickings for the real agents of the era.

Here is an example:

``Then, as ``money and pleasure'' cede to the imperatives of ``blood and soil,'' only the traditions, the way of life, and the genetic principles SIC defining them as a people will matter.''

``Genetic Principles''? This is idiotic. They wish to replace materialism of the commodity with commodification of the biological. As though the whole universe of Existence should be subject to the rigidity of the physical world. Great way to set up your own people for an unwinnable battle.

As though these mental deformations were not enough on their own, these groups diminish their small chances of prosperity even further by stimulating the occult war with their attacks on Islam. Their criticism of the behavior of uncouth African and Arab urban youth in Europe is lumped under a generally paranoid criticism of `Islam' -- the reproachment of a whole civilization undertaken by a rag-tag bunch of `thinkers'. Could anyone with even a pretention to metaphysical understanding be so foolish as to aim such generalized attacks at a traditional civilization without expecting occult repercussions?

Needless to say, Cologero, that your project is an exception.

\hfill

\texttt{Perennial on 2011-08-11 at 05:27 said:}

I presume, Spir, that your deprication of Victorian style society stems more from a concern in baseless restorations, than in the thing itself. I personally favour the style of society you describe to a degree, but only if based on realities, and not on some exaggerated moralism.

\hfill

\texttt{Spir on 2011-08-11 at 09:17 said:}

Perennial, suffice to say that different settings offer different possibilities, and at a certain point one can view all possibilities as neutral; any personal preference is thus acceptable, precisely because it is a preference -- people prefer a setting in which they can do what they wish to do.

But in the context of Tradition, an objective standard is raised, and if activist groups agitate for a restoration of the West to resemble conditions one or two hundred years ago, are these groups not simply composed of reactive nostalgic moderns? In and of itself this desire is not illegitimate (especially if we are discussing mere style); however, trying to pass such activism off as `traditional' betrays a lack of understanding, if nothing more.

To be fair, in the domain of action most need to work with tangible things, and so such social goals as those that I appeared to deprecate might be the only thing many people can really strive to attain. I suggest to such activists that their goals are modest enough to be decoupled from extensive association with the arcane theorists of Tradition without it adversely affecting their pursuits.

\hfill

\texttt{I on 2011-08-11 at 12:30 said:}

Spir: A few thoughts came to my mind from both of your posts: First of all, I think you are wholly right in your basic evaluation of the rightist groups who are using and more or less distorting Evola and Guénon. They ar thaking this or that from their thinking and debunking what they don't like or what doesn't inspire them in their partisan actions. This is very dangerous indeed. I think this is also the very thing that the upholders of Tradition have warned about: in the end of the current cycle the counter-initiation will change its tactics from those of the anti-tradition (modernity, liberalism, enlightment values etc.), and its agents will apparently pose themselves as upholders of Tradition, while in reality working vehemently against Tradition -- the latest Borean bomber being somewhat a textbook example of this agency in its most visible form. The more hidden and vicious forces will propably try to collide different factions against each other by inspiring them to use and misuse the thoughts of traditional thinkers: for example taking from Evola his ``spiritual racism'' and turning it into materialistic, deterministic and biological racism and inspiring irrational hate and violence from this background.

Yet, I think that you somewhat took the ``genetic principles'' quote out of its context: it is true that genes (materialism, biologoy) cannot be the basis of this sort of awakening, but it is possible that genes (and other more occult forces traveling in blood and soil of a people) can be to some extent act as catalysts and as the activating biological / phenomenal side of a spiritual awakening. I have to add, though, that this is mere speculation, and if the spiritual principles are not latent but have been throughly erased as seems to be the case in most of modern men, who are largely ``demos'' instead of ``ethnos'', then no amount of ``superior genes'' can help. Only ``transcendence within'' can, in the last analysis.

\hfill

\texttt{Perennial on 2011-08-11 at 13:26 said:}

Excellent analysis, Spir and I. I concur whole-heartedly.

\hfill

\texttt{Spir on 2011-08-11 at 21:32 said:}

I, your observations are sound. There are ways in which a biological base, such as a particular race of the body, can catalyze a spiritual development*. Myths play an important part in conveying the keys and animating the soul. But (with the dominance of demos, as you explain) this can never apply to whole masses of people in our time, only to rare special men, so we would see further deformations or lukewarm possibilities if activists attempted to mobilize masses around this idea; the generality of people do not even need to know that these horizons exist.

Yet, on the mundane plane, if a people want to coalesce around a shared biological heritage, and/or customs, this is not an absolute error. From the highest vantage, the Origin, one can point out omissions and shortcomings (human affairs are small from on high); but if a group has established goals particular to some limited domain, and feels that it has the means to attain them, then this special criticism can be ignored.

*Not only biological, but even social and other individual factors; since these factors predominate in the life of the person, and the person is (or more accurately, appears to be) the point of departure for spiritual development, any combination of factors can catalyze the commencement of spiritual development where it is destined. Thus, there is not such an urgent need to impose reforms modeled on the ideal society; a less lifeless reaction to turbulent times involves opening new paths.

\hfill

\texttt{Admin on 2018-08-06 at 12:51 said:}

Looking back at these older posts, I find it curious that there are no longer any discussions in the comments.


\hfill

\end{sffamily}
\end{footnotesize}

