\subsection{Guenon/Evola Letter 7 Introduction}

The correspondence between Rene Guenon and Julius Evola was broken after 1934 and resumed in 1948, where there was a regular and frequent correspondence between them. Letters 3 through 6 and the first several paragraphs of Letter 7 were concerned primarily with topics only of interest to authors: books in progress, reprints, translations, etc. The second part of Letter 7, however, addresses several important topics. Unfortunately, we don't have Evola's letter which I am sure would be of great interest. We can only infer its contents, based on Guenon's responses. Here we see some significant differences between the two, and Guenon is taking the teaching role. My opinion is that Evola's works would be even stronger had he been able to absorb some of Guenon's corrections.

Guenon can be curt and often shows impatience with Evola's views. The points addressed are important and need to be taken seriously. There should be ample material for those interested in a discussion. but for those concerned about our comments policy, bear in mind we are not interested in infinite discussion, but in coming to a decision. Please try to understand what Guenon is saying, either by reading our material on the topics, or better, his directly if you have the time. By coming to terms with their fundamental principles, one's understanding of both Guenon and Evola can only be deepened.

Since the letter is rather long, it won't be published until tomorrow evening. But, in anticipation, the following topics will be addressed:

\begin{itemize}
\item whether the metaphysical identity of the possible and the real is an error 
\item the esoteric character of early Christianity 
\item whether Masonry has a true initiation 
\item the current existence of Christian Hermetic orders that go back to the Middle Ages 
\item discussion of Eliphas Levi, Meyrink, Bo Yin Ra, and Kremmerz 
\item the difference between the true man and the transcendent man (jivan-mukti), a topic recently addressed here 
\item Guenon took several initiations at a young age 
\item Evola was still looking for a valid initiation (so much for the ``Sufi of Rome") 
\end{itemize}


\flrightit{Posted on 2012-07-29 by Cologero }
