\section{The Golden Book \& Saint Clement}

Boris Mouravieff, when teaching on polar-beings in \emph{Gnosis}, references the existence of the Golden Book, an oral (if also incomplete) compilation of Jesus' teachings to the inner circle of disciples. It is only fair to point out that (like the exoteric-esoteric paradox of Christianity itself), the idea of a written compilation of oral teachings is somewhat odd -- nevertheless, then, its very incompleteness testifies to its legitimacy, as some teachings must be ``discovered''.

Clement of Rome\footnote{\url{http://www.scribd.com/doc/13561289/-The-Authentic-Peter-The-Preaching-of-Simeon-Kefa-from-the-Journal-of-T-Flavius-Clemens-Clement-The-Clementine-Recognitions-Recognitions-of-Cl}} was involved in the ``inner'' circle of disciples following Christ's death -- the ``inner'' circle perpetuated itself and thereby endorses the reading of the Gospels in this light: Jesus' parables are given because ``they have ears to hear, eyes to see, but cannot... ''. A \emph{parable} (as Ea points out in the collected Ur-group writings) is a ``carrying with or beside'', and thus less direct than a \emph{symbolon}, which is a ``carrying into''.

\begin{quotex}
Clement was a highly educated young man \emph{with a photographic memory}. He was also a member of the powerful Flavian dynasty (and cousin to emperors Vespasian, Titus and Domitian). Though young, he was a man of considerable financial means. He had no background whatsoever in Judaism and would not have come into the circle of Nazoreans in Jerusalem had he not by chance net Barnabas in Rome and followed him to Palestine.Clements high intellect and great curiosity were not lost on James the Just (Yaakov haZaddik), overseer of the Nazorean Assembly in Jerusalem. He officially assigned Clement to Peters group of students and commanded Clement to send periodic reports of Peters activities back. These reports are often mentioned in the book, and are sometimes indistinguishable from Clement s own journal entries. As an unbaptized gentile proselyte, Clement and others were not allowed to eat or sleep with the baptized until judged spiritually and doctrinally fit for immersion.
\end{quotex}


I am not sure his relation to Clement of Alexandria, a point I raised in a previous essay when using this source. Jackson Snyder appears to confuse the two Clements, unless (perhaps) they were actually related by blood as well as by body of myth and name. Regardless, Clementine literature\footnote{\url{https://en.wikipedia.org/wiki/Pope\_Clement\_I}} portrays him (and Clement of A) as:

\begin{quotex}
the Apostles' means of disseminating their teachings to the Church
\end{quotex}

I will remark here, two things: the presence of an elite (``highly educated... with a photographic memory'') and the doctrine of secret teaching (the ``Church'' was a cult tolerated by the disciples as a condescension to man's weakness, but a body which was subjected to impulses from an elite to ``regulate'' it's health).

The beginning of the text rings true:

\begin{quotex}
I Clement, who was born in the city of Rome, was from my earliest age a lover of chastity; while the bent of my mind held me bound as with chains of anxiety and sorrow. For a thought that was in me whence originating, I cannot tell constantly led me to think of my condition of mortality, and to discuss such questions as these:Whether there be for me any life after death, or whether I am to be wholly annihilated: whether I did not exist before I was born, and whether there shall be no remembrance of this life after death, and so the boundlessness of time shall consign all things to oblivion and silence; so that not only shall we cease to be, but there shall be no remembrance that we have ever been. This also I revolved in my mind: when the world was made, or what was before it was made, or whether it has existed from eternity. For it seemed certain, that if it had been made, it must be doomed to dissolution; and if it be dissolved, what is to be afterwards? Unless, it may be that all things shall be buried in oblivion and silence, or something shall be, which the mind of man cannot now conceive.
\end{quotex}


What noble young man, starting upon life, who is not clouded in the fogs of lust, has not felt this yearning? At first, Clement desires to consult a magician who will raise a spirit from the dead to prove the immortality of the soul, but a friend dissuades him, saying that the Elohim set themselves against those who trouble dead souls (here in the idea/reality of General Law).

He defends a visitor to Rome who tries to speak on behalf of the disciples, and invites him to his home. Clement sails to Caesarea, and there is brought to meet Kefa (Peter). Peter has decided to ``choose'' him as his successor. Peter teaches him the cause of man's ignorance:

\begin{quotex}
The will and counsel of YHWH has for many reasons been concealed from men; first, indeed, through bad instruction, wicked associations, evil habits, unprofitable conversation, and unrighteous presumptions.~ On account of all these, I say, first error, then contempt, then infidelity and malice, covetousness also, and vain-boasting, and other such like evils, have filled the whole house of this world, like some enormous smoke, and preventing those who dwell in it from seeing its Founder aright, and from perceiving what things are pleasing to Him. What, then, is fitting for those who are within,excepting with a cry brought forth from their inmost hearts to invoke His aid, who alone is not shut up in the smoke-filled house,that He would approach and open the door of the house, so that the smoke may be dissipated which is within, and the light of the sun which shines without may be admitted... 
\end{quotex}


Peter is saying here, that the Kali Yuga predominates. This poetic metaphor resonates even with Western ecclesiastical\footnote{\url{https://en.wikipedia.org/wiki/Edwin\_of\_Northumbria}} history:

\begin{quotex}
The present life man, O king, seems to me, in comparison with that time which is unknown to us, like to the swift flight of a sparrow through the room wherein you sit at supper in winter amid your officers and ministers, with a good fire in the midst whilst the storms of rain and snow prevail abroad; the sparrow, I say, flying in at one door and immediately another, whilst he is within is safe from the wintry but after a short space of fair weather he immediately vanishes out of your sight into the dark winter from which he has emerged. So this life of man appears for a short space but of what went before or what is to follow we are ignorant. If, therefore, this new doctrine contains something more certain, it seems justly to deserve to be followed.''
\end{quotex}


The souls of Edwin and Clement were worthy to receive the Truth, \emph{therefore, it appeared}. Those who think all souls are equal are servants of Satan -- all souls are in one sense equally valuable, but there is a reason certain people yearn towards the truth, and others do not. Clement desires very much to know the answers to his questions, enough to humble himself (outwardly and inwardly) -- this ``creates'' a higher Truth for him than his current spiritual state has manifested (the true un-Semitic definition of humility).

Simon Peter prepares for an upcoming controversy with Shimon the Magus, which is delayed, giving them time to discuss things which may come up during the debate. Clement is somewhat saddened, but Peter teaches a truth similar to Boethius' arguments in \emph{Consolatio}:

\begin{quotex}
He who believes that the world is administered by the providence of YHWH El Shaddai ought not, O Clement, my friend, to take it amiss, in whatever way particular things are done, being assured that~ of YHWH guides to a favorable and fitting issue even those things which seem superfluous or contrary in any business, and especially towards those who worship Him more intimately; and therefore he who is assured of these things, as I have said, if~anything occur contrary to his expectation, he knows how to drive away grief from his mind on that account, holding it unquestionable in his better judgment, that, by the government of the good YHWH,even what seems contrary may be turned to good.
\end{quotex}


Peter sees an opportunity to expound oral teaching, which cannot be taught any other way, including \emph{Sola Scriptura}.

\begin{quotex}
for I believe that it has been done by the providence of YHWH,for your advantage; that I may be able, in this interval of seven days,to expound to you the method of our faith without any distraction, and the order continuously, according to the tradition of the Navi Emet, who alone knows time past as it was, the present as it is, and the future as it shall be: which things were indeed plainly spoken by~Him, but are not plainly written; \emph{so much so, that when they are read, they cannot be understood without an expounder, on account~of the sin which has grown up with men,} as I said before.
\end{quotex}


This is absolutely true for all except those of whom the mystery of the Just applies, and it still affects them. Peter teaches a truth which resonates with Guenon's work on the Manifest/Unmanifest:

\begin{quotex}
There always was, there is now, and there ever shall be, that by which the first Will begotten from eternity consists; and from the first Will precedesa second Will. After these came the world; andfrom the world came time: from this, the multitude of men; from the multitude the election of the beloved, from whose oneness of mindthe peaceful Malkuth of YHWH is constructed... 
\end{quotex}

In other words, there is something underlying God's manifestation that is unmanifest. The Orthodox still distinguish between God's essence (unknowable) and His energies (manifested and experienced). Peter then teaches about the ``upper waters'' which divide heaven and earth. The point here is that Peter is \emph{initiating} Clement. We learn about Nimrod and the worship of fire, as well the adoration of each nation's \emph{sarim}\footnote{\url{http://www.archangels-and-angels.com/misc/sarim.html}} as gods. Here is given the rudiments of a cosmology. The Western tribes drove the middle tribes into the East by means of violent warfare. Abraham the astronomer recognizes the Creator in the stars and is taught by the angel of the Lord, who shows him that his inheritance lies in Palestine (where usurpers dwell). This is connected with the ``boundaries of men'' set by the \emph{sarim} and General Law. What follows is a summary of the Old Testament. Then comes the Advent. It is worth noting that the 12 disciples and 70 elders mirror the angelic orders, in the intention of Christ, the image of the \emph{Imperium}:

\begin{quotex}
When Elohim had made the world, as Master of the universe, He appointed chiefs over the several creatures, over the trees even, and the mountains,and the fountains, and the rivers, and all things which He had made,as we have told you; for it were too long to mention them one by one. He set, therefore, a malach as chief over the malachim, a spirit~over the spirits, a star over the stars, a demon over the demons, a bird over the birds, a beast over the beasts, a serpent over the serpents, a fish over the fishes, and a man over men, who is Moshiach Yeshua. But He is called Moshiach by a certain excellent rite of obedience; for as there are certain names common to melekim, as Ahashwerosh among the Persians, Caesar among the Romans,Pharaoh among the Mitsrayim, so among the Yahudaïm a melek is called Moshiach. And the reason of this appellation is this: Although indeed He is the Son of YHWH, and the beginning of all things, He became as man; Him first YHWH anointed with oil which was take from the wood of the tree of life: from that anointing therefore He is called Moshiach... \end{quotex}


We also learn that the schisms of the Judaic sects were work of the devil to obscure the truth, and that baptism was a Christian rite because man is ``taken from the waters'', as the chosen one is. More teaching is given about the internal disputes of Jewish sects, including Saul and John the Baptist. After James is thrown down the steps of the Temple by Saul, and persecution breaks out, Peter is sent to Caesarea to refute Shimon Magus, who is misleading many of ``our people'' by pretending to be the Primordial man come again -- the Messiah. Peter thus weaves divine history into a personal narrative of why Clement and he are there in the city.

\emph{More of Clement and Peter, to come... .}



\flrightit{Posted on 2012-08-04 by Logres}

