\section{When the King Enjoys His Own Again}

When Cromwell presided over the Puritanification (and therefore, the ultimate modernization) of England, the common people were not necessarily fooled\footnote{\url{http://en.wikipedia.org/wiki/The_World_Turned_Upside_Down}} as to what the end result would be. For them, it was mainly about the lack of charity and feasting, yet the beginning of the song makes clear that they understood something of what was at stake:

\emph{Why should we from good Laws be bound?}
The meaning here, is, Why should we be proscribed from being under good Rule?

This is a good question to ask, \& to meditate upon. Why, indeed, should (in either the inner or outer worlds) men be separated from good governance? Just as one may perceive that the overthrow of monarchies across the West did not result in the promised reign of sweetness \& light, so can anyone who wishes perceive that man is ``fallen'' and that this lack of assimilation to God does not produce freedom or liberty, but rather, disorder and misery. Why should I not be resurrected and regain control of my own body? Is this not a disorder? It is indeed. The King should be restored, and the royal path leads to the throne.

The Regal Path was the subject of correspondence\footnote{\url{http://www.fondazionejuliusevola.it/Documenti/Evola_Palamidessi.pdf}} between Julius Evola and Tomasso Palamidessi (who has been cited\footnote{\url{http://www.gornahoor.net/?p=3115}} here before). To read one of Palamidessi's pamphlets (published\footnote{\url{http://www.lulu.com/shop/tommaso-palamidessi/the-occult-constitution-of-the-man-and-the-woman/paperback/product-18794493.html}} by the Archaeosophical Society) is to be drawn into a current which unites esoteric Christianity, yoga, occult lore, \& ancient pagan mysteries. Just as he eschews a fundamentalistic Christianity (which stops at ``belief'', a ``belief'' in Sola Scriptura, the formal exoteric Church alone, personal conscience or conviction, etc.), so does he eschew ``mere Paganism'', which amounts to saying ``who needs God to have God?''.

\begin{quotex}
Christian esotericism is in its essence, the bringing to light of the Archaic tradition, and has always existed since Christ to today, even though the Church of Rome and the Protestants have never consented to its recognition.
\end{quotex}

Palamidessi believes that the distinguishing feature of a Christian archeosophy is the ``gradual'' initiation. The spiritual imprint of ``masters'' is used to a minimal extent (as a midwife), and (rather) the effort of the disciple is what becomes prominent. He mentions that both baptism and confirmation are rituals in which the ``subtle body'' (still impressionable) is imprinted with the power of the initiator. However,

\begin{quotex}
in the present condition of the official religions, also those endowed with Sacraments, they are incapable of guaranteeing (but maybe not of effecting?) the salvation of souls and their spiritual happiness; nor can anything concrete or decisive be achieved by the various societies of esoteric culture...
\end{quotex}

Instead, what is needed is the experiential knowledge of the Divine, gained through discrete, concrete practices which are well-tested and known, and also through the initiation from above which can operate at the behest of angelic forces and ultimately God, but which can be assisted horizontally by a legitimate and charismated ``master''. Tomberg addresses these possibilities (along with rituals, examples, historical cases, etc.) at length in his \emph{Meditations}. Gornahoor has made it very explicit here\footnote{\url{http://www.gornahoor.net/?p=4005\#more-4005}}.

Archaeosophy was his effort to (to some extent, perhaps as much as is possible or has been done) codify a vast realm of occult practice \& lore under the cross of Christ -- hence the Lotus-Cross Order. He saw himself as (essentially) doing what the Church could (and should) do.

Palamidessi makes clear that the ancient Church knew of such in the following remarks on Clement, Origen, and others (familiar territory for Gornahoor readers, but worth noting again):

\begin{quotex}
... in the beginning (of Christianity, he means) Jesus Christ had given his disciples the keys of the archaic tradition, that is of Archaeosophia, and therfore of the the truly total Ascesis that is the Ascesis biophysical, mystical in the superior sense, and initiatic... and what else could have been meant by men like Saint Basil, Father of the Greek Church, who lived in close contact with the initiated monks of the Orient, in the year 374 AD in Caesarea of Cappadocia, today Kaisarije in Anatolia, by saying: ``we receive the dogmas that have been transmitted to us by right and those that have come to us from the Apostles under the veil and under the mystery of an oral tradition. How could be diffused publicly that which non-initiates are forbidden to contemplate? ... that is exactly why many things were transmitted in an unwritten form... '' (\emph{Treatise on the Holy Spirit}, XVIII). And Clement in \emph{Stromata} I, chapter XII, specifies: ``Since the holy tradition could not be something common and public, at least if one realizes the greatness of its teachings, it is necessary to hide `this wisdom expressed in mystery', that the Son of God has taught us.'' After having recalled that among the Jews there existed oral teachings, after having reviewed Greek philosophy: ``Gnosis is a deposit which has been transmitted and has arrived to very few people: it is the `Wisdom' and was communicated orally to some Apostles who had received it from the mouth itself of the Son of God'' (Stromata VI, 7). In the year 203 AD, Clement was succeeded by his disciple Origen to continue the work of the master. In his polemical work \emph{Contra Celsum}, Origen confirms the existence of a secret doctrine within the Christian Church, divided into an exoteric or externalsection, and an esoteric or internal section. ``It is necessary to conceal the royal secret... '' (V, 19).

\flright{Excerpt from \emph{Pamphlet \#1, The Archaic Tradition and Foundations of Archaeosophical Initiation}}
\end{quotex}

The bishops and monks of the early Church were intimately familiar with the possibility and reality of what Palamidessi claims to teach, \& he goes as far as to note that the modern Church has largely lost the knowledge of her own teachings (which places us firmly in the Kali Yuga, just as Christianity emerged at the end of a miniature Kali Yuga within a smaller epicycle, around 30AD). Daniel's dream of the statue had come true.

Since Tradition \emph{is what persists}, the fact that the Church is either unaware, or else ignorant, does not obliterate the Pattern from Above -- Tradition will persist, and God needs neither a Church nor a ``people'' if they think they can forget Him, because ``out of these rocks He can raise up children of Abraham''. The ``Spirit blows where it lists''. Modern Christianity cannot avoid this collision with its own Scriptures, in the ``unmarked verses'' of its own Bible.

``He who comes to God\footnote{\url{http://bible.cc/hebrews/11-6.htm}} must believe that He is, and that He is a rewarder of those who diligently seek Him''. Contained in this verse is a prescription (or at least a permission) for esoteric practice, for it demands a knowledge of personal Divinity, a metaphysic of Being (what ``is''), and a theory of praxis involving devotion (\emph{bhakti-yoga}). Saint Paul himself refers to this when he tells his little Church that he yearned for them until ``Christ is formed in you\footnote{\url{http://bible.cc/galatians/4-19.htm}}''. There is also baptism for the dead\footnote{\url{https://en.wikipedia.org/wiki/Baptism_for_the_dead}}.

The formation of the Christ-body in the personality of the individual wishing to be ``saved'' involves the tripartite nature of man, the chakras (including the heart-chakra, where the Christian initiation is situated), \& the various ``bodies'' attached to one another in their ``fallen'' and un-assimilated configuration. Mouravieff discusses all of this, using diagrams and a science of numbers, in \emph{Gnosis}, as well.

It would be interesting to study further the correspondence of Evola with Palamidessi, but after going over several of his pamphlets in detail, it is clear that what this ``other Italian'' teaches is uniquely ``Christian'' in a way that Franz Bardon's system of magic (for example) is not quite; it is more ``fully'' Christian in a way that Steiner's evolutionary theories lack the metaphysical heft for; and it is more ``truly'' Christian than the doctrines of Max Heindel, who was connected with Steiner, and whose work had more to do with ``revelations'' than with initiations. Palamidessi also emphasized the harmony of magic, theurgy, and meditation, as well as the place of art and breathing exercises, in a very simplified (and possibly, safer) form.

All of these things lead me to conclude that he is a good place to begin esoteric practice with, especially for someone who is a Christian.

Tomasso lists many kinds of Ascesis. There are:

\begin{enumerate}
\item Physiological and Psychosomatic Ascesis (see Eliade's book\footnote{\url{http://www.amazon.com/Yoga-Immortality-Freedom-Bollingen-Series/dp/0691017646}})

\item Social Ascesis (see the Right Hand Path\footnote{\url{http://www.righthandpath.org/}})

\item Mystical Ascesis (Alchemy and the Saints)

\item Theurgical Ascesis (Magic\footnote{\url{http://www.scribd.com/doc/24362696/How-to-Attract-your-Desires-Eliphas-Levi-1854}} -- eg., Peladan, Eliphas Levi)

\item Magical Ascesis (here, we may place the pagan magicians)

\item Cosmic Ascesis (true astrology)

\item Sapiental and Initiatic Ascesis (the Royal Path)
\end{enumerate}

Palamidessi also delves into ``the three powers of the heart\footnote{\url{http://argivephoenix.blogspot.com/2011/05/three-powers-of-heart.html}}``, how each of the three centers of man can be saved, particularly the last and most dangerous one, the eros-dynamic soul, which contains the ``gravity'' in fallen man that will drag him away from God.


\flrightit{Posted on 2012-04-14 by Logres}

