\section{Hell and other lovely thoughts}
My plan this summer is to read \textbf{Dante}'s Inferno, but this time as initiatory work, rather than as a work of literature or the source of \emph{la lingua nazionale}. Dante insists that it is the beginning of a journey that he must make in his body, that is, the reader must participate in it and not regard it from the outside as a visitor at a museum views a sculpture. This is a journey everyone must make, dead or alive, although few will complete it.

It may explain the dream I had recently. I attended a funeral, where I was greeted by an old man I used to know, but never think about. He allowed me into a room, where a sick priest was lying in bed. I thought to ask him to hear my confession, but he was rambling incoherently and asking strange questions. Meanwhile, his foot kept appearing from beneath the sheets, at right angle to his body; it tried to grab my leg, as though it were a hand, in order to pull me in. I decided to leave, neither forgiven nor condemned. Perhaps it was Minos, trying to determine my place in hell.

\paragraph{Feminization of Thought}

I was invited to dinner last night by two dear friends on the occasion of the visit from New York of an old university buddy. After dining on Japanese food, where pleasantries were exchanged and our lives caught up, we finished our sake and the conversation got more serious. Somehow, we acknowledged the feminization of American ``culture'' and how many young men feel alienated (although not enough). I explained the difference between the male and female thought process. Women create a fictitious thought world which they then inhabit, and attempt to draw others in through constant chit chat and attempts at consensus building. When that fails, they resort to shaming and guilting. Men, on the other hand, use thought as a tool, supervening over experiential reality. To accomplish their ends, they understand the need for concrete motivations or else coercion, both subtle or overt.

Since Nietzsche, we can no longer deny this: there are different perspectives, all masking and underlying the lust for power. As an anti-egalitarian, I do not regard all points of views as equivalent, or all opinions as worthwhile; the reasons for this have been touched on previously and will be worth revisiting at a later time.

While there is no disputing facts, thought constructions, by their very nature, automatically invite dispute, arousing an opposition between the yes and the no. Unless a thought is anchored by reference to the material world via the senses, or to the supernatural world via spiritual, intuitive vision, it creates a web of illusion. Such thoughts latch onto each other in a pattern resembling chainmail. As it tightens, it obscures any higher vision, and a man becomes a slave to his own opinions.

\paragraph{The American Election}

As might be expected, the conversation moved onto the American elections. Of course, our view is orthogonal to all the current alternatives, and is impossible to explain to the unprepared. We see one side representing the Shudra caste, functional atheists, promoting egalitarianism, formlessness, and lack of differentiation. Opposed to them, is the remnant of the Vaishya caste, advocating a hierarchy based on financial resources, and following a degenerate religious form.

To those without a sense of transcendence, their spiritual instinct is often replaced by an emotional bond to their political position, the highest form they can acknowledge. Unconcerned with facts, they are satisfied by repeating predigested responses to any objection. This makes any dialog, in the Socratic sense, impossible; in fact, it could be driven by robots.

Those who have any serious interest in chess will know that the most common opening moves have been analyzed in depth. Thus, in a tournament, you will see the first several moves, sometimes even a dozen, take place in rapid succession. That is because the best opening moves have been memorized. It is only when that has been exhausted does the real game begins at a much slower pace. The same can be said about contemporary political discussions. There are memorized responses that are hurled with complete confidence that they are defeaters to every argument. This continues to the point where original thought must begin; since most find that impossible, the end game is usually an emotional tirade. In chess, there is an objective standard, so the delusion cannot continue forever. In dialog, with no chessboard or established rules, very few are willing to submit to the laws of logic.

\paragraph{Choosing Obama}

One man seemed completely bewildered by white blue collar men opposing Obama, since it is ``in their interest'' to support him. Either he, or the blue collar workers, are delusional. I have an opinion on that. Earlier in the day, I watched a discussion program on a news channel supportive of the democrats. Since the so-called left wing is souring on Obama, one commentator defended him. The commentator was writing a book and had been granted interviews with Obama. According to the commentator, Obama said all his programs, such as healthcare, were designed to help minorities. Unfortunately, he claimed, he couldn't come out and say that. But, curiously, neither will his political opponents.

\paragraph{Sufi Poets}

By then I had had enough, although it is instructive to observe the human thought process. Fortunately, our old friend was an admirer of \textbf{Hafiz} and \textbf{Rumi}, so we moved the conversation in that direction, at least the two of us did. As I was leaving, she promised to send me some recent translations of their poems. That I look forward to.


\flrightit{Posted on 2012-06-06 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Kamchatka on 2012-06-06 at 01:56 said:}

I just want know, ``Who is Hafiz?''. I know Rumi and He is one of my torches of my twiligth way.

\hfill

\texttt{Cologero on 2012-06-06 at 07:00 said:}

\textit{Kenosis and Hafiz}\footnote{\url{http://www.gornahoor.net/?page\_id=4324}}

\hfill

\texttt{Kamchatka on 2012-06-06 at 07:05 said:}

Oh I see... Persian sufi people and their works are not main source of me... 

\hfill

\texttt{Cologero on 2012-06-06 at 07:37 said:}

Wasn't Rumi a Persian Sufi?

\hfill

\texttt{Kamchatka on 2012-06-06 at 07:47 said:}

He was born in Belh, Afganistan and He had written in persian and spoken in turkish. ``Rumi'' name means Rome or Byzantium or Easter Roman Empire, at the very end days of that empire. ``Rumi'' nickname is given to him by other islamic and turkish people because of his last home city, Konya. His name is Mevlana Celaleddin-i Rumi, in fact.

When I talk about my sources, I mean Anatolian people and cultures... 

\hfill

\texttt{Michael on 2012-06-06 at 09:39 said:}

I am re-reading the Commedia as well--mostly because of this blog. Please keep us apprised of insights you run across as you make your journey.

\hfill

\texttt{Avery on 2012-06-06 at 23:19 said:}

Your perspective on American politics is quite enlightening. It was not long ago that America's conservative class was openly opposed to equality and direct democracy. Now, the value the conservatives want to celebrate most is ``freedom'' and legal equality, which seems like a very fast devolution, but actually makes sense for a country that purposefully rid itself of tradition. They are still fighting populism-- only the populism has become more absurd than it was a hundred years ago.

\end{sffamily}
\end{footnotesize}

