\subsection{Second Letter}

The Truth is too high to receive the least insult.

\begin{quotex}
It is unfortunate that we don't have Evola's letter to Guenon, although we can surmise what it contained. We see in this dialog, that Guenon is always the master. We have to agree with Guenon that Evola misunderstands certain principles that discolor his work in unfortunate ways. As for the inability to clearly distinguish the esoteric, or metaphysical/initiatic aspects from the religious/exoteric, Evola has created confusion in those whose understanding of Tradition is limited to Evola. It is not a matter of choosing or preferring one exoteric form over another. Hence, there are the mindless debates of paganism vs Christianity that are pervasive in counter-Traditional circles; the very fact of debating is itself an indication of the counter-Tradition.

It also makes Evola's understanding of the Middle Ages quite confused, since he admires that period as Traditional, yet fails to adequately grasp its spiritual foundation. Yet Evola makes it very clear that what is required is an inner transformation, so this corrective would not radically alter his overall project.

Evola apparently wasn't clear about the distinction between mysticism and the ascetic or initiatic path, although he referred often to it. For Guenon, they belong to different spheres, so there is no point in criticizing mysticism, if that path is proper for a given person. This is also related to the distinction between salvation, which is meant for the many, and liberation which is restricted to the few. Again, there is no value in disputing this question.

In practice, Evola's choice seems to lead to a dead end, viz. to ``ride the tiger". Yet that is not at all the goal of an ascetic, heroic, or initiatic path. Rather, it seems to me, a man should devote his efforts toward Liberation, the Supreme Identity, that is, an overcoming of the modern world, rather than an uneasy accommodation to it. The rest of the letter speaks for itself, although I do not know who ``P. A." is.

Added 22 July MMXX

(1) Re Guenon's aphorism: in the \emph{Act of Contrition}, the fear of hell or the desire for Heaven are considered imperfect motives. Make of that what you will.

(2) Re RTT: The comments were harsh on this point. The question is does it proceed beyond individual possibilities or not. 

\end{quotex}
23 February 1934

Cairo, Egypt

Forgive me once again for being so late in responding to your letter that I received with pleasure after such a long silence. But I have suffered from an acute eyesight weakness, and your letter reached me exactly at the moment when I was able to read it only after a very long time. The quantity of things of every type that had accumulated while I found myself in that impossible situation of working is such that, ever since, I have not yet succeeded in freeing myself from it and to regain that lost time.

I thank you for all your appreciations in regard to my works; and I think that in effect we can find ourselves in agreement, at least, on what concerns the conditions of the current world and the necessity of a return to the tradition and spirituality, if indeed it is still possible for the West, at the point in which things have currently reached.

As far as living far from Europe, I cannot perhaps precisely take into account certain tendencies; I must confess that I do not excessively trust in a ``renewal" that, as much as I know of it, remains up to this point very superficial and somewhat confused: above all, except for rare exceptions, it is about vague and poorly defined aspirations, and it is very difficult to say what will result from it. But what is certain, is that we notice in very general lines how the people are no longer so satisfied with their own modern ``civilization", and that some begin to doubt the ``progressive" pretense: as far as that goes, it is insufficient, nevertheless it is already at least something …

Regarding the problems brought up in your letter, permit me to tell you with great frankness that these difficulties appear to me especially to derive from the fact that you do not make a very clear distinction between the religious point of view of the one hand, and the metaphysical or initiatic, on the other. Whatever their relationships might be in certain respects, it is never necessary to confuse or mix them, since they refer to totally different domains, and they cannot consequently interfere with each other. The domain that defines religious truth belongs to what Hindu doctrine calls ``non-supreme" knowledge; it is sufficient to put everything in its place and in its order because there is no conflict possible. Above all, it is necessary not to forget that mysticism belongs totally to the religious ambit; any comparison is therefore not possible between mysticism and metaphysics.

The two ways, without considering the very relevant differences of their modalities, are not, in reality, absolutely marked out to reach the same goal; and the ``mystical union" is not the \emph{jivan mukta}, no more than ``salvation" is ``Liberation".

Everything that is religious, including mysticism, concerns individual possibilities, in the indefinite extension of which they are susceptible and does not go beyond them; on the other hand that is its reason for being. On the contrary, the reason for metaphysical realization is to proceed beyond [individual possibilities]; and this is why the one can serve as the base of the other. Nor was it the case for Christian esoterism of the Middle Ages, as it was always for Islamic esoterism; and, in this regard, I cited this aphorism that seems to me to be perfectly adapted to the argument: ``As long as a man desires Paradise or has fear of Hell, he will not be able to aspire to the least grade of initiation".

I must moreover bring to your attention the fact that the religious point of view is necessarily tied to certain historical contingencies, while the metaphysical point of view refers exclusively to the order of principles. To speak of ``multiple \emph{avatars}", is to stick to the domain of appearances; nevertheless, in absolute reality, they are ``the same thing"; the Christ principle is not multiple, whatever it can be made of its terrestrial manifestations or other types. The ``Mediator", according to all traditions, is the ``Universal Man", which is also the Christ; whatever the name by which he is called changes nothing, and I do not see what difficulty there can be in regard to this.

The ``ascetic" way would be, in its type, more comparable to the initiatic way of what is not mysticism, if only in that it implies a method and a positive effort. Mysticism, for its part, is instead totally the opposite because of it passive character. The ascetic way can therefore be a preparation for a realization of another order, much more that the mystical way, which would even seem even incompatible with that goal.

But I do not think on the other hand that we can assert that some of what passes beyond elementary religion is open to all; asceticism belongs only to some, and mysticism to some other. As to what is beyond the religious domain, it is obvious that it concerns a much more restricted number of persons. Whoever finds his fulfillment at a certain level would make a very great error to try to pass beyond it. That concerns the question of a necessary hierarchy, against which all sophisms of democratic egalitarianism are impotent, though many of Catholics themselves today unfortunately are affected by it: and there are perhaps still a few of them who even suspect it.

In regard to your objection inherent to the domain of pure intellectuality, is it quite certain that it is even what it has as a goal? In that case it is still necessary to make an essential distinction: the texts that you cite are revolts against profane knowledge, not against sacred knowledge; and we absolutely do not confuse what is simply rational with what is purely intellectual. When I speak of profane knowledge, I understand by it, naturally, everything that is philosophy; the less the spirit is blocked from all those things, the better, certainly, and from the initiatic point of view even more than the religious. It would be necessary perhaps to include also a good part of theology, insofar as it contains many useless subtleties and is of a still quasi-philosophical nature. In any case, everything that is discussion and controversy is of a purely profane spirit. That said, it is necessary to add that pure intellectuality eludes on the other hand the religious domain; this is another thing and it stands to reason that sentiment and action have their part in it. Yet again, it is necessary to put all things in the place that belongs to them, without allowing them any influence over a domain that is not their own.

Finally, pure intellectuality is in the same way indifferent in regards both to pride and humility, two opposed notions that are of a sentimental order in the same manner; those who pretend the contrary show clearly in that case that they do not have the least idea of what is truly intellectuality.

I see that you consider as valuable the incomprehension of P. A.; it would be somewhat difficult to find a more limited spirit than his. Moreover, what a fine way of defending Christianity by continuing to deny that its doctrine reaffirms a higher meaning, instead of the nonsense of moral and social character that one admits to often see in it! I don't see in what a similar vulgarity would presuppose the intervention of a superhuman principle; fortunately, I have for my part a better idea of Christianity than his.

It is sad to see how persons of this type look to diminish all that is higher than them … The Truth is too high to receive the least insult.



\flrightit{Posted on 2012-07-22 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Aghorable on 2012-07-23 at 04:38 said: }

``So long as a man desire Paradise or fear Hell, he cannot aspire to the least grade of initiation"

Cf. Evola's statement that the aspirant must be able to sustain ``severe blows upon one's self, without being hurt thereby."

P.S. Thanks for the articles and translations.


\hfill

\texttt{Mihai on 2012-07-24 at 01:30 said: }

``In practice, Evola's choice seems to lead to a dead end, viz. to ``ride the tiger" "

To ride the tiger doesn't have to be an ``uneasy accomodation to the modern world". It is a necessary step to take in the disordered, anti-spiritual world of today, for the person who cannot yet escape it, until the conditions for an effective initiation can be reached. There are not really many options for the western man, so riding the tiger can also be seen as a challenge of worth, of patience, and of serious orientation.


\hfill

\texttt{Mihai on 2012-07-24 at 06:08 said: }

PS: Who is the P.A. that Guenon refers to in the last part of the letter ?


\hfill

\texttt{Avery on 2012-07-24 at 07:52 said: }

Cologero doesn't know. Obviously it would have had to have been someone in the circles of both Guenon and Evola in 1934.


\hfill

\texttt{Cologero on 2012-07-24 at 08:25 said: }

That is the point, Mihai. If it is a necessary ``step", what, then, is the next step after it?

It has been 50 years since ``ride the tiger" was published, so it is time to move beyond. Certainly, there should be agreement by now about what is the opposite of the ``disordered, anti-spiritual world", but, due to the great influence of pseudo-Traditionalists, that is not the case. Unfortunately, there is a rather vocal segment who know nothing of Tradition except what they find in Evola.

That is why this letter is important. It is the only time when Guenon provides a clear critique of what is lacking in Evola's perspective. And Guenon's points are necessary and fundamental. Properly understood, they make most of Evola's points stronger, not weaker.

The task now is no longer just to criticize the modern world, but rather to describe what comes next in terms of order and spirit.


\hfill

\texttt{Mihai on 2012-07-24 at 10:27 said: }

``It has been 50 years since ``ride the tiger" was published, so it is time to move beyond."

For us, who live in the modern world, Ride.. is an important principle to understand and apply. It is about living in the modern world, participating in some of its activities because of necessity without letting this brake your will and right orientation.

This is most important from the spiritual point of view. Nowadays, all that is left to western man is a formalized and bureaucratic Christinity, almost entirely devoid of its deeper, intellectual elements. It is not easy to find one's path under such conditions. So until the path reveals itself, one will have to keep one's cool amidst the chaos and survive it without despairing. This is why I consider this to be Evola's most important work. 

Also, it is important to realize that Evola refined and deepened his views with the passage of time. However, he lacked a participation to a spiritual path, as far as I know he had no initiation, so it is obvious that he couldn't penetrate into the deepest meanings of symbolism and Tradition as Guenon did.


\hfill

\texttt{Cologero on 2012-07-24 at 12:44 said: }

I have been revealing a path, Mihai. Do you think I do this for fun and profit?


\hfill

\texttt{Mihai on 2012-07-24 at 15:25 said: }

I would never have thought of accusing you of anything like this, nor did I ever say anything that would hint such a thing, so I don't know why you made this assumption.

As for the path, I think it is necessary to clear up what I mean by this. I am referring to an active participation to a traditional spiritual path. This is a personal quest and the answer to it certainly cannot be found on the internet.

I think the website provides a starting point in seeking and an orientation.


\hfill

\texttt{Cologero on 2012-07-24 at 21:02 said: }

Transcendence, by definition, means to be above the forces of the world, whether destructive or positive. If you cannot abandon the world, you lack power. If you do not wish to abandon the world, then you are still attached to it.

But as a preparatory stage, the neutralizing of negative forces, otherwise called drinking poison, handling snakes, or transmuting lead into gold, is certainly necessary. But it is likewise necessary to understand it is still far from the ultimate goal.


\end{sffamily}\end{footnotesize}
