\section{Let no one ignorant of Geometry Enter}

\paragraph{Signs of intelligence} 

No one ignorant of geometry was allowed to study at \textbf{Plato}'s Academy. Topology is a contemporary and more rarefied form of geometry. Mastery of geometry demonstrates some ability to think abstractly, that is, apart from the objects of sense, as well as the talent to follow and formulate a logical argument. So if maths is the most basic of sciences, then what are the highest sciences? They are politics and theology, since they are the sciences of temporal power and spiritual authority, respectively, and hence the most sublime knowledge.

So we see that politics and theology could be properly understood only by the few. Our knowledge of man, society, being, and God are as exact and as certain as a theorem in algebraic topology. In the modern world, there is quite the contrary notion as we pointed out in an early post\footnote{\url{https://www.gornahoor.net/?p=37}}. We moderns seldom understand maths, yet we suppose everyone is entitled to discourse on the topics of politics and religion. Thus, they become mere matters of opinion rather than the objects of infallible knowledge.

\paragraph{The Middle Way}
Last week, John vacationed in South Beach. While there, he sent a postcard to his Mom in Boston describing the beautiful beach, the warm, calm ocean, the fine restaurants, and so on. His Mom was happy to receive the card and reading it brought her happiness. However, she had no direct experience of South Beach, unlike her son.

So, what does it mean to understand \textbf{Nagarjuna}'s Middle Way? First, a man must experience his own consciousness and ways of thinking. Then, the teachings of an enlightened being can make us appreciate its significance. But without the initial efforts, the teachings can only make a man confused. By observing how thinking one thought leads to its opposite, we can become frustrated with the constant interplay of irreconcilable opinions. That may lead a man, and very few of them, to seek something beyond the samsaric world. This is not the place to discuss the specifics of the Middle Way, but rather to make clear what it means to understand such a path. It is quite the opposite of glibness.

\paragraph{Shibboleths}


The \textbf{Book of Judges} describes a conflict between the Gileadites and the Ephraimites. The latter had no \emph{sh} sound in their dialect, so they pronounced the word ``Shibboleth" as \emph{sibboleth}. Thus the Gileadites were able to recognize Ephraimites by asking them to pronounce the word, or else meet their death.

The Gileadites had no interest in the metaphysical sophistication of their enemy, just their status as enemy. The spirit of the Gileadites is quite active today. The use of words ``as a test" is still in force: if you look a certain way and hold to certain opinions, then you are ``one of us". As such, it is a low level of discrimination. Yet, in a war, it may be the most important.



\flrightit{Posted on 2012-02-02 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Michael M on 2021-02-02 at 10:19 said: }

``The use of words ``as a test" is still in force: if you look a certain way and hold to certain opinions, then you are ``one of us". As such, it is a low level of discrimination. Yet, in a war, it may be the most important."

What does it mean to hold certain opinions? i.e. They profess a certain opinion when asked and will be held ``accountable" when acting in a contrary manner? or more so actually how someone functions in behavior?

I'll put forth the view of both are valid, as the outcome of such a test is the desired data. If one speaks a certain way and acts differently, his character is revealed. If one does not speak but his behavior is seen as in tune, his character is revealed. All is settled in both accounts. Thus it is most unclear with those who are secretive or vague and never put themselves out into the light.


\end{sffamily}\end{footnotesize}
