\section{The Meaning of the Past}

\begin{quotex}
This is third section of an article by \textbf{Guido de Giorgio}, titled The Instant and Eternity, first published in Diorama Filosofico in 1939.

I am constantly struck by so many people today who only perceive darkness, ignorance, or intolerance in the past. They tell me about dictators oppressively their people, who must be saved; some unfortunates who cannot afford this or that without a new government program; some marginal group that must be accepted; and so on. They long for some mythical future when all will be well with the world. Yet they do nothing for their own family, their own children and grandchildren; they assume their current life will continue, just as it is now. They cannot conceive that someday they will be oppressed, impoverished, or marginalized, with nothing to hope for. Those who cut their ties to the past don't have a future, because they don't even know what it is.
\end{quotex}

We said many times that ancient cultures are immobile or seem to be such; but that is precisely their greatness, that fundamental stability that removes all contrasts, integrates all rhythms into the central vein, to the \emph{traditional type}, who alone remains in the wholeness of its formative efficacy. That is why whoever intends to remain in the pure domain of traditional truth, always turns, logically, toward the past, to retrace the stages of certitude and add them to his experience. This, under that angle, is then recapitulative and conclusive: it is not an exterior repetition, but grants its rhythm to that which is none other than is own face, today still ignored, from now on found again and vivified. It is very difficult to explain certain things to those who hold onto dualist positions and who think that there is something besides the Truth, which is God eternally present. Truth: there only can \emph{one become what he is}, i.e., one transcends the sphere of human limitations to live the same beat of the infinite.

When we say \emph{ancient} we mean everything that is valuable, perennial, traditionally authentic in the past of the East and the West, whether it is about a distant or near, doctrinal or poetic, past: it matters little, provided that it reflects, in the truth of its expression, the great light of the Superworld. Beyond the Sacred Books, there is Poetry and sacred Art. There is finally all the forms of activity which, in the past, always relate to a truth of a higher order, be it in a modest utensil, and in the fabrication and the destination of objects of current usage. That past, as we mean it, and all those who are searching only for the truth of God should mean it, is life, creative rhythm, an inexhaustible deposit of wisdom which renews itself each time it is actualized by a new experience. But it is especially the reality of a vibrant life because vivified by the perennial breath of traditional energy. Moderns consider the past as a relic whose ruins they borrow and around which they prowl with the curiosity of photographers and archeologists: who, among them, accepts the past totally, assumes it in all its fullness, not in order to seize form it some fragments and exalt them, but in order to integrate it in one's experience of life while recapitulating it in a creative manner?


\flrightit{Posted on 2012-06-11 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Egemen on 2012-06-12 at 04:51 said:}

A nice dealing about the sacred past of civilization. I want to share a song and poem of one of the esoteric orders of the that is bekta?ilik (\url{http://en.wikipedia.org/wiki/Bektashi\_Order}), (\url{http://it.wikipedia.org/wiki/Bektashi})

And this is the youtube link of song;
\url{http://www.youtube.com/watch?v=m-aWLCGLCNo}
And these are the lyrics which will translate and explain by me later... 
\begin{verse}
Ondort bin yil gezdim pervanelikte\\
Sidk-i ismin buldum divanelikte\\
?ctim sarabini mestanelikte\\
Kirklarin ceminde dara düs oldum\\
Guruh-u naci'ye özümü kattim\\
?nsan s?fatindan cok geldim gittim\\
Bülbül oldum firdevs baginda öttüm\\
Bir zaman gül icin zara düs oldum
\end{verse}

I had travelled in space for fourteen thousand years (pervanelik means space or outer space and also pervanelik means planets; if a object turning around another thing, you can say for object; it is the pervane of other thing)

I had found the true meaning of your name (sidk-i ism) in madness (divanelik means madness or extacy)

I had drunk your wine at mestanelik (mestanelik means a place which people is getting transcendence state of mind)

I had forced to be brought to book at forty's court (in ancient asian tribes or little orders there was a court that call sometimes ``gathering'', which judges some problems of society. Forty's are prophets from Adam to Muhammed)

I had put in the forty's to my essence (güruh-u naci is the persian idiom of Forty's)

I had come and gone from the presence of human (or Adam Kadmon), (that means i had many experience of transfiguraiton, transformation or metamorphism... )

I had had form of nightingale and sing at the vineyard of Firdevs... (Firdevs is the highest level of the heaven; maybe as mentioned in Dante's Comedia)

i had fall to a garden for rose, for a moment... After all this explanations, there is a little story about origins of that song:

Before Islam and other monotheistic religions, ancient central asian nations had their pantheons, their celestial beings as elohim in torah. One of them called as ``Gabriel, Cebrail or Cibril, most probably after influence of Islam, and this Gabriel is humanlike creature of the god.

This human's breed was in danger, because of their green sun which Gabriel's home planet rounds around it (pervanelik), had come to hers ver end of life.

If Gabriel could not hade found a way to rescue his breed, they will have been destroyed with their planet. Hereby, Gabriel or Cibril had get a quest for a new planet to live on. And Cibril had prepared his spacecraft and hit on the roads of outer space, after that he had met with the god. And god asked to the Cibril:

``Who are you and who am I?''

Cibril answered was a little bit defiant;

``You are You and I am me''

God had not liked that answer and He punished Cibril by rampant traveling in outer space about fourteen thousand years.

In this punisment era, Cibril was getting into the madness and suddenly God appears and ask his same question again:

``Who are you and who am I?''

After all these hard years, Cibril had come to his senses and answers:

``You are the Creator, I am the creation''

After being excused, Cibril encounters a blue planet, they land to that planet and search of the lands and waters. They recognize, a creature that has similar genetically similar to them is living that blue planet. They decide to make a mergence between their breed and that creature. They had get some of these creatures with them and return their home planet.

At the Forty's Court; they had made some genetical experiments about mergence of breeds and they found a new species that are similar to them. After that, the creators of human generation had get that new species and return to the blue planet and so on as you know.

P.S. I listened that story in 1987 from a Bektasi person. After several years, I met with Zecharia Sitchin and his Works

\url{http://www.sitchin.com/}

After some research about that story, I found it is not only belongs to Bektashi order, and you can find similar stories in Dharma and Ancient Chinese Traditions... 

\hfill

\end{sffamily}
\end{footnotesize}

