\subsection{Third Letter}

\begin{quotex}
Looking back from 1949 to 1927, I wonder about some things. De Giorgio and Guenon often complain about health problems. In this case, there is no implication that De Giorgio's illness is the result of sorcery. Actually, Guenon brings up that topic only after having relocated to Cairo. Something was going on with him that he never made clear.

Another point is that Guenon always apologizes for being ``busy". Does anyone else wonder about that? I'm sure he was doing a lot of research, working on his books, raising a family, etc. 

Once again, the topic of a Western initiatic group comes up. There is no mention of that mysterious group of twelve. Perhaps Guenon discovered them subsequent to 1927.

My comments are inside footnotes.

\end{quotex}
Paris, 12 June, 1927

What must you be thinking of my silence? I am truly sorry for leaving your last letter without a response, and I blame myself so much more since you have not been feeling well. I want to think that your health will improve before long.

I went to the Bosse bookstore right away as you requested; they couldn't tell me if your money order had arrived, because it appears that that they had failed to do the research. Nevertheless, they tell me that you needn't worry about it and they will issue a refund.

The Italian translation of the King of the World has appeared; I recently received it. I don't know if you know M. Hackin\footnote{Specialist in Oriental art.}; we spoke of you last time that I saw him. He also sent me Stcherbatsky's\footnote{\url{https://en.wikipedia.org/wiki/Fyodor_Shcherbatskoy}. Probably Stcherbatsky's \emph{The Conception of Buddhist Nirvana}, published in 1927. Guenon took little interest in Buddhism, as this comment shows. However, note below re Avalokitesvara.} book, but until now, I never had time to read it. It has been a long time since that translation was made.

I saw Masson-Oursel\footnote{Masson-Oursel was a positivist in the Comtean tradition. To adopt Guenon's perspective would indeed have been quite a conversion, and its public revelation undoubtedly would have destroyed his reputation.} some time ago; he told me that he has been coming closer and closer to my point of view and that he took account that the orientalists had certainly committed many errors. After all that he has written up until now, I was somewhat surprised by this declaration. I ask myself if he will have the courage to say that publicly someday. He is truly too indecisive and fears being jeopardized by clear-cut affirmations.

Did you ever receive a response from \textbf{Tahar Kheireddine}?\footnote{A sheik greatly respected by Guido de Giorgio, and who apparently initiated him into Sufism in Tunisia.}

Here is the explanation that you had asked me on the subject of Avalokitesvara. This name literally means ``the Lord (Ishwara) seen from below"; but, in spite of the use of the past participle \emph{lokita}, it is most often interpreted as if it meant ``the Lord looking down". In reality, these two interpretations, far from being contradictory or excluding each other, complement each other perfectly, because there is somehow a reciprocity of relationship.

One could speak in this regard to a longing from the lower to the higher, prompting the descent of spiritual influences. It is that descent that represents the symbolism of Avalokitesvara; it is that which can be called ``Cosmic Love" (certain Muslim schools also use an expression with that meaning), while employing, of course, this word ``love" outside of every sentimental meaning of the word. One of the most used symbols, and which is found again almost everywhere, is the triangle whose apex is pointing down. The comparison which was sometimes made of Avalokitesvara to a feminine principle is also in connection with the same idea and the same symbolism: in India, the inverted or descending triangle is one of the signs of the Shakti.

The Rex Nemorensis which you speak to me about must be the same person who is often called the ``priest of Nemi" (I believe that Renan wrote something under that title [true]). I know that there exists a lake of Nemi, but I don't know exactly in what region of Italy it is found. In any case, it is Nemi which is the current name of the locality; but at its origin, that could not be a proper name, because it is simply the Latin word \emph{nemus}, which means ``woods", and especially ``sacred woods". This word is closely related to the Celtic \emph{nemeton}, and also, by inversion to the Greek \emph{temenos}, which means some ``consecrated place". The roots of these words, like those of \emph{templum} and \emph{sacratum} express principally the idea of putting apart, of separating from the profane world.

I haven't had the chance to see Mardrus' translations but based on what I know of him, I mistrust him a little; furthermore, I don't believe that the Koran is truly translatable.\footnote{The translation was commissioned by the French government. In general, any translation of the Koran is considered an ``interpretation" at best. The mistrust probably arises from Mardrus' less than faithful translation of \emph{1001 Nights}.}

Evola always sends me his journal; he would really like me to give him an article, but I have hardly any time for it. There are things in it of rather uneven interest, yet there is much less of ``philosophy" than I had feared. Moreover, Evola wrote me that he would soon leave aside that ``philosophical" form that he had adopted up until this point. I can only applaud that intention, but I fear the influence he underwent will persist in spite of everything, although he defends himself and claims to have had special reasons to adopt this language and mode of exposition.

Everything you tell me on the subject of magic is completely right; it seems that Evola takes this word in a somewhat different sense from what it normally has, and the way he employs it has some drawbacks. In the proper meaning of the word, it is, after all, only a science of the experiential order. It can in effect serve as the point of departure for something else, but that is true of all traditional sciences, such as they are, and I would even say willingly that they are especially made for that. Only the sciences conceived in the modern Western way can lead to nothing else and which are constituted in a way that do not permit the passage to knowledge of a higher order.

As for what Evola wrote you, I agree with him that there used to exist a Western initiatic tradition; but, unfortunately, I strongly doubt that it can be considered as still currently living. I certainly encounter, from time to time, the assertion of the existence of spiritual centres in this or that region of Europe, but, up until now, I have had no proof that that assertion is justifiable.



\flrightit{Posted on 2012-08-27 by Cologero }
