\section{Tradition \& the General Law}

Before beginning this post, several definitions of terms peculiar to Boris Mouravieff's work:

\textbf{A Influences}: sensate pressures and resulting spiritual tendencies, along with the cosmic hierarchies which keep them in ``tutelage'' (see St. Paul in \emph{Galatians 4})

\textbf{General Law}: the karma which keeps the men in bondage to the above from ever escaping the spiritual entropy of their own lusts. The General Law aims for the common good -- it ``uses'' mankind to accomplish aims which it itself, by itself, cannot accomplish. General Law is itself subject to Higher Laws, which are ``B Influences''.

What this essay will postulate is that submission to Authority (A Influences at first, then B influences, if man succeeds spiritually) is the only way according to Tradition that the ``masses'' can be helped (primarily, by avoiding a cataclysm of fiery judgement, but also by making ``quiet lives of godliness and peace'' possible through the organization of man. The Revolution, on the contrary, postulates that Authority does not exist, or (if it does or is ``felt'') ought to be ``broken''. Thus, a third definition:

\textbf{Revolution}: the idea/ideal that A Influences (and by extension, B Influences as well) can be ``broken'' rather than transcended or undone. Mouravieff takes pains to explain that ``The Great Game'' means that A Influences are legitimate, \& that they can only be overcome through purity of heart (forgive your enemies, turn the other cheek, etc., all of which has esoteric meaning as well). It is not possible to defeat Satan by reifying Evil as Authority, and attacking Authority. Saint Paul concurs: ``the weapons of our warfare are spiritual, not physical''. ``Love the brotherhood, Fear God, Honor the King''.

So a reading of Mouravieff's Gnosis\footnote{\url{http://www.scribd.com/doc/4549935/Gnosis-Book-I-Boris-Mouravieff-Exoteric-Cycle1}} makes clear that Christian Esotericism is not an \emph{ersatz} for those wishing to avoid other traditions or authority, nor an escape for those who want to remain purely exoterically Christian, without ever advancing (or preparing to advance in the postmortem state). One cannot substitute political ``Revolutionary'' progress in external matters for inner development. To explain this, he notes after a long discussion of man's ``three centers'' (head, heart, motor) and how they inter-relate with the higher centers which remain (for him) un-actualized , that ``practically speaking'' for most men, the only break upon their unbridled self-domination of the false I of their mini-personality is actually the ``links'' or ``shocks'' which this personality induces upon the other parts of him which are (at the moment) not acting, or not directly involved in the instantiation of the false I. Mouravieff then links this with the General Law -- in the absence of self-government, men come under the ``General Law'' (which he later specifies is the Adversary, cosmologically). So, since the ``fear of God'' is missing in these people, their governments reflect the ``brakes'' or links within their own \emph{chakras}, which are all that is holding them back from
\emph{publicly} proclaiming themselves to be as God. Therefore, strong

government is required to hold them in check. Penalties and checks are required to restrain them, because (and he is clear on this) ``morality alone'' cannot restrain the extreme and anarchistic tendencies of the unbridled false I/personality. Mouravieff is saying that Revolution by itself (and that is the definition of Revolution) is ``Hell on earth''. The Revolution destroys little Satans, \& sets up even bigger Ones. How is this Progress?

This is precisely the ancient, traditional doctrine of government, endorsed by none other than the apostle Paul in Romans 13, when he proclaims the secular arm or Caesar to be the ``sword of God'', whose purpose is to bring terror to evildoers. Christ taught this also -- ``render unto Caesar''.

Bondarev's interpretations of Steiner are accurate here, in that the Counter-Initiation will seek to proclaim an Anti-Christ, which will counterfeit the Kingdom of God and the integrity of Tradition, falsely liberate man, and suck all of Creation (or at least the earth) into a black hole (since the Lie can only live until the ``fullness of evil'' is reached, after which, it will implode). Mouravieff teaches that Peter warns against this in his own epistles -- either redemption (the rise of a Christian elite, or ``salt and light''), or another cataclysm.

It is not difficult to connect this with either the French Revolution or with the modern ideal of Progress: in fact, Mouravieff explicitly states in the early pages of \emph{Gnosis 1} that ``progress'' (as such) can have no absolute meaning at all, in the slightest. In fact, there is no ``progress'' in material affairs, when measured by the yardstick of true Enlightenment. Material man is subject to Death, and this is all. \emph{The Earth returns to the earth}, says Scripture. The ``straight line'' of Progress goes against the Logos, because only the spiral (in accordance with the Rays of Creation) can resist entropy, which is Christ's mission. Tomberg would call these ``vertical energies'' which transcend the Serpent's eternal horizontal circle of purely physical reality. Satan offers man the chance to ``die'' by refusing self-government and embracing Revolution (Lust). Thus, he promises life without Life, life through illusion: ``sorcery-States''. Man can, therefore, stay as he is, as he wishes to be and remain. This is an illusion, the General Law will still win -- thus man is eternally frustrated in ``Revolution''.

It should be un-necessary (here) to connect these teachings with the Anti-Christ. That is not our concern for now, instead to defend the idea of a Christian knighthood. In fact, ``Superman'' is a term which Mouravieff says applies properly only to a ``new elite'' (pages 185ff in \emph{Gnosis 1}). If the false elite succeed in spreading this corruption (and find fertile ground for it), the Scriptures promise a baptism of avenging fire for the earth, rather than a redemption. Christ (says Mouravieff) warns of many wolves and false shepherds.

It is not hard to draw the conclusion that almost every trend in the modern world (I say almost, because Mouravieff also makes clear that the ``transmissions'' of B influences from God-ward towards man never cease, nor does Providence's spirals of upwardness) is towards the sensate or ``New Man'' of the Revolution. In Mouravieff's terms (as we shall see) this is an evolutionary ``dead-end'' (Tomberg) or a ``one-sidedness''.

Mouravieff's work, when read alongside of Evola's treatises, or without them, should point us towards the idea that our stand on behalf of mankind has to be made, and can be made, only upon the path that has been revealed -- that of individuals who attain higher levels of Being which render them immune to the general process of A influences and the Counter Initiation. The other option is a ``Progress'' that substitutes for legitimate victory over A influences (the Serpent) \emph{through collective slavery to the endless cycle of A influences in their totality, a totality which still amounts to a Zero}. The General Law, in fact, rewards those whose ``go along''. Brilliant careers, much money, fame, fortune, good reputation, etc. Their esoteric energies (teaches Mouravieff) are preyed upon by higher powers in order to ``feed the Moon'' (evolve a new world for the common good, which will one day be inherited by the ``meek of the earth'' who resist A influence and opt for personal revolution over exoteric activity).

This is the reason that Donoso Cortes and Oliviera, as well as Taparelli\footnote{\url{http://bonald.wordpress.com/2010/07/19/taparelli-dazeglio-forgotten-founding-father-of-the-right/}}, van Prinsterer\footnote{\url{http://en.wikipedia.org/wiki/Guillaume\_Groen\_van\_Prinsterer}}, \& also even someone like Althusius\footnote{\url{http://en.wikipedia.org/wiki/Johannes\_Althusius}} should be studied as models for practical men of action seeking to understand against what they are struggling. Evola is useful for the very same reason.

And what do these good gentlemen have to say?

\begin{quotationx}
We are living in a condition of permanent revolution ... revolutions are here to stay and will grow much worse in scope and intensity unless men can be persuaded to return to Christianity, to practise its precepts and to obey the Gospel in its full implications for human life and civilized society. Barring such a revival, the future would belong to socialism and communism, which on this view were but the most consistent sects of the new secular religion. To Groen, therefore, the political spectrum that presented itself to his generation offered no meaningful choice.

In terms of his analysis, the `radical left' was composed of fanatical believers in the godless ideology; the `liberal centre,' by comparison, by warm believers who warned against excesses and preached moderation; \textbf{while the `conservative right' embraced all those who lacked either the insight, the prudence, or the will to break with the modern tenets yet who recoiled from the consequences whenever the ideology was practised and implemented in any consistent way}. None of the shades or `nuances of secular liberalism represented a valid option for Christian citizens.'' Groen called for a rejection of the entire available spectrum of political positions, calling for a ``radical alternative in politics, along anti-revolutionary, Christian-historical lines 
\flright{\textsc{Harry Van Dyke, Groen van Prinsterer}, \textit{Lectures on Unbelief and Revolution} (1989, pp. 3-4)}
\end{quotationx}


Here is the same thought, from Clement Maria Hofbauer\footnote{\url{http://www.catholicism.org/clement-maria-hofbauer.html}}:

\begin{quotex}
``Almost as a postscript to the heavenly warning issued at Fatima in 1917, Saint Maximilian Kolbe, two years later, reviewed the three Great Evils of the latter times, noting: ``In 1517, the Protestants rebelled against the Church; in 1717, the Freemasons rebelled against Christ; and, in 1917, the Communists rebelled against God.''

In a single sentence the Polish martyr had exposed common origin and natural succession of each of these Apocalyptic nightmares.
\end{quotex}


What Christian esotericism (Tomberg \& Mouravieff) does is offer a way that explains this in terms that most modern people are capable of understanding and accepting, at least in their saner moments and if they are trying to understand (granted, which assumes a lot), thus re-linking exoteric Christians to the Tradition. \emph{The peculiarly Christian hope that exists here is to make the process less dangerous}, both in terms of premature or unnecessary rejection or going astray down a side alley. Whatever else it does, it reaffirms Tradition's \emph{veredictum} upon the French Revolution -- this was not (at least) a permanent way to help the ``masses''. The ``masses'' are to be trained and educated into being a ``folk'' and a ``people''; the creation of the ``masses'' was itself a diabolical act, and no good will come of it, either for the false elite who did it, or those who wish to accept it as a \emph{fait d'accompli}.

Many other points throughout \emph{Gnosis 1} make this clear:
\begin{quotex}
As for the ordinary man of the new era, he will continue to live outside the Truth, as he does today, as long as he can dissimulate his thoughts. The latter will however be easily read by any person who has attained a degree of culture which can be compared, all proportions maintained, to those who give University teaching today.
\end{quotex}


This reading of thought will no doubt be intrusive \& uncomfortable -- it might even allow an ``upper hand'' (in many maneuvers) by the ``elite''. And what is the implication of knowing the ordinary man's thought? A countervailing action will be taken, at the least, to make provision for his duplicity, and perhaps it will simply be made known what he is. Is this in accord with ``human rights'', as understood today? It is not.

Mouravieff concludes, \& does not hesitate to speak of an elite:

\begin{quotex}
Esoteric Tradition teaches that any civilization is none other than a projection of the consciousness of the `I's of elite man onto the exterior world. The `I' of the intellectual already differs from that of a knight. In the coming civilization, characterized by Art inspired by the sacred, the elite man will possess an `I' consciousness altogether different from the three preceding epochs. He will have, as has already been said, the consciousness of the real `I', of a permanent and unshakable `I', not that of the personal, unstable and composite `I' which our times accept and glorify. Thus the edifice of the future civilization will not be built by the elite on sand, but on the rock14 of the onsciousness of the real `I', the spark.
\end{quotex}


Mouravieff even puts his finger on the problem of the ruling class of our time:

\begin{quotex}
For man 3 in a `world 3', this is much more difficult. For his mental16 constitution, reinforced by education, instruction, and the intellectual environment, ends by making a perfectly onesided being of him. This is the underlying cause of the weakness of the governing class, which today is not able to stabilize and equilibrate the life of human society even though technical progress offers all the necessary material means to reach this end.
\end{quotex}


\begin{quotex}
Man 3 (intellectual man) lives in an excessively intellectual age. Thus, we have the global age of the manipulator and technocrat, the best of whom are undoubtedly aware of what is happening. It is their one-sided (my noted added: one might say, in Tomberg's words, ``evolutionary dead-ended'' development) which propels them to power \& renders them an obstacle to world peace at the same time.
\end{quotex}


Man 2 (emotional or heart-driven man) is at a disadvantage, and cannot attain power, but the Ray of Creation is turning the wheel back to the heart, in order to ``recalibrate'' human nature. In the coming Age, those who rule will be those who are obviously superior (a spiritual elite). Even the masses will turn to them, just as they turned to the intellectual to end the age of rule by overt physical force. The Revolution is not eternal, Progress (as Mouravieff also states) is not a straight line. Nothing in Creation is a straight line, because straight lines attain their goal, and thus come to a natural end.

Here is a warning for those who want to ``go Modern'':

\begin{quotex}
It is the same with the woman who, in this modern world, is endowed with refined emotions, and finds herself in conditions which permit her to realize rapid progress on the esoteric plane. In fact the tendency to develop the intellectual qualities in our civilization favours the equilibrium of her centres, on condition, however, that even if dazzled by science she does not lose her feminine emotionality and does not become too calculating. For her, this preservation of her femininity is an obstacle, a competitive test by which selection is automatically made. She must beware of acquiring a masculine mentality and identifying with this. A male mind in a woman's body excludes the possibility of esoteric development. This type of woman is unfortunately widespread in our days, as is that of the effeminate man, representing what the Tradition calls the neutral sex. Union between these people who have thus deviated from the normal represents the opposite pole to the Androgyne, the summit of deified human power. The Kingdom of God is closed for them.
\end{quotex}


The Kingdom of God is closed to them. Tomberg makes precisely the same point, more gently, when he warns of becoming over-developed in one aspect, in \emph{Meditations}. The only way, according to Tradition, to overcome Authority and General Law, is to gain victory over the Self. Any other path sets the stage for impurity of heart and delusion, and an actual triumph of A Influences \& General Law.

\emph{To be continued, with more on ``General Law'' and overcoming it... }


\flrightit{Posted on 2012-06-16 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{apeiron on 2012-06-16 at 11:14 said:}

A good article on the reactions we often encounter from Men who were witnesses in their day to the early onslaught of Satanic decline. While they offered hints and examples of how they lived and brilliant analizations of our circumstances, they have also left us as lone soldiers without an army. No impermeable institutions/orders have lasted their passing on.

``The only way, according to Tradition, to overcome Authority and General Law, is to gain victory over the Self.''

Now what of the disciplines, actions and practices to prepare for the postmortem states, to overcome the self and to attain the kingdom of god? No one here has a word to say about these. If you do, then please state them, because it appears that Tradition today is studied, contemplated and over-rationalized from every angle other than lived by the most important and foundational issue at hand: What actions to take.

\hfill

\texttt{Logres on 2012-06-16 at 13:33 said:}

I understand the frustration, yet unfortunately, it is a lot easier to ``do'' what is wrong than what is right -- therein is the rub. Mouravieff, if you interested, has many instructions in Gnosis for spiritual exercises, including postures, breathing, etc, as well as directions for discerning A influence in your life. This is the first, most vital element -- ``detoxifying''. Only then can one act with comparative safety of attaining the goal. He also offers guidelines for marriage, as well as explaining the reason behind differing castes and characters of men. This is in addition to a cosmology interpreted by Tradition (Tomberg also has small hints and practices buried in Meditations, including how to use the sign of the cross). However, all of these practices are geared towards preparation -- they are preparatory, and thus, not on a grand ``world saving'' scheme or in the nature of direct action. Their intent is merely to ``clear ground''. If you interested in more ``direct'' rituals, you should read Magic:

\url{http://www.amazon.com/Introduction-Magic-Rituals-Practical-Techniques/dp/0892816244}

But read it in conjunction with Meditations or Gnosis, because it has spiritual dangers. Personally, I would say that almost all ``action'' in today's world is necessarily ``hidden'' or ``occult'' -- otherwise, it invites the retribution of General Law on a grand scale. Are you prepared for that?

\hfill

\texttt{Logres on 2012-06-16 at 20:25 said:}

Apeiron, your question is a fair one. Mouravieff actually teaches that faster means are necessary, \& in Volume III (which I haven't read lately, I finished the first two) talks about ``polar beings''. The situation is unsatisfactory; I'd be interested in what Cologero or others have to say.

\hfill

\texttt{HOO on 2012-06-17 at 14:50 said:}

The masses shall return to the tortured bowels of the earth from which they came, they malformed subhumans. Begone, and good riddance. Now Godspeed the First Men.

\hfill

\texttt{logres on 2012-06-17 at 22:39 said:}

Actually, the ``masses'' will never do anything good on their own, or even with leadership, until they are made into a ``folk'' or a peuple or a commoner. But if there is an elite that is ready when Rome begins to burn, then individuals from lower castes will be drawn to them.

\hfill

\texttt{Cologero on 2012-06-18 at 12:25 said:}

\emph{The masses shall return to the tortured bowels of the earth from which they came, they malformed subhumans. Begone, and good riddance}.

This is the sort of ersatz machismo that passes for ``virility'' really has no place in Tradition. The Traditional organization of society must be understood as organic, with the different segments interrelated in a hierarchical manner. The Kshatriya rules and fights not for their own end, but rather for the whole. Disdain for those who butcher our meat, bake our bread, mow our lawns, or care for our children needs to be replaced by the ideal of magnanimity.

The ``masses'' are driven by the passions of eros and thymos, unchecked by reason. Without the spiritual authority to provide that check, their fate is the first rings of hell. The deeper levels are for those who act out of deliberate malice, i.e., the higher castes who know better.

\hfill

\texttt{scardanelli on 2013-04-16 at 10:30 said:}

After getting a bit further into Mouravieff while also learning a bit about Orthodoxy, i'm curious as to how those in the Orthodox church react to Mouravieff. Is his work seen as basically in line with the church or as heretical new age speculation? I suppose i'm wondering how those in the church view his claim that his work represents a synthesis of esoteric knowledge and practices found scattered throughout the writings of the Church Fathers. It seems that many Orthodox writers strongly recommend the help and advice of a spiritual father from within the church. Would the work of Mouravieff be something to discuss with this spiritual father, or would it be frowned upon and better kept to oneself? Is this perhaps an instance where one would remain silent lest he provoke the General Law?

\hfill

\texttt{scardanelli on 2013-04-16 at 10:46 said:}

I know that one must take into account the differences that exist between the esoteric and the exoteric, but it seems that the exoteric Orthodox church has a well developed esoteric and ascetic tradition, and i'm curious as to how this tradition regards Mouravieff.

\hfill

\end{sffamily}
\end{footnotesize}

