\section{Clement’s Initiation: Peter and Shimon Magus debate}

This is a continuation of a previous post\footnote{\url{https://gornahoor.net/?p=4533}}. My source document is here\footnote{\url{http://www.scribd.com/doc/13561289/-The-Authentic-Peter-The-Preaching-of-Simeon-Kefa-from-the-Journal-of-T-Flavius-Clemens-Clement-The-Clementine-Recognitions-Recognitions-of-Cl}}, and I have already noted the problems with dates -- it is part of the Clementine literature, involving at least two Clements, but its contents are noteworthy.

Simon Magus is an under-appreciated character, even though modern sensibilities have tried to paint him as someone who fights poverty or performs tricks as good deeds -- instead, he is a pivotal figure in the history of the Church.

To continue, Peter (Simon Kefa) is recounting the Apostles' arguments with certain Jews, in which Matthew and Phillip both speak, along with James the Just:

\begin{quotex}
Then a certain Prush, hearing this, chided Philippos because he put Yshua on a level with Moshe. To whom Natanyahu bar Talmai,answering, boldly declared that we do not only say that Yshua is equal to Moshe, but that He is greater than he, because Moshe was indeed a navi, as Yshua is also, but that Moshe was not the Moshiach, as Yeshua is, and therefore He is doubtless greater who is both a navi and the Moshiach, than he who is only a navi. After following out this train of argument, he stopped.
\end{quotex}

Here we see (early on) the Christian claim to primacy which caused such trouble within the religion of Judaism (and in Rome). However, it was also more complex than this, as Peter will explain. James the Just actually argues that Jesus ratifies the angels, rather than the other way around, and Mathias (in place of Judas), claims that (in any case) the Messiah is worthy of great love, no matter what one thinks. Each apostle makes a uniquely personal and uniquely intelligent argument:

\begin{quotex}
After him Yosef bar Naba, who also is called Mattityahu, who was substituted as a Shliach in the place of Yahudah of Keriot, began to exhort the people that they should not regard Yshua with hatred, nor speak evil of Him. For it were far more proper, even for one who might be in ignorance or in doubt concerning Yshua, to love than to hate Him. For YHWH has affixed a reward for loving and a penalty for hating. For the very fact, said he, that He assumed a Yahudai body, and was born among the Yahudaïm, how has not this incited us all to love Him? When he had spoken this, and more to the same effect, he stopped... 
\end{quotex}


The apostles argue that the Messiah is heaven-born, otherwise, John the Baptist would truly be the greatest born of women, and thus Messiah, rather than Jesus. It's a very interesting look inside the ``sects'' of Judaism at that time. Indeed, Gamaliel is said to secretly be a follower of Christ, who remained within Jewish ranks in order to moderate their fury, which he does, giving his famously pragmatic speech ``If these things be of God... ?''. James proceeds, and has almost convinced everyone to receive baptism when Saul intervenes with agitation and violence, and James is cast down the temple steps and left for dead.

Peter says he fled to Jericho, then Caesarea, and intends on going to Rome following the debate with Shimon Magus. No women are present at the debate, and Peter opens by explaining his practice of night-time meditation, which Clement approves of:

\begin{quotex}
I believe, therefore, that you also have acquired the habit of wakefulness, as you state; and you have wished at a fitting time to explain this to us, that we also may not grudge to throw off and dispense with some portion of our sleep, that we maybe able to take in the precepts of the living halakah. For when the food is digested, and the mind is under the influence of the silence of night, those things which are seasonably taught abide in it.
\end{quotex}


Peter wants to know what character Shimon is, that he might know how to argue without casting pearls before swine. Niceta warns them (he was once a co-worker with Shimon):

\begin{quotex}
but because I well know the impieties of this man, I think of your reputation, and at the same time the inner-beings of the hearers, and above all, the interests of the truth itself. For this magician is vehement towards all things that he wishes, and wicked above measure.
\end{quotex}


The description sounds like a high ranking initiate:

\begin{quotex}
This Shimons father was Antonius and his mother Rachel. By tribe he is a Shomroni from a village of the Get tones; by profession a magician yet exceedingly well trained in the Greek literature;desirous of glory, and boasting above all the human race, so that he wishes himself to be believed to be an exalted power, which is above YHWH the Creator, and to be thought to be the Moshiach, and to be called the Standing One. And he uses this name as implying that he can never be dissolved, asserting that his flesh is so compacted by the power of his divinity, that it can endure to eternity. Hence,therefore, he is called the Standing One, as though he cannot fall by any corruption... \end{quotex}

Shimon has replaced Dositheus\footnote{\url{https://en.wikipedia.org/wiki/Dositheos\_\%28Samaritan\%29}} as head of the thirty, and taken Luna as a consort; he also uses the dead spirit of a boy to accomplish his ``miracles''. In fact, he even explains to Niceta how it is done, by adjuring Malakim to allow a spirit to come back. Peter is moved to tears, and says that spirit of the boy of air is not what it seems to be:

\begin{quotex}
But, in truth, he is deluded by demons...
\end{quotex}


\emph{Incidentally, Bo Yin Ra claims that ``lemurian physical entities of putrefaction'' use people's souls to create illusions of human spirits for their own energetic purposes, in perfect agreement with Peter}. Peter answers objections to God's goodness in exposing man to the Kali Yuga, when almost all men are lead astray -- he does so simply by teaching that God permits it to reveal who are truly His. \emph{Clement is asked to leave, prior to Shimon's actual arrival, because he has not yet been baptized, and Peter needs to enter God's presence for prayer.}

Shimon arrives and challenges Peter, refusing \emph{Shalom}. Peter attempts to show what he meant: that the inquiry would be peaceful and quiet, not violent and contentious. Shimon shoves this aside, and says Jesus came to bring a sword, not peace; Peter joins battle and quotes the peacemaker beatitude. The conflict begins in earnest. Peter ends up saying ``do not rashly take exception to things you do not understand'' quite a lot. Shimon shows his hand, and argues that Christ contradicted himself (after it is clear that Peter actually has lived with Jesus long enough to know what he is talking about). They swap rhetorical acrobatics over Jesus' saying about ``dividing a house'' and Peter comes off as very well-spoken, although not as naturally so as Shimon, who appears to be merely squabbling over rhetorical terms. In a somewhat ``modern moment'', Peter tells Shimon not to speak of what the learned or unlearned are willing to believe, \emph{but what he Shimon really thinks of the navi}, Christ, whose Gospel Peter is compelled to teach authoritatively. Shimon presents his creed of many elohims, with one above all: ``\emph{elohim hagadol of elohim}``. Shimon teaches that the Jews were ignorant of him, and Peter insists that the Creator YHWH is indeed this very Elohim of Elohims.

We are in the middle of the birth of Gnosticism, here.

Peter answers by claiming that all \emph{malachim} (godlike angels) are messengers to the tribes of men:

\begin{quotex}
For every tribe has a malach, to whom YHWH has committed the government of that~tribe; and when one of these appears, although he be thought and~ called elohim by those over whom he presides, yet, being asked, he does not give such testimony to himself.
\end{quotex}


The Bible is not anti-polytheistic, but henotheistic\footnote{\url{https://en.wikipedia.org/wiki/Henotheism}}. \emph{See Galatians: angels are tutors}. Peter claims that the \emph{malachim} are appointed of God, and sometimes a mortal man is chosen instead, but that the worship that goes to them is carried to God. Of course, Shimon thinks this is a delusion: there is a hidden Elohim that no one knew until now; YHWH is deluded, self-deluded, into thinking He is the true Creator; only the enlightened ones can escape this delusion and go to join the true Elohim. He cleverly quotes Christ -- ``No man has known the Father, but the Son... ''. Moses, then, and all the rest, were ignorant of true Elohim. When Peter points out that Shimon doesn't believe in Christ's sayings, and that Moses prophesied of Christ (and therefore, accepted Christ's pre-eminence and was not entirely ignorant of truth as claimed), Shimon points out that ``having a Son'' involves the Creator in passions. Familiar ground, but freshly done in this exchange. Shimon (like many Gnostics) is very slippery here. Or subtle, depending on your loyalty. The rest of the debate is quite interesting, and very complex, very rhetorically brilliant on both sides (another reason I think the letter is genuine). It prefigures the great Gnostic-Christian divide of those early centuries quite well; this encounter may have symbolically actuated the great divide. Suffice it to say there are many strokes of genius by Peter as well as Shimon; Peter asks why the immortal, invisible New Power does not confer a new sense on men to recognize him, and if He has given it to Shimon, then why not to them? And can Shimon read their thoughts, since he can read God's? Predictably, Shimon argues that he learned from the Bible's own story that the ``Creator'' cannot be good -- He made a manifestly imperfect world. Thus, the Problem of Evil undermines faith in the ``Creator''. Peter mocks the rejection of Torah on the basis of Torah, but Shimon frames the \emph{Euthyphro} problem brilliantly and classically:

\begin{quotex}
But even if Torah had not given indications from which it might be gathered that the Elohim who made the world is imperfect, it was still possible for me to infer from those evils which are done in this world, and are not corrected,either that its creator is powerless, if he cannot correct what is done amiss; or else, if he does not wish to remove the evils, that he is himself evil; but if he neither can nor will, that he is neither powerful nor good. And from this it cannot but be concluded that there is another \emph{elohim} more excellent and more powerful than all. If you have aught to say to this, say on... .
\end{quotex}


Peter of course does say on. He not only counters with a very traditional statement about choosing wise instructors:

\begin{quotex}
O Shimon, they are liable to conceive such absurdities against YHWH who do not read Torah with the instruction of masters, but account themselves teachers, and think~that they can understand Torah, though he has not explained it to them who have learned of the Master.
\end{quotex}


... but uses a \emph{reductio ad absurdum} to show that the Ineffable light lies under the same charge brought against the ``Maker'' or ``Architect''. They are now at the very heart of the matter, metaphysically speaking, and all those who wrestle today\footnote{\url{http://www.scribd.com/doc/57492796/Gnostic-Fragments-by-Nimrod-de-Rosario}} between accepting Gnostic doctrine or Christian tradition (or any other) would do well to ponder this collision, so vividly painted in the ancient literature.

We will end this post with Peter's objection to the Ineffable God, which is one of the strongest I have seen:

\begin{quotex}
Whence that new power of yours is not only found liable to a similar charge, but even to a worse one, if, in addition to all these things, it is believed to be, when it is not. \emph{For He who created the world, His existence is manifest by His very operation in creating the world, as you yourself also confess. But this power which you say that you alone know affords no indication of itself, by which we might~perceive, at least, that it is, and subsists.}
\end{quotex}


\emph{More to come later, about the continued journey of Clement with Peter, and the initiation of Clement, which Peter says he did ``moved by certain unknown motions of the heart''.} Also, the notes of Tomberg on Cyprian the Mage\footnote{\url{http://en.wikipedia.org/wiki/Cyprian\_and\_Justina}} are worth reading --- the magic that was being done was not a ``joke''.


\flrightit{Posted on 2012-08-11 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Gabe Ruth on 2012-08-13 at 15:43 said:}

Thank you for the posts.

I mentioned it in a recent comment so I hesitate to bring it up again, but have you read A Voyage To Arcturus? That is a vividly painted portrait of Gnosticism, and I truly wish I'd never read it.

The question those who wrestle with these options must answer is whether you would give your allegiance (for whatever that's worth) to a good but not all-powerful God if there was a higher power that was not good, and whether it would make a difference if the non-omnipotent good God was your creator. I've never been particularly worried that atheists are right, but the gnostics trouble me, because their answer to the problem of evil seems so reasonable.

This piece also got me to thinking about Voegelin's use of the term gnostic, and a way to unify his meaning with the historical gnostics: they both posit a way of being, a new truth, which is wholly opposed to everything that we see around us and have inherited from our ancestors, in fact casting the old ways as completely evil, the source of the evils that we now experience.

\hfill

\texttt{logres on 2012-08-13 at 16:03 said:}

I saw your comment on Arcturus (I think) because I looked the book up -- Lewis uses it for his space trilogy, so I marked it to buy and read, at some point, thank you. The Gnostics trouble me more than atheists, likewise -- as an acquaintance pointed out, the lesson of Gnosticism may be ``let the dead bury the dead'' -- sometimes traditional forms have to be resurrected rather than resuscitated (if that's a good analogy?). Which means a lot of hard work. And they can certainly help us in dealing with the current climate (which is actually becoming very Demiurgic, what with the rampant PC and reigning gods of multicult, etc.). If the Gnostics had the courage to go against God, what excuse do we have for not standing against false gods? They certainly have praiseworthy points. I chose the debate because Peter's answers are so good -- if the Demiurge really is real, and ``revelation'' is from him, how does this get the Ineffable off the hook for the problem of evil? It doesn't (it seems to me) -- it just pushes the problem back by giving an emotionally charged and short-term satisfying answer (defers the cost). Same as atheism with final causes.

\hfill

\texttt{Dominion on 2012-08-14 at 00:14 said:}

Professor James Cutsinger provides a good explanation of the role of the exoteric ``religion'' in his lecture entitled ``The Noble Lie'', available on youtube. Taking this viewpoint, it is not so much the religion, be it Christianity or Gnosticism, that matters, so much as it is the potential of its use as a spiritual aid. The doctrine of the Triune God and the creation undergoing apotheosis and the doctrine of the Demiurge and the ``true Light'' or ``laughing Christ'' can be taken as conflicting narratives being used to attain the same end, at which point the doctrines themselves (their `literal' truth or untruth not being the point) can be surpassed and salvation or liberation attained. In that respect, the initiate among the Cathars and the one receiving the Eucharist at Divine Liturgy or Mass can embrace each other as brothers in a sense beyond their fellows in the exoteric faith. That said, Gnosticism's ability to be of much use as a faith on which civilizations can be founded seems limited, precisely because of its ambivalence to matter and the material world.

Also, at what point do you believe that Traditional esoteric doctrines and practices became known among some in the Church? Personally, I have always found it difficult to see how the scant if not nonexistent hard evidence can lead us to be certain that Christ and his Apostles themselves were in fact initiates in the proper sense (they are ``esoteric'' doctrines, after all), and at best it seems as if no conclusion can be drawn either way.

The appearance of the Euthyphro is certainly a treat.

\hfill

\end{sffamily}
\end{footnotesize}

