\section{Players and Pugilists}

It is of considerable significance that the profession considered most sinful in classical, medieval, and even early modern European society was the actor, that theater was consistently denounced by both civil and religious authorities, and that actors were not considered citizens in France until the great turning point of 1789. Even following this legal inclusion, actors continued to be excluded from polite society. Before the 1940s, appearing on stage was the surest way to destroy your good graces, and returning to society required renouncing one's past as a performer; but since the 1960s, newspapers have eliminated their ``society pages'' entirely and replaced them with coverage solely of actors and singers, with the effect that the society of gentlemen has been totally exchanged for a society of performers, with few even noticing the world-upending magnitude of this transformation.

The prohibition on stage performance has now vanished from society, except among the dwindling readership of the \emph{Social Register} and surviving members of the European aristocracy. What is surprising is how long the taboo against acting held out: even episodes of \emph{I Love Lucy}, and films through the early 1950s, exhibit this disdain for ``show business''. It is equally surprising how completely the taboo is forgotten today, considering its universality up until the 20th century. It is as if actors have always been worshipped, but throughout written history the reverse is standard.

In Rome, acting was the ars ludicra, basically implying a pointless and irreverent burlesque, and equivalent to prostitution. The state did not grant any legal rights to the actors, who were also forbidden from entering the military. In the Lex Julia, masters were forbidden from forcing their slaves to become gladiators, prostitutes, or actors. Likewise, senators and emperors were forbidden from marrying into families of slaves, prostitutes, and actors. Cornelius Nepos, examining the culture of Greece, reported that ``to appear on the stage and exhibit oneself to the people'' appeared to be almost honorable. But in Nepos' wording we see the key Roman criticism of acting: that it degrades one's virtue by creating a spectacle with the self at the center, drawing in not only friends and neighbors but that unfamiliar mass, ``the people''.

St. Augustine, denouncing theater as a distraction from higher callings, suggested that actors be forbidden Communion, citizenship, and the ability to hold public office. The Decretum Gratani shows that actors were forbidden even from filing suit. In France they could not be buried alongside Christians. In Germany, the \emph{Mirror of the Swabians} (\emph{Schwabenspiegel}) and the \emph{Mirror of the Saxons} both rendered actors outcast, since ``wandering minstrels and professional fighters ... were described as taking goods for honor, indeed as giving up their personhood for money.''

The use of ``personhood'' here is striking, and recalls Evola's explanation of the etymology of ``person''. To reinforce how Evola is far from alone in his opinions, here is Hobbes on the same subject:

\begin{quotex}
The word person is Latin: instead whereof the Greeks have \textgreek{προσωπον}, which signifies the \emph{face}, as persona in Latin signifies the \emph{disguise} or \emph{outward appearance} of a man, counterfeited on the stage ... and from the stage hath been translated to any representer of speech and action, as well in tribunals as theatres.
\end{quotex}


Acting in its very language is the essence of becoming. The subject of a becoming is an actor, and to become is to act. A Brahmin may be the village patriarch, or he may \emph{act as} the patriarch; he may be a permanent institution, or \emph{play a role} in an impermanent arrangement. In the modernist mindset there is basically no being, only becoming, which is why human beings do not exist as high and low, men and women, but are rather said to ``perform'' an ``act'' in ``social roles'' and ``gender roles''. It is shocking to realize that many of these theatrical metaphors, now so common that the underlying metaphor is vestigial, scarcely even existed in English before the year 1900.

Personhood is the ability to represent a being, either oneself or another, and should properly be employed honorably, facing towards virtuous things and ultimately towards the divine. In a setting of initiates these representations can take on great value. The self is something that should be minimized so that greater things can be maximized. To perform on the secular stage is to face away from virtue and towards ``the people'', beautifying the self, and exhibiting oneself for fame and material gain. It is thus a metaphysically degrading profession, as is prostitution and democratic politics. Having written these words, we begin reading an essay by the American Founding Father John Witherspoon on ``play-actors'', and we discover that he is in complete agreement:

\begin{quotex}
All powers and talents whatever, though excellent in themselves, when they are applied to the single purpose of answering the idle, vain, or vicious part of society, become contemptible... Music has always been esteemed one of the finest arts, and was originally used in the worship of God, and the praise of heroes. Yet when music is applied to the purposes of amusement only, it becomes wholly contemptible. And I believe, the \textbf{public} performers, from the men-singers and the women-singers of Solomon, to the singers in the present theatres, are considered as in a disgraceful calling. Emphasis added --AHM 
\end{quotex}


The example of the gladiator or professional fighter, with whom the actor is often twinned in these sources, should provide a useful analogy. A Christian or Muslim may flagellate his own body as a demonstration of his submission and selflessness. The only audience of his action is God, and indeed modern groups which practice mortification, such as Opus Dei, do so privately. A fighter, on the other hand, aims to attract a large, paying audience to see him batter the body of another and be battered himself. The Roman gladiator lacked a monetary incentive, but the direction of the activity was the same: facing away from the sources of virtue, and towards ``the people''.

The traditional attitude towards this variety of degradation is a subject which deserves more historical study, as do the changes which produced modernity. What we find consistently in the few accounts we have assembled is that actors are banned from any role in hierarchy, as above, but they are still granted basic privileges. In the Ancien Régime, actors could get married under clearly symbolic conditions: they would renounce their profession, marry, and then quickly be reinvited to their profession by the valets de chambre. This was one of the institutions eliminated by the French Revolution, and the consequent dispute over the social rank of actors can be seen as one of the first battles of the occult war, albeit one that is completely forgotten today, and part of the process by which tradition became limited to the religious sphere in the West.

One example will suffice to show how closely this battle resembles modern battles over similar issues. In December 1870, an actor named George Holland died in New York City. An Episcopal reverend named W.T. Sabine was called upon by a comedian friend of Holland's to perform the funeral. Rev. Sabine initially agreed to it, but then the comedian remembered that not all clergymen were fond of actors, and informed Sabine of the deceased's profession. Rev. Sabine responded that he would not permit a funeral for an actor in his house of worship, but that he would officiate at a home funeral if need be, and that there was a ``little church around the corner'' which would be more likely to accommodate him. This simple commitment by a single reverend to his values caused a national uproar, with newspapers around the country commenting on it, and Mark Twain dubbing the reverend a ``crawling, slimy, sanctimonious, self-righteous reptile''. The ``little church around the corner'' held the funeral and became famous for doing so, and still retains that nickname today. The news media did not speak a word in Rev. Sabine's favor, while those who approved did so privately, behind the closed doors of gentlemen's clubs and parlor rooms.

We must consider, not the ``right thing to do'' in this situation, but what metaphysics was incarnated in Reverend Sabine. For this man of God, the Church existed for worship and submission. By making a living on the popular stage, which is quite a different activity from acting as religious service, the actor failed to concern himself with the realization of his innate capacity for worship, which we call faith. Like every populist throughout history, his executor demanded the benefits of Christendom without fulfilling any of the responsibilities incurred by being a Christian. The Reverend followed his conscience to deem the actor a sinner; the doctrine of mercy could allow him a home funeral, but holding one in the church would do an injustice to the ancient standards of the body of Christ, dating back to St. Augustine. But the Reverend had unfortunately been born into modernity, where the people dictate their own laws and choose their own gods.

Metaphysics in the sense of Guénon and Evola can often be confusing, but if you understand its rhythms it is easy to recognize its human forms. When we speak of Tradition, we mean to suggest to you the instructions Reverend Sabine was following when he refused the actor a church funeral. We aim to dispel the emotional clouding of populism which caused Twain to dub the Reverend a ``reptile'', and to give the reader a clear and unclouded window into the beautiful structure of the Reverend's universe, the same beauty that is reflected when the modern church refuses marriage to homosexuals. Tradition is not a confusing relic of the past, but a permanent structure that gives access to divinity; and while it is almost vanished in the present day, it still exists for anyone with the courage to recognize its forms and appreciate them.

The New York incident was the end of official discrimination between actors and Christians. Unofficially, though, high society remained outside and above show business for many decades more. Newspapers continued to print denunciations of theater up until roughly 1900. In the private spaces of those clubs and parlors, which were quite influential in that era, marriage to an actor or singer was anathema through roughly 1940, as we have said. Following the mass production of film, however, the cult of the actor grew in America to such an extent that an unholy marriage between acting and nobility was produced, and birthed ``celebrity''. The pugilist, to be sure, is still not welcome at a party, for the degrading nature of his work remains clear; but the player has become the star of the show, so to speak. The private aristocrat is no longer interesting to the common people, except when it is time to hate him. The public actor, on the other hand, shares with the common people an open desire for mindless entertainment and vanity, and is thus beloved.

The ennobling of the traditional outcaste of actors must be seen as a key step in the transition from tradition to modernity. As an outcaste, actors formally play no role in the traditional metaphysics. This enables them to create their own, rivaling metaphysics and ridicule the values of the traditional system. When actors are exclusively male and employed by a patron or an institution, as was the case in Greece, on Shakespeare's stage, and in the Catholic passion plays, they may work as a visionary force within tradition. Left to their own devices, though, actors and actresses will display themselves to the world for money and fame, and this celebration of vanity is destructive to tradition.

Works referenced

Hobbes, \emph{Leviathan}.

John Witherspoon. \emph{A Serious Inquiry Into the Nature and Effects of the Stage: And a Letter Respecting Play Actors}. 1812.

Cleveland Amory. \emph{Who Killed Society?} Harper, 1960.

Jonas A. Barish. \emph{The Antitheatrical Prejudice}. University of California Press, 1985.

Kathy Stuart. \emph{Defiled Trades and Social Outcasts: Honor and Ritual Pollution in Early Modern Germany}. Cambridge University Press, 2000.

``Pharisaical Delicacy\footnote{\url{http://query.nytimes.com/gst/abstract.html?res=F00715F83F551A7493CBAB1789D95F448784F9}}.'' New York Times, 29 December 1870.

Recommended further reading:

\url{http://www.imdb.com/keyword/all-male-cast/}

Claudia Durst Johnson. \emph{Church and Stage: The Theatre As Target of Religious Condemnation in Nineteenth Century America}. McFarland, 2007.

You may read my blog\footnote{\url{http://avery.morrow.name/blog/}} if you like.


\flrightit{Posted on 2012-06-09 by Avery Morrow}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Izak on 2012-06-09 at 20:57 said:}

This was a damn fine little essay --- very interesting and thought-provoking!

Definitely adds a much-needed perspective for anyone who's ever been exposed to postmodern ``performativity theory'' in the American academic establishment.

Good job.

\hfill

\texttt{Cactus on 2012-06-10 at 20:31 said:}

A fine essay indeed!

``Left to their own devices, though, actors and actresses will display themselves to the world for money and fame, and this celebration of vanity is destructive to tradition.''

Ultimately, every activity left to its own devices (that is to say, performed outside of a tradition) can be degrading, `sinful'. So, in Cavalcare la Tigre we read that the problem for the differentiated man is not the search for a basis, which he ought to possess by virtue of his special differentiation, but his ``expression and confirmation'' in this life. Hence the treatment of `pure action' in the same text, for those who must remain in the world and choose from a variety of desecrated activities, for whatever purpose.

It is no wonder that acting and singing gained prestige at a time when traditionally respectable professions were undergoing final subjection to secular humanism, scientistic positivism, and their attendants, as all understanding of true ethics was eclipsed.

Now it is the Arabs' turn to handle the hot potato (this from today's news):

``A Saudi city known for its ultra-conservatism has created its own version of the `Arabs Got Talent' television reality show, but with no music and women banned from taking part.

Instead, competitors will be permitted to perform religious chants, recite poems and engage in sports events.''

\hfill

\texttt{Cologero on 2012-06-12 at 19:07 said:}

It was probably Grace Kelly's marriage to the Prince of Monaco who changed all that; at least, unlike today's entertainers, she learned to act the part of a Princess. Since then, the ``beautiful people'' and those connected to power began their increasingly more entwined alliance. The result is that, over time, the values of the prole and the outcast have become the ``in'' values: the need to shock, to overturn, to revolt; the quest for the novel and the bizarre; the pursuit of pleasure rather than depth; the elevation of appearance over reality; playing the part rather than being authentic.

By definition, popular culture must appeal to the masses. As a consequence, this involves
\textit{``épater le bourgeois''}\footnote{\url{http://en.wikipedia.org/wiki/\%C3\%89pater\_la\_bourgeoisie}}. Since the bourgeois class lacks any support from above, they are made to look ridiculous. Unfortunately, we meet the same inclination from those above; for example, the likes of Evola and Nietzsche do the same, albeit from a higher point of view. However, that subtlety is lost on the crowd, and the overall result is a pulling downward rather than an uplifting.

\hfill

\texttt{Michael Rayborn on 2012-06-28 at 14:51 said:}

While I understand the article being written as well as its viewpoints, I disagree with the idea that pugilism is a degenerate practice. It is valuable for many reasons which I will list below:

In a Traditional Society, the Military or ``Second Function'' is considered one of the higher functions because a person is willing to sacrifice themselves for their particular homeland and culture.

In order for a Military to be effective in the practical sense, it must be able to Teach a curriculum of the highest order. Part of this curriculum is Hand to Hand combat.

Sporting activities such as pugilism (MMA, boxing, ect.) provide impetus through which techniques can be explored and tested within a laboratory. Competition through the promise of fame and fortune provides motivation for the fighters to excel and develop these techniques to the highest level. These methods can the be extrapolated and then incorporated into military training.

My views on this are not unique or even my own. The modern military utilizes Mixed Martial Arts as its combat method. Ancient Greece (and Sparta) utilized an ancient form of MMA known as ``Pankration'' as its fighting style.

All sports, particularly combat sports (hockey, football, ect.) are useful for instilling aggressiveness, competitive attitudes, willingness to overcome and face pain, fighting spirit and many other attributes that are necessary for warrior ship. These activities also gives warrior folk something to do during times of peace so that their bodies, minds, and souls will not atrophy.

The pursuit of wealth and fame are not necessarily at odds with Tradition. A person can be Wise, Just, Courageous, and Noble and know how to utilize fame and wealth towards constructive ends. People who are wealthy and famous have more power, and people with more power are better able to cause change (for good or ill). A person who possesses fame and wealth while having a noble disposition can do very positive things. Of course, I am not saying that someone should absolutely pursue things things (I know I don't). I am simply stating that such things can be used to serve the Needs of Tradition itself.

\hfill

\texttt{Cologero on 2012-07-21 at 00:18 said:}

This is from Louis de Bonald:

\begin{quotex}\footnotesize

Yet when in this past century some thought that comic actors should be raised to the rank of public men because they performed for the public, \emph{opinion protested against the absurdity}. The profession of acting has in fact abased itself all the more in spite of the efforts that have been made to render it honorable, and we may be assured that the opinion that condemns it still exists in all of its force.
\end{quotex}


This is the precise reason:

\begin{quotex}\footnotesize

it is noble to devote oneself to the public utility and ignoble to sell oneself to the public's pleasures.
\end{quotex}


It light of this, it is clearly a sign of the times that in the USA, young people get their \textit{news from a clown!}\footnote{\url{http://www.cbsnews.com/2100-207\_162-603270.html}}

\hfill

\texttt{Ash on 2013-07-29 at 23:21 said:}

A late read, but here are a couple of comments anyways:

Are you distinguishing here between High Society (American and European) and the actor's degraded role? I want to be sure that's the case. I don't think one can overlook the public spectacle that High Society many times itself became -- you yourself mention the Society pages of the papers. The extravagant pandering to the media and hotlists of the time has been well documented from the 19th century onward, from the works of Dostoyevsky and Tolstoy in Russia to the Bloomsbury group in Britain to the authors of works like the Great Gatsby in America.

Also, I think Cologero's example of Jon Stewart is rather apt. It is only when the proper authorities have failed in their duties that the forces which the ``actor'' here represents can take over. Stewart's role as a court jester of the modern United States is valuable indeed, even if America were a functioning society. But the ``proper'' news outlets are utterly vile sources of information and their journalism is barely extant. What can one do but turn to the court jester, when he at least gives us a far better substitute? In the interregnum, it's the best there is.

\hfill

\texttt{JA on 2013-07-30 at 13:15 said:}

We still have some ways to go before our ``aristocrats'' become men akin to this bloke : \url{https://en.wikipedia.org/wiki/Elagabalus}

\hfill

\texttt{Avery Morrow on 2013-08-03 at 06:06 said:}

Ash, yes, I did not mean to imply that old aristocrats could not be society maniacs and love to see their name in print. But such behavior was deplored when aristocracy was by blood. Today it is significantly less so. ``Celebrities'' (the very term is a neologism) have only a pale imitation, almost a parody, of the restraint and decorum shown by those who they wiped from the entertainment pages of the newspapers.

\hfill

\texttt{JPV on 2015-03-28 at 20:24 said:}

I am an actor, director and a playwright. I am also a conservative Catholic. For years I struggled to conciliate those two facets of my identity. In acting school I had to do perform in two plays and in those two plays I had to play homosexual characters (at the time I was an atheist but I was already sick of the agenda).

When I became a Catholic I had to leave the whole circuit behind, because I just couldn't conciliate my faith in Christ with the modern ``values'' exposed by 95\% of my peers. I became an outcast. Still, I felt I was running away from my vocation.

Now I only direct and act in morality plays, or plays from Christian authors or plays I find transmit wholesome values I agree with. My last production was ``Magic'' by GK Chesteron and I was surprised by people's reaction. You have no idea the amount of people who came to me and said ``Finally! A play who is not about a homosexual coming to terms with his homosexuality! Finally a play I want my children to see!''.

This was a very enlightening post. And I do agree with the sentiment: we give actors and artists in general too much credence.

God bless.

\end{sffamily}
\end{footnotesize}

