\section{My Meandering Spiritual Journey}

The gnomic will is when you feel compelled to deliberate between multiple choices before deciding what to do. This is opposed to the True Will which arises spontaneously and intuitively from one's real inner being. This post started out as a prequel to the upcoming translation of Evola on Christian mysticism. One of the sections was to be on the possible alternatives. Hence, I relied on my own explorations and it morphed into what follows.

I should be reluctant to post this because it is too personal, but to give in to reluctance would be to fall victim to the gnomic will. But, you may object, I am not infallible so I have no True Will. In that case, consider what follows to be my Confessions and Retractions. I retract all the false paths I followed. I mention them to save the reader possibly years of wasted effort. Many of you are quite fortunate to have discovered Tradition at a young age. As for the confessions, if anything written here sounds enticing, I can assure you it was not so good in reality and the game was hardly worth the candle.

What are the alternatives to the dominant Western tradition? Neo-paganism and new ageism are hardly serious alternatives. I have known quite a few and even participated in them. The \textbf{Course in Miracles}? Yes, I did the 365 exercises in a year.

``Nothing I see in this room (or from this window, or on this street, or in this area) means anything." Isn't that what Guido De Giorgio wrote? What does a blog mean?

Yet a leader of a Course study group said he canceled his health insurance, although he still wore glasses and could not grow any more hair on his head. I have known Theosophists, Anthroposophists, Freemasons, Rosicrucians, Martinists. I know people who have followed the Builders of the Adytum (BOTA) for years and painted their Tarot cards with colored pencils. I have attended classes at the \textbf{Kabbalah Center} (yes, the same one as the stars). I know \textbf{Gurdjieff} groups, and those who have danced his dances for years. I ran into a man I knew 20 years ago, still involved in fourth way work. He hadn't changed his mind about anything.

There is one man I knew, who was most knowledgeable about such orders around the world. He was even an expert on Nordic mythology, which interested him from boyhood. He was a Russian Jew.

I have tried to follow the teachings of Don Juan and Richard Leary and Ram Das. I studied \textbf{Religious Science} and \textbf{New Thought} and still got sick and never got rich. I know the Secret and it is not worth telling. 

But I did meet a woman in a religious science class. She told me she had moved past her childhood Catholicism into true spirituality. She said she would do anything for me. Since she was a dancer I asked her to do the Egyptian alma dance\footnote{\url{https://en.wikipedia.org/wiki/Ghawazi}}. Alma also means ``soul". She claimed sexual experience brought her close to God. Now a better pickup line cannot be imagined, but it is a longshot. So I don't recommend it for you the reader, since no one will likely mistake you for a god. It is a tough role to live up to.

I met a woman who said she ``dabbled" in quantum theory, whereas I can actually follow the math; I was not impressed. I have witnessed channelers channel their mysterious beings. A friend taped the sessions and then typed them up; they were totally forgettable. I knew a woman who could see dead Indians, and they came with her to my study groups; I was never sure who she was talking to. A man told me that humans were planted here by extra-terrestrials. Who planted the ETs, I asked him? I learned Red Tantra\footnote{\url{https://www.online-tantra.com/white-tantra-red-tantra/}} from a goddess priestess; I didn't get enlightened, but it was better than spending an hour at the dentist.

I did a journaling retreat\footnote{\url{https://intensivejournal.org/}}. I still have the journals; that is why I am doing this.

I have listened to gurus from India and dined with them. I have spent time with their followers. Everyone seemed to be confused. ``But there is no one to be confused," is the usual response. Then there is no one to say it and no one to hear it.

I heard a Tibetan lama give a long speech in very bad English. Everyone was excited because he kept communicating with them by mental telepathy. I got bored. At question time, the women always ask the lama about the karmic consequences from weird situations or else what will happen to their dog when he dies. There is a pained look on the lama's face, since the dog is on his way to the hell realms.

A man tried to hypnotize me at a party; despite the noise, I really tried. I even pretended it worked just to make him feel good, but he just got angry with me. A feng shui expert analyzed my house. So I now keep the toilet cover down so energy does not escape the house; it seems to be working. An astrologer told me my fortune, but none of it came true. I still have the cassette tapes, but no way to listen to them. Nevertheless, I did get a date with her but it wasn't in the stars.

One time I was getting my hair cut and I mentioned to my stylist that I had two tickets to a play since my date backed out on me. On my way home, I got a phone call from her. The woman in the chair next to me overheard the conversation and wanted to go with me to the play. Lesson: maintain good relations with your hair stylist.

I have sat zazen for hours upon hours and chanted the Heart Sutra in a sort of Zen conga line. That I really miss. I have meditated on the Medicine Buddha and imagined I was an enlightened being. That is a good meditation, actually. Except place yourself in the midst of the angels and saints and the Virgin, and give up your identification with your sensations, feelings, and thoughts. Stop pretending to be what you are not.

Fortunately for me, I had done several novenas as a boy so I have been under the watchful protection of the Virgin ever since. It is a miracle enough that I am still alive, but some stories I cannot tell publicly. However, the gift of corporeal life is to provide the opportunity to overcome spiritual death. I don't have many chances left.

When you pray, be very still. You may, to your surprise, hear another prayer spontaneously arise; that is the one you need to pay attention to.



\flrightit{Posted on 2013-05-14 by Aeneas }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Avery Morrow on 2013-05-14 at 01:11 said: }

``I learned white Tantra from a goddess priestess; I didn't get enlightened, but it was better than spending an hour at the dentist."

I went to the dentist and she suddenly broke off her regular spiel and told me about Eastern Orthodox ecstasy and of her brother who became a priest at Mt. Athos. I don't remember anything that happened in December 2011 besides that one conversation. I'm not sure if the bill was more or less than what you paid the goddess, but I did get my teeth cleaned.


\hfill

\texttt{Jason-Adam on 2013-05-14 at 13:45 said: }

I've been through similar – though as a Kshatriya my path led me through various political cults – first as a libertarian Republican then a monarchist then as a communist then a fascist then a national bolshevik before I finally returned to the Catholic faith of my ancestors. I also was a punk rocker before returning to Tradition.

True Will – shades of Thelema anyone ?

The upcoming Evola article has my interest……..


\hfill

\texttt{Iñaki Luna on 2013-05-14 at 13:53 said: }

To Avery Morrow : We do appreciate your Humour !

However , this article on the Gnomic Will (and its relationship to ``A Course in miracles" , which is a strange book indeed …) is very interesting too …


\hfill

\texttt{Constantine on 2013-05-15 at 15:56 said: }

Great article but just so you know a Jehovah Witness would call that spiritual fornication and you might have contracted spiritual aids.


\hfill

\texttt{Aeneas on 2013-05-16 at 00:38 said: }

Not just them, Constantine, but a Guenon would say I put myself under the influences of evil or counter-initiatory influences. Yet there is always the possibility of conversion and atonement; that is why I was clear that it is a confession and a retraction.


\hfill

\texttt{White Rabbit on 2015-12-06 at 00:59 said: }

Hagakure (The Book of the Samurai): 

``In the words of the ancients, one should make his decisions within the space of seven breaths. Lord Takanobu said, `If discrimination is long, it will spoil.' Lord Naoshige said, `When matters are done leisurely, seven out of ten will turn out badly. A warrior is a person who does things quickly.'

When your mind is going hither and thither, discrimination will never be brought to a conclusion. With an intense, fresh and underlaying spirit, one will make his judgments within the space of seven breaths. It is a matter of being determined and having the spirit to break right through to the other side."


\hfill

\texttt{Alistair Fraser on 2015-12-06 at 08:00 said: }

Wise words, check out the film Le samourai with alain delon, there is a very powerful existential message relating to certain disciplines portrayed in that film


\end{sffamily}\end{footnotesize}
