\subsection{Ennead}

Iamblichus has taken us on a metaphysical tour of the Numbers; is it too much to claim that any metaphysics or religion worthy of the name ought to be measured by the yardstick he has provided? I do not think so. That is, if a supposed ``philosophy of Life'' or doctrine cannot provide a legitimate and profound account of how it accounts for and accomodates the spiritual value of Number which is revealed in his meditations (which are the sum of centuries of thought upon such in Greece), then it is inherently suspect, and ought to be condemned. The condemnation would rest on the fact that it does not partake of Logos, Measure, or Number. The Logos is the structure of manifested reality, the pattern of higher things: as the Scripture puts it, it is the ``evidence of things not seen''. By faith, guided by Reason, we see that the Numbers are actually markers or seeds and guidelines which reflect the unmanifested One above, as well as the manifested but higher planes of existence which are not so obvious to the the untutored. A great and valid religion should be able to explain its exoteric dogmas in terms of Number(s), so that it demonstrates a correspondence with actual Creation, rather than wish-fulfillment and delusion. The same would hold true for such
\emph{ad hoc} philosophies as neo-paganism or Nietzsche's philosophy of the hammer. Where, in most modern worldviews, is there any effort to harmonize with the Logos? Most often, we only see expressed the virulent hatred for other points of view, even if justifiably so in terms of pure intellect. The same would hold for a certain kind of Traditionalism that restricts itself merely to the rejection of what is Modern, thus (in a weird way), acknowledging its opponent to be an anti-Monad which it is rejecting in a Dyadic rebellion.

Corrollaries to this are obvious. Obviously, the Stoical tradition is superior (for instance) to the Epicurean philosophy (a fact acknowledged in the epistles of Saint Paul, who doesn't mind quoting certain philosophers as against others). In our own day, a similar form of sifting might occur with the many Thoughts we encounter in the ``Marketplace of Ideas''. That is, some men are more sane, balanced, and normative than others, and can be taken as sound guides on certain subjects. Isn't this what Gornahoor has undertaken?

In regards to the Ennead, it is so short, that I simply recommend readers to peruse it\footnote{\url{http://www.scribd.com/doc/48888397/Iamblichus-the-Theology-of-Arithmetic}}. Iamblichus makes the point that Nine is the last of the numbers, since Ten is simply a Monad once Nine is taken away, and so on and so forth. Nine is the cube of 3 (in the sense of 3+3+3), and so is a kind of natural completion, or ending point, which Iamblichus again compares to a goal post that is raced up to and turned around to head home. Even its name is a play on words resembling Henad (Hen means One, so a Monad). It is Oceanus, or ``horizon'': moderns might say, Event Horizon. It is also called Hyperion, because it is the supremely last manifested Number before the repetition begins (strictly speaking) with the Decad. It also contains all harmonic ratios, as 4+3+2 = 9 (sesquialter, double, and sesquitertian).

But the chapter is very short, and if you are going to read a chapter of this work, this is a good place to dive in.

I hope it is not necessary to point out that Plotinus wrote The Enneads\footnote{\url{http://en.wikipedia.org/wiki/Enneads}}, to readers of Gornahoor.


\flrightit{Posted on 2013-04-05 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{David on 2013-04-06 at 09:08 said:}

I thought some of you would be interested about some numbers hidden in the Gospels. So here it goes :

-- In the Gospel of John, in the ressurection part, the 7 disciples appear 7 times in the text. 7 being number of perfection (4 cardinal points + triad)

-- Again in John, it is said that there is 153 fish that were collected. Note that it's not written «about». Number 1 to 17 added up in a pythogerean fashion = 153. 17 is composed of 10 (the Multitude, the Totality) and 7 (the Perfection). Thinking about Jesus and the «God made flesh so that flesh could be made God» make the meaning quite interesting.

-- There is also a number of small numbers that goes everywhere in the Gospels and are repeated for a structural reason (ex: Peter that deny 3 times Christ, and Christ who asked him 3 times to guide his sheeps, as to absolve him).

-- In the genealogy of Matthew in the youth of Christ, he make 3 lists of 14 names (or 6 of 7, again 7). 14 is the numerical number of David (DWD in hebrew) added up : D=6, W=4, D=6; 6+6+4=14. And it is again paired with 3, the Triad. And this is evident for two reasons : Jesus is presented as the son of David (royalty), and David is exactly in the center of the genealogy. Also, Jesus (through Theotokos Mary) is presented as the breaking point of the last name in the serie (the verb used is total passivity to the Divine instead of an active principle like the other): he is thefore the number who add up to make everything complete, to be the last monad that forge the last Triad (4+3=7), to forge the last number in the sequence so that everything add up as I explained it. There is a lot more to say about the genealogy, which is full of symbols, but nothing else with numbers.

-- In Luke, the number 3 appears for symbolical meaning through his Gospel (3 days mostly), and it's generally a feature only present in Luke\footnote{For those of you who wants to read the Gospels other than meditativly, you should try to read them in parallels : that way, you'll unveil difference of message and therefore meaning between the Gospels. They do not want to teach us the same face of Christ. Do not say ''I've already read this!'', or you'll miss sometimes big details.}.

-- In the genealogy of Luke : 77 names; 11 list of 7 names. The 7th name is Enoch, the first to have ascended to Heaven.

I could go on. There is alot of numbers around the Gospels, alot more that needs to be unveiled.

\hfill

\texttt{Logres on 2013-04-06 at 22:28 said:}

Thank you, David. I used to dismiss ``Bible code'', but there is certainly something encrypted, although I would hesitate to say that it is paramount in the sense that it could over-ride the results of meditation. However, they should operate in a parallel fashion, no?

\hfill

\texttt{David on 2013-04-06 at 23:26 said:}

I think they do. And I think they convey both different message. Trying to tell people that Jesus is son of David through numerical alchemy is not meant (I think so) to meditate on the «way» rather than on the message in itself. I think it's supposed to give motivation, to answer to human concerns, to develop exoterism (a lot of the numerical things answer to either greek socio-historical concerns or «jewish» ones). I think we should keep in mind that some of the things in the Bible, for us, being in Kali Yuga in the second millenium are motivation and esthetics : alot of it is either veiled or forgotten, or just not meant for us. Tradition always adapted to its own historical context, and the Bible is not (for me) apart from this. Jesus is the revelation, not the book.

\hfill

\texttt{Logres on 2013-04-07 at 06:17 said:}

That makes sense: this is sort of how I've treated it in my own mind, since obviously the code is so complicated and the meaning somewhat lost, however, it undoubtedly was meant to enhance the meditation for some readers and/or confirm the exoteric religion for others. But your answer helps clarify this.

\hfill

\texttt{David on 2013-04-10 at 18:02 said:}

\url{http://glory2godforallthings.com/2013/04/08/reading-in-communion/}


This text is very thoughtful on the subject of reading the Bible. Sorry to have disgressed with numbers and Iamblichus to the Scriptures.

\end{sffamily}
\end{footnotesize}

