\section{The Triad}

When one refers to the Triad, one must remember that:

\begin{quotex}
The Triad has a special beauty and fairness beyond all numbers, because it is the very first to make actual the potentialities of the Monad -- oddness, perfection, proportionality, unification, limit.
\end{quotex}

We recall that the Dyad is a second Monad, which begins the process of the primordial's manifestation, which process reaches a form of completeness with the Triad; this is because the Triad is the first number to be ``odd'' (``Odd'' here meaning that when divided, one part is greater than the other, or differs from the other). It is the first number to be equal to its predecessors, because 1+2=3. Recall with the Dyad that all we can say is that it does not succeed the Monad in the same way that the Triad succeeds both Monad and Dyad, because 0+1= 1, not 2; the Dyad is an ``as-if'' Monad, not a true successor. With the Triad, a true harmony is reached.

In the Christian tradition, the Son does not claim equality with the Father, nor does He retain dominion and lordship at the end of time: rather, He delivers all back up to the Father, or Arche, through power of the Spirit. It is with the Spirit that true balance within the Trinity is reached: The Father generates the Son, and the perfect submission and harmony from the Son to the Father, the perfect Love for the Son from the Father, is so intense it generates a Third Person, and this forms the perichoresis of both inner essence and outer energies of the Divine ``Three-in-One''. The Son is associated with the Dyad, as is Queen Mary, the Queen of Heaven: Water, the most submissive element (as the Tao teaches) symbolizes the Feminine \& the baptism of the Son.

The Triad is unique in that 1+1+1=3, where the middle term is the same as the two extremes. Also, in 1+2+3, the middle term is equidistant from the two extremes. So it succeeds (perfectly) two sources, and is the system of the two sources. 3 exists between 2 and 4 in a special way, additionally: 2 is the first double, and sums up the idea of being ``greater than'' the One (0+1=1, less than 2), while 4 is sesquialter, that is, it is less than 1+2+3 (=6). 3, as pointed out, is precisely equal and balanced, because 1+2=3.

So we see the ``Son'' is highly exalted, although being perfectly submissive. The ``Son'' (2) is not successive to the number(s) preceding it (and is therefore less), while He is greater in that He is ``more'' or actualized, because He ``doubles'' the Father. This is the way it is expressed in Christian dogma, but Iamblichus is teaching something that is at the common root of this dogma with other traditions, seeing that even in Taoism, there is a Divine Three\footnote{\url{http://en.wikipedia.org/wiki/Three\_Pure\_Ones}}. We might say that Christianity is the religion of the Son, externalized, out of the bosom of the Father, but that the Monad retains the power and the glory to sustain, unmanifested and manifested, something else that shares the common root in that Monad. If this makes your head hurt, or sounds too Christian, just concentrate on the Numbers.

\begin{quotex}
The Monad is like a seed in containing within itself the unformed and unarticulated principle of every number, the Dyad is a small advance towards Number, but is not Number outright because it is like a source; but the Triad causes the principle of the Monad to advance into actuality and extension. `This' belongs to Monad, `either' to Dyad, and `each/every' to the Triad.
\end{quotex}

Iamblichus also uses etymology: Trein means piety, so it is called prudence \& wisdom; in the Christian tradition, Divine Sophia (as well as in Gnosticism). Knowledge (true knowledge) follows the Triad. The Triad is also beginning, middle, end. This is why prayers are made three times, in formulas of threes, and why even the Triangle (the first planar figure) has three different types (equilateral, scalene, isoceles). Adding the Triad up we get 6, which is the first perfect Number. It is unyielding and cannot be worn down, because you cannot divide it into two equal parts. There are three Fates, three levels of Creation (sky, land, water), three motions of the stars (summer, winter, ecliptic), and three phases of moon (waning, waxing, full). Giving, receiving, and requital is the pattern of Generation, both divine and human. He even quotes Homer: ``all is divided into three'', which becomes (with Aristotle) a meditation on virtue being the mean between two extreme vices.

What is crystal clear is that even something as simple as the relations of the very most basic Numbers is revelatory, to the meditative spirit, of the fabric of the universe. And this makes perfect sense, for how could God not be revealed in that which is most humble, the smallest of things, the very pebbles or calculi which we move around in our first attempts at numeracy?


\flrightit{Posted on 2013-01-11 by Logres}

