\section{A Higher Imagination}

Typically, as Mihai has pointed out, the imagination functions to degrade man. It offers him pornography, violence, and all other species of lust: a 24/7 cinema strip of running images. If it's not doing this, it's running narcissistic movie clips of nonstop happy endings for the engrossed slave, who lives in the Matrix and desires to conform to his own ignorance and those around him.

Every so often, writers come along who seem to break free from this, and challenge our perceptions. This can occur in either direction. Obviously, Lucifer ``imagined'' a world in which God was irrelevant and distant, a ``sub-creation'' that was independent of God's kingdom. The Left has made great use of such imaginary images and art, which it terms ``higher'' in relation to the ordinary mechanical cinema that runs in men's head: artists like Bunuel\footnote{\url{http://en.wikipedia.org/wiki/Luis_Bu\%C3\%B1uel}} have attacked the Church with brilliant subtlety and abandon. This is ``higher'', but only in relation to the degradation previously induced by Leftist ``ideals''. So it is a carrot-and-stick ploy: man is driven by lusts, and lead on by illusions.

Sometimes (however) the writers challenging our perceptions happen to be operating from a magnetized or sanctified area of their own personality. It is the Inklings who were the true ``rebels'' towards the System, not the Surrealists or Existentialists or Dadaists, given that the ``System'' is not even what it is portrayed to be. They proposed that a portion of man's personality could be elevated in such a way that it would ``drag'' other portions of that personality behind it.

Is this an ``un-traditional'' Idea? We do not think so: Mouravieff in Gnosis suggests that any portion of man can be magnetized and provide a ``core'' which then magnetizes the rest of his personality. Furthermore, various physical and national races have various ``make-ups'' -- the Anglo-Saxon, one, for instance is relatively balanced and placid, whereas the Slavic personality is not. This probably explains the Anglo-Saxon capability for ``watering down'' his religion to the lowest common denominator, a fact that made him uniquely suited to further the Kali Yuga through Anglo Empire. The German and the Slav would have made a harder choice.

There is a specific example of the doctrine of polar beings which occurs in \emph{Phantastes}, which we will excerpt partially and summarize, along with other passages, next week. How did MacDonald know to have Sir Anodos learn love by being willing to lay down his desire for the white lady because ``she belongs to another, better man''? By pure intuition?


\flrightit{Posted on 2013-05-18 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Scardanelli on 2013-05-18 at 11:16 said:}

Thank you for these posts Logres, they're always insightful and I appreciate them. I'll have to read up on MacDonald in anticipation of your

\hfill

\texttt{William Zeitler on 2013-05-18 at 12:25 said:}

There's nothing like actually reading the man. Phantastes and quite a number of his other books are free for Kindle. I got the complete works for \$2.99. Been working through them waiting for the subway, etc. He definitely has an elevating effect I have not experienced with other authors (even C.S.Lewis).

\hfill

\texttt{Lemonheadz on 2013-05-18 at 13:05 said:}

``Is this an ``un-traditional'' Idea? We do not think so: Mouravieff in Gnosis suggests that any portion of man can be magnetized and provide a ``core'' which then magnetizes the rest of his personality.''

What exactly is ``traditional'' about Mouravieff?

\hfill

\texttt{Logres on 2013-05-18 at 21:47 said:}

1. Doctrine of Unity

2. Praxis of Awakening

3. Subtlety of analysis

4. Comprehensive theory explanatory of more spiritual ``facts''

5. Doctrine of angelic/demonic hierarchies

6. Esoteric Christian core

7. Breathing exercises/yogic techniques/exercises of prayer

8. Mathematical expression

There is more.

\hfill

\texttt{Logres on 2013-05-18 at 21:49 said:}

Mouravieff expressly states he writes for the modern man -- that is, he writes to reach him where he is capable of being reached; this perhaps gives a certain flavor which is unpalatable to those used to reading ancient texts (which Mouravieff includes), Guenon \& company, and other sources. He writes expressly to convince in a practical way, modern man (that is, it is not merely rhetorical).

\hfill

\texttt{Logres on 2013-05-18 at 21:50 said:}

Thank you both (Scardanelli): I hope to be very precise and also original in going through portions of MacDonald and even Lewis (on the Anti-Christ).

\hfill

\end{sffamily}
\end{footnotesize}

