\section{Death}

\begin{quotex}
This is the chapter ``La Morte'' from \textit{Dio e il Poeta} by \textbf{Guido De Giorgio}; this is the first of two parts. The influence of neoplatonism is obvious. De Giorgio has several excellent turns of phrase; I hope I have done them justice.
\end{quotex}

There is an interruption, a withdrawal, a fracture: Death. Universal, necessary, conclusive, that everyone knows, everyone is present at it, it is continuous, hammering, unforeseen. It gleans madly, one could say disorderly, most times against justice, with a cruelty that often offends the petty human calculation that does not know and overlooks the mysteries of divine justice and rails and springs against the hard law of necessity, as if God were imperfection, casualism\footnote{\url{http://www.merriam-webster.com/dictionary/casualism}}, vindictiveness, and not intelligence, perfection of justice and above all of mercy, of love! They do not think that everything is measured, ordered, foreseen in creation but in a sphere and an abundance that passes beyond common human discernment, beyond every egoistic partialism, every eudemonistic calculation. They do not think that one dies because one must die, that human existence without death would be folly, desperation, pure and simple impossibility. If death did not exist men would pray to God for it, as a liberating escape, as a flight that finally interrupts the existential desperation derived from the \emph{taedium Vitae}, from the disgust of life.

What is death in reality? It is the abandonment of form, of the visible, of sensory concreteness and the passage to something that is unknown, i.e., that which cannot be made precise, fixed, defined. In the human, no one can assert or deny what we will be when we no longer will be and men could conjecture nothing without the Sacred Scripture that---in every religion---speaks above all of death and of that which begins when terrestrial existence terminates. One should say instead that all the religious dynamic puts death in the first place as the limit-point for the intelligence of the divine and the attainment of true wisdom, that which leads to God. The various philosophies, instead---with the vacuity that characterizes them---disregards death, as if it did not exist, as if it fell into the ambit of an experience too common to be considered important, as if its necessity were the indicator of fatalism of becoming, universal recurrence, that it does not depend on the human will, that it is not explained, and is only accepted, coasts along, ignoring it. Yet it a culminating point in the evolution of man that he is viciously split from it with a hurling into the void, the terrible abyss of what is not known and is feared to be known.

Why does Religion make death almost the center of the true vision that man must have of life? The answer is simple: because only religion \emph{sees} after death, sees what follows earthly existence and puts men on guard and admonishes and orients them to this existence in accordance with the necessity of death. To live thinking of death, to always comport oneself in life in view of the end and only in respect to this abrupt void that suddenly spreads wide in front of our soul, our only soul since the corporeal shell is separated and motionless forever and is returned to the earth.

The ``force'' no longer exists, the pneumatic-psychic-somatic unity is ruptured and what remains can, furthermore, constitute a shadow that begins the voyage of eternity. The detachment will not be rescinded: in the first days the soul of the dead will desire to return to the immobile skin and will roam ravenous around the cadaver then, after every attempt to take back possession of the corpse is rendered useless, he will being his judgment, his examination and---according to a traditions of the Orient---he is placed facing a great light, in front of which most will close their eyes. His destiny depends on the degree of light that the soul is able to support. The symbolism is obvious: the light is the divine truth and the degree of splendor is the level at which the soul stopped on earth, hence the judgment by which the soul will be fixed for eternity.

The value of prayers and funereal rites that sustain and help to passing over is immense, they serve to facilitate the escape to give the soul in pain the strength necessary in order not to be extinguished, to not lose consciousness of himself, to not let himself go to the vertigo of the void. It seems that the whole cohort of the Saints and archangels is invoked to salvation and it makes the crown of the soul that abandoned the body and that remains fearful, without a home. Only the Saints and the Just who are already \emph{dead} in life do not fear the detachment but instead wait for it with joy because the flight toward God is more certain.

Sanctity is above all the experience of the voluntary death through the realization of the divine and the certain flight of the soul in the formless void. In order to do that the Saint already has prepared himself, by this earth, the wings in order to not be afraid of the void and to be freed in the immensity of the ``world without people''. Whoever, while still in life, has habituated himself to the detachment of the soul from the body, who has glimpsed his naked transience from the contemplative height, who lives of the soul, does not fear death and once this happened, he is ``placed'' in the order of the spirits who are assigned to him by the Divine Will and that he himself had chosen on earth renouncing every vain desire of choice to conform to the will of Heaven.

For the greater part of men instead, death is a leap, a fearful leap into a bottomless abyss, a laceration, a shout, a falling without knowing how or where one falls, the tragic epilog of a sterile life, and a life without a reason. Even if these men adhered tepidly, superfluously to a Religion in which they did not truly believe, who accepted it as a norm transmitted secularly, as an inert formula, a dead weight, a useless heredity, and difficult to access.

To a life not sanctified by prayer, by thought of God, by the gaze of the heart gathered in Him who alone is true, to a human, only human, life, there follows a poor death, the death of those taken by the throat by the cruel bite of necessity. They have no excuse: on the one hand human reason alone is sufficient to make them glimpse the necessity of a higher, divine Order that justifies the cruelty, the insanity, the enormity of our earthly existence, always hoping but never reaching, a rejoicing, a thinking about the \emph{void}, a construction that is abandoned forever, a love to be what we will never again see, a hooking on walls that fall on us with us but beyond common reason, they have the Religion that extends a hand, supports, explains the precariousness of this world in view of \emph{another} which invites us with songs of hope, expectation, joy.

Religion aims at \emph{everyone}, it has hope wisdom, strength for everyone, it has its rites, its prayers, it extends a very valid hand to all, from birth to death, it has an exercise of ministry that dispense the word of God in order to understand what does not need instruction, study, quality of criticism, only the will to participate, the will to believe, the will to hear, the will to recognize, suffice. To understand does not count, is not important, is not worthwhile: it is enough to
\emph{feel}, to love, to give oneself: in the night darker than ignorance, the most alive stars shine, the holiest wisdom is offered, the most resolute certainty, because the things of God are so different from the human, just as sacred wisdom is different from the profane. For the things of God only love, only faith, only the gift of oneself to the Truth is demanded, the Good, Love that everyone believes, that everybody supports, that is lowered to everybody, who gives himself to all, who runs, gallops, flies, rises up, gives himself in.

To refuse the help of Religion means to blaspheme against intelligence, reason, light, it means to dehumanize, to descend to the level of the brute, the filthy beast, the stone, to have a heart of stone, \emph{cor lapideum}, insensible both to what is in life and to what is beyond life. They are the men who don't even know they are men, who believe, by rejecting and disdaining Religion, they are elevating themselves to a rigorously higher vision of life, they who collect only the more infantile detritus, refuse, and illusion, admitting the real existence of matter which is less real, denying being to spirit which is more real, without which there is nothing that is.

The illusory materiality of matter is sufficiently proved by the miracle that breaks its laws, that is a fracture, a suspension in the realm of necessity that it annuls the most rigorous laws temporarily and subverts any certainty of causal link, definition, norm. This and nothing else is the reason for the miracle, to show men like lightening, through twinklings, that matter is an illusions if its laws can be violated, even if only for a few instants. They are warnings, rebukes, that the Lord throws to men of good will because they see, they understand, they put judgment, they become aware that outside appearances there is a reality that they don't know, but it is not to be denied outside of their ignorance, if the best men, those equipped with sacred science, affirm its existence.

Miracles are for those who do not believe, do not see, do not want to hear the word of the wisdom of God. The miracle in an instant beats down the illusory structure, falls through material matrix, makes redeeming action pass through it that it gives back to God what is of God, showing that from God who is Spirit, nothing can come that is not spirit and that materiality is not other than the illusory condensation of spirit that is denied and is hidden because the world is the mask of God and it is opposite to Him just as, in some way, the mirror is opposite to the image because it is reflected in it. If matter truly exists, the miracle would not be possible, but especially the liberation of the material yoke of those men would not be possible, who not only deny, but actively supplant the cohesive laws of terrestrial nature, living supernaturally in the same nature, penetrating the impenetrable, dissolving the ordinary, solidifying the subtle, crossing the uncrossable mountain, and annulling temporarily the iron web of necessity. Whoever does not want to be convinced of this remains in his error where Death, the tardy revealer, surprised him.

\paragraph{Death and Deliverance}

\begin{quotex}
This is second and concluding installment of the chapter ``La Morte'' from Dio e il Poeta by \textbf{Guido De Giorgio}. Here we see De Giorgio moving past salvation to liberation (or deliverance as Guenon translates it). Evola draws upon similar traditional teachings in his distinction between the path of the ancestors and the path of the gods. For liberation, even Heaven is a prison. De Giorgio is forced to use apophatic language\footnote{\url{http://orthodoxwiki.org/Apophatic\_theology}} to describe this process of deliverance. He leads us to the God beyond God, non-Being beyond Being, or in Vedantic terms, Brahman behind Ishwara.
\end{quotex}


Divine knowledge is like a voice that passes by and does not return: whoever does not subscribe to it in freedom, will have his own negation as fate, i.e., the darkness, the void, death. And we arrive in this way at the crucial point: what is Death? Is it truly a cessation \emph{in toto}, an irremediable end, a no longer existing, or a simple, even if complex, break in the natural process of transformation? For common belief, it is the end, the stop, which follows what is not known nor does it want to be fixated on it, but that is, in sum, a not-life, a no longer being what one was. This opinion is common as much to the deniers of God as well as to the lukewarm adherents of Religion. But for Religion, instead, Death is only a crucial point from which another life begins, one more certain, more secure, and truer than that false life that is experienced in the world, the world being a bridge, a necessary passage that conditions the future destiny of man who, we affirm it, is its judge to the extent he \emph{wants} to be it.

How, in fact, could there be light for the man who obstinately rejects the light and how would there be salvation for the man who is convinced that he must be saved from nothing and nobody has to be saved in the inexplicable drama of life? Certainly, when a man passes from the essence of God, from His being all that is, to men's \emph{post mortem} destiny, to the diversity of that destiny and above all to the rather vast exclusion of most from the life in God, he is seized by doubts, he is chased by protests, he is here ontologically out of place, as in the face of the partial negation of a total Affirmation, to an exclusive part of the All, to a tragic void in an absolute fullness, to an inexplicable residue of the divine Universality, to a clashing collision, to a radical injustice. But that depends on the limitations of our vision that is only human, and therefore is in a certain sense
\emph{passionate}, and of our reason that is confined to fixed fields of logic.

One thinks only that God offers Himself to those who want Him and denies Himself to those who deny him and that, at least and only in this, that ``reciprocity'' is manifestly just, in asking and giving, as a man can claim nothing, if, when he returned home in the evening, he did not find what he didn't want to find, rejecting the unpopular guest to whom he always denied access to his own home, slamming the door in his face.

God cannot give himself to those who do not want Him, to insist with those who are deaf and blind. Religion requires men of good will who adhere to and want to adhere to the Law of God. Whoever rejects it is judged by himself before being judged by God. This personal way of seeing things can be very useful only to those who adopt it, even if necessarily limited and partial. Because, examining the thing from a higher and truer point of view, the problem of his own salvation vanishes in the face of that more universal idea of \emph{liberation} which is the dissolution \emph{in toto} of the human-cosmic illusion in God. It corresponds to what St Paul says; when he speaks of the crucifixion of the world, he alludes precisely to the entire and universal process of liberation that is a definitive escape from every limitation, even the supernatural, for the realization of God in the negation of everything, even of Paradise.

But this integrating God in oneself and for oneself, this centering Him in His purest Essence, which is the exclusion of every perfection, for the perfection of all the perfections, this ontological rarefaction that is the negation of every affirmation, this surpassing of every praise, of every eulogy, of every superlative, is truly the intuition of the True God, of the Deity about Whom nothing can be said, not even that He is, if being is only being, and if beyond being, there is the unthinkable absolute, the unintelligible absolute, the absolute ineffability of the Ineffable. In this total super-eminence, God is more than God, since even God is surpassed in the vertiginous exclusion of every perfection. The realization of so much muchness that coincides with the mininumness of every minimum, it is the highest degree of divine realization and it is the silence of silence in silence: here God says of Himself: \emph{Ego sum quod non est}, the neuter absorbing in his indeterminative infinity every personal veil, every aseity, every formulation is only tautological.

There is no more mirror---\emph{Ego sum qui sum}---not even of God in God, not a name, not a voice, not a song: nothing. The Being that is, is the non-being that is. It suppresses the last, the universal determination, which is that of being, of aseity, nothing remains if not the Nothing, the non-Being, the Ineffable of the Ineffabile. This is liberation and those who attain to it will find ``that even Paradise is a prison'' and that there is no longer anything to attain for those who ask for nothing because they can have nothing more. This is not an absurdity of words, but rather a radical absurdity, a divine absurdity that climbs over every limit, every peak, that fills up every chasm, that sinks in the intimate, hidden beauty of the Infinite. For so high a liberation, so total, Death is nothing, barely a transit, not even a passage, but the door through which the Wise, the Saint, abandons the cosmic prison, climbs through all the heavens, and penetrates into the ultimate, \emph{Coelum Coeli}, where God is more than God, He is the infinite totality of negation. Here all the songs are silent and the Abyss alone is silent.

Death is a great master of wisdom and virtue if it is known like that, to what sublimities and what terrible ruins it leads, if one familiarizes oneself with its advent, if it is obtained from God who prepares us to receive it as the necessary means for the exodus from this ``false world'' and the access point to the highest degrees of Being, to the boundless possibilities of new worlds where the soul can sing its loftiest victories. God's true wisdom prepares us to accommodate it as the liberation from the corporeal burden, as the abandonment of the cadaver that we drag on earth, since the life of the multitude, deprived of God's light, is foolish, dull, a worthless necrobia, a dragging of artificial superstructures that hide vermin, the putrefaction of a dead man who dreams.

Dying before dying prepares us for death, feeling oneself a cadaver before becoming one, separating before detachment, sanctifying the corpse, purifying it, making it the marble temple of the Living God. The sanctification of the body is the sanctification of life and the sanctification of the soul that only the Holy calls Holy, i.e., God in us. The divine integration, the knowledge that God alone thinks, feels, and works in us, is the most certain means to attain freedom and serenity at the transition.

Whatever pain, whatever sacrifice, St Paul says, is nothing in the face of the infinite richness of the prize that is now pure of God in comparison to the meager and false earthly riches. Freeing oneself from the tyranny of the I, one can gradually and laboriously attain to peace, to the triumph of the eternal in us, when all the interior springs flow with fresh water. It is a type of progressive denudation, of spiritual desquamation, of the gradual frustration of transience, of no longer feeling the demon that roars in us, the tormenting hammering of the I that arrogantly grabs the openings of the soul and twists them to pain, to crash, to laugh, to tears. To empty life of every content, to give it back to God, through contemplation, prayer, invocation, pneumatic elevation, radiant of Love, the inward conversation with Saints and Masters who become our Guides, our helpers, our mirrors of virtue. This reducing oneself, this abolishing the inner thunder, this listening to the perennial sources of peace flowing, this flowing of silence, this caressing the void in its fullness, this perceiving the darkness of oneself until the Light bubbles, this darkening, this crumbling and incorporeal re-emerging, this confident abandonment in God, this acceptance of all only for the love of the Lord, this No to the world and this Yes to God, this reflecting oneself in other-than-oneself, no longer recognizing one's own face, and ignoring everything except for Him, here is how rebirth is realized after the end, when Death, death killed, is the triumphal song of the soul that rests in God.

\flrightit{Posted on 2013-02-26 by Aeneas}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Izak on 2013-02-27 at 16:21 said:}

This really is very beautiful. I realize that your `musings' posts get more comments, but I think the reason is because these translations are harder to digest and demand a higher degree of contemplation, so everyone risks looking like a fool so much more before saying anything.

But I think everyone appreciates these very invaluable translations, perhaps more than anything else on the site. The fifth paragraph is especially nice. I'm assuming that its poetry is even more strong in the original Italian than it is here.

\hfill

\texttt{Cologero on 2013-02-27 at 22:12 said:}

I appreciate your interest, Izak. As you may imagine, it is not necessarily easy to translate, to find the right word, to be sure that the meaning is correct. It is actually a transformative process to do so, and I hope the readers of the translation will experience the same process. It is important to avoid being presumptuous, i.e., by assuming the meaning is obvious and clear cut. Some facebook comments claim the text is ``life denying''; if so, their lives must be quite spectacular.

\hfill

\texttt{george on 2013-02-28 at 00:22 said:}

The perspective is so thoroughly antimodern that it is hard to contemplate without letting go of modern perspectives.

\hfill

\texttt{Izak on 2013-02-28 at 14:49 said:}

Yeah, I'd say you did about the best translation that the English language can permit, at least for that paragraph. I can get a sense of what Guido was doing with language.

It's really nice to see Guido de Giorgio in English either way. I can see why people would call this passage ``life denying,'' but real metaphysics always kind of does walk the tightrope, doesn't it! It's very hard to say the truth without appearing Manichaean or pessimistic or whatever. Not to totally switch topics, but I always did appreciate when John Woodroffe took pains to explain that ``maya'' does not mean ``illusion'' in the way Westerners think about it, with negative or moralistic connotations. If only more people could be alleviate of such confusions!

\hfill

\texttt{Cologero on 2013-03-03 at 08:00 said:}

Someone called me last week and in the course of the conversation, he claimed that ``God would not send him to hell.'' At the time, I thought it was an odd comment from someone who frequents Gornahoor, not least of all because of the rather blasphemous view of God as some sort of remote ``sky god'', occasionally poking around in the world rather arbitrarily and even unjustly.

Coincidentally, since this particular post was scheduled, rather than attempt to answer him directly, I asked him to read this post when it appeared to see if it addressed his particular concern. I assume it has or else he lost interest. De Giorgio claims that in the post mortem state, the soul will itself seek its own level based on how much ``light'' it can bear. He claims to be repeating traditional teachings.

It is true that Plotinus said that different souls would exist on different levels, reflecting their large diversity, Dante mapped out these levels in the Divine Comedy. It seems obvious that a soul attached to a certain vice will have an affinity to a certain level of hell. It is pointless to put him on a higher level since he will just fall back to his own lever; there is no need to ``send'' him there as though it were against his own will. This is not a theory, it can be verified in your own consciousness, at least for those who have achieved the proper level of self-awareness.

To avoid such a fate, the important issue is to understand the ``light''. It is well worth the effort. According to traditional teachings, the Self is self-effulgent.

\hfill

\texttt{Michael on 2020-03-24 at 10:14 said:}

Just the way GdG forms this writing is amazing, even more so is his light which the writing attempts to convey.

And yet he gives us the parts of a process, specifically the last two paragraphs. All of the statements made there convey exercises to be done by us the reader. Perhaps more can be elaborated on them, will sit with them to digest.

\hfill

\end{sffamily}
\end{footnotesize}
