\section{March Madness}

This is an announcement from Obama that you may hear before the year is out:

\textbf{Obama}: I have determined that all the problems in the world stem from certain events involving Adam and Eve in the Garden of Eden. Therefore, I am proposing legislation to ensure that something like that never happens again.

\textbf{Reporter}: Mr. President, what measures will you take to prevent a future Adam and Eve from eating the fruit from the Tree of Good and Evil?

\textbf{Obama}: Adam and Eve? Who said anything about them? No, my legislation will be directed against God to prevent him from imposing those unjust punishments on humanity. I would have done this by now, except the Republicans refuse to compromise.

\paragraph{Sweet Whispers in your Ear}

Presidents should be limited to a single term. After four years, the sound of his voice starts grating on your nerves, it no longer communicates, it becomes predictable, it begins to make commercial messages sound intelligent. Apart from the occasional Mussolini, few leaders are still able to inspire after a couple of years.

\paragraph{Sports Talk}

The National Football League had their combine this week, i.e., a camp for aspiring footballers who are rated on their skill, agility, speed, and strength. Several teams have been subtly inquiring about the sexual orientation of the aspirants. The media are shocked at that and the NFL has since banned such questions with the threat of dismissal. Now there is no uniform law in the USA banning such discrimination, although everyone simply assumes it to be true. The Federal government has taken a rather broad interpretation of the ban on discrimination on the basis of sex to include sexual orientation; this has been done through executive orders and regulations rather than any actual law.

Apparently, team managements are concerned about gays in the locker room. That is why they are reassured when players are arrested for beating up on their girlfriends. It shows they care. Female athletes are following in their footsteps. Recently a woman basketball player was charged with attacking her girlfriend\footnote{\url{http://www.tmz.com/2012/11/15/wnba-chamique-holdsclaw-arrested-gunshot-gun-fight-girlfriend/}}. She fired a gunshot and missed. Perhaps that is why her career is over.

Much has been made about the success of Danica Patrick at the recent NASCAR race in Daytona, as though it broke some sort of natural barrier. All it proves is that NASCAR is not really a sport. Now NASCAR is now trying to appeal to women and blacks to grow their audience. Apparently they haven't given much thought about why they are appealing to their current audience.

\paragraph{News Commentary by the Numbers}

What chance does the historiography of the Right have to enter into the group mind?

Bill O'Reilly and Laura Ingraham on Fox News recently discussed the Catholic Church. They spoke about the ``crisis'' in the Church, as though it was caused by some natural event. They rightly pointed out that the secular news media takes glee in that crisis. However, the two pundits failed to mention what everyone else is able to see: the ``crisis'' arose allegedly because of the large number of homosexuals in the priesthood. Had they done so, the media would have been entwined in Catch 22\footnote{\url{http://rorate-caeli.blogspot.com/2013/02/fr-dariusz-okos-major-article-with-pope.html}}. Even secularists grasp that\footnote{\url{http://lionoftheblogosphere.wordpress.com/2013/02/27/catholic-church-not-cool/}}. Miss Ingraham's ``explanation'' that we are all fallen, as though your average guy will solicit sexual activity from a young teenage boy because of something Adam did 6000 years ago. But she reaches millions of people per day, and Gornahoor reaches about 600.

Miss Ingraham then pointed out all the ``good'' that the church does in the world. That is to confuse the spiritual mission of the church with mere humanitarianism, which is the only modern touchstone of the good. Albert Schweitzer was a great humanitarian and his writings on his experiences in Africa are much more interesting than the typical slide show of a Catholic missionary delivered in the parish hall.

\paragraph{Where is Al Jazeera?}

I had heard that Al Jazeera English would start broadcasting in March, so I turned on Current TV last night, only to find out that nothing had changed. No sports event and no litigation is conceivable without there being a referee or judge to fairly arbitrate the proceeding. Unfortunately, that policy is not followed in political debates. One American news channle claims to be ``fair and balanced'' because it allows competing points of view to be expressed. But that is not sufficient; a referee will also call ``foul'', and even issue a yellow or red card. Hence a debate referee would, for example, stop the proceeding when he hears a logic flaw or error of fact. Since I have no interest in the content of these shows, I shall, as an experiment, appoint myself the referee of the programs I saw.

I caught the final segment of a show on Islamophobia in which the anti-Islamist Pam Geller\footnote{\url{http://atlasshrugs2000.typepad.com/}} was interviewed. When confronted with the fact that most Muslims have no desire to do her harm, she pointed out that such an attitude brings no credit, since it should be the expected norm. Besides, she said, she is not against people but only against a certain ideology.

\textbf{REFEREE CALL}: When certain authors write against certain Jewish ideologies, they are accused of anti-Semitism, just as she is accused of Islamophobia. The game rules need clarification.

The narrator is unconvinced and points out how Islamophobia is another example of the American experience of racism and xenophobia.

\textbf{REFEREE CALL}: According to the sociobiologist, E O Wilson, all life is xenophobic, i.e., the fear and hatred of interlopers. So is this a uniquely ``American'' experience or a more universal experience? Do predominantly Muslim countries allow Europeans to immigrate and build churches? Does Israel allow Christians to proselytize? Is there Islamophobia in Nigeria? Why is there no Christophobia?

Finally, Miss Geller mentions that Muslims were responsible for the 9/11 attack on the World Trade Center. Curses, said the narrator, people like Geller always mention the worst of the Muslims to make their point.

\textbf{REFEREE QUESTION}: Is that why so many progressive liberals, including our erstwhile commentator Charlotte, prefer to believe that 9/11 was an inside job initiated by George Bush? I.e., so they avoid being secretly Islamophobes?

Apparently Current TV liked that line of attack, so they decided to choose the most obscure situation to make their point about American racism and xenophobia. The next show is about Harrison, Arkansas, a nearly all-white enclave in the Deep South. Because of a purge a couple of generations ago, the city became bereft of black residents and thus it attracted white nationalist groups including the Ku Klux Klan, a Neo-Nazi group, and Kingdom Identity Ministries, a Christian Identity Group. The Klan leader was the most cooperative since much of the air time was focused on him. He seemed more interested in selling his merchandise. The Nazi leader seems to advocate violence, so he could be considered dangerous. Little was said about the minister but they believe that Jews and the spawn of Satan and blacks have no souls.

\textbf{REFEREE CALL}: Jesus said that the devil is the father of the Jews and Orthodox Jews believe that non-Jews have no soul. This needs to be cleared up.

Needless to say, blacks do not feel welcome there. Of course, there is a counter current of opposition who want to bring diversity to the city.

\textbf{REFEREE CALL}: By ``diversity'', they mean they want Harrison to look like all the nearby cities; that sounds like sameness.

The mayor, who is a carpetbagger from Chicago, leads that opposition.

\textbf{REFEREE CALL}: Then why did he move to Harrison? What qualities attracted him there?

The mayor says that businesses won't locate there, although there is a large FEDEX plant there. He says the quality of life would be better.

\textbf{REFEREE CALL}: To make his case, they should have pointed out measurable quality of life factors. For example, what is the crime rate compared to nearby cities? What is the high school graduation rate, and so on. These would be interesting numbers ... or not.

The mayor went to a black college to deliver a speech; he brought along a busload of high school students for support. Good thing he did, since there were virtually no black students at the speech, but his white students gave him a standing ovation. He ended his speech by quoting yet again MLK's dictum to judge people by the content of their character, not the color of their skin.

\textbf{REFEREE CALL}: Then why is the mayor inviting blacks specifically to move to his city and encouraging diversity? Shouldn't he be inviting everyone of good character to his city regardless of their race? If everyone in Harrison had the same good character, would it be more or less diverse?

\paragraph{Being a Better Blogger}

One of our contributors has accused me of actively discouraging readers from reading. It is not deliberate, but I can point out that views have dropped nearly 50\% since my posts by Guido De Giorgio. Is it too morbid? Or too difficult? Or too irrelevant to people's lives? There is a blogger who got dozens of comments about Orange Juice. But I don't have that sort of charisma or reader loyalty.

But if there are twenty people who understand, and I believe there are, I will continue on. I think for a while I will just provide some translations and book reviews of some old, obscure, little read books. That is because I have the impression that people think I just make all this up. I promise to reveal everything on Post \#1001.

\paragraph{Note to bloggers}

It is impolite and really lazy to copy a text just to increase traffic to your site, not to mention it would annoy me if you got more visitors than I do. The bloggers with better manners post an excerpt and then a link to the full text. Of course, the best course is to add value. For example, some of the posts are being translated into Polish or Turkish. An intelligent option is to add commentary and incorporate selected, fair use, quotes. Please let me know.

The most vulgar thing is to solicit tips for copying someone else. One such site recently changed its policy to allow tips to go to the translator as an option. That should be the default option, nay, it should be the only option. Not that I want your money, since I'll just spend it on cigars and alcohol, or 9 mm bullets, but I will gladly accept a Maserati Quattroporto.

\paragraph{Kick Starter}

Perhaps we need a traditional Theory of Everything. It would have to be scientifically plausible, historically accurate, and true to tradition. No, it couldn't just blame the jews, or liberals, or whatever. There is a Kick Starter web site\footnote{\url{http://www.kickstarter.com/projects/1870526265/the-god-problem-how-a-godless-cosmos-creates-book}} to get funding. I just need a title idea, or, as they say, a ``hook''. Please provide suggestions.


\flrightit{Posted on 2013-03-02 by Cologero}

