\section{The Finger of Father Sergius}

\textbf{Leo Tolstoy}'s story about \emph{Father Sergius} begins with the ambitious, intelligent and handsome Prince Stepan Kasatsky. In everything he tried, including speaking French and playing Chess, he strove to excel. He began a military career and tried to enter into the stratified court life around the Czar. Unable to break into the highest circles on his own, he determined to accomplish that through marriage. He fell in love with a beautiful countess.

The Prince idolized femininity and regarded unmarried women as possessing great purity. Women, in return, flattered by such an evaluation, strove to be goddesses. Men, on the other hand, were permitted to satisfy their sexual urges. However, since men were so overpowered by sexual urges, it was left to the virgins to draw the line. This is the opposite from today when women are now free to pursue their sexual lusts and it is the responsibility of men to respect the line under their own power.

Before their wedding, she confessed that she was not a virgin. This so disturbed the prince that he gave up the world and entered a monastery to live a life of purity and devotion to God. Of course, as Father Sergius, he followed the monk's life with single-minded attention. At this point, the story is open to misinterpretation unless Father Sergius is understood from the inside.

Obviously, the simple act of becoming a monk does not lead to immediate sanctity. Rejecting the lures, enchantments, and glamour of the world, just exposes more deeply rooted lures. Two sorts of cynicism arise from this.

The secularist will expose the false consciousness of the monk and declare that the life of the prince is the real one. Sexual satisfaction, social life, and nihilism are his values. One's inner life is irrelevant. The problem for the secularist is that the game of life is stratified, as exemplified by the court life of the Czar. Sexual life is also problematic, since some are desirable and others not, some are considered natural, others disordered. Hence, the inner battle becomes externalized as the fight for economic equality and sexual normalization.

A Lutheran would say that the justified is still a sinner. While true, in one sense, it does not distinguish between the prince and the monk. The prince is a sinner spontaneously, while the monk has reached a higher stage. This spiritual egalitarianism, that is, all being equal in their depravity, is more insidious than the secularist. Until he can become master of his own soul, the monk relies on exoteric teachings, the regulated life of the monastery, and the authority of the Abbott as legitimate aids.

Nevertheless, Father Sergius struggles with feelings of pride and superiority. This is a paradox of the spiritual life. From the objective point of view, the life of the monk is indeed superior to the life of the soldier. Probably unintentionally, Tolstoy has written a parable of the superiority of the Brahman over the Kshatriya. At a higher stage of spirituality, not found until the end of the story, such feelings are themselves transcended.

The celibate path embraced by the monk frees him the strongest attachment to the world. As Nicolas Berdyaev puts it\footnote{\url{https://gornahoor.net/?p=234}}:

\begin{quotex}
Man is greatly dependent, sexually, on woman: he has a weakness for the other sex, a radical weakness which may be the source of all his weaknesses.
\end{quotex}



By overcoming this ur-weakness, a man is free to sublimate his eros in service to spiritual evolution. On the other hand, this becomes a challenge to a particular sort of woman. In this case, an attractive divorcee, knowing of the monk's sanctity, vows to seduce him. She tricks him into allowing her into his cell where she begins to disrobe. Overcome by lust, Father Sergius takes an axe and chops off one of his fingers. Strongly moved by that act, the woman fled the cell and shortly thereafter entered a convent.

At this point, the life of Father Sergius aligns with that of St Anthony of the Desert. He is sent off to a more worldly monastery. As this reputation for sanctity grows, a throng of people begin pestering him, the opposite of the life of a hermit. They ask for healings, blessings, spiritual direction, advice, and so on. Now he is considered a saint and a staretz. Since miracles seem to be happening, more people come and he begins to believe them himself.

This demonstrates a peculiar problem in spirituality. Instead of following the example of the monk, most people prefer to live their spiritual life at second hand by attaching themselves to who they believe to be a holy figure. In the case of a Saint Anthony, this may be somewhat beneficial, but it opens up a path to darker forces, as recently occurred with Rampal in India. On a smaller scale, this is happening in the case of Valentin Tomberg. Instead of executing the difficult tasks of meditating on the Tarot, the focus shifts to the person of Tomberg. Speculation then arises about his role, is he a bodhisattva or some sort of savior. Some will claim to be receiving special inspirations from the deceased Tomberg. This is a peculiarly feminine phenomenon for the most part.

After years of this, Father Sergius begins to tire under the demands placed on him. Despite decades of prayer life, his faith wanes and he considers leaving the monastery to live as a vagabond. Finally, left alone with a simple woman whom he was really supposed to help in some way, he allows himself to be seduced. This is not uncommon psychologically, since apostasy is very often associated with a fall into sexuality. Examples, are too numerous to mention.

He sneaks away, and looks for his cousin whom he has not seen in decades. She is a simple woman, once wealthy but now a commoner; yet, she is devoted to her family and to fulfilling her duties. This makes a final impression on him. Some months later, he is arrested on a minor charge and exiled to Siberia. Symbolically, Siberia is on the edge of the world. There,

\begin{quotex}
he settled down as the hired man of a well-to-do peasant, in which capacity he works in the kitchen-garden, teaches children, and attends to the sick.
\end{quotex}

\paragraph{Postscript}

Tolstoy was opposed to the Russian Orthodox Church, so in the final stage Kasatsky rejects ``churchiness'', ritual, healings, and other siddhis. Perhaps this choice is not ultimately opposed to the esoteric path outlined by Rene Guenon. Nevertheless, there is no shortcut in this. Kasatsky lives the life of a Warrior, then a Priest, before he is able to transcend all possibilities open to man.

There is an echo of this in \emph{Siddhartha} by \textbf{Herman Hesse}. After successive disillusionments with financial success, the love of a beautiful woman, and family life, Siddhartha ends up as a ferryman, ferrying people from one side of the river to the other, and back again, world without end.


\flrightit{Posted on 2014-11-21 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Gert on 2014-11-21 at 07:40 said:}

If avoiding sin requires such drastic measures for men, it should make leftists think about what's the real rape. Millions of deeply religious males have had their innocence SLAUGHTERED on the Feminist altar of sexual liberation. If a woman drunk on alcohol cannot consent, how can a man drunk with sex hormones do the same? Sex captivation BULLDOZES intelligence and rational thinking in males, prosecute SEDUCTIVE females for complicity in rape

\hfill

\texttt{August on 2014-11-21 at 18:56 said:}

It's not so much seductive females -- these days, there is an abstract, collective sexual atmosphere, heavy, hostile and not really oriented toward fulfilment, hence the provocative exhibitionist neurotic, much worse than true seductive women.

If only people were more suited, the `bulldozing of intelligence and rational thinking' could lead upward, and sex could be restorative. How would Kasatsky's life have looked if sex had instead been a positive point of departure for him? There is a potential shortcut -- through the mountain, rather than over it. No wonder women dread the highest type of man.

\hfill

\texttt{Pickman on 2014-11-22 at 17:23 said:}

Gert,

``prosecute SEDUCTIVE females for complicity in rape.'' True in some of the cases. However, why not just wrap women up in dark blankets so men can safely keep their penis in check without the lure of sin. This is the method of the Levant and removed from the Aryan spirit. Beyond seclusion to concentrate on the inner work

or any personal work of a professional creative disposition.

I would be more inclined to August's commentary on the matter.

\hfill

\texttt{August on 2014-11-22 at 23:27 said:}

The style of the Levant can be criticised, but emphasis on female modesty was also a feature of the Aryan world, where it was considered to be a distinguishing custom of the well-bred.

``When someone inquired why they took their girls into public places unveiled, but their married women veiled, he said, ``Because the girls have to find husbands, and the married women have to keep to those who have them!''

--- Plutarch -- Sayings of Spartans

\hfill

\texttt{David on 2014-11-23 at 01:35 said:}

In the end, so-called ''hormones'' and deviation are not product of external features. They are internal response to specific perception of ''external'' realities. Therefore, any (re)action to those are coming from the interior of the subject and therefore bound to be an act of passivity (in the case of reaction to stimulus) or will (in the case of self-control). This is of course the higher state -- which I do not say that I possess -, it is the goal. Even if you remove all women from earth, you do not adress the direct problem of self-control and higher aspiration.

\hfill

\texttt{Scardanelli on 2014-11-23 at 22:00 said:}

Very true David. The solution must be sought within rather than without. Yet rather than simple self control, wherein one restrains an impulse, perhaps what is needed is a transmutation of that impulse, raising its vibrations as the hermeticists say. Once can transmute desire for the flesh and physical union into desire for the spirit and spiritual union. This, I think, is what is truly meant by the transmutation of sex, rather than any potentially dangerous ``sexual gymnastics.''

\hfill

\texttt{Lu Dongbin on 2014-11-29 at 17:55 said:}

Good post. Though I wonder, if we know that what Berdyaev says is true, which seems to meld quite well with Indian traditions with its emphasis on brahmacharya as a necessity for spiritual growth or Daoism's emphasis on celibacy to conserve vital essence needed for alchemic work and transcending the need for woman for the same purpose, what are the best ways to go about achieving this aim?

We live in an era where pornographic and semi-pornographic material surrounds us and in a generally hyper-sexualized society, and as your post pointed out with woman who are free to pursue their lusts and thus temptation everywhere for man. So what are the best methods or lifestyle changes from the world's wisdom traditions for obtaining perfect self-restraint in this area? I see a lot of emphasis on dietary changes and the practice of constant prayer in Christianity or japa mantra in the Indian traditions. Perhaps Augustine would have some insights into this matter since it was one he apparently struggled with but overcame.

\hfill

\texttt{August on 2014-11-30 at 04:27 said:}

If that is the aim, start by avoiding women and situations where desire can be aroused and consummated. Do you find that difficult to do? Or are the consequences hard to bear? If there is difficulty, the next step depends on understanding it somewhat.

\hfill

\texttt{Pickman on 2014-11-30 at 09:05 said:}

The real answer is transmutation. Restraint through the power of will, and ``sexual gymnastics'' as Scardanelli has pointed out is not the ideal (even though the tests of ``love'' and some tantric sexual rites of coitus). Neither would physical dismemberment. Even after consulting Metaphysics of Sex and trying to point to a populist spiritual figures in history is troublesome. Anyone have any other suggestions on the transmutation process?

\hfill

\texttt{August on 2014-11-30 at 18:04 said:}

I think most here are aware that transmutation is the answer. But that is the goal. Abstinence is a means to that end, and initially* requires a greater or lesser application of the will.

Some traditions have specific methods aimed at sexual transmutation, others simply require abstinence with the knowledge that sex will follow the general development of ascesis from gross to subtle, outer to inner; this change of direction is transmutation.

If you really need more drastic means to cut yourself off from sexually, there are plenty of things one could do before taking the knife. For example, if any girl has an eye for you, sit her down and start talking about the degeneration of the castes, your medium-term aim of restoring the Primordial State, the spiritual death indicated by modern music and culture, and a hundred other topics that we are familiar with. If she still wants you after that, you perhaps ought to keep her.

* ``Initially'' could mean decades, and the enthusiasm those with an appropriate vocation feel at the outset should be used as a boost, but without courting illusions about the extent of the task ahead.

\hfill

\texttt{Cologero on 2018-11-21 at 09:50 said:}

his review arose from a correspondence with a Russian woman living in Kazakhstan. She was a student of Agni Yoga and claimed that Fr. Sergius was running away from life. My response:

The only way to run away from life is to let it live you rather than you living your own life. Whatever prevents that needs to be cut off. The finger is not a dead end, just one of the many temptations he had to overcome. Tolstoy's views on sexual abstinence are those of Fr Sergius. In the end, the aristocrat Kasatsky works for the peasant. So the simplicity of the peasant is superior to the court life and church life of the aristocrat.

For me, Natasha, a man has to be a whole. His personal, political, and spiritual life has to be a consistent whole. By experience, I often see that women don't care so much, as long as the man looks good or says nice things.

\hfill

\texttt{Cologero on 2018-11-21 at 09:51 said:}

Fr. Sergius aka Prince Kasatsky insisted on virginity and was immune to seduction. That makes him, in a sense, not easily controllable.

\end{sffamily}
\end{footnotesize}

