\section{Scientific and Political Myths}

\begin{quotex}
It may some day happen that a traveler from the Greeklands will again lodge in this palace and read the book. Then he will talk of it among the Greeks, where there is great freedom of speech even about the gods themselves. Perhaps their wise men will know whether my complaint is right or whether the god could have defended himself if he had made an answer. \flright{\textsc{C S Lewis}, \emph{Till We Have Faces}}

\end{quotex}
\paragraph{The Flat Earth}
Whenever someone wants to dispute a hoary truth, he often mentions that the ancients believed that the earth was flat. Never mind that he can't identify who used to claim that or when they stopped, that accusation serves, in his mind, to discredit any other viewpoint they may have had.

An intelligent man would take another approach. He should be more concerned about all the ridiculous things he believes and work to purge them. This is the method of elenchus\footnote{\url{https://en.wikipedia.org/wiki/Socratic_method\#Method}}.

\paragraph{The Sky God}
A common claim, often made on social media sites, is that Mr X can no longer believe in a ``sky god" or an old man in the sky. Presumably Mr X assumes that is some brilliant insight. Of course, the true Traditional teaching is nothing of the sort and usually goes by the name ``classical theism" in the West or ``panentheism" in the East. Actually, if anyone believes that today, including but not limited to the Mormons, it is actually a relic from paganism, with its belief in Zeus.

\paragraph{Polytheism}
Another not uncommon claim in some circles is the alleged superiority of polytheism. On \textbf{Rene Guenon}'s principle that religious concepts can be re-expressed as metaphysical concepts, ``polytheism" means ``metaphysical pluralism", a complete absurdity.

The other problem is that no pagan of intellectual merit was a polytheist in that sense. Celsus\footnote{\url{http://www.newadvent.org/cathen/03490a.htm}}, for example, defended pagans against the accusation of polytheism by pointing to the Platonic understanding of God (i.e., a form of classical theism). Hermann Wirth's research on the ancient Nordics\footnote{\url{https://www.gornahoor.net/?p=3205}} also confirms it.

\paragraph{Geocentrism}
Another often made claim is that the ancients used to believe that the Earth was the center of the universe. Thus, it is said, the heliocentric theory disoriented man, making him somehow anxious and insignificant. On the contrary, God, or Heaven, was regarded as the center of the universe (metaphysically understood) and the Earth was quite distant from that center. In its sublunar position, the Earth was in a bad place, subject to malevolent influences. Furthermore, Earth was closest to Hell, located deep at her center. This is quite a different picture.

\paragraph{The Future of Intelligence}
One use of intelligence, or rational thought, is to determine how best to achieve a goal. This is instrumental intelligence. True intelligence, however, is to know what goals to aim for.

Traditional anthropology says that man has a rational, animal, and vegetative soul. To be properly human is for the rational soul to dominate the other two. In other words, the rational part of man should be determining goals. The function of intelligence is not simply to find the means to satisfy emotional and sensual needs. If science is the touchstone of rationality, can anyone truly live based solely on science? Who should you marry, what career should you follow, even which is the right side of the bed—these are all questions that no science experiment can answer.

Therefore, every behavior needs to be justified. If the only answer is ``to satisfy a desire", then intelligence is not truly involved in the decision.

\paragraph{Adamic Man}
Another Traditional principle is that science and metaphysics do not conflict. Any apparent conflict is because one or the other discipline has overstepped a line. The origin of man, as the rational animal, is one of those issues that seems to straddle that line. Biologically, man is an animal, metaphysically, man is rational, at least virtually. A viewpoint that is becoming more mainstream is that there existed hominids\footnote{\url{http://edwardfeser.blogspot.com/2014/12/knowing-ape-from-adam.html}}, i.e., animals that look human, but were not. Specifically, they have an animal and vegetative soul, but lack the rational soul that would make them fully human. At some point, somehow, some of them, or exactly two of them, were endowed with a rational soul.

The existence of these pre-Adamic hominids was also proposed by the Jewish physicist \textbf{Gerald Schroeder} in \emph{The Science of God}. Even C S Lewis held a similar view, as he wrote in \emph{The Problem of Pain}:

\begin{quotex}
If by saying that man rose from brutality you mean simply that man is physically descended from animals, I have no objections…. For long centuries God perfected the animal form which was to become the vehicle of humanity and the image of Himself. He gave it hands whose thumbs could be applied to each of its fingers, and jaws and teeth and the throat capable of articulation, and a brain sufficiently complex to execute all material motions whereby rational thought is incarnated. The creature may have existed for ages in this state before it became man…. We do not know how many of these creatures God made, nor how long they continued in the Paradisal state. 

\end{quotex}
\textbf{Boris Mouravieff} has revealed that the existence of pre-Adamic hominids is part of Orthodox esoteric teaching. Moreover, he claims that they still exist on the Earth, although they cannot be recognized by appearance alone.

\paragraph{Brain Use}
My son and I saw a commercial for the film Lucy in which the claim that we use only 10\% of the brain was repeated. We pondered that and my son finally came up with the figure 98\%, since there is not really much room for improvement in most people. Thinking about it again, I think he was much too generous.

\paragraph{Don't Tell Me What to Do}
A woman recently told me that the medieval Church was totalitarian, always telling people what to think and do. In the next breath, she told me she was confused about the current world and did not know what to think or do. Does anyone see the irony?

In point of fact, the medievals simply tried to determine rationally how best to live a human life based on its nature. As such, there can be intelligent discussion about it. The alternative is to live like an animal.

\paragraph{Dog on Bus}
There has been a story circulating in the news about a dog that allegedly gets on a bus alone in order to get to a dog park. People find it fascinating because it would seem to indicate some measure of the dog's human-like intelligence. On the contrary, the real lesson is how little intelligence it takes humans to get through their day. So if you are getting on the bus at the same stop and getting off at another, day after day, you have demonstrated the intelligence of a dog.

There is no denying that animals, and even plants, have intelligence. But that intelligence does not determine ends but merely achieves them. Matter itself is intelligent, since it follows certain laws, although science cannot say why that must be so.

\paragraph{Jesus as comic hero}
On a radio show recently, I heard someone (he may have been a priest) claim that Jesus allowed himself to be captured by the chief priest's soldiers when he could have rained down fire from Heaven to prevent it. I think he may be confusing Jesus with a Marvel comic superhero.

\paragraph{Speaking and Civilization}
I watched some of the Sunday talk shows this morning. They are continuing the sham claim that Charlie Hebdo is defending freedom of speech, following a week in which there were 54 arrests in France on hate speech charges.

Another sham, which no one cares to mention, is that a cartoon is not ``speech" in any meaningful sense. In the epigram, Orual defends free speech by writing a book, based on rational thought. That way a reader can evaluate her case reasonably. There is no rationality in a cartoon: it appeals to emotion, not to intelligence.

No, the purpose of the cartoon is to silence the opposition, not to promote reasonable dialog. The new editor of Charlie Hebdo announced that their purpose was really to promote secularization or laicity, which is the absence of any religious interference in political affairs. Such views must simply be banned, not rationally, but by mockery.

Do not for a moment believe that laicity is merely a concept for bookish intellectuals or ribald cartoonists. It came at a steep price. France invented Terror as a political tactic. Vendee was brutally destroyed, much more brutally than the attack on \emph{Charlie Hebdo}, to promote the same goals as those of the cartoonists. If history repeats itself, those who oppose the cartoonists may face a similar backlash.

\paragraph{Killing for Civilization}
It is also said that \emph{Charlie Hebdo} represents the pinnacle of Western civilization. Of course not, Western civilization predated the Revolution for millennia; to not include all that is simply ignorant. The Revolution is anti-civilization. The talking heads are quite sure that vulgarity, secularism, sheer meanness, and irrationality represent the best of Western civilization. Can anyone seriously spend a day at the Louvre and then claim that the cartoon art is at the same level?

Now, I'm sure that many are willing to die for such values, actually anti-values. That is not the real issue. The fact that many are willing to kill for such anti-values is an evil impulse.

\paragraph{Race and Decadence}
When \textbf{Mahatma Gandhi} was asked, ``What do you think of European civilization?" he famously replied: ``I think it would be a good idea and they should try it."

Rene Guenon had a slightly different, but not necessarily incompatible, idea of race from Evola's. He claimed that every race at its origin had a Tradition. So what we call today the undeveloped races are, in his opinion, actually the decadent remains of once superior races. For example, the Ramayana describes a battle between the North and the darker races of the south which had a great civilization.

By this standard, the various white races of Europe, by rejecting Tradition, are in a process of slow decline and decadence. Hence, the response to Gandhi should be: ``They did try it, but gave it up."



\flrightit{Posted on 2015-01-19 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{IA on 2015-01-19 at 08:24 said: }

The European continent and culture will only survive as a distinct entity if they expell non-euros. Otherwise, France and others will become Lebanon.


\hfill

\texttt{Cologero on 2015-01-19 at 10:19 said: }

IA, there is another way to look at it, something I learned from a Chinese colleague. China and India, which may appear monolithic to us from the outside, are really multicultural entities consisting of nations with various languages, customs, and histories. I was telling him of my trips to Europe and how different the countries are from each other. Given his Asian perspective, he looked puzzled and was unable to accept that. To him, ``Europe" was a single entity, like China, and something that the various costumes, languages, cuisines, etc., could not efface. So it is instructive to see ``Europe" the same way my Chinese colleague did.

So, in a sense, Europe already is, and has been, ``Lebanon" for a few centuries. This is especially so once the spiritual unity of Europe was lost after the Middle Ages.

We too easily forget history. The Europeans killed some 60 million or more of their own people last century, and that was prior to the mass immigrations that began in the 60s. You may be proposing a treatment for a symptom, but I doubt that will cure the disease.


\hfill

\texttt{IA on 2015-01-19 at 11:31 said: }

I appreciate your response but having lived in both India and China (and Europe) for many years I have a different viewpoint. The fact is, Europeans produced a very unique and valuable culture. A high-trust culture for one thing, unlike the Rest of the World. (Don't ingest anything from China.)

Islam is hell on earth. If you want to destroy civilization you can't do better than importing large numbers of blacks and Muslims. Its insane what's happening. Why you think Islam is going to cut down on violence I have no idea. There is nothing in its 1400 year history to suggest that. Just the opposite. They haven't done more damage because they lacked the weaponry. Same with blacks but they're unable to coordinate large-scale attacks. I think you need to separate wishful thinking from a dispassionate appraisal of cultural artifacts and human nature.

The Chinese and Indians especially loathe both blacks and Muslims having had intimate contact for hundreds of years. That prof. Gates character had his meltdown after returning from a trip to China. My guess is they treated him in a way he wasn't at all accustomed to and the scapegoat mechanism kicked in when he got back to the self-flagellant West.

Cheers.


\hfill

\texttt{David on 2015-01-19 at 18:21 said: }

``The Christian ideal has not been tried and found wanting; it has been found difficult and left untried" – G. K. Chesterton


\hfill

\texttt{August on 2015-01-20 at 06:54 said: }

``Islam is hell on earth."

Don't such judgements always depend on one's position in a given society?

It's noteworthy that many of those who have a problem with violence by Muslims will, in the same breath, fawn over the Hindus, as though they earned their place on the continent using kittens and flowers. And yet, the Hindu world is in many ways closed to them, unlike the Islamic – at least in principle, given the present situation.

Shall we ask the men of the East who recall their subjection, or today, the remaining natives of the Americas and Oceania, and the Africans, whether they think the European man violent? And if we find that they view him thus, shall IA accept the condemnation? And extend the same to, say, the Romans? What is good for the goose…

You live in the past IA, look at the modern West. Moaning about violence by Islamic militants, but no comment on the 60 million that Cologero raised, or the Vendée. Your style of criticism is poor because it opens your own to attack. You sound like a bourgeois pacifist afraid of those who would tear the comforts of egalitarian civilisation out from under his posterior. This is understandable, but should be seen for what it is.


\hfill

\texttt{IA on 2015-01-20 at 09:19 said: }

Interesting comments, August, although you might be projecting your own feelings.

I really don't care if I'm on the ``wrong side of history." I don't at all worry about what the West has done to other cultures. European men have contributed far more good than bad and have created the greatest culture in history. Black men cannot be european. Muslims don't want to, and are Arab anyway. But they do want our land and women.

I have no problem living in the past since the western past is both pagan and Christian and its this combination of blood, soil and God that define who I am. 

Charlie Hebdo has always been anti-western. It is the anti-Louvre. Charlie Hebdo is the cancer while the great art in the Louvre is the cure. Modernism, the Enlightenment, is a disaster.

Your Muslims would love to get their hands on the Louvre and do a major Bamiyan. The blacks would just sell it all to the Chinese and Japanese.


\hfill

\texttt{Jacob on 2015-01-20 at 11:41 said: }

I believe Cologero is saying the problem is much deeper than simply a ``rising tide of color", IA. If you find a many dying in the middle of the road, chasing away the buzzards won't keep the man alive. He obviously is there for some other deeper problem. 

I myself am a American, and have never been to Europe. However, from what I can see and what I have been told; The West is dying. Likely because it has, at least forgotten, it's inner life. Change this and I believe the rest should fall into place.


\hfill

\texttt{IA on 2015-01-20 at 12:25 said: }

Jacob, I couldn't agree with you and Cologero more about changing inner life. If Europe or anywhere else becomes Muslim the State will kill you for a change of conscience. Or, if they allow you to live, you will be a dhimmi. A second-class citizen. You need only look at what is happening now in Muslim lands to see this. You cannot change inner life if you are dead. Nor can you live an honorable life as a dhimmi.


\hfill

\texttt{Paulo on 2015-01-27 at 19:30 said: }

``Vendee was brutally destroyed, much more brutally than the attack on Charlie Hebdo, to promote the same goals as those of the cartoonists. If history repeats itself, those who oppose the cartoonists may face a similar backlash."

Was the opposite Cologero,Vendee promote the opposite values of those cartunists!


\hfill


\end{sffamily}\end{footnotesize}
