\section{Homage to Giuseppe}

\paragraph{News for the Week}

\subsubsection{Principles of negotiation}

A deal was apparently made between the G6 and Iran over its nuclear program. Anyone who has ever negotiated in good faith understands that a deal will involve concessions on both sides. Those opposed to the deal argue that economic sanctions should have been tightened in order to coerce a better deal. In other words, negotiations are unnecessary. A legal principle is that a deal is invalid if made under duress or coercion. Hence, the real goal is unconditional surrender. Just say it.

\subsubsection{Climate Change}

Obama is back on climate change. There is a simple solution to reduce greenhouse gases that no one is bringing up, at least not on the political talk shows. According to the {World Bank}, the per capita emission of CO\textsubscript{2} (in metric tons) in the USA is 17.5. Other sample countries: Pakistan (0.9), Mexico (3.8), India (1.7), Nicaragua (0.8). The USA could easily reduce its emissions by 5\% by reducing the number of inhabitants by 5\%, without excessive taxation or burdensome regulations. Instead, it insists on increasing those inhabitants.

\subsubsection{Irrationality}

A young woman claimed that being in the presence of irrationality makes her feel ill; that is something she learned from Rudolf Steiner. Unfortunately, in her case, the illness arose from her own irrationality, but she is too irrational to notice that. On the other hand, most people experience no ill effects from their irrationality.

\paragraph{Null Hypothesis}

The null hypothesis\footnote{\url{https://explorable.com/null-hypothesis}} is the presumed truth in the absence of any evidence to the contrary. Francis Fukuyama\footnote{\url{http://en.wikipedia.org/wiki/Francis_Fukuyama}} claims that liberal democracies and free-market capitalism represent the highest point of human evolution and furnish the definitive answers to all previous political and social questions.

It may be regarded as a class marker since it is the default position, with some leeway, considered obviously true by educated and sophisticated people today. Moreover, the denialists of this consensus reality are considered insane or evil. The past is the history of injustices which need to be corrected.

\paragraph{A Day at the Zoo}

A couple of weeks ago I visited the Jacksonville Zoo\footnote{\url{http://www.jacksonvillezoo.org}} with my family. My son was impressed by how ``ripped'' the chimps were. Of course they were, since they seldom went to the ground and never stopped moving and climbing.

The next exhibit was for the gorillas. I asked the zookeeper why they kept the chimps and gorillas separate. She explained that they were too much alike and would end up competing with each other.

A biologist has speculated that humans involved from the mating of a male pig and female chimp\footnote{\url{http://www.macroevolution.net/human-origins-2.html\#.VSSfOvnF9Bk}}. I can't imagine how horny that chimp had to be although I have known woman with rather poor judgments about their boyfriends.

But that would be an interesting experiment: put some pigs in the chimp exhibit and see if any humans arise. And maybe add a talking snake to see what happens.

Boris Mouravieff says that organic life is subject to three factors: hunger, sex, and fear. That does seem to describe animal life exactly.

\paragraph{Alternative Hypothesis}

The alternative to the Null Hypothesis is what sane and normal people used to believe. They were closer to their natural life in the sense that it promoted fitness. Sex was channeled into procreation and the continuation of the gene pool (``the selfish gene''). Family life was recognized as the way to safeguard this, so any behavior that was opposed to it was considered taboo. Children were considered a blessing. Nowadays pregnancy is often considered an undesirable, if not dangerous, side effect of natural sexuality. Strangers threatened the gene pool.

Hunger drove the acquisition of resources and the social structures to maintain and keep them. This further enhanced genetic fitness. Unfortunately, the unfit often could not survive.

The fear of the lack of food and other resources, or encroachments from other tribes or peoples, led to rule by the fathers. Their duty was to protect against invaders and organize the acquisition of resources. Since that was often dangerous, leading to premature mortality, boys were particularly valued.

Since there was an awareness into the future through one's descendants, there was also a reverence for the past. One's ancestors were revered and their cultural and spiritual inheritance passed on.

\paragraph{Giuseppe Salvo}
Giuseppe Salvo was my great grandfather. Below is the letter I recently sent to several of his descendants in the generation after mine.

\begin{quotationx}
I'm sending you guys a picture of your great great grandfather, taken in 1880. His name was Giuseppe

The first Salvo arrived in Sicily from Tuscany around the 15th century, where he was made a nobleman. From there the Salvo family spread throughout Sicily, especially in the province of Trapani where it is not an uncommon name. In Salemi, where our branch came from, it is a not an uncommon name.

Salemi is uniquely positioned, high on a hill where you can see for miles around. The countryside is gorgeous, dotted with farms, olive gardens, vineyards, etc. For that reason, it was also the perfect location for a military encampment, since a sneak attack would have been impossible.

The Arabs were there briefly, some 1000 years ago or so. They built a castle atop the highest hill. A couple of hundred years later, the Vikings arrived by way of the Normans. (The Normans were descendants of the Vikings who had settle in the province of Normandy in France.)

The Normans moved on to Palermo, where there was a much larger castle. From that position, the Normans managed to re-establish the Holy Roman Empire. I'll tell that story if anyone is interested. There was a lively school of Romantic poetry, from whom Dante learned his craft. Again, let me know if you are interested.

The bottom line is that there is a lot of nobility, both from the Tuscan side as well as the Norman side, whose genetic and spiritual influence must have been quite extensive in a small village like Salemi. You can see it in the picture ... Giuseppe looks nothing like those caricatures of Italian organ grinders that Hollywood portrays.

You can treat this as an interesting, but ``so what'', story. Or you can use it as a model for your own lives, and the lives of your children: to achieve, to explore, to act nobly, and to avoid the vulgarity of contemporary decadent culture.
\end{quotationx}


\flrightit{Posted on 2015-04-08 by Cologero}


