\section{Too Late for Groundhog’s Day}

Thoughts that should never have seen the light of day.

\begin{quotex}
The great majority of human beings grope about during their whole lives among these data of their own existence rather as one gropes one's way between heavy chairs and tables in a darkened room. And what is tragic about their condition is that perhaps only because their lives are passed in this shadowy gloom can they bear to live at all. It is just as if their seeing apparatus had become finally adapted to this twilight state: it is a state of non-vision which is not, however, a state of quite complete non-awareness. ... all of us tend to secrete and exude a sort of protective covering within which our life goes on.
\flright{\textsc{Gabriel Marcel}, \emph{The Mystery of Being}}

When we set out to speak about truth, as when we set out to speak about God, we are in danger of speaking about something which is not truth, but is merely its simulacrum.
\flright{\textsc{Gabriel Marcel}}

Truth is the whole Universe realizing itself in one aspect.
\flright{\textsc{Frances Bradley}}
\end{quotex}

\paragraph{How to Pronounce My Name}

See How to pronounce Calogero\footnote{\url{https://www.youtube.com/watch?v=Pe_1mgizPaY}} for the correct pronunciation. The most common spelling is ``Calogero'', which in Greek means ``Beautiful old man''. For some reason, our family uses an ``o'' instead of the ``a''; so make a minor adjustment to the first syllable.

\paragraph{They Call me the Trickster}

Apparently the word is out that ``the Gornahoor guy is a trickster''\footnote{\url{http://forum.socialmatter.net/discussion/70/last-book-you-read-and-rate-it}}. Of course, a real trickster would put in his writing unacknowledged references to obscure works, deliberate mistakes to hide his real intention, and half-completed thoughts that need to be teased out.

Someone else wrote this:

\begin{quotex} 
His review of Submission is pretty good but the author seems to think that Rediger is a real person and that Guénon lecteur de Nietzsche is a real dissertation, which he claims to have accessed and read. Does he not know that Rediger and his dissertation are both fictional?
\end{quotex}

Obviously, I was the victim of an elaborate hoax. Someone claiming to be in possession of that dissertation showed it to me, but without leaving me a copy. It fooled me because the quality was quite high.

\paragraph{Sophistry and Philosophy}

Socrates distinguished between the Philosopher, the lover of wisdom, and the Sophist, a commercialized simulacrum of wisdom. The philosopher's quest is the knowledge of the Whole, the Sophist is a partisan whose goal is to promote partial knowledge as knowledge of the whole. The philosopher relies on justice, the sophist on power. The power of the sophist resides in the widespread confusion between what we would like to be the case and what is actually true; the confusion is the perversion of the understanding.

The entire news apparatus in the West is based on sophistry. That is, commentators are paid to promote specific opinions. Often, there will be two opposing opinions, so the passive viewer believes he can synthesize a knowledge of the whole out of that. Rather, what is more common, is that the viewer will cheer for one side or another, as though watching a sporting match.

But what sporting match is played without referees or umpires? Hence, a news show should have referees who will issue red and yellow cards for errors of fact and logic. Yet, that assumes there is an objective truth and that logic has power. Curiously, the ideal of pure reason is now under attack as a trick used by white males as a form of control\footnote{\url{http://opinionator.blogs.nytimes.com/2015/07/02/looking-white-in-the-face/?_r=0}}.

\paragraph{On Being Intelligent}

The philosopher \textbf{Thomas Nagle} famously posed this question: What is it like to be a bat? He didn't ask the more mysterious question, What is it like to be intelligent?

\paragraph{The Quest for Adolph}

I saw episode one (of several) on the History Channel about the quest to find Adolph Hitler in Argentina. Since the obsession of that Channel on Hitler and space aliens puts Miguel Serrano in the mainstream, I decided to watch it. Its premise was based on some recently released 70 year old FBI files, full of hearsay. They even hired \textbf{Robert Baer} of \emph{Syriana} fame as the project leader. Has anyone seen the \emph{Syriana} outtakes that claim Baer's first wife was a secret agent?

Well, the show was so bad, I stopped watching. Meanwhile, \textbf{Wikileaks} uncovered the results of a secret quest to find a Nazi base in Antarctica.\footnote{\url{https://wikileaks.org/gifiles/attach/49/49783\_.pdf}} They must be hiding something.

\paragraph{Sniffing your Neighbor's Anus}

Last month I stumbled upon a ``bright'' blog, having been mistakenly informed that it was a serious site. It was the same old, same old: atheism, science, consciousness as ``nothing but'' the firing of neurons. He made an interesting point that experiments have shown that there are neurons firing even before there is conscious awareness of a thought. Now from the esoteric perspective, that is not surprising, since the True Self is hidden and what we take to be our ``conscious'' self is secondary. Of course, the goal is to reverse that. I thought about adding a comment with that thought, but fortunately a wiser neuron dissuaded me.

The unanswered question, however, is why neurons have such different effects: what is the point of excruciating pain and sexual ecstasy? Do the neurons know? That is just a revival of ancient Gnosticism, viz., an evil Demiurge trapped consciousness in matter, while leaving it totally ineffective. What then is the point of pain? Why do we feel fear and anxiety if the conscious self is helpless to do anything about it?

Clearly, he believes he is just a mammal. I am happy to concede that about him, but that does not imply that all beings who look human are just mammals.

He then claimed that homosexuality is ``natural'', since animals in nature perform such acts. ``That is the only definition of `natural'\,'', he asserted. Now it may be natural, but mimicking dog behavior is not a good argument. Our religious leaders used to tell us to imitate the saints, or even Christ, but never animals. So this is what they call progress.

Apparently he is a flop with chicks, based on some of his rants. Perhaps he should try sniffing the behinds of the women in his office to get in their good graces. After all, it works for dogs, so it must be perfectly natural.


\flrightit{Posted on 2016-02-03 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Matt on 2016-02-04 at 00:45 said:}

Taking a look at the other thinkers interviewed by that writer on the nytimes site, and what does one notice? They all share the writer's ideological viewpoint. I guess he hasn't given the Aghori of the Mind experiment a try. Maybe that experiment is too difficult for individuals like him, or maybe he'd deem it racist- like everything else.

The apparent ``point'' one is to take away from the conversation between those two sophists: if you're a person of European descent and you avoid speaking about race, you're a racist, but then if you speak about race, that still makes you a racist. Forget about shifting the goalpost, the goalpost has been completely removed from the field.

\hfill

\texttt{Jacob on 2016-02-04 at 13:13 said:}

One of the best things about the right is the lack of over-seriousness. On the Left there is no room for laughter, because you either mock the enemies of revolution, or you are the enemy.

We are all tricksters. We've all deluded ourselves into some kind of grand performance, but some recognize it as a trick and some don't. I guess I'll just have to be a fool continuing in my folly.

\hfill

\texttt{Cologero on 2016-02-04 at 22:41 said:}

@Matt, the ``true north'' example is absurd, since there actually is a true north. Sailors will get lost without knowing it. Post-modernists are lost because they have no true north.

\hfill

\texttt{Max on 2016-02-05 at 18:55 said:}

The ``grid north'' is what maps or people tell you, it is a human convention. Magnetic north is dependent on earths magnetic field and is in motion, a living thing, more ``psychic'' in character. To establish magnetic north one first needs to account for compass deviation which is dependent on local tendencies, and then to account for magnetic declination which is dependent on global tendencies.

Although there is a ``True North'', the same can not be said for east and west. What is truly straight may appear bent, because when sailing the shortest route, projected onto a flat map, the course will take the form of a curve.

\hfill

\texttt{Alistair Fraser on 2016-02-06 at 07:23 said:}

On Magnetic north pole whether psychic or physical, here is an ultimate test that should define either / or existence . Go to Magnetic north pole (remembering that your not actually allowed to go there) When standing on the so called magnetic north pole, with palm of hand flat out horizontal , with a compass on it, should that compass needle now just spin or stall because it cannot locate the actual north direction ? Then according to the scientists version of the root of that magnetic force... .. if you then tilt that hand and thus compass 90 degrees, then surely the compass needle should now point at your feet , into the earth ? (or is the sky ? ) Then for a double proof of magnetic root , you can go to the south pole (remembering that your not allowed to go there) and repeat the same experiment which should result in the same result , i.e., tilt hand 90 degrees of horizontal and compass needle should point at your feet .

\hfill

\texttt{Mark Citadel on 2016-02-06 at 08:01 said:}

I've been pronouncing that name wrong for so long, I thought it was a pseudonym, not an actual family name! Good to remember for the future, though through force of habit I make no promises.

As for the old `homosexuality is natural' argument, this makes the dual failure of applying the dubious concept of `homosexuality' itself to animals who have absolutely no sense of sexual identity, nor any kind of identity. It is also to equate natural with moral, which is a gross distortion of natural law theory. You never see early Catholic theologians proposing that cannibalism is acceptable after observing the behavior of spiders.

\hfill

\texttt{Anonymous on 2016-02-06 at 23:31 said:}

@Scooby: what a convoluted lunatic comment! If I understand you correctly, what you're saying is that the brain is a passive condition through which the True Self consciousness manifests. This individual has laid out that point of view here:
\url{http://www.bernardokastrup.com/2011/12/brain-as-knot-of-consciousness.html}

\end{sffamily}
\end{footnotesize}

