\section{The Misuse of Intelligence}

\begin{quotex}
We have written this book, not for the irrational animals who, in their exterior, have the form a man, but in their image, in spirit, are evil and wild beasts, which is disclosed and exhibited by their properties; but for the image of man, for those who are budding forth out of the animal image with a human image that belongs to God's kingdom, and who would fain live and grow in the human image, in the right man. \flright{\textsc{Jacob Boehme}, \emph{Six Theosophic Points}}

\end{quotex}
\paragraph{Pleasing One's Masters}
At that time, I was tutoring a middle school student at a swank private school. Tuition is more than \$30,000/year, so it is populated with the children of the wealthy and the powerful. The graduates move on to Ivy League and other top ranked universities.

The topic was the number line, with both positive and negative numbers. A typical exercise was like this: if the temperature rises from -2 degrees to +2 degrees, how much did the temperature rise? Obviously, the answer is 4. Then we came to this problem: a baby was born in 2 B.C.; how old will he be in 2 A.D.? The quick answer given by the student was 4 years old. Aha, I pointed out. There is no year 0, so the baby will be just 3 years old. He eventually grasped the point, so I dared him to speak up in class the next day. I tried to convince him it would be to his own benefit.

The next time I saw him, I asked how it went in class that day. He sheepishly admitted that he had said nothing, but went along with the consensus answer. \textbf{The real lesson he learned was this: The purpose of intelligence is not the search for the truth, but rather the repetition of what the system wants to hear.}

\paragraph{Progressivism as Intellectual Decline}
The sociologist \textbf{Christopher Lasch} described the rise of the new class that seized the role formerly held by the aristocracy of the \emph{ancien regime}, although it lacked the sense of \emph{noblesse oblige}. This sentence describes that class with precision:

\begin{quotex}
Progressivism represented the ascendancy of secularized, deracinated élites enamored of the state, dismissive of tradition, suspicious of community, hostile to the traditional family, and antipathetic to the America that had endowed them with privilege. … secular rationalists who took art and politics as their religion. 

\end{quotex}
Lasch did not have a deep philosophical position, so his analysis ultimately falls short. That, however, does not discredit the historical trends that he documented. In opposition to these pseudo-aristocrats, Lasch advocated a form of populism. We don't know how he would have reacted to contemporary trends in populism, but we do know that the pseudo-aristocracy is vehemently in opposition to it.

There is an unfortunate misuse of intelligence. The promise of near universal education should have led to the rule of the best, a truly meritocratic state. Instead, this new elite was trained to despise the very people over whom, presumably, they were educated to rule. Although the quote above was made a few decades ago, the description is even more apt now. The points speak for themselves, so there is no need to go into more detail at the moment.

Those who are neither populists nor progressives can see both positions clearly. Unfortunately, just like the number line boy, the progressives have no interest in undermining the system that gives them legitimacy. Would that I could numb myself and become a progressive with a clear conscience. My social and family lives would become so much easier to navigate.

\paragraph{Types of Men}
\begin{quotex}
Man is as if he were three men, an \textbf{animal man}, as he is according to the senses, then a \textbf{rational man}, and finally the \textbf{highest god-like man}. . . One is the external, animal, sensual man; the other is the internal, rational man, with his rational faculties; the third man is the spirit, the highest part of the soul. \flright{\textsc{Johannes Tauler}}

\end{quotex}
Although Tauler's remark is consistent with St. Paul's teaching, it is never heard in exoteric preaching. Although in most cases this teaching is either unknown or poorly understood, there are sound reasons for suppressing it. Obviously, it is incompatible with the egalitarian attitude so prevalent today.

The main problem is that people will tend to overestimate their own progress. Anyone can claim, for example, to be the highest type of man, apart from any actual attainment. The Cathars used it to create a privileged class. Although people of the spiritual class claimed to transcend personal desires, in actuality they were libertines. This was justified by the claim that such liberties had no effect on them. The greatest danger is that the ``animal man" may be associated with a particular race or ethnicity, i.e., a biological rather than spiritual state.

We had a recent comment from a reader who asserted that his spiritual director advised him not to read the mystics directly. This demonstrates the point about suppression.

There are Greek terms—hylic, psychical, pneumatic—for Tauler's classification, although ``carnal" is often used for ``hylic". The effect of religious teachings is different when they affect the different types. For example,

\begin{itemize}
\item \textbf{Hylic}. The hylic personality will be most impressed with material and physical issues. For example, he or she will be interested in miracles or other unusual phenomena, patterns of clouds in the sky, historical events (which are for reason called ``literal"), and so on. 
\item \textbf{Rational}. This type loves to argue. He misunderstands spiritual combat to be some sort of intellectual battle of ideas. Instead of seeking wisdom like the philosophers, he enjoys ``making a case", as the sophists would do. He maintains centuries old divisions, regarding them as inspired somehow, instead of looking for reconciliation. 
\item \textbf{Spiritual}. The spiritual man wonders whether he should laugh or cry. I hope you find some sections of this post funny, or at least, you can tell which is which. 
\end{itemize}
\paragraph{Psychic Reality}
Like \textbf{Meister Eckhart}, Tauler understands that the real birth of the Son is in the soul when he writes:

\begin{quotex}
God accomplishes all His works in the soul and gives them to the soul; and the Father brings forth His only-begotten Son in the soul, as truly as He brings Him forth in eternity, neither less, nor more. 

\end{quotex}
That would disturb the materially minded, the hylic personalities, who believe that the physical is more real than the psychical. Whatever his faults and limitations may be, Carl Jung seems closer to Eckhart and Tauler in his insistence that the psyche is \emph{real}. For many, the notion that something happens in the soul means that it is less real, it is imaginary, it is all ``in your head". In opposition to that notion, Jung makes this claim:

\begin{quotex}
It does not matter at all that a physically impossible fact is asserted, because all religious assertions are physical impossibilities. If they were not so, they would necessarily be treated in textbooks of natural science. But religious statements without exception have to do with the reality of the psyche and not with the reality of \emph{physis}. 

\end{quotex}
Now a \textbf{Rene Guenon} might object that religious statements are disguised metaphysical statements. Nevertheless, Jung's observation is a good half way point, and may actually bear more fruit in practice.

\paragraph{Science and the Carnal Man}
Rudolf Steiner, in his commentary on Tauler, describes the limitations of science:

\begin{quotex}
Man is entangled in the world of the senses and in the laws of nature, by which the world of the senses is dominated. He himself is a result of this world. He lives because its forces and substances are active in him, and he perceives and judges this world of the senses in accordance with the laws by which it and he are constructed. 

\end{quotex}
This is the circularity of the deracinated, secularized, rational, scientific man. He is fundamentally a carnal man, a highly intelligent animal. However, all his science is self-confirming, so he is never able to break out of the circle to reach transcendent, or higher, knowledge. Steiner notes this anomaly:

\begin{quotex}
Do we not stand above all mere conformity to natural laws when we survey how we ourselves are integrated into nature? We \emph{see} with our eye in accordance with the laws of nature. But we also \emph{understand} the laws in accordance with which we see. We can stand on a higher elevation and survey simultaneously the external world and ourselves in interplay. Is not then a nature active within us which is higher than the sensory-organic personality which acts according to natural laws and with natural laws? In such activity is there still a partition between our inner world and the external world? 

\end{quotex}
A recent example is this study that finds it might not be consciousness that drives the human mind\footnote{\url{https://www.sciencealert.com/research-finds-it-might-not-be-consciousness-that-drives-the-human-mind}}. The research notes that the carnal mind ``gives us a feeling of ownership and control over the thoughts, emotions and experiences that we have every day". On the contrary, the esoterist, the Hermetist, the yogi has known that for millennia, so the scientists are wasting their time. The mystics understand that thoughts and feelings spontaneously arise, in a manner not unlike our experiences of the external world, so that they are not part of our ``inner being". They have developed practices to observe the arising of thought and the means to resist the internal programming that they represent.

Their conclusion is quite odd:

\begin{quotex}
the contents of consciousness are generated ``behind the scenes" by fast, efficient, non-conscious systems in our brains. 

\end{quotex}
We are left wondering how exactly does the brain produce thoughts. Is there some chemical formula or equation of physics that describes each thought? Since our lived world is the creation of our thoughts, it would be incredible if someone could describe a modern city as the result of some efficient system in the brain.

The denial of the immateriality of the soul is a misuse of intelligence.

\paragraph{A Liberal Mother Watches TV}
A young mother complained to me recently about the current crop of Netflix shows. It seems that they all portray the parents as bumbling idiots while the children are bright and sassy. (I have never watched these shows.) She considered them to be a bad influence on her young son, who was already prone to ``talking back". Genuinely surprised, I exclaimed, ``Wow, I thought you used to be a liberal."

She indignantly asserted that she was still a liberal, but that it is a matter of making ``good choices". Although this woman is quite intelligent in the conventional sense, she is like the number line boy who does not think very deeply about things. Unlike her, I see that Liberalism is the rejection of authority. This has been its history.

In this case, it is the rejection of the authority of the parents. For the liberal, this parental authority is not self-validating. From the child's perspective, I suppose, he depends on his parents for survival. However, beyond that need, they might still be idiots. Ultimately, authority depends on God, and parental authority is the first earthly commandment. The mother cannot see the long term consequences of her views, even when they are staring her in the face by way of Netflix shows.

The denial of the existence of God is a misuse of intelligence.

\paragraph{Menstruation}
In the seventh grade, a boy dared me to ask one of the pretty girls in the class to explain a Kotex to me; I was genuinely ignorant. This caused embarrassment both to her and to me, so I have since became cautious about being ``set up". Needless to say, because of that lesson, I remain quite ignorant, despite intimate relationships over the course of my life (i.e., before I decided to become ``saintly").

Now the Eastern Orthodox forbid menstruating women from entering a church. I presume, or at least hope, that is done on the honor system\footnote{\url{https://en.wikipedia.org/wiki/Honor_system}}. Now the Romans don't have that restriction, so it is probably one of the many heresies they accuse us of. Presumably, that custom derives from the ancient Hebrews, who had several rules relating to that topic.

I recently learned, via a youtube video, that the Brahmans of India forbid menstruating women from public worship. This is not surprising, considering that the Hebrews ultimately descend from Hindus. Now this Brahman offered a likely story\footnote{\url{https://en.wikipedia.org/wiki/Timaeus_(dialogue)\#Nature_of_the_physical_world}}. He said that back in the olden days, the scent of a bleeding woman would have attracted wild animals, so keeping her at home was a matter of safety and precaution.

Around the same time I learned that at the French courts the scent of menstrual blood was considered erotic. I don't know how the courtiers controlled the scent or flow before the days of Kotex and Tampons, but I don't think men today have the same reaction. Moreover, it seems more likely from a Darwinian perspective that the scent of an ovulating woman would be more of an aphrodisiac. But then again, the French aristocracy were approaching degeneracy.

\paragraph{Fapping on the Web}
I was invited recently to a secret facebook group that claimed to be transgressive, on the assumption apparently, that would appeal to me. Now a few members would post soft porn pictures of women every evening … it must have taken them hours of searching through other facebook pages to find the most transgressive pictures.

In order to encourage them to desist from that practice, I commented: ``You guys sure like to fap!" To my surprise, instead of an expression of remorse, shame or embarrassment, they boasted about how much pleasure they derived from fapping. After all, they pointed out, they would have no chance with such attractive women in ``real life", i.e., carnally.

Since the leadership of the group was in the hands of converts to Orthodoxy, I should have pointed out the Orthodox teaching on fapping prior to my resignation. After fapping, in order to get right with God, a man needs to make 50 prostrations while making the sign of the cross. I suppose that for those callow fellows, that is a small price to pay.

The lesson: no matter their level of intelligence or learning, most men can't think of anything better to do than to masturbate.



\flrightit{Posted on 2017-11-27 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Fred on 2017-11-27 at 15:05 said: }

As I understand it he told me ``If you are still Hylic you will misunderstand the mystics".

He didn't say hylic but that's the meaning: you can't understand what they mean by studying what their language says literally, to understand the Fathers you need a partially purified intellect.

No illumination without purification first.


\hfill

\texttt{Max on 2017-11-28 at 18:41 said: }

The average person's limited intellect is exhausted by keeping up with what you are supposed to think. If directly commanded what is right and wrong those resources could be freed up for more worthwhile pursuits. The social consensus system ensures that it is only possible to rule through lies, whereupon the new elite assumes that the lie is a sign of superiority while it actually signifies privation. Empiric science cannot recognize authority as that is spiritual, a ``nothing", and more often than not the recipients themselves are unaware of its true source and nature.

Here comes a contemporary lesson in how not to become ``king", unusually perceptive for a presumed [recovering?] progressive:

``To help make my point, let's assume for a moment that Ken, instead of going for his singing career, had set out to climb Mount Everest — and failed. In that case, the public would not have set out to break him, and he wouldn't have allowed himself to be broken. The horrible thing about challenging the monarchy is that it is not `real' in the way a mountain is real. There is no worse defeat imaginable than loosing to `nothingness,' because not only have you lost, you've also had to grovel and beg for forgiveness from `nobody.' Ken set out to fight windmills — and lost. It's no surprise that after such a loss, one hardly remains a man.

Thus, today's monarchy is no laughing matter, and definitely not something to be toyed with. It has overwhelming metaphysical authority. Challenge the monarchy, and your losses will be greater than you ever imagined. Winner takes all and the looser is screwed for good, left in a state of eternal disgrace from which there is no remedy." http://www.askergren.com/monarchy.html

Authority even this late comes down to us from a primordial sacred source no matter how erratic and wayward, – justified or unjustufied, consciously seized or not – its descent may have been. The King, as Cologero pointed out via Carlyle in an earlier post, is he who can, making a necessity to resort to lies indicative of the opposite. In case someone can ``tell it like it is" and remain unscathed take notice.


\hfill

\texttt{Gabe Ruth on 2017-12-01 at 17:13 said: }

Conceiving as a result of intercourse during menstruation would be pretty miraculous, so there is no evo-psych story to justify the degeneracy of the French court. 

Those crazy Orthos are something else. I wonder if Rod Dreher makes his wife stay home at that time of the month. I'm partial to the purported approach of the Alawites to this thorny issue. If you don't have a soul, why would you need to go to church?


\hfill

\texttt{Farewell on 2017-12-05 at 11:32 said: }

``apostates to Orthodoxy"

I want to thank you for this, as lately I've been realizing just how heretical this site truly is. and with the latest post, my doubtrs have been finally confirmed. Thank you for providing occasionally some good insights and sources. It was nice reading you all and may God have mercy on our souls.


\hfill

\texttt{Cologero on 2017-12-06 at 21:27 said: }

Mr. Farewell, clearly you don't know what the word ``apostate" means.

And it is sweet that you don't even get the irony of calling me a heretic, something predicted in the text. Of course, you have provided no example of heresies.

I suppose a ``farewell" is good coming from you, since your fathers were more interested in fratricidal wars over alleged heresies. You can sink back into your safe world, thereby avoiding any future ``good insights".


\hfill

\texttt{My apology on 2017-12-08 at 13:41 said: }

I apologize for the previous comment. I was going through hard and turbulent time, and still am, so my mind wasn't clear. It's okay if you ban me from this site. Or if you can delete comments, that would be even better. Sorry once again.


\hfill

\texttt{Caveman on 2017-12-10 at 12:48 said: }

I like your application of the term `pseudo-aristocracy' to the new elites. Analogous to Guénon's categories of pseudo-initiation and counter-initiation, perhaps we might add the term `counter-aristocracy' to designate bloodline descendants of the old nobility of ancient right who fell or sold out into consciously working against Tradition and the Christian project. Such counter-aristocrats (``revolting kshatriya" as Guénon would say) might have preceded the rise of the pseudo-aristocrats. 

«But then again, the French aristocracy were approaching degeneracy.»

`Approaching' is a bit of an understatement. The royalty and the more urban nobility in particular were extremely degenerate. I have understood that much of the rural nobility was healthier and more traditional still. 

«The promise of near universal education should have led to the rule of the best, a truly meritocratic state. Instead, this new elite was trained to despise the very people over whom, presumably, they were educated to rule.»

A universal education that is itself not universal—representing sacred and holistic knowledge (where something like mathematics for example is not allowed to develop into an isolated quantitative science)—could never achieve such an aim. It just proliferates an ignorance that sees itself as enlightened: and what could be worse than an inflated mental `intelligence' deprived of the light of the Heart? This is the kind of `intelligence' that leads gradually to the robotization of the world. The results of the educational system providing the formation of the present elites reveals that the development of this system even from its earliest seeds has not been driven by a force that is in service to the True and the Good.

«The Cathars used it to create a privileged class. Although people of the spiritual class claimed to transcend personal desires, in actually they were libertines. This was justified by the claim that such liberties had no effect on them.»

This is very much alike to left-hand Tantrics in the East.


\end{sffamily}\end{footnotesize}
