\section{Tempted by Calypso}

\begin{quotationx}
If you see anyone blinded by the illusions of his empty and Calypso-like imagination, seized by the desire of scratching and delivered over to the senses, it is a brute, not man that you see.

If you come upon a philosopher winnowing out all things by right reason, he is a heavenly, not an earthly animal.

If you come upon a pure contemplator, ignorant of the body, banished to the innermost places of the mind, he is not an earthly, nor a heavenly animal: he more superbly is a divinity clothed with human flesh.

\flright{\textsc{Pico della Mirandola}, \emph{Orations on the Dignity of Man}}
\end{quotationx}

\textbf{Man the Brute}. In popular culture, the handicapped person who excels at athletics, an elderly woman who parachutes out of an aeroplane, or an aging model who poses in a bikini, all elicit widespread admiration. While this a testament to the human will, it still falls short of man's higher possibilities. A man who has lost his physical prowess can still become a philosopher or a contemplator, although the public will ignore him.

Note how Pico anticipates \textbf{Rene Guenon}. In his categorization, the philosopher actualizes all his possibilities as man and the contemplator actualizes his transcendent possibilities.

\textbf{Rock `n Roll Models}. If you can't become a philosopher, then how about a scientist, i.e., a ``natural philosopher''? Metalheads have some role models: Brian May of Queen became an astrophysicist as did several other musician/scientists\footnote{\url{http://www.thebestcolleges.org/doctors-of-rock-10-musicians-with-phds/}} .

\textbf{Pornography for the young}. Plato was distrustful of the corrosive effect of the Poets in society. Although Ariana Grande grew up just a few miles from where I live, I was not familiar with her work. Apparently her songs and show, although targeted at young girls, are somewhat pornographic\footnote{\url{http://www.barnhardt.biz/2017/05/27/manchester-the-brutal-truth/}}. Yet we fear those who kill the body, yet remained unconcerned about those who kill the soul.

\textbf{Death of Cultures}. \textbf{Oswald Spengler}'s masterpiece, \emph{The Decline of the West}, is on every new/alt right reading list.
So why is everyone surprised to witness Western culture apparently self-destruct, not necessarily with a bang, but certainly with a whimper?

\textbf{Conspiracy Theory}. As for the noted symbiosis between liberalism and Islam, this explanation may seem far-fetched; nevertheless, a well-crafted conspiracy theory is an art form. Manly P Hall --- who knows something about secret Masonic orders --- leaves some clues in \emph{The Fraternity of the Rose Cross}. He says this about Christianity:

\begin{quotex}
Whatever profundity is found in Christian philosophy is due primarily to this infusion of pagan ideas. Conversely, the shallowness of Christianity is the direct result of the clumsy efforts made to improve upon the ancient doctrines.
\end{quotex}

On the other hand, his opinion on Islam is quite the opposite; it improved on paganism.

\begin{quotex}
In his cavern on Mount Hira, Mohammed prayed not for new truths, but for old truths to be restated in their original purity and simplicity in order that men might understand again the primitive religions: God's clear revelation to the first patriarchs. ... The Prophet was but the reformer of a decadent pagandom, the smasher of idols, the purifier of defiled Mysteries.
\end{quotex}

Hall then lists several Europeans who were initiated in the mysteries by Arabians: Paracelsus, Cagliostro, Comte de St. Germain. Those Rosicrucians were, according to Hall, the precursors to the overthrow of the established order of Crown and Altar. Those are the dots ... \textbf{The Opposite of a Revolution}. Reactionaries, then, look for a restoration of the established order of things.

\textbf{Christendom}. So-called ``White Nationalism'' is a misnomer, because there is no white race, but rather white races. Instead, Europeans used to identify themselves as ``Christians''. For example, Queen Elizabeth granted Walter Raleigh the right

\begin{quotex}
to explore, colonise and rule any ``remote, heathen and barbarous lands, countries and territories, not actually possessed of any Christian Prince or inhabited by Christian People
\end{quotex}

Note that she did not refer to lands possessed by ``white people'' and, a fortiori, not even to ``European people''. Nationalism cannot be based solely on genetics, but solidarity is achieved only with a spiritual unity.

\textbf{Differentiation and the Kali Yuga}. One of the prophecies about the Kali Yuga is the lack of differentiation. That is, the traditional categories of caste, nation, religion, even gender become blurred. So there should be no surprise about that. The difference comes in the valuation: liberals promote undifferentiation as a sign of progress, but reactionaries see it as a symptom of spiritual decline.

However, the West is leading the way into the Kali Yuga. According to Tradition, primitive people are not evolving upwards, but are rather the residues of once great civilizations. That may be the fate of Europeans. What will kill the project is that other peoples are not at that point. Hence, they manage to retain their identities, unlike the majority of Europeans.

The way out is for a small group to live the opposite of a revolution. Judge your allies by traditional standards. Are they brahmins, warriors, or merchants? Not the latter, because the opposite of a revolution is not a business model.

\textbf{Universality and identity}. The dogmas about the characteristics of the Church have been transferred to Humanity as a whole. For example, traditionally only the baptized could be considered to be ``children of God'', now that applies to everyone. The Church says that all its members are ``one'', in the sense that everyone plays a role in the Body of the Church, which is hierarchically arranged. The Liberal says that we are all ``one'' in our humanity; we are all equal, so there cannot be a hierarchy.

As recently as a hundred years ago, St. Pope Pius X could categorize those outside the Church: ``Infidels, Jews, heretics, apostates, schismatics, and the excommunicated.'' Obviously, this is an embarrassment to any bishop today. Nevertheless, even the liberal regards some to be beyond the pale of humanity; these are called the deplorables.

\textbf{Sexual Revolution}. Promoters of sexual revolution are no part of any ``Right'' whatsoever, but are located somewhere on the Left.

\textbf{Orthodox Apostates}. There has been a minor movement of apostates from Rome to Orthodoxy. Now this may be legitimate in a ``state of emergency'', i.e., there is no local Roman parish that can be trusted. That does not mean that vitriol aimed back at the Roman Church is justified, especially when it includes false witness. From the esoteric perspective, there is but one Tradition; jurisdictional disputes among bishops cannot alter that.

Also, from a Traditional perspective, one follows the religion of one's ancestors or legitimate leaders. If you are Nordic or Roman, that religion is Rome. There is too much to give up: e.g., Thomism, Chivalrous orders, mystics of a high order like Meister Eckhart, Dante, centuries of Art, etc. I suppose if Bruce Jenner can decide to become a woman, and Rachel Dolezal can become black, then a Western European can wake up and become a Ukrainian.

By the way, I sometimes attend a Uniate Byzantine Rite mass, in which the \emph{filioque} is omitted. So it is hardly a ``deal breaker''.

\textbf{The Power of Hate}. The Voice of the People wants to restore hate to its rightful place\footnote{\url{https://voxday.blogspot.com/2017/06/darkstream-hate-is-christian-virtue.html}}.

\begin{quotex}
There are six things that the LORD hates, seven that are an abomination to him:

\begin{itemize}

\item haughty eyes

\item a lying tongue

\item hands that shed innocent blood

\item a heart that devises wicked plans

\item feet that make haste to run to evil

\item a false witness who breathes out lies

\item one who sows discord among brothers.
\end{itemize}

\flright{\textsc{Proverbs 6:16-19}}
\end{quotex}

\textbf{Hormone-free Humans}. Supermarkets charge premium prices for hormone-free dairy, eggs, and meat. Soon, hormone-free humans will be a rarity.

\textbf{Imitation and Conviction}. Arthur de Gobineau theorized about culture creators and culture consumers.

\begin{quotex}
The masses are invincibly ignorant of the Western world and its movements. ... There is a great difference between imitation and conviction. Imitation does not necessarily imply a serious breech with hereditary instincts; but no one has a real part in any civilization until he is able to make progress by himself, without direction from others.
\end{quotex}

Have any new/alt right groups been able to define a High Culture for our age? Hereditary instincts do not suffice.

\textbf{Spiritual Combat for the Masses.} Paula White in \textit{The Power of Thoughts}\footnote{\url{https://wwwt.youtube.com/watch?v=g4CBkEqPVGg}} teaches the power of thought and dominion. Before dismissing it, consider whom it motivated. It is the opposite of resentment and victimhood for the masses.

\textbf{Marquis of Queensbury rules for online debates}. Online arguments and faux television news debates are pointless because they lack a referee to monitor the arguments and a judge to declare a winner. And, unlike diving or boxing, for example, there is no standard for judgment. Who would be foolhardy enough to box or play American football without a referee?

\textbf{Faux Outrage.} It is easy to write a post pandering to a sense of outrage. It is even easier to excite prurient interest. It is more difficult to appeal to the intellect, although it is worth the effort.

\textbf{The Feminine Universe.}
Aristasia is a female-only Utopia\footnote{\url{https://aristasia.wordpress.com/2007/09/17/the-feminine-universe/}} based on the writings of Rene Guenon and Ananda Coomaraswamy. Perhaps a more traditional reconciliation of the sexes may be possible if the Men's Movement rejects MGTOW, PUA, or game, and embraces innocence, elegance, cuteness (kawaii), and modesty\footnote{\url{http://www.daughtersofshiningharmony.com/}} instead.

\textbf{The Inevitable End}. If you've reached this far, you are on the way to becoming a Philosopher, a heavenly animal.


\flrightit{Posted on 2017-06-05 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{James on 2017-06-06 at 11:55 said:}

Thanks so much for another great post.

\hfill

\texttt{Andrew G Taylor on 2017-06-06 at 13:39 said:}

``... if the Men's Movement rejects MGTOW... and embraces innocence... '' Inner and outer. This is something I'm trying to balance in my life currently.

\hfill

\texttt{Auld Wat on 2017-06-10 at 10:39 said:}

The last link appears to be dead. Thanks, its been fixed:

\textit{Daughters of Shining Harmony}\footnote{\url{http://www.daughtersofshiningharmony.com/}} 

\hfill

\texttt{Cologero on 2017-06-11 at 21:02 said:}

``At this point it is not about destruction, but of constructing oneself at every moment, through a renewed act of love and renunciation,
\textbf{on a plane higher than oneself} where there is the impassibility
of the Watcher or, better, the Lord, outside any storm or tumult whatsoever, whether interior or exterior.''

\hfill

\texttt{geirón on 2017-06-12 at 08:10 said:}

Faith is many things. There is fideism and there is Fides, the goddess of trust and bona fides (good faith) in Roman Faith. 1  2 She was one of the original virtues to be considered an actual religious ``divinity''.

\hfill

\texttt{Max on 2017-09-05 at 04:56 said:}

There is no reason to beat down on Christianty while simultaneously promoting Mohammedanism unless the aim is precisely the destruction of the faith. If one were simply a lover of Mohammedans there is no need to waste energy on Christianity since the option lies open to move to the Middle East in order to pursue the dream, so to speak. Negative obsession with Christendom demonstrates a dependence, and the sympathy is due to both denying such a relationship.

\end{sffamily}
\end{footnotesize}

