\section{A Necessary Interruption}

\begin{quotex}
In the right order of nature, the flesh is subject to the spirit and not the reverse.
\flright{\textsc{The Cloud of Unknowing}}
\end{quotex}

\paragraph{UPDATE 12 March 2021}
Cologero's surgery was a success. He is now in great health and, God willing, should be for some time ahead. He thanks Saint Elizabeth of the Trinity for her intercession.

Since Cologero is dealing with a serious health issue, he is undergoing rigorous tests. On top of that, he got dragged into a family squabble related to a small inheritance. So he asked me to fill in until he is able to get back to work. He has several posts in various states of readiness. I won't presume to complete them. However, I'll update some news items and discuss the future of our project.

\paragraph{Board meeting}

We took advantage of the Full Moon last week to channel \textbf{Losang Shenphen}, Cologero's spiritual director, to call a board meeting to discuss the last post as well as the possible future of Gornahoor. Cologero wants to offer a full confession, including a list of errors and refutations. I don't see that as beneficial at all. I proposed instead to commission a writer to summarize the entire web site.

My opinion is that it would help to re-organize the writings into a coherent system. Unfortunately, few readers know how to decode an esoteric text. Moreover, we have heard from a great many readers, who would be interested in the material, except that they find it too difficult to read. Cologero chalks that up to the nature of things. I, on the other hand, believe it can be pre-digested to make it more practical ... i.e., to make it more in the style of a Deepak Chopra or an Oprah Winfrey.

Losang prefers not to worry about it, since a web site is by nature impermanent. He claimed he can gather the best posts into the Golden Book described by \textbf{Ann Catherine Emmerich} and hide it in the Mount of Prophets. We often have trouble determining when he is kidding or not.

\paragraph{Kidney Failure}

Cologero was informed last week that his kidney function has further deteriorated, to the point that he was placed on the Transplant List, waiting for a kidney, whether quick or dead. This has obviously changed his priorities. He still would like to translate some essential Hermetic texts, but that would require a sustained effort. Maintaining multiple web sites is a burdensome distraction. Of course, the Gnosis group is high priority, since a text alone is insufficient.

\paragraph{Post 1001 and Planning for Death}

The Board arrived at an impasse, and post 1001 may never appear, at least not soon. Perhaps in a generation from now, a professor, desperate for a project, may try to unravel Gornahoor.

One proposal was that after Post 1001, Gornahoor should continue in a different format. The obvious choice would be to integrate all the sister sites into Gornahoor, since it has established to a tiny degree, what is called a ``brand''. Then we would invite other writers to contribute, as used to be the case a couple of years ago.

We thought about buying a 5-year hosting plan, in the event Cologero becomes unable to continue the maintenance functions. It would still require regular WordPress updates. Instead of that, we could convert the entire site to static HTML pages, but that would make it difficult to update. No final decision has been reached. Apparently, we would need to appoint a Trustee for the site, assuming that is even a legal option.

\paragraph{Personal Library}

As you can imagine, Cologero has amassed quite a nice library on topics discussed on the web sites. It is certainly not full of popular texts, but a certain type of personality would be excited by it.

He is concerned that his heirs will just box them up and take them down to the used book store to cash them out. That possibility saddens him. He is not ``famous'' enough that a university would want the books. He is open to ideas.

\paragraph{Frater M}

\textbf{Francis Mercuri}, who published some posts here under the moniker Frater M\footnote{\url{http://www.gornahoor.net/?author=160}}, died in police custody two weeks ago in Daytona Beach under mysterious circumstances. There has been a news blackout; I can't find a notice in the local newspapers not in the police blotter. In other circumstances, the death of someone in custody might lead to street demonstrations. I guess Francis does not fit the profile.

A mass will be said for him at the Basilica of St. Paul on April 3\footnote{\url{http://www.legacy.com/obituaries/postgazette/obituary.aspx?n=francis-anthony-mercuri\&pid=188392519\#sthash.Mnbea84N.gbpl}}.

\paragraph{Entryism}

The \textbf{Society of St. Pope Pius X} has recently bought a church property in West Palm Beach. There are a mere couple of dozen participants at the moment, although the new building can accommodate 100 or more. If only there could be a concentration of people who could use it as a spiritual base. Ask Cologero for details.

\paragraph{City of the Sun}

In the Utopian society, The City of the Sun\footnote{\url{http://www.gornahoor.net/?p=518}}, \textbf{Tommaso Campanella} suggested that older women could volunteer to sexually service the young men. The USA has also adopted that system, although here they are called ``high school teachers''\footnote{\url{https://www.bing.com/search?q=high+school+teacher+male+students\&form=EDNTHT\&mkt=en-us\&httpsmsn=1\&refig=bcffabfdaa9e44a7c11a858a0f253fdc\&sp=-1\&pq=undefined\&sc=0-27\&qs=n\&sk=\&cvid=bcffabfdaa9e44a7c11a858a0f253fdc}}.

\paragraph{Electric Dreams}

Amazon Prime has recently adapted some Philip K Dick stores in the series Electric Dreams\footnote{\url{https://www.amazon.com/Season-1-Official-Trailer/dp/B075NTXMN9/ref=sr_1_1?s=instant-video\&ie=UTF8\&qid=1520825550\&sr=1-1\&keywords=electric+dreams}}. In \emph{Episode 7 The Father Thing}, they have slyly inserted an alt-right meme:

\begin{quotex}
our young hero Jack must save the human race while coming to terms with the disillusionment of childhood as he is among the first to realize that humans are being replaced by dangerous aliens
\end{quotex}

The alien who replaced Jack's father gave quite an inspiring speech about how their superiority to the indigenous humans would be better for the Earth. I'm sure they had some valedictorians and entrepreneurs in their flying saucers. Nevertheless, Jack honored his father and rejected the proposal.

\paragraph{Epicenity}

If you believe that boys and men are becoming more effeminate, the trend originates in the 16\textsuperscript{th} century, if you believe Campanella. He pointed to the prevalence of pederasty, sodomy, and whoring among the literary elite. That sound quite contemporary. Or perhaps, it has always been that way, so there is no ``trend''.

\paragraph{Rule of Women}

International Women's Day just passed a few days ago. All day, we were treated to statistics and exhortations about giving women more power and money in society. That is a generally accepted good, except that the God of the Old Testament regards it as a punishment. He will remove the strong and reliable men as this threat makes clear:

\begin{quotex}
For behold the sovereign the Lord of hosts shall take away from Jerusalem, and from Juda the valiant and the strong, the whole strength of bread, and the whole strength of water. \emph{The strong man, and the man of war, the judge, and the prophet, and the cunning man, and the ancient.} The captain over fifty, and the honourable in countenance, and the counsellor, and the architect, and the skillful in eloquent speech.
\flright{\textsc{Isaiah 3:1-3}}
\end{quotex}

This will be their punishment:

\begin{quotex}
And I will give children to be their princes, and the effeminate shall rule over them.
\flright{\textsc{Isaiah 3:4}}
\end{quotex}

This is the result:

\begin{quotex}
As for my people, their oppressors have stripped them, and women have ruled over them. 
\flright{\textsc{Isaiah 3:12}}
\end{quotex}

Is it any wonder that Marcionism has taken such hold?


\flrightit{Posted on 2018-03-12 by Aeneas}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Sometimes on 2018-03-12 at 12:35 said:}

Frater M had numerous run-ins with Daytona police, including reckless driving while evading a police officer; a racially motivated beating of a Jewish teenager; beating a fellow skinhead; as well as an arrest in Pittsburgh when the apartment he was staying in was raided and weapons were seized in connection with a neo-nazi group.

He was an excellent acupuncturist, and an insightful Hermeticist.

\hfill

\texttt{james on 2018-03-12 at 15:21 said:}

My heart is really saddened by this news.

\hfill

\texttt{Ana Madriz on 2018-03-12 at 20:45 said:}

I am very sad to hear about Cologero's illness. I pray for your fast recovery.

\hfill

\texttt{Bernard on 2018-03-13 at 11:35 said:}

My prayers to Cologero and his family.

\hfill

\texttt{Bill on 2018-03-14 at 12:35 said:}

My prayers are of course with Cologero, but also with everyone, living and dead, who have been involved with this website. It is a source of unending inspiration and has been invaluable in my own education and formation. My every best wish to Cologero.

\hfill

\texttt{geirón on 2018-03-16 at 12:21 said:}

Best wishes to Cologero's health.

\hfill

\texttt{Malcolm Pollack on 2018-03-16 at 12:55 said:}

Oh dear -- I am very sorry to hear of Cologero's illness. Please give him my heartfelt wishes for a swift and lasting recovery.

\hfill

\texttt{geirón on 2018-03-17 at 14:49 said:}

Why not allow an old man his confessions?

\hfill

\texttt{sindre on 2018-03-17 at 18:15 said:}

I hope you recover.

\hfill

\texttt{Mikkel on 2018-03-28 at 11:53 said:}

Feels like a flood of cold water. I hope Cologero's health improves as much as he wishes, my regards to him and his family.

To that eventual time, may we suggest a book sale through this site? A book competition entry contest? Perhaps a book lottery?

\end{sffamily}
\end{footnotesize}

