\section{Living in Interesting Times}

\begin{quotex}
When man knows himself in his deepest essence, that is, in the center of his being, then at the same time he knows his Lord. \flright{\textsc{Rene Guenon}, Know Thyself\footnote{\url{https://gornahoor.net/?p=58}}}

Buddhist meditation, but above all that of Zen, seeks not to explain but to pay attention, to become aware, to be mindful, in other words to develop a certain kind of consciousness that is above and beyond deception by verbal formulas, or by emotional excitement. \flright{\textsc{Thomas Merton}, \emph{Zen and the Birds of Appetite}}

\end{quotex}
\paragraph{Know Thyself}
I'm not opposed to book learning, since I have read a great many books, but the mere accumulation of erudition brings one no closer to gnosis. That is why futile debates about paganism, Orthodoxy, etc., are of secondary interest. This is the primary issue: which path brings one closer to self-knowledge?

\paragraph{Death to the World}
To regard the world process as causally closed is to condemn oneself to be blown around by the arbitrary whims of opinion and power. There is no solution to the problems of life within the world system. The world is the play of shadows on the wall. One must be attentive to the one way of escape, which is dying to the world, but living a higher life.

\paragraph{Love by Government Fiat}
\begin{quotex}
Because your attack thoughts will be projected, you will fear attack. And if you fear attack, you must believe that you are not invulnerable. Attack thoughts therefore make you vulnerable in your own mind, which is where the attack thoughts are. Attack thoughts and invulnerability cannot be accepted together. They contradict each other. \flright{\emph{A Course in Miracles, Lesson 36}}

\end{quotex}
New Ager Marianne Williamson doesn't seem to understand \emph{A Course in Miracles} [ACIM], on which she bases her spiritual and political positions. Her use of attack thoughts is totally contrary to the teaching. Is it merely a coincidence that her political platform is identical to all the other candidates, or is it more likely that ACIM has nothing at all to do with it.

For better or worse — and I'm not recommending ACIM — I did all the ACIM exercises for one year, in the 80s. The result is that I avoid speaking the way Marianne does, with all that name calling. She uses LOVE as a weapon to claim a faux superiority, as though no one else is capable of Love. Or worse, those who love gather together in a partisan way.

God commands us to love, but the government cannot command love. The idea that reparations, or any other government program for material gain, will lead to love is delusional.

\paragraph{Unfulfilling sex}
Bonald, Jr. wrote recently:

\begin{quotex}
John Zmirak admitted that unmarried people like he then was, struggling to obey the Church's teachings on chastity, were getting annoyed by converts going on about all the hot, mind-blowing sex they had before finding Jesus, and how it was so unfulfilling. Who doesn't want to be the guy who's so unfulfilled, so jaded, with being able to get pretty girls enthusiastically into bed with him? \flright{\textsc{What Could be Better}\footnote{\url{https://bonald.wordpress.com/2019/07/26/what-could-be-better/}}}

\end{quotex}
Even Roosh, who had established for years a reputation as a pick-up artist, has had a religious conversion and repudiated his earlier life of non-committal sex. His conversion to Eastern Orthodoxy seems genuine, actually more genuine than the cradle Orthodox whom I've known. On the other hand, now in his 40s with a beard turning grey, his luck as a PUA has undoubtedly run out. Without money and fast cars, the young pretty girls won't have much interest in him.

Now he is complaining that he can't find a good woman to marry, despite all his alleged seductive skills. Teresa Chicalinda made the point that only a certain type of woman gets ``picked up"; the PUA has only learned how to identify them.

Back to Bonald's point:

\begin{quotex}
Sex is a natural desire, and is a good, unlike the murder of innocents, for example, which is intrinsically evil. It is attractive because it is fulfilling, it is fun, and you meet interesting people. The issue is not sex itself, but its appropriateness; that distinction is too abstract for most people. It is the most difficult drive to control. 

\end{quotex}
Yet, the Primordial Tradition demands mastery over the sex instinct: e.g., Hinduism and Buddhism both teach sexual control in order to achieve deliverance. This is not because sex is evil in itself, but because it dislodges the Real I from its rightful place and often leads to unpleasant outcomes. Abstention from meat is a form of intentional suffering that aids in developing attention and self-mastery. Over time, that led to the idea that vegetarianism is superior for spiritual development. In that case, however, giving up meat is no longer a sacrifice.

This is analogous to sexual restraint One masters the sex drive, not because celibacy is superior, but because there is value in intentional suffering.

\paragraph{Mary Magdalene}
The legend is that Mary Magdalene was a repentant prostitute, who eventually ended up in France where she lived a life of penance. A New Zealand professor was interviewed about Mary for Crux\footnote{\url{https://cruxnow.com/vatican/2019/07/27/could-mary-magdalene-be-part-of-the-modern-metoo-movement/}}, but I heard the interview on the radio.

It was a typical postmodern performance, short on argument, but relying mostly on psychological nonsense and alleged power relationships. She asserted without ground that ``sexual sins are the worst". That is hardly the case for the reasons cited above. Dante placed sexual sinners at the first ring of Hell proper (there is no suffering in Limbo), precisely because it is the least serious sin. The interviewer was clearly inadequate to the task. What exactly did the professor mean? Is she implying that adultery, for example, is perfectly normal and harmless? How about pornography, pederasty, or bestiality? Ultimately, it is not a question of drawing a line, but rather of where exactly to draw the line.

The professor did end on a sound note when she claimed that the sin of pride was worse. Of course, pride is claiming to know better than 2000 years of Tradition.

\paragraph{Chief Feature}
An important aspect of spiritual growth is Finding one's predominant fault\footnote{\url{https://rorate-caeli.blogspot.com/2019/07/finding-ones-predominant-fault.html}}. Gurdjieff often referred to finding one's ``Chief Feature", by which he means the same thing. This is how the essay begins:

\begin{quotex}
Each of us has a particular temperament which encompasses our whole manner of feeling, judging, sympathizing, willing, and acting. This temperament is to be perfected in each one of us by the practice of the Christian virtues. What can impede this work of perfection, and even bring each of us to our eternal ruin, is what is known as ``the Predominant Fault". 

\end{quotex}
It is easier to see someone else's predominant fault than one's own. Nevertheless, the importance of recognizing and overcoming it cannot be over emphasized. This is the result:

\begin{quotex}
Before conquering the predominant fault, our virtues are more like good inclinations than true and solid virtues that have taken root in us. When, with God's help, we have overcome it, the virtues become firm and strong in the nourishing rays of Charity. Charity, the love of God and of souls, comes to reign in our souls through our predominant virtue. It transforms our temperament, making us more truly ourselves: ourselves without our defects, ourselves in Charity, in God. Peace will enter the soul, and the interior joy that it brings with it, because peace is the tranquility of order which we have re-established in our souls by our mortification, that is, by our struggle against our own evil. 

\end{quotex}
\paragraph{Communist Jesuits}
The Jesuits finally revealed their true motivations with the publication of two articles promoting communism in their periodical, \emph{America}\footnote{\url{https://www.americamagazine.org/faith/2019/07/23/catholic-case-communism}\newline\url{https://www.americamagazine.org/politics-society/2019/07/23/why-we-published-essay-sympathetic-communism}}. To wit:

This reduces a supernatural religion to the natural level, i.e., the belief that the solution can be found within the world process. Progress can be made only by individual efforts, by overcoming ones on faults and by becoming conscious of the forces opposing that. The communist way states that individuals can be made good by altering social structures and the material conditions of life.

Although the communist solution is abhorrent, there can be no objection to a just arrangement of things and to the improvement of economic circumstances. Nevertheless, the reduction of religion to a mere humanitarianism is a waste of time. The honest thing to do is to discard religion as do the real communists. But that would take away the privileges that the Jesuits currently enjoy.

\paragraph{Orthodox Purgatory}
\begin{quotex}
The life in Christ originates in this life and arises from it. \emph{However, it is perfected in the life to come} … It cannot attain perfection in men's souls in this life, nor even in that which is to come, without already having begun here. \flright{\textsc{Nicholas Cavasilas}, \emph{The Life in Christ}}

\end{quotex}
A theolougemon is a theological opinion. Fruitless debates arise over opposing opinions; the only solution is to penetrate to the inner reality that the opinion tries to explain. The reality is that there are three stages of the interior life: \emph{purgation}, \emph{illumination}, and \emph{union}. One must embark on the path through these three stages in this life. Any unfinished work must be completed in the next, as Cavasilas pointed out. In Man's Spiritual Evolution, Constantine Cavarnos points out that there will be continual progress in the afterlife, and points to the teachings of the Fathers to support that view.

But isn't Cavasilas illustrating a process of purgation in the afterlife? When you bypass the opinions about the exact nature of postmortem purgation, the idea of Purgatory, as a state of being, is the same in both the East and the West.

\paragraph{Random Thoughts about the World Process}
Bill Clinton, back in the 90s, signed a crime bill that greatly reduced the crime rate in the USA. Twenty five years later, Trump signed another bill that overturned many of the features of the original bill, including the early release of some felons from prison.

I watched part of a Trump rally in Ohio last week. He boasted about the crime bill on the basis that it helped the minority community. His nearly all white supporters then clapped and squealed like trained seals. You could see them looking around the arena, and even the TV cameras panned the crowd. However, there seemed to be no black felons in the audience to show their appreciation to Trump.

Trump pointed out that Baltimore is infested with rats, has a high crime rate, poor schools, and a high rate of poverty. The debate took two forms: one was that Trump is a racist for pointing it out, the other was how to ``solve" the problem. I propose an alternative: Baltimore residents should seek asylum in some other country, like Canada or Germany, in order to seek a better life.

It is claimed that Sweden has become the rape capital of Europe. Now, I don't know if all that is true. But there are some social experiments that can be tried. For one, send a gang to Swedish youths to Saudi Arabia to rape the local women, and then report what happens. Or English rowdies can go to Pakistan to groom school girls for a life of prostitution. Text me the outcome.

Some commentators on Fox News often complain about how the tech companies are deplatforming those whose opinions offend them. However, we don't see them, e.g., Alex Jones, E Michael Jones, inter alia, ever appearing on Fox News. In other words, they deplatform by omission rather than explicitly.

\paragraph{Eccentricity}
\begin{quotex}
In Mill's open society, \textbf{the individual with his absolute right of expression is then instructed that eccentricity is a positive good and that there is a duty to pursue it. That is, there emerges a public orthodoxy of eccentricity and this will drive }\textbf{individuals to the making of exorbitant and impossible demands upon society.} This in turn will lead to confrontation and the disintegration of society, for there is no center that can hold; more importantly, there is no obligation or duty for anybody to have since all things political are conceived wholly in terms of individual rights and demands. Into the vacuum created by disintegration will move force and coercion—in a word, tyranny. \flright{\textsc{John East}, \emph{The American Conservative Movement}}

\end{quotex}


\flrightit{Posted on 2019-08-04 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Clive on 2019-08-05 at 16:35 said: }

``[sex] is attractive because it is fulfilling, it is fun, and you meet interesting people."

….what….

I don't think we should go easy on sexual sins as though they're not serious. Dante may have put them in only the first ring of hell…..but they're still in hell.


\hfill

\texttt{Cologero on 2019-08-05 at 16:39 said: }

Good point, Clive. I was just giving the ``worldly" attitude, but I guess I did not make it clear enough.


\hfill

\texttt{Ernest on 2019-08-07 at 03:34 said: }

One wonders whether Cologero, Teresa and Losang are all just multiple personalities playing about in the very senile, yet interesting mind of this site's author. I guess we'll have to wait for 1001 to find out!


\hfill

\texttt{Logres on 2019-08-08 at 20:26 said: }

Magic is playing an increasing role in politics:

\url{https://thehill.com/homenews/campaign/455496-marianne-williamson-supporters-organize-occult-task-force-on-her-behalf}

And there were the Trump workings, which backfired spectacularly during the Supreme Court justice hearings. So misunderstanding and incompetence seem to be the order of the day. It's almost as if people are wanting to use occult Enlightenment as if it was a scientific technology, which includes as a weapon of war. 

``The professor did end on a sound note when she claimed that the sin of pride was worse. Of course, pride is claiming to know better than 2000 years of Tradition."

Zounds. Great point. Being logical enough to spot irony is an essential precondition for locating your ``chief feature". Failing that, to follow sound teachers who point stuff out, but few like to follow any more.


\end{sffamily}\end{footnotesize}
