\section{Love and Work}

\begin{quotex}
Love and work are the cornerstones of our humanness. \flright{\textsc{Sigmund Freud}}

\end{quotex}
\paragraph{Work}
The True Self is always subject and never an object in this world. Thus, platitudes like ``contacting your higher self" make no sense; rather one must ``become" one's higher self. The corollary is that the Self is not static. It must be created in our acts. What we do simultaneously expresses the Self, reveals the Self, and creates the Self. As in everything that exceeds book knowledge, everything that requires know-how rather than knowing-about, the process of self-creation requires trial and error, a lifelong task to manifest all its potentialities.

A constrained life cannot do that. Forty years spent in a cubicle, or teaching kindergarten, or as a government functionary limits one's experiences, thereby leaving many of life's possibilities unfulfilled. Your life choices will be unique to you.

I don't believe in the anonymity espoused by some esoterists, including Guenon, Evola, and Tomberg, as though one's metaphysical views arose spontaneously, like babies in a cabbage patch. St Augustine and Rousseau are better guides with their revealing confessions. A man's mundane loves and work form his foundation; his connection to the spiritual world builds on it. Without that foundation, the symbols cannot be understood. Life is a struggle, the end of struggle is death, not peace.Analogously, life requires spiritual combat; there is no end to it. Only snake oil salesmen will promise instant enlightenment.

Third rate thinkers tend to dismiss superior thinkers with a single line or two, something that would fit on a placard. Instead, one must learn to wrestle with ideas, come to terms with them, and be willing to reject them if necessary. In my case, I had done a close reading of Benedict Spinoza in my youth. I decided then on a life course inspired by that reading: be independent, become proficient in a technical field (he was a lens maker), maintain a liberal (in the word's original sense) cast a mind, and strive to be God-intoxicated (as he was called).

Hence, I became a software consultant, fulfilling the first two conditions. I did not rely on a corporate pay check, but rather on my own abilities. That gave me a wide range of experience, so I learned how many systems work: telecommunications, medical equipment, radio, computer networks, manufacturing systems, security systems, financial systems, inter alia. I've had to travel, often to undesirable locations: e.g., remote locations in the South, driving 100 miles from airports in the dark, without GPS. Those places are called ``fly-over country"; without discounting the people who choose to live in such places, I understand the disdain of the ``elite". I've also been sent to more glamourous locales: LA, Santa Barbara, San Francisco, New York, Munich, and so on. In spite of that, I never absorbed the world view held by the pseudo-sophisticates in such places.

The consultant is not permitted to fail; he is usually called in as the last resort. Few people like to live under such pressure, but you do find out who you are.

\paragraph{Love}
Much spiritual literature uses Romance or Love to express the desire for union with God. The obvious examples are Dante, Hafiz, Rumi, even Solovyov. The \emph{Romance of the Rose}, an esoteric work translated into Italian by Dante himself, is an extended metaphor based on the quest for romantic love. Obviously, the symbolism of that text cannot be understood without the prior understanding of romantic love in this world. The quest is worthwhile despite the risks.

I've been playing poker monthly for several years with the same group of guys. In all that time, I barely know anything at all about their personal lives. With women, the situation is quite the opposite: it is all about personal revelation. Having had many loves is a veritable school of life. I grew up in a stable and loving family, but that is probably not the norm. From women, I've learned in great detail what goes on behind closed doors in families, things that used to shock me: abuse, neglect, fights, deceit, bad decisions. Sometimes there is malice, but often there is just poor judgment. There are too many examples: an arranged marriage by fundamentalist Christians, the CIA agent who beat his wife, the mother-in-law stomping on her daughter-in-law. 

Of course, there are infidelities. I was often surprised at how willing women are willing to cheat on their boyfriends and husbands, even to my supposed benefit.

Women like to lie in relationships, even when we both know it is a lie; it is part of the charm. Yet, women also know a secret, which is more difficult to extract. I'm still trying to decode it.

A dozen years ago, I had decided to create three blogs: one for musing (this one) and one for politics. Since contemporary politics is so absurd, I neglected the political blog. Moreover, there are many sites who do the job more or less competently.

The third blog was for the so-called Manosphere, a movement that has fizzled out in the meantime, with web sites disappearing and others having second thoughts, even religious conversions. That blog would have offered suggestions, for example, in seduction. I always had reservations about revealing so much and some guilt about leading others into sin, but it was only this year that I finally surrendered the domain name I had planned to use.

In the meantime, I've had to confess all the sins associated with my liaisons, for which I will suffer much in Purgatory. Moreover, I will have to take on the women's punishment for enticing them into their sins of fornication, adultery, and unnatural acts. So this is my one and only piece of manosphere advice: don't do it.

I often have dreams about the perfect woman; she appears to me and I experience great love and joy. The dream seems so real, so that waking up feels like entering a fake world. Then I rack my mind, reviewing my life to try to remember when and where I had seen her before. Actually, it is an encounter with my Anima, a foretaste of the Alchemical Wedding, the union of the Holy Spirit with the Virgin Soul, from which the True Self, as Christ, is born.



\flrightit{Posted on 2019-06-09 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Melthorg on 2019-06-14 at 18:02 said: }

Great article as always, Cologero. Thanks.

I have one question on you. Do you think it is worth to try to find that perfect woman in the real world? As you said a woman who represents your Anima, or female counterpart, and union of both of you can lead to higher mystical experience. Or it is better to experience my Anima just in my mind, as Solovyov do with Sophia?

I ask because I am a young man and dealing with a woman on a practical level(meeting sexual and sentimental needs) isn't motivating me to do it actually. It just can't change my existential state and it's no more than meeting any other physiological need.

But to find really valuable woman seem almost unreal. Most women, if they are not sluts, are perfectly average, they care only for everyday petty things, nothing really deep. I don't know if I should just seek better, or my expectations are too high.


\hfill

\texttt{Cologero on 2019-06-15 at 22:12 said: }

Melthorg, you are asking me to make choices for you that only you can, and should, make. Moreover, I don't know anything about you or your situation. Hence, specific advice is out of the question, nevertheless, there are some general principles.

First of all, you need to learn to live without expectations, i.e., to be indifferent to results. You live and act, the best you can, come what may.

To find the perfect woman, you yourself would have to be perfect. The Cabbala explains that all pre-existing souls have a male and female part, that separate at birth. Only if the human being is pure and his deeds are pleasing to God, will the union fall to his lot which he possessed before birth. In the opposite case, he receives for a spouse a soul that does not correspond to him.

On a more pessimist note, Adam found the perfect woman for him, yet things did not work out the way he had hoped. You can look it up if you don't believe me.

But on a more practical note, don't make any woman the centre of your life. Find an interest … a destiny … and pursue it with all your heart and energy; that should be your central focus. Good things will follow.


\end{sffamily}\end{footnotesize}
