\section{The Hedgehog and the Fox}

\begin{quotex}
A fox knows many things, but a hedgehog knows one big thing. \flright{\textsc{Archilochus}}

\end{quotex}
People can be like foxes or hedgehogs. The fox has a wide-ranging mind that is able to synthesize many ideas. The hedgehog is more plodding, focusing on one major idea. The world needs both types, even if most people are neither.

\paragraph{Tacit and explicit knowledge}
\textbf{Michael Polanyi}, in his book \emph{Personal Knowledge}, made clear the distinction between tacit and explicit knowledge. Explicit knowledge can be communicated through the written or spoken word.

Tacit knowledge, on the other hand, cannot be communicated in that way. It is more like ``knowing how" rather than in knowing. Tacit knowledge comes only by doing. Polanyi provides examples like riding a bicycle, learning to swim, becoming a wine taster, and medical diagnosis. Just reading a book or attending a lecture on those topics will not make you proficient in any of those activities. With rare exceptions, it is necessary to learn under the tutelage of someone else with more experience. Nevertheless, many people prefer to bloviate, as a simulacrum of tactic knowledge, without actually knowing how.

This distinction is even truer in esoteric matters. Reading a text has its place, but it is no substitute for the know-how of reaching higher states. The latter requires trial and error, with feedback from someone who has.

\paragraph{Faith Journey}
In this Lenten season, you may hear people discussing their effeminate sounding ``faith journey". That sounds like some sort of horizontal motion, whereas the Tradition always spoke of an Ascent to God. Most famous perhaps is \textbf{Dante}'s \emph{Divine Comedy }which describes the soul's ascent to Heaven.

\textbf{St Bernard}, who was a direct spiritual influence on both Dante and the Templars, made that the centerpiece of his \emph{De Consideratione}. In the first book, Bernard criticizes the Curia for its worldliness and carnal spirit, even criticizing the Church as a den of thieves. Clearly, not much has changed in the past thousand years. The Church today, as an institution, seems more concerned about air conditioners and immigration than spiritual enlightenment. However, that has no effect on the Tradition itself, which is solid.

In the fifth and last book, Bernard turns his attention to God and the Angels. In contemplating them, the soul transcends the realm of becoming to find her true home. He said that the higher things can be known in three ways:

\begin{itemize}
\item By opinion 
\item By faith, which knows truths in a veiled way 
\item By understanding (or intuition or gnosis) which knows the same truths openly 
\end{itemize}
Thus, we see that salvation comes not from ``faith alone", but from the higher intellect or gnosis. Faith sees God through a glass darkly, while gnosis sees Him face to face. \textbf{Giovanni Gentile} points out that these three stages correspond to Virgil, Beatrice, and Bernard in Dante's own spiritual journey.

Bernard's teaching is not unusual, as I can provide many examples. But that just adds to the store of explicit knowledge. Tacit knowledge requires that one actually embark on the contemplative path described by Bernard.

\paragraph{The Crucifixion}
In the early years of TV, there was a show called The Millionaire. The premise was that an anonymous donor would donate one million dollars to a different person each week. The effects of that windfall on the various recipients created the drama of the show.

The Crucifixion pays off the debt mankind incurred from the Fall, restoring the human race to the status quo ante before the Fall. What have we done with the windfall?

Suppose Frank went into debt from gambling losses. Most of his pay check services that debt. Suppose that someone then pays off Frank's debt. Is Frank grateful? Check with a bankruptcy attorney and you'll learn that most people just go into debt again.

In the symbolism of the cross, the vertical bar has a twofold effect. First Christ descends into Hell, purifying all the lower elements. This initiates the process of regeneration that restores the human race to its purified state.

In the opposite direction, Christ opens up the way to ascent to God to everyone, not just a few initiates.

\paragraph{Chastisement}
Divine chastisement is the punishment of the human race. As such, it is the opposite of salvation. This may sound contradictory or even harsh from the viewpoint of explicit knowledge. Tacit knowledge, on the other hand, comes from actively contemplating the notion of chastisement.

The Hermetist \textbf{Eliphas Levi} explains it this way:

\begin{quotex}
Man is himself the creator of his heaven and hell, and there are no demons except our own follies. Minds chastised by truth are corrected by that chastisement, and dream no more of disturbing the world. 

\end{quotex}
In other words, truth itself is the chastisement. If one rejects the Logos or Cosmic Order, he will experience chastisement from that folly. Of course, God is also the True, the Good, and the Beautiful, so in a sense, it is God who chastises through his willing the cosmic order.

\textbf{Valentin Tomberg} makes this claim:

\begin{quotex}
Is mankind therefore solely responsible for its history? Without a doubt — because it is not God who has willed it to be as such. God is crucified in it. 

\end{quotex}
The world of everyday experience is anthropogenic, not God created. The crucifixion is the result of man's rejection of God's appearance and influence in the world. This is the result:

\begin{quotex}
Man is from now on alone, hopelessly alone, left to himself, at the mercy of other men! Thus, it is vain to expect a further chastisement, as the absence of God is the true and supreme punishment of our crime of treason. God's Justice is exercised with just reason. As man in his mad arrogance has shown to God his will to do without Him, man's terrible punishment consists in being granted what he has asked for. Without God, man rapidly becomes the slave of his own pride and passions, a beast that lives only to satisfy his instincts. \flright{\textsc{Chastisement and Punishments\footnote{Dead link.}}}

\end{quotex}
\paragraph{Post-mortem States}
\emph{Pace} this bizarre rant Debunking the myth of hell, hell is not an arbitrary or cruel punishment inflicted on the unwary. A person who has no interest in God, or the things of God, in this life, will have no interest in God after death. We have all we need to be saved; take it or leave it as you choose.

Hell is the absence of God, and man is the creator of his own Hell. It is not a punishment, but a matter of justice. As Dante points out, the souls willingly take their place in Hell. They blame each other for their predicament, but not God.

\paragraph{The Religion of Europe}
We wrote previously about the Religion of Europe as envisioned by the New Right. While claiming the desire to ``save" Europe, they reject the notion of Logos, cosmic order, and the true religion of Europe. Moreover, not to speak ill of the dead, the West cannot be restored by rejecting its real past and promoting a mythical unreal future.

By rejecting the religion of Europe, they necessarily reject the Tradition of saints, doctors, philosophers, poets, artists, knights, and so on, who created the West. They embody the eternal truth of ``Catholicism", not the Curia whose corruption today is no different from the Curia of a millennium ago. Whatever is of value in other Traditions can be found at home. Read the essay Acres of Diamonds.

\paragraph{Cultural Marxism}
You can love him or hate him. but Marx was right in many of his predictions. There can be no dispute that the technological and material changes in society have altered social behavior and mores. For example, better hygiene and cleanliness have made oral and anal sex more acceptable than what it was when bathing was infrequent. The wide availability of mind-altering drugs has had its consequences. And the mass media has made propaganda easier to promulgate to vast audiences.

The term ``cultural Marxism", although popular, is certainly redundant. The Communist Manifesto asserts:

\begin{quotex}
The Communists everywhere support every revolutionary movement against the existing social and political order of things. 

\end{quotex}
Hence, it is not merely a question of a new economic system, but the overthrowing of every hierarchical order. Don't we see that today when every alleged hierarchy is under attack? If the question is why that is, Marx had anticipated the answer:

\begin{quotex}
A spectre is haunting Europe—the spectre of Communism. 

\end{quotex}
He could see the evil spirit of communism and its long-term effects. Even those who deny being communists nevertheless work to implements its programs. Lenin was incorrect in his belief that a violent revolution is necessary for the communists to take power. Remarkably, we see the advanced liberal countries voluntarily accepting the communist tenets. A voluntarily chosen democratic socialism is hardly superior to socialism imposed by force.

\paragraph{Survival of the Species}
Evolutionary psychology seems to have become the latest craze, even if it appears at times to be little more than a collection of just-so stories. One of the most recent goes by the name of the Jolly Heretic, although he has co-authored a book with Bruce Charlton. His hedgehog idea is that religious belief has a survival value for evolution. Obviously, then, the cultured despisers of religion are at an evolutionary dead end.

The Jolly Heretic asserts that behaviors that are evolutionary imperatives are then attributed to the Will of God, giving them a moral force. Here, in brief, are some of his main points:

\begin{itemize}
\item Believers have, on the whole, better health than the irreligious. 
\item Religion teaches that life has meaning, which has better survival value than the beliefs of secularists that life is random with no absolute meaning. 
\item Religious people have an in-group preference, unlike secularists. For example, marriages with someone of a different religion are discouraged. 
\item Reproduction and family are encouraged, unlike secularists who promote abortion, contraception, and infertile sexual activity. 
\item The more religious a society is, the more ethnocentric it is. 
\item One of his more startling claims involves the role of the Executioner. He claims that up to 2\% of the population was executed for various crimes in the Middle Ages. This had the salubrious effect of eliminating psychopaths, sociopaths, and other reprehensible types from the gene pool. 
\end{itemize}
The atheists and agnostics of the alt-right, with their love of genetics and evolutionary psychology, should take heed.



\flrightit{Posted on 2019-03-07 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Han Fei on 2019-03-10 at 00:09 said: }

The final part in particular piques my interest. To say that our ethical principles are tied quite literally, to the matter of our biological survival, is probably true, but not in the way that author intended it. God is hard working but not so hard working as to not let nature run its course once in a while. In fact I would go so far as to say the universe is designed in such a way as to not tolerate certain possibilities from coming about and persisting in a stable state. With that in mind, much of the moral compass of modern society depicts a road to ruin equally from the standpoint of judgment-detached science as that of a religious morality. I wouldn't consider it at all anti-traditional to contend that the two aren't necessarily opposed to each other. Perhaps we did not know how this could be possible 200 years ago, but now we do.

The problem of course is that having come to accept (or argue from the standpoint) that nothing has meaning and substance beyond a hard code ingrained by nature in DNA, it would be very difficult to ensorcell ourselves to believe in any solution that directly contradicts this assumption. No matter how much he's convinced in the veracity of this causal relationship, he can't be one to partake in it.


\hfill

I've made this available: \textit{Essays on Catholicism, Liberalism, and Socialism}\footnote{\url{http://www.gornahoor.net/library/CortesEssays.pdf}}

Years ago, I wrote a series on economic systems for a forum that is not longer online. I'll try to dig it up if I still have it.

Here is an article I translated: Economics and Spirituality. I have a few other translations on that web site.

There is a recent book, which I haven't read, that may be relevant to your concerns: Why Liberalism Failed, by Patrick Deneen.

Then there is the obvious option. If no one is doing a necessary task, they you could take it on yourself.

Many people have looked foolish by failed predictions of the end of the world. Science has predicted the end of the world in the past. E.g., Malthus, DDT, the population explosion (I can't remember the guy's name, but he was influential in the 70s), global cooling, the ozone layer, the book On the Beach, not to forget the heat death of the universe.

If you truly believe the world will end relatively soon, what exactly do you think I can do to prevent it? Or perhaps I am not even opposed to the end of the world. The end of one world is the beginning of another.


\hfill

\texttt{James on 2019-03-20 at 04:03 said: }

I'm only going to assume that Arthur Konrad has a particular ``message" he wanted to present via his reply rather than to authentically claim that Dante's poetry was motivated and inspired by personal rather than superpersonal considerations. Surely, his comment was simply weaponizing a single line and not an actual attempt at contradicting the obvious and intuitive fact that Dante's Hell is populated by those who willingly wish to be there. Surely this is some kind of strange and purposeful reading of his.

Still, I can't help but be nauseated by the idea. Assuming for a moment that Arthur Konrad as a person doesn't exist and merely the idea of ``Dante places [the damned] in Hell according to his own personal grudges" is presented by some bystander: this is actually useful to examine. What kind of reaction is this? Is it not a purely superficial reaction? It is as if one were to read the Bible literally and only literally. It is the lowest and most primitive of readings.

Again, this has nothing to do with Arthur Konrad, but to simply examine what this comment might mean if we take it at face value.

The beauty of Dante's poetry is that it speaks of the universal in light of the personal. It is the very act of alchemically synthesizing the PERSONAL with the SUPERPERSONAL. It is an imitation of the Incarnation for it is in the Incarnation where we have the UNIVERSAL assuming a PERSONAL character. It is in the Incarnation that the infinite decided to be one gender, one man, one race, one place in time. He baptized the personal.

But look at the comment. It assumes that Dante makes arbitrary and personal choices. What this demonstrates is that the speaker of this comment *would not have recognized Christ*. What I mean by that is that the LOGOS became a SUBJECTIVE person and made SUBJECTIVE decisions and teachings. His parables, His examples, His miracles, His healings were all meant to be understood not *just* `his' own decisions but HIS decisions as well. They had universal import and relevance. The curing of the blind was not just his pity on the blind man, but His saving power on those who seek to see.

To be unable to recognize the universal and ideal in light of the imminent and personal is to fail the ``test" of the Incarnation.

It is to fail the ``test" of the Eucharist.

``Surely that's just bread." ``Is this not the son of the carpenter?"

Dante speaks of lofty Florentines that we have never heard of nor perhaps have never been interested in. And yet to dismiss that as ``personal" is to miss the very mode of God.

To be unable to understand this is akin to listening to old stories about a small race millennia ago escaping the much mightier Egyptians and wandering through the desert and saying, ``I should not be interested because this is just a personal account of Jews" and ``there is no value to be learned from the Exodus because it is purely the subjective grudge of the Jewish race."

The simpleton thus doesn't understand that it is also the story of the soul coming out of the slavery of sin. Again, I am not calling anyone a simpleton in any demeaning way. There are indeed people who possess the intellectual ``dharma" of a simple person and that is totally fine.

For those who are not intellectual simpletons, there is a hidden and majestic universality to the intensely personal accounts of Dante's closest associates and rivals. Dante expresses fidelity to God's mode in that THE REALITY AROUND HIM WAS PLACED FOR HIS DISCERNMENT. The men and women placed in his life were not ``random" but represented an EDUCATION given to him by God.

Thus, to extol and valorize–to purify and elevate the intensely subjective until it painfully reveals the universal is the intellectual participation IN THE BLESSED SACRAMENT. For the secret of the universe is that God has made this reality as a sacrament: that with the eyes of faith, we can see the miracle of God consecrating the world and accepting upon the tongue and tasting it imminently if we are properly disposed. Dante's proximal relationships and his milieu were placed there by a guiding Hand and not a matter of chance. To be faithful to this ``school of community" demonstrates whether we truly worship the God of order or if we secretly admit to being concubines of Chaos and Chance.


\end{sffamily}\end{footnotesize}
