\section{Waiting for a Miracle to Come}

You can hope for a catastrophe or wait for a miracle\footnote{\url{https://www.youtube.com/watch?v=lL2DzqrUNQc}}, but you cannot expect normalcy.

I am becoming more like Don Quixote. I live in a world that does not really exist and I fight imaginary battles. I hope I win one someday.

Wittgenstein confessed, ``It's damned hard to write things that make blank sheets better!" … he got that right, but I don't see it discouraging anyone.

Adam named the animals, which implies he understood the essences of things. Lesser men today engage in magical thinking. They change the names of things, believing, thereby, that they are changing essences.

\begin{quotex}
Everyone sees what you appear to be, few really know what you are, and those few dare not oppose themselves to the opinion of the many, who have the majesty of the state to defend them … the vulgar always judge by appearances and by results, and in the world there are only the vulgar \flright{\textsc{Machiavelli}}

\end{quotex}
I've been watching physics videos to see if the world makes sense. It doesn't. In \textit{Is faster-than-light travel possible?}\footnote{\url{https://www.youtube.com/watch?v=-2c0P2CEU9A}}, Sabine tells of a fellow who went back in time, accidentally killed his grandfather, and immediately ceased to exist. Also, a particle can't tell if it is moving backwards or forwards in time. I think I could tell.

We've learned that we can live just fine without professional sports, but not without the workers at power plants. Which of these would you choose: shut down all the ``news" channels or the meat packing plants?

We poker players have an adage: ``If you can't recognize who the chump is at the table, it is probably you." Likewise, you don't want to be the one with a knife at a gun fight.

The Argyraspides\footnote{\url{https://en.wikipedia.org/wiki/Argyraspides}} were a feared and revered division of Alexander's army, consisting of veterans past the age of 60 years. They won many a battle. They could save the US army beaucoup bucks since they pay themselves out of their plunder, and they may actually win a war or two.

\paragraph{The Inverse of Truth}
The effects of Adam's fall are from a higher state to a state of ignorance, malice, and concupiscence. Tradition teaches that knowledge of the good ameliorates malice and concupiscence. The collected wisdom of the group provided that knowledge to the ignorant.

The current belief is that man ``evolved" upwards to ignorance, malice, and concupiscence. This is the solution of the modern world:

\begin{itemize}
\item Universal education as the solution to ignorance. It has made things worse, because the ignorant believe they are knowledgeable. 
\item Social reforms as the antidote to malice. Result is permanent revolution, since no reform is ever enough. 
\item Promiscuity as solution to concupiscence. The expression of each and every sexual desire has not made anyone happier. 
\end{itemize}
\paragraph{Animal Intelligence}
Joe Rogan\footnote{\url{https://www.youtube.com/watch?v=TlzY_KvGSZ4}} was impressed by the intelligence of an octopus who climbed out of his tank, crossed the floor, ate some fish in another tank, and returned to his own. Conclusion: If you can go to the fridge, make a sandwich, take it back to the couch and turn on the TV, then you have reached the intelligence level of an octopus.

Scientists have been trying to decode animal talk. It's bad enough listening to people, but do you really want to listen to your dog complaining all day long? Unfortunately, animal can't make vowel sounds\footnote{\url{http://theconversation.com/examining-how-primates-make-vowel-sounds-pushes-timeline-for-speech-evolution-back-by-27-million-years-128514}}, so they can't speak. Danish has too many vowels, and it has not helped their speech.

\paragraph{The Allure of Women}
Axel Heyst in \textbf{Joseph Conrad}'s novel \emph{Victory} was a Christ-like figure. Axel lived a solitary idyllic life on an Indonesian island. On a trip for supplies, he meets the most beautiful woman in a touring orchestra. He reluctantly brings Alma (his soul, get it?) back to his Island, where they fall in love. Heyst's worldly detachment is challenged by his desire for Alma and he succumbs to the world of action. Then she is tragically murdered by a gang of thieves with ``mud souls".

\begin{quotex}
I do not understand why they keep talking of love when they are around me, when they look into my eyes and kiss my hand. … But when the red light is on, at the midnight's hour, and everyone hears my song, the reason becomes clear to me: My lips they give so fiery a kiss, my limbs are supple and white. It is as if it were written in the stars: You shall kiss! You shall love! My feet glide and float, my eyes lure and glow, and I dance as if in a trance because I know: My lips give so fiery a kiss.

\end{quotex}
My son ties me down to my chair before I watch Patricia Janečková sing My Lips\footnote{\url{https://www.youtube.com/watch?v=pzF3ubtimGE}}.

\paragraph{The Media Now and Then}
Rachel Maddow, a popular TV ``news" analyst, won a defamation lawsuit on the grounds that her statements were so hyperbolic that any rational listener should not take her statements as literal facts … yet so many do. You would think that such a decision would discredit her, but apparently that is considered normal. Rachel was just kidding, get over it\footnote{\url{https://www.nbcsandiego.com/news/judge-dismisses-sd-based-one-americas-defamation-suit-against-rachel-maddow-msnbc/2331027/}}.

In the classic movie Stagecoach (1939), starring John Wayne, the banker sounded like an episode from Fox News: ``What's good for banks is good for America", he said, and then complained about the National debt reaching \$1 billion, and finally ``we should have a businessman as president." 75 years later his every desire has been fulfilled. Watch it before it gets canceled.

The John Wayne character proposed marriage to a woman he just met on the stagecoach the day before. (i.e. life before Tinder) Love at first sight was common in old movies. That actually is rather easy, but love 6 months later is a much different movie.

\paragraph{Education Advice}
Teach your children — or yourself to play music\footnote{\url{https://getpocket.com/explore/item/the-benefits-of-playing-music-help-your-brain-more-than-any-other-activity}} — it is one of the seven liberal arts. Playing Music helps your Brain. It will help your brain as much as grammar, logic, math and cosmology.

When I started Junior High School (NKA middle school), they divided the students into tiers: e.g., 7A1 to 7A4, 7B1 to 7B4, then the C levels. It wasn't meant to hurt anyone, just to help them find their best job. Presumably, there was a college track, a business track, and a trades track. At that time, you could make a living from any one them. They must have had that system for decades. Then egalitarianism crept in, so they renamed the groups 8L6, 8O3, etc., until someone figured out that LOQUACIOUS represented the correct order. Now, I suppose, they don't even do that anymore, so few students know what their best skills are.

\paragraph{Plausible if not True}
The US government was supposed to release their UFO files this month. A source says it is more like invasion of the body snatchers.

In the movie version of Marat/Sade\footnote{\url{https://en.wikipedia.org/wiki/Marat/Sade_(film)}}, the Marquis de Sade directs a play starring the inmates of an insane asylum in which he

\begin{quotex}
subtly manipulates both the players and the audience to create an atmosphere of chaos and nihilism that ultimately brings on an orgy of destruction.

\end{quotex}
Now it repeats as farce. Nevertheless, it would be great to be one of the inmates; then you could take solace in being on the winning side of history.

\paragraph{The Divine Comedy}
Rene Guenon can sometimes be so tedious as, for example, in his treatise on Dante. He sees symbols and correspondences everywhere, while ignoring completely Dante's own explanation. Guenon writes nothing about contemplation, self-knowledge, etc. Purgatory is, for him, the ``trials" of achieving the Masonic grades. From what I see, there are no trials, just expenses. Richard of Saint Victor repeats that purity of heart and self-knowledge are the conditions for mystical elevation. The soul is a mirror that reflects the invisible things of God, and a purified soul reflects best. Theosis is the overcoming of the human state of being. It does no harm to keep repeating this.

\url{https://www.youtube.com/watch?v=HBdq4mTk6Tk}



\flrightit{Posted on 2020-08-16 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Mike on 2020-08-17 at 06:13 said: }

This whole country is about to burn to ash, past November. NYC's ``expiration date" is a foreshadow of the country's own ``expiration date". If Trump wins, the whole country will be set ablaze by useful idiots such as antifa, extremist-liberals and blm. As a response to this the national guard will be deployed, martial law will be enforced and a civil war, maybe as a precursor to WWIII will break out. If Biden wins, together with his crackpot sister, Harris, then say hello to communism, insane taxes, death of privacy, increased degeneracy, no more guns or other weapons to protect oneself, unbridled immigration and just generally bowing down to whatever whims the globalists have. Not to mention the violent responses from right-wing extremists that would also erupt on account of an ultra-leftist, ultra-progressive government.

Things might turn out differently, but this is not too far afield from the realm of possibility. I can only wonder how the plans at gornahoor and its affiliates will be affected by this. 2021 will be a very interesting year…


\hfill

\texttt{Cologero on 2020-08-17 at 07:20 said: }

Don't the Chinese consider this a curse: ``May you live in interesting times."

2021? You're more optimistic about Gornahoor than my doctors are.

My plan is like Dante's plan for the Divine Comedy:

\begin{quotex}
The purpose of the whole and the part could be multiple, that is both remote and proximate. But leaving off subtle investigation, we can say briefly that \emph{the purpose of the whole as well as the part is to remove those living in this life from the state of misery and to lead them to the state of bliss}. 

\end{quotex}

\hfill

\texttt{burningfireeye on 2020-08-18 at 19:28 said: }

UFOS are spiritual realities, I know a friend who was with another friend of his, they both saw a great lightning at the same time while inside of the car, both of them were perfectly sober. He told me he felt a great energy inside of him. It is the Astral Light spoken of so much by Eliphas Levi.


\hfill

\texttt{jalexandermaximilian on 2020-08-18 at 22:36 said: }

Burningeyes.

UFOs are most certainly a Reality. I see them almost everyday, generally in the dawn and twilight. I had a similar experience in which I was meditating, with my wife (we are both pupils of Serrano and that is what brought us together) in the cemetery by my house, it was nighttime. I was visualizing a stream of energy coursing from a sphere above my head, when the energy actually manifested and I felt a great electrical current, a strange vapor, which tasted of ozone was rising out of my stomach and out of my mouth. My wife mentioned me reeking of ozone as well. We felt we should begin back home… I was still on fire with same energy your friend mentioned, and it was 30 degrees outside, my wufe was getting cold and I was able to put the warmth into her as she got cold. Suddenly about 100-150 feet in the air a White Disk, quite large, and alarming flashed at us 3 times, it was very close, and illuminated the entire area, and flew away. When we got back home we were elated, and just mind blown! It was now late so we went to bed, it was hard to sleep as I was charged and when I'd close my eyes I'd see a field of typical ``grey alien" faces. That was a bit disconcerting to be honest. But we now see the white Disk, and orbs as well quite regularly, not as intense as that one, but I do have photos and footage. Quite a bit. I'd swear on Christ this true, but He said not to swear such oaths so I will just say… I tell you no fib.


\hfill

\texttt{Mike on 2020-08-20 at 09:31 said: }

@Cologero

I suppose an esoterist's manner of prepping is all-inclusive and the items on his list are ordered differently. (I doubt toilet paper would be your number one priority and I doubt communing with God through effortless contemplation would be on the ordinary person's list, even in ordinary times).

@burningfireeye

I recall watching a surprisingly interesting documentary done by Vanity Fair, that documented domestic terrorism fueled by left-wing groups in the 70s and 80s, mainly in liberal areas like NYC. These were children of the wealthy and the bourgeoisie that went about making things like nail-bombs and also blowing up buildings. Eventually, their group fizzled out, but some of them made it into academia and just continued funneling whatever poison had destroyed their minds into the new generation of kids we see today. So maybe this is some sort of cyclic-resonance, since it's been 50 years or so since the 70s and it's the same cushioned children of the wealthy and bourgeoisie that are causing mayhem. Perhaps they too will fizzle out and some other bigger distraction will stave off the animal-like impulses of the masses, for a little while longer…


\hfill

\texttt{Arthur Konrad on 2020-08-20 at 19:20 said: }

I was of impression that your health was improving Cologero. I derived much value from your writing once, and your translations of Evola and his correspondences helped us all learn more about him.

Music, as opposed to what Schopenhauer said of pity, makes life worthy of praise. Romanticists devalued it somewhat with their infantile love of decibels. The world owes Italy a great deal for its conception of taste in music. The country was nearly immune to all sorts of folly and excess in this domain, and many Italian musicians remain so to this day.


\hfill


\end{sffamily}\end{footnotesize}
