\section{Audrey of the Sun}

Audrey del Sol was a pisser, as we say in Boston. She was alive in her body, and in her soul, and in her spirit. Never without an opinion, her mind wandered over whatever topic intrigued her. Born in Tajikistan (née Biliana Dimitrova), her interests hardly coincided with those typical of American women. She listed her religion as the Big 5: Christianity, Islam, Judaism, Hinduism, and Buddhism.

She told me she was ``hugely aligned'' with Evola's philosophy even before she heard of him. So she found a kindred voice in Gornahoor, often promoting posts and actively commenting on articles. She liked guns and was a 2\textsuperscript{nd} Amendment fanatic.

She claimed a Sicilian connection. A relative lived in Italy and it was her family's tradition to honeymoon in Sicily. She claimed to have worked with the Italian intelligence agency SISMI.

Her train of thought could be exasperating, probably because she stood beyond expected norms. This is how I described her in one post:

\begin{quotex}
I know Audrey, and she does have an unusual, but not unintelligent, mind. She not only thinks `out of the box', I doubt she even knows where the box is.
\end{quotex}

But, in retrospect, perhaps she did know where it is when she told me, ``I also live outside of the world of library, university and home.'' If those institutions define the box, what lies beyond? For her it was the world in which Gurdjieff, Evola, Samael Aun Weor, Franz Bardon, Valentin Tomberg make sense. Some whom I know claim to live in that world, but few actually do. It may not even be possible ... except perhaps for Audrey.

Outside of library, university and home, even outside the natural state of the lion, she sees something beyond nature, a sexuality that is neither pleasurable nor procreative. She swears an ``oath not to use her body as a vehicle of another's desires.''

You can find her Gornahoor comments with a site search\footnote{\url{https://www.google.com/search?channel=cus2&client=firefox-b-1-d&q=audrey+site\%3Agornahoor.net}}. When I noticed she had studied creative writing, I asked her to write an article for Gornahoor. She never did write the article; perhaps they could have explained her worldview coherently.

Her grasp of the whole boy-girl thing was neither feminist nor traditional, but derived from odd insights and her personal experiences. One day, she would send an anti-contraception article by Samael Aun Weor\footnote{\url{https://anael.org/es/anticonceptivos-parte-2}} and then one on Gurdjieff's notion of the Formatory Apparatus, and then relate them to the left-hand path of vamacharya. This sounds like a rather crazy party in DC.

\begin{quotex}
This is the same apparatus that gets engaged when the male starts ``barking'' from his crotch and shouting ``slick slogans'' that are supposed to act as ``banana peels'' for the ``sexually naive and gullible'' females. But I love you... . because it IS trying to wall-paper its butt crack because it DID NOT want to go on the niva fields and wipe its own butt. Not only that, but they confound this abuse of the left-hand path with the practices of vamacharya. It is not ``wide-spread''; it's ludicrous among computer workers, and yes I mean Microsoft, etc. They disrespect that the female actually has legs and they pretty much ``rush'' them into the house where a ``party'' begins, usually at her expense. As one of them said it: svara was all about ``putting the monkey on the moon''. This is called human ROT.
\end{quotex}

Toward the end, she was getting restless. She sent me this message:

\begin{quotationx}
The Earth is not getting colder, it is getting warmer, and may in time become like the sun. There you go. The majority of humanity is simply being burned up like briquettes for that event. Consciousness means that you become ``a'' Star or ``a'' Sun and live immortally. ``Food for the moon'' or ``dacha by the sea'' means exactly that, as described in \emph{Yoga of Power} by Julius Evola. THAT is what money is. THAT is why Albert Wass wrote ``I am not interested in your money, but how close you are to god''. That is the actual meaning of the ``veil'' in Islam.

Most humans do not awaken to their ``star'', Hence they are locked in a mechanism of how much and what they are getting. Hence ``cops'', ``wars'', ``stupid food attitude'', ``violence'', ``poverty'' and the multitude of problems we are plagued with.
\end{quotationx}

This is her ultimate opinion of the human condition:

\begin{quotex}
The fact is that the `fallen' condition of mankind, is one of animals. We must work and struggle to become human, i.e., conscious.
\end{quotex}

She saw herself as a conscious human being forced to live among animals --- lustful, money grubbing animals.

She scorned the fastest growing industries: retail, healthcare, hospitality. The only interesting work was being done by the government and military, she said, and sent me a video of the Italian Special Forces in action. She has extensive formal university education as a cognitive scientist, mathematician, linguist, philosopher and art historian.

Yet her work history is quite varied and does not always reflect that education. Her last civilian job was in hospitality, as a cocktail waitress. Nevertheless, I was surprised at her decision to enlist in the USA army, especially as a Specialist, not even an Officer; she was near the maximum age for enlistees. She seemed to like the idea of carrying a gun.

Prophetically, Audrey of the Sun made her last facebook post on 27 January 2013:

\begin{quotex}
Lacking nothing, contemplative, immortal, self-originated, sufficed with a quintessence: he who knows that constant, ageless, and ever-youthful Spirit, knows himself and does not fear death.
\flright{\textsc{Shankara}}
\end{quotex}

Two days later, on 29 January 2013, she was dead, struck by an SUV as she was walking along the street.

I hope she became a ``Sun''.

\url{https://gornahoor.net/decameron/wp-content/uploads/2020/12/Audrey-Del-Sol.mp4}


\flrightit{Posted on 2020-12-06 by Admin}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Juliano on 2023-12-15 at 11:42 said:}

I hope you also became a ``Sun'', Cologero, and you and Audrey are sending us help from above. Rest in peace.


\hfill

\end{sffamily}
\end{footnotesize}

