\section{A Solitary Man}

\begin{quotex}
`Informer' ... means one who gives information. It means what `journalist' ought to mean. The only difference is that the Common Informer may be paid if he tells the truth. The common journalist will be ruined if he does.
\flright{\textsc{G.K. Chesterton}, \emph{Utopia of Usurers}}
\end{quotex}


Back in the day, I aspired to be a public speaker, so I attended a conference of the Florida Speaker's Association. They told me that I had to be funny if I wanted to be paid. I figure the information is funny enough on its own, so I don't need to add a joke.

\paragraph{Isaias}


It has taken TV newscasters three days to learn how to pronounce the name of Hurricane Isaias correctly. Years ago, Hurricanes all had female names. Then the feminists objected, so traditionally male and female names alternated. People were still unhappy, since the names were mostly Anglo. So the pool of names was enlarged, and where it ends I don't know.

As for Isaias, I'm sure the forecasters think they did pretty well, predicting the storm with 75 km or so. But for us who live here, 70 km offshore (as it was) is quite different from a direct hit (as was predicted). Hence, we didn't get much bad weather. The mathematics of weather forecasting involves what are called ``chaotic systems''. It goes well for a while, but at some point, the equations break down and there is no solution possible. Most people are ignorant by disposition, design or their own choice. But even for those who know, there are inherent obstacles to complete knowledge built into the cosmos.

\paragraph{The New Decameron}


\textbf{Giovanni Boccaccio} wrote the \emph{Decameron} about 10 (upper class) people who evaded the Black Death by moving out of the city into the countryside. But they each had to come up with a story to entertain each other. I've had some friends who suggested the same thing, but none of them was clever enough to create a story. Boccaccio's are pretty good, but since this lockdown has gone on longer than anticipated, I think I'll post some of my own. One them will begin like this:

\begin{quotex}
Brianna had a lust for life, ok let's be honest, she just had lust. When I walked into the kitchen for coffee and breakfast, she was with her 15 year old daughter Jessa. Just to be clear, if I had a pretty teenage daughter, I would lock her up in the bell tower. To my surprise, the two of them were dancing around singing together, ``me so horny, me so horny'' over and over, all smiles. Just then, I realized that there is a secret wisdom passed down through the generations from mother to daughter, of which men have no inkling.
\end{quotex}


\paragraph{Speaking Truth to Power}


\begin{quotex}
We are dealing with real attack of globalists on all kind of independent media. That is the sign of their weakness they cannot stand the truth.
\flright{\textsc{Alexandr Dugin}, after being deplatformed.}
\end{quotex}


On the contrary, that is a sign of their strength and the truth does not matter. They have adapted the traditional Chinese model of statecraft:

\begin{quotex}
In classical Chinese statecraft there are two tools for gaining and maintaining control over ``the mountains and the rivers'': The first is \emph{wu} (\textbf{weapons}, \textbf{violence} -- \begin{CJK*}{UTF8}{bsmi}武\end{CJK*}) and the second is \emph{wen} (language, \textbf{culture} -- \begin{CJK*}{UTF8}{bsmi}文\end{CJK*}). \emph{Chinese leaders have always believed that power derives from controlling both the physical battlefield and the cultural domain}. You can't sustain physical power without discursive power. \emph{Wu} and \emph{wen} go hand-in-hand.
\end{quotex}


What matters is who holds power and controls discourse. This has always been the case. This may disappoint idealists who try to win an argument rather than win. Mao followed the antique tradition which is still applicable.

\begin{quotex}
Mao's discursive advantage was Marxist-Leninist ideology. \emph{Language was not just a tool of moral judgment. It was an instrument for shaping acceptable behaviour and a weapon for distinguishing enemies and friends.}
\end{quotex}


You still don't get it?

\paragraph{Poets and philosophy}


Poets have a different perspective on things. So when \textbf{Charles Peguy} asserts that \emph{Plato}, \emph{Henri Bergson}, and \emph{Rene Descartes} are the three greatest philosophers, that is not a ray of light. Rather it is an irritant like a grain of sand in an oyster. It is not obvious that a pearl will result unless you believe that Peguy offers a special insight to life.

Ditto for \textbf{Ernst Junger} placing \emph{Saint Francis}, \emph{Saint Anthony}, and \emph{Max Stirner} among the great saints.

They teach detachment from worldly concerns and relief from the gnomic will.

I'm looking into becoming an anchorite, just to make my lifestyle official. Demons attacked the solitary Saint Anthony. They can feast on my carcass.

\paragraph{New Morality Memes}


\begin{quotex}
What thus impressed me was the chapter in which Bentham passed judgment on the common modes of reasoning in morals and legislation, deduced from phrases like `law of nature', `right reason', `the moral sense', `natural rectitude', and the like and characterized them as dogmatism in disguise, imposing its sentiments on others under cover of sounding expressions which convey no reason for the sentiment, but set up the sentiment as its own reason... \emph{The feeling rushed upon me that all previous moralists were superseded and that here indeed was the commencement of a new era in thought.}
\flright{\textsc{John Stuart Mill}}
\end{quotex}


Mill discovered the ``meme'' and it was life altering. We are undergoing another new era in thought, in the same way. There has arisen a new set of dogmatic moral views, of which no past moral philosopher had ever taken notice. Argue against them if you must; let me know how it turns out.

\paragraph{Overcoming temptation}


Is there temptation if there is no desire in the first place? There is no masturbation or homosexuality among the Aka and Ngandu people of central Africa. Instead a man and his wife have sex several times a night to create children. How many American men long for a wife like that (or perhaps vice versa)?

The Bronze Age Pervert claims that animals fap only in captivity. Do you want to hunt gazelles all day or work in a cubicle for 40 years?

\paragraph{Life as Allegory}


\begin{quotex}
Every man's life is a perpetual allegory.
\flright{\textsc{John Keats}}
\end{quotex}


Life is like a symphony in which the same theme recurs with slight variations. Actually, your life has been lived before by someone else, probably in a different key. I find the keys to my life in great literature. If your life is rather stunted, you my find the keys in mediocre literature or B movies.

\begin{quotex}
The things which happen to a person, and the order in which they happen, may be as much a part of him as his physical organism.
\flright{\textsc{Owen Barfield}}
\end{quotex}


Actually there is no ``maybe''. Before you retire for the night, review your day. What happened? What was your role in it ... and for that you must be honest. How about the past year, the past 10 years, etc. The first time that your will was totally aligned with God's was the choice of the circumstances of your birth. I was probably the only time, if you continually rebel against the idea.

\paragraph{The Horror of Reincarnation}


Someone who claims to know, told me that the people you know before the age 35 were known in a past life, but the people you meet after that age will be part of your future life. My past life must have been pretty good, but I have reason to dread my next life. I sure hope she is wrong.

\paragraph{High school experience}


I was not one of the popular kids in high school for obvious reasons. I was eager to graduate and hoped things would be better. But life is still like high school. The ``popular kids'' still get all the attention no matter how bland and ordinary their opinions are. They get all the ``likes''.

\paragraph{Clear Ideas for Young Men}


\begin{quotex}
For an individual, however, there can be no question that a few clear ideas are worth more than many confused ones. A young man would hardly be persuaded to sacrifice the greater part of his thoughts to save the rest; and \emph{the muddled head is the least apt to see the necessity of such a sacrifice}. Him we can usually only commiserate, as a person with a congenital defect. Time will help him, but intellectual maturity with regard to clearness is apt to come rather late. This seems an unfortunate arrangement of Nature, inasmuch as clearness is of less use to a man settled in life, whose errors have in great measure had their effect, than it would be to one whose path lay before him. \textbf{It is terrible to see how a single unclear idea, a single formula without meaning, lurking in a young man's head, will sometimes act like an obstruction of inert matter in an artery, hindering the nutrition of the brain, and condemning its victim to pine away in the fullness of his intellectual vigor and in the midst of intellectual plenty}.
\flright{\textsc{Charles Sanders Pearce}}
\end{quotex}


\paragraph{Truth and Reality}


\begin{quotex}
And the independent reality of the individual, when we examine it, is in truth mere illusion. Apart from the community, what are separate men? It is the common mind within him which gives reality to the human being, and taken by himself, whatever else he is, he is no human ... If this is true of the social consciousness in its various forms, it is true certainly no less of that common mind which is more than social. The finite minds that in and for religion form one spiritual whole have indeed in the end no visible embodiment, and yet, except as members in an invisible community, they are nothing real. For religion, in short, if the one indwelling spirit is removed, there are no spirits left.
\flright{\textsc{Francis Herbert Bradley}, \emph{Truth and Reality}}
\end{quotex}


The Anarch, on the other hand, is real.

\paragraph{Sources}


\begin{itemize}


\item \textbf{Charles Peguy}, \emph{Notes on Bergson and Descartes}


\item Engineers of the Soul\footnote{\url{https://sinocism.com/p/engineers-of-the-soul-ideology-in}}

\item Where Masturbation and Homosexuality Do Not Exist\footnote{\url{https://www.theatlantic.com/health/archive/2012/12/where-masturbation-and-homosexuality-do-not-exist/265849/}}

\item Eumeswil by Ernst Junger\footnote{\url{https://gornahoor.net/?p=9186}}

\item \textbf{Charles Pearce}\footnote{\url{https://gornahoor.net/library/HowToMakeOurIdeasClear.pdf}, \emph{How to Make our Ideas Clear}}


\item \textbf{Johnny Cash}\footnote{\url{https://www.youtube.com/watch?v=e\_OZrUo-7m4}}, \emph{Solitary Man}
\end{itemize} 

\flrightit{Posted on 2020-08-02 by Cologero}


