\section{The Nihilist and the Traditionalist}

Over the weekend, I met up with my essence friend Tony, who drove up from Miami Beach to pick up some of his girlfriend's stuff. I met her briefly in the driveway; she dropped him off and set off to do her errands. She was quite attractive and well spoken, with a lithe trim figure. We found a shady spot in the backyard near the pool and settled in. I had whipped up a pitcher of mojitos with mint I had grown in the garden, and poured two glasses. He pulled out two Cuban cigars and we were ready to relax for a couple of hours.

Curiosity got the better of me so I asked how he had met Lorraine.

``Actually, I met her two months ago, the last time I was in this area. I stopped by Trader Joe's downtown to pick up some lunch items. She was checking out, just ahead of me. I couldn't keep my eyes off of her, as her shorts were too short, just enough to show her cheeks dancing every time she shifted her weight. From her body type, I figured she did Pilates or yoga. Her order was not too large so it looked like she lived alone. There was no pet food, which was an added bonus.

``So I casually asked her, `Are you going to make me lunch?' She turned around, looked me over, and responded, `Only if you pay for it!' With no hesitation, I added my stuff to hers on the conveyer belt, paid for everything, and followed her home.''

``I guess the rest is history, as they say.''

``Ha. I say you make your history. The moment was there and I seized it. There was no guarantee. But, more importantly, there was no regret. There won't be a time in the future when I look back and think to myself, `What if?'\,''

``OK, I was just speaking metaphorically.''

``Were you really? You hide behind your books and poetry. Has a woman ever really loved you? For no reason, I mean. Not because she sees you as a meal ticket or a prize to display to her friends. You have a theory for everything, but do you have the reality?''

``I think so, I can see it.''

``That you are good at: you can experience things vicariously through others.''

``I know the most intelligent and creative minds in history, you know that. I believe they are good teachers.''

``Sure you do. You know I am a nihilist; I don't believe in anything. Only what I can experience directly.''

``I am likewise a nihilist. That means I can choose to believe anything I want. And that means a world populated by Knights, Ladies, Poets, Prophets, Philosophers, Sages, Mystics, Artists, Heroes, even angels, gnomes, jinns, and gnomes. It is a difficult worldview, requiring an understanding of maths, physics, metaphysics, history, and so on. Not everyone is capable of it.''

``I see the same world as you, except that it is populated by criminals, psychopaths, usurers, gluttons, adulterers, idiots, and worse.''

I wasn't pleased with his sudden mood change, but he wouldn't let up.

``I know your circle of faux traditionalists who don't actually follow a real tradition. They think they are bad boys for being banned by fb for a week. For all the talk of Heidegger, they are Cartesians at heart, men without qualities, with a detached ``I'' that sits in judgment on the `modern world'.

``They are not a Dasein, with mysteriously given attributes, who is thrown into a particular world at a time and place with unchosen relations. That is the battle to be fought, not the imagined one in their heads.''

As he had never spoken so harshly to me before. Perhaps it was the sun and the alcohol. I have known him a long time, sometimes closely, but often with gaps in our relationship. Usually, we are so close that it seems that we share the same brain, but use it for different thoughts.\href{https://www.gornahoor.net/wp-content/uploads/2021/07/BrainInsideBrain.jpg}{}

Since our smokes had gone cold, we had to pause momentarily to re-light them. That gave us a chance to change topics.

I still have unpublished notes from the summer he spent as a dance host so he could get free Mediterranean cruises. Or the three times, he was selected as a poster boy for matchmaking services catering to women; he got an invitation to the Marine Ball at the Jamaican embassy out of it. I'll clean them up and post them someday, but I was interested in a simpler topic that afternoon. As our moods changed from Saturn back to Jupiter, I brought it up.

``I got involved in Twitter thread posed by some young women discussing what is the best first date. I tried to interject some thoughts but got nowhere.''

``You don't seem like the Twitter type, my friend,'' he responded.

``Apparently, most people agree with you. I am totally ignored on Twitter. That is why I decided to write a post about it and would like your input.''

``I do have a lot of experience in that area. It will make for a nice conversation.''

We continued talking until dusk, when we had to go inside to avoid the mosquitoes. Soon after, Lorraine arrived to pick him up.\footnote{DISCLAIMER: This is a work of fiction and any resemblance to persons, living or dead, is probably coincidental.} 

\url{https://www.youtube.com/watch?v=QBUOKnLdJpc}

\flrightit{Posted on 2021-07-13 by Cologero}

