\section{White Boi Summer}

\begin{quotex}
``There's no use trying,'' she said: ``one can't believe impossible things.'' ``I daresay you haven't had much practice,'' said the Queen. ``When I was your age, I always did it for half-an-hour a day. Why, sometimes I've believed as many as six impossible things before breakfast.''
\flright{\textsc{Lewis Carrol}, \emph{Alice in Wonderland}}
\end{quotex}

For today only, believe that humanity has signed a peace accord with alien intelligences from outer space.

\paragraph{From Predestination to Providence} Our Saturnine philosophers represent a decline: from Spinoza's God of Reason, to Schopenhauer's blind Will, to Santayana's meaningless movement of matter. From the planet Saturn, looking down at the Earth, humanity seems to be driven by forces beyond their awareness or even common understanding. The philosopher can do nothing but observe detachedly from afar. But Dante reveals what lies far beyond the Sphere of Saturn: A long ladder leads to the Heavens, but his deep love for Beatrice, makes his ascent instantaneous.

\paragraph{Virginity}. For a boy, unless he is a born saint, his plan to lose his virginity is ``first available.'' It is my impression, however, that a girl gives more thought to it, as to person and circumstances. Surprisingly, the decision does not necessarily depend on meeting some great love. In effect, then, the woman is the selector in the game of pseudo-Darwinian selection.

\paragraph{Three Initiates} Rene Guenon mentions, in various essays, the works of Shankara, Ibn Arabi, and Dante as prime exemplars of initiatory texts. You would think that anyone ``influenced'' by Guenon or members of the so-called ``Traditionalist School'' would rush out to understand those texts. I haven't found that to be the case, however.

\paragraph{Understanding a Text} Reading an esoteric texts requires understanding its meaning, not just the grammatical arrangement of the words. In a letter to L C d'Amiens of 26 June 1937, Guenon explains:

\begin{quotex}
Reading the Quran can certainly ``open'' a lot of things, but, of course, on condition of being made in the Arabic text and not in translations. For this and also for some esoteric writings, this is something that has no relation to the external and grammatical knowledge of the language; I was also told the other day of the case of a Turk who admirably understood Mohyid-din Ibn Arabi, while in his life he had never been able to properly learn fluent Arabic. On the other hand, I know teachers from El Azhar (Cairo) who cannot understand a single sentence.
\end{quotex}

\paragraph{Statistical distribution} Although metaphysical understanding never changes, that is not true of our knowledge of mathematics and physics. For example, many assume that qualities, or ideas in the Divine Mind, are unalterable in their descent into manifestation. The truth, however, is that these qualities are distributed statistically, usual as a Normal Distribution, but not always. Hence, in a given population, a quality varies both in terms of count and of intensity. Otherwise, it is difficult to understand how psychological qualities like Intelligence, the Big Five, and disorders are distributed in a population.

\paragraph{The Top Dog} Some women are attracted to the Alpha dog and others to the rescue dog. You want the former type, since the energy between them is strong. The latter type makes her man into a ``project'', but if that does not work, she becomes an SJW.

\paragraph{Deliverance} It is important to insist that G is not proposing a new ``philosophy', or worse, a new ``school of thought''. Rather, he is describing a path to Deliverance, which is unachievable just by ``thinking'' about it. That teaching is basically invisible to most readers. Guenon explains that the true Self transcends all states:

\begin{quotex}
The being who manifests himself as a human individual in a certain state also contains the possibilities of all the other states.
\end{quotex}

\paragraph{Explorers of the Mind} I was tempted, at one point in life, by the likes of Alan Watts, Timothy Leary, William Burroughs, and Ram Dass, by the idea that the next stage of exploration was of consciousness, since the Earth has already been explored. This led to a period of risky behaviour and involvement with some oddball people. Ultimately, I concluded that natural means are the only effective tools. Moreover, the use of any compound that darkens the Intellect is sinful, since it inhibits the single characteristic that defines the Human State.

\paragraph{Gnosis} The Quran says, ``And he who has been given hikmah (gnosis), has been given much good.'' For the Christian, Gnosis is a gift of the Holy Spirit. Do not be dissuaded by those super-correct people who will call you a heretic. On the contrary, those who have been given this gift, and spurn it, are mistaken.

\paragraph{Christian polytheism} The Trinity is a difficult doctrine, and many Christians are material polytheists without realizing it. One fellow explained that the Holy spirit is a ``Being'', which is explicitly denied by the Nicene Creed. It smacks of polytheism.

\paragraph{The Absent Minded Professoressa} Years ago, I met a college professor who, one might suppose, has a certain measure of intelligence. We seemed to hit it off and she offered me her phone number. After a suitable interval, I called and left a message. A maelstrom erupted. She changed her phone number and reported me to the police for ``harassing'' her. Although I've known a lot of creepy guys, who actually do harass women, but in my experience, they can't always be believed.

\paragraph{A Gentleman's Knowledge} The prole mind is interested in knowledge only if it has ``cash value'', as William James described it. So he watches ``how-to'' videos on youtube. A gentleman, on the other hand, is full of general knowledge of literature, art, poetry, science, mathematics, history, metaphysics, etc. An evolutionary psychologist might say that it demonstrates that he has the intelligence and the resources to pursue such knowledge. Yet he is not a specialist in any of them, for that would make hima hedgehog and he is a fox.

\paragraph{Unexpected death} In the olden days, an Eastern Orthodox town had a mass burial ground for foreigners, visitors, suicides, heretics, etc., who could not be buried in hallowed ground. Even those believers who died suddenly, perhaps by accident, were buried there because they did not have the benefit of a final confession and the last rites.

\paragraph{Living in the Fast Lane} It is said that a drowning man experiences his whole past life. Rudolf Steiner provides a more interesting perspective. A man who dies a sudden or accidental death, experiences his \emph{future} life as it may have progressed had he lived. This all happens in an instant. Perhaps he even sees his final confession and last rites.

\paragraph{Til Death do we part} Marriage ends at death, but how many believe it? If that be true, then why do people feel bad when their spouse dies? Maybe they have been with their polar being. But there is but one Polar Being, all others are just biding time. That is a hard teaching. But the feeling of meeting one's polar being is unforgettable.

\paragraph{The Prince} Naida Muslić contrasts the goals of completion and perfection\footnote{\url{https://www.orphicinscendence.com/post/momma-how-could-you-do-that-to-me}}, applicable to the feminine and masculine elements respectively. The devouring mother, however, shames the daughter into perfection, which is opposed to her normal psychological development leading to completion. The perfect male (the ``Prince''), on the other hand, is incomplete. In the fairy tale \emph{Rapunzel}, the Prince finds his dream lover, but then becomes blind (i.e., unconscious), thereby losing her. When he does ultimately find her, she restores his sight, thus making him a complete being.

\paragraph{Leftover Scotch} Do you know the feeling when you fall asleep, leaving half a glass of fine Scotch on your desk, and you awaken in the morning and think it would go good with your bowl of Cheerios, but then discover that it does not tase good because all the alcohol has evaporated?

\paragraph{The Human State} I often enjoy the Human State and tend to spend much time there. I have learned not to take it so seriously. Yet, there is something much better, so I am now exploring the planetary and celestial spheres, with a sweet little star as a guide.

\flrightit{Posted on 2021-06-01 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Arthur Konrad on 2021-06-01 at 04:58 said:}

''In effect, then, the woman is the selector in the game of pseudo-Darwinian selection.''

That would only be true if the female in the Darwinian paradigm enjoyed freedom to choose a weaker male as well. Since females never bind themselves to a loser in a duel, then the correct term for her bondage to a stronger male is ``submission'', not ``choice''.

\end{sffamily}
\end{footnotesize}

