\section{Happy Wife, Unhappy Life}

I watched the movie, \textit{Let Him Go}\footnote{\url{https://www.imdb.com/title/tt9340860/?ref_=nm_flmg_act_3}}, so you don't have to.


George is a retired sheriff living with his wife Margaret, their son James, his wife Lorna, and little Jimmy. James dies in a ranching accident, so Lorna eventually marries Donnie who belongs to some evil family. Donnie, Lorna, and Jimmy move away. Margaret is upset about that and has concern for Jimmy. She packs her bags with the intention of finding the new family and ``rescuing'' Jimmy. That is called ``kidnapping'' because grandparents have no custodial rights over the grandchildren.

Nevertheless, she stubbornly cannot be dissuaded. So George agrees to go with her, against his better judgment. Then stuff happens.

The lesson is that if you follow every crazy idea of a woman, you will do things you don't want to do, go to places you don't want to visit, and deal with people you don't like. Then some bad guys will chop off your hand and you will die a horrible violent death trying to get her what she wanted in the first place.

Then she lives happily ever after with Jimmy, keeping the bloodline intact.

\url{https://www.youtube.com/watch?v=GfMkjdIc24I} 

\flrightit{Posted on 2021-11-12 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Allogenes on 2021-11-12 at 20:30 said:}

Brilliant

\hfill

\texttt{Presbyter Johannes on 2021-11-13 at 13:55 said:}

Seems pretty much in line with my relationship experience.

\hfill

\texttt{Balder on 2021-11-14 at 15:58 said:}

Love's cruel fate that tears the worlds asunder.

\hfill

\texttt{Joe on 2021-12-03 at 13:12 said:}

So he had to make the ultimate sacrifice by giving up his dearest possession, his retirement. Beautiful story, glad everything worked out in the end. I read yesterday that ``men manipulate the world, women manipulate men.'' What does that say about patriarchy?

\end{sffamily}
\end{footnotesize}

