\section{Naima}

\emph{Ida}\footnote{\url{https://www.musicboxfilms.com/film/ida/}} is a 2013 Polish film about a young woman orphaned during World War II and raised in a convent. In 1962, she has to decide whether to leave the convent or remain. Before taking her vows, the wise Mother Superior sends her out into the world to learn of her past. Her only living relative is a hard-living, chain-smoking, heavy-drinking, promiscuous, cynical aunt named Wanda. Ida learns that she is Jewish.

So Ida and Aunt Wanda go on a road trip to discover the fate of Ida's parents. On the way, they pick up a hitchhiking jazz-rock saxophonist, Lis, on his way to a gig. They listen to him play, although there were no sparks with Ida. He was playing \emph{Naima}, which was a nostalgic moment for me. I recalled hearing John Coltrane as a teenager which was a shock to this suburbanite. I had never heard sounds like that before, so I spent hours listening to him and Pharoah Sanders. Of course, I had the advantage of playing them on vinyl through a high-end audio system with expensive speakers. Listening to mp3 files through a computer audio system is almost criminal.

Together they visit the family's home to find it occupied by a Polish family. Ultimately, the owner, Felix, admits murdering Ida's parents to take over their property. In return for a promise not to try to claim the house, the wonder takes them to where the bodies were buried. Along with Ida's parents, Wanda's young son was also buried. Felix explained that the boy was too swarthy---Poles are rather pale---and circumcised. Since he could not pass as a Pole, he had to go. Ida, on the other hand, looked Polish so she was spared.

Ida and Wanda took the family's remains to be buried in a Jewish cemetery. After the war under the communist regime, the tables turned and the aunt became a bloodthirsty state prosecutor who was able to extract revenge for crimes committed during the war. But that was not enough.

Ida returned to the convent. On the other hand, Aunt Wanda, still unhappy with her life, defenestrated herself. Now a life of tobacco, alcohol, and promiscuous sex can reveal a lot about a person. For many people, that sounds like just another Saturday night. They have high time preference. Aunt Wanda is more intelligent and sophisticated and has low time preference. However, as an atheist, she had no higher values to live by. The future would be the same as the past. She had no prospect of salvation to atone for such a life.

Ida realizes she is still not ready to take vows. Having confronted her past, she leaves again to confront her future. She hooks up again with the saxophonist Lis and spends the night with him, sacrificing her virginity.

Naked in bed together the next morning, she asks, ``Now what?'' He suggests, ``we get a dog, get married, have children, buy a house.'' And ``then what?'' she asks again. ``Life as usual,'' he responds. And then what? Then we bicker over money and sex, fight about the children, have our friends over for dinner, And then what? And then we die.

That possible future has no appeal to Ida. She has a much lower rime preference, which only eternity can satisfy. We see her walking on the road with her bag. She is returning to the convent to devote her life to her God.

\url{https://www.youtube.com/watch?v=ikZ\_0BEk\_rA}

\flrightit{Posted on 2021-11-04 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{veritycacciatrice on 2021-11-05 at 04:12 said:}

Sounds great on paper.

I lived each of those lives (above more or less) and made the choice in my very late twenties to settle only for my other half, who I eventually found. It seemed hopeless for years and that was fine too.

Yes we do all that sniping stuff at each other but it's harmless banter, but also yes I dedicate my life to God simultaneously and he and the children know it. The balance is usually glorious.

So. In a nutshell this film reeks of (... ) propaganda.

Because- and this is what changed my mind about being a cat-lady Hermit (which I still consider myself to be in part), I doubt I'm perfect, therefore I will need to return.

Therefore I would like to return to my own family... 

Therefore I ought (and do) educate my own blood with every wisdom and foolishness learned this life, so they're enabled to recognise the illusions and adjust their choices accordingly.

Fingers crossed they breed well. I hear a third cousin is ideal.

\hfill

\texttt{Arthur Konrad on 2021-11-05 at 11:52 said:}

`Ultimately, the owner, Felix, admits murdering Ida's parents to take over their property, apparently not an uncommon practice'

It is a Jewish film, and Jews never waste an opportunity to address their present political agenda


\hfill

\end{sffamily}
\end{footnotesize}

