\section{The Empty Man}

\begin{quotex}
An empty man uses the completeness of a woman all the time. A complete woman is dangerous in her completeness, more so than a man. She is unreliable, moody, nervous, but also capable of great changes. Women like that can pick themselves up and go anywhere. They'll do nothing there, but that's because they had nothing going to begin with. Empty people, on the other hand, can't jump like that anymore, but they're more reliable. The Nagual said that empty people are like worms that look around before moving a bit and then they back up and then they move a little bit more again. Complete people always jump, somersault and almost always land on their heads, but it doesn't matter to them.
\flright{\textsc{Carlos Castaneda}, \emph{Second Ring of Power}}

\begin{verse}
Nothing real can be threatened.\\
Nothing unreal exists.\\
Herein lies the peace of God.
\end{verse}
\flright{\textit{A Course in Miracles}}
\end{quotex}


I wasn't expecting much from this film about the Empty Man\footnote{\url{https://www.imdb.com/title/tt5867314/?ref\_=fn\_al\_tt\_1}}, assuming unjustly that it is yet another teenage slasher film. That is, until I noticed the students were attending ``Jacques Derrida High School'';. that caught my attention.

The story begins in Bhutan, where the experience of Buddhist emptiness is quite unlike California happy talk Buddhism. Paul, one of four hikers, discovers the Empty Man and is never the same again. His three companions meet their demise in a murder-suicide, presumably suggested by the Empty Man.

\paragraph{Emptiness}


Several years later, in the US, James, a private detective, is alone at a Mexican restaurant, celebrating his birth day ... literally. Then begins Day One. James becomes self-aware, his I is born\footnote{\url{https://gornahoor.net/?p=57}}; he remembers growing up in San Francisco, he recalls the deaths of his wife and young son. They cry to him, ``Where were you?'' Of course, how could he be ``there'', since ``they'' are just a memory? And bringing the past into the present just causes pain.

Sitting on a park bench, his girlfriend's daughter, Amanda, approaches him. She explains that she found a wonderful teaching that helped her realize that nothing can hurt you, because nothing is real. Or maybe she meant, ``nothing'' is real. Reality does not create thoughts, just the reverse:

\begin{quotex}
What's real starts here she points to her head and ends up out there. What we think about with focus, attention, and repetition, we manifest. But what if there is a secret truth? What if our thoughts begin somewhere else and they travel through us like a signal traveling down a wire? Thoughts that are old and hidden and singular.
\end{quotex}


When she suddenly goes missing under suspicious circumstances, he tries to find her. Among her effects, he finds a flyer for the ``Pontifex Institute''; pontifex refers to a bridge that may cross an abyss.

\paragraph{Abyss}


He attends a lecture at the Institute and hears:

\begin{quotex}
You are complete in yourself. There is no struggle; struggling has failed; struggling is what wakes you up every day. There are no distinctions. Therefore, we deny that there is such a thing as right or wrong. They are evolutionary constructs that foster the illusion of separateness. The Empty Man beckons you to discover the true face of the world.
\end{quotex}


Privately, the speaker tells James:

\begin{quotex}
Our thoughts feed and are fed by the noosphere, which is the sum of all conscious thought transmission from which can be stimulated through the application of certain vectors very much like a virus.
\end{quotex}


There are three intakes required for proper development:

\begin{itemize}


\item Atmosphere: Air

\item Biosphere: Food

\item Noosphere: Thought and images
\end{itemize}


This is especially important for children who need pure air and nutritious food. What may be forgotten is that they need to fill their minds with positive thoughts and healthy images. Sitting them in front of the TV all day is not at all helpful.

\begin{quotex}
The Noosphere is real. People do indeed tune into it, ingesting thoughts at random. They then mistakenly take the thoughts as representations of reality. Quite the contrary, the truth is that what they take to be ``reality'' is nothing but the projection of their fears onto the experienced world. Through repetition, they become collective stories and worldviews that are superimposed on the actual world. The fear and anger they fell toward the world is misdirected.
\end{quotex}


Constant repetition turns ideas into gibberish, even more so of entire bodies of thought. The leader mentions, in particular, Nietzsche's quote about the abyss looking back at you. It has been repeated so often by yobs on the Internet, that is has lost its impact.

\begin{quotationx}
I have frequent dreams about the abyss, but not just dreams. In the twilight states between sleep and awakening, disturbing images of an abyss enter my consciousness. It may be a building or a natural structure. I look down, straight down, and experience the same vertigo as if it were there physically in front of me.

It is not the fear of falling that is the cause. Rather, it is the fear of jumping that raises alarm. The return to the unreal status quo is a bad option, since it is just more repetition of the same old. On the other hand, there is no bridge to cross it. At such moments, it is necessary to make a decision and accept what life has offered you. So you jump.
\end{quotationx}


\paragraph{Tulpa}


The Empty Man is the abyss. On Day 2, you see him and on Day 3 he stares back at you. Then you know fear! If James is born on Day 1, then he comes to full life on the third day, when he learns his true identity.

From the Pontifex Institute, James is led to a secret camp in the countryside. There he discovers a dossier containing a complete and factitious biography of his life. He is ultimately led to a hospital where the aforementioned Paul lies comatose, on the verge of death. James learns he is a Tulpa, that is, an artificial thoughtform created by the cultists, as Amanda described. James kills Paul and becomes the new transmitter of the Empty Man. The cult members then bow to him, as they seek the Primordial Thought that created the world.

Unbeknownst to them, apparently is that the purpose of Tulpa creation is to learn to face one's fears. That accomplished, the Tulpa is destroyed the instant the illusion is destroyed.

\url{https://www.youtube.com/watch?v=0D4V5awe-PA}

\flrightit{Posted on 2021-07-22 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Paulo Adolpho on 2021-07-24 at 01:35 said:}

There is the so-called luminous emptiness, and its opposite, Buddha called himself empty, but the so-called emptiness of today people, are not empty, they are full of shit, or rather, they have a big wound in them, a hole that sucks their energies. Can you understand the difference, and how words can confuse us!?


\hfill

\end{sffamily}
\end{footnotesize}

