\section{Pumpkin Spice Muffins}

I woke up so happy on Venus day this morning, so I decided not to go back to sleep. I was thinking about Pumpkin Spice Muffins. I made some earlier in the week and they attracted family, friends, and foes, before I could even photograph one for this recipe. They look like Double Chocolate Chip Muffins\footnote{\url{https://gornahoor.net/?p=14871}}, except that they are brown, and they don't taste at all the same.

Like so many other things, pumpkins originated in the New World, yet another legacy of forces set in motion by Christopher Columbus. No, a random ``discoverer'' counts for nil. That is why Lief Erickson has no enemies. A man in the world who accomplishes something quickly finds out that he will have friends and foes; he must learn how to deal with both.

So I leave you with a recipe for Pumpkin Spice Muffins. Once you learn the basic Mother Muffin, then you can easily adapt that knowledge to make specific muffins of your choice.

\paragraph{Pumpkin Spice}

If you can't find pumpkin spice, then you can make your own using 1 teaspoon cinnamon, 1/2 teaspoon ginger, 1/4 teaspoon ground nutmeg \& 1/4 teaspoon ground cloves. Variations of pumpkin spice have been found in North American cookbooks since the 18\textsuperscript{th} century.

It is so popular in the Autumn that a variety of products flavored or scented with pumpkin spice can be found. These include muffins, pies, candles, coffee, lattes, beer, vodka, cookies, aerosol spray, toothpaste, and even Spam.

\paragraph{Ingredients}

Metric equivalents are within parentheses.

\begin{itemize}

\item 11/2 cups (240g) all-purpose flour

\item 3/4 to 1 cup (150g to 200g) sugar

\item 1 Tablespoon (15ml) baking powder

\item 2 teaspoons (10ml) of pumpkin spice

\item A little salt

\item 6 tbsp (84g) melted butter

\item 2 large eggs, beaten (a US large egg is 57 g. A European large egg is 63 g, so close enough)

\item 1/4 cup (60 ml) sour cream

\item A little Vanilla (optional)

\item 1 15 oz. can of pumpkin puree. About 1 cup (250 to 500ml)
\end{itemize}

\paragraph{Mixing}

Combine all the dry ingredients in a bowl and mix thoroughly.

In the meantime, melt the butter, either stovetop or microwave (easier), and beat the eggs.

Add the wet ingredients (sour cream, butter, eggs, pumpkin puree) and mix thoroughly. It should turn into a shade of orange (also called brown) loose paste. It will be less viscous than chocolate muffins, so cooking time may be extended.

\paragraph{Baking}

Spray oil into a muffin pan and put some of the muffin mixture into each of the holes.

Bake at 375°F (190°C) about 20 to 22 (20 to 22) minutes

I use a Ninja toaster oven which heats up quite quickly and holds heat well. You may need to cook longer. A trick that looks really cool if you are trying to impress friends is to insert a wooden toothpick into the muffin. If it comes out clean, the muffin is ready; otherwise, it needs longer. In practice, the technique is not very helpful, so I just look see if the top is encrusted.

This makes a dozen (12) American muffins or 10 metric muffins.

This is a Halloween favorite. It was banned by the BBC when it first came ou on the grounds that it is to morbid.

\url{https://www.youtube.com/watch?v=l2PoSljk8cE}

\flrightit{Posted on 2021-10-29 by Cologero}

