\section{Brief Encounter}

\begin{quotex}
The two lovers, conscious of their presumably integral polarity, are called upon to straightway renounce carnal love. They must do this consciously and of a common accord, at the same time cultivating the Sacred fire of their Love, which then takes the form of \emph{courtly love}.
\flright{\textsc{Boris Mouravieff}, \emph{Gnosis Book II}}
\end{quotex}


Schopenhauer's criterion for a good novel is not the amount of action in it, but rather how well it reveals the inner life of the characters. This can be more difficult to achieve in film. The highest revelation is always in words, not pictures, because hearing is the most subtle sense, associated with the element ether (akasha).

For the functionally illiterate of today, who prefer action to thought, that is hardly an issue. CGI characters don't have an inner life; one suspects that a large part of the population has a rather limited inner life. Or else, much of it is tedious and unpleasant. That is why there is such a large usage of mind-numbing agents in the West like alcohol, as well as legal and illegal drugs.

A common technique in film is to use a voiceover as, for example, in the first film version of Dune. In most action-packed movies, it is difficult to understand the motivations for the actions of the characters. Usually, the motivations are so trite that the viewer simply assumes one. Film as art, rather than mere entertainment, must rely on dialog or facial expressions for that purpose.

The film \textit{Brief Encounter}\footnote{\url{https://www.imdb.com/title/tt0037558/?ref_=fn_al_tt_1}} is a 1945 British film which takes place just before WWII. It is considered one of the best British films ever made, so it is worth the trouble. I don't expect it to be superseded given the current cultural decline of the UK.

Laura is a dissatisfied housewife who realizes that an essential element is missing. Laura, it must be noted, is the name of the mysterious woman who was the subject of so many of Petrarch's sonnets. Hence, she must not be understood as a contemporary woman eager for sexual experiences, although she is never quite sure of what is missing in her life. The return to the Primordial State requires the union of the man and the woman who are uniquely suited to each other, just as were Adam and Eve. Her real desire is to be known, but she can be known only by the one who knows himself. There is no distinction at that point.

\paragraph{Plot}

What is it like to be a middle class housewife, who realizes that she is married essentially to a stranger, and is bored with her uneventful life of grocery shopping, watching movies alone, etc.; in short, she is trapped in the illusions of ordinary life? Then she meets the charismatic general practitioner Alec Harvey at the train station, who seduces her. She hears Rachmaninov's Concerto \#2 whenever her imagination wanders into an alternative future. Concerto \#2 is arguably the most beautiful concerto every composed.

Half of the film is her inner monologue and she is frustrated because she cannot share her innermost thoughts with anyone else: they would either misunderstand or else judge her harshly. Her delightful thoughts of love arouse her kundalini which then fills her imagination with visions of vacations in the tropics while dancing by the ocean at midnight. Boris Mouravieff reveals the source of such dreams which must never be ignored.

\begin{quotex}
With the aid of impulses proceeding from the sexual centre via the motor centre, the emotional centre has erotic dreams which can seem entirely true. On the other hand, if it dreams with the pure and direct aid of the sexual centre it creates ideal images of its polar being, founded on the experience of its present life at previous experiences. With the sexual centre's direct intervention the emotional centre can also create idealized images of living beings that the man has met, or who resemble his polar being.
\end{quotex}

Then her lover forces a decision. She realizes that, in order to love him, she will have to be a liar to her husband. But isn't a return to her current life also a lie? She implores her lover to release her since she lacks the strength herself. Had they undergone some esoteric training, the result might be different. Mouravieff describes the case similar to that of Laura and the Doctor;

\begin{quotex}
Two polar beings who meet at a mature age, when life has already burdened each of them with a karmic load. In such cases, the first sacrifice demanded is the renunciation of a physical relationship, and the first exploit consists in the methodical liquidation of the respective karmic burdens, If they ardently and effectively undertake some esoteric work in the same direction, the moment will come when they will be purified. Having become courtly, their Love will assume all its objective force, and in the purity rediscovered in this way, they will finally be convinced of the reality of a polarity that they had felt intuitively. There is no possibility of error at this stage. At this moment, the Second Birth will unite them forever in the midst of the vivifying Love; and death, finally conquered by this, will lose all semblance of catastrophe for them.
\end{quotex}

Their final meeting is at the cafeteria at the railway station. He had made plans to leave England, boarding his train, as she boards the train home, expecting never to see him again. Suddenly she disembarks, but it is too late; he is gone forever. She is forced to return home to her husband.

\begin{quotex}
\begin{verse}
Tis better to have loved and lost,\\
No matter how high was the cost,\\
The pain of loss at the cafeteria hall\\
Is less than never to have loved at all.
\end{verse}
\end{quotex}

\url{https://www.youtube.com/watch?v=rEGOihjqO9w} 

\flrightit{Posted on 2021-11-14 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Presbyter Johannes on 2021-11-14 at 22:45 said:}

In the subject of films, this might be of interest to you... \url{https://youtu.be/vqnjzVX8EKA}

I'd have a lot to talk about this particular subject as well, but don't quite have the time. I hope you like it though.

\hfill

\texttt{Auld Wat on 2021-11-23 at 07:48 said:}

I have definitely experienced dreams in which the `ideal' woman for me has appeared. The sense of loss on waking is intense, and the desire is to find that woman immediately. Mouravieff suggests, as you point out, that it is possible to use the sexual centre to directly intervene in the emotional centre to create `idealisations' of those who we have met. So we don't necessarily need to physically find that ideal woman, although if we did that would be a great blessing if the opportunity to purify themselves is taken. As was said to Heloise regarding Peter by the Abbot of Cluny, you will meet again where beyond these voices there is peace.

\hfill

\texttt{Tom C on 2021-12-02 at 08:58 said:}

Great post (and great movie), but surely you mean Petrarch, and not Plutarch? (Plutarch was the Greek historian). Thanks for the plug to Geist magazine!

\end{sffamily}
\end{footnotesize}

