\section{The Fog Travelers}

There is an episode in the \textit{Neverending Story}\footnote{\url{https://en.wikipedia.org/wiki/The\_Neverending\_Story}} about people who sail an ocean of fog. They steer their ship by a collective chant and their collective will. They don't have names nor do they ever do anything as individuals. Bastian feels attracted to this life at first, because he had become the strongest and best in everything --- thus he felt very lonely.

But then he witnesses the sudden disappearance of certain crew members on the ship --- they are taken by a giant crow. Nobody is interested in the missing people, the remaining crew members don't even notice that people are missing! He starts feeling uneasy and realizes that what he is looking for is not losing himself in anonymity, but rather the desire of being recognized and loved for who he is. He then leaves the people of fog travelers to look for that love.

I'm mentioning this, because sometimes I get the feeling that no matter how brilliant and intelligent people are, there is a tremendous fear of saying something ``wrong'', not as a simple mistake, but to reveal a thought that comes from somewhere other than the source of the chants of the fog travelers. Something beyond the fog, where you don't just wander around in other people's thoughts and deal with, or ignore, what ``happens'', but rather create a thought that breathes away the fog.

This clinging to the fog may be the only way for the fog travelers, because their survival depends on their collective will to navigate the fog. But you know I was in love with ``Bastian'' as a girl and I did understand his desire to move beyond what everybody else was ``singing''.

\url{https://www.youtube.com/watch?v=lHytjEj7B9g}

\flrightit{Posted on 2022-07-17 by Sibylle}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Balder on 2022-07-19 at 08:27 said:}

Only an awakened person can breathe the fog away. The thing is, first you have to pass through the inner fog before you can sing another song apart from the collective.

\hfill

\texttt{Balder on 2022-07-19 at 08:30 said:}

This came to mind also:

``It is certainly a good idea to plead to reason and common sense, but what if you're audience is in the midst of a collective psychosis?''

-- C.G. Jung

\hfill

\texttt{Arthur Konrad on 2022-07-24 at 18:18 said:}

Things that I've clamored for were often ushered in on a white horse right in front of my face, but it made no difference because I was hopelessly self-absorbed. Perhaps everything that is `wasted' in life too has some sort of redeeming power. I honestly admire attentive people, the kind of people who are not content to sit by when important things are taking place.

I loved, and will always love Mary.


\hfill

\end{sffamily}
\end{footnotesize}

