\section{Charade or the Primordial State}

\begin{quotex}
The important thing about the chivalric ideal is the double-demand it makes on human nature. The knight is a man of blood and iron, a man familiar with the sight of smashed faces and the ragged stumps of lopped-off limbs; he is also a demure, almost a maiden like guest in hall, a gentle, modest, unobtrusive man. He is not a compromise or happy mean between ferocity and meekness; he is fierce to the nth and meek to the nth.
\flright{\textsc{C S Lewis}}

When a traditional form is on the point of becoming extinct, its last representatives may deliberately entrust to the collective memory what would otherwise be lost beyond recall; that is, in fact, the only means of saving what can, in some measure, be saved. At the same time, the natural incomprehension of the masses is a sufficient guarantee that what has an esoteric character will not be laid bare and profaned, but will remain only as a sort of witness of the past for those who, in later times, will be capable of understanding it.
\flright{\textsc{Rene Guenon}, \emph{Fundamental Symbols}}
\end{quotex}

\emph{Charade}\footnote{\url{https://www.imdb.com/title/tt0056923/?ref\_=nv\_sr\_srsg\_0}}, starring the seductive Audrey Hepburn and the perennial leading man, Cary Grant, was a romantic comedy released in 1963. Historically, marriages were often based on convenience or were even necessary for survival. However, as a means of keeping alive the idea of the Primordial State, ideas of Romance was made part of popular culture.

The reunion of the male and the female represents the return to the Primordial State of Eden; the Edenic stage is the integration of the human being brought to his original centre. Thus Adam-Eve is the ``Universal Man''. This teaching was consciously preserved in the Middle Ages by Dante and the Romance of the Rose. The authors of the grail romances were probably less conscious of any hidden meaning to their poems, but preserved the teaching intact.

\paragraph{My One and Only}

Reggie, diminutive for Regina (the Queen), was on holiday when she announced to her friend that she no longer loved her legal husband, Charles. There she met the mysterious and handsome stranger, Peter Joshua. Lovestruck, she tried to seduce him, unaware that he was seducing her.

Back in Paris, she sees that her flat was totally emptied and that her husband had been murdered and thrown off a plane. Summoned to a CIA office, she learns that her husband was involved in a caper with four other men; they stole \$250K from the US Government, which was trying to recoup it.

They all believe that Reggie knows where her husband Charles had kept the money. They each threatened her, but they don't trust one another. Meanwhile, Peter kept showing up to protect her. Through a series of complexities, which are better watched than explained, Peter undergoes several identity changes. Although they rattle her, Reggie remains in love with the bearer of those name changes. She is able to discern the one soul that she loves despite the multitude outer identities.

\paragraph{Hidden in Plain Sight}

While everyone is looking everywhere for the money, it turns out that Charles was hiding it in plain sight all along. So many times, what you are seeking is right there in front of you, but you don't see it.

\paragraph{The Knightly Ideal}

Peter demonstrates the knightly ideal. Among the men, he displays cunning, strength, and power, while with his one and only love, Regina, he is ``a gentle, modest, unobtrusive man'', always kind to her and very protective.

Let's say that all's well that ends well. Peter's ``true'' identity is revealed, they marry, expecting the sort of bliss than can only be fully realized in the Edenic state.

\paragraph{Knights, Knaves, and Mansplaining}

As an interesting diversion from the movie plot, Peter describes a version of the logic puzzle named ``Knights and Knaves'' by the logician \textbf{Raymond Smullyan}\footnote{\url{https://en.wikipedia.org/wiki/Knights\_and\_Knaves}}, several years later. In frustration with Peter's lies, she asks how can she tell when someone is lying. He engages in mansplaining with this puzzle.

Suppose there are two Indian tribes: the white-footed tribe always tells the truth and the black-footed tribe always lies. How can you tell which is which?

Reggie, not illogically, suggests, ``why not look at their feet?''

Sometimes you have to break free of the system to finally see the truth. Love cannot be explained by the logic of the world. The Queen found her knight.

\url{https://www.youtube.com/watch?v=C6T2Q4XO7uA}

\flrightit{Posted on 2022-01-06 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Anon 7777777 on 2022-01-10 at 19:50 said:}

I'd like to humbly recommend Diary of a Country Priest if you haven't reviewed that yet

\end{sffamily}
\end{footnotesize}

