\section{Power and Random Acts of Kindness}

Growing up in the Ten Hills section of Somerville, Massachusetts, up to the age of 11, was a pleasant experience. Somerville before my time was a lovely suburb, with a beach on Mystic River, and everything within walking distance. There were a large number of Italian and Irish families, the former from the North End of Boston and the latter from the South End. My grandfather told me that occasionally there were pitched battles between the North and South Ends, but those antagonisms did not pursue them into Somerville.

We were allowed freedom of movement, and our only instructions were to return home when the street lights turned on. That must have provided the mothers with some relief, unlike suburban life today when children are always underfoot. There was less traffic during the day because two car families were rare. Kids were everywhere, so there were plenty of things to do. We learned street games -- passed down from the older kids to the younger -- like Hide and Go Seek, Giant Steps, Red Light, Freeze, Kickball, and various versions of Tag.

Boys and girls played those games together, unless we went to the park to play baseball or just wander through the streets. We took pee breaks like feral cats, usually in one of the many alleys. The girls had to go home, except for Susan who liked to come and watch the feral boys do it. Only boys played Cops \& Robbers or Cowboys \& Indians; I presume the latter came is frowned upon. Of course, we all had cap guns.

\paragraph{Hernia Surgery}

I was born with a congenital hernia that was repaired when I was 5 or 6 years old. It was not outpatient at the time, but required several days at the hospital. In the children's wing, the beds were lined up against the walls, more like a barrack. The boys were on one side and the girls on the other. They used to take our temperature in the ass, probably to humiliate us. Sometimes a girl would come to our side with the pretext of asking the nurse a question. God forgive us if a boy walked into the girls' section.

My father often told me that I almost died on the operating table, due to the anesthetic and an asthmatic condition. I did not understand why he would have mentioned it. However, over the course of my life, I have been rescued from illness, death, financial ruin, and incarceration by forces beyond my control. I am certain that it is because of all the novenas we attended as children. The church was filled with children, and after the nine days, we would receive a scapular or holy medal. I have always had a strong devotion to Mary and the Salve Regina is my favorite prayer.

\paragraph{My Best Friend}

My best friend, Bobby M., and I were the same age. We got to organize the games as the older kids moved on to other activities. Part of being buds was the occasional fights which always ended in a draw with no hard feelings.

On one occasion, we were seriously grappling on the ground in front of Lorraine's house. Her older sister, Elaine, came out and ordered us to stop. She screamed that we could die from fighting. Well, I was certainly getting out of breath and I figured that a teenager must know something. Nevertheless, neither of use paid her any mind. I learned a lesson that day: I would rather die than surrender. The battle ran its natural course and we both survived. There was another lesson that I did not understand until much later: a woman is perfectly willing to lie to you to make her point.

\paragraph{Back to the Hernia}

Coincidently, Bobby had to undergo hernia surgery. That gave me a good idea, I thought, which I ran by my parents. I told them that the next time we had a fight, I would punch him in the hernia area where he would be vulnerable. They were appalled and forbad it. I suppose that it was unethical to hit below the belt; it was also risky because it would leave my upper body and head area vulnerable to a counterattack.

I suppose that some of you 21\textsuperscript{st} century men, and most women, would be wondering why I wasn't planning some random act of kindness to welcome my bud out of the hospital. All I can say is that you and I have traveled in different circles.

\paragraph{Hierarchy}

You more astute esoterists will recognize the battle for hierarchy. As the two oldest boys, Bobby and I often struggled to achieve dominance. Nevertheless, hierarchy is not an evil, as people today believe, rather it is, and has been throughout history, the norm. Evil results from the failure to recognize the natural hierarchy; it such cases, physical force becomes the last resort.

Friend against friend, brother against brother, have existed since the first generation of the human race. Cain murdered his brother Abel rather than acknowledge that God preferred Abel. This is Cain's punishment:

\textbf{God}: you shall be a fugitive and a wanderer on the earth.

\textbf{Cain}: from thy face I shall be hidden; and I shall be a fugitive and a wanderer on the earth

By refusing to acknowledge the hierarchy established by God, Cain is condemned to wander the earth aimlessly and without purpose. Does that remind you of anyone you know?

Cain immediately grasped the consequences. He would never again know God. He will try one thing and then the next, but nothing will ever bring him back to knowledge of God. No number of pretty words on social media, nor chasing butterflies will ever compensate.

\url{https://www.youtube.com/watch?v=SmqLKr6gF-Q}

\flrightit{Posted on 2022-12-11 by Cologero}

