\section{Fernweh: I belong not in this place}

\paragraph{I belong not in this place}
\begin{verse}
I think that I belong not in this place,\\
Separated by an abyss of time,\\
I hear a sweet song from another space\\
Floating on the breeze that changes the clime.

With the Cherubim's song, it is admixed.\\
The ethereal voice makes my soul soar\\
Beyond the world that has had me transfixed\\
To fly with the angels who her adore.

Joined not in space, but by affinity\\
What I see and hear abides in my heart\\
Love lasts not in time, but infinity.

What would I really gain to taste and touch\\
What no cosmical force can tear apart?\\
Unlike the world I used to love so much.
\end{verse}

\url{https://www.gornahoor.net/wp-content/uploads/2022/11/Fernweh.mp3}


\flrightit{Posted on 2022-11-29 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Mike M on 2022-11-30 at 11:20 said:}

Once you jump you realize it was a comedy all along.

\hfill

\texttt{De Bosit on 2022-12-01 at 15:21 said:}

``Paradise is still on this earth'', this was one of the earliest sentences I read in Evolas Hermetic Tradition, just gotta find my spot... 

Nice to hear your voice.

\hfill

\texttt{Balder on 2022-12-02 at 06:35 said:}

The Harmony of the Spheres is througly a sight once you get a touch of the Ogdoad's transcendent impulse that moves them from the super-worldly Kingdoms of Light.

\hfill

\texttt{Sibylle on 2023-05-22 at 02:39 said:}

I have some blurred memories of an imagined ride on a carpet somewhere in my childhood home. I used a lot of pillows to secure the edges of the carpet, so I wouldn't accidentally fall off, since I remember I was flying very high, too high for a little girl on her own... You know when you get homesick for something that is so far away and so deeply hidden that you don't recall it properly. But you know that is your real home. You can call it Fernweh -- and that's what I thought it was when I was little, and I imagined I was flying to the stars... Now I am no longer afraid to fall, because I realized what no cosmical force can tear apart. And I know you are there, flying with the angels. You abide in my heart.

My thoughts and prayers are with the Salvo family.


\hfill

\end{sffamily}
\end{footnotesize}

