\section{UFO Research: Aliens}

UFO research\footnote{\url{https://en.wikipedia.org/wiki/Ufology}} is usually met with derision on one side, and passionate belief on the other. But in keeping with revolutionary science, it is necessary to be neutral, or zetetic. For the purposes of this article, only data that is consistent with the extraterrestrial hypothesis\footnote{\url{https://en.wikipedia.org/wiki/Extraterrestrial\_hypothesis}} --- i.e., UFOs are spacecraft operated by intelligent beings --- will be considered. There are many stories of not just visual sightings, but also physical contact with such aliens.

There must be thousands of contact stories but a cursory search did not reveal any organized body of evidence. We are only interested in physical entities, not any claims about demons or psychic entities. The assumption that such contacts are real, raises many questions. Somehow the aliens have solved these problems.

\paragraph{Navigation}

How does a craft navigate in outer space? Everything is moving in relation to each other, so there is no fixed point like the Polestar on earth. Even supposing we could send a craft to Alpha Centauri, how would the crew find their way back to earth? ETs must have solved the problem in some way.

\paragraph{Relativity}

As far as we know, it is not possible to surpass the speed of light, and even reaching a fraction of that speed would require large expenditures of energy, more than we can generate. The craft would need to be huge, which brings new problems, e.g., getting bombarded with space dust. There is no reason to doubt the Theory of Relativity, since is has never been empirically refuted. Hence, there are technologically insoluble problems with accelerating a craft to make space travel feasible.

But suppose Relativity is not true, so aliens are not actually limited by the speed of light and the energy required to reach it. That would be good to know.

\paragraph{Human beings}

Despite the diverse odd-looking depictions of aliens, they all seem to resemble human beings. They are upright and have sense organisms analogous to ours. It may be that that people, who have claimed contact with aliens, are unable to conceive of any being other than humans. Or it may be that human like qualities are pervasive in the universe, and are necessary to any intelligent beings. These properties may be metaphysical, as we believe, or they may simply be emergent properties of certain material arrangements, as physicalists believe. These are:

\begin{itemize}

\item They are material beings, presumably also based on DNA.

\item They are alive in a basic sense, as are plants.

\item They experience emotions as do animals.

\item They have intelligence. Thus, the aliens are rational animals like use; that is what permits fruitful contact between humans and aliens.
\end{itemize}

If we avoid biological definitions of humans, this all makes sense.

\paragraph{Moral Universe}

Whether you believe that the universe is fundamentally moral, as does this site, or that it is a human construct, reported alien contacts involve moral issues. This is necessary for rational animals. There is no point to intelligence, if the intelligent being is unable to choose between the alternatives presented by the intellect.

The motives of the aliens have often been in question, most famously in the \textbf{Twilight Zone}\footnote{\url{https://www.dailymotion.com/video/x80j7it}} episode \emph{To Serve Man}. Even the atheist Stephen Hawking was concerned about letting aliens know about our existence because they may have evil intents toward us.

\paragraph{Limits to Intelligence}

Aliens may reveal inherent and intractable limits to our intelligence. As a graduate assistant, I had to teach basic algebra to elementary school majors. It was a frustrating experience, because many of the students seemed to be constitutionally unable to learn the topic. I tried explaining in several different ways, but it was futile. Premed students have had to drop out because they could not understand organic chemistry, and business majors often can't pass an economics class. I am fortunate, because there is no academic subject that I have not been able to learn.

But what if aliens have knowledge the simply exceeds the capacity of our brains. There have been stories of alleged visits of aliens to earth, such as the one described in \emph{The Sirius Mystery} by \textbf{Robert Temple}. But in none of these stories, has anything been taught of much significance. To wit, they haven't taught unknown physics or cures for all diseases. Nor have they taught the best way to organize societies. You would think that they would have mastered those subjects. Perhaps, we simply cannot grasp what they know, so that we are forever limited in our knowledge.

\paragraph{Theological Issues}

Sometimes, people raise theological problems, viz., about the ``fall of man'', and how it relates to aliens. There are no difficulties to the esoteric understanding of events. The entire universe is fallen, not just on earth. We would be surprised to discover a civilization of immortal, totally good inhabitants.

Fleshly human beings, understood in the sense described above, are corruptible. The Adam of Eden was in a degree of existence above ours, which is why there was a ``fall''.

\paragraph{Sources}

Theories of Everything with Curt Jaimungal\footnote{\url{https://www.youtube.com/watch?v=eouxDJMlPuM}} has provided the following list of UFO research projects. I could not listen to much of them, because they ramble too much and struggle to get to the point. But maybe others will find good material.

\begin{itemize}
\item Need To Know (with Ross and Zabel): \url{https://www.youtube.com/c/BryceZabel}

\item Alien Girl: \url{https://www.youtube.com/c/Aliengirl111}

\item That UFO Podcast: \url{https://www.youtube.com/c/ThatUFOPodcast}

\item Cristina Gomez: \url{https://www.youtube.com/c/ParadigmShifts}

\item Red Panda Koala: \url{https://www.youtube.com/channel/UCvSbzThCfsiETLp3eOdVkNw}

\item GUFON: \url{https://www.youtube.com/c/GUFONRadioStream}

\item Thomas Fessler: \url{https://www.youtube.com/c/thomasfessler}

\item Truthseekers: \url{https://www.youtube.com/c/Truthseekershow}
\end{itemize}

\flrightit{Posted on 2022-07-24 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Balder on 2022-07-25 at 06:26 said:}

The UFO phenomenon never stops amazing me. In this issue otherwise intelligent and serious people seem to fall for the post-modern mythology / demonology more easily than an illiterate peasant; perhaps they are desperate inside.

The UFO cult has penetrated even into religion and esotericism. New agers are especially drawn to aliens, as proven by the likes of Benjamin Creme and his ilk. Even the esoteric Hitlerist, Miguel Serrano, claimed that the superior blonde Aryans literally came from space, namely from Aldebaran, and the supreme cosmic tragedy was that they intermingled with the dark races created by Yahweh. All this goes on to show how deeply `materialized' the concept on of a heavenly descent has become in modern minds.

Now I have seen some strange things in the sky like mysteriously moving lights that then suddenly disappear from sight but I'd rather opt for secret military intelligence as a cause than actual UFOs.

Nevertheless, Ancient Aliens must be the best program that has ever been broadcasted on TV for its deeply humoristic aspects (although there are actually people who take it seriously, and I'm guessing most of them are Americans). ``Now I'm not saying it was aliens but it was aliens.'' ?

\hfill

\texttt{Michael DeJoy on 2022-07-25 at 10:04 said:}

My take on the ufo story is that our physical universe is a closed system existing for the benefit of humans to carry out its destiny. All the discoveries of planets, solar systems, etc simply occur for the development/entertainment of man. It appears that the universe is infinite so that humanity will have hope of ``progress'' and that their beliefs in progress are good and correct. The universe that we see is like a painting existing in a museum along with many others. They are all different from each other existing in their own frame of reference. Probably the best of the sci-fi movies is Tarkovsky's ``Solaris'', where humanity finds an intelligent planet capable of interacting with people only on a scale which defies our understanding as to why it does what it does.

\hfill

\texttt{Balder on 2022-07-28 at 19:29 said:}

@Michaeldejoy: we are absolutely not living in a ``closed system''. Our world is interpenetrated by forces from below or above every SECOND.IDEA OF a closedworld can only appear on a closed mind. Every one who walks upon his own path in some point has to realize the different forces and powers, the national spirit around, above and below him.

WHEN THE KSHTRITA HAS OENETRTATWD HIS OEN HAART.

It is said that the modern requirement of openness is only veiled forum of hatred upon secret. There are good reasons for gentlemen not to indulge in a basically profane ponderings upon inclinations of one's own character. There is the point from where free Men act.

\hfill

\texttt{Zagreus on 2022-07-31 at 19:44 said:}

Any thoughts on Revelation 16:12-16

\hfill

\texttt{Balder on 2022-08-02 at 17:57 said:}

I must have been very drunk when writing the previous post and I apologize for it. Cologero, you are a very patient man.

\hfill

\texttt{Cologero on 2022-08-02 at 19:10 said:}

The spam detector blocked it, and I don't know why. Should I delete it or not?

\hfill

\texttt{Balder on 2022-08-02 at 20:00 said:}

There was a right idea but it was dispersed between a drunken thought and bad expression.

\hfill

\texttt{Balder on 2022-08-02 at 20:05 said:}

Did the spam detector also block my post on love post? Maybe it detects when a man is drunk which is exceptional! Does the AI know when I'm posting something when binge drinking? It must be the aliens!

\hfill

\texttt{Dimitri on 2022-08-03 at 10:17 said:}

I don't think mankind has encountered any physical, tangible aliens (yet), but many people have had strange interactions with ``psychic'' beings from different spheres of reality that they believed were aliens (this would probably include ``Lovecraftian'' beings). Charles Upton has written books on the subject as I'm sure you're aware. I do find it interesting that prominent theologians used to discuss these sorts of things openly, like Nicholas of Cusa, who wrote the following:

``Life, as it exists on Earth in the form of men, animals and plants, is to be found, let us suppose in a high form in the solar and stellar regions. Rather than think that so many stars and parts of the heavens are uninhabited and that this earth of ours alone is peopled -- and that with beings perhaps of an inferior type -- we will suppose that in every region there are inhabitants, differing in nature by rank and all owing their origin to God, who is the center and circumference of all stellar regions.''


\hfill

\end{sffamily}
\end{footnotesize}

