\section{Ides of March}

\begin{quotex}
The father loves his children, the mother loves her children, the children love their father and mother; but this is not like that, brothers. The wild beast also loves its young. But a man can be related only by similarity of mind and not of blood… \flright{\textsc{Gogol}, \emph{Taras Bulba}}

\end{quotex}
The Steiner version of the three Vedic gunas — Ahrimanic, Luciferic, and Christic forces — may be more expressive for us. How do events appear using those categories? Appealing to the superstition of facts misses the forest for the trees. This is a recent comment I wrote:

The temperature and pressure of a gas can be measured without knowing the positions and motions of each individual atom. That is because the gas is a totality, so the differences in individual atoms turn out to be irrelevant as they average out. Moreover, the individual atoms are not even aware that they are participants in a larger process.

The theologian \textbf{Michael Heiser} wrote the book, \emph{The Unseen Realm}, a few years ago. His thesis is that the stories about angels, Nephilim, demons, divine councils, etc., in the Bible need to be taken seriously. In practice, this means that there are influences in the world that transcend the combined actions of individuals. Whatever their individual intentions, higher intelligences, both benign and malign, are manipulating them for their own purposes. They have no power over physical events, but work through the hearts and minds of individuals.

I heard one pastor claim that the seraphim are dragons with wings and the cherubim are bulls with wings. That makes for beautiful symbolism, but that is not how we recognize their activity in thw world. Rather they reveal themselves — both angels and demons — through constellations of thoughts. It is necessary to become aware of those worldviews. Here, Nietzsche can offer us some insights. The worldviews are expressions of the Will to Power, not topics suitable for intellectual debate.

On the individual level, the first temptation of sinister forces is doubt — double-mindedness. Like every negative thought, detach yourself from the thought, find an anchor point, and observe it. THERE IS NO INTELLCTUAL ARGUMENT TO MITIGATE YOUR DOUBT. It is an existential problem, not an intellectual problem. Recall Tomberg's discussion of the fall, for reference.

Esoteric teaching — as least true esoteric teaching and not pop spirituality — is not anti-theology. Actually, just the opposite, because it teaches a worldview in which theology makes sense. The alternatives are inscrutable theological conundrums, like this discussion between intelligent and educated theologians. If your dog goes to heaven, then do his fleas also go to heaven? If you want to be a Zen Buddhist, then treat it as a koan and let me know.
 

Today, on the Ides of March, we celebrate the assassination of the Roman worthy Julius Caesar. Instead of killing Julius Caesar today, purge your mind of useless and vain thoughts.


\hfill

\begin{verse}
\textbf{What happiness}: it is night and we are alone!\\
The river is like a mirror, all ablaze with stars,\\
And there – lean your head back and look up –\\
What profundity and purity there is above us!\\
Call me mad! Call me what you will:\\
At this moment\\
Reason fails me,\\
And in my heart I feel such an outpouring of love,\\
That I cannot be silent, I won't, I can't!\\
I'm sick, I'm in love, but anguished and in love –\\
Just listen! Hear me well! – I cannot hide my passion.\\
And I want to say that I love you,\\
It is you, you alone that I love and desire!
\end{verse}

\url{https://www.youtube.com/watch?v=qr7BJ67xv3o"}



\flrightit{Posted on 2022-03-15 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Balder on 2022-03-16 at 10:54 said: }

I'm constantly surprised you refer so often to Steiner; wasn't he the very epitome of anti-tradition and a prophet of neo-spirituality? It's not along time I read Evola's `Mask and Face of Contemporary Neo-Spiritualism' in which the Steinerian current is given a very harsh treatment.

I don't think the Steinerian forces of Ahriman, Lucifer and Christ refer to the three gunas, at least not in Steiner's own thought in which they depicted the personalized fallen forces / entities of past humanities (manvantaras). They are more like the qlippoth of Qabalah (The ``kings of Edom that fell" to quote Guénon), the unusable residues of past manvantaras that still haunt men.

To not engage in discussion whether Steiner's theories of these fallen entities is correct or not, I still think they depict quite well the impulses that a human may undergo in one's trials. In Lord of the Rings they are the ``shadows of ancient kings", the Nazgul (dupes of Sauron). In a dream a few years ago I fought them (successfully); in that dream they were first ``Nazi black magicians" who turned precisely into ``shadows of ancient kings" and then vanished into dust when I stroke them with my sword.


\hfill

\texttt{Balder on 2022-03-16 at 12:18 said: }

Correcting myself: Christ was of course not one of the fallen hierarchies, the inverted trinity is satans, ahrimans and lucifers. They all depict a spiritual impulse wrongly conceives and used for selfish ends, that is why they are written in the sources with small initials.

This is how I see those forces acting in man: 1) satans: power and violence; power for powers sake (``Might is Right" philosophy and Nietzshean will-to-power). 2) ahrimans: technocratic materialism. 3) lucifers: self-aggrandizement, narcissistic self-worship, ego-inflation, megalomania. To refer to LOTR again, fallen Saruman represents all of these in one person while Gandalf is of course the risen Christ Victorious.

temptation of the ahrimans: \url{https://www.youtube.com/watch?v=9PHxTEoVbjk}

temptation of the lucifers: \url{https://www.youtube.com/watch?v=KQKlh7UxaRQ}


\hfill

\texttt{Drew on 2022-03-20 at 20:25 said: }

So my problem is not that I cannot reason my way to faith but that I have fixated on reason. I've built a tower when I should have cultivated a garden.


\end{sffamily}\end{footnotesize}
