\section{Zabriskie Point}

\emph{Zabriskie Point}\footnote{\url{https://www.imdb.com/title/tt0066601/?ref_=nv_sr_srsg_0}} is a 1970 film by the famed director \textbf{Michelangelo Antonioni}. Hence, the film is about
Antonioni, which makes a complex plot and great acting unnecessary. Unfortunately, the only version that I had
available was a poor quality video I found on youtube. That did not do justice to the cinematography of the film, so I
was compelled to rely on imagination and memory.

The place Zabriskie Point is near Death Valley in California. It was once a lake, which dried up leaving the badlands.
It is not something that humans would normally see, which explains its alien yet compelling look. At one time, it was
home to camels, mastodons, and other fauna, all of which left their marks. Of course, there must have been humans whose
psychic residues probably still haunt the area.

A film should be an experience, not a story. There must be scenery to look at. The plot of the story must be secondary
to the inner lives of the characters as per Schopenhauer. Mark and Daria were a young couple whose paths crossed
momentarily in an unexpected, and impossible, way.

The film begins with a meeting of revolutionaries debating whether blacks or whites are more oppressed. Some 70 years
later, the argument continues even though the revolutionaries are now in total control. Mark is not impressed; he
claims he is willing to die, but not of boredom. He then embarks on an adventure that few would describe as boring.

In a single day, he gets arrested, steals a Cessna 210 airplane, flies over the desert, spots Daria on the highway and
``picks her up" by impressing her with airplane tricks. They make love in the dunes of Zabriskie Point, paint the
aircraft with brightly colored slogans and designs; he then flies back to the Los Angeles airport where the police
shoot him to death as he landed the craft.

Most men would need a week to accomplish all that. Actually, on second thought, most men might fantasize bout that for
years, expect for the part about dying in a hail of bullets.

Obviously, it is not at all an action flick, so we do learn about their inner lives. Daria if clearly infatuated by
Mark's rebel without a cause persona. He tells her about revolutionary philosophy that divides
people into heroes and villains. When she questions such a sharp distinction, he explains that it the only way to tell
them apart. He was wise beyond his years and understood instinctively the friend-enemy distinction described by Carl
Schmitt. Even today there are the Darias seeking peaceful resolutions and the Marks advocating for unpeaceful
revolutions.

Mark tells her that they are all alone on the lake floor. She soon yields to his seduction as he slowly unbuttons her
dress. Suddenly, dozens of other amorous couples appear. They are the primordial couple but centuries of human
visitation are revealed to them. As a side note, the authorities investigated the film. It was a federal crime to bring
women across state lines for the purpose of sex; it was, and might still be, a state crime to have an orgy at a state
park. So, guys, if you ever find yourself in such a situation, just get a room. Antonioni was cleared because the
scenes were simulated.

Daria hears of Mark's execution on the radio, but she continues her trip to Phoenix to meet up with
her boss Lee, at his awesome desert home. Lee is a real estate developer whose current project is to create a desert
paradise with a lake; but has trouble attracting investors. Little did they know what Phoenix would turn into.

Daria finds her life emptied of meaning, her new love lost forever, and unable to return to the past. She drives off,
turns around, and fantasizes Lee's home exploding and bursting into flames.

Who is living in an illusion, Daria or Lee? Whose world will end? As Rene Guenon concludes:

\begin{quotex}
It can be said in all truth that the end of a world never is and never can be anything but the end of an illusion.

\end{quotex}

\url{https://www.youtube.com/watch?v=Lb1h4RsmIM0}



\flrightit{Posted on 2022-10-09 by Cologero }
