\section{The Unforgiven and unforgivable}

The Unforgiven was a movie released in 1960. Directed by John Huston, and starring Burt Lancaster and Audrey Hepburn, one should expect a superior film. That depends. On the one hand, it could be a history of how the West was won, in its raw brutality, or it could be an allegory about racism and why the West should not have been won.

The film begins in the middle of things, since the patriarch of the Zachary family, Ben, had been killed in a battle with the Kiowa Indian tribe. He left behind his wife, Matilda, his eldest son Ben (Burt Lancaster) as the acting pater familias, the hot-headed middle son Cash (Audie Murphy) and the lovable dim-witted Andy.

Finally, there is the mystery meat Rachel\footnote{\url{https://www.urbandictionary.com/define.php?term=Mystery\%20Meat}}, played by the unduly attractive Audrey Hepburn. The family legend is that Ben found the baby Rachel alive, after a raid by the Kiowa. He brought her home to the delight of Matilda who desired a daughter and raised her. Rachel, approaching her prime, was pretty and spunky. She also had an affection to Ben which bordered on the incestuous.

But then the mysterious Kelsey appears on the scene and, like the even more mysterious Khidr\footnote{\url{https://en.wikipedia.org/wiki/Khidr}}, reveals the true origin of Rachel to whoever is willing to listen. He even reaches the ear of Lost Bird, of the Kiowa tribe, who learns that Rachel is actually his sister.

What's fair is fair, so Lost Bird offers to buy back Rachel from the Zacharys for a few horses. Ben will have none of it. As the rumor spreads, Ben's business partner Zeb turns against the Zacharys, whom he blames for his son's murder.

At that time there was no DNA testing, so no one suspected that the very Belgian looking Rachel could possibly be a Kiowa. Cash became furious to learn that he had been living with an Indian all those years. He used a slur that can never be spoken aloud; like the secret name of Rome, it must be protected lest enemies use it to gain power.

In a huff, he leaves home to spend a night in the hay with the very blonde, very pretty, and very horny, Georgia. Before Tinder, it was risky to swipe left because there were few eligible men within a reasonable radius. So Georgia was desperate to snag the manly Cash, offering to him man's supreme pleasure, right on the spot. That left him with the moral dilemma: stay with Georgia or return home to protect his family.

Meanwhile, Lost Bird returns to the Zachary homestead, with a bunch of his bros. Ban hunkers down and prepares for battle. He even convinces Rachel to shoot at her own tribe. After an extended battle, the situation looks gloomy for the Zacharys. Meanwhile, still in the hay with Georgia, Clay hears gunshots from 10 miles (16 km) away. To her disappointment, he gathers up ammo and weapons and heads home, now contrite from his earlier outbursts.

Unlike the John Wayne move, Stagecoach, in which the calvary shows up at the last moment to save the white people, Cash rides in firing his six-shooter all alone, thus dispersing the hated Kiowa. Lest that sound implausible, Audie Murphy was a legitimate real-life hero in World War II. He was awarded several medals from three different countries. This was his best:

\begin{quotex}
Murphy received the Medal of Honor for valor that he demonstrated at the age of 19 for single-handedly holding off a company of German soldiers for an hour at the Colmar Pocket in France in January 1945, then leading a successful counterattack while wounded and out of ammunition.
\flright{\textit{Wikipedia} \url{https://en.wikipedia.org/wiki/Audie\_Murphy}}
\end{quotex}

In a touching scene, Rachel is confronted by her blood brother Lost Bird. After a few moments of thought, she shoots him in the belly in cold blood. As Ben had convinced her, although she was a Kiowa by blood, she was Anglo in every other way, thus settling the nature vs nurture debate.

Finally at peace, Ben makes plans to marry his sister Rachel. After all, they aren't blood relatives and she was never legally adopted.

\url{https://www.youtube.com/watch?v=UHUa\_svkuU0} 

\flrightit{Posted on 2022-01-26 by Cologero}

