\section{Gone, Baby, Gone: Man's Justice}

\label{sec:GoneBabyGone}

This is an ancient conundrum:

\begin{itemize}
\item Is a command just because God commanded it? 
\item Or did God command it because it was just? 
\end{itemize}
On the human level, governments establish a set of laws and policies in order to achieve just decisions. This involves a complex bureaucracy with established procedures. The conundrum can be rephrased:

\begin{itemize}
\item Is a result just because the procedures have been faithfully followed? 
\item Or does following the procedures define justice? 
\end{itemize}
In the first case, God is justice, there is no distinction between the alternatives. However, in the human case, there can be a privation between the ideal of justice and its implementation in specific cases. For the purposes of this discussion, let us assume that the law is not inherently unjust.

\paragraph{Outline of the Story}
Amanda is the four year old daughter of the tamasic Helene, who is addicted to alcohol, cocaine, and sex. While she is out carousing for several hours, she returns home to discover that Amanda went missing. Since the police investigation is making no progress, Amanda's aunt Bea and uncle Lionel come to Patrick and Angie for help. The latter two specialize in finding deadbeat debtors, but not kidnap victims. Nevertheless, they take the case.

Since Patrick has buddies in the Bostonian underclass, who never talk with the police, he is able to extract information unknown to the detectives. He is even introduced to the malevolent underworld of pedophiles, now known as ``minor attracted persons". Leaving aside all the plot twists, misdirections, and machinations, we can cut to the chase.

Amanda's death was faked. However, eventually Patrick figures out what really happened. Those who loved Amanda conspired to protect her from her mother: Uncle Lionel often saved Amanda from her mother's neglect; Detective Remy was disgusted by his experience with abused children; the police captain always wanted a daughter.

Patrick visits the captain at his country home and spies Amanda; she seems healthy, happy, and loved. He has to make a choice:

\begin{itemize}
\item Leave Amanda with the Captain for her own good. 
\item Do the ``right" thing and report the kidnapping to the proper authorities. That would return Amanda to a sad life with her mother, and bring bad ends to the Uncle, the Detective, and the Captain. 
\end{itemize}
 

\paragraph{Mother}
The justice system has a process in place that almost always keeps the child with her biological mother. But what is a mother? Is she the birthing person or the one who desires what is best for the child?

Jochebed was willing to give up her son, Moses, in order to save him from death by the Pharaoh.

To resolve the dispute between two women who each claimed to be the mother of the son, the wise Samuel decided that the one who wanted to protect the baby was the true mother.

The justice system has no room for wisdom, since everything must follow the law. And Helene was not willing to provide Amanda with a safe, stable, and loving home life.

\paragraph{The Decision}
Patrick has to decide what to do.

Jochebed disobeyed the Pharaoh's law to protect Moses. But if everyone disobeys the law, then anarchy ensues.

Wisdom may have selected the new family as best for Amanda. But the justice system follows a process for which the question of Amanda's future is not considered.

The Captain warns the young Patrick that he will have to live with his decision for the rest of his life. In a display of bravado, Patrick denies concern. But the Captain was right.


\hfill

\url{https://www.youtube.com/watch?v=Wzh4Q88sthY}



\flrightit{Posted on 2022-01-20 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Matt on 2022-01-20 at 23:41 said: }

Is it even possible to make a just system? To be just, in many cases, would require a discernment from a moral being that no set of laws, no matter how complex, could possibly capture fully. But the more room for human discernment you place into a system, the more room for corruption. Too bad we can't just place God in charge of such things.


\hfill

\texttt{Arthur Konrad on 2022-01-21 at 05:44 said: }

There is a distinction between justice as an institution, and justice as a concept in philosophy, and once this distinction is examined, a question appears as to whether their subjects are related at all. In reduced language – the former is concerned chiefly with outcomes (or more precisely, ought to be). In theory, it is possible to practice law as a moral exercise – that's how law works today for the most part – but in that case a legal science is inconceivable.


\hfill

\texttt{Mike M on 2022-01-21 at 22:42 said: }

@Matt This came up in group, which I enjoyed your answer the best, and my question is, if in fact, humans cannot make a ``perfect" system in which the idea of justice is enacted in every circumstance (that is typically the definition I see of people calling something ``perfect"), is it true impossible? Because if it is not possible to manifest, is it truly the ideal at all? An analogy is how Christ did not cure suffering in the world because suffering is part of the human condition for a purpose of perfection in eternity. 

Perhaps what humanity does make out of a system of justice over time is the representation of a truly just system depending on their achievements to align with the idea of justice. If so, then mustn't we subject ourselves to the system to some degree? Hence I liked the answer given that one should do one's best to satisfy the system while at the same time pushing for the goal that is most ``just" intuitively, without breaking the rules. This would take personal responsibility for the man in the film to involve himself and not just pass the decision onwards either way. This would satisfy Gospel as well: ``Do not think that I am come to destroy the law, or the prophets. I am not come to destroy, but to fulfill." I believe a good example of this in film is Ikiru, in which a man with determination utilizes the system without breaking it to achieve a common good.


\hfill

\texttt{Logres on 2022-01-22 at 12:07 said: }

This conundrum reminds me of Tomberg's passages on ``gilded silver" or ``silvered gold" : if one is aware of the ``trap" (or varieties of possible outcomes for choices, and giving justice to all sides), there is a chance of seeing a way through (with courage and intuition) to a satisfactory conclusion, even if, the result itself is still slightly imperfect as Tomberg describes, or rather, has an emphasis on one element of either Justice or Mercy, slightly. It does occur to me, that the choice is not presented to ``everyone", but to an exception or a unique person. And it is not possible to ``refuse" it, except by abdicating morally. If the person rises to this exceptional element, there is a hope that their choice will neither ignore Mercy nor break Justice, but bring them closer together. However, most people will not even understand their dilemma.


\end{sffamily}\end{footnotesize}
