\section{Fallen Angel}

\begin{quotex}
\begin{verse}
We were born to tread the earth as angels,\\ 
To seek out heaven this side of the sky.\\ 
But they who race above shall stumble in the dark,\\ 
And fall from grace.\\ 
Then love alone can make the fallen angel rise.\\ 
For only two together can enter Paradise.
\end{verse}
\flright{\textsc{Marty Holland}}
\end{quotex}

\textit{Fallen Angel}\footnote{\url{https://www.imdb.com/title/tt0037691/?ref_=fn_al_tt_1}} was a 1945 film noir based on the novel by Marty Holland (nom de plume for Mary Holland). Like every film noir, there is some drama, a murder, a mystery, and ultimately a confession. You can watch the film if you like, or read the plot on Wikipedia. For now, I prefer to focus on the character archetypes, since they fit in so well with the so-called manosphere. Yet, as the film shows, there is nothing new under the sun.

\paragraph{The Chad}


Eric Stanton is a man down on his luck. He had sold his business in New York but lost it all gambling on his way to San Francisco. Unable to afford the last leg of his bus trick, he was forced to stop in the small town of Walton. Down to his last buck, he stops at a greasy spoon\footnote{\url{https://en.wikipedia.org/wiki/Greasy_spoon}} for a late meal.

Eric is what is called today the Chad, that is, the man desired by the most woman. He is tall, handsome, romantic, assertive, and a smooth talker. Nowadays, he would be in the top 10\% to 20\% of men who get all the attention on dating apps. For such guys, women often become disposable commodities, while the bottom 80\% of men get hardly any attention.

In a small town, however, the options are fewer. Nevertheless, Eric manages to seduce two women in a week, even proposing marriage to both. Since juggling two women won't end well, Eric finds himself involved in a situation beyond his control.

\paragraph{The Town Flirt}


While awaiting his meal at the greasy spoon, Stella suddenly waltzes in. He is immediately smitten. It turns out that he is not alone. She is the Town Flirt\footnote{\url{https://www.youtube.com/watch?v=AS4bqgSFWAw}}, a seductive dark-haired beauty with Siren-like qualities, turning heads, stopping men in their tracks, and insinuating her image into their private fantasy lives. Even Pops, the old guy owner of the dive, is not immune to her charms. Same for the rest of men in the town, since Stella has a different date every night.

Eric manages to get some quality time alone with Stella. The tension between the Chad and the Beauty is palpable. For once, Stella meets a man that she would like to be with. Eric pushes himself on her, but she parries his advances. Here is an example of the Hermetic teaching:

\begin{quotex}
Man wants to enjoy before possessing; woman wants to possess before enjoying.
\end{quotex}


Drunk with love like a rutting iguana, Eric even proposes marriage in order to enjoy her; he even promises to gather a nest egg to make her life comfortable.

The archetype of the Town Flirt is not uncommon. Stella tells Eric that she has had enough of riding the carousel\footnote{\url{https://notesfromaredpillgirl.com/2016/03/18/dont-ride-the-carousel/}}, and expresses her need to finally settle down. Men should beware of getting involved with a woman with a high body count.

Some attractive women date around in their late teens to late twenties. Eventually, they stop getting the same type of attention; perhaps they see their friends settle down or they get baby fever. Then they have to settle for what's available. Stella is juggling three marriage proposals; the jealousy that such a situation arouses will not turn out well for her.

\paragraph{The Alpha Widow}


Often a woman will meet a Chad and fall madly in love. Of course, he leaves her in a state called the Alpha Widow\footnote{\url{https://www.youtube.com/watch?v=6ClzNJglxVU}}. She becomes totally heartbroken by her boyfriend and the resulting sorrow can last a lifetime. Clara Mills is part of the upper crust of the town. Now an aging spinster, she reveals that she had once been madly in love with a Chad. So she recognizes the type and tries to protect her younger sister from his charms.

\paragraph{The Girl Next Door}


Eric plans to embezzle some money from June Mills by seducing her. June, a blonde of course, represents purity. She plays the organ at the local church, and immerses herself in the high culture of books and music.

Eric challenges her to live a real life, with real people, and real events, not second hand through books and music. She falls in love with him at a night picnic at the beach, eating hot dogs, as the waves break the silence in the background. She even agrees to elope with him.

Pure and innocent, June can only see the best in Eric, even when he is accused of murder. Unable to deal with her naivety at first, he eventually returns her love. That is when she recites the poem that is the epigraph to this review.

\url{https://www.youtube.com/watch?v=LIHcAkeSlP0}

\url{https://www.youtube.com/watch?v=6ClzNJglxVU}

\url{https://www.youtube.com/watch?v=\_y6WMHZv7pw}


\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Auld Wat on 2022-02-21 at 18:46 said:}

True love's the gift which God has given

To man alone beneath the heaven; 

It is not fantasy's hot fire, 

Whose wishes, soon as granted, fly; 

It liveth not in fierce desire, 

With dead desire it doth not die; 


It is the secret sympathy, 
The silver link, the silken tie, 

Which heart to heart, and mind to mind, 

In body and in soul can bind.
\end{sffamily}
\end{footnotesize}
