\section{The Secret Garden}

\begin{quotex}
To know is to remember: let us therefore remember together, immortal Souls, who on earth sigh for Man's celestial kingdom, and desire the divinity of life. 
\flright{\textsc{Saint-Yves d'Alveydre}}
\end{quotex}

\paragraph{Poem by Soloviev}
\begin{verse}
Dear Friend, don't you see\\
That all we see is only\\
A reflection and shadow\\
Of what cannot be seen?
Dear friend, don't you hear\\
That this discordant roar is\\
Only a distorted echo\\
Of joyous harmonies?
\end{verse}

\paragraph{Living by one's choices}

A man lives or dies by his decisions. He accepts the rewards or consequences, and neither rejoices nor cries about the outcome.

\paragraph{Secret Thoughts}

He is out every day, as is she, doing her chores. She looks longingly thinking he would never accept her. He stares wondering how to approach her. Her secret thoughts, the walls are high to protect them.

She: I hear the birds sing of love, too, I look at the cloudless sky and watch the shepherd guide his herd into the neighbor's garden, I feel the sun, the passion and I can't stop thinking of you. You are everywhere.

\paragraph{World Without Love}

Too often, one fantasizes about the circumstances --- time, place, etc. --- where love can be found. The one fails to seize the opportunity when it presents itself, since it does not resemble the fantasy.

In the novel \emph{Zanoni}, Mejnour and Zanoni were initiates who knew the secret to eternal life. Mejnour was living forever, although he was nobody in particular, moving around to no place special. But his student, Zanoni, fell in love with a mortal woman. He then had to make the choice to live at a particular time and place with one special woman. For that one love, he gave up his immortality.

The lesson is that one moment of love is preferable to eternity without love. How many people live a life without love? More than you know.

\paragraph{Living with Mystery}
\begin{quotex}
All that He wrought among men, all that He said and suffered, He disposed in such wise that not a single moment was passed without mystery, not a single letter was devoid of some mystery.
\flright{\textsc{St. Bernard}, \emph{Sermon III on Palm Sunday}}\end{quotex}

\paragraph{Sacred History}
\begin{quotationx}
Historical facts conform to the law of correspondence, and thereby translate higher realities, of which they are a human expression. It is this that gives to these facts their significance. This symbolical character is bound to be particularly clearcut in the case of events connected with sacred history; thus it is recognizable in a most striking way, in all the circumstances of the life of Christ.

If the foregoing has been properly grasped, it will at once be apparent not only that there is no reason for denying the reality of these events and treating them as mere myths, but on the contrary that these events had to be such as they were, and could not have been otherwise; it is clearly impossible to attribute a sacred character to something devoid of all transcendent significance. \flright{\textsc{Rene Guenon}, \emph{Symbolism of the Cross}}
\end{quotationx}

\paragraph{Changing the Past}

The demon appeared to three monks and He told them: if I gave them power to change something from the past, what would they change?

The first of them, with great apostolic fervor, replied: ``I would prevent you from making Adam and Eve fall into sin so that humanity could not turn away from God.''

The second, a man full of mercy, said to him: ``I would prevent you from God and you will condemn yourself eternally''.

The third of them was the simplest and, instead of responding to the tempter, he got on his knees, made the sign of the cross and prayed saying: ``Lord, free me from the temptation of what could be and was not''.

The devil, giving a raucous cry and shuddering with pain, vanished.

The other two, surprised, said to him: ``Brother, why have you reacted like this?''

He replied: ``First: we must NEVER dialogue with the enemy. Second: NOBODY in the world has the power to change the past. Third: Satan's INTEREST was not to prove our virtue, but to trap us in the past, so that we neglect the present, the only time in that God gives us His grace and we can cooperate with it to fulfill His will ``.

Of all the demons, the one that catches the most men and prevents them from being happy is that of ``What could have been and was not''.

The past is left to the mercy of God and the future to his Providence. Just the present it's in our hands. ``Live today''

\url{https://www.youtube.com/watch?v=jR1gerSNTYk}

\flrightit{Posted on 2022-06-12 by Cologero}

