\section{Artificial Intelligence is better than none}

A reader recently pointed out that the AI tool ChatGPT\footnote{\url{https://openai.com/blog/chatgpt}} knows about Gornahoor and provides a not unreasonable evaluation of it. Actually, it is better than my family and most of my friends, who ignore me altogether. At this point, Chat AI technology is, at best, only capable of paraphrasing text that already exists. This is ChatGPT's summary.

\begin{quotationx}
Gornahoor is a blog that explores traditional metaphysics and the perennial philosophy, particularly in the context of the Western esoteric and spiritual traditions. The blog features articles and essays written by various authors, covering topics such as traditionalism, spirituality, philosophy, history, and culture.

The blog emphasizes the importance of the traditional worldview, which holds that there are timeless principles and truths that are accessible to human beings through spiritual and intellectual development. This perspective contrasts with the modernist and postmodernist worldviews, which often reject traditional knowledge and emphasize individualism, relativism, and materialism.

The authors on Gornahoor draw on a variety of sources, including the works of traditionalist writers like Rene Guenon and Frithjof Schuon, as well as the esoteric and occult traditions of the West, such as Hermeticism, Kabbalah, and alchemy. The blog also engages with contemporary cultural and political issues, applying traditional insights to current events and trends.

Overall, Gornahoor provides a platform for exploring the perennial philosophy and traditional metaphysics, and offers a perspective that challenges the dominant materialist and relativist assumptions of modern Western culture.
\end{quotationx}

In contrast, the description from Bing AI Chat\footnote{\url{https://www.bing.com/search?q=Bing+AI&showconv=1&FORM=hpcodx}} is more terse, yet not incorrect.

\begin{quotex}
Gornahoor is a blog that explores metaphysics, physics, religion and high culture, connecting Ancient Wisdom to the modern world. The blog covers topics such as the religion of medieval Europe, metaphysics, spiritual knighthood, sophiology and the common task, romantic Christianity, high culture.
\end{quotex}


\flrightit{Posted on 2023-03-27 by Cologero}


