\section{Nana’s Sauce}

\begin{quotex}
Sono poeta. Que faccio? Scrivo.
\end{quotex}

All the recipe web sites promote complex recipes where simplicity should suffice. They also tend to ramble on. I can ramble, but the recipe will be simple. Peasant food does not depend on a well-stocked kitchen, but rather makes do with what is on hand. You've probably heard stories about the women slaving over ``Sunday gravy'' with all its ingredients. My mother, however, was an Anglophile, so we always had a roast on Sundays. Pasta with tomato sauce and meatballs was a midweek specialty.

As a master of negotiating the vagaries of the modern world, I have simplified the recipe. After a little practice, this can be put together about 20 minutes; cleanup is easy. You will love the result and wonder why you ever bothered with those heavy sauces. It is even halal.

\paragraph{Lonely Girls}

\begin{quotex}
Venus is the source of the power of taste. She governs the stomach whence emerges the ability and taste for eating and drinking.
\flright{\textsc{Picatrix}}
\end{quotex}

If you happen to be in the supermarket between 5 PM and 6 PM, you are likely to see a lonely girl or two. She is still in her daytime uniform as she shops for dinner, her face vacant. In many cases, she has, in her cart, a box of pasta, a bottled imitation Italian tomato sauce, cheap shelf-stable grated cheese, and a bottle of wine. The giveaway is the pie or cake from the bakery --- there is just one slice.

This is the nigredo stage. If she only knew how simple this is; maybe should could entice a man and being the albedo stage.

\paragraph{The Sauce}

Crops, such as the potato, maize, pumpkin, chocolate, vanilla, and tobacco, originated in the New World yet radically transformed the Old world. Imagine Ireland without potatoes, England without chips, Russia without good vodka, Belgium without chocolates, or dolce without vanilla. What did Europeans do with their hands before tobacco?

New World Tomatoes came to define Italian cuisine, at least in the South. So let's prepare the sauce first. I can reveal a personal secret: I cook it in a wok these days for obvious reasons.

\begin{itemize}

\item Chop up a small onion and sauté it in some olive oil.

\item When the onions are translucent, add two boxes of Pomi strained tomatoes, or 1.5 kg of something else.

\item You can add a little water to the emptied boxes, shake them a bit, and add to the sauce.

\item Add some ground black pepper and some fresh or dried basil.

\item Finally add some wine, such as Marsala. This serves two purposes. It is better than adding sugar to remove some bitterness from the tomato. Moreover, just as vodka is added to penne alla vodka, the alcohol brings out more of the tomato essence. The alcohol will boil off, if that is a concern.

\item Bring it to a light simmer.
\end{itemize}

\paragraph{Dual Use}

The sauce can be used as a marinara sauce, say for use with ravioli or gnocchi. Add some oregano if you want to make pizza. Slicing a baguette in half makes a quick pizza like snack. Toast it a little, then cover with some sauce, mozzarella and grated cheese. Bake until cheese melts. Kids will eat it up.

\paragraph{Meatballs}

Simmering some meatballs in the sauce will continue the alchemical process.

\subsubsection{Ingredients}

\begin{itemize}

\item 1 pound or 500 grams of quality ground beef such as sirloin. To make them more tender, you can use 25\% of ground lamb and/or ground veal. However, if Bambi made you cry, the skip this step. Don't use ground pork; try to maintain some dignity.

\item 2 large eggs, beaten. Euro eggs are slightly larger than American eggs so don't mix American and metric measures. No one has explained the reason to me.

\item Add some ground black pepper and salt if you like it. For me, the cheese is salty enough

\item Locatelli Pecorino Romano Cheese: ½ cup or 125 ml. Don't go cheap on the cheese.

\item ¼ cup or 60 ml of plain bread crumbs. A nice alternative is a slice of white sandwich bread, without the crust, cubed into small pieces and then soaked in a little water for a few minute.
\end{itemize}

\subsubsection{The Process}

It should have taken you perhaps 5 minutes to gather the ingredients. Now you need to mix them all together by hand. Clean your hands thoroughly or use food gloves; don't get nail polish on the meat.

Stop when the mixture is homogeneous; do not overmix the ingredients.

Now roll the meat mixture into balls and drop into the simmering sauce. This will make 12 American or 10 metric meatballs.

Cover the pot or wok and let everything simmer for about 90 minutes or two hours. Stir occasionally with some flair and sing along as you do so. If people are around, pretend to taste it and add a little more spice with great fanfare. Be sure to skim excess oil and fat off the top.

If you concentrated as you cooked, you may get a taste of the rubedo stage, red like the tomato!

\paragraph{Service}

You've made enough sauce for 1 pound or 500 grams of your favorite pasta. Cook it al dente; don't embarrass yourself. Drain the pasta and mix it with some sauce.

Serve the pasta with some extra sauce on the side, which I like to do. Sprinkle liberally with some Pecorino cheese until the plate looks like the peaks of Mount Zalmon.


\flrightit{Posted on 2023-03-30 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{taivasvaeltaja on 2023-03-31 at 09:36 said:}

Sounds very tasty and elegant. Just ate some mama's meatballs with an oatmeal-rice mix. I could add or suggest a small change into the meatball recipe to make it more nutritious, filling and tasty: instead of bread crumbs or flour, use oatmeal instead. And if the recipe says two eggs, use at least three.

\hfill

\texttt{Cologero on 2023-03-31 at 14:58 said:}

There is no one perfect sauce, taivasvaeltaja. You can use your own creativity to enhance it in a way that seems suitable to you.

\hfill

\texttt{taivasvaeltaja on 2023-04-03 at 15:42 said:}

May I then perhaps present my classic tomato sauce?

Cook 500 grams of minced beef\\ Add some salt, black pepper, white pepper, oregano and basil according to your taste\\ Add 2 sliced onions and 6 nails of minced garlic and let it cook\\ Add 1 can of kidney beans\\ Add 5 sliced tomatoes and one can of tomato sauce\\ You can also add ketchup according to your taste

Cook middle heat for 15-20 minutes and roll it up now and then. Don't forget to cook the pasta with it. I also like cooked potatoes, carrots and broccoli.

\hfill

\texttt{Cologero on 2023-04-03 at 16:08 said:}

I had a girlfriend who added ketchup to her tomato sauce, unknownst to me. After I tasted a bite, I put down my fork and walked out of her house.

\hfill

\texttt{taivasvaeltaja on 2023-04-04 at 08:26 said:}

Haha! I knew the ketchup part was controversial. I forgot to add that it is only allowed little and when the label is Heinz.

\end{sffamily}
\end{footnotesize}

