\section{Statistics Ruin the Game}

At that time, my father, mother, sister, and I were playing a family game together. I don't recall the game, but I was scorekeeper. Of course, I started to calculate various statistics and announced them at the end of each round.

I could see my father squirming and looking upset. Finally, after one round, he snapped at me, ``You're ruining the game!'' That ended the computations, both then, and for the future.

Perhaps he was expecting some type of Victorian era family chit chat. But we knew each other too well to engages in small talk and I was probably the only one capable of a more stimulating discussion. Even at that age, I knew elementary science, history, and geography. I used to read the encyclopedia and almanac; every day, I would read the all the sports stats in the newspaper.

I did learn when to be quiet, but boys are going to count. I still count even if I don't let people know.

\paragraph{Fast Forward}

If you fast forward to today, you will see the domination of statistics even if few understand them. The biggest adoption is in the sports world. I have always been a sports fan, so I've seen how Sports Analytics\footnote{\url{https://en.wikipedia.org/wiki/Sports\_analytics}} have come to dominate in the past decade or so.

In the past, for example, a basketball coach would want the ball passed to the player with ``hot hand''. Now the pass goes to the player with the opportunity for the highest percentage shot. It's funny to hear sportscasters talk about ``sample sizes'' even if they had never heard the term until 20 minutes ago.

The downside is that, with computers, they repeat all sorts of statistics that have no bearing on the game. The old story about the ignorant using technology.

As Galileo noted, ``the language of science is mathematics.'' But this is true for all manifestation. When writing a sonnet, the poet counts lines and syllables in each line. In a good sonnet, the counting will not intrude.

Boys will count.


\flrightit{Posted on 2023-02-25 by Cologero}

