\section{Simulation and Yellow Cards}

At that time we had just moved to the suburbs. There I learned of an abandoned crabapple orchard which exist nowhere in the city. At harvest time boys would often go there and engage in a war of all against all, using crabapples as weaponized projectiles. They could really sting, but the potential joy of inflicting pain overcame the aversion to experiencing pain oneself.

Jacques got hit hard by one such globose pome. To my surprise, he started bawling like a baby. Since he was past the age when it was appropriate for boys to cry, I was confused; perhaps it was a defect of country folk. Nevertheless, since he was in my sixth grade class, I felt an obligation to comfort him. But then, I was even more surprised when he suddenly stopped. He said to himself aloud, ``Where is Mrs. F.? Doesn't she hear me crying?''

So he was just trying to get attention from some mother on the street, whom he apparently knew. Perhaps she knew his ruses or was reluctant to go outside during the war, but she never appeared.

That was when I learned about ``simulation'' for which he should have received a yellow card. Now I never assume that outward emotional manifestations necessarily correspond to inner states. Jacques was never a playmate after that incident.


\flrightit{Posted on 2023-04-11 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{taivasvaeltaja on 2023-04-19 at 02:38 said:}

Perhaps he didn't get hugs at home from his mother? From experience I can say that children usually want their mother to comfort them; it's written in our biological make up. Men and fathers just don't suffice, we're not built for it and the children know it.

\hfill

\texttt{Logres on 2023-05-20 at 12:17 said:}

In case the news has missed anyone here not on lists.

I am grieved to have the duty of letting the bees know that their beekeeper has passed away, peacefully, after a month long battle with Covid and attendant complications. Joel Salvo his son was kind to reach out. He passed on Ascension Eve. The website will be maintained by Mr Joel Salvo. If I receive a funeral date and time I will add it here. The beautiful old man is dead to this world but is with the Resurrection and the Life. Long long indeed, shall he live. Had thought of canceling meetings in his honor. But some may wish to commemorate him with attendance and reminiscing. Either way, but he would wish us carrying on, as the Spirit leads.

\hfill

\texttt{Arthur Konrad on 2023-05-20 at 13:28 said:}

I'm very saddened by the news. Cologero was inimitable and he will be greatly missed.

\hfill

\texttt{Candido Catão on 2023-05-20 at 23:15 said:}

This was the last article I read by C.S while he was still physically present in our world. It was a great honour and privilege for Humanity to have had the opportunity for such a valuable soul as like his to have shared great teachings with us.

\hfill

\texttt{Ismo Meinander on 2023-05-21 at 01:08 said:}

R.I.P. Cologero. You were a treasure island of knowledge and wisdom, and you'll be greatly missed.

\hfill

\texttt{Juliano on 2023-05-21 at 06:18 said:}

Rest in peace Cologero, you will be greatly missed.

\hfill

\texttt{Fergus on 2023-05-21 at 12:59 said:}

Rest in peace Cologero.

\hfill

\texttt{Noct on 2023-05-21 at 13:03 said:}

His writings have been a great contribution in the formation of men called by the Spirit. We could have differences, but you have been a great contribution in my own way and in that of many others. May the grace of the Lord be with you.

\hfill

\texttt{Santiago on 2023-05-21 at 17:51 said:}

Thank you for everything.

\hfill

\texttt{PS on 2023-05-21 at 21:26 said:}

Thank you for everything, Cologero.

\hfill

\texttt{Sunrise on 2023-05-22 at 03:04 said:}

Thank you for sharing your wisdom with us, Cologero.

You will be greatly missed. RIP.

\hfill

\texttt{De Bosit on 2023-05-22 at 10:32 said:}

I hope he is in Islamic paradise now.

I will miss his posts.

\hfill

\texttt{Avery on 2023-05-23 at 13:55 said:}

``By giving birth to true virtue and nourishing it, he would be able to become a friend of divinity; and if ever a human being could become immortal, he could.''

\hfill

\texttt{Sullen Eyes on 2023-05-23 at 17:50 said:}

Forgive me for never compensating you in any way for the hundreds of articles you gave away for free. Farewell, Cologero

\hfill

\texttt{T. Freitas on 2023-05-23 at 22:05 said:}

Rest in Peace, C.

I learned a lot from your writings. Thank you very much.

\hfill

\texttt{Dimitri on 2023-05-25 at 11:05 said:}

Rest in Peace, Cologero. The work you've done here has been indispensable and invaluable to so many. Thank you for carrying on the torch of Tradition, rest assured your memory and mission will carry on. God bless all here at Gornahoor.

\hfill

\texttt{Aria Veritas on 2023-06-03 at 16:57 said:}

This post was for me I believe. He was acknowledging that he'd stepped over the line and knew I was taking his request seriously- I wasn't coming back out of the house until he knocked politely.\\ I first heard of Charles' death yesterday through the O.I. newsletter.

For the last fortnight I've been expecting an email from him every day because he's been so strongly about, but decided to completely obey him for once and `leave him be,' which was the last thing he wrote to me on April 4th. I wrote back, `Consider it done my friend,' on the 5th.\\ One month is by far the longest break we'd had since correspondence began.

I needed time to write and he didn't altogether approve of my themes; he thought it almost betrayed the church and his beliefs. But I must write what I see as true of course, intending to finish a few posts before contacting him again to explain my position in detail (through said posts).\\ Several times I thought something had happened but just as often overrode the feeling, figured he was messing with me (been hit by a globose pome, haha..) as he'd done a few times in the past.

He had told me of a dream about his father around mid-March. His father had asked him to sign a document or pay a fine to stay. He decided to pay the fine. We both suspected he was about to either jump or die but didn't voice these possibilities. That was what Gornahoor was about anyway- and he knew I had low tolerance for fears of death.\\ I'd also dreamt about a similar situation a few months prior, beautifully situated and surrounded by friends. But it was still rather a bell tolling in the distance and we both heard it, wishing we hadn't.

We were writing to each other for over a year. He found me wandering about in the worlds he described on this site back in 2021, and his (usually) charming and unexpected attachment enabled me to see God face to face during the 2022 Burning of the Ram, making universal sense of literally everything to my profound surprise and utmost gratitude.

\hfill

\texttt{MasterRaider on 2023-06-05 at 05:56 said:}

May god embrace him as a friend, what a brilliant and great man he was. Rest in glory my brother, thank you for what you so generously taught us all these years.

\hfill

\texttt{tom on 2023-06-05 at 11:01 said:}

\url{https://www.godsongs.net/2021/09/celtic-song-of-farewell-may-choirs-of-angels.html}


\hfill

\texttt{Max on 2023-06-05 at 15:13 said:}

Thank you, Cologero.

\hfill

\texttt{Verity on 2023-06-06 at 02:54 said:}

Thank you for everything. Eternal respect.

\hfill

\texttt{quantum lament on 2023-06-10 at 09:12 said:}

Thank you for everything, Cologero.

\url{https://youtu.be/CxCreGRH61M}


\hfill

\texttt{Juliano on 2023-06-13 at 11:01 said:}

Anyone knows about Cologero's translation of Man and His Becoming? I hope this important work is not lost.

\hfill

\texttt{DA on 2023-06-16 at 16:12 said:}

The work is finished, God called for his servant. Rest in peace.

\hfill

\texttt{J on 2023-06-17 at 23:56 said:}

I'm very saddened to hear this. I'm thankful to have been able to read Cologero's posts for many years now. It truly was a gift.

Rest In Peace Cologero.

\hfill

\texttt{InIlloTempore on 2023-08-10 at 08:44 said:}

First of all, Cologero, thank you for opening for us the fountain of Tradition. That light which some wanderers have been seeking for so long. Before arriving this site, I believed that there was no hope of quenching that odd thirst of depth... Rest in peace, beautiful old man.

I pose a couple of questions to those who are still around.

First, I am making slowly my way along the posts of this site. That arises the feeling of compiling and ordering them in a manageable format for study and meditation. The purpose is that compiling becomes itself a way of meditation and inner work. The task is vast, since the amount of content may fill at least ten books. The pont is that some of the readers of this site may be interested in sharing such task.

In particular, the project would be collecting posts in pdf's according to main issues: philosophy, metaphysics, translations, reviews, the Ancient City and family values, practical work, etc; and trying to track a path within the collection to make its inner thought a bit more apparent. Edition can be rendered via LaTeX or similar. It is worth the try, to give proper design to such a excelling content.

Needless to say, such project is to be done between friends. I mean, the final output is just to be shared between those interested. Only spiritual profit intended.

\hfill

\texttt{InIlloTempore on 2023-08-10 at 08:46 said:}

Second, I read that there were forums, mail lists, even a study group to put into practice inner work. Have they been discontinued? If not, how can anyone join them?

PS: I beg your pardon for any misunderstanding above, English is not my native tongue.

\hfill

\texttt{Cae on 2023-09-29 at 19:35 said:}

I'm so very saddened to hear of Cologero's passing... .He was truly a source of light, and I have deeply appreciated the wisdom of his writings here.

May it be that he has been welcomed into Heaven -- Amen.

\hfill

\texttt{Woody on 2023-09-30 at 09:40 said:}

Blessed repose and eternal memory. Christ is risen.

\hfill

\texttt{Presbyter Iohannes on 2023-11-25 at 11:59 said:}

AVM

Only now I came back here and read the news. I'm heavy hearted now. I will be sure to include him in my prayers and urge everyone who was a regular here to do the same. The work conducted by this man played a significant part in transforming me for ever. Thank you, Cologero. Condolences to his family and may he find the way of union with his heavenly archetype, for Eternal rest and Glory in the Absolute.

And for those who still dwell in here. Memento Mori. Let us die before we die, so that when we die, we don't die; and may we have the grace of meeting him and all of those who might be worthy in the Last Day.

Requiem aeternam dona ei, Domine, et lux perpetua luceat ei.

Requiescat in pace.

Amen.

\hfill

\texttt{William on 2023-12-10 at 11:20 said:}

Hello,

As I understand it, from now on Mr. Joel Salvo will be taking care of the site. Does anyone know how can I get in contact with him?

Thank you in advance

\hfill

\texttt{Jean M.M. Bernard on 2025-11-21 at 08:05 said:}

I had, regrettably, been out of touch with this site and its mastermind for a while, and did not learn about Cologero's passing away before today. I was compelled to check in here after Cologero seemed to appear to me in a dream this morning, communicating certain things which included giving me some kind of old clock in purple. He said that he loves all of you: emphasis on all.

May God bless him.

\end{sffamily}
\end{footnotesize}

