\section{Initiatic Paths}

A couple of weeks ago, we asked about the available options for initiation for Western men today. Although the first two options got the most votes, the third one seems to have generated the most interest.

\paragraph{Read Dante's Divine Comedy in Italian}

Men attracted to Tradition today are, or should be, interested in the idea of initiation, that is, how to begin on a spiritual path that goes beyond mere book knowledge. However, in our situation we cannot just let George to it\footnote{\url{http://en.wikipedia.org/wiki/Let\_George\_Do\_It\_(radio)}}. Now, Rene Guenon, who certainly has read a great deal of initiatic texts, asserts that the two Medieval poems -- \textbf{Dante}'s \textit{Divine Comedy} and \textit{The Romance of the Rose} -- are of the same quality. Where then are those who claim to be interested in a Western initiation?

There are three obstacles. So called Christians in our day are embarrassed by or indifferent to their spiritual masters, and prefer a watered-down version. The most disheartening are the self-loathing Europeans who reject anything from that period in the name of some fictitious past. They regard that period as an unfortunate interlude from a glorious past, rather than as the development and deepening understanding of that past.

The real reason may be that there are no men capable of understanding such texts. The Comedy represents a long symbolic journey incorporating ancient paganism, Medieval philosophy, and a deep metaphysic. The developmental stages are described in detail. The obstacles to success are enumerated. It also incorporates a true eroticism with the vision of Beatrice. Dante was influenced by the Sicilian poets who, in turn, had incorporated erotic elements from the Troubadours of Provence and Islamic poets.

The rehabilitation of these texts should be among the prime concerns for those interested in the Western Tradition.

\paragraph{Meditate in a Cave in Tibet}

This turn to the East is often made today, whether to some school of Hinduism or some form of Buddhism. This requires an understanding of linguistic elements much more alien than Medieval Italian or Latin. Furthermore, cultural accretions get mixed up with genuine teaching, especially since the former are much easier to adopt. It reminds me of \textbf{Ananda Coomaraswamy}'s turn to Tradition after he noticed Indian women dressing like Europeans in his native Ceylon.

Dressing up like a swami or a Tibetan monk does not make you such. In our day, one should know several people who have turned to the East. What have they gained beyond a new set of dogmas? Please observe such people with detachment before choosing this route. If you don't know anyone personally, just check out some of the most influential Westerners associated with Tibetan Buddhism.

\paragraph{Practice Sex Magic}
This certainly has more appeal than the first two choices. It has the advantage of being egalitarian and democratic. Unlike reading long poems or spending hours in meditation, sex (and drugs) can be done by anyone. However, we suggested that references to sexual union should be taken as symbolic\footnote{\url{https://gornahoor.net/?p=3767}}, just as we do with alchemical texts. We also suggested that, if it is not merely symbolic, then it is reserved to those who don't desire sex\footnote{\url{https://gornahoor.net/?p=51}} or treat it instrumentally rather than erotically (e.g. Crowley). We even referenced \textbf{Miguel Serrano\footnote{\url{https://gornahoor.net/?p=70}}}, as the expositor of the highest, i.e., non-physical, Tantra teachings.

These references bring little interest. As the contrarian, \textbf{Julius Evola} brings in another view. In his essay ``Libertà del sesso e libertà dal sesso'' (Sexual freedom and freedom from sex), included in \emph{L'Arco e la Clava}, he writes:

\begin{quotex}
We affirm that in every complete traditional civilization ... the ascetic and warrior values were able to be cultivated along with erotic values, with reference to different paths and different vocations. So for example in India we find a ``path of desire'' (kama-marga) along with that of knowledge (vidya-marga), that of ascesis (tapas-marga) and that of action (karma-marga). India itself, as well as China and Islam, show us the co-existence of an androcratic regime, that is, supported on masculine supremacy, with a very highly developed erotic life.
\end{quotex}

Although this is not quite right, since the three paths of renunciation are those of knowledge, devotion, and action, Evola introduces here a major blunder. The path of desire is not considered a way of liberation or deliverance. To the contrary, it is the path of the ordinary man; such a man, whose mind and will are dominated by desire, ought to fulfill such desires rather than go against his own nature. On the other hand, the man who sees the emptiness in the pursuit of sexual desire, wealth, and fame is able to move on to one of the paths that lead to deliverance. This blunder renders moot everything Evola writes on this topic. For example, in the \emph{Metaphysics of Sex}:

\begin{quotex}
The ontological principles of Shiva and Shakti or of other equivalent deities present in the bodies of the man and woman are realized; it is necessary to arrive ritually and sacramentally at a state where the man actually feels himself to be Shiva and the young woman Shakti. This is the precondition of a sexual union that must transcend the limitations of physical and carnal nature and assume a magical character. Its center of gravity must be transferred to a subtle plane where the mutual magnetism of intense sexual love and desire consists in an intoxication or congestion of astral light. Its climax and supreme ecstasy correspond to the rupture of level of individual consciousness and the sudden realization of the nondual state. This is the ultimate goal that path of desire or kama-marga which has been considered in Hindu tradition alongside the other paths of liberation ... 
\end{quotex}

This is the heart of the matter: the path of desire is not considered one of the paths of liberation. In any event, even this paragraph confirms what we have written in the comments section. The ``center of gravity'' must transcend the merely physical and carnal act in itself. As a practical matter, this requires the man to remain conscious and detached during the ``sacrament'', \emph{which may make it magical at the cost of destroying its entertainment value as a sexual act}.


\flrightit{Posted on 2012-02-29 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{logres on 2012-02-29 at 23:52 said:}

Option 1 is unlikely, but least so.

\hfill

\texttt{Cologero on 2012-03-01 at 03:39 said:}

So the modern worldview is ineradicable and there is no return?

As for option 3, it seems more likely than not: \textit{Yoga and Sex Scandals}\footnote{\url{http://www.nytimes.com/2012/02/28/health/nutrition/yoga-fans-sexual-flames-and-predictably-plenty-of-scandal.html?\_r=1}}

\hfill

\texttt{logres on 2012-03-01 at 08:32 said:}

Perish the thought. I must try -- I've marked the two works for purchase. How important is it to have a parallel Italian translation? I know a little German and a little Latin, but Italian might be too much to learn quickly. In fact, I know it is.

\hfill

\texttt{logres on 2012-03-01 at 08:38 said:}

I think I see what you mean, looking at the article -- the ``nobler'' or higher path (and therefore, the more traditional one) would be to pursue Option 1. Option 3 is more ``unlikely'' in terms of preserving all of what we wish to preserve. It would be a society increasingly defined by casual sex, enjoying an uneasy truce with modern practice. It's one thing to come to traditionalism while already engaged in certain practice, and then implement a change of context, and another to begin to be drawn to the practices after one already understands Tradition, as a means of synthesis?

\hfill

\texttt{Cologero on 2012-03-01 at 12:35 said:}

No, the parallel translation is nice, but not necessary. The language is a secondary issue. The real issue is knowing enough Virgil, Aristotle, and Aquinas, inter alia, to get drawn into the world view. But who has 40 years to meditate on the Divine Comedy?

\hfill

\end{sffamily}
\end{footnotesize}

