\section{The Knight as a Bibliophile}

This post is In response to a ``challenge" on to name 10 influential books. The primary criterion used in this selection is that the book provoked a fundamental change in world view. That is why I have seldom been able to reread any of these books, since a change like that is not repeatable.

\paragraph{Dead White European Males}
My parents had a collection of classic fiction in the house, so I was able to read some of the greatest pre-twentieth century European novelists. Among them was \textbf{Rudyard Kipling} (\emph{Captains Courageous}, \emph{Kim}), \textbf{Mark Twain}, \textbf{Nathaniel Hawthorne}, \textbf{Robert Louis Stevenson} (\emph{Treasure Island}, \emph{Kidnapped}), \textbf{Jack London}, \textbf{Daniel Defoe} (\emph{Robinson Crusoe}), \textbf{Jules Verne}, \emph{Alexandre Dumas} (\emph{Three Musketeers}, \emph{Count of Monte Cristo}), \textbf{Herman Melville} (\emph{Moby Dick}), \textbf{Joseph Conrad}. My favorite was \emph{Ivanhoe} by \textbf{Walter Scott}. As a whole, they provided the foundation for the worldview of a white, European male. They were about Heros, Knights, Chivalry, Conquest, Ingenuity.

\textbf{James Fenimore Cooper} (\emph{Deerslayer}, \emph{Last of the Mohicans}) provided the image of the solitary knight\footnote{\url{http://www.claremont.org/article/in-defense-of-cowboy-culture/\#.VA0tSJXTCAg}} in the early American setting. A few years ago, a friend of mine sent me the DVD of film version of \emph{The Last of the Mohicans}. It was her favorite movie, so she asked me to explain it to her. This is how I described it.

\begin{quotex}
Clearly a magnificent movie, simultaneously a great adventure and a convincing love story, while demonstrating the variations of human character and the agony of the tough decisions we sometimes need to make.

I looked at Amazon and imbd … many people are captivated by this film. But what grabs Lori? What secret does she harbor, a secret she holds tight while longing for someone to pry it loose. It reminds me of one of those challenges the medieval knights had to conquer on the quest for the grail.

It must be more than the long hair and skill in the martial arts. The turning point for Cora came when they came upon the massacred family. Cora was distracted from her task and wanted to take time to give them a proper burial. Hawkeye, undeterred from his goal and brusquely dismissive of Cora's request, forces them to press on. She is angered by what she considers his cold-heartedness and moral failing. It is only when he brings her to see the situation in larger terms that her infatuation begins. Suddenly, she is no longer the protected bourgeois daughter of a British officer. She is exposed to man's world with all its glory, cruelty, sinfulness, and stripped of its veneer of cultured, ``civilized" life. She also sees that a life with Duncan is part of that old, false world, and that real Love can only exist in the midst of the real world, with a man at home in that world, while still holding values that transcend it.

Hawkeye's and Cora's love is the beginning of their salvation. In the cave, the knowledge of Hawkeye's love fuels Cora's will to live through the faith he will keep his promise. For Hawkeye, all other concerns are secondary as he single-mindedly looks to rescue Cora. He also has the backing of his father and brother who demonstrate their loyalty. Yet love alone is insufficient and the couple needs to rely on the sacrifice of Duncan who, while not altogether the innocent victim, is yet made innocent by his love for Cora, a non-egoic love as he is willing to give her up for her own happiness.

Unlike Cain and Abel, Uncas dies for his brother … and for Cora. But he is unable to save Alice who, not understanding or loving like her sister, yields to her despair. As all the agents — the English, the settlers, the French, the Hurons, the Mohicans — act out their little parts on the stage, they are oblivious to the greater change they inadvertently are bringing about. Only the wise Chingachgook can see it all coming, \emph{sub specie aeternitatis}, as it were. 

\end{quotex}
\paragraph{Math books}
My father used to take me to science classes that the \textbf{Boston Museum of Science}. After class, I waited in the library where I discovered the beauty of mathematics. I recall two specific books: \emph{The World of Mathematics}, \textbf{James R Newman} (4 volumes) and \emph{Great Ideas of Modern Mathematics}, \textbf{Jagjit Singh}. They were popular collections so they were comprehensible to an eleven year old boy.

I was struck by the beauty and elegance of the symbols. In math, there was truth, logically necessary truth, and that appealed to me. It was above the disputes of human opinion, which to this day I am reluctant to engage in. The pure air of mathematics is immaterial and can only be known by an immaterial soul.

Math teaches you how to think logically and dispassionately. Through its abstract structures you learn to discern the minimum conditions for something to be true. No one ignorant of mathematics was allowed to study metaphysics and political philosophy at \textbf{Plato}'s Academy, and for good reason. The reasons still hold true, but they are ignored in the democratic atmosphere of today.

\paragraph{Chess Books}
My father gave me his chess pieces, made of wood, which he kept in a crumpled brown paper bag. I'm sure he had had it for quite some time. After I began to beat him, he stopped playing with me. That is when I began reading chess books.

The books where replete with analysis of moves and possible moves. So I learned to make the distinction between the possibilities of manifestation and those possibilities not manifested. But the more important part was learning synthetic thinking. Strategy is much more important than tactics. It is impossible for the human mind to consider all possible moves, so the player learns instead to ``see" the game as a whole, to aim to establish a certain positional advantage rather than a tactical gain.

Chess and math provide a hint of angelic thinking. The higher orders of angels have fewer ideas, since they are more universal and apply to immense regions of reality. \textbf{Fr. Garrigou-Lagrange} makes this comparison:

\begin{quotex}
A human parallel is the sage, who, in a few simple principles, grasps an entire branch of knowledge. The stronger is the created intellect, the more it approaches the preeminent simplicity of the divine intellect. 

\end{quotex}
To be good at chess or math requires a knowledge of principles which, when known, explain whole realms of knowledge. There is an intuition necessary. Fr. Garrigou-Lagrange explains :

\begin{quotex}
The nature of his ideas, at once universal and concrete, makes the angel's knowledge intuitive, not in any way successive and discursive. He sees at a glance the particular in the universal, the conclusion in the principle, the means in the end. 

\end{quotex}
It is a good experiment to try thinking like the angels every now and then. It should be clear how to do so.

\paragraph{The Book on the Taboo against Knowing Who You Are}
At university, the brother of a friend used to pass out copies of \emph{The Book on the Taboo against Knowing Who You Are} by \textbf{Alan Watts}. I never knew why, but I did read it. It was my very first introduction to Eastern ways of thinking. I became totally smitten by it. I then read every book he wrote and became intrigued by the possibilities of seeing the dominant Western tradition through Eastern eyes. Through Watts, I came across the name Rene Guenon, although some time passed before I could ever locate a book that he wrote.

Although Watts can be dismissed as a mere popularizer, one should never turn down a gift. The Buddhist Tantric literature is full of lessons learned from unlikely sources, such as low caste men or even women.

\paragraph{On the Road}
There is little to say about this book by \textbf{Jack Kerouac}, other than to repeat \textbf{William Blakes}'s Proverb of Hell, ``The road of excess leads to the palace of wisdom." It also can lead to an early death, as it did to Jack Kerouac. In my case, I hope it led to the proper destination.

\paragraph{Libertarianism}
After the excesses of the 60s, the New Left fragmented. Some, the Progressives, continued on that path, although it was the liberals who managed to implement them. Some becomes neo-conservatives led by the former Marxist David Horowitz. A few turned to libertarianism. An odd little monograph titled \emph{Individualism and the Philosophy of the Social Sciences} by \textbf{Murray Rothbard} had an undue influence on my intellectual development. The book tied together three philosophical viewpoints that I still adhere to: \textbf{Thomism}, \textbf{Phenomenology}, and \textbf{Praxeology}.

Praxeology denies that social science can be practiced as an empirical science like physics, since it is impossible to perform experiments on populations. Rather, the socials sciences should determine first principles and derive consequences from them; i.e., in the human realm, ideas create facts.

Rothbard referenced several Thomist texts, which I painstakingly accumulated since they were all out of print. I also learned about phenomenology, from which I learned to pay careful attention to inner states. It is an effective antidote to a scientistic worldview which attempts to leave out the human subject. It is also important in esoteric practice.

\paragraph{Atlas Shrugged}
\emph{Atlas Shrugged} by \textbf{Ayn Rand} was obligatory reading for libertarians. She claimed to be the rightful heir to Aristotle and Thomas Aquinas. Certain themes are recurring. Fundamental is the idea that there are certain objective truths in philosophy and ethics. This appealed to my mathematical mind. There is the ideal of rationality; since a human being is defined as the ``rational animal", she is just proposing a version of a universal truth. In the commitment to the ethical life, which follows reason rather than whim, she is also close to Tradition. Of course, there is the idea of the elite 5\% who actually make the world work. (This is not at all like the alleged 1\% who run things now.)

Unfortunately, her adherence to the Cartesian dualism of matter and consciousness led her to a false view of the soul. Hence her entire system is world-oriented. Nevertheless, while making the dollar sign the highest symbol, the hero John Galt spurns riches to maintain his own independence. There is also an element of anti-feminism when the powerful and intelligent Dagny Taggart prefers to be the simple housewife to Galt.

There are obviously faults in Rand's system, and they showed up in the psychological conflicts of her organization, but the virtues of ethics, rationality, independence, and creativity are praiseworthy. Her critics would do well to write a new novel with those same or better ethics, but without her faults.

\paragraph{Herman Hesse}
At the time I encountered him, reading \textbf{Herman Hesse} was not like reading a book but more like reading my own heart.

\textbf{Steppenwolf}: That was my life: a solitary man with a room full of books. At night I would go out looking for what? Magic? Actually a magical woman. I can't complain about all the problems I encountered, because the woman you are attracted to is the reflection of your own soul.

\textbf{Narcissus and Goldmund}: I thought I was like Narcissus, the intellectual and monk, but somehow I transformed myself into Goldmund over time. I assume that is the point, and the two are really the same.

\textbf{Siddartha}: When asked his skills, Siddartha replied, ``I can \emph{think}. I can \emph{wait}. I can \emph{fast}." Those are skills I continue to develop.

He entered the world of business, he won the love of the beautiful courtesan, had a family. She disappointed him. He ended his days as a hermit ferryman. I knew way back then that the future course of my life had been laid out.

\paragraph{Ethics}
I went through \textbf{Spinoza}'s \emph{Ethics} point by point working out all the proofs. Here again was the promise of an ethical life based on reason and developed with geometrical certitude. Spinoza had been described as both a God Intoxicated man and a atheist. Of course, anyone with a true understanding of God risks being considered an atheist but those who can move beyond a Zeus like being.

In more than a few ways, I modeled my life on Spinoza. Reason, logic, an understanding of Biblical stories beyond the literal meaning, were all features. What was more important was to be independent. So I joined no parties, organizations, and worked as an independent consultant. Spinoza earned his living as a lens grinder. That was a precise, technical field based on science. So I entered the software field since it had analogous features. It is important to work on tasks that have actual effects in the material world. Jobs that involve no more than the manipulation of words — journalists, lawyers, etc. — are bound to lead to distortions in worldviews.

\paragraph{Meditations on the Tarot}
After Vatican II, I no longer bothered with the church because it embraced the modern world. What was its point, since I was already involved with it? Reading \emph{Meditations on the Tarot} when it first came out was a complete shock. What kind of mind could be so precise, incisive, spiritual, all while embracing such a range of authors and teachings? I saw clearly that abstract ideas alone were insufficient; rather they need to be appropriated in one's consciousness through a direct intuition.

I was familiar with several of the primary sources that \textbf{Valentin Tomberg} used (e.g., \textbf{Thomas Aquinas}, \textbf{Teilhard de Chardin}, \textbf{Carl Jung}, \textbf{Nicholas Berdyaev}). I followed up on those I didn't know, particularly Vladimir Solovyov and the various figures in the French Hermetic tradition. At that point, I realized that a Western esoteric tradition still exists and I began to move away from Asian traditions.

\paragraph{Crisis of and Revolt against the Modern World}
The influence of these two books is obvious. However, they are not museum pieces, so they need to be incorporated into a larger whole and all the logical consequences teased out.

The following books were also in consideration, but were left out for lack of space: \emph{Godel, Escher, Bach: An Eternal Golden Braid}, \emph{Memory, Dreams, Reflections}, \emph{Alexandria Quartet}, \emph{Stranger in a Strange Land}, \emph{Transcendental Meditation}, \emph{Confessions}, \emph{Phenomenon of Man}, \emph{Insight}, \emph{In Search of the Miraculous}, \emph{The Conservative Mind}, \emph{Sex and Character}



\flrightit{Posted on 2014-09-08 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Wayne Ferguson on 2014-09-08 at 09:19 said: }

Thanks for sharing this list — much overlap with my own, which I have just revised (unable to limit myself to 10 either).

\url{http://jwayneferguson.wordpress.com/about/the-most-important-books-in-my-life/}


\hfill

\texttt{Scardanelli on 2014-09-08 at 21:58 said: }

The best books, I have found, are those that point beyond themselves; those that, if taken seriously by the reader, eventually render themselves obsolete and unnecessary. As LCDSM says, ``Books are the windows of truth, but they are not the door; they point out things to men, but they do not impart them." Meditations on the Tarot has been such a book for me, and I look forward to the day when it becomes obsolete.


\hfill


\end{sffamily}\end{footnotesize}
