\section{Attack Thoughts and Cold Hearts}

\begin{quotex}
O you, sweet warm soul, how inspired I was in your arms.
\flright{\textsc{Richard Wagner} in a letter to Judith Gauthier}

\begin{verse}
And when man falls silent in his torment.\\
A god allowed me to say what I'm suffering.
\end{verse}
\flright{\textsc{Goethe}, \emph{Marienbad Elegy}}

\begin{verse}
Und wenn der Mensch in seiner Qual verstummt.\\
Gab mir ein Gott zu sagen, was ich leide.
\end{verse}
\flright{\textsc{Goethe}, \emph{Marienbader Elegie}}

\begin{verse}
I carry peace in me,\\
I carry in myself the forces that strengthen me\\
I want to fulfil myself with these forces of warmth,\\
I want to permeate myself with my will power.\\
And I want to feel how peace pours through all my being,\\
When I am strong, peace as power.\\
To find within me through the power of my striving.
\end{verse}
\flright{\textsc{Rudolf Steiner}}
\end{quotex}


\paragraph{Attack Thoughts}

The recent guest post about correct thinking\footnote{\url{https://gornahoor.net/?p=10120}} coincides synchronistically with our group exercise on observing negative thoughts. On the same day, the quote from the Orthodox priest, the \textbf{Elder Thaddeus of Vitovnica}, about the effect of attack thoughts also showed up.

Although I don't like to promote New Age ideas, the Elder's teaching that our thoughts determine our lives is not so incompatible with the \emph{Course in Miracles}. It is no secret that I diligently performed the exercises daily for one year, more than a couple of decades ago. I did learn the role that our thoughts play in our lives. I certainly don't recommend the main text, but some of the exercises are worthwhile. What is relevant are the exercises on Attack Thoughts below.

Conversations and discussions often devolve into mock legal battles. Instead of a dialectic leading toward a higher aim, there is nothing but attack and defence, in an endless loop. If the initial attack is fueled by intense passion, even a guilty plea will often not suffice. Ideally, at least between friends, the attack should not even start. When there is a misunderstanding on one side, or a poorly phrased statement on the other, a request for clarification should be the first step. That most often will rectify the issue before hard divisions are formed.

After all these years, those daily exercises I performed are still fresh in my mind. In a way, this demonstrates the wisdom of Rudolf Steiner's assertion that a mere 5 minutes per day of careful thinking can change your life. Those exercises took little more time than that. I'll leave you with this idea:

\begin{quotex}
Today's idea accurately describes the way anyone who holds attack thoughts in his mind must see the world. Having projected his anger onto the world, he sees vengeance about to strike at him. His own attack is thus perceived as self defence. This becomes an increasingly vicious circle until he is willing to change how he sees. Otherwise, thoughts of attack and counter-attack will preoccupy him and people his entire world. What peace of mind is possible to him then?
\end{quotex}

\section{Attack Thoughts}

\begin{quotex}
Your thoughts are burdened because you are influenced by the thoughts of your fellow men. Pray to the Lord that He might take this burden from you. These are the thoughts of others which differ from yours. They have their plan, and their plan is to attack you with their thoughts. Instead of letting go, you have allowed yourself to become part of their plan, so of course you suffer. Had you ignored the attack, you would have kept your peace. 
\flright{\textsc{Elder Thaddeus of Vitovnica}, \emph{Our Thoughts Determine Our Lives}}
\end{quotex}

\paragraph{LESSON 21}

\textbf{I am determined to see things differently.}

The idea for today is obviously a continuation and extension of the preceding one. This time, however, specific mind searching periods are necessary, in addition to applying the idea to particular situations as they may arise. Five practise periods are urged, allowing a full minute for each.

In the practise periods, begin by repeating the idea to yourself. Then close your eyes and search your mind carefully for situations past, present or anticipated that arouse anger in you. The anger may take the form of any reaction ranging from mild irritation to rage. The degree of the emotion you experience does not matter. You will become increasingly aware that a slight twinge of annoyance is nothing but a veil drawn over intense fury.

Try, therefore, not to let the ``little'' thoughts of anger escape you in the practise periods.

Remember that you do not really recognise what arouses anger in you, and nothing that you believe in this connection means anything. You will probably be tempted to dwell more on some situations or persons than on others, on the fallacious grounds that they are more ``obvious''. This is not so. It is merely an example of the belief that some forms of attack are more justified than others.

As you search your mind for all the forms in which attack thoughts present themselves, hold each one in mind while you tell yourself:

I am determined to see \_\_\_\_\_\_\_\_ name of person differently.

I am determined to see \_\_\_\_\_\_\_\_ specify the situation differently.

Try to be as specific as possible. You may, for example, focus your anger on a particular attribute of a particular person, believing that the anger is limited to this aspect. If your perception is suffering from this form of distortion, say:

I am determined to see \_\_\_\_\_\_\_\_\_ specify the attribute in \_\_\_\_\_\_\_\_\_ name of person differently.


\paragraph{LESSON 22}


\textbf{What I see is a form of vengeance.}


Today's idea accurately describes the way anyone who holds attack thoughts in his mind must see the world. Having projected his anger onto the world, he sees vengeance about to strike at him. His own attack is thus perceived as self defence. This becomes an increasingly vicious circle until he is willing to change how he sees. Otherwise, thoughts of attack and counter-attack will preoccupy him and people his entire world. What peace of mind is possible to him then?

It is from this savage fantasy that you want to escape. Is it not joyous news to hear that it is not real? Is it not a happy discovery to find that you can escape? You made what you would destroy; everything that you hate and would attack and kill. All that you fear does not exist.

Look at the world about you at least five times today, for at least a minute each time. As your eyes move slowly from one object to another, from one body to another; say to yourself:

I see only the perishable.

I see nothing that will last.

What I see is not real.

What I see is a form of vengeance.

At the end of each practise period, ask yourself:

Is this the world I really want to see?

The answer is surely obvious.

\paragraph{LESSON 23}


\textbf{I can escape from the world I see by giving up attack thoughts.}


The idea for today contains the only way out of fear that will ever succeed. Nothing else will work; everything else is meaningless. But this way cannot fail. Every thought you have makes up some segment of the world you see. It is with your thoughts, then, that we must work, if your perception of the world is to be changed.

If the cause of the world you see is attack thoughts, you must learn that it is these thoughts which you do not want. There is no point in lamenting the world. There is no point in trying to change the world. It is incapable of change because it is merely an effect. But there is indeed a point in changing your thoughts about the world. Here you are changing the cause. The effect will change automatically.

The world you see is a vengeful world, and everything in it is a symbol of vengeance. Each of your perceptions of ``external reality'' is a pictorial representation of your own attack thoughts.

One can well ask if this can be called seeing. Is not fantasy a better word for such a process, and hallucination a more appropriate term for the result?

You see the world that you have made, but you do not see yourself as the image maker. You cannot be saved from the world, but you can escape from its cause. This is what salvation means, for where is the world you see when its cause is gone? Vision already holds a replacement for everything you think you see now. Loveliness can light your images, and so transform them that you will love them, even though they were made of hate. For you will not be making them alone.

The idea for today introduces the thought that you are not trapped in the world you see, because its cause can be changed. This change requires, first, that the cause be identified and then let go, so that it can be replaced. The first two steps in this process require your co-operation. The final one does not. Your images have already been replaced. By taking the first two steps, you will see that this is so.

Besides using it throughout the day as the need arises, five practise periods are required in applying today's idea. As you look about you, repeat the idea slowly to yourself first, and then close your eyes and devote about a minute to searching your mind for as many attack thoughts as occur to you. As each one crosses your mind, say:

I can escape from the world I see by giving up attack thoughts about \_\_\_\_\_\_\_


Hold each attack thought in mind as you say this, and then dismiss that thought and go on to the next.

In the practise periods, be sure to include both your thoughts of attacking and of being attacked.

Their effects are exactly the same because they are exactly the same. You do not recognise this as yet, and you are asked at this time only to treat them as the same in today's practise periods. We are still at the stage of identifying the cause of the world you see. When you finally learn that thoughts of attack and of being attacked are not different, you will be ready to let the cause go.

\paragraph{LESSON 26}


\textbf{My attack thoughts are attacking my invulnerability.}


It is surely obvious that if you can be attacked you are not invulnerable. You see attack as a real threat. That is because you believe that you can really attack. And what would have effects through you must also have effects on you. It is this law that will ultimately save you, but you are misusing it now. You must therefore learn how it can be used for your own best interests, rather than against them.

Because your attack thoughts will be projected, you will fear attack. And if you fear attack, you must believe that you are not invulnerable. Attack thoughts therefore make you vulnerable in your own mind, which is where the attack thoughts are. Attack thoughts and invulnerability cannot be accepted together. They contradict each other.

The idea for today introduces the thought that you always attack yourself first. If attack thoughts must entail the belief that you are vulnerable, their effect is to weaken you in your own eyes.

Thus they have attacked your perception of yourself. And because you believe in them, you can no longer believe in yourself. A false image of yourself has come to take the place of what you are.

Practise with today's idea will help you to understand that vulnerability or invulnerability is the result of your own thoughts. Nothing except your thoughts can attack you. Nothing except your thoughts can make you think you are vulnerable. And nothing except your thoughts can prove to you this is not so.

Six practise periods are required in applying today's idea. A full two minutes should be attempted for each of them, although the time may be reduced to a minute if the discomfort is too great. Do not reduce it further.

The practise period should begin with repeating the idea for today, then closing your eyes and reviewing the unresolved questions whose outcomes are causing you concern. The concern may take the form of depression, worry, anger, a sense of imposition, fear, foreboding or preoccupation. Any problem as yet unsettled that tends to recur in your thoughts during the day is a suitable subject. You will not be able to use very many for any one practise period, because a longer time than usual should be spent with each one. Today's idea should be applied as follows:

First, name the situation:

I am concerned about \_\_\_\_\_\_\_\_

Then go over every possible outcome that has occurred to you in that connection and which has caused you concern, referring to each one quite specifically, saying:

I am afraid \_\_\_\_\_\_\_\_ will happen.

If you are doing the exercises properly, you should have some five or six distressing possibilities available for each situation you use, and quite possibly more. It is much more helpful to cover a few situations thoroughly than to touch on a larger number. As the list of anticipated outcomes for each situation continues, you will probably find some of them, especially those that occur to you toward the end, less acceptable to you. Try, however, to treat them all alike to whatever extent you can.

After you have named each outcome of which you are afraid, tell yourself:

That thought is an attack upon myself.

Conclude each practise period by repeating today's idea to yourself once more.

\flright{Lessons are from \emph{A Course in Miracles}}

