\section{Amor Fati and the Heroic Life}

\begin{quotex}
My formula for the greatness of man is amor fati — to change nothing, neither before nor after, throughout all eternity. Not only to bear Necessity, and still less to hide it — all idealism is a lie in the face of Necessity — but to love it. \flright{\textsc{Friedrich Nietzsche}}

A happy life is impossible; the highest to which man can attain is an heroic course of life. \flright{\textsc{Arthur Schopenhauer}}

\end{quotex}
\paragraph{Amor Fati}
The highest form of Amor Fati is to follow the will of God. The first act of will is the acceptance of the circumstances of your birth. To wish it to have been otherwise, is to wish to have been someone else.

When \textbf{Saint Anthony the Great} thought about the depth of the judgments of God, he asked,

\begin{quotex}
Lord, how is it that some die when they are young, while others drag on to extreme old age? Why are there those who are poor and those who are rich? Why do wicked men prosper and why are the just in need? 

\end{quotex}
He heard a voice answering him,

\begin{quotex}
Anthony, keep your attention on yourself; these things are according to the judgment of God, and it is not to your advantage to know anything about them. 

\end{quotex}
\paragraph{The Heroic Life}
To resist the modern world, embrace what is most despised by the modern world and, a fortiori, also despised by some who fantasize about resistance. Embrace the world that valued:

\begin{itemize}
\item Logic, grammar, rhetoric 
\item Geometry, mathematics, music, cosmology 
\item True philosophy in the mould of Plato and Aristotle 
\item Mysticism 
\item Chivalry 
\item Art, Poetry, Architecture 
\item Social Order 
\item Care widows, orphans, the sick, and infirm 
\end{itemize}
Dante showed the path from the bug life (mere animation) to Union with God. The seven stages are described in their all-too-human detail, not as some lifeless abstractions. This path will lead to results and save you years of fruitless seeking. Do not just read, but actually apply, what is contained in these texts:

\begin{itemize}
\item \textbf{Richard of Saint Victor}, \emph{De Contemplatione} 
\item \textbf{Saint Bernard}, \emph{De Consideratione} 
\item \textbf{Saint Augustine}, \emph{De Quantitate Animae} 
\item \textbf{Saint Thomas Aquinas}, \emph{Summa Theologiae} 
\item \textbf{Dante}, \emph{The Divine Comedy} 
\end{itemize}


\flrightit{Posted on 2020-08-07 by Cologero }
