\section{Right Thinking: Impartiality}

Training in the path of the Heart aims for the thinking, feeling, and willing functions to be animated by the Higher Self. Then there can be gradual improvement in these areas:

\begin{itemize}
\item Thinking and Feeling: \textbf{Positivity} 
\item Thinking and Willing: \textbf{Impartiality} 
\item Thinking, Willing, and Feeling: \textbf{Harmony} 
\end{itemize}
In this case, we want to focus on the development of impartiality, which means that thinking is objective and disinterested. Here presented are some suggestions to develop impartiality and then some obstacles to impartiality. The thinking faculty must develop right opinion and right judgment. This is as much a matter of the will as it is of thought.

\paragraph{Right Opinion}
Pay attention to one's ideas.

Think only significant thoughts. Learn little by little to separate in one's thoughts the essential from the nonessential, the eternal from the transitory, truth from mere opinion.

While listening to the talk of others, try to become quite still inwardly, foregoing all assent, and even more so, all unfavourable judgments (criticism, rejection), even in one's thoughts and feelings.

\paragraph{Right Judgment}
Make a judgment on even the most insignificant matter only after fully reasoned deliberation. All unthinking behaviour, all meaningless actions, should be kept far away from the soul. One should always have well-weighed reasons for everything. And one should definitely abstain from doing anything for which there is no significant reason.

Once one is convinced of the rightness of a decision, one must hold fast to it, with inner steadfastness.

Right judgments are formed independently of sympathies and antipathies.

\paragraph{Dunning-Kruger Effect}
\begin{quotex}
The \textbf{Dunning–Kruger effect} is a cognitive bias whereby people with low ability, expertise, or experience regarding a certain type of a task or area of knowledge tend to overestimate their ability or knowledge. \flright{\textit{Wikipedia: Dunning-Kruger Effect}\footnote{\url{https://en.wikipedia.org/wiki/Dunning\%E2\%80\%93Kruger_effect}}}

\end{quotex}
This is opposed to the principle of coming to a conclusion \textbf{only after fully reasoned deliberation}. The complementary effect is that high performers may have a tendency to underestimate their skills. That is because high performers understand how difficult it is to gain knowledge. That means they recognize ignorance so that new ideas motivate them to learn even more.

The low ability person, when hearing a new idea, tries to interpret it in the framework of his own limited knowledge. Hence, he thinks that he already knows it.

Those involved in esoteric studies need to be extra vigilant about this effect. Don't presume that you understand an author so easily; esoterism and metaphysics can be both tricky and difficult. Although they may appear to be dealing with common or philosophical issues, they are not. If you encounter references to other texts, then follow them up.

\paragraph{Gell-Mann Amnesia Effect}
This effect was first described by the author \textbf{Michael Crichton}:

\begin{quotex}
He describes the phenomenon of experts believing news articles on topics outside of their fields of expertise, even after acknowledging that articles written in the same publication that are within the experts' fields of expertise are error-ridden and full of misunderstanding. \flright{\textit{Wikipedia: Michael Crichton, Other Speeches}\footnote{\url{https://en.wikipedia.org/wiki/Michael_Crichton\#Other_speeches}}}

\end{quotex}
This should be easy, but it isn't. How often have you read in a newspaper, or heard on a TV news show, something that you know to be false or at least misleading. In legal terms, that would be called ``impeaching the witness", so that the witness need not be believed. If a journalist or new organization is willing to present false information in one field, there is no reason to believe them uncritically in other fields.

Unfortunately, few people have enough expertise in a field for this effect to take hold. They assume that the news source has authority and are willing to believe whatever they say.

In esoteric studies, people often go by reputation, and are willing to believe those with a good reputation whether deserved or not. Following a false path or idea will cost you possibly years of wasted effort.

\paragraph{Myside Bias}
\begin{quotex}
Myside bias occurs when people evaluate evidence, generate evidence, and test hypotheses in a manner biased toward their own prior opinions and attitudes. Research across a wide variety of myside bias paradigms has revealed a somewhat surprising finding regarding individual differences. The magnitude of the myside bias shows very little relation to intelligence. Avoiding myside bias is thus one rational thinking skill that is not assessed by intelligence tests or even indirectly indexed through its correlation with cognitive ability measures. \flright{\textsc{Keith Stanovich}, \emph{Myside Bias, Rational Thinking, and Intelligence}\footnote{\url{http://www.keithstanovich.com/Site/Research_on_Reasoning_files/Stanovich_CDPS_2013.pdf}}}

\end{quotex}
Professor Stanovich is an expert on Myside Bias and its insidious effects. Although we recognize accept this in practical affairs, such a bias is even more insidious for those engaged in esoteric studies. Such work needs to be experienced as totally new. If you believe a text because it confirms the viewpoint of your particular side, you will be led astray.

\paragraph{Critical Thinking}
Although critical thinking often leads to cynicism and doubt, that need not be the case. Rather, at its best, it simply means to test all ideas in respect to logic, common sense, experience, and prior knowledge. In this video, clinical neurologist \textbf{Steven Novella} explains the neuroscience of critical thinking. Although we have no interest in his physicalist and materialist viewpoint, there is some value in hearing him out. It is just necessary to transpose his ideas to a more metaphysical plane. In the end, he cannot help himself and starts referring to the brain as if it were a person.

\url{https://www.youtube.com/watch?v=MHgevPKZfJQ}



\flrightit{Posted on 2022-10-23 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Taivasvaeltaja on 2023-10-23 at 09:03 said: }

In one aspect of the process we must essentially try to become masters of citta, the cosmic thought substance that manifests through us as samsaric beings and because of the result of the Fall as impulsive and uncontrollable imaginative forms of maya, and thus steals our unification and individuation process.

The process of becoming the masters of our true fate as the awakened ones who are detached, free as birds of the wind, born again from the spirit and water; the bearers of the torch of liberated humanity, is done by straightening your posture, breathing deep, detaching from your ordinary state of an animal man with passions and bonds, seeing through the illusion of ordinary life, only apparently dis-connected from the supra-Human realms, and then manifesting the Divine Order into your life and to your surroundings.

For the spiritual warrior now as Mars slowly moves out from its initial sign Scorpio, this is the time to ascend to the Polaris and bring the Dharmic Order of the spiritual Cosmos alive in one's life with the Transcendent Guardian of the Heavenly Order…Tiwaz, Taivas, Deiwaz, Deus Pater, Ju-Pater, Divspatri!


\end{sffamily}\end{footnotesize}
