\section{Moral Purification of the Will}

\begin{quotex}
Man's mind is rapt by God to the contemplation of the divine truth in three ways:

\begin{enumerate}
\item He contemplates it through certain imaginary pictures. 
\item He contemplates the divine truth through its intelligible effects. 
\item He contemplates it in its essence. 
\end{enumerate}

Now when man's intellect is uplifted to the sublime vision of God's essence, it is necessary that his mind's whole attention should be summoned to that purpose in such a way that he understands nothing else by phantasms, and is absorbed entirely in God. 


\flright{\textsc{Mary of Agreda}, \emph{Mystical City of God}}
\end{quotex}

Mary of Agreda has given us the three stages of Hermetic contemplation, which are related to their Yoga counterparts in the Table~\ref{tab:PurificationWill}:

\begin{table}[h]
\label{tab:PurificationWill}
\begin{tabular}{cccc}
\toprule
\textbf{Hermetism} &
\textbf{Yoga} &
\textbf{Theophan} &
\textbf{Description}\\\midrule
Concentration &
dharana &
Spoken prayer &
Visualization\\
Meditation &
dhyana &
Mental prayer &
Mental Prayer\\
Contemplation &
samadhi &
Unceasing prayer &
Prayer of the Heart\\\bottomrule
\end{tabular}
\end{table}


In his booklet, \emph{The Path of Prayer}, \textbf{St. Thephan the Recluse} provides some basic guidance on these stages. \emph{Spoken prayer} is when we pray using the prayers of others. In \emph{mental prayer} we raise our mind to God through reflection on divine things. But these are prepatory stages to \emph{unceasing prayer} which is the constant turning of mind and heart to God.

In \emph{Meditations on the Tarot}, Tomberg describes a type of meditation based on visualization which I call Hermetic Meditation\footnote{\url{https://www.gornahoor.net/?p=898}}, which culminates in a state of Deep Silence.

\begin{quotex}
Concentration without effort, which means there is nothing to suppress and where contemplation becomes as natural as breathing and the beating of the heart, is the state of consciousness — of the intellect, the imagination, the feelings, and the will — a state of perfect calm, accompanied by the complete relaxation of the nerves and muscles of the body. It is the deep silence of desires, concerns, imagination, memory, and discursive thought. We would say that the entire being has become like the surface of calm waters reflecting the immense presence of the starry sky and its inexpressible harmony. And the waters are deep, oh how deep! And the silence increases, always increasing, what SILENCE! Its growth takes place in regular waves which pass, one after the other, through your being: one wave of silence followed by another wave of deeper silence, then yet another wave of even deeper silence … Have you ever drunk the silence. If so, you know what concentration without effort it. 

\end{quotex}
Beyond the images and thoughts of the first two stages, the third stage is a direct intuition, or gnosis, of the divine essence. It involves the Will and the Heart. But Purity of Heart is to Will one thing. A divided self or impure heart cannot attain this stage. Tomberg explains:

\begin{quotex}
It is futile to attempt to be concentrated if the Will is passionate about other things. The oscillations of the mind will never be able to achieve silence unless the the Will itself infuses it with silence. Only the still Will can render the imagination and the intellect silent in concentration.

St. John of the Cross and St. Theresa d'Avila never tire of repeating that the concentration necessary for spiritual prayer is the fruit of the \textbf{moral purification of the Will}

\end{quotex}

\hfill

Originally published on 12 April 2011 in medtarot.



\flrightit{Posted on 2023-01-02 by Cologero }
