\section{The Test}\label{sec:TheTest}

Although Evola always writes in a discursive style, he advises us that it is only the rishis, the seers, the clairvoyants who can fully understand. In many ways, Miguel Serrano picks up on some of Evola's themes, though from the perspective of a clairvoyant rather than a metaphysician. In the following passage, Serrano describes the highest Tantric teachings while bringing in the major elements of the esoteric teachings from the Middle Ages: the Grail legends, the Templars, troubadours.

The ultimate end is identical for both Evola and Serrano. Both reject the mysticism of pantheism which looks to the annihilation of the Self and its merging into the Totality. Rather, they proclaim the Absolute Self, the raising up from the merely human to the higher states, the process of deification, atman=brahman.

\begin{quotationx}
Thus were completed the different stages of this most ancient Hyperborean Initiation of A-Mor, revealed in the mystery of the Grail, in the esotericism of the troubadours and the Minnesänger of the High Middle Ages. Transported to the icy wastes of the south of the world, with Parsifal, in a Templars' ship, with the Vermilion Cross on its white sails and all its lights on, as the saga tells us, and from `whence it never returned'. To the true Kingdom of Hyperborea of the White Gods of America-Albania. 

\emph{And while the ultimate test of this initiation was taking place in that ancient night, with a man and a woman lying naked side by side, separated by a sword, without taking possession of each other's physical body, she explained to him in her musical voice full of longing for eternity}: `The light doesn't come from the east. Light is only truly light in the depths of midnight. Now is the depths of midnight. The followers of Lucifer, of the Morning Star, do not beg to be allowed into heaven. They demand to be, because they feel that they have done everything possible to merit being deified. At the end of our road, no fusion with a god or redeemer awaits us. Our way is not the way of ecstasy of the saints but the way of separation of the magicians, of the White Gods who have become absorbed into the sources of creative energy. Creating worlds, loving each other inside and outside eternity. We do not beg, like the lunar troubadour: “Take us back to where you took us from!” We are going to try and change God, giving him a face. Therefore, my love, do not take possession of my body. Let us not create children of the flesh. I will make you pregnant with the son of death. And we will both remain virgins.'

`I understand,' he whispered. `The chastity of the sacred warrior is the nobility of his sexual act, the refusal to tolerate all that is brutal, because his feeling for the beauty of that act prevents him doing so. The Wounded King also said this.'

`It is an immaterial irrevocability. The Grail doesn't tolerate unbridled passions, it loves pious reticence, a reverent attitude. And I will not destroy your magic virility, dividing your flesh and mine, giving you children of the flesh to bring new opportunities to other individuals, when such a great possibility already exists for us. I will not lure you into loving my body in the only way known to the dark age, because that way death will swallow you up. I will never be the Great devouring Mother, the Primordial Female, who will turn you into a vanquished warrior, living in a dream of unfulfilled glories. I will be the She who leads you to heaven. Because it is your magic virility which will enable us to travel along the river of death. Your sacred virility will enable us to return to life. Do you remember the words of the ascetic of the Grail?: “You will become a woman if you love the body of a woman.” This is so: because \emph{only by becoming effeminate could you satisfy the erotic sensibilities of a woman's physical body}. The chaste warrior is the more virile one. To take physical possession of the beloved is to lose one's soul. True possession is the mental possession of all the other bodies. With the memory or your beloved in your heart, you will achieve the Grail. \emph{The genuine orgasm isn't a physical one but another which is endless, and which is produced by your contact with my invisible bodies}, where you will find the perfume of my visible body, the warmth of my lips, the amorous racing of my blood intensified. As I will find in yours. We must discover this love together, when we are no longer made of mortal flesh but of red imperishable matter. By loving my body, which lies beside you, you will make it even more material, you will turn it into a body made of lead.' 

\flright{\textsc{Miguel Serrano}, \textit{Nos: Book of the Resurrection}}
\flright{[emphasis by editor]}
\end{quotationx}


\flrightit{Posted on 2008-08-30 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{imperator on 2008-09-08 at 18:08 said: }

Wonderful section, packed with philosophical fury. Warrior-ascesis, something like a new Templar order to combat the mafias and anarchists of modernity, is what shall bring salvation to the West, not masochistic limpwristed priests.


\hfill

\texttt{JA on 2013-07-25 at 01:56 said: }

Serrano was a strange man, his books sometimes contain profound insight, more profound than even Evola at times, and more wise than even many eastern sages of our day yet…….his overall worldview is insane, no other way to put it, Hitler as God ? I wonder how such a gifted person could err so wrongly…..


\end{sffamily}\end{footnotesize}
