\section{I'm Just a Girl}

A few years ago I was watching an Argentinian movie about the Tango with Cologero. The main character, the aging movie director, was seducing a young dancer. He told her, ``Sometimes I feel like a lion wandering aimlessly in the savanna." Of course, that turned into a teaching moment. The lion does not wander aimlessly. He established the boundaries of his territory by pissing on the bushes; the scent of his urine is a signal to other lions to stay away.

Within those boundaries the lion is a presence, a ``being there". He creates the feeling of security for his harem so they can search for game and raise the cubs. I objected that the lion is lazy, since the lionesses are responsible for hunting. Cologero insisted that the lion is like the unmoved mover and the action of the lionesses are really the female force. It is only our age that regards women as knickknacks; there is no reason for them to be useless or incompetent.

I'll consider that but I'm not quite convinced. On the other hand, unlike the prairie dogs\footnote{\url{https://gornahoor.net/?p=4898}}, there are no interlopers that dare approach his harem. Even Clint Eastwood would not risk his life for a few nights with a lioness of average looks and intelligence.

I sort of get that ``presence" thing. This past spring, I saw Pitbull at an outdoor concert. Like a lot of women, I find Pitbull incredibly attractive and compelling. I was hardly alone, as there were thousands of screaming women, yelling at him to take his clothes off. Cologero said that if ever saw him in street clothes at Target, I wouldn't look at him twice. I suppose that is true. Yet on stage, he is totally compelling. There is something about a man at the center of attention that excites a woman's passion.

I can hear the complaints now, just as I have so often in therapy sessions: ``Women want the bad boy, and the good guys are left wanting." That is not quite true, as there are women who want the good guys; it's just that the good guys don't want them. Even if it isn't Pitbull, different women will be attracted to different qualities, but they will always prefer the one that stands out from the crowd. Everyone else just seems interchangeable. If you need a toothpick, does it really matter which one you pick out of the box?

I know everyone thinks they need to be ``nice", say by buying flowers and chocolate. That is nice, but it would be much nicer from the equivalent of Pitbull. I expect an email from Cologero because I am using this wrong: Flowers are the path of the ancestors, but Pitbull is the path of the Gods.

That is simply the reality of sexual attraction, so intelligent men need to deal with it. I'm sure you can make the case that the wrong qualities are at the center of attention in our time. Intellectually I may agree, but primally I feel otherwise. If it is true, as I have been reading here, that intelligence is at the heart of reality, why is the world as it is arranged so unintelligently? Why don't the intelligent men take control and redo society in their own image? I think it is because they lack the qualities of the Lion. It seems to me, instead, they are expecting the lionesses to change. That won't happen. So how can that be done, you ask? Well, don't ask me, I'm just a girl.



\flrightit{Posted on 2012-09-29 by Teresa Chicalinda }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Audrey on 2012-09-29 at 22:50 said: }

Ye, see that is the basis of war.. those other guys pants are poopy, and they breed like rabbits! Mussolini was awesome about it.. perfume your underwear with roses y'all. Then the Germans came along, mis-read Evola, and lacking any sense of humor about coffee and the underwear washing thing, turned that into crotches of roses and bread baking.. and tgen there was Auschwitz. I hear Lithuanian ho's make the best Svara rabbits.


\hfill

\texttt{Audrey on 2012-09-29 at 22:54 said: }

Born on a cock, died po Svarog. Germanz are ppl two. I have indeed released a record calked Iron, and dedicated it t the US Army Special Forces, who were founded by ex-Russians trying to get out from behind the Iron Curtain.


\hfill

\texttt{Audrey on 2012-09-29 at 23:11 said: }

Pitbull is the path of the Gods? I am laughing here .. Apropos of the lioness commentary, that is simply the male enslaving the female skeletal structure to his reproductive impulses. There is no normal female who would concent, hence Nature invented boobies, and pearls..ie, drugs. The first principle of authentic Tantric sex is for the female to give up orgasm, and second is the oath not to use her body as a vehicle of another's desires. That means no kids. Society is so geared towards the family and pressuring the female it is not funny. The extremes are sacred prostitutes and military rape, which they call `mead'.


\hfill

\texttt{Audrey on 2012-09-29 at 23:26 said: }

Moreso, until a real Guru ties your hands with an authentic, real mala, and you are taught mudras, you will always only and only be a playbunny.


\hfill

\texttt{Andrew on 2012-09-30 at 02:47 said: }

This site has been hijacked by a troll in Cologero's absence. I hope.


\hfill

\texttt{Audrey on 2012-09-30 at 18:29 said: }

I am perfectly fine keeping my thoughts to myself, Ansrew, because most humans do not care to face the fallen aka predatorial nature of human sexuality. Nor the fact that things need to be faced squarely befoee they can be refined, which us the only way to change.

Look at our society… Clinton lies to his wife, and others condone rape as a form of conception.

Maybe you need a dose of grow up, and face the music because due to prim behavior we have \$5 his beaten and raped by middle class white millionaire wannabes and taco hell.

Just sayin' Andrew, but I also live outside of the world of library, university and home.


\hfill

\texttt{SW on 2012-10-01 at 15:31 said: }

So it's actually the lionesses that go out and hunt and the lion king that stays home. Interesting.


\hfill

\texttt{Graham on 2012-10-02 at 23:07 said: }

``Why don't the intelligent men take control and redo society in their own image?"

An intellectual could be defined as a man who attempts to do just that. Obviously nothing could be less traditional and more revolutionary: the intellectual, in this sense, is a Paine, a Marx, a Rousseau. Sites like this one exist because intelligent men have succeeded in redoing society in their own image. We attempt something quite different.

``I think it is because they lack the qualities of the Lion. It seems to me, instead, they are expecting the lionesses to change. That won't happen."

And I think that that's a platitude followed by a blame-shift typical of `conservative' women with ensconced feminist tendencies. I wonder if you actually believe what you're implying, namely that women are devoid of agency and responsibility for themselves. Strangely that seems to be a point on which machos and feminists effectively agree. Myself, I suspect they are in symbiosis. 

Perhaps women will have to learn the hard way what happens when from pride they refuse obedience to men: with voluntary submission the yoke is easy, and the burden light; with involuntary submission, not so much! [D83D?][DE09?]


\hfill

\texttt{Audrey on 2012-10-03 at 04:17 said: }

From pride? What does pride have to do with experiencing sex as an energy of liberation and knowledge, and nt as a yoke, easy or hard? Mind you the reason why women get yoked is so that they get entrapped by the male structure anyway. What I mean by this is the fact that during reproduction, esoterically speaking, the male skeleton, brain and skull reproduce, whereas the female body is only sacrificed as blood and flesh. After that, the child feeds on the mother's body until it is time to separate.. think about it.. it is a lateral extension of the structure into a temporal `future'. Not only does it tie up the hands of the female, but it dictates to her rhythm, life direction, and she has no future. Hence, the perceived triumph of females over males, when the child is female, and the whole notion chiefly propagated by what I call matt patatt of child bearing as golden lamb and if women do not reproduce they will disappear bullshit. The whole German-Russian thing, which is bullshit, yet rocks and undermines society like a motherfucker.

It is not only that the female body is regarded as receptacle, but there is no actual registration of the female body at all. In other words, females have no awareness of themselves as bodies at all, physically and intellectually. It is classic black Tantra.

Briefly, and with not much precision, there are considered to exhist three level of knowledge, oral tradition based on breath energy, the book, and that knowledge which is deposited and transmitted by the skeletal structure .. note hiw females are excluded from gun wearing, sword practices, being generals, ranger school, all of that `male' physical exercise.. and we are to be sluts in the bedroom, goddesses in the kitchen, and oh yeah, fetch home the bread like a good bitch too. All concepts of `femininity' are servile to the male, and intellectually I am to salute little dicks who only see as high as mynleft breast..mmkay no.

I do have extensive formal university education as a cognitive svientist, mathematician, linguist, philosopher and art historian, and formally have the intellectual respect of the general manager of microsoft, harvard graduates in non business areas, and so forth. I find these to be mental gymnastics of the kid variety at best. At worst, reference Hadot, par example.

Gurdjieff is accurate about the kleenex part, and again during birth the last dot on the picture is the nose.. it is a symmetrical funny shape, and is much responsible for why males try to `hermeticize' female hands, male fantasies about mixing semen with excreta (sorry if yall not comfortable with graphicness), but also the popularity of say the shovel in military context, the playbunny phenomenon, the attractiveness of pitbull, olfactory smell sexual fntasies, such as bread, the smell of babies, perfumes, as well as a good portion of the non-mithraic aspects of wine, and the popular fantasy of omen as apples.. not to mention why females like to lie on chaises nd discuss their sex lives.. Dick Couch anyone?

Females are externally yin and internally yang, with the male opposite. During sexual discharge, aka orgasm, the female brain is flooded with pleasure, just as when she is breastfeeding, so she does not actually self-actualize the sexual energy, but gives it away. Purely to feed the external, yang shell of the male. If she provides food for the male, she is also feeding her yin energy to his internal yin. Very schematically speaking.

That is why I carrry my m16 left arm, and keep and berretta on my right thigh and a knife around my belt.

And if you think I will salute to your head senor coconut, instead of you mine (Evola also commented on the active and passive male/female in those terms), we are going to tango.

And you are right Cologero, I do not really care about the (sand) box. I think a bit of distraction in the way of examining the meaning of Grammar in Sanskrit would give you a small hint into somethng else.


\hfill

\texttt{Graham on 2012-10-03 at 12:04 said: }

Audrey you're crazier than Charlotte ! I for one never realized that the shape of the nose was responsible for so much, up to and including the popularity of the shovel in a military context.

Anyway, yes, Audrey. From pride. 

Pride that leads some women to see fathers, brothers, and husbands as tyrants and children as parasites … leads them to call loss of chastity `liberation and knowledge’ … leads them to brag about their degrees when nobody asked. Rather than admiring the Lion we could all do with more study of the Lamb. Submission: it's not just for women !

Out of curiousity, what are the (written) sources of this `esoteric feminism’ ?

To return to Terisa's post. The proffered symbolism of the lion and his harem, applied to Pitbull and his hysterical fans, is an example of the current organic symbiosis between machos and feminists, who both seek to `liberate' women's libidos. Game bloggers call this hypergamy and fittingly explain it in animal terms – rather like a lion and his harem, in fact.


\hfill

\texttt{Gabe Ruth on 2012-10-03 at 15:37 said: }

Though he cannot (yet?) be categorized as a traditionalist, this writer does have a clue about what ails the world: kinglessness.


\hfill

\texttt{Saladin on 2012-10-04 at 20:48 said: }

Well said!


\hfill

\texttt{Just a Guy on 2012-10-18 at 18:45 said: }

So much of people's lives is wasted in this. It is time to arrange marriages again. Get it over and done with early and move on with life, not decades of emotional drama and societal masturbation.


\hfill

\texttt{Andrew on 2012-10-18 at 20:27 said: }

Thumbs up.


\hfill

\texttt{Lucas on 2015-08-30 at 08:25 said: }

I understand your intentions, hahaha.


\hfill


\end{sffamily}\end{footnotesize}
