\section{Who do you love?}

The New York Times had an article this week about the latest dating trend, which is to ask about your date's credit score before you get too involved. That is something pretty girls don't really have to worry about because we know that if this one gets away, there will be another to take his place. I never really had to ask that question myself, since there are more subtle ways to glean this information. I always order a glass of cognac for a cocktail and watch their face for a tell. If they flinch, even a little, that puts my guard up. My worst experience with that was when one date didn't care about the \$20 glass of cognac, but objected vehemently when I ordered it on the rocks. He said it was a sad way to treat cognac and made me appear unsophisticated and prolish.

We dated for a while nevertheless, but he did treat me rather shabbily after that incident, and I believe because of it. I've tried it without ice, but can't seem to get it down. Well, back to the point, there are certainly more subtle ways to gauge this. If he pays with a gold Amex card, I know that they have run their own credit check on him. I'm a bit more suspicious when a date pays for everything with \$100 bills although I do find it somewhat exciting.

The experts say that you should verify a man's driver's license if you are meeting him for the first time. That is hardly sufficient, in my opinion. After all, virtually anyone can get one. No, what I ask for is his \textbf{Concealed Weapon Permit}. Then you know he has no criminal record and has passed a background check!

But, yes, I do have a serious point. After our first date, when it seems that there is something more that we mutually want to pursue, I ask for a background check. No, not one of those crappy Internet thingies, but a more personal one. You may think this is impudent, but trust me, it is worth the trouble. And boys are more cooperative than you might expect.

From my work with female clients, I understand the frustration women feel when their men do not open up with them. Boys don't like to talk about that girl stuff. However, if you ask him for evidence, you may be surprised what you find. My dates have shown me grades, SAT scores, and even IQ test results. You would be surprised, too, how many keep scrap books. They may keep clippings of their sports exploits, promotions, charitable activities, and so on. Having something concrete in front of them seems to get boys to open up more, so they become more talkative. You can find a lot about a person in this way. Photographs, too, are a great tool. Girls, you will do this at some point anyway, so why not learn all you can at the beginning? If he has no documentable past, is that really something you want to take a chance on?

But hold on, that is not all. I also ask for three references, obviously not related to him. Again, boys, if they want you, are more accommodating than you might imagine. And if he can't find three people who admire him, do you really want to belong to a club of one? Usually, I get to talk to close friends, sometimes even the man's minister. If you get his AA buddy, you'll have to decide how best to handle it.

The fellow with the hangup about the ice cubes gave me the names of his childhood friend and two female friends. I wasn't sure if they were ex-girlfriends, or women he kept on the bench in case he needed a sub. To my mind, it showed a lot of élan, as well as a subtle warning that I was replaceable. Neither said a bad word about him, even after an hour, although I got the impression they would be happy if our relationship faltered. It did shake my confidence, since he seemed ready to dump me over ice in a glass of cognac. I've never been dumped, so it just made me want him more.

I'll end with a warning: like a good lawyer, never ask for information you don't want to hear. One date took me to an Italian film festival and then to a nice Italian restaurant afterwards. I always felt a couple should be open about their past experiences and how it affects them. So I spoke to him about my recent divorce, how it had torn me up, even to the point of not even telling my family for several months. I then asked him to tell me about his past relationships and what he had learned.

To my shock, he went on for the better part of an hour. He related a history of the past 15 years of his several girlfriends, how he met them, why they broke up, their personal histories (and some of them were bizarre), even their sexual practices. He seemed to take great pleasure in comparing the sizes, shapes, and tints of their body parts. He told me he skipped some because he couldn't remember their names. What made it so disconcerting, is the complete lack of ordinary emotion as he went on, nor any empathy for the feelings of those women. He may as well have been describing a deer hunt with his hunting buddies.

I was getting quite uncomfortable, fidgeting in my seat, yet not daring to interrupt. At a suitable stopping point, I told him how I felt. He gave me a quizzical look, responding, ``That's what I thought you asked about.'' He then returned to his meal with great gusto, and I pondered whether I would be the end of the road for him or the topic for a future conversation.


\flrightit{Posted on 2012-06-01 by Teresa Chicalinda}

