\section{What Does Woman Want}

\begin{quotex}
Happy wife, happy life.
\flright{\textit{Old wives' tale}}

The great question that has never been answered, and which I have not yet been able to answer, despite my thirty years of research into the feminine soul, is `What does a woman want?'
\flright{\textsc{Sigmund Freud}}

\textit{On se lasse de penser, meme d'agir, jamais on ne se lasse d'aimer.} (We get tired of thinking, even of acting. But we never get tired of loving.)
\flright{\textsc{Auguste Comte}}
\end{quotex}

\paragraph{Women of Power and Virtue}

Since most biographies at the time were about men, \textbf{Giovanni Boccaccio} redressed that imbalance in \emph{On Famous Women}. He wrote about 100 or so women, mostly from antiquity. However, he was not simply interested in telling tales or repeating gossip, but rather in drawing moral conclusions from their lives. Hence, there are stories about virtue, greed, duplicity, adultery, loyalty, and so on. He did not regard women as pets who needed to be shielded from the harsh realities of life. In the story of Sulpicia, he makes that perfectly clear:

\begin{quotex}
Women should not always be resplendent with gold and jewels; they should not always indulge in ornaments; they should not always flee the sun in summer and rain in winter; they should not always dwell within their changers; they should not always spare themselves.
\end{quotex}

To the contrary, the women he admired were as strong as men in their own way. When her husband Cruscellio, whom she loved dearly, faced hardships, she shared them also. Boccaccio explains:

\begin{quotex}
When the blows of Fortune demand it, women must endure toil, go into exile, bear poverty, and face danger bravely with their husbands. This is the admirable service of women; these are their battles and their victories. Their striking triumphs are to overcome ease and luxury, as well as domestic difficulties, by means of their virtue, steadfastness, and purity of mind. In this way they win perpetual fame and glory. Let, therefore, not only those women be ashamed who follow the shadow of happiness with all their efforts, but even more let those be ashamed who shun seasickness, are worn out by light work, are frightened by foreign lands, and become pale on hearing the bellowing of an ox.
\end{quotex}

Women became rulers not because they crashed some imaginary glass ceiling, but rather because they displayed the qualities of leadership and fairness. Thus he write this about Dido, the Queen of Carthage:

\begin{quotex}
She cast aside womanly weakness and hardened her spirit to manly strength, and for this she later deserved being called Dido, which in Phoenician means `heroic'.
\end{quotex}

Women could also be warriors, followers of Diana, like Camilla, Queen of the Volscians. Of course, a warrior requires self-discipline. Boccaccio draws this lesson from Camilla:

\begin{quotex}
I wish that the girls of our time would look at Camilla and, seeing that virgin already adult and free amusing herself by running through fields, forests, and the dens of animals with her quiver, curing the pleasures of lascivious desires with work, refusing pleasures and amusements, abstaining from elaborate food and drink, and with constant spirit rejecting not only the embrace, but even the company of young men of her age, would learn from her example.
\end{quotex}

Men were often at fault, forcing themselves on women through their lust. Circe, the daughter of the Sun, affected men through her remarkable beauty. Boccaccio describes her effect on men:

\begin{quotex}
She made many men become haughty because of her matchless attraction, and so those who through this wicked woman's deeds appeared to have lost human reason, seemed to have been deservedly changed into beasts through their own crimes.
\end{quotex}

In this regard, she was hardly unique. Boccaccio continues:

\begin{quotex}
If we consider human behaviour we can well understand from this example that there are many Circes everywhere and many other men are changed into beasts by their lustfulness and vices.
\end{quotex}

It is up to men themselves to overcome their lustfulness.

\begin{quotex}
Ulysses, instructed by Mercury's advice, i.e., by Hermes , obviously signifies the wise man who cannot be bound by the trickery of deceit.
\end{quotex}

Although he praised the chastity of many of the women, Boccaccio was well aware of the sometimes overwhelming power of sexual concupiscence. Chastity, therefore, had to be the free choice of a mature woman, and he opposed forcing virginity on girls. Rhea Illa was forced by her brother to become a Vestal, pledged to eternal virginity. Therefore he rejects sending young women to the nunnery against their will. That leads to unhappiness:

\begin{quotex}
To console their saddened hearts, they have recourse only to thinking how they can destroy their prison and flee, or at least bring their lovers inside, trying to take furtively the pleasure which has been denied them.
\end{quotex}

Hence, he concludes:

\begin{quotex}
Girls must not be consecrated to God while ignorant, or young, or under coercion. On the contrary, they should be well brought up from childhood in the parental home, taught honesty and praiseworthy behaviour, and then, when they are grown and with their entire mind know what they are doing, let them submit to the yoke of perpetual virginity of their own free will and not under duress.
\end{quotex}

\paragraph{The Double Standard}

The term ``double standard'' is hardly used any more. Not so long ago, it indicated that men and women were held to different standards in regards to their sexual activity. Presumably, men were admired for their sexual conquests, whereas women were condemned as sluts for being promiscuous. But life is not so simple.

There simply is not the same degree of difficulty. Almost any woman can walk into a bar and find a casual sex partner for the night. For men it is not so easy, since a few men get all the luck, but most would go home alone. Women are most attractive when young, but men mature later. After leaving the world of boys, they have to learn to deal with rejection, and only hard-earned experience will teach them what women want.

At a pool party some time ago, I met a fellow who was married to a TV sports executive. Although he had a normal build, she looked like a wrestler, a male wrestler to be clear. Seeing my pretty girlfriend, he must have felt he owed me an explanation. He had gotten drunk at a party, took her home, and one thing let to another. She got pregnant so he married her. His explanation, however, did not suffice, since they had a second child, probably not the result of a drunken orgy. He did the manly thing, rejected abortion, and married her. He really should not have been concerned about what other men think.

The British rag, the \emph{Sun}, recently quoted a ``study'' claiming that sex on the first date is best to get into a romantic relationship. That just means that most men are so hard up for sex that they'll latch onto any women who offers them a steady supply.

I used to be shocked at the number of women who were willing to cheat on their boyfriends or husbands with me. Were the husbands aware that their wives were in love with another man? Nevertheless, they were steadfast, presumably because they knew the fling would not endure for long.

On the other hand, we have the example of Hosea and Gomer. God wanted Hosea to marry her, despite her adulterous nature. So we should perhaps admire the cuckold for his faithfulness.

On the other hand, it is said that the angels rejoice when a man marries a prostitute and leads her to virtue. The ideal of the good-hearted prostitute persists into our own time, as the success of the movie, \emph{Pretty Woman}, attests.

\paragraph{The Unhappiness of Women}

Despite the overcoming of the double standard, about 20 to 25\% of American women take anti-anxiety or anti-depressant medications. Another group self-medicates with alcohol, marijuana, or illegal drugs. Women were supposed to have become deliriously happy since they could engage in sex at will, take birth control, and abort unwanted pregnancies. The sexual revolution has not been an obvious blessing.

\paragraph{Loss of Interest in Sex}

At a lunch meeting with some old friends, a woman mentioned a ``study'' that claimed that youths were engaging in less sex. She thought that was marvelous, assuming that young people were returning to some sort of virtuous life. She was shocked, however, when I explained that the lack of interest in sex was due to masturbation. These youths are jerking off to online pornography.

When I was a youth, birth control was illegal except for married couples in Puritanical Massachusetts. When I got my driver's license, I drove to New Hampshire to buy a pack of condoms. I kept one in my wallet, ``just in case'' the right occasion arose. A few years later, the day --- or rather night --- arrived. When I opened the package, the entire condom fell apart. I did not take into account that body heat would eventually dissolve the latex.

\paragraph{Starship Troopers}

The film, \emph{Starship Troopers}, has often been described as ``fascist''. That term is falsely used to describe any sort authoritarian rule. The correct term is ``tyranny'', since Fascism has quite a different agenda.

Actually, the story is based on ideas from \textbf{Plato}'s \emph{Republic}, whether consciously or unconsciously, I cannot say. The hidden Guardians rule the state. The Producers are those in trade, commerce, etc. Although the Producers can achieve personal wealth, they have no part in the political life of the State. They are regarded as weak and their lives as dishonourable, lacking in meaning.

In between them, there is the group of auxiliary Guardians, or warriors. As in the Republic, one does not become an auxiliary through birth, but rather in must be earned. Another feature is that warrior births are regulated by the Guardians, so the warriors require a license to bear a child. In the Republic, the auxiliaries, both men and women, would do gymnastics together in the nude. That is shown in the famous co-ed shower scene in the film.

Some may despise that political arrangement, even though it was successful in repelling an invasion of alien intelligent bugs. Democracies most likely could not accomplish that today.

\paragraph{Mediocrity and the American Dream}

There is constant talk in the media about achieving the ``American Dream'', which, as far as I can tell, is to be a prosperous consumer with no higher values. After observing the behaviour of American politicians, it seems that the American Dream is that those of mediocre talents and virtue can obtain wealth and prominence.

A Korean professor claims that space aliens are mating with earthlings in order to bring about a superior race that will ``save the planet''. I heard that there is a dating app for who would like to hook up with those aliens. So even if you yourself are mediocre, perhaps you could have an indigo child\footnote{\url{https://en.wikipedia.org/wiki/Indigo_children}}.


\flrightit{Posted on 2019-05-02 by Cologero}
