\section{The lover becomes the thing he loves}

\textbf{Luís Vaz de Camões} was Portugal's greatest poet, on the order of Shakespeare, Homer, Virgil, and Dante. He wrote several sonnets in the style of Petrarch. And like Petrarch, he seemed to have had a mysterious muse, as this beautiful sonnet reveals. All attempts to identify her are inconclusive. In any event, such attempts are doomed by the fault of literalness; as such, they miss the deep symbolic meaning of the sonnet.

The demigoddess is revealed in Camões's soul. The sonnet reveals his sense of longing or Sehnsucht, which can only arise from experience. Camões was not just an academic, as he lived an active life as a bohemian, traveler, and adventurer, or better said, a real Renaissance man.

\newlength{\saveleftmargini}
\setlength{\saveleftmargini}{\leftmargini}
\setlength{\leftmargini}{0em}

\begin{table}[h]
\begin{tabular}{cc}\footnotesize
\begin{minipage}{.5\textwidth}
\begin{verse}
Transforma-se o amador na cousa amada,\\
por virtude do muito imaginar;\\
não tenho, logo, mais que desejar,\\
pois em mim tenho a parte desejada.

Se nela está minh'alma transformada,\\
que mais deseja o corpo de alcançar?\\
Em si sòmente pode descansar,\\
pois consigo tal alma está liada.

Mas esta linda e pura semideia,\\
que, como um acidente em seu sujeito,\\
assi co a alma minha se conforma,

está no pensamento como ideia:\\
o vivo e puro amor de que sou feito,\\
como a matéria simples busca a forma.
\end{verse}
\end{minipage}
&
\begin{minipage}{.9\textwidth}\footnotesize
\begin{verse}
The lover becomes his beloved\\
by virtue of much imagining;\\
since what I long for is already in me,\\
the act of longing should be enough.

If my soul becomes the beloved,\\
what more can my body long for?\\
Only in itself will it find peace,\\
since my body and soul are linked.

But this pure, fair demigoddess,\\
who with my soul is in accord\\
like an accident with its subject,

exists in my mind as a mere idea;\\
the pure and living love I'm made of\\
seeks, like simple matter, form. 
\end{verse}
\end{minipage}
\end{tabular}\end{table}

\setlength{\leftmargini}{\saveleftmargini}% restore original value

by \textbf{Luís Vaz de Camões}, 1595

Translation: © 2006, Richard Zenith



\flrightit{Posted on 2021-12-22 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Presbyter Johannes on 2022-06-26 at 00:24 said: }

This translation is significantly inaccurate. Is this a professional publicised translation?


\hfill

\texttt{Cologero on 2022-06-26 at 13:40 said: }

Yes, it is a professionally published translation, because I don't know Portuguese. Poets often take liberties with translations, so I usually do my own whenever possible.

However, Portuguese is a Romance language, so I get the drift of it. It is certainly not ``significantly" inaccurate.


\hfill

\texttt{Presbyter Johannes on 2022-06-26 at 15:49 said: }

It's good that you can tell.


\end{sffamily}\end{footnotesize}
