\section{Doctor Zhivago Unveiled}

\begin{quotex}
Then lovers hastily are drawn

To one another, vague and dreaming, \flright{\textsc{Boris Pasternak}, \textit{Unique Days}}

\end{quotex}
There is no need to repeat the details. Doctor Zhivago\footnote{\url{https://www.imdb.com/title/tt0059113/?ref_=nv_sr_srsg_0}} was a novel written by Boris Pasternak, which was banned in the Soviet Union, smuggled into Italy, and finally published in 1957. This is a review solely of the film version, which deviates from the book in some ways. Despite its length, the movie was very popular when it was released.

\begin{quotex}
You lay life on a table and cut out all the tumors of injustice. Marvelous. Ah, but cutting out the tumours of injustice, that's a deep operation. Someone must keep life alive while you do it. By living. Isn't that right?

Only the solitary seek the truth, and they break with all those who don't love it sufficiently.

All of the Communist Parties, upon attaining power, have become completely merciless. But at the stage before they achieve power, it is necessary to use disguises.

\flright{\textit{quotes from Doctor Zhivago}}

\end{quotex}
The story is ostensibly about a Doctor Zhivago, who was a general practitioner like Dr. Harvey — after all, specialization is for insects. He also wrote popular poetry; you all know a doctor like that. Part of the movie is about the individual lives of the main characters and the other part is about the historical events that affect their lives in oh so many ways. 

The personal part glamorizes the doctor's adultery, infidelity, and fatherhood of an illegitimate daughter. He abandoned his son by his wife, who was ``inconsolable" to be without a father. He also abandoned his daughter with his enticing mistress, Lara, played by more than beautiful Julie Christie. In the end, he lost both his wife and his mistress. Years later, old and in poor health, the Doctor, looking out through the window of the street car, spots Lara wandering on the street. He disembarks, hoping to have his Sehnsucht satisfied. While chasing after he, he suffered a heart attack. He died alone in the streets, watched over by strangers. But, heck, love the one you're with and it's all good. He turned out to be more popular dead than alive.

\paragraph{Historical Background}
While the Doctor was on the German front during World War One, the Bolshevik revolution was nearing completion back home. When Zhivago returned home, he was surprised to find 13 other families living in his home. He, his wife, child, and father-in-law were confined to a small upstairs room without sufficient fuel to keep warm.

Marx is dangerous not because he is wrong, but because he is often right. Marx predicted that a late capitalist system would smoothly embrace communism as the next step. Lenin had readjusted expectations and believed that a revolution would be necessary to make the transition. It turns out that Lenin was wrong, and Marx may still be right.

Matter, in the Marxist sense, is not the matter of the physicist. It has a direction that is expressed through the world process. The sequence starts with feudal society, enters the capitalist phase, and culminates in the equality of communism.

The process is ineluctable and there is no meaning to free thought. One's worldview is a creation of the material process, so there can be no thought that transcends it or judge it. If you think the revolution, you are the vanguard of the revolution. Otherwise, you retard it so your thinking needs to be readjusted.

Bourgeois thinking cannot lead to happiness, since it depends on subjectivity and inequality. Since the material circumstances of life creates thought, then one's thought becomes free only when the social structure in consistent with the absence of coercion by capital. With inequality eliminated, so also is envy, and everyone should be happy.

\paragraph{The Bourgeois Poet}
On Schopenhauer's principle, the important thing is to understand the inner states of the characters in a book or a movie. Hence, we want to understand the inner life of both the revolutionaries and the bourgeoisie. The remarkable thing is that we can see the same inner life in the revolutionaries of our own time.

The Doctor's alter ego is the mysterious Strelnikov, the idealistic revolutionary and husband of his mistress Lara. Strelnikov turns out to be too intense even for the Bolsheviks. Strelnikov confronts the Doctor and explains the revolutionary mindset. The Doctor is too bourgeois with his poetry, since it is concerned with his feelings and inner life. The revolutionary needs to be guided by the forces of historical materialism. After some tense interrogation, Strelnikov lets him go. We are led to believe that, in a moment of weakness, Strelnikov lets his past association with the Doctor affect his judgment.

In the course of the story, the Doctor seems to be frequently menaced by the authorities. Nevertheless, he remains free, even publishing his poetry to popular acclaim. On the other hand, the two women in his life are not so fortunate. His wife flees to Paris with their children, and Lara escapes to the East, abandoning their daughter in the process.

In the book, the Doctor starts a new family, which he also abandons, and Lara returns to Russia after his death. The film version realizes that Lara has to be the love of his life, with no other. Hence, he catches one final glimpse of her. We may hope that Lara is his Beatrice, the revelation of Sophia, assuming that his life and poetry had made him wise.

\url{https://www.youtube.com/watch?v=M1iQ5hQTR5s}



\flrightit{Posted on 2021-11-21 by Cologero }
