\section{Fundamentals of Sex Magick}

\begin{quotex}
The road of excess leads to the palace of wisdom.
\flright{\textsc{William Blake}}
\end{quotex}

Several would-be esoterists have tried to convince me that sex magick, particularly the Crowleyian version, is the fast track to enlightenment. They do so apparently on the grounds that since their social life already revolves around sex and drugs, that therefore it would be a natural fit for them. If it were that simple, then any American inner city would be a centre of spiritual enlightenment. Even an experienced Mason has told me that the search for the Grail is really \emph{chercher la femme}, on the grounds that the shape of the cup reminds him of a vagina. I could only conclude that we travel in different circles.

Although there are reasons for limiting the male orgasm --- and not out of prudery or some Victorian moral code --- there is no doubt sexual energy has been used as a tool for self-realisation. Since Crowley has left records of his experiences, it seems useful to pay attention. First of all, Crowley's ``method of science'' means that the sexual experience --- even one's own --- must be treated with detachment and objectivity. Any one who has attempted this will notice that regarding a sex act as merely physiological significantly changes the experience itself; he will also notice how much of the sexual experience is controlled by the imagination and not the biology.

Crowley's essay \emph{Liber DCCCXI: Energized Enthusiasm} describes his experiences. He points out that the ceremonial use of sex has restrictions that don't apply to the natural physical purpose of sex.

\begin{quotex}
If the act in itself is to be the sacrament in a religious ceremony, this act must be accomplished solely for the love of God. All personal considerations must be banished utterly.
\end{quotex}

And again,

\begin{quotex}
I need hardly emphasize the necessity for the strictest self-control and concentration on their the celebrant's part. As it would be blasphemy to enjoy the gross taste of the wine of the sacrament, so must the celebrant suppress even the minutest manifestation of animal pleasure.
\end{quotex}

Crowley further notes:

\begin{quotex}
It is, indeed, of the first importance for the celebrant in any phallic rite to be able to complete the act without even once allowing a sexual or sensual thought to invade his mind. The mind must be as absolutely detached from one's own body as it is from another person's.
\end{quotex}

Nowadays, Tibetan Buddhism is a rival for the sex magick crowd. Its appeal to its Western adherents is that it is cooler than their parents' religion and compatible with science. Then there are all the thangkas depicting the yabyum, again depicting something they fancy themselves to be skilled at when those interminable meditation sessions become fatiguing. Never mind that Tibetan Buddhist sexual morality is identical to that of the despised Roman church; they presume that is a mere detail that arises from the culture of Tibet and not fundamental to Buddhism.

Nevertheless, the Dalai Lama asserts that initiation into Buddhist Tantra is reserved to those who have achieved total indifference to experience. He gives the example that if you don't care whether you are served an elegant meal or a serving of faeces with a glass full of pus, then you are ready for Tantra. It seems that Crowley may have taken this dictum quite literally.

In his book \textit{Carnal Alchemy}, \textbf{Stephen Flowers} notes that on one particular sex magick session, Crowley was ``forced into an act of coprophagy as a kind of Holy Eucharist.'' (p. 20) Now Dr. Flowers is a brilliant linguist, a promoter of Germanic languages, and somewhat kinky himself. Nevertheless, he prefers the clinically sounding Greek term, rather than the more colourful Anglo-Saxon equivalent. I myself prefer the elegant French --- \emph{manger la merde d'amour} --- since it sounds as though it deserves a glossy monthly magazine with scratch-off sniffers.


\flrightit{Posted on 2008-05-21 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Daniel on 2010-09-27 at 22:56 said:}

``he prefers the clinically sounding Latin word caprophagous ''

... you mean, of course, the clinical-sounding GREEK word.

\hfill

\texttt{Leonardo Cavalcanti on 2020-09-07 at 11:57 said:}

Crowley was in many ways a charlatan. If he wrote many things that are true, that's because he was merely repeating, but what he really interested in was the counter culture of the subverted West, he acted as an agent of subversion, proving that he was the worst kind of charlatan, because he acted as a wolf in sheepskin.

\hfill

\texttt{Cologero on 2020-09-07 at 12:25 said:}

I wouldn't disagree. I told a thelemite that Thelema is the de facto religion of the West, or at least of the USA. He couldn't see it.

\end{sffamily}
\end{footnotesize}

