\section{Stages of Initiation}

\textbf{Louis Claude de Saint-Martin} separated men into four categories.

\paragraph{Man of Torrent}
The Man of Torrent is caught up in the world process. He accepts things as they appear. He identifies with his thoughts, feelings, likes, dislikes, and desires. However, since they are so variable, he lacks a stable identity, thinking one thing today, another thing tomorrow, willing one thing today, another tomorrow.

\paragraph{Man of Desire}
The Man of Desire is no longer satisfied with ordinary life and the desire for something else begins to burn in him. He is a seeker, he reads and studies metaphysical ideas, he questions consensus opinions about religion and political systems. He learns to meditate. He challenges and tests himself, physically, emotionally, intellectually. He seeks to overcome the prison that fear, lust, greed, and guilt keep him in. Still tied to the physical, he is still tempted to enter debates with the men of torrent, believing that his rationality will persuade them. In many areas, he is still attached to the tastes and values of the men of torrent.

\paragraph{The New Man}
The New Man is more developed and avoids the errors of the men of desire. He has calmed many of the distractions that perturb consciousness. Through introspection, he gains self-knowledge. This allows him to understand human nature and how man is moved by the random thoughts and feelings that pass through his mind.

\paragraph{Spirit Man}
Or Guenon's True Man\footnote{\url{https://gornahoor.net/?p=1039}}. This man has realized all his possibilities as a man. He has mastered the personal element: the arbitrary thoughts, feelings, likes, dislikes, and desires no longer perturb his consciousness. He is in the Primordial State and his consciousness is a clear reflector of transcendent ideas. No longer tormented by doubt and inconsistent thoughts, he has true Will. Since he wills one thing, he is pure of heart. He avoids inter-religious debates, since he follows the primordial tradition. He is unconcerned about solar or lunar initiations, since he has undergone both.



\flrightit{Posted on 2010-11-10 by Cologero }
