\section{The Dawn of the Gods}

We all know about the Twilight of the Gods and how they have abondoned us, but few dare to ask where they came from in the first place.

\textbf{Auguste Comte} proposed that humanity has progressed from belief in polytheism to monotheism, then to an all embracing metaphysical or ideological principle, and finally to positivism. While the sequence may be somewhat correct, it represents not progress, but rather a fall. There is a primordial consciousness of unity, inwardly united with the divine Logos. The role of the human being is to bring this unity into factual realization in nature, or manifestation. As such, the human being is divided, with the consciousness of unity that unites him with God, and a material existence that unites him with the natural world.

By rebelling against the divine principle, human beings fall under the power of the material principle. The Logos was once the center of consciousness, but afterwards is experienced as an external law and hence a burden. Enthralled by the material world, he loses his center and becomes just another entity in nature. Initially, he experiences this disunity in consciousness and the idea of multiple gods, or polytheism arises. Rather than seeking unity back at the source, he looks outside himself and creates various philosophical and ideological systems based on an external principle, be it matter, thought, evolution, genetics, race, and so on. Still dissatisfied, he thinks that the facts and the laws explaining these facts will suffice. But not accumulation of facts of or theories about nature can recover the lost unity.

In this passage, \textbf{Vladimir Solovyov} describes how the gods arise in the cosmogonic process:

\begin{quotationx}
For consciousness that has lost the inner unity of the all in the divine Spirit, the only external unity that becomes accessible is that produced by the cosmic action of the divine Logos upon the world soul, which is the matter of the cosmic process. The consciousness of humanity seeks to reproduce in itself those determinate forms of unity that had already been generated by the cosmogonic process in material nature. The unifying forces of material nature (the offspring of the Demiurge and the world soul Anima Mundi) appear now in consciousness as principles that determine it and give it content. These forces gradually manifest themselves and reign in consciousness as lords not only of the external world but even of consciousness itself, as genuine gods. Thus, this new process is, first of all, a theogonic process. This does not mean, of course, that these dominant principles were created in this very process, nor does it mean that humanity invented its own
\emph{gods}. We know that these principles existed prior to humanity as cosmic forces.

In that capacity, however they were not gods, for there are no gods without worshippers. They become gods only for the human consciousness that recognizes them to be such, after it has fallen under their dominion as the result of its separation from the one divine center.
\end{quotationx}


\flrightit{Posted on 2010-04-20 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}

\texttt{Cologero on 2010-04-23 at 22:47 said:}

If by polytheism, one means the absolute and irreducible plurality of principles, that is no part of any Traditional teaching.

Polytheism is not the denial of monotheism, but rather the polymorphous manifestation of the Divine unity. That manifestation takes the form of a celestial hierarchy, which is experienced as various cosmic forces. When they become objects of worship, they are called gods. Of course, then, there are gods for all aspects of life, starting from the national gods which give identity to a tribe, nation, or race.

However, all these forces are ultimately ordered by the Logos.

Therefore, polytheism is inconsistent and ultimately incoherent. Polytheists claim that they are most open to and tolerant of foreign deities due to their doctrine. However, they cannot be tolerant of monotheism, since that would destroy the whole basis of their beliefs.

On the other hand, monotheism is consistent with polytheism, properly understood. Of course, it rejects the ``worship'' of these cosmic forces, but not their reality or importance. It is only unorthodox and non-Traditional deviations of monotheism that fail to understand this.

\end{sffamily}
\end{footnotesize}

