\section{Prolegomena to any Future Tradition}

As a thought experiment, we shall attempt, in several posts, to specify the characteristics of what Tradition in the West will have to look like, based on the 400+ preceding posts. In particular, we will look at the following elements:

\begin{enumerate}
\item Foundation 
\item External structure 
\item Internal Structure 
\item Practical Dualism 
\end{enumerate}
These characteristics will be looked at from the point of view of a spiritual race. Features based on biological or psychic race are not part of this consideration, since they are contingent and arbitrary. Specifically, the originary races no longer exist as such, due first of all to migrations and secondly because the race of the soul alters over time.

\paragraph{Foundation}
A tradition always has a divine or supernatural founder who is its lawgiver and formative spirit. Thus any and all attempts to build on merely human thought forms are destined to fall short. This divine person reveals the cosmic law and is the unifying factor among the various segments of the members of the tradition.

\paragraph{External Structure}
The external structure follows the divine plan. There must be a spiritual authority that understands, interprets, and teaches the revealed law. Beneath that is the temporal power that implements the law for the benefit of all, justly, according to the merits and needs of the members. A producing segment provides for the economic well-being of the group. This requires a caste based system of the regal-spiritual caste, a warrior caste, and a productive caste. An Internet comment brought to our attention claims that “Gornahoor loves castes”. It is not a question of “loving” castes for their own sake, but rather a love for the revelation of cosmic law. A caste system is the simple recognition that people have different talents and interests and thus need a place in society where they fit in. The opposite of this is so-called egalitarianism, which is never such in practice since some will attempt to gain power by unprincipled means based on money or mere physical prowess.

\paragraph{Internal Structure}
Men will understand and experience themselves as threefold beings: spirit, soul, and body. The soul and body are natural elements, irreducible to each other, while the spirit is transcendent. The spirit, or rational soul, understands the external and internal structures directly and intuitively, without the intermediary of discursive thought.

\paragraph{Practical Dualism}
Men learn to fight the greater and lesser battles, the former for their individual souls, the latter for each other. Every man capable engages in this battle in his own consciousness, siding with the forces of the cosmos against those of chaos. He must learn to assert the primacy of the spirit over the soul, and the soul over the body. His tools are meditation, prayer, rites, sacraments, clairvoyant vision, hence this can never be reduced to a mere system of thought. We call this practical dualism, since it is part of this experience, although it is not a real dualism. Chaos had no independent existence, but as long as we experience is as resistance, it is a privation, not a thing in itself.

To be victorious, a man must know how to think, how to fast, and how to wait. While instrumental and logical thought have their place, man must, however, learn again how to think. Discursive thought, with each thought connected to another, creates a tight web, like chain-mail, that obscures the light of Being. Thoughts need to come in isolation and understood more as acts of perception rather than as a secondary experience of being.

While caloric fasting leads to loner physical life, mind fasting is the path to eternal life. Thus, the first trial is to purge the mind of opinions, beliefs, theories, unexamined assumptions that are connected together in the flow of consciousness. When freed of this lower mind, then the mind of spirit can gain gnosis, a direct intuition of the world of ideas. Man must also fast from the stories he uses to organize his life. Habitual ways of looking at the world, repeated tales of wrong doings, injustices, excuses, and justifications are dropped in the mind fast.

Then there is the waiting. When the mind is freed of false ideas and emotional perturbations, in can act as a perfect reflector of the Spirit, allowing it to come into consciousness. There is no technology for this, no technique, no drug, no practice, no gene, no new theory, no weekend seminar, no new age “mind treatment”.

\flrightit{Posted on 2011-09-26 by Cologero}