\section{The Evil Eye}

It is nice to know that once upon a time in Europe, Emperors, Kings, Knights, Templars, Hermetists, Saints, Mystics, Troubadours and so much more were abounding. Now we know there were also \textbf{Sorcerers} and probably there still are. I don't recall \textbf{Rene Guenon} mentioning the activities of sorcerers in any of his ``official" writings, but he dwells on the topic in three letters to \textbf{Julius Evola}. It is safe to assume that the latter was open to the idea.

Yet, we must wonder where these sorcerers came from and what kind of initiation they receive. Regarding that, I can say nothing. Guenon provides with the knowledge of three levels of attack and their defense.

\begin{itemize}
\item \textbf{Impeccability}. The rarest of saints, such as Mary in Catholicism or the Prophet in Islam were impeccable, and hence immune to any sort of spiritual attack. 
\item \textbf{Psychic defense}. Those with a true spiritual vocation are immune from psychic and mental attacks. 
\item \textbf{Corporeal attacks}. It seems that no one is totally immune from spells and curses on the physical plane, although the cornu is said to ward of the curse of the evil eye (just in case any one is getting any ideas). 
\end{itemize}
Now the motivation for such attacks on Guenon, or even Evola, is unclear. Guenon hints that Evola's injury may have been the result of such forces. However, Evola's risky behaviour is probably sufficient explanation. Leon de Poncins, on the other hand, certainly had enemies, since he wrote about alleged Masonic and Jewish conspiracies against the Church. Nevertheless, the ill effects were not long lasting against him. Perhaps the nuns prayed for his recovery.

Guenon was often in poor health and it seems he didn't attribute it solely to natural causes. Since he was never overtly political or militant, I don't know which sorcerers would target him. In particular, the incident with the Jewish lawyer looking for his photograph seems odd. Guenon most likely had a secret. His family never allowed his papers released; that is why we have just one side on the Guenon-Evola correspondence.

I can relate a personal incident. A few years ago in Paris, after a day in the Louvre, my pretty hostess and I stopped at a nearby cafe to relax \emph{al fresco} with some coffee and sweets. Shortly thereafter, a wizened Gypsy woman on the street stopped at our table and asked for a handout. Apparently, she had never learned that ``two's company and three's a crowd", so I tried to ignore her. She kept mumbling, but I could not understand her heavily accented French. Perhaps on another day, I may have dropped a Euro on her, but at this particular moment, my mind and attention were elsewhere, so my heart hardened and I ignored her. She nevertheless persisted and as she was leaving, I heard her utter some malediction. Slightly astonished, I muttered a Hail Mary under my breath, and continued on with my date.

Coincidentally, I spent the next two years with a series of ailments, in and out of hospitals. Unfortunately, I wasn't wearing my father's cornu at that time in Paris, but I do now. (actually, I don't: 20 Aug 2020)



\flrightit{Posted on 2012-08-20 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Audrey Del Sol on 2012-08-22 at 01:06 said: }

Now the motivation for such attacks on Guenon, or even Evola, is unclear. Guenon hints that Evola's injury may have been the result of such forces. However, Evola's risky behaviour is probably sufficient explanation.

I probably speak a lot, yet. I am of the opinion that Evola's injury was on such basis; hence I brought up Franz Bardon \{whose external expression was so `flowery' it almost attracts froufrous\}.

Physical level harm comes out of the fact that these ``occult" forces are invariably concerned with

material goods and they seek to destroy healthy, independent human bodies. It has to deal with mechanization of humanity also mention ed by G. I. Gurdjieff. Modern medicine and popular abuse of computers fall under this category. Can elaborate if requested.


\hfill

\texttt{Audrey Del Sol on 2012-08-22 at 01:12 said: }

Vicomte Léon De Poncins was a French journalist descendant of a noble family. 

He describes Freemasonry: ``.. , to establish in its place an atheistic rationalist society which, in fact, leads straight to materialism although it is supposed to have science and reason as a religion. The role of Freemasonry is to create the revolutionary state of mind rather than to act directly."

There you go. Evola (and I admire him for it, hence my adamant defense) was a man of action who DID.

Is this ``risky"? As opposed to abusing female torso's, arms, and hands as ``babycribs"?


\hfill

\texttt{Audrey Del Sol on 2012-08-22 at 01:17 said: }

Also, hence my such physical presence. In our ``intellectual baby" state, when bodies are atrophied and malnourished heads rule, it is so easy to ``brainwash" people to ``think" things and have their bodies act accordingly. Therefore, I also temporarily believe in being loud, and telling people when they are being merdre, despite the fact that it goes against good manners.


\hfill

\texttt{Cologero on 2012-08-22 at 08:10 said: }

OK, Audrey Del Sol, I confess to often missing your allusions. I searched for ``babycribs" and found nothing about the type of abuse you describe.

I'm not sure what Evola DID, perhaps at one time he was active, but late in life he became more passive (e.g., just to ride the tiger and survive.) Nevertheless, the real task at this moment in time is to create the ``counter-revolutionary" state of mind. That is the greater battle that must precede the lesser battle. 

Feel free to elaborate on occult forces and popular abuse of computers.


\hfill

\texttt{George on 2013-02-03 at 09:46 said: }

There is a large ``transhumanist" agenda, that argues essentially:

Progress is impossible to stop, but humans are handling it well. Therefore, we need to adapt humans to progress by the ``next evolutionary step". Whatever that is… it seems they want to use technology and genetic engineering to create the new human.


\end{sffamily}\end{footnotesize}
