\section{Sex and Tropism}

\begin{quotex}
Man is directly a natural being. As a natural being and as a living natural being he is on the one hand endowed with natural powers, vital powers --- he is an active natural being. These forces exist in him as tendencies and abilities --- as instincts. On the other hand, as a natural, corporeal, sensuous objective being he is a suffering, conditioned and limited creature, like animals and plants. That is to say, the objects of his instincts exist outside him, as objects independent of him; yet these objects are objects that he needs --- essential objects, indispensable to the manifestation and confirmation of his essential powers.
\flright{\textsc{Karl Marx}}

Sexual behavior is a complex of chemotropism, or automatic reactions evoked by chemical stimuli in the bloodstream... Sex in its ordinary manifestations among civilized human beings is not the product of natural instinct that is generally supposed to be but is a chemotropism evoked or conditioned reflex (in Pavlov's sense) evoked in response to aphrodisiacal stimulation by foods and beverages, especially animal proteins, alcohol, coffee, and also tobacco. This tropistic reaction, in both its physical and psychical aspects, is subject to voluntary control through diet, an alkaline-forming, low-protein vegetable diet reducing it, while an acid-forming high-protein meat diet increases it.
\flright{\textsc{Raymond W. Bernard}, \emph{Nutritional Sex Control and Rejuvenation}}

All actions of the animal (including human) organism are determined by external and internal or hormonal stimuli, and the concept of purposiveness in behavior and free will must be completely discarded and replaced by one of automatic reactions in terms of quantitative laws of mechanical causation.
\flright{\textsc{Jacques Loeb}, \emph{Forced Movements, Tropisms, and Animal Conduct}}

From its first moment, we can say that the ``I'' still lives only as if in a dream: it is not yet a self-consciousness, nor an autonomous principle of action: immersed in an immediate, indistinct coalescence with nature and the world, we can say that it is not so much he who thinks, speaks, and asserts himself, but rather that various forces and impulses think, speak and assert themselves in him. He therefore is only a type of medium, a passive instrument that has his very life outside himself, and he experiences everything as grace, as spontaneity, as the immediate self-manifestation of something that transcends him.
\flright{\textsc{Julius Evola}}

Logic can descend even further. When this happens, human beings are beginning to use a type of logic that is directed from the subterranean spheres. This may even be an irrefutable logic and may have an enormously convincing effect ... Then, one no longer thinks independently, but Ahriman is thinking through that person.
\flright{\textsc{Valentin Tomberg}}
\end{quotex}

Recently, I was interviewed by a team of Martian scientists and metaphysicians who were studying human life on earth. They were curious about the objects we've been sending to their planet; in particular, they wondered why we didn't just use astral travel. You see, the Martians never went through the process of solidification experienced by earth humans. Thus they are able to see the ``inner'' nature, or essence, of things. That they regard as much more important than phenomena and mere appearances. They can discern the inner states of conscious beings through a process they call ``external considering''.

They explained that metaphysical principles are true always and everywhere and that gave us somewhat of a common language. I discussed certain texts like the Tao Te Ching, the Vedas, Medieval philosophy, and so on. They listened with interest but then conceded that, although they were ``in the ballpark'', they were still quite primitive. Nevertheless, they prepared their final report using the terminology I had given them. Since it provides some suggestions for mystical evolution, there may be some value in it. It has the advantage of providing and objective evaluation of the human condition

I have no hard or electronic copy, but they gave me permission to publish the main points of their report from the Akashic record. Those texts will be marked with a green vertical line.

\begin{quotationx}
Because of the solidification of their world, the earthlings have grave difficulties with metaphysical concepts. Their thought patterns are stuck on discursive thought and direct intuition does not come easily. Actually, it almost always depends on sense experiences first followed by a difficult climb to grasp the essence. Hence, they tend to see timeless and spaceless concepts as mere extensions of phenomena. They then misunderstand their myths and symbols as referring to physical events.

With all that in mind, this initial report will focus on the earthlings as biological beings. Like many other beings in the galaxy, their physical organism is DNA-based. Although their biologists have made some progress in analyzing DNA, they have not advanced at all from the ancient texts in understanding the interiority of their being. This creates inner conflict. Therefore, we have provided some suggestions to match their inner and outer natures and have encouraged our human contact to disseminate this material widely.
\end{quotationx}

\paragraph{General Law}

The Macrocosm is the same everywhere; that is why the Martian experience is of value to us. What they reported next sounds very similar to ideas from the Hermetic revival of last century in figures like Evola, Mouravieff, and Tomberg. They would neither confirm nor deny any prior communication with those earthmen. As Rene Guenon points out, Life is one of the essential elements of the manifest universe. Hence, certain life issues arise in all life forms. The report continues:

\begin{quotex}
In the universal phenomena of life, there are three infra-rational factors that maintain the existence of life. In their teachings, these factors are related to what they call chakras:

\begin{itemize}

\item \textbf{Fear}. The fear factor is centered in the root chakra; this serves to protect the organism from threats.

\item \textbf{Sex}. The sex factor, centered in the chakra of the sex organs, ensures the creation of new and distinct organism.

\item \textbf{Hunger}. The hunger factor, centered in the chakra of the solar plexus, enables the organism to maintain and improve its state of existence.
\end{itemize}

In a healthy being, these factors serve for the benefit of the organism. Yet, due to the solidification of their planet and life forms, these factors come to dominate the consciousness. Each of them has an intellectual component, which serves their own purpose with a level of intelligence. Unfortunately, what seems rational from the perspective of a given chakra is far from rational for the being as a whole. Hence, earthlings often get themselves involved in situations that seemed expedient from a limited perspective, yet eventually result in serious problems.
\end{quotex}

\paragraph{Tropism}

The Martians were very interested in our discussions about tropism. They certainly understood the concept, yet had trouble imagining the state of consciousness of a being dominated by a tropism. I suggested that the law of analogy may be helpful, if tropism is understood as an attraction that is expressed on the physical level as electrical attraction, at the lowest life level as tropism, then as sexual attraction, all the way to the attraction of the metaphysical dyad. Here is their report:

\begin{quotationx}
At its most basic level, tropism begins in plant life as a response to a stimulus. Although many tropisms have been observed, the best model is the heliotropism of a sun flower which bends during the day to follow the movement of the sun. In addition to the normative expression of the tropism, two variations may arise: the artificial and the abnormal. In the former, there is an artificial stimulus that mimics the normative stimulus and the abnormal, which is a deviation of the response to a natural stimulus.

For example, a plant may respond to an artificial light rather than the sun. In the worst scenario, the plant will turn away from the stimulus because of a genetic mutation. Most often, but not always, that will result in the failure of the plant to pass on its genes.

There is an interesting twist in tropisms for earthlings. The tropisms, which operate at the lower centers, interfere with the thought process. This interference causes the being to believe it is consciously choosing the stimulus response, so it then fabricates confabulations to attempt to explain its response. Our earth contact explained to us that it is quite difficult to lead such a being out of the intricate web of his confabulations once they become deeply rooted.
\end{quotationx}

\paragraph{Hunger and Fear}

The Martians gave me some interesting suggestions about control of the lower chakras which are the centers of hunger, sex, and fear. All intelligent life forms have a foundation on these general laws of life although their effects may be quite different. Specifically, as in the description of tropisms, these laws should result in intelligent and normative effect. Unfortunately, in the human condition, they often take an artificial, factitious, or even abnormal path. I will have more to say about this in a follow up post, but I will mention here the key point.

Many systems try to raise inner psychic energy, often called kundalini, from the bottom up; i.e., from the root, sex, and solar plexus centers, and then to higher centers. The results of that have seldom been good. They suggested, on the other hand, that the movement should be in the exact opposite direction. Specifically, there needs to be mastery over the sensations of hunger, sexual attraction, and then fear, in that order. This is a downward movement of the higher intellectual part of the being to dominate the lower parts.

\paragraph{Sex}

This is always a difficult topic since everyone, it appears, has an opinion about it. Besides, I often hear about it in private communications and comments, many of which need to be rejected. There are several stages of indifference, which will be discussed in a forthcoming post. (Some of these stages are discussed in Love is but a dream\footnote{\url{https://gornahoor.net/?p=3356}}, and in the previous chapters.) One of the last stages is the ability to discuss this topic with no emotional response. That is called the ability to ``drink poison and handle snakes''. Nevertheless, for others it may be a remote, or even a near, occasion; some of them will unsubscribe from mailing lists or ``unfriend'' me.

Nevertheless, the objective, scientific viewpoint of the Martians, based on solid metaphysical principles, will ultimately prove of value. The Martians bring birth to new beings consciously. That is, they take into consideration both heredity and the environment before initiating conception. They are astounded by the low level of our sexual development. They attribute it to the lush conditions for life, for which the ``fittest'' is a rather low bar to reach.

\begin{quotationx}
The sex of life of earthlings is pursued independently of metaphysical principals for the most part. There is some understanding, according to our contact, of the metaphysical dyad, which is the archetype for all the polarities of manifestation. But for most, sex is little more than a biological function, a natural force like any other, which can be used for any suitable purpose. Even those who can discern a higher source often remain on the psychical level rather than the spiritual or metaphysical.

The male and female earthlings are manifestations of the metaphysical principle of the dyad. Even still, since the earthlings have not mastered the art of manifestation, this process is often incomplete, resulting in a privation of the primordial dyad. That is, for some beings there is an imperfect sexualization which causes psychological distress and even social turmoil.

There is also huge waste of energy involved in the sex function, both material and psychical. The females lose their eggs every month. The males lose their sperm involuntarily through nocturnal emissions or, more often, voluntarily with no real biological purpose.

Along with physical dissolution, the psychic energy, which could have been sublimated to a higher purpose, is likewise dissipated.
\end{quotationx}

Since I can't reproduce the entire report here, I will highlight the main points. Keep in mind that sex for the Martians is understood on two levels. The first is the means to create a new being on the biological level. The second is their equivalent of the androgyne in which the dyad is resolved in their individual consciousness. The pleasure and joy associated with that process is so overwhelming that the physical sex act pales in comparison. Hence, they use it for its natural and organic function.

Of course, our own Traditions are not unlike the Martians', at least in principle. The Yoga Sutras require sexual restraint. Buddhist initiations prohibit masturbation, oral, and anal sex. The Western tradition is not unlike the Tibetan in that regard.

To make it simple for me, they explained it in terms of tropisms, or orientations. Subsequently, I found some biologists who explain sexual activity in terms of chemotropism. Hence, the distortions in the imagination and thinking centers caused for sexual desire are merely confabulations. Looked at objectively, the biological function of sex is simply the continuation of the species. All other functions are then artificial, factitious or abnormal, as difficult as that may be for us to accept. From the male point of view, they provided the real tropisms of orientations, in order of increasing specificity.

\begin{enumerate}

\item Orientation toward a vagina (rather than an ersatz substitute for it)

\item Orientation toward a specific vagina (i.e., that of one's polar opposite, rather than a random female)

\item Orientation toward fertilizing an egg (rather than for pure recreation)
\end{enumerate}

That sounds quite convincing on an abstract level, but quite difficult to accept for everyday life.


\flrightit{Posted on 2015-10-22 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Sparrow on 2015-10-22 at 05:40 said:}

Perhaps you are simply joking and I'm just too dim to grasp, but are you serious about being contacted by Martians? As someone who's been wondering about this for a few years, I find it extremely interesting.

\hfill

\texttt{Aleksandar Gueric on 2015-10-22 at 08:20 said:}

I have been contacted by Jupiterians few years ago, but it just didn't made any sense. They were a matriarchal society -- on Jupiter (at least, I understood it that way), and told me that it would take 93 generations for humans to be able to live in there. Then they just showed me mental images of planet Earth and talked more, but I cannot remember anything. I understand that it wasn't physical Jupiter nor did they meant living on physical Jupiter since it's a gas giant. I never gave much atention to all of that, but this post reminded me of it.

\hfill

\texttt{obscure on 2015-10-22 at 10:34 said:}

The Martians are to Cologero what the Atlantians were to Plato.

\hfill

\texttt{veritycacciatrice on 2021-10-24 at 11:42 said:}

A female perspective;

I can attest first-hand to the difference in quality of natures created through these top-down or bottom-up means of purely physical reproduction.

Lust in proximity to a man is sensed electro-magnetically, yes, drives one towards the masculine, pulling the flesh (and will) with it, but it can be disengaged easily enough with practice.

I wasn't practiced with the father of;

The first daughter, now 26, conceived with a virile opportunist- we were under the influence of Dionysian lower natures.

This daughter is wild, her current beau is equally wild, they are directionless and without discipline.

The second daughter, who turns 8yrs old next week, was consciously bred down to stars, hours of conception (none of this was discussed), diet of each parent, and specific physical and mental attributes raised in request to the heavens.

This daughter completes adult puzzles and reads/writes fluently in english and Italian. She has a vein of the wild streak but is disciplined and wise already.

Her father was capable of escaping every trap, disentangling himself from every snare and answering every metaphysical conversation-starter I set to judge his quality. I could see out the back of this fellows mind to another world, the greater universe, like the wardrobe in The Lion and the Witch. This man originally prompted gold sparks in the very air, which proved to be a sound augur.

We burned the last of our egos off each other. (Though being southern Italian- his is coded into each cell and renews endlessly.)


\hfill

\end{sffamily}
\end{footnotesize}

