\section{The Bondage of Spirits}

In the essay \emph{On Magic}, \textbf{Giordano Bruno} discusses the existence and action of non-human spirits. The practice of invoking god, demons, and heroes is the branch of magic called theurgy. He calls this the ``magic of the hopeless" because they often become the vessels of evil demons. A safer practice is to learn to command and control lower demons through higher spirits.

Valentin Tomberg describes theurgy in a similar way:

\begin{quotex}
With respect to the cult of the ``gods" and the iconolatry that this cult entailed, the ``pagan" initiates and philosophers saw in it the practice of theurgy, i.e. that of intercourse with entities of the celestial hierarchies either by raising themselves to them, or by rendering possible their descent and presence on earth.

\end{quotex}
As sources, he lists Plutarch, Plotinus, Proclus, and Hermes Trismegistus to show that the highest theurgy did not try to evoke demons. Tomberg is apparently an Euhemerist, as was Giovanni Boccaccio by the way. He explains the origins of the pagan gods this way:

\begin{quotex}
The ``paganism" of the poets—symbolic and mythological paganism — was, in so far as it was not a symbolic version of the wisdom and magic (theurgy) of the mysteries, a universal humanism. Its ``gods" were, truth to tell, human personages — heroes and heroines, divinised or poetised, who were prototypes of the development of the human personality

\end{quotex}
Our interest here is not in magic, but rather in understanding the sources of thought.

\paragraph{Thinking and the Laws of Physics}
Several years ago, we wrote about free will and the laws of physics\footnote{\url{https://gornahoor.net/?p=210}}, reaching the conclusion that free will and physics are compatible. The confusion is the failure to understand that free will is not the same as power. A person cannot break the law of Gravity, or any other physical law, through will alone.

Another confusion is the idea that everything that exists is ultimately reducible to physical laws. The human world is not the physical world. In all likelihood, most everything around you is an artefact, not a natural kind. Even assuming that all natural kinds are ultimately reducible to physics, artefacts are not.

That is because human activity is the result of thought, and there is no physical law of thought. That is true not just of artefacts, but of all of culture. That raises the question, then, about the origin of thought.

Some physicalists are persistent, and assume that thoughts are simply the epiphenomena of biochemical processes. In that case, the works of Shakespeare should be explicable in chemical terms, and a fortiori, by a physico-chemical law. That is because those works cannot arise by chance, as shown by the infinite monkey theorem\footnote{\url{https://en.wikipedia.org/wiki/Infinite_monkey_theorem}}.

The next attempt is to attribute thoughts to ``evolution". The social systems arising from thought are said to have favoured by the ``survival of the fittest". Of course, this just begs the question. An evolutionary explanation, in order to be scientific, must identify a gene, or complex of genes, that are associated with a given system. Not only has that not been done, no one is even trying. I volunteer for research for 23andme, and they have never asked me such questions. Moreover, different cultures would have to have genetically different populations. Yet such a genetic population can create a patriarchy and an anti-patriarchy at the same time. Besides, no one has yet explained how genes create thoughts.

The more educated seem content to attribute thoughts to ``culture" or ``society". Yet that hardly makes sense, since culture and society are the results of thought, not the creators of thoughts. How can such abstract notions inculcate thoughts in people?

We could, I suppose, remain content with the concept of the ``meme". It is useful as a description, but there is no indication of the source of the meme or why certain people, or groups of people, are more influenced by a given meme that others.

We could also speculate that there are unknown forces that create memes. That leaves us with three choices.

\begin{enumerate}
\item The force acts at random. If so, then it cannot explain the origin of the works of Shakespeare. 
\item The force is a rigid law. That goes against all experience, since thought is fluid. One thought does not necessarily lead to another. 
\item The force is controlled by a conscious being. This would explain why thought is neither random nor rigid. This is the esoteric, Hermetic, and Traditional explanation. 
\end{enumerate}
\paragraph{Giordano Bruno}
Giordano Bruno explains quite a bit about such conscious beings. He explains how spirits affect our inner life:

\begin{quotex}
Demons penetrate through bodies and initiate thoughts in us. The convey impressions directly to our internal senses, just as we sometimes seem to think of something suggested by the internal senses.

They send not only dreams, voices, and visions to be heard and seen, but also certain thoughts which might be hardly noticed. They communicate truths sometimes through enigmas and sometimes through sense impressions. Sometimes they may even deceive.

\end{quotex}
There are many types of spirits, which explain the different influences on human thought. Bruno elaborates:

\begin{itemize}
\item Some are brute animals and cause injury without reason. Mark 9:25 describes them as ``deaf and dumb", i.e., they are without reason, recognize no commands, do not hear threats or prayers. 
\item Demons which are fearful, suspicious and credulous. They do not distinguish the possible from the impossible, or the appropriate from the inappropriate. They are like humans who are dreaming and disturbed by fantasies. 
\item Wiser demons reside in pure air. They distort worship and prayers, create Illusions of fear, anger, religion, and introduce confusion and doubt. 
\item Ethereal spirits are pure and luminous. Not hostile. 
\item Aqueous and terrestrial spirits are hostile, or at least not friendly. Less rational and more fearful. 
\item Spirits of fire are heroes and gods, and are the ministers of God. These are the Seraphim, Cherubim, and Thrones. They have very subtle bodies made of fire. 
\item Spirits occupy the bodies of humans, animals, stones, minerals. This is a form of panpsychism. Whatever his defects, \textbf{Teilhard de Chardin} insisted that a true science had to deal with the inside of things as well as the outside. 
\item No body is totally devoid of spirit and intelligence. 
\item Spirits never perish; rather, there is continual change of combinations and actualisations. 
\item There are various loves and hates. 
\item Everything desires to remain in its present state of existence and does not comprehend or think about another state of being. 
\item The diversity of natures and drives gives rise to a variety of bonds which affect spirits and bodies. 
\item There are subterranean spirits. 
\item Demons have emotions, desires, angers, jealousies 
\end{itemize}
Tomberg's definition is straight forward:

\begin{quotex}
``An autonomous being capable of interfering with the intentions of the ego" and which ``possesses a mental life of its own" is nothing other than what we understand by a ``demon".

\end{quotex}
We can't go into further detail in this introduction. Bruno claims to have experienced many of these spirits. The human race has always suspected their existence; this is shown in all the Hollywood films about spirits and in the works of \textbf{H P Lovecraft}.

\paragraph{Lucy and Harry}

\hfill

There are two particular spirits, or actually classes of spirits, that play a prominent role in Anthroposophical literature: Lucifer and Ahriman, nicknamed Lucy and Harry by the anthros. In an early work, \emph{Inner Development}, Valentin Tomberg describes the activities of these spirits in the human being, as shown in the diagram on the right. The Luciferic being is bound to the astral body in man's fallen state. He is called a doppelganger since it is like an evil twin. He is the father of lies, and creates idle fantasies in the mind that mask the imagination of Sophia, or Divine Wisdom. A moral universe requires hindrances, which are provided by this double. This is a clue to the meaning of the Book of Job.

Another doppelganger is the strictly human subconscious. In the Meditations, Tomberg uses the ideas of \textbf{Carl Jung} to explain the content of the subconscious. As for Ahriman, let us use his own words:

\begin{quotex}
And just as our angel, standing above us, represents, one could say, our archetype, so it is that the Ahrimanic double, on the other hand, is our caricature. This caricature, the Ahrimanic double, is a very intelligent being, one who does not manifest through arousing wild passions, but instead manifests particularly whenever subordination to an aim works strongly in a person. Precisely those individuals, therefore, who are highly respectable (from the external, bourgeois point of view) can be tools of this Ahrimanic double. It is precisely such people who can be subordinated, to a very high degree, to the being of the Ahrimanic double—who thereby takes on the leading role. \flright{\textsc{Valentin Tomberg}, \emph{Inner Development}}

\end{quotex}
As a reminder, the angel is our guardian angel whose prototype is the same as the person. This notion has even penetrated Orthodox theology, for example, in \textbf{Sergei Bulgakov}.

In the \textit{Meditations}, Tomberg became more critical of the anthro's obsession with the Luciferic and Ahrimanic beings. In that, he is justified, since their tendency is to reduce, even in its highest forms, the activities of the human spirit to manifestations of evil.

He barely mentions Ahriman at all, in fact. However, in the Letter on the Devil, he brings up Mephistopheles, whose characteristics are actually identical to those of Ahriman described in his earlier work. For example, consider the following quotations:

\begin{quotex}
Even Doctor Faust—who made a pact with the devil (and this concerns all ``pact-makers" of this kind, ancient and modern)—was only the naive victim of a prank on the part of Mephistopheles (who is a rogue well-known to all those who have knowledge of the ``occult world").

Some do it in the tragic sense of having to do what they no longer want to do and they no longer believe; others do it with fierce conviction and passionate indignation; there are also entities from the hierarchies of the left who accuse by making use of ridicule —\emph{farce}— as a means of demonstrating their prosecuting thesis.

It is primarily human pretension and snobbery that he turns into ridicule.

Mephistopheles; and the whole revelation is only a farce made by him for the credulous? no — for spiritual snobs.

Mephistophelian accusation against those who do not seek the truth as such, but rather extraordinary circumstances of revelation

\end{quotex}
The Ahrimanic doppelganger, aka Mephistopheles, is a \emph{caricature} of the human being. Other features are spiritual pranks, ridicule, and farce. He can be recognized by spiritual pretension and snobbery, i.e., the desire to appear to be ``spiritual" to others.

As for being well known, Mephistopheles may reveal himself in a dream rather than to the awakened consciousness. That I know from experience.

\paragraph{Personality Disorders}
Another indication of the influence of demons are personality disorders and psychopathy. Tomberg explains how that is possible:

\begin{quotex}
One engenders an elemental being and one subsequently becomes the slave of one's own creation. The ``demons" or ``evil spirits" of the New testament are called today in psychotherapy ``neuroses of obsession", ``neuroses or fear", ``fixed ideas", etc. They have been discovered by contemporary psychiatrists and are recognised as real — i.e. as ``parasitic psychic organisms" independent of the conscious human will and tending to subjugate it.

\end{quotex}
This is not the case with organic conditions such as bipolar disorder or schizophrenia, whose effects can be somewhat mitigated by medications. However, personality disorders like the paranoid, histrionic, borderline, etc., are different. Just because psychologists can categorize certain complexes of behaviour does not necessarily mean they are strictly medical problems. There is no real understanding of their cause nor is there any universally effective treatment. Tomberg's description, therefore, is pretty apt. He goes on to say:

\begin{quotex}
One need not fear the devil, but rather the perverse tendencies in oneself! For these perverse human tendencies can deprive us of our freedom and enslave us. Worse still, they can avail themselves of our imagination and inventive faculties and lead us to creations which can become the scourge of mankind.

\end{quotex}
If you have ever had a close relationship with someone with a personality disorder, you will probably agree that it feels more like dealing with evil than with a sickness.



\flrightit{Posted on 2018-10-08 by Cologero }
