\section{The Demon Lover}

\textbf{Naida}, at \textit{Orphic Inscendence}\footnote{\url{https://www.orphicinscendence.com/post/the-sacred-center}}, reveals a strange aberration of the Animus archetype that she calls the Demon Lover\footnote{\url{https://www.orphicinscendence.com/post/the-demon-lover}}. She describes one woman's experience this way:

\begin{quotex}
The demon lover, the immature and cruel animus who was guiding her towards manic pursuit of external goals, was feeding and taking her life energy. While she was chasing all of those things, she was actually feeding him. Eventually, he showed his face and came to have a laugh at her inability to recognise his tricks.
\end{quotex}

The cruelest trick of all is to pretend to be the ``good guy'' in her life.

\begin{quotex}
Almost universally, religions and spiritual traditions speak how demons do not always appear as ugly and fearsome, nor do they appear with knives and guns. Often, they are seductive, eloquent, beautiful, offering everything we think we need.
\end{quotex}

The danger of a woman's encounter with the demon lover should not be minimized. It can lead to disease, rage, depression, and even suicide. This encounter adversely effects family and love relationships. Yet, the demon lover is not undefeatable.

For the complete description of the demon lover and the path to spiritual and psychological healing, please visit \emph{The Demon Lover} at \textit{Orphic Inscendence}.

\flrightit{Posted on 2021-03-11 by Cologero}

