\section{View from the Primordial State}

\begin{quotex}
What heaven wishes\\
Fortune ordains\\
Reason demands\\
and, above all, what my will desires. \flright{\textsc{Don Quixote}}

In general, grace does two things. First, it remedies the defects in the natural order that have resulted from original sin, at least partially restoring what would have existed had the Fall not occurred. Second, it directs nature to an even higher, supernatural end — the beatific vision.\flright{\textsc{Edward Feser}}

\end{quotex}
\paragraph{Human Stratification}
The esoteric understanding of human differences is not horizontal, in the sense that people differ by races or nationality. Rather, the more important differences are in the vertical direction. Thus the notion of castes is used to explain it. However, caste has the connotation of being hereditary, so it is misleading. Castes can also be defined by function, as in the trifunctional classification of Georges Dumezil. But even that distinction does not take into account that people may functionally be in a caste that is inappropriate to them. That is noticeably true in decadent epochs like the Kali Yuga. So ``class" as used here has no other connotations. Hence, it is incumbent to understand the inner state appropriate to each caste. These are defined by how they relate to the three primary influences in the world: Providence, Will, Destiny.

\begin{itemize}
\item \textbf{First class}: they are guided by ``what Heaven wishes", or Providence. They are the spiritual leaders. Those in this class have realized a permanent sense of the transcendent Self. They live only for humanity and have come to help. These are in direct contact with the 1st angelic hierarchy. They have chosen their incarnation to accomplish a specific task. 
\item \textbf{Second class}: they are guided by their Will, which is aligned with the Will of God. They include the political leaders and warriors. People in this class consciously take part in the conflict between good and evil. They have developed, to an extent, a strong sense of Self-awareness and are led by the 2nd angelic hierarchy. They fight for causes that are larger than their own individuality, which they learn to recognize over time. 
\item \textbf{Third class}: they are subject to Destiny which they have been given by God. They have not yet developed a real sense of I and are victims of the curses resulting from the Fall. The people of the third class are all still experiencing their own karma and looking for their real purpose in life. They are under the leadership of the 3rd angelic hierarchy. 

Since they don't have a true sense of Self, their psychic center is dominated by one of the three soul functions:

\begin{itemize}
\item \textbf{Thinking}: Those dominated by the Thinking function seek out rational knowledge and love to dispute and debate. They believe that the road to knowledge is paved by a selection of books. 
\item \textbf{Emotions}: Those dominated by the Feeling function are moved by strong emotions. Since emotions are so fickle, it is difficult to establish a stable sense of Self. 
\item \textbf{Sensations}: They seek out ``experiences", and are guided by what they like or dislike. They are dominated by the Will, but a will untrammeled by intellect or sound emotions. 
\end{itemize}
\item \textbf{Fourth class}: Those in this group do not have any type of center and rely on outside influences for stability. The Church has always held a preference for these people, protects them from arbitrary political and economic powers, and leads them to salvation. 
\end{itemize}
The ideal distribution of power is from the first class downward, but deformations occur in history. The understanding of historical events needs to include the roles of the four classes at that time. This arrangement is not oppressive. Those of the First Class are called to serve the others, not to exploit them for their own benefit. Those of the Second Class take responsibility for administering the state, distributing justice, and protecting against enemies. The remaining classes are not conscious enough to take on those burdens.

\paragraph{Human State}
A human birth is rare and privileged. The principle of the human being is the Self, his center; a person is human to the extent that he participates in that center. Only then can he progress vertically. To sum up: our task as human beings, in order to actualize our possibilities, i.e., achieve Salvation, is to move horizontally toward the center and then vertically to reach even higher states.

In the essay \emph{Deliverance and Salvation}, Rene Guenon points out that there is a danger in failing to at least reach the center. His explanation needs to be quoted exactly:

\begin{quotex}
When a being must pass to another individual state, nothing guarantees that there he will again occupy a central position relative to the possibilities of that state as he does in its present state.

On the contrary, there is even an incomparably greater probability that he will encounter one of the innumerable peripheral conditions comparable in our world to those of animals or even vegetables.

One can immediately understand how serious a disadvantage this would be, especially from the point of view of the possibilities for spiritual development. 

\end{quotex}
Referring to the first part, those in the third class have not achieved the primordial state. Since the intellectual soul distinguishes the human from animals and vegetables, those dominated by the Thinking function still have the possibility of developing their possibilities. On the other hand, those dominated by Emotions and Sensations are in danger of remaining in a state of being comparable to animals or plant life.

\paragraph{Diseased Souls}
A fundamental metaphysical principle is that Knowing is Being. For example, in order to Know the Primordial State, one must be in, i.e., experience, the Primordial State. It is insufficient to just read and speculate about it.

Hence, the state of the soul determines what the being understands to be real and true. A disordered soul will not see reality, but only the projection of its disorder. 20\% of adult Americans, or about 50 million people, suffer from some degree of mental illness, including 10 million with a serious mental illness. One of the symptoms is a distorted understanding of reality.

For example, you wouldn't ask Sylvia Plath if life is worth living. Nor would you put a paranoid schizophrenic on the security council to identify the most serious threats facing a nation. Those with serious illnesses like personality disorders or psychopathy are similarly unable to relate to the world accurately. For example, the Mayo Clinic says:

\begin{quotex}
you may not realize that you have a personality disorder because your way of thinking and behaving seems natural to you. And you may blame others for the challenges you face. 

\end{quotex}
That describes a great deal of comments on social media and explains why communication can be so frustrating. This is not to belittle those who are in therapy, especially for conditions that may be situational or temporary. Nevertheless, such widespread mental illness reveals a deep flaw in the collective psyche of a nation.

Moreover, one's level of development also shapes a worldview and opens oneself up to various distortions.

\begin{itemize}
\item Those dominated by the Thinking function may fall victim to various ideologies or conspiracy theories. 
\item Others will react emotionally, rather than rationally, to events. 
\item At the lowest level, all that matters is pleasure vs pain, like vs dislike, attraction vs repulsion. 
\end{itemize}
Anyone not making the horizontal journey to the center will be inconsistent. He or she will have little self-awareness.

\paragraph{Lost in the Wilderness}
So living in the fallen world is just as described in Genesis. You will encounter a jungle of vegetation of various types. You will always be on the lookout for animals, some of which are harmless, others dangerous. Some are cute and cuddly, others are repulsive.

You will encounter pockets of proto-humans, able to lead and organize the animals or cultivate the vegetation. The blather of competing opinions sound like the random sounds heard at a watering hole in an African savannah.

On the positive side, fully functional humans are working for things that transcend their own self-interest. They may be organizing for humanity, such as, building nations, administering justice, creating hospitals, churches, etc.

Further off in the distance, you may catch a glimpse of saints and initiates. They are in higher states, but from time to time you may reach that state and recognize them. There are even higher beings. For us, the influences of angels and demons are experienced directly.


\hfill

\paragraph{Note}
\textit{The State of Mental Health in America}: 

\url{https://www.mhanational.org/issues/state-mental-health-america}

\textit{National Institute of Mental Health}: 

\url{https://www.nimh.nih.gov/health/statistics/mental-illness.shtml}

\textit{Personality Disorders}: 

\url{https://www.mayoclinic.org/diseases-conditions/personality-disorders/symptoms-causes/syc-20354463}



\flrightit{Posted on 2021-01-10 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Solphomeron on 2021-01-10 at 17:10 said: }

Hi I'm a mentally ill man, have been diagnosed with Borderline personality disorder, Narcissism and most recently Dissociative identity disorder (this last one is self diagnosed).

Finding Gornahoor and esoteric knowledge has been a double-edged sword. On the one hand It gives me tools to deal with my diseased soul, for example: Living a chaste life, in obedience to the Catholic faith have helped me to develop self respect and boundaries between me and other people, being vigilant helps me to not identify myself with my intense emotions, praying the rosary constantly have made me grow a more healthy internal dialogue, instead of violent and lustful daydreaming I find myself repeating the Our Father or the Ave Maria.

but on the other hand, My delusion feeds itself with the information found here, and creates more intricate, labyrinthine alternate realities..

So it has become difficult to distinguish between a synchronicity and a delusion of grandeur. My mind is always looking for associations, deeper meanings, symbolism. I often feel many layers of meaning in a conversation or interaction, and the deeper it always have some archetypal/mythic feeling. 

Can I trust myself? I find the only way out is to surrender completely To Christ. I'm a cup brimming with demonic goo and a few angelic sprinkles. But they are not me. Who am I? . I'm an interface. I'm a battleground.

\hfill

\texttt{Auld Wat on 2021-01-11 at 07:44 said: }

A personal reflection:

I believe I was in the fourth class until I was about 22 years of age. I then entered into the Third Class, on the lowest rung, when I became aware of what I liked and disliked. I cannot remember being overly emotional but I do remember enjoying debates and discussions. The changes in my awareness were brought about by external factors, women, political events, recommendations. 

I believe I moved from the Third to the Second Class only recently, despite being a practicing Catholic for the last 5 years, as it was only last week that I came to the realisation of what my purpose in life is. I had known what it was to some extent for over 2 years but I didn't recognise it fully until now.

Thank you for writing.


\end{sffamily}\end{footnotesize}
