\section{The Seven Steps of Esoteric Christianity}

\begin{quotex}
What had so often occurred individually within the Mystery schools, far from external events, took place openly within Jesus Christ in outer world history.
\flright{\textsc{Rudolph Steiner}}
\end{quotex}

Despite Evola's condemnation of Steiner as a ``hopeless muddle'', a closer look into Christianity can uncover patterns that might make us take Steiner more seriously. Steiner's seven steps of initiation closely correspond to passages from 2 Peter \& The Beatitudes (mirrored in the \emph{Shepherd of Hermas}) which represent a very ancient template for The Way (of Christ), a template also duplicated in shorter form in St. Paul's ``Faith, Hope, (Knowledge) \& Love'' (in which ``Hope'' can be read to stand for the middle virtues of Goodness, Knowledge, Self-Control, Fortitude, Godliness, \& Brotherly Affection listed in 2 Peter.

\begin{quotex}
5 For this very reason, make every effort to add to your faith goodness; and to goodness, knowledge; 6 and to knowledge, self-control; and to self-control, perseverance; and to perseverance, godliness; 7 and to godliness, mutual affection; and to mutual affection, love. 8 For if you possess these qualities in increasing measure, they will keep you from being ineffective and unproductive in your knowledge of our Lord Jesus Christ.
\end{quotex}

As George Heart points out\footnote{\url{http://www.amazon.com/Christianity-Dogmatic-Gnostic-Vivifying-Knowledge/dp/1412062284}}, this formula was supposed to be used by the Church to progress from naked Pistis \& Dogma, through to actualized positive spiritual knowledge, and finally the Pleroma of Love, which is not merely a state of Being, but actually the highest state of Being.

The \emph{Ekklesia} (for a number of reasons) has been unwilling \& unable to perform this mission -- to admit that the current Matrix would not exist without the Church would be the first step, but to strip down to truth is painful, although it would allow the Church to perceive that which could be done, which still inheres in the potentialities of our decline as ``intimations of deprivation''.

Rudolf Steiner\footnote{\url{http://wn.rsarchive.org/Lectures/EsoCosmo/19060601p01.html}} lists the seven steps of initiation, which can be accomplished through meditation and the final step of a specific dream:

\begin{quotex}
The neophyte who meditates on this theme for months and years has a vision of the Washing of the Feet in the astral world during sleep. Then he is ready to pass to the second stage of the Christian initiation.
\end{quotex}

The entire Schemata of Steiner follows (my suggestions in parentheses -- Faith is assumed to begin the possibility of initiation):

\begin{enumerate}
\item The Washing of the Feet. (Goodness?)

\item The Scourging. (Knowledge?)

\item The Crowning with Thorns. (Self-Control?)

\item The Bearing of the Cross. (Perseverance?)

\item The Mystic Death. (Godliness?)

\item The Entombment. (Brotherly Affection?)

\item The Resurrection. (Love?)
\end{enumerate}
As we are trying to see, these steps are roughly equivalent to the seven steps of growing Faith (\emph{Pistis}), which must proceed towards \emph{and even through} knowledge to stay alive. For example, Scourging can be viewed as a trial of \emph{Knowledge}, since only those who truly understand \& value rightly the external world (\emph{Gott und die Welt}), holding it with a light hand, can accomplish this heavy task; let us look in more detail:

\begin{quotex}
The Scourging, --- At this stage man learns to resist the scourgings of life. Life brings sufferings of all kinds --- physical, moral, intellectual, spiritual. Life is felt to be a dreadful and incessant torture. The disciple must endure it with perfect equanimity of soul and heroic courage. He must cease to know physical or moral fear. When he has become fearless, he sees, in dream, the scene of the Scourging. In another vision he sees himself in the Christ Who is scourged. Certain symptoms in physical life accompany this event. There is an intensification of the life of feeling, a wider sense of life and of love. We have an example of heightened sensibility transferred to the world of intelligence, in the life of Goethe.
\end{quotex}

In this ``matching'' of the virtues with Steiner's steps, Faith is assumed \& comes previously to the first initiation, and so Faith is but the beginning; initiations follow to allow the disciple to ascend the seven ``following'' heavens.

Interestingly enough, the ``Scourging'' was rumored to have occurred during the torture of Knights Templars during their suppression --- the intense doubt \& pain caused by the ordeal supposedly brought some of the knights to manifest Baphomet, who would have been one of the guardians of the lower regions of the soul, a double to the higher soul. It was taken as proof of guilt, although it may have simply represented the failure (under duress) of a Templar to pass a stage of growing initiations.

\begin{quotex}
In the course of the stage in the Christian initiation known as the Crowning with Thorns, there arises the phenomenon known as the Guardian of the Threshold --- \emph{the appearance of the lower double of man}. The spiritual being of man, composed of his impulses of will, his desires and his thoughts, appears to the Initiate in visible form. It is a form that is sometimes repugnant and terrible, for it is the offspring of his good and bad desires and of his karma --- it is their personification in the astral world, the Evil Pilot of the Egyptian Book of the Dead. This form must be conquered by man before he can find the higher Self. The Guardian of the Threshold which has been a phenomenon of astral vision from times immemorial, is the origin of all the myths concerning the struggles of Heroes with monsters, of Perseus and Hercules with the Hydra, of St. George and Siegfried with the dragon.
\end{quotex}

The rest of the initiations can be matched up nicely\footnote{\url{https://gornahoor.net/?p=2619}} with the template of virtue begun in, and common to, the Beatitudes, 2 Peter, Paul's project of gnosis (which summarizes the middle virtues as ``Hope'' and makes explicit the command to begin in Faith), \& the Shepherd of Hermas. Regardless of whether one accepts Steiner's account or links it with the early Christian community, this pattern which intertwines Exoteric expression (Faith is an exoteric ``virtue) with esoteric ``fullness'' (Faith must be regarded as ``passing away'') is present in the foundational documents of the ancient Christian religion. It may be possible to also relate the ``nine true parts of the fully human'' to this template of virtues. Steiner also lectures on this mystery\footnote{\url{http://www.doyletics.com/arj/christia.htm}} (which by the way matches Mouravieff's division of the human person into three centers with three sub-divisions).

\begin{enumerate}
\item Physical -- Corporatio -- Embodiment

\item Ether or life body -- Nativitas -- Birth

\item  Astral -- Circumcisio -- Circumcision

\item I --- sentient soul -- Apparitio -- Appearance

\item I --- intellectual soul -- Passio -- Suffering

\item I --- consciousness soul- Mors -- Death

\item Manas --- spirit self -- Resurrectio -- Resurrection

\item Buddhi --- life spirit -- Gloria -- Glory

\item Atman --- spirit body -- Regnum -- Kingdom
\end{enumerate}

Even if the pattern is contested, the idea of Faith the seed growing organically into a celestial body\footnote{\url{http://members.ozemail.com.au/~moorea/peter\_keys\_kingdom.html}} of Love is virtually incontestible. The virtues are living \& active (or should be) as their Lord the Spirit.

These patterns abound in Scripture, and elsewhere in the Christian Tradition. The last words of the Lord's Prayer are ``thine is the kingdom, the power, \& the glory forever''. Here we see the repetition of 3's -- the Kingdom is the Archetype (and also the return, the end of the mystery, when God is ``all in all'' according to Colossians), the Power is the Incarnation (Christ is not just ``an'' avatar, but the Lord of avatars), \& the Glory is the Spirit (who is the more subtle and feminine element in the Godhead). Here again we find a ritualistic formula whose real meaning has been lost; there are others, such as the Spirit, the water, \& the blood. Here we may hazard a guess that water is associated with soul-death, the moon, and Old Testament rites -- hence, Yahweh, or the Archetype. The blood is flesh and man, so -- the Incarnate Form of God. And so on. More research in esoteric sources (or helpful or knowledgeable readers) may be able to add to this.\footnote{A Catholic theologian who could serve as a bridge: \url{http://www.scribd.com/doc/5994483/Jean-Borella-Gnosis-and-AntiChristian-Gnosis} (like Steiner) between esotericists and the \emph{Ekklesia} is Jean Borella: \url{http://en.wikipedia.org/wiki/Jean\_Borella}}.


\flrightit{Posted on 2011-08-12 by Logres}

