\section{Kind Hearts, \& Norman Blood}

\begin{quotex}
Gewusst, Erkannt, Geahndet
\flright{\textsc{Schelling}}\footnote{\url{http://books.google.com/books?id=k-\_r3feIiNoC&pg=PA101&lpg=PA101&dq=gewusst+erkannt+geahndet+schelling&source=bl&ots=YveYm9wWiV&sig=K5bIoWF-Hb4W9r14wYWhn4QclIA&hl=en&ei=SosLTpC2PIS-tgfnt8CSAQ&sa=X&oi=book\_result&ct=result&resnum=1&ved=0CBYQ6AEwAA\#v=onepage&q=gewusst\%20erkannt\%20geahndet\%20schelling&f=false}}

What is now called the Christian religion already existed among the ancients, and was not lacking at the very beginnings of the human race. When Christ appeared in the flesh, the true religion already in existence received the name of Christian.
\flright{\textsc{Augustine of Hippo}}
\end{quotex}


So much comes back to the union of opposites, the basis of alchemy -- the turning of lead into gold, the union of male \& female, etc. One might suggest that the fourth chakra is the balancing chakra (the heart chakra\footnote{\url{http://en.wikipedia.org/wiki/Anahata}}) which unites heaven \& earth. Perhaps one reason that Alistair Crowley is merely mined by Evola for insight, rather than endorsed as a master, is that Crowley essentially attains deification rather than divinization -- that is, he bypasses the heart chakra totally (and all considerations of morality, even in the act of unconsidering it), dedicating one's self to ``service-to-self''. This is a difficult path; most humans have some feeling at least for their kith \& kin, which (of course) is why Chesterton argued that the gods of Rome\footnote{\url{http://www.worldinvisible.com/library/chesterton/everlasting/part1c8.htm}} were ``human'' in a way that the gods of Carthage were not

\begin{quotex}
The Roman augurs and scribes who said in that hour that it brought forth unearthly prodigies, that a child was born with the head of an elephant or that stars fell down like hailstones, had a far more philosophical grasp of what had really happened than the modern historian who can see nothing in it but a success of strategy concluding a rivalry in commerce. Something far different was felt at the time and on the spot as it is always felt by those who experience a foreign atmosphere entering their own like a fog or a foul savor. It was no mere military defeat, it was certainly no mere mercantile rivalry, that filled the Roman imagination with such hideous omens of nature herself becoming unnatural. It was Moloch upon the mountain of the Latins, looking with his appalling face across the plain; it was Baal who trampled the vineyards with his feet of stone; it was the voice of Tanit the invisible, behind her trailing veils, whispering of the love that is more horrible than hate. The burning of the Italian cornfields, the ruin of the Italian vines were something more than actual; they were allegorical. They were the destruction of domestic and fruitful things, the withering of what was human before that inhumanity that is far beyond the human thing called cruelty. The household gods bowed low in darkness under their lowly roofs; and above them went the demons upon a wind from beyond all walls, blowing the trumpet of the Tramontane. The door of the Alps was broken down; and in no vulgar but a very solemn sense, it was Hell let loose. The war of the gods and demons seemed already to have ended; and the gods were dead. The eagles were lost, the legions were broken; and in Rome nothing remained but honor and the cold courage of despair.
\end{quotex}


I do not presume to judge the Crowleys of the world who invoke Koranzon\footnote{\url{http://en.wikipedia.org/wiki/Choronzon}}; I am sure that God can do that all on His own.

However, it does seem to me that the Pagan/Christian split is rather out of date today; one can find plenty of neo-pagans \& neo-Christians who agree that the world is essentially ``secular''. Likewise, one can find plenty of pagans \& Christians who actually desire to raise a standard against the modern world, although the means for doing so are disagreed over. Most Christians will view the attempt to forge a deeper unity as potential syncretism. A large-minded Christian friend of mine wrote that ``I see certain challenges with the presentation of Christianity as an esoteric doctrine which is rooted in a kind of pan-religious, metaphysical syncretism''. Certainly. If we explore the fear (which I share) of straying from the ``Way'' (for those who are rooted in Christianity as I am), I am lead to conclude that there is finally a kind of violence which is practiced on the self; Jesus gave this as ``taking heaven by storm'', or ``violence carrying it away''. One can either obviate the need to struggle by cutting out the heart chakra entirely and proceeding by way of bizarre practices which are not strictly speaking necessary for those who are not ``seared'' or ``vulnerable masters'', engage in fundamentalist fantasy, or integrate the Left \& the Right hand paths at a much higher level than that given to us by the conditions we live in. In other words, ``syncretism'' is only a necessary problem at the exoteric level -- it is a potential problem at the esoteric.

Now even Evola, in \emph{Revolt of the Modern World}, speaks of the beginning usefulness of drugs to create an alternate sense of reality, but a fading of that need as progress is made. Here we have the ``Christian'' tendency from the other direction, just as there is a legitimate ``pagan'' tendency from ``ours''. \emph{This places Evola squarely in the camp of those who wish to preserve the entirety of the human being.} This is close to what St. Paul must have meant when he said ``all things are lawful, not all are profitable'' -- the keeping of Torah was a re-affirmation of a concrete and concretely discerned ``path of heaven''. It seems to me that such a deep drive to preserve the entirety of human aspiration and integral wholeness is ultimately what is meant by purity, or true religion. Thus, the division innate between exoteric meanings only carries us so far -- a pagan who communes with Odin, yet who wishes to preserve the full meaning of ``Englishness'', is committed inescapably to certain Christian aspirations.

Defining what that looks like is difficult, even humanly impossible. If we abandon the ``we'' of exotericism, \& focus upon the ``I'' of transformation, the difficulty of finding the ``we'' will become considerably less. This is the task facing the West -- re-embrace metaphysics, or have the metaphysics of the alien imposed upon us, as Guillame Faye\footnote{\url{http://guillaumefayearchive.wordpress.com/2010/07/09/190/}} has argued will occur in Europe\footnote{\url{http://guillaumefayearchive.wordpress.com/2007/07/07/the-islamic-conquest-of-europe/}}.

Whereas Christianity and the Church and ``paganism'' may ultimately decide that some measure of separation is beneficial, it cannot be the case that a hard line divides those who practice esoteric paths. This is not ``syncretism'' -- provided the sage\footnote{\url{http://www.gornahoor.net/?p=2314}} does not trouble the faithful with their wisdom. This is simply the attaining of poetic opposites in a manner which dissolves the need to have a ``war of ideas''. As John Ruskin points out, the ``pathetic fallacy'' in poetry is only a fallacy\footnote{\url{http://www.ourcivilisation.com/smartboard/shop/ruskinj/}} in the hands of an amateur. Likewise, here, the ``reunion'' of Christianity with the faithful pagans is attainable by the non-vulnerable master (one committed to totality in totality, with the whole of the balancing chakra). The common person can grasp this purely on the level of poetry, but in the supreme heights of God's kingdom, in Valhalla, heaven and earth will kiss, righteousness and peace will make peace. This dream built the \emph{imperium}\footnote{\url{http://www.youtube.com/watch?v=mraO8JZbSkg}} of Europe, from the times of Boethius through Zwingli's insistence that Socrates would be found in heaven, all the way to the days of Kitchener's storming of Khartoum and Lord Cromer's administration of Egypt. This dream did not begin with the Catholic Church being ``syncretic'' or harboring esoteric doctrine (which it did). It did not die, in spite of such disasters as the crusade against the Cathars in Southern France. It began with the touching of Christ's teaching with the Mediterranean world around Alexandria (at least in our memory). St. Peter and St. Mark transmitted the teaching of the perfection of the kingdom to Clement, who endeavored to ``conceal'' it openly in the \emph{Stromata}. The fact that this secret teaching was both ``secret'' and ``Judaic'' or based on Torah is not a contradiction to the true aspirant, should he choose it. The teaching\footnote{\url{http://jacksonsnyder.com/arc/bible-news/apostolia/pdf/CLEMENT-Excerpt-Book-IV.pdf}} was concerning the architect of the Cosmos, and the path of those who desired to set themselves free from ``demons''. This was because the ``flesh'' had to be superseded. Man's destiny was to be more than human. ``Greater than the angels''. This teaching is both ``pagan'' (in the sense that it accepts the reality of the supernatural), ``Gnostic'' (in the sense that this world is not made ultimately for bourgeoise comfort\footnote{\url{http://www.arctogaia.com/public/eng/gnostic.htm}}), and ``Christian'' (in that it aims for the Father of Lights in Valhalla). It is inimitable. It is what it is. It is the Tradition. Here is a ``Gnostic'', Alexander Dugin:

\begin{quotex}
The mankind has always had two types of spirituality, two paths --- ``Right Hand Path'' and ``Left Hand Path''. The first one is characterized by the positive attitude to the surrounding world; the world is seen as harmony, equilibrium, good, peace. All the evil is viewed as a particular case, a deviation from the norm, something inessential, transient, without deep transcendental reasons. Right Hand Path is also called ``The Way of Milk''. It doesn't hurt a person, it preserves him from radical experience, withdraws from immersion into suffering, from the nightmare of life. This is a false path. It leads into a dream. The one going by it will reach nowhere... 
\end{quotex}


In the apostle's day, the way of milk was using demons for fortune-telling to make a buck or two.

Tennyson made the point that ``kind hearts are more than coronets, and simple faith than Norman blood''; Heaven's Truth, being additive, can simply add that the perfection is both the attainment of Norman blood, and the purity of simple faith with kind hearts, crowned with the corona. Likewise, esoterically, we seek to listen to those who can ``discern the spirits'' so that true North, all of it, comes into view, if happily we search for it, and find it. This is why Gornahoor attempts to harmonize Donoso Cortes with Julius Evola.

There is a great line in a very modern movie -- \emph{focus is somewhere in between rage \& serenity}. In our decaying age, what is needed is focus, beginning in the Self. The one who truly wishes to find the Tao will listen to whomever speaks at hand with wisdom, judging in their heart, the heart. Here, in the kingdom of Heaven, ``there is neither Jew nor Gentile''. That is the meaning of Saint Paul's comments; since Grace perfects, and does not eradicate Nature, he is not teaching One-World Utopianism, in which a sexless, raceless, classeless drone inhabits the void created by technology on the earth. He is saying that Aratus and Socrates are in heaven, or could have been. It is finally only the struggle with the Self -- not only ``every sin'' but ``every weight'' must be laid aside, for all things ``are lawful, but not all are convenient''. \emph{Upaya}, skillful means, or the balancing of the fourth chakra must occur, for in no other way can ``Thy Kingdom Come''. There are plenty of ``Christians'' who have no idea what the \emph{Basileia tou Theou}\footnote{\url{http://en.wikipedia.org/wiki/Kingdom\_of\_God}} means, or even that it exists. In our day, the brightest line is between those of bad will who are trapped by the self, and those of good will who have risen to the struggle. To tell the difference, begin to look deep inside.


\flrightit{Posted on 2011-07-01 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Kushiel on 2011-07-02 at 07:37 said:}

For the deeper Christians who, against linear logic and cultural taboos, find high value in Guenonian Traditionalism and its modification by Evola, what can the ``Left Hand Path'' mean, as followers of Jesus Christ? How can the ``Left Hand Path'' be reconciled to Christ and consistent moral theology?

\hfill

\texttt{Kushiel on 2011-07-02 at 09:46 said:}

Isn't the proper response of the pure sacred warrior to the ``Left Hand Path, as seen, for example, in the dehumanizing ``spermo-gnosis'' and malefic activities of the O.T.O... PURE ANNIHILATION? I feel like quoting the Korn here: ``Slay the Pagans, wherever you find them''--isn't this the normal moral response of the outwardly violent but inwardly calm Ideal Traditional Warrior to immorality and decadence, no matter how dressed-up in pseudo-mysticism? Should we even attempt to reconcile what should not be reconciled? Is a Hegelian sublation possible where antipodes exist--God as LOGOS-AGAPE, transcendent, selfless Wisdom and Love vs. God as compassion-free CHAOS?

\hfill

\texttt{Logres on 2011-07-02 at 10:48 said:}

Well, Dugin's definition seems pretty sound to me -- the Left Hand path (for the Christian) would be the Dostoevsky path, or the path of violence against the self. This might look un-Christian to some; Evola himself advocates practices which would seem counter-intuitive (drug use, the right to suicide), but only under higher conditions. Alexander Dugin's definition captures some of that; and there are Biblical examples: literally, Ehud, and others in the ``Hall of Faith'' who (if you read their story) you would think were ``pagans''. Barak was a coward following a woman, Samson with his Philistine women, etc. Christ opened up a ``broad'' narrow path (a safe path) based on the balancing of the heart, but some people are so mangled they can't use that. There are some Left Hand paths that are not available to Christians, because they have already entered the Kingdom and begun the journey; to them, it is sin. To another, with no purpose, it may not be sin, but simply degeneration which may or may not issue in ultimate salvation of a kind.

\hfill

\texttt{Logres on 2011-07-02 at 12:51 said:}

In a literal sense, yes, you are right. However, Christ's coming in Revelation is very similar to this -- as are all the promises made to those who do not seek God. Yes, so in a sense you are absolutely right. The difference between Hegelian synthesis and poetic apprehension is that the poetry doesn't have to be literally crammed together (again, this is precisely what Crowley does in the Thelema, right? Good and evil are the same?). It's a synthesis that preserves particulars. In Christian dogmatic terms, this has to do with the difference between being part of Christ's head (or One) \& being recapitulated as part of his body. This is the mystery referred to in Colossians -- first, a Lordship of Christ, then the return to the Father. I don't think this mystery permits us to say that there is some kind of absolute Torah; the Torah becomes ``relative'' in order better to relate. Hence, there is a mystery of evil as well (Judas). All this does, for the Christian, is to preserve the possibility that other paths may have their own God-given destiny. It isn't a forced syncretism or even comprehension. At a deeper level, is there evil in the city, and the Lord has not done it? I realize that most people will see this as a betrayal of Agape; I see it as a strengthening of it. Part of the human nature of Christ, which freely chose to serve under a yoke he by right could have transcended. Again, this is explicit in Scripture. So while I share the concern (if you are a Christian), and remain puzzled, I don't agree that we can always know with certainty what is taboo. Was the Iron Guard anti-Christian? Most would say so. I am not so sure.\\
\url{http://en.wikipedia.org/wiki/Iron_Guard}


\hfill

\texttt{Will on 2011-07-03 at 09:30 said:}

While Dugin's criticism is certainly applicable to modern `spiritual' sensibilities, it is incorrect to say that this is the Right Hand Path. The difference between the Right and Left hand paths is not, as he asserts, between false and true, but rather between what Evola calls `dry' and `wet' paths, respectively. He regarded both as valid, for different types of people and different historical periods. His examples were early Buddhism and Indian Tantra.

The Right Hand Path is based, as Dugin notes, on withdrawal from the world. This is the principle behind monasticism, and if it is anti-traditional, then Christianity has a lot to answer for, since it has made use of monasticism perhaps more than any other tradition.

\hfill

\texttt{logres on 2011-07-06 at 12:08 said:}

Dugin is taking poetic license -- his point is more emotional than metaphysical (I think). Here is a good summary of how I feel about Crowley:

``The cosmic constellation of the human being, particularly in modern times, is extremely complex. On the one hand he remains within eternity, in the lap of the Divine Trinity, only unconscious. On the other hand, if he is to become conscious of his situation -- which is the main task of the spiritual ontogenesis of man, if one may so express it (because we can develop ourselves only in the totality of body, soul and spirit) -, he must for a period of time work upon himself within a further trinity: that of Christ, Lucifer and Ahriman. And if he errs (development always entails risks), he may quickly find himself within yet another trinity: that of Lucifer, Ahriman and Asura (a retarded spirit of personality, archai). He will then fall out of the evolution intended for him by the divine worlds, and take a different path of which the Gospel says that there reigns outer darkness; there will be weeping and gnashing of teeth (Matthew 22; 13). At some point this world too will be transformed into good, but today no-one can say whether the human being who takes this path will then still exist.

To the spoiled but also intimidated human being of today, it is highly uncomfortable to learn of truths like these. And yet he must either develop within himself the strength and the wish to face up to the truth, in order to recognize the true laws of development and work upon himself (then God Himself will be his helper), or sooner or later he will find himself on the path of `continuous Fall'. In some black-magical occult circles this path is even considered the best, but such opinions belong to `black' romanticism, which is a special kind of egoism.''

This is Gemmdy Bondarev:
\url{http://www.altanthroinfo.9f.com/GoodEvil.htm}


\hfill

\texttt{logres on 2011-07-09 at 01:07 said:}

``The soul then perceives how Christ holds Ahriman back in the sphere of the mineral forces (the gesture of the right hand), and directs him to his realm of measure, number and weight. With the gesture of the left hand Christ keeps Lucifer back from the fall into darkness, and thereby opens up for him the possibility of salvation '' -- Bondarev

\hfill

\texttt{Mr Green on 2017-08-07 at 18:17 said:}

Logres said:

``I don't agree that we can always know with certainty what is taboo. Was the Iron Guard anti-Christian? Most would say so. I am not so sure.''

A shame that this statement went unchallenged. It is understandable that one is dazzled by an archetypal warrior personage such as Corneliu Codreanu, but a closer look at the history of the Iron Guard, especially after Codreanu's `martyrdom' in 1939, should dispel every doubt that the `Christianity' of this organisation was anything more than a merely nominal creed misused for nationalistic ideological purposes.

\url{https://en.m.wikipedia.org/wiki/Legionnaires\%27\_rebellion\_and\_Bucharest\_pogrom\#}


The legacy left behind by Codreanu's organisation of `Christian' legionnaires is mainly an anti-semitism so blindly fanatical that it resulted in a mass orgy of such monstrous violence that it escapes comprehension how any human being, let alone a self-professed Christian, could be capable of such a depth of evil. It reveals that the organisation was massively under demonic influence. It was merely a parody of the Christian warrior principle: at times beautiful insofar as its image reflected an archetype, but ugly at heart.

\hfill

\texttt{Logres on 2017-08-08 at 22:11 said:}

The pogrom occured after the murder of Codreanu, and when Eliade had left for England. The most that you can claim is that the pogrom was the natural fruit of the Legion's work, and I think that would be difficult to establish, although I suppose it's possible (as my comment indicated so). I was speaking of Codreanu and his men, not the subsequent policies of the Iron Guard. What about the Legion of the Archangel Michael? And isn't it true that corruptio optima pessimi?

\hfill

\texttt{Mr Green on 2017-08-09 at 09:19 said:}

I know that it occurred after the death of Codreanu. But how could possibly a truly spiritual order deviate so completely in such a short period of time after the death of its founding leader? Does not that tell us something about the general character of its membership body? How do we know Codreanu himself wouldn't have participated in such excesses had he lived a little longer? Since I find much to admire in the man, I hope this wouldn't have been the case. But in any case, it is indisputable that Codreanu actively forged an intellectual anti-semitism so extreme that these cruel excesses committed so shortly after his death with the sanction of the order's elite cadres indeed, and unfortunately so, must be considered the fruit of his work. Ideas have consequences, and especially when fed to a paramilitary organisation prone to direct action. If the membership body at large had been genuinely permeated by the spirit of Christ from the beginning, as befits its official intention, these evils committed by that same body so soon after its beginning would not have been possible. If Codreanu had not approved of it (and this is an ambiguous matter), he must have utterly failed in the spiritual formation of this body, so that there was no one to succeed him, letting loose the lowest and most vulgar level of men. All these people considered themselves Orthodox Christian believers. It goes to show how irredeemably fallen people are when they are capable of justifying the most extreme evils as self-professed Christians. What do we need satanism for then?

\hfill

\texttt{Logres on 2017-08-25 at 23:43 said:}

Well, apparently the entire world degenerated rather quickly after Creation, so I find it hard to believe that some kind of ``Fall'' (perhaps the acts referred to by yourself) did not intervene. How else do you reconcile the obvious nobility with the obvious crime? I'm not as familiar with Codreanu as I could be, So I'm not as eager to judge his entire heart; there was something noble there which went off the rails. If the head of a hierarchy is chopped off, I don't think you can automatically blame what follows on the leader. That would be like blaming the texture of Reconstruction on Abraham Lincoln without making a specific case..That's not the popular view of history, or of ideas, but Codreanu's ``fruit'' cannot be necessarily considered as the actions of certain elements in the Iron Guard following his violent murder.

\hfill

\texttt{Logres on 2017-08-25 at 23:54 said:}

It's possible you're right: I am not as certain as you are. I don't think spiritual formation (as you say) is quite so straight forward: a man may do great things, and ``lose his own soul''.

\hfill

\texttt{Ilo Stabet on 2021-03-16 at 06:41 said:}

I have tried finding the source of that St. Augustine but it's proving difficult. I find it reproduced a lot on the internet, but only a few mention its origin. The ones who do say it is from Contra Faustus, 33:6 -- but I cannot find it there (nor within any of the other chapters). So it seems everyone that provides this source is also reproducing it from the one place and taking it at face value. Do you have any clues as to the real source?

Thank you.

\hfill

\texttt{Cologero on 2021-03-16 at 15:36 said:}

It should not be difficult to find. It is in the Retractions, Chapter 12.

\hfill

\texttt{Ilo Stabet on 2021-03-16 at 16:09 said:}

Thank you.


\hfill

\end{sffamily}
\end{footnotesize}

