\section{General Law}

After finishing \emph{Gnosis 1,2,\&3}, the concept of \textbf{General Law} stands out as very nearly the most important paradox in the book we can grasp, if we want to grasp daily practical implications. This concept has political implications, as well as personal: No one escapes General Law -- it is impossible. In the words of Phillip Dick in the novel V.A.L.I.S.\footnote{\url{http://en.wikipedia.org/wiki/VALIS}}, ``those who fight the Empire, become the Empire (the Black Prison)''. Or if you prefer Saint Paul, ``we struggle not against flesh and blood... bodily exercise profiteth little... we are in the world, but not of it... ''. General Law is the daemonic/angelic/demonic hierarchy which determines natural history and cosmology, as well as the fate of the person who is ``under the Law''. Now, for Saint Paul, the law is ``holy and good'', yet it is ``\emph{given} for the lawbreaker''. One might say, it is \emph{a given}. But Mouravieff does more, and attempts to explain the inner workings of this law. For instance, how can it be both good \& evil? What does Saint Paul intend to teach, here? We are knee-deep in the mysteries of the epistles with this teaching. One principle is worth mentioning as an aside: The General Law allows you to escape it \emph{only if you train (or create/provide for) your replacement}. All sorts of applications could be made here, for instance, with the duty to raise children, or a master teaching a pupil.

This Law is iron -- the more one attempts to confront it with human means in space and time, the more powerful the Law becomes. It grows through defiance. This is the existential defeat and despair of fallen man, and it is, coincidentally, \emph{precisely what the illusory dream of Revolution is truly all about}. Man will confront the cosmic tyrant and break his chains. Man will stand against the gods, on behalf of all of our sorrow, rage, fear, suffering, and hunger. Many would portray Christ as the new champion of this. And man will thereby strengthen Satan.
But what did King David\footnote{\url{http://www.biblegateway.com/passage/?search=Psalm+2\&version=KJV}} see of this?

\begin{quotex}
Why do the heathen rage, and the people imagine a vain thing?

\textsuperscript{2} The kings of the earth set themselves, and the rulers take counsel together, against the Lord, and against his anointed, saying,

\textsuperscript{3} Let us break their bands asunder, and cast away their cords from us.

\textsuperscript{4} He that sitteth in the heavens shall laugh: the Lord shall have them in derision.

\textsuperscript{5} Then shall he speak unto them in his wrath, and vex them in his sore displeasure.

\textsuperscript{6} Yet have I set my king upon my holy hill of Zion.

\textsuperscript{7} I will declare the decree: the Lord hath said unto me, Thou art my Son; this day have I begotten thee.

\textsuperscript{8} Ask of me, and I shall give thee the heathen for thine inheritance, and the uttermost parts of the earth for thy possession.

\textsuperscript{9} Thou shalt break them with a rod of iron; thou shalt dash them in pieces like a otter's vessel.

\textsuperscript{10} Be wise now therefore, O ye kings: be instructed, ye judges of the earth.

\textsuperscript{11} Serve the Lord with fear, and rejoice with trembling.

\textsuperscript{12} Kiss the Son, lest he be angry, and ye perish from the way, when his wrath is kindled but a little. Blessed are all they that put their trust in him.
\end{quotex}

The General Law \emph{is the missing link} between the testaments of Law \& Grace. Jesus, also is, because he satisfied that General Law, and thereby provided an opening. But the Law is not dead, but is active and alive still, unless one is under the aegis of some symbol. It is also a warning to Modern Times: the power of the angels will only bind our chains tighter if we presume to liberate man in any other way than the one way given, the narrow way. This way is to transcend the human condition, through vertical energies, \& to enter the Kingdom of Heaven. There is no other way which the Guardians of the Threshold\footnote{\url{http://en.wikipedia.org/wiki/Guardian\_of\_the\_Threshold}} will permit, personally or collectively. As seductive as the idea of building everyone a duplex with running water and free phone lines is, it is actually a grave sin against every human being \& one's own soul to directly confront the Guardians of the General Law with human desire and physical means, \emph{promoting the Revolution on earth}. Revolution will only build more Revolution, because it will only build the Black Prison higher and stronger. I will repeat this -- to directly confront the hierarchies in a human and physical (read ``political'') manner is to spit in Christ's face (who Himself refused this temptation, though he might have done it for his own sake). Ultimately, all Revolutionists are building their own Hell, and it will be all the more secure for that, since they built it, who imagine that anything in this life on its own level can ``save'' man, political or otherwise.

In fact, as Luther apocalyptically\footnote{\url{http://creation.com/battle-quote-not-luther}} said, it is the gravest sin in the world today, because this is the occult point of attack:

\begin{quotex}
If I profess, with the loudest voice and the clearest exposition, every portion of the truth of God except precisely that little point which the world and the devil are at that moment attacking, I am not confessing Christ, however boldly I may be professing Christianity. Where the battle rages the loyalty of the soldier is proved; and to be steady on all the battle-field besides is mere flight and disgrace to him if he flinches at that one point.
\end{quotex}

The Moderns are traitors even to their founding saint, as the above amply illustrates; sadly for them, Luther may be saved anyway, but they will not, because God will laugh them to scorn and mock them, and Christ Himself will break them in pieces. Christ permits the General Law, because He himself satisfied it, and has directed in what manner it may be overcome. It is \emph{hubris}\footnote{\url{http://en.wikipedia.org/wiki/Hubris}} to imagine another way.

There is a warning for traditionalists also -- the Gods of these times permit the modern excess in an open manner, and thus to directly confront this in an overly physical manner is to become guilty of the same modern \emph{hubris} the Moderns practice (think the Confederate States of America, here). Rather, warn those who need warning, \& save yourselves, first, and allow Revelation and history to unfold until ``it is time''. Do not confront Satan directly (for even the archangel Michael did not\footnote{\url{http://www.biblegateway.com/passage/?search=Jude+1\%3A9\&version=KJV}}), for the power of evil has not yet been exalted. This is the Kali Yuga. This is why the founders of Gornahoor have spent so much time on ``theory'' and personal practice, rather than ``direct action'', although one can certainly prepare the groundwork and framework against the chaos that is sure to inevitably transpire.

Instead, we should spend our excess power and time in seeking to legitimately (read, traditionally) escape General Law in the approved manner, through the narrow gate. The jealous and all powerful (within their spheres) hierarchies seem to give man much ``leeway'' but it is actually a total illusion, designed to keep him in chains to further the Common Good and destiny of the species, rather than liberate his soul prematurely. Man is so fragmented and insane, spiritually, that God permits him to be ruled with illusion and lies in order that this generation will perish in ``feeding the moon'' (see Gurdjieff's teachings), but that the Divine Plan will go on. Man will think himself free under the Revolution -- the Revolution is a deepening of the Lie to its most powerful point to counterbalance the excess of disintegrative energies that is loose in our time.

And he will die and perish, and his states and societies will perish with him, because Man is more important than he dreams, or than even the Revolution promises. Man will endure anyway.

Even the Demons delight in the Great Game, but they secretly yearn for heroes who are equal to tilt the course against them, teaches Mouravieff. Is this what you, too, desire? Then bow down in the temple of Rimmon, \& repent of speaking ill of high authority, with all of its attendant woes and evils, and seek the narrow gate. Reject Revolution, embrace Authority; submit, and bow down in the temple of Rimmon\footnote{\url{http://en.wikipedia.org/wiki/Naaman}} if you have to, but pray without ceasing to the Illuminator of the Heart, that you might defeat the adversary, and be set free eternally. This is the meaning of the saying, \emph{if the Son makes you free, you shall be free indeed}.

But for those who resist, the Absolute III and the energies of the occult servants of the Holy Spirit will devour them.

This, and much more, is taught in Mouravieff's \emph{Gnosis}.


\flrightit{Posted on 2012-07-13 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{apeiron on 2012-07-14 at 17:20 said:}

Nice, but that link you provided to Luther's famous battle quote argues against the legitimacy of it. The quote was attributed to him in a fictional 19th Century novel by Elizabeth Rundle Charles however poetic it may sound. Luther would have done well to follow his own advice had he not come so late to the Jewish question, but by then the Revolutionary Reformation had blown full swing and Europe was dangerously partitioned with disastrous future consequences to follow.

Particularities aside, and to simply take what the essence of the message you are trying to convey on an objective level this is what we conclude: Don't fight the Satanic forces of the Kali-Yuga head on, (as to do so would bring the wrath of the Zeitgeist/Universal/General Law which at this point in time favors the Adversary) but instead form an elite of the few who remain awake in the night to save your own asses from the last judgement and to hell with the rest of the goyim. I'm still trying to figure where the heroic part comes in. Did Christ not incur this wrath under crucifixion(perhaps as initiation)? Did not the Iron Guard and so on?

The old' days still had men worthy of a higher caliber to oppose it, because despite all, the essence of that opposition carried a spiritual significant meaning that rings true to us.

\hfill

\texttt{logres on 2012-07-14 at 22:17 said:}

For the sake of clarity, your construal is not what I intend to mean. Saving yourself has nothing to do with it. Since I've just lost two lengthier replies, I'll content myself with noting that Mouravieff's concept of Authority in consonant with Tradition, \& that the issue here is: What are you capable of forcing the General Law to concede? I am not discouraging anyone who is capable from acting; what I wish to avoid is encouraging those who think that there is no warning here for us, as well as the Revolution.

\hfill

\texttt{logres on 2012-07-14 at 22:22 said:}

You fight spiritually just as you would physically -- using an Art of War, rather than acting without a precise knowledge of times and seasons, inner or outer. I am not trying to discourage that, but rather note that there is a warning here to us all. It was a point I did not want to make, but thought that knowledge and logic from the doctrine of General Law required be made. That's certainly not the last word on the subject, and would require another assay.


\hfill

\end{sffamily}
\end{footnotesize}

