\section{Making the Modern Mind}

Sometimes it helps to meditate on what is good by looking at the opposite, at least for those of us who only have ``intimations of deprivation\footnote{\url{http://www.dunedain.net/discussion/viewtopic.php?f=3\&t=1070}}'' and have not yet found our true center. This kind of ``cautionary tale'' meditation or meditating ``by negative example'' is very easy in Western societies at present. One need only take the nearest objects at hand in the mental furniture and stare at them awhile. You just need to know how to reverse the negative plate photographically in your own soul-mirror, which can be challenging. Still, if one knows roughly what the room should look like\footnote{\url{http://www.righthandpath.org/the-church-and-the-common-mind/}} by appealing to authority, then perhaps there is value in it.

What, one may ask, is ``self-evident'' about all men being created equal? That they are equally unique? Or equal in outcome? Equal in ability? It is not self-evident at all, on the contrary. The only thing truly self-evident is that men are singular, unique, personal\footnote{\url{http://www.amazon.com/MODERN-AND-AMERICAN-DIGNITY-Contemporary/dp/1935191896/ref=pd_sim_b_1}}, even when they are deprived of Being and live at the level of the animal. But this is a far cry from what Jefferson meant.

Locke (along with Ockham \& Bacon) is also regarded\footnote{\url{http://www.nytimes.com/books/first/p/porter-modern.html}} as self-evident, for it was the British Enlightenment along with ``classical liberalism'' which undermined the Anglo-world, rather than just the French Revolution. Yet what is self-evident about ``property rights''? If one takes it mystically as a ``given'', certainly, but that is not the position of the modern mind, which is the only possible position to effect a stand. Rather than respect the ancient homestead, modern states generally bulldoze them\footnote{\url{http://www.landrights.org/ocs/SocioCultural/BuffaloRiverInholders_2.htm}}. Property is not the ``family farm'' but ``my stuff''. The invisible hand of the market supposedly irons out any disparities or lack of balance, and Nature's bountiful table has room for all? But Locke wrote when the New World's resources were being discovered and settled, just as Rousseau wrote from the vineyards and shade trees of France. Property rights today no longer have the ``home as castle'' basis they once did. A friend of mine recently argued that we just have to take the mystery of property as a ``given''; this same friend rejects monarchy, whose basis was precisely this ``given'' and mystical ``thingness'' of the King owning his property, the kingdom.

Another friend thinks ``infinity'' is self-evident\footnote{\url{http://www.amazon.com/Beginning-Infinity-Explanations-Transform-World/dp/0670022756/}}, a given, that the horizontal and evolutionary world of the serpent will last, and last forever. This (however) is precisely what one could fear -- that the illusion of being God-less can perpetuate itself indefinitely in the long, slow decline\footnote{\url{http://thearchdruidreport.blogspot.com/}}. Others place their faith in ``singularity\footnote{\url{http://en.wikipedia.org/wiki/Technological_singularity}}'' that Science will soon give us. If a new planet, replete with earth-like resources, were to wander into our solar system, these same people would still deny the Destinies and Fates and the God who had brought it. It would be an excuse to continue on the present course.

All of these modern nostrums can be traced to handfuls of thinkers, which have been individually and collectively refuted, over and over again, to very little avail. Karl Marx, Max Weber, John Locke, etc. \emph{Yet the Modern Mind remains because it is a mindset and a compulsion}, a generator of self-sustaining egregorae, the spider of the fable told by Swift\footnote{\url{http://en.wikipedia.org/wiki/The_Battle_of_the_Books}}, rather than the bee.

If Chaos and Law are co-equal, this means that Chaos is pre-eminent. But the other point of view, prevails, for now. We are not arguing against an argument, but an ``anti-argument'' which refuses both reason and revelation. It can't even discover the riches of its own pagan heritage, because it's willing to borrow the worst elements of exoteric Christianity in order to destroy any supernaturalism which might unsettle its pathological existence. And over and over, in the face of the most terrible catastrophes, like General Haig ordering another push over the fields of Flanders, it proclaims it has made ``progress'' when all it has done is to waste even more time and resources achieving pitifully small objectives. This is the meaning of Progress, the anti-Middle Ages, the negative photograph of what should be made to exist.


\flrightit{Posted on 2012-03-24 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Caleb Cooper on 2012-03-24 at 13:35 said:}

As regards the horizontal world of the serpent (huge thanks to Gornahoor and especially Cologero for making me aware of Tomberg, studying the eighth arcanum right now) lasting forever:

It's my understanding that while universes die, the physical cosmos is an increasing infinity forever, but in a beautiful way to the glory of God.

Imagine an individual universe as a vertical cone, infinitely wide at the top where it's time begins, and narrowing as time passes until it converges on a singularity at its end time. The singularity is God, manifesting His full glory \_within\_ the physical universe. ``And I beheld his face from whom the heaven and earths fled... ''

One of the most arresting images from Revelation is God's full glory manifesting, which is so intense that it dissolves the universe. From the perspective of physics this is correct. Observation causes the wavestate to collapse. And within the wavestate, time does not exist, that is, a wavestate contains all it's past (and future, which is irrelevant in this case of the end time). So what would be the consequence of of God turning His gaze fully a universe?

Total wavestate collapse, which would cause all moments of time to simultaneously exist. And on the physical plane, all those instances would have mass, multiplying the mass of the universe by the number of plank units it existed for. There are about 1.85 x 10 to the 43rd power plank units in a second, so the number is ridiculous. More than enough mass to instantly destroy a universe in the final epikrosis, every unit of space potentially collapsing into a black hole large enough to form a universal seed, turning inside out and exploding in a big bang in a new dimensionally separate universe on the black holes other side.

(I suspect this may finally be the mercy that allows some of the `wandering stars' to die and return to the wheel that turns toward the true God. This is just speculation, but I suspect some of those immortal diamond body's are still anchored in the mid-world; being able to travel to the Celestial realms is not necessarily the same as fully transferring the seat of your consciousness there).

The upshot is that for every second the universe exists, 1,855,094,832,000,000,\-000,000,\-000,000,\-000,000,\-000,000,000 universes will be born. The suffering of the cosmos is ever increasing, the glory of God is every increasing, unto eternal infinity.

As it was in the beginning

Is now and ever shall be

World without end.

Amen. Amen.

\hfill

\texttt{logres on 2012-03-24 at 16:02 said:}

Thanks for the well thought comment, which is very beautiful, also. There is (of course) truth to ``infinity'' (and Tomberg insists on this) but it is only possible to appreciate and fully benefit from it within a vertical perspective (which you just gave). Most of those I know who make these kinds of arguments are making them precisely so they won't have to address God, the Self, the soul, the cosmos, etc. They are rejecting cosmology by hiding within infinity's sands.

\hfill

\texttt{Andrew on 2012-03-25 at 00:49 said:}

@ Caleb Cooper

A good image of Samsara, and while it may be the case, verily not something to celebrate from my own Buddhist perspective.

\end{sffamily}
\end{footnotesize}

