\subsection{Labors of Hercules, Part 2}

The story continues\footnote{\url{http://en.wikipedia.org/wiki/Lernaean\_Hydra}}: Hercules has an adversary, Eurystheus, who is the current king and holds regnum over Hercules' territory. Now, in the esoteric Christian tradition, Satan is not our ``enemy'', but our ``adversary'' -- there is a subtle, but distinct difference (Alain Benoist has come to the conclusion that the USA is an adversary, but not an enemy of Europe -- I'll let you draw your own conclusions). The adversary wishes to joust with Hercules, \& so he assigns him twelve labors. Hercules, having conquered his passions and mastered his thoughts, now becomes a serious contender. He, further, draws the ire of Hera, that champion of the eternal Feminine, who will test all men, \& who fashions the Hydra specifically to destroy the Hero. The adversaries are many of the spiritual combatant. Demons will lay specific traps, not for men in general, but for YOU specifically. They know your weaknesses, perhaps better than yourself. This is why the first adversary is the man's Ego or mind -- detaching himself from his ``stream of consciousness'', he begins to examine thoughts. Who am I, really? I am not Me.

Hercules will be tempted to play to his Alpha strength -- sheer power. The power of the man, the virile, masculine, war-making instinct, which levies eternal defiance to heaven and earth, if need be. As that old Scottish reaver\footnote{\url{http://books.google.com/books?id=JZYLAQAAIAAJ&pg=PA555&lpg=PA555&dq=redcloak+ballad&source=bl&ots=cCnQdzDt7h&sig=TVjb42Epzx\_JHuQIg9SBeHccTqo&hl=en&sa=X&ei=N9oKUsWiCMa02gWdi4DYBA&ved=0CCkQ6AEwADgK\#v=onepage&q=redcloak\%20ballad&f=false}} told Queen Elizabeth, ``What is there, that a man will not dare?'' Indeed, the power of man is great: witness the captive English hero who was invited to duel to the death in front of the Spanish court. Hercules, like the Englishmen amid his foes, will have to choose wisely\footnote{\url{http://ejmas.com/jwma/articles/2001/jwmaart\_docherty\_0501.htm}}: Peeke ended up killing one Spaniard, gravely wounding a third, and driving his rapier-wielding Alpha foes from the ring. He was set free, with honors. Or, as Robert Jordan wrote:

\begin{quotex}
Hammar moved to stand beside Galad, still groaning on the ground and trying to push himself up. The warder raised his voice to shout, ``Who was the greatest blademaster of all time?' From the throats of dozens of students came a massed bellow. ``Jearom, Gaidin!'' ``Yes!'' Hammar shouted, turning to make sure all heard. ``During his lifetime, Jearom fought over ten thousand times, in battle and single combat. He was defeated once. By a farmer with a quarterstaff! Remember that. Remember what you just saw. During his lifetime, the greatest blademaster fought over ten thousand times, in battle and single combat. He was defeated once. By a farmer with a quarterstaff! Remember that.''
\end{quotex}


In this combat, Satan can be viewed as an honorable but cunning adversary -- the test can be passed, if wisdom is added to power. Since Hercules will need wisdom, his kinsman Ioalus will provide it (or the gods themselves).

Hercules covers his mouth and nose, symbolically giving precedence to the eyes. He is no one-eyed monster who cannot distinguish good versus evil, but watchful and alert. He approaches the den of the Hydra, and awakens it with flaming arrows, which show that the hero is aware that he will need to draw the beast forth, rather than climbing down into its hole, like Beowulf, after the dragon.

Like a true man, he begins combat, but finds that for every head of the Hydra he severs, two grow back. This shows that fasting, prayer, and mortification of any kind are inferior to watchfulness. They only awaken the passions: they draw it forth. A man who has fasted learns that he is made of carbon, water, and other base materials. If he relies solely upon his own strength, his combat will fail.

His kinsman has the idea of supporting his combat by burning the heads. The aid of other men in spiritual combat is great, indeed. While Hercules concentrates on destroying each head, Ioalus sears the stumps. In an alternate telling, Hercules uses the blood of the Hydra to sear the stumps. This is why the innocent who refuse to undertake combat are inferior to the man who fights -- only he who fights can take the blood of demons and use them to cauterize the wounds of the monster. Additionally, further ``shocks'' are necessary than individual effort -- by the law of Three (discussed in Gurdjieff/Ousepensky's systems), man requires the aid of group and cultural effort to sustain him over the natural gaps in his own line of spiritual effort. They fill in the critical gap (see Psalm 73\footnote{\url{http://www.bartleby.com/108/19/73.html}}), allowing the awe and wonder of external comradeship carry a man over the desert of his own failure, which is innate. Even Hercules requires this help. Even Hercules has to have a town to defend, or ruined temples to haunt, or just assistance. Lone voices only attract wolves, unless a man is born in the mystery of the just, in which case, he will find himself constituting both a Church, a Culture, and a State. This is not a self-elected position -- others will confirm this for you. Until that time, it's safe to assume otherwise (if you're wrong, angels will show up to reason with your modesty).

After his hard-won victory over the multitude of the Passions (Satan's deputies to handle his petty work), Hercules buries the head of the Hydra in a sacred place on the royal road, dedicated to Zeus. The victorious soldier gives credit to St. Michael \& the host (or whatever name they assume in your mythology). This is additionally a good idea, because this enemy is not the last that Hercules will undertake -- it always helps to have the odds and the gods on your side, or better, that you are on theirs.

The assault on hell (and the storming of heaven) is not the work of a moment, nor that of the drifting cowboy. It also entails the risk of making enemies which may succeed in turning you into a martyr: if you are constitutionally opposed to self-sacrifice in all instants, then you are not ready to be recruited for the Legion\footnote{\url{http://en.wikipedia.org/wiki/Iron_Guard}}, \& it will be difficult to rise (although it does sometimes happen). Even outsiders have to have a ``meteoric rise'' if they come in \& eclipse the ``regulars''. In this sense, the Christian path is not ``one path of many'', but indeed the only way such things could happen, in any case. How else can God overturn Hell, but from the inside out? Will there not be casualties? Dangers? Risks? Of course there must be. Christ came as a lamb, because He knew this was the way to become the Lion.

At the level of the Church, the Hydra is heresy and error, which leads to apostasy. At the level of Society, the Hydra is ideology or ersatz religion, which always threatens to whelm organic growth. One might say, the ``anti-State'', in our day. The Hydra is a cancer, and it devours all that it feasts upon. Passions, heresy (Hyper-Calvinism tweaked)\footnote{\url{http://unqualified-reservations.blogspot.com/2007/06/ultracalvinist-hypothesis-in.html}}, the mega-State... ..they offer us, at first, escape from our individuality and personhood, our discipline. They offer to initiate us in the multiplicity of delusion.

A prophecy\footnote{\url{http://www.catholicism.org/clement-maria-hofbauer.html}} of the era of~ ``new men without chests'':

\begin{quotex}
Almost as a postscript to the heavenly warning issued at Fatima in 1917, Saint Maximilian Kolbe, two years later, reviewed the three Great Evils of the latter times, noting: ``In 1517, the Protestants rebelled against the Church; in 1717, the Freemasons rebelled against Christ; and, in 1917, the Communists rebelled against God.'' In a single sentence the Polish martyr had exposed common origin and natural succession of each of these Apocalyptic nightmares.
\end{quotex}


There is a story about Clement Maria Hofbauer, how he went into a bar to beg for money for his orphans, \& a clientele spit into his face. He said to all there, ``That was for me. Now what about for my boys?'' Needless to say, a man like this leaves the bar with the money he came to find. I doubt if he spoke with any fear, lack of dignity, or craven Semitic bearing. And somewhere, in heaven, his angels slew several legions of demons, in honor of his conduct. Is this really degenerate spirituality? You want that man heading the Church, or someone very much like him.

Now I would ask those who are consciously ``not-Christian'' (many for very sound reasons) to please consider this appeal carefully. Obviously, you cannot acquiesce in what is without honor, that is true. So the appeal is to honor -- where is the assault falling today? \textbf{Even if Christianity is the ``vehicle'' for the Kali Yuga, \& destined to be transformed into something else, the rear guard is where the fighting is at}: over the fallen altars, at the throne. Sure, martyrdom is a possibility, physical or otherwise, yet, isn't this what the path of honor smiles upon? The new age will be determined by the blood spilled at the crux of combat: regardless of Christianity, it is spiritual blood that will seed the future. Sacrifice is a Christian-Semitic principle only in the sense that it was given high-articulation within those vehicles. The Ideal is true. It may need altering, or qualification, or balancing, but Christianity apprehended something in Tradition that was true, \emph{regardless}.

When the Franks where whelmed at the mountain passes, Roland\footnote{\url{http://en.wikipedia.org/wiki/The\_Song\_of\_Roland}} was given the honor of leading the stubborn \& doomed rear-guard action. Today, the clueless conservatives\footnote{\url{http://www.chiltonwilliamson.com/articles/wwwtw\_leftist\_rage\_conservative\_hate.html}} are leading an ineffectual rear-guard action, and there is no intellectual order. \emph{What else do we need today other than the men who are not welcome in the Church to insist upon being in it?} It is not only an option to request service in the rear-guard, \emph{but is in fact a high honor}, almost a duty to request a post there (even if denied). Chivalry, Western high culture, was merely the high and strong and noble apprehension of Truth, accentuated by the vehicles used to attain it, colored by the nationalities and souls involved, articulated in the languages and thoughts of those who served.

It is this Truth that we serve, all of it, no one part at the expense of the other, and I cannot forget what happened in the deserts of Thebaid, when the passions were tamed by the wild monastic fathers, \& the fruit was grown and reaped in the great ``Romance'' that swept over the barbarian tribes.


\flrightit{Posted on 2013-08-16 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Michael on 2013-08-17 at 20:08 said:}

Powerful post Logres. It doesn't take too much discernment to see that the assault is falling primarily on Christianity. When is the last time you heard a New Atheist attack Islam? And it is one Christian communion in particular that bears the brunt of the attack --- the Catholic Church --- and I don't mean the SSPX.

I know that many readers here don't think there is enough of tradition left in the Catholic Church to be worth joining and defending her. But the forces of the counter-initiation seem to think differently.

\hfill

\texttt{John Curley on 2013-08-18 at 02:03 said:}

Probably one of the most compelling arguments I've ever read for the conversion to Traditional Christianity.

\hfill

\texttt{Doug on 2013-08-18 at 11:53 said:}

It is indeed an excellent post. We should remember that ``Assuredly, I say to you, whoever does not receive the kingdom of God as a little child will by no means enter it.'' The ascetic discipline of choosing(or being chosen by it in sphynx-like fashion-see Coomaraswamy) the path will go far to aid the aspirant if he is capable of slaying the several heads of the Hydra (the illusion of separate paths). That Orthodox Christianity is the current example of the Fullness of Truth(that is, every viable path is contained within it) should be apparent by the specific and focused attacks of the Adversary who recognizes a legitimate target when he sees it. The path of honor is clear. Those who are willfully negligent through a polemical view of pseudo-Christianity given to them by a Radical Secular culture are guilty, as I was at one time, of a rejection of a hard truth that is not in accord with second-function life.

\hfill

\texttt{David on 2013-08-18 at 11:56 said:}

I will present some observation on the subject of Herakles that may or may not be in direct link with what you already wrote. I had written those observations down to take them up eventually to write an article. As Logres already did (and did it well), I will only expose some of the information here.

Herakles is a «mute» character; that is, he doesn't partake in a major form of greek writing by himself (that is, tragedy, comedy, etc.). For this reason, he was genuinely taken up by the philosophers, the warriors, the Christians, etc. as a representation of positive value. Beyond comparison (which are very interesting) between Thor and Herakles, or Gilgamesh and Herakles, there is some interesting feature to shed light on, as Evola already did.

One of the main themes that are associated with Herakles, especially in the Christian world is that of his meditation on the fight between Virtue and Desire. As Logres already explored, there is an ascetic form displayed by the various fights of Herakles reminiscent of a more ancient Tradition.

As hinted by Logres and developed in other texts, Christianity can be seen as a warrior religion. It is not to diminish Christianity to a purely warrior religion, but rather to see in Jesus Christ a Royal Principle that both transcends the priestly and warrior class. At moments, he embodies both of them perfectly. There is even some comparison that could be drawn between the figure of Herakles and Jesus Christ. Also, the use of exterior (godly) help by Herakles is a theme dear to Christians : we cannot do everything by ourselves, we need Divine Love/Help.

Dumézil pointed out that Herakles was a foundation and representation of the second indo-european function, that is, the warrior. In his passing of his weapons, he made sure that there was a line, a tradition per se of the warrior ethos and spirit.

To Evola, Herakles is a pure principle of the warrior class. He links to several other interesting ideas linked with Herakles or heroes in general :

-- The Female (as mystical power of life married or conquered) as the source of regal power (i.e. Heracles and the Tree); see also the tree of the World in general linked with heroes of Order

-- The Hero representing cosmic principle of Order winning over the Chaos (on which I will return shortly when discussing the twelve labors) sometimes for the gods, sometimes even with the gods.

-- The Hero as founding figure of a tradition (mysteries, initiatic order)

-- Rig Veda states that the sacrifice/rite was performed by a «virile hero». The ascetic and the Hero (Warrior) as representation of spiritual virility.

-- Hyperborea hidden from the earth (only opened to heroes such as Herakles)

-- Heroes may attain a state similar to that of the Primordial Age

The twelve labors are twelve civilizing labors. They bear witness of the primordial principle of Order against Chaos. The hero, that is Herakles, becomes the channel through which the Order can unite itself with the world to sanctify it. The Indo-Europeans in general were a civilizing people, bringing social order with them (class/caste). It is therefore fitting that one of the most recognizable hero of the Greco-Roman mythology is civilization in itself. There is an ambiguity in Herakles reminiscent of the very principle of civilization and warriorship : there is a facinans and a tremendum, a fascination and a fear toward Herakles. He is able to chain Chaos and bring Order, yet at the same time, he channels both of them in himself and finally he transcends both of them (i.e. his ascension to godhood).

To the stoic, Herakles was one of the first of them. His acting through detachment, for the greater good of the cosmos, through sheer will and representation of the Will was for them a clear exposition of the doctrine. As we already spoke of the Virtue contra Desire theme, we will not return on it, but it is quite fitting. He was also seen by some as a proto-philosopher (as in Socrates description of the philosopher).

There is also a possibility of more profound meditation on Herakles journey through greek mythology and philosophy. First of all, there is a human part in Herakles (that is, an animal part) that makes him demi-human (touched by eros and thumos) : in eros he rapes Aule, indulge in gluttonery, etc. and in thumos he kill his teacher out of rage, commit act of violence, etc. Yet at the same time, those things do not control his life; they are subordinated to the Self. In this he regain his demi-God part, performing deeds fitting a warrior-philosopher, ultimately that of a King (being divinized), in killing lawless monsters, founding cities, civilizing (through execution of religious rites), etc.

As he is a human-god, he is also a god-human, that is, a divine being. We already hinted to this in his participation of the Principle of Order and his acting as a warrior-philosopher-king. But more than that, he is, in some version of the story, elevated as a god himself, and married to Hebe, goddess of youth: therefore Orderly Power is linked with Life and Youth.

There would also be some important things to look between the link existing with Herakles and the initiatic mysteries of Elysium. But that will be for someone else to think about and work upon. Evola talked a bit about the resurrection and afterlife of Heroes in «Revolt» p.49-50.

\hfill

\texttt{michel on 2013-08-21 at 13:13 said:}

``You got the sharks on the left and the snakes on the right, and anything you do they wanna get a little bite, it really doesn't matter if you're wrong or if you're right coz, once they get their teeth in, NOTHING REALLY FIGHTS!''.


\hfill

\end{sffamily}
\end{footnotesize}

