\section{Phantastes}

It must be said (for subscribers to the \emph{Sintesi}) that the Inkling project manifests a decidedly ``Dionysian'' accent: that is, it is the spiritual striving of those who do not possess the Truth innately, but are able to aspire, yearn, and perhaps awaken to that Truth through art and particularly the vegetative mystery rites which survive in the Dionysian mysteries.

\begin{quotex}
The other characteristic of solarity is that of a light that arises and fades, that it has death and resurrection and a new death and a new dawn and, in other words, a law of becoming and self-transformation. This is Dionysian solarity as opposed to the Apollonian principle. It is a virility that aspires to the light through a passion that cannot free itself from the sensual and telluric element and even from the ecstatic-orgiastic element typical of the lowest forms of the Demeterian cycle. The association, in myth and symbol, of feminine and lunar figures to Dionysius is, in this regard, rather significant. Dionysius does not accomplish the transition, the change of nature. It is a still earth-bound virility in spite of is luminous and ecstatic nature. The fact that the Dionysian and bacchanal mysteries were associated with the demeterean mystery, instead of with the purely Apollonian mystery, clearly indicates to us the final point of the Dionysian experience: it is a ``dying and becoming'' in the sign not of the infinite which is above form and the finite, but of that infinite that is fulfilled and delights of itself in the destruction of form and the finite, harking back therefore to the forms of telluric-demeterean promiscuity... 
\end{quotex}

JI Packer, a Protestant theologian, remarks somewhere that the Celt manifests a deeper apprehension of spiritual realities than their Anglo-Saxon counterparts, which is bourne out in the history of Puritan theology (think John Owens). Indeed, Evola cites the Celt as specifically prone to the matriarchal side of spirituality, and therefore, the Dionysian path may represent the closest collective approach such a spirituality can make to Tradition. So,when we praise the Inklings, please keep in mind that their approach is in some ways indirect -- Williams, Barfield, Lewis, and MacDonald were all to some degree Celts. Everyone but Tolkien, who of course had a Catholic or more Roman approach to spiritual matters.

Mouravieff comments on racial differences in \emph{Gnosis}: the Russian (for instance) often has two magnetic poles in his personality, which stymie each other. Since Dionysian striving represents a common situation for the aspiring--noble, modern man of the 19th and 20th century, Mouravieff attempts to explain Tradition as if speaking to Dionysius; that is, he tries to strengthen the longing, and to properly diagnose the poles within the seeker, so as to enable magnetization.

\emph{This is precisely what the Inklings, and more specifically MacDonald, are doing in their fiction.}

Here is MacDonald on self-knowledge, or introspection:

\begin{quotex}
Afterwards I learned, that the best way to manage some kinds of painful thoughts, is to dare them to do their worst; to let them lie and gnaw at your heart till they are tired, and you find you still have a residue of life they cannot kill.
\end{quotex}

Renunciation:

\begin{quotex}
Joy cannot unfold the deepest truth, although deepest truth must be deepest joy.I wished to be... no longer a man beside myself.
\end{quotex}

Evil:

\begin{quotex}
... I was almost glad that I had sinned...
\end{quotex}

Death:

\begin{quotex}
``The hot fever of life had gone by, and I breathed the clear mountain-air of the land of Death. I had never dreamed of such blessedness. ... I lay thus for a time, and lived as it were an unradiating existence; my soul a motionless lake, that received all things and gave nothing back; satisfied in still contemplation, and spiritual consciousness. 
\flright{\emph{Phantastes}, 314}

Thus I, who set out to find my Ideal, came back rejoicing mat I had lost my Shadow. If my passions were dead, the souls of the passions, those essential mysteries of the spirit which had embodied themselves in the passions, and had given to them all their glory and wonderment, yet lived, yet glowed, with a pure undying fire. They rose above their vanishing earthly garments, and disclosed themselves angels of light. But oh, how beautiful beyond the old form! 
\flright{\emph{Phantastes}, 314-15}
\end{quotex}

Sin:

\begin{quotex}
Alas, how easily things go wrong! A sigh too much, or a kiss too long, And there follows a mist and a weeping rain. And life is never the same again. (235)\footnote{\url{http://www.snc.edu/northwind/documents/By\_work/Phantastes/The\_Sources\_of\_Phantastes\_-\_John\_Docherty.pdf}}
\end{quotex}

Owen Barfield points out that MacDonald is allusive, \textbf{but not elusive -- he puts his finger on the truth and does not merely have an ``inkling'' but speaks whereof he knows}:

\begin{quotex}
Is there not in this poem a certainty, a grounded knowledge? It is not content to stop in imagination and hint and suggestion. One feels that its meaning, its openly expressed meaning, reaches right down into the solid earth and right up into the empyrean. It is the resurrection of the body---in terms of the body.
\end{quotex}

Just because MacDonald focuses upon the Phantastes or feelings does not entail a neglect of will and thought:

\begin{quotex}
Anodos is oblivious of any evil menacing \emph{his} House of Alma, the ``fairy palace.'' However, this palace contains twelve halls of dancers/statues representing shades of feeling. (These twelve satellite halls reappear as the mood chambers of the wise woman in \emph{The Lost Princess}.) Presumably it is because MacDonald is describing only the areas of life which are the concern of Phantastes that his fairy palace is solely a palace of the feelings. Anodos, however, does experience ``wilful'' aspects of feeling when he pursues the marble lady out of the palace, and ``intellectual'' aspects of feeling in the library of the palace.
\end{quotex}

In one of MacDonald's letters to his wife, the melding of Christianity with vegetative rites becomes somewhat plainer, breaking out into the clear daylight of his prose: 

\begin{quotex}
Some of the dark closes \& entries look most infernal, and in the dim light you could see something swarming, children or grown people perhaps, almost falling away from the outlined definiteness of the human, . . . Dearest, you must come here with me, you would be so interested. It is like no other place . . . You know Edinburgh is built very much up and down hill; and so in some places narrow closes, some so narrow that your little arms could touch both sides, run from top to bottom of the hill through these great, tall houses. Glancing down one of these I was arrested. It was Very narrow and went down, as if to Erebus, and suggested bad and dangerous places, down into the unseen and unknown depths. But across the upper part was barred the liquid hues of the sunset, against which stood the far off hill with some church, tower or something of the sort in relief against the infinite clearness. . . . Dearest, I hope you will not be frightened tonight. God, the Sky God---the Green Earth God be with you, our own God, as David says.
\flright{\emph{Letter from MacDonald to his wife}, 1855, in Sadler 87-88}
\end{quotex}


\emph{MacDonald seems to be wrestling with reconciling the mystery religions with Christianity (female with male).}\footnote{\url{http://www.snc.edu/northwind/documents/By_work/Phantastes/Chthonic\_Aspects\_of\_Phantastes=\_From\_the\_Rising\_of\_the\_Goddess\_to\_the\_Anodos\_of\_Anodos\_-\_Fernando\_Soto.pdf}}

Indeed, Anodos himself confesses this to be his own project during his sojourn through Faerie Land:

\begin{quotex}
Perhaps, like a geologist, I was about to turn up to the light some of the buried strata of the human world, with its fossil remains charred by passion and petrified by tears... 
\end{quotex}

I won't unveil every detail of the plot and structure of Phantastes, although in brief it is the story of a man's quest for the ``White Lady'' \& his renunciation of her in favor of a ``better man'', partly due to the fact that the protagonist continually tries to ``touch'' or ``hold'' her prematurely, thus destroying the magic that is re-creating them both. Sir Anodos, in this story, is unable to distinguish between legitimate and auspicious impulses which lead to salvation, and those which lead to damnation (eg., the summoning of his shadow in the chapel of darkness wtih the ogress). It is a fugue or dream state: Sir Anodos acts on blind impulses, which are automatic -- some are good, others are bad, although, the ending reveals that even the bad is the ``shape which the Good was forced to assume'' at that time and place to bring greatest help.

MacDonald's world corresponds to a post-mortem state -- the main character has ``died'' and is journeying through the intermediate worlds, looking for his polar being, and for heaven. 

I do not think that the Inklings' work is for every reader; some will be put off by the ``childlike'' qualities of Narnia, Middle Earth, or Faerie Land. However, I am maintaining (for all that) that what we have here is not ordinary children's fantasy, but complex and spiritually subtle allegories which are meant to ``land upon the precise and needed point'' in order to generate spiritual change. That this change would occur in the Dionysian individual is another, less obvious point, which would explain why their work was largely rejected as the century wore on. Men increasingly dominated the landscape with a sub-Dionysian spiritual state. 

What the Inklings are really doing is charting a course through ``Fairie Land'' (either in this life, or the next) which is valid for a certain kind of spirituality. As such, those individuals can and should read them to great advantage.


\flrightit{Posted on 2013-05-24 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Jacob on 2013-05-25 at 13:29 said:}

Hi, I really enjoy reading your articles.

As I began reading Phantastes, I thought it at least in some ways reminiscent of Dante. Mainly because of the plot of chasing a hard to obtain lady through fantastical settings. Do you believe some of the difference between the inklings and say a writer with a better understanding of tradition like Dante could have to do with the influence of Steiner?

Also, where did your first quote come from?

\hfill

\texttt{Jason-Adam on 2013-05-25 at 15:29 said:}

Crowley could also be perceived as a Dionysian ?

\hfill

\texttt{Logres on 2013-05-25 at 22:43 said:}

Jason-Adam: qualified Yes; from what I know of Crowley, \& without sitting in judgement on his soul (which the Church prays for), Crowley seems to present the challenge of someone who didn't want to save his soul, just his spirit. But I am not qualified to speak of Crowley, since I've only hearsay, plus the one essay on him by Evola here at Gornahoor's library. Still, there is a difference between someone who remembers the Sun during the night, and someone who attempts to make themselves burn as a Sun to the night itself.

Jacob: the first quote is from the Society of Independent's translation of Evola's Sintesi, done by Cologero. And, yes, Dante's grasp of Tradition/Truth seems stronger, and there is no doubt that Steiner is problematic in ways that Dante is not, although MacDonald came slightly before Steiner's career, in any case.

\hfill

\end{sffamily}
\end{footnotesize}

