\section{The Lutheran Reformation}

It's October, and I am thinking, inexorably, of the Protestant Reformation, where my spiritual roots lie. In order to do those roots justice (it does no good to saw them in two), I have to understand the contention made against the Church of the West, the Latin Roman Catholic Church. Now there are two types of attacks usually made on the Church: one of these points to specific abuses (in the modern day, these have to do with altar boys, but I qualified this ever since I found out\footnote{\url{https://en.wikipedia.org/wiki/John\_Jay\_Report}} that most of the perpetrators were priests the Church let in under pressure from Vatican II reforms and the Cultural Wars, after the 60s, in the decade 1970-1980), and the other has to do with the doctrine of Sola Fide. So that it is important to remember that the RCC was liberalizing when it got burned: that is, it let in a lot of alternative candidates who would not ordinarily have been anything but rejects, following the pressure of the spirit of the Times, and predictably, got what it deserved. But this is not really an argument against the RCC per se, so much as it is a cautionary morality tale: better to have stood against the world and gone under, than to have compromised and then tried to hide the sorry result. I assume that indulgence sales must have operated along the same lines: lots of pressure from secular rulers to loosen up those tight demands, and make things more accessible to the money changers and to Mammon.

It is easily seen that Sola Fide was an attempt to cut open the circle of God's decrees at one proper, precise point, the point where the energies of God touched the response of man. The circle of God's decrees met at this one point, in the response of faith. This was done so that man could take a minimum of credit (belief, or trust, is a very responsive, passive, spiritual reaction to the action God had already objectively accomplished in the finished work of the cross), and so that God could get a maximum of glory (the faith response only proved that the Spirit had regenerated a man. From above, God reaches down in election, but the regeneration operates from below: both are accomplished through that eternal love that operates from the foundations of the world. Thus, even in salvation, man is caught in the pressure of God's Providence acting from above, and God's decretal purposes entering from below (AA Hodge said that regeneration occurs in the subconscious). The design of this was to preserve second order acts (Free Will and the apparent synergism of having to ``come to Christ'' in an act of Faith), without sacrificing (or at least sacrificing a minimum of) God's sovereignty. Of course, inevitably there were theologians and sectarian leaders who wondered, out loud and in print, that if there was no mystery in the Eucharist, why should there be any mystery in the \emph{Ordo Salutis} either? So their conclusion was that no Law at all was required: if you were saved, you were saved, and if not, you weren't. This only pushed the mystery back to why there was such a ``salvation'' or Supernature at all, treating the Incarnation and all its effects like a cancer that needed shrinking. As we shall see, Revolution does precisely this: it is, at root, a vast oversimplification (which explains the apparent emphasis on sophistication in ``matters indifferent'', ie., technology). Luther himself was no Revolutionary, and didn't anticipate fully where his ideas lead, although he saw enough\footnote{\url{https://en.wikipedia.org/wiki/German\_Peasants\%27\_War}} to smell trouble. Likewise, Calvinism itself is a kind of ``shorthand'', which is true as far as it goes. Here is what Evola said of it:

\begin{quotex}
As for Christianity in its less popular forms, it presents an aspect of the tragic doctrine of salvation, which to some extent preserves an echo of the ancient truth: the idea---pushed to extremes by Luther and Calvin---that man on earth stands at the crossroads between Salvation and eternal damnation. This point of view, if lived intensely and coherently, could create the conditions for liberation at the moment of death or in postmortem states.
\flright{Note 2, page 96, \emph{The Hermetic Tradition}}
\end{quotex}

Note that Evola (unlike Luther or especially Calvin) is not committed to the idea that this ``shorthand'' represents a comprehensive treatment of spiritual matters. In all of this, the intent was flawlessly impeccable. The Reformation presented itself as a restoration of the full orbed, original Gospel, over against the new Phariseeism of works religion that had forgotten what the point was. Indeed, Luther's Anfechtung\footnote{\url{http://www.orlutheran.com/html/anfecht.html}} was actually an experience of moral collapse in the face of the demands of the Gospel, a collapse so total that only self-despair could provide the impetus to embrace Christ in an act of flying to the Cross. Luther was, for instance, interested in the \emph{Theologica Germanica}\footnote{\url{https://en.wikipedia.org/wiki/Theologia\_Germanica}}, which spoke of such spiritual states.

What could be wrong with this? Well, in a sense, nothing. In fact, it is true that Calvinism and Lutheranism potentially were Restorations of the Gospel. That is exactly what Europe needed: not a Renaissance, but a Restoration. Instead, it got a Revolution. If Lutheranism and Calvinism had preached the comfortable truth that God calls a man as he is, and declares him righteous by imputation to begin the process of salvation, the Church might have been Reformed. Instead, it wasn't Reformed at all, and when the Reformers split off from the Church, by definition, they were admitting that the goal was no longer Reformation. Instead, in places like England, a minority of nobles decided to use it as a pretext to rob the Church and strip the altars\footnote{\url{http://www.amazon.com/Stripping-Altars-Traditional-Religion-1400-1580/dp/0300108281/ref=sr\_1\_1?s=books&ie=UTF8&qid=1414554060&sr=1-1&keywords=eamon+duffy+stripping+altars}}. In the Netherlands\footnote{\url{http://www.amazon.com/Calvinist-Preaching-Iconoclasm-Netherlands-1544-1569/dp/0521088836}}, mobs of idiot villagers lead by hedge priests smashed works of art that they couldn't even comprehend, let alone create. And in Germany, all hell broke loose.

What could have been an opportunity for the Gothic, North-West European culture to assume its rightful and perhaps even dominant place within Christendom over against an overly Baroque, ossified and undoubtedly corrupt leadership emanating from Rome, instead became a long Deformation and Degeneration of the Church, which has ended in our day by the secular hyper-Calvinization of the religion of Progress. This was, to be clear, the fault of all parties in Europe, on either side.

In addition, the study of the Ordo Salutis\footnote{\url{https://en.wikipedia.org/wiki/Ordo\_salutis}} and the doctrines of Grace (TULIP\footnote{\url{http://www.theopedia.com/TULIP}}) lead to a conviction on the Protestant side that if you didn't speak the precise language of the debate, and agree with the way of framing the question in a specific and penultimate way, you were a heretic and anti-Christ. This was partly understandable, given the Catholic intransigence and stubborn insistence that ``surely we have to contribute something to our salvation!'', which amounted to agreeing with the way the Protestants framed the debate. In fact, both sides framed the debate in essentially legalistic terms (right down to words like imputation and justification), and conducted it like a court room trial. This was merely a continuation of the Western fascination with innovative abstract categories (you thought the Greeks were bad?) that had lead to the East-West split to begin with. The fall of Rome was so damaging in the West that they had literally lost on a large scale (not totally) the doctrines of illumination and glorification for instance. It was so bad that, for awhile, Irish Christians were in closer communion with the spirit of Egypt than most of the West. Whole provinces (eg., Spain) remained Arian for extended periods of time.

So this was partly a hangover from the time of the Invasions, when God was pictured as the just judge\footnote{\url{http://en.wikipedia.org/wiki/Dies\_Irae}} coming to town for the hanging assizes, because this was the day-to-day experience to begin with, the raw theological data of human experience. Meanwhile, the Eastern Church acquired a Byzantine varnish that further confused the newly civilized barbarians. The West never really gotten over the fall of Rome, and Luther came along to finish the job. This was a lost opportunity, historically speaking.

So where does this leave me, contemplating the Reformation? I think that the theology of the Cross (as opposed to a theology of Glory) paid some dividends, that the particular way of cutting open the circle of God's mysterious decrees that the Reformers preferred yielded priceless insights, and that the Northern European Gothic spirit\footnote{\url{http://anglicanrose.wordpress.com/2012/07/23/northerness-respondi/\#more-2822}} was probably going to erupt at some point, unfortunately choosing the spiritual arm of Society as its main target (later, would come the turn of the kings). John Ruskin discusses some of this in his \textit{Bible of Amiens}, which I have posted on, and readers can refer to Cologero's posts on the Three Orders\footnote{\url{http://www.gornahoor.net/?tag=the-three-orders}} as a start. John Romanides' work is invaluable for illuminating the debate between East and West but reaches very unhelpful conclusions: what the West was, it was, and it might have been much more. Unfortunately, rather than attain spiritual pre-eminence, the Northwest of Europe chose the path of Empire, which (of course) was no longer Holy, but based on mercantile and raiding trader powers, the hated ``Anglo-Atlanticism'' which Alexander Dugin wishes to bring low. The Reformation set the stage for this by shattering spiritual unity, and further muddying the waters in its outcomes. There is still much to be gained and seen by what it did, both positively and negatively (as we do not wish to imitate the party spirit).

However helpful Reformation theology is (and most people don't know nearly enough about it), it is not helpful to have a naive belief that a particular facet of the Truth contains the sum of all Truth. However true this might be, were men angels, and could they see with angelic sight, mortals (even with the perspicacious Word of God to guide them) have to consider ALL facets of God's truth. Luther wanted to throw out the book of James, for instance, an ``epistle of straw''. And he seemed to have no knowledge of the ancient, ancient doctrine (both classical and Catholic) that the ``Word of God'' was much, much more than merely the enscripturated text of the Bible, or even the act of Incarnation: it was everything in between as well, because the worlds were created through a Living \emph{Logos Tomeus} (this is in your Bible). Additionally, the doctrine of Recapitulation raised the question of the archetype of the Father in relation to the summing up Word or Logos, which means that at the end, it is not the Word that is Supreme because the Word will return everything to the Father. Luther hated philosophy, so he couldn't be bothered with such trifles.

What he did know and realize, was that something had gone very, very wrong in the West. On this account, he was most certainly even more right than he understood. Luther was living in an era in which the Church had already distanced itself off from the German mystic tradition in the North (Tauler, Boehme, Eckhart) and almost cut itself off from its own tradition in Italy (Italy was prepping for Machiavelli and the Renaissance). So his legitimate spiritual experience found no outlet, and he had to formulate the dogmas of the Reformation almost single-handedly, forging them in the fires of his conscience. This would have been a difficult task for a man much better prepared than Luther, and with more help. Like Lenin after Marx, Calvin came after Luther and ossified the Reformation, guaranteeing that the mystery of the Incarnation in its full aspect would be viewed through a very narrow lens, however accurate for purposes of establishing Christian security and comfort, a lens which would be used to focus the fires that would engulf Europe in Revolution.

And this is the tragedy: God had intended a Reformation. However, He has also revealed that He intends to show every work for what it ultimately is. \emph{In this case, Reformation laid the ground work for Revolution. We have to Reform the Reformation.}

To my understanding and knowledge, there is only one Protestant thinker who has ever made or attempted a conclusive study of this process of Reformation, which he failed to see was inherent in some of its doctrines and emphases: Groen van Prinsterer\footnote{\url{http://en.wikipedia.org/wiki/Guillaume\_Groen\_van\_Prinsterer}}. This is because in most cases, Reformation was accomplished through Revolutionary tactics, and its bad form to bite the hand that feeds you. Van Prinsterer is an anomaly, although Alexander Vinet is an interesting thinker.

It is, for instance, a Revolutionary tactic to throw out babies with bathwater. A legitimate grievance is advanced, and then it is used to justify a comprehensive programme of rebellion. This is precisely the pattern that we see in the Reformation. It started with legitimate insights into truth which badly needed to be shared and spread, and righteous objections to horrific abuses, which badly needed to be stopped, and ended up by tearing apart the body of Christ, killing a quarter or third of Germany and other parts of Europe in the process. The office of pope was tarnished forever, the safety of tradition was undermined, hierarchy was threatened in any form, and most of all, Protestantism lost all contact with mystical and ancient traditions, which resulted in the fomenting of a potent brew of materialism (the Dutch Empires gave rise to other Empires), secularism (particularly in England, where sects had their heyday), and revolution in France along with scepticism and ultimately atheism in Germany (who gave us higher criticism).

How can we honor Luther during Reformation October? We can embrace the truths he so tenaciously clung to, but we won't really embrace the spirit of Luther until we learn to see farther than he did, and to rectify the errors he inevitably set loose. We have to Reform the Reformation. Luther would surely be the first to proclaim that God must have other Reformations in mind, that he was not Christ, that if popes and councils can err, so can Reformers, and that surely his discovery and recovery of Truth (if real) must of necessity lead to even more Truth?

We have slept in Zion. The secularists have claimed Luther as their own. The revolutionaries are faithful to his spirit. The atheists praise him for making every man his own priest. The schismatics and lunatics thank God for Luther's Bible, which enables any idiocy to take the stand along with the most venerable and sacred mysteries in God's Creation. But we, the possessors of his lineage, do nothing but repeat verbatim what he said. Luther once said, The Letter Kills. Let us recover and discover Luther again, and let us find what he was looking for: absolute certainty in the almighty grace of God to pardon unspeakable sinners, but let us understand that the mysteries are not thereby done away with, but rather grown exceedingly great. Let us recover a Theology of both Glory and the Cross, and eschew the revolutionary tactic of pitting one thing against the other in the body of the living Christ, the Man-God, the eternal Word, whose Kingdom shall have no End, and who is coming again like lightning from East to the West.

What we need is a Protesting Catholic Western Church; what we are getting in America, instead, is a late Catholic form of degenerate Protestantism. Is it too late to Reform the Reformation? Our spiritual giants and ancestors showed us how to make a beginning. With their example, we can perhaps avoid the ending.


\flrightit{Posted on 2014-10-31 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{X on 2014-11-01 at 00:55 said:}

Each time I read what you write, Mr. Logres, it feels like eating a rubber.

\hfill

\texttt{X on 2014-11-01 at 07:20 said:}

No, sorry, after I have eaten it I discovered that it is actually meat which looks like rubber.

\hfill

\texttt{Joseph O'Sullivan on 2014-11-01 at 07:53 said:}

An excellent and thought-provoking piece. Many thanks!

\hfill

\texttt{andros on 2014-11-01 at 08:22 said:}

Yes, thank you for the tasty fare, seasoned well with nourishing links -- I look forward to finding the Duffy tome in particular. One link perplexed me though: re. `most of the perpetrators were priests the Church let in under pressure from Vatican II reforms' links to the John Jay Report. The wiki link states `Almost 70\% of the abusive priests were ordained before 1970, after attending pre-Vatican II seminaries or seminaries that had had little time to adapt to the reforms of Vatican II'. Can you resolve the contradiction please?

\hfill

\texttt{William on 2014-11-01 at 09:18 said:}

Might you suggest a link (a Gornahoor page or something else) on ``Theology of Glory''? Thanx!

\hfill

\texttt{IA on 2014-11-01 at 10:15 said:}

Very thought-provoking article. I was raised as a protestant too but certainly not progressive. I remain unconvinced that protestantism begat liberalism. Today we see status-marking Eloi egging on race and sex identity group human rights hucksters. Self-deification is also popular it seems.

These people to my eye seem more like descendents of millenarianism, associated with Irenaeus who's writings concerning millenarianism came into opposition to church doctrine established by Origen and Augustine (Norman Cohn, The Pursuit of the Millennium, Oxford University Press 1970). Church doctrine began to emphasize individual (and not collective) spiritual progress continuing into the next world, as opposed to this one. Admittedly, modernists do not maintain the original nationalist or racially collectivist mindset of imminent salvation. They hold to a kind of class and quasi racial solidarity though. They are the good whites opposed to the haters and bigots.

\hfill

\texttt{logres on 2014-11-01 at 11:46 said:}

Andros, thanks for clarifying my poorly stated point. I had read the report quite some time ago: the best way to make the argument is to note the elective affinity, or the correlation, rather than stating it the way I did. The ``crop'' coming up after the 1960s should have been treated differently, as they represented a different wave of souls, with affinities to the times they lived in. As with all such movements, their formative years were in a period of decay immediately preceding the change. The same thing happened with Barzun and the old professors at Columbia, at the university: They should have stood up to their prodigy. Instead, under pressure to treat them ``as we were treated'' or out of blindness to what had happened, or laziness, or what have you, they caved, and ushered in the revolutionaries. Even though the new era is always born under the aegis of the old, it is still true to say that the new rulers, in spirit, are always people seeking to create conditions favorable to their new inner state. So there is a correspondence between the result and the previous state.

IA, good point: studying the millenial cults would yield a lot of insight into Church history. However, the evidence for connecting Revolution and Reformation is not merely philosophical (or Catholic), but historical. Cromwell combined millenialism and Reformation. The American Revolution was a Scottish Presbyterian affair (at least the British thought so), and the French Revolution had a connection to Huegonot movement. You see atheist ex-Catholics, but most of your modernist critical theorists, philosophers, ``community organizers'', etc. tend to come from a liberal Protestant background. At least, that's my impression. But I'm not interested in tarring Protestantism, rather in protesting against its flaws.

\hfill

\texttt{andros on 2014-11-01 at 18:21 said:}

Logres, thanks for the clarification. However, the linkage between sexual abuse within the church(es), Vatican II and/or the new `crop' is still a weak argument, tending to scapegoat a particular section and exculpating the church itself.

The sexual abuse of the vulnerable by those in authority is a heinous crime. The crime is exacerbated when the abusers are sheltered by an institution which demands of its adherents a moral standard which it signally fails to uphold itself. Therefore the primary failure is of the institution as it is, not any one part (as `fall guy'), and has to do with the nature of power, self service and moral turpitude.

So let us start the true reformation by recognising the cause of our own failures, and most especially within those institutions where such failures are magnified.

Defend Christianity by all means, but let not the behemoth of churchianity lurk in silence to feed off the unwary.

\hfill

\texttt{IA on 2014-11-01 at 19:27 said:}

One thought about protestants: they were all northerners. Maybe the easygoing mediterranean culture couldn't fertilize the ice people in their own lands. If nature is equal to nurture that might explain it.

\hfill

\texttt{Logres on 2014-11-01 at 21:46 said:}

Andros, granted. I am not making an argument that the Church is not heinously culpable. What I am saying is that they are, but wound up in that position for a very specific, formal reason: namely, they caved in to ``the Times''. This was reflected in ``conservatism-for-conservatism's sake'' (Churchianity), which then provided a gateway or opening of weakness for compromise with the arrival of the barbarians within the gates. Abdication, scandal, and endless compromise has followed. My point in mentioning the report was to draw attention to the fact that the very thing most secularists beat the Catholic Church with (the Satanic abuse of children), has a connection with a formal act which they applaud out of ignorance (being ``up to date''). This shows the incoherency and chaos of using Revolutionary categories to do theology, social criticism, or organic gardening. I am not defending the RCC on the matter of what happened or how it is handled. It's the biggest reason I am not a Catholic. But we should indict or diagnose the root problem, along with the symptom. Historically, celibacy has not created this problem in the Church. That was my point, and I didn't make it clearly. The power, self-service, and moral turpitude you mention are valid points, and I suppose you could see Vatican II and the scandals as both on the level of symptoms, with a deeper root cause. However, some symptoms can cause other symptoms, and themselves be exacerbating causes, so I think it deserves exploration.

IA, the Northern peoples deserved to have their cultural heritage baptized and assume its proper place within Christendom. Eamon Duffy's book looks interesting: his thesis is that in England, the peasantry and poor folk clung resolutely to their Catholic heritage, which was all they had, having had the Celtic tradition suppressed by the synod of Whitby. A minority of nobility and the rising classes were set against this. This explains the survival and staying power of the Anglican Church, which was the compromise eventually reached. But that won't last forever either, given Revolutionary social conditions. The purpose of this blog as I understand it is to begin to do consciously what has been done (lately) unconsciously, and to uncover and begin to use the actual living roots of Tradition where they still survive, and to begin to grow from them.

\hfill

\texttt{Logres on 2014-11-01 at 21:51 said:}

No Church can survive without the heart, which is the Church of John, the esoteric lifeblood of even the body of Christ. But esotericists should have an interest in what happens to the formal body of Christ, where unity is still preferable to division. You can see that unity, although not formal, when you look at Scupoli's book on Spiritual warfare and compare it to the Philokalia, the traditions of the Irish Church, and even some of the writing of the Puritans.


\hfill

\end{sffamily}
\end{footnotesize}

