\section{La Revolution Devore Ses Enfants}

During the years leading up to 1789, Paris had increasingly become the exclusive city in France, a city-among-cities, with no peer. One of the dynamics behind the destruction of the ancien regime was Paris' desire to subordinate the legitimate and ancient diversity of the \emph{pays} (regions) of old France. Paris would become the truly capital city, rather than the body of the king, since in the medieval world, the city or seat of sovereignty, was in the actual person of the king, wherever he happened to be touring. With the Enlightenment justifying its bourgeois aspirations, and Art turning to the bizarre and degenerate, the entire city descended from famine into chaos.

We can see an analogous process at work today. America has decided to be the ``Paris'' of a global order which encompasses nothing less than the world plantation. All of the world's goods flow into American territory, just as French produce and goods flowed into the environs of the burgeoning Parisian enclaves. With a philosophy behind it to justify its aspirations (multiculturalism, tribalism, and the continuance of Revolution), the ``new American'' seems poised to plunge his country into the same chaos that engulfed France in 1789. This is because, as we all know, the Revolution must play on.

With notable differences (Paris had a lot more potential to control the local French countryside than America has to control all of the world, and Paris was more homogeneous than America is today), it is true to say that America and its satellites and imitators comprise a global arrangement increasingly aspiring to universal ``values'', which is to say, global sway. This is what the neo-liberal consensus comes to mean. When they say ``human rights'', what they mean is the ``customary rights of those over whom we hold sway, our serfs''. There are no visible aristocrats, just beautiful people making large incomes living in good style, in cosmopolitan centers of ``Westerness'' all over the world. Oswald Spengler points out how the\footnote{\url{https://journals.lib.byu.edu/spc/index.php/CCR/article/viewFile/12966/12830}} \emph{megalopoloi} come to dominate the countryside in late phases of civilization, reaching their tentacles into the land to suck wealth out of it and legitimize itself by aggrandizement \emph{vis a vis} the provinces. In the final analysis of this cycle, it is the ostentatious display of arbitrary power and wealth, at the expense of everything else, which marks the final phase. The outer shell of civilization, precisely as it grasps power, wastes away.

In something reminiscient of pyramid schemes and cultic orders, with perhaps a bit of sheer bee-like and herd-like behaviour thrown in, everyone's greatest concern is to escape the smaller town and go ``where something is happening'', which is to say, money, power, and other attractions.

This phase is operates to help America become a ``world city''. More and more people are invited; more and more countries are invaded\footnote{\url{http://isteve.blogspot.com/2012/09/invade-world-invite-world-gag-american.html}}. The entire country, from the Everglades to Gnome, Alaska, becomes a truly diverse \emph{cosmopolis}. This befits America's status as ``most advanced region'' of the planet, future home of the ``world-city'' that turns all other places into backwaters, from which to draw more staying power and ferment. True, others imitate us (Europe, most of all), yet half-heartedly. Some regions even attempt to un-imitate us: witness Putin's intransigence over the Ukraine. \textbf{And wouldn't the invasion of the Ukraine make Russia more diverse?}

What happens when 1789 comes round again, \& the countryside (this time) is prepared to resist, not just in the Vendee\footnote{\url{https://en.wikipedia.org/wiki/War_in_the_Vend\%C3\%A9e}}, but all across the planet, when China and Brazil and Portugal all say ``no'' to the unbridled spread of the end of history? Or will they? Will America be forced to fulminate more Revolution at home, in order both to project enough power, and to maintain enough solidarity to convince the other regions to cooperate? What happens when the Revolution begins to devour its own children\footnote{\url{https://en.wikipedia.org/wiki/Jacques_Mallet_du_Pan}}? My guess is that the stakes have risen, since America is becoming divided, and losing its power to project grandiosity and force. So there is a lot on the table, to lose, at this point, for the Revolution, with Russia thinking about dusting off the crown of Christ and China waffling over whether liberal democracy is really a valid replacement for a civilization that lasted over 4000 years\footnote{\url{https://en.wikipedia.org/wiki/Joseph_Needham}}.

What ought to be done to prepare, in an outward sense?

Plinio Correa de Oliveira\footnote{\url{https://gornahoor.net/?p=6893}} makes some tantalizing remarks about the Revolution and Counter-Revolution which should help the contemporary ``American''.

1. Those who are ``of the Reaction'' are, by their very nature, exposed to ridicule and a feeling of helplessness, sometimes even in their own eyes, as their ignorance contributes to a lack of realization of who they are: in this sense, the ``wholeness'' they partially lack is mimicked by competing systems of thought which may confuse their soul, providing a sense that although they can never fit, yet what they cannot fit into possesses some of the qualities they intuitively seek, such as legitimacy (in this case, illusory). It forces them to view themselves in a tragic light. This, of course, is all the better for ``the Revolution'', as it consigns their worst enemies to the outer darkness, partially self-imposed. We get to pay the upkeep on our own soul-prison.

2. Many are latent Reactionaries: that is, their upper levels of soul are dominated by the current thought paradigms, but deeper down, they are instinctively ``men of the Right''. These people simply need firm guidance, and to be shown strong moral character, coupled with principled action and depth of soul. They will be drawn, just as instinctively, to those of the first category who have stabilized themselves and emerged with the Hyperborean spirit. This second category could be termed ``Demetrian\footnote{\url{https://gornahoor.net/?p=7233}}``, instinctively drawn to those who can incarnate the power that they seek and lack. This would be a warrior class, maybe even a priest class, and in converting these classes, the natural capos or leaders would achieve a preponderance of force that would allow for an overthrow of the Revolution, and a change in the balance of power. The Revolution succeeds by hiding its aims; the Counter Revolution can only succeed \emph{by revealing its aims to this class}, but the manifestation has to be multi-dimensional, profound, and utterly authentic. Nothing less would convince the sturdy, hapless enforcers of Revolution who unwittingly serve the very forces they hate at a deeper level.

3. He compares the Revolution to the vine in the Amazon rain forests called the strangler fig\footnote{\url{https://en.wikipedia.org/wiki/Ficus_aurea}}. I have often made the mistake of thinking (and so have others, I suppose) that ``this couldn't last''. But in a startling metaphor from the
\emph{Liber Naturae}\footnote{\url{http://blog.adw.org/2010/11/the-word-of-god-is-more-than-a-book-a-reflection-from-the-post-synodal-exhortation-verbum-domini/}}, (Buch der Natur\footnote{\url{https://de.wikipedia.org/wiki/Buch_der_Natur}}), our author shows us that the goal of the strangler fig is to kill the host \emph{and establish itself as a free standing tree}. Of course, the original tree will not be allowed to live after it has served its purpose: this is why the Revolution hid and covered its aims. It wanted ``this and that'', or ``old and new'', together: tolerance for the lower castes, that's all. But the true aim is the annihilation of the original tree, which suggests that envy and hatred are, in reality, the real motivating causes of the Revolution after all, rather than compassion. The two organisms are incompatible, and only one has the essence of Life. This dissembling is a weak point in the Left, as they are now the ``authentic and legitimate order'', and our author says that firm and unwavering contradiction in a manly spirit, even against overwhelming social disapproval, is the only way of unmasking the lie -- as Veuillot\footnote{\url{https://en.wikipedia.org/wiki/Louis_Veuillot}} says, \textbf{\emph{the emptiness flees contradiction}}. Any real resistance will be met by deceived men of the number two category, and their involvement in persecution will only serve to enlighten them as to the real aims of the Revolution and the real worth of the ``rebel'' (who is the opposite of one). The Left has been winning without even trying for so long, that a real dose of suffering of any kind, at whatever level (including the wallet) would quickly make it clear that only the men of the Right have the freedom to overcome circumstances. This is quite clear right now in the crisis over Crimea: Russians are willing to sacrifice to restore the Third Rome, however imperfectly, and Americans are too lazy to even understand the issues involved, let alone drip any blood. Russia wins. America has been purging herself of the men who are of the mettle to make American force stick overseas; it remains to be seen whether she can perfectly finish the Revolution at home, as more and more of the aims come out of the closet in order to ascend the decayed altars like demons haunting the throne.

4. It is imperative to promote the association of men who are truly of the Right. Cologero has gone to great lengths to do just that, \& it will be up to us to continue this push. By providing a ``home'' and the camaraderie that all men need, noble or not, but which we alone tend to lack, given the circumstances, an important element of morale is introduced, as well as the opportunity to learn from one another, and no longer feel ``alone''. Many here have voiced how lonely it is out there, and this is true. But the hearth fire burns, despite the cold winter night. There are various ways of seeking this kind of fellowship, but it is fundamental to mental sanity during the apex of Revolution, the height of its power, just before it begins to turn upon itself. We are going to need all of our wits and resources in the years ahead, as Chaos is standing at offstage right, and the raving lunatics of the Left are congregating at stage left. After all, when Satan is on the throne, what can he really say about Revolution?

That's probably insulting to Satan. What we've got on our throne is a collective demon or elemental force, generated by centuries of passions and muddled thought. Satan is a man of Law, and wouldn't dream of having dinner with revolutionaries. He'd rather joust with a man of the Right.


\flrightit{Posted on 2014-04-04 by Logres}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Roman Bernard on 2014-04-06 at 04:45 said:}

Typo: La Révolution dévore ses enfants

(with an ``e'')

\hfill

\texttt{JA on 2014-04-06 at 13:53 said:}

One thing I must keep pointing out is Guenon's writing in ``The Reign of Quantity and the Signs of the Times'' that the present atheist-materialist epoch -- based on the absence of God -- will be replaced with a counterfeit spiritual society -- based on a false God -- which accords perfectly with the Apocalypse of St John explaining how the Antichrist will be worshiped as God... ... ... ..which is why I believe that the American consumerist global society will fall to a Russian-Eurasianist faux ``traditional'' world order that will be the reign of Antichrist... ... ..until the Grand Monarch arises from France to restore true Catholicity... ... ..

\hfill

\texttt{Pickman on 2014-04-06 at 14:37 said:}

True JA. Also consider that the angle of critique coming against the scientific-mechanistic milieu is from the new-age spiritualist clique.

Tradition must assert itself.

\hfill

\texttt{Logres on 2014-04-06 at 18:47 said:}

Why would the true monarch come from France?

\hfill

\texttt{JA on 2014-04-06 at 19:47 said:}

Logres, dear brother -- Catholic prophecy predicts the coming of a French grand monarch to save Europe and Christendom from evil, if you would like I can post some links to various writings from the saints mentioning this.

Also, using Evola -- things come full circle, the revolution began in France, it will end in France... ... ..

\hfill

\texttt{Timotheus Lutz on 2014-04-06 at 20:27 said:}

`Also, using Evola -- things come full circle, the revolution began in France, it will end in France . . .'

That seems rather short-sighted. The subversive explosion of 1789 was the result of trends that had for very long been present all over Europe.

\hfill

\texttt{Michael on 2014-04-06 at 22:27 said:}

JA, If you don't mind, please post a good source for the prophecies concerning the grand monarch. I'd like to learn more. I wonder why it would be the king of France and not the emperor.

\hfill

\texttt{Logres on 2014-04-07 at 00:34 said:}

It would be fitting for the French to atone for 1789. King Arthur, not to mention (as noted) the Emperor under the Mountain, make for fine candidates as well. Prophecy is not to be despised.

\hfill

\texttt{JA on 2014-04-07 at 08:35 said:}

\url{http://www.todayscatholicworld.com/great-catholic-monarch.htm}

\hfill

\end{sffamily}
\end{footnotesize}

