\section{Heredity and the Law of Attraction}

The metaphysical foundation of the Law of Attraction, reincarnation, race, heredity.

\paragraph{The Law of Attraction}
A lot of pop spirituality comes from deformations and misunderstandings of Traditional teachings. Among them is the so-called \textbf{Law of Attraction}, the notion that our thoughts can attract persons, things and event into our lives simply by thinking of them. Obviously, this usually fails since no one can control and concentrate his thoughts to extent necessary, and his mind is a mixture of some positive but mostly negative thoughts. When it does appear to produce desired results, it is because its true source is much deeper than thought.

\paragraph{Being and Manifestation}
As we have previously mentioned spirit is the non-formal manifestation of a being\footnote{\url{https://www.gornahoor.net/?p=1539}}. That is, it can never be the object of consciousness. Another way of putting it is that it is noumenal, not phenomenal.

Now, the human state is just one possible state of the being. The being is not fully determined, but has an indefinite number of possibilities of manifestation. The human state is one such state where those possibilities can be made manifest or actualized. Since every being has his own particular nature, destiny, and possibilities, when he manifests in the human state, he must necessarily attract those conditions which are compossible with his own nature and possibilities.

So there are two factors to be considered. The first is what originates from the being-in-itself, that is, the being as transcendent to manifestation, or the active factor. The second is what proceeds from the totality of the environmental influences arising from immanent manifestation; this is the passive factor.

We are all familiar with the notion of physiological heredity. The being is manifesting in a certain environment in a certain current of time. Hence, he will necessarily depend on genetic and other material factors that form his corporeal state, or gross manifestation.

But there is also the question of subtle manifestation, which is itself subject to heredity. So the being not only inherits a physical body, but also an astral body, by which we mean no more than the sum total of thoughts, emotions, desires, and so on that properly are part of subtle manifestation.

\paragraph{Family, Nation, Race}
The manifestation of the being as a member of the species homo sapiens is not arbitrary, but depends on factors intrinsic to the spirit. Hence, his gender, sexuality, family, nation, and race — the factors that constitute his corporeal and psychic hereditary — as not contingent, but are attracted to the being as the field where he can actualize his possibilities or life-project. With this understanding of heredity, we can see the metaphysical foundation for Evola's distinction of the races of the body, soul, and spirit.

\paragraph{Accidents and Relationships}
Our roles, identities, pasts, and so on cannot be understood as accidental, since they are manifestations of who we are. Everything that ``happens" to us in life, whether good or bad, needs to be seen from that perspective.

Similarly, all our relationships, both positive and negative, are elements of the cosmic environment that we attracted, both synchronically and diachronically.

\paragraph{The Problem of Births}
Traditional societies relied on arranged marriages and even astrological considerations to better plan births, at least for the higher castes. Nowadays, everything is allowed to occur at random, allowing the incarnation of beings that would have not been possible at other points in the cosmic cycle.

\paragraph{Reincarnation}
Psychic residues left behind by a being at his death can form part of the subtle heredity of another being. If this influence is strong enough and organized to a sufficient extent, the newly incarnated being may interpret this as evidence of reincarnation. This is so particularly since his self-identity is usually restricted to the psychic realm. This is why the Hermetist \textbf{Valentin Tomberg} can state that reincarnation is an objective fact, but not a spiritual reality. It is a fact in the sense that the astral remains of one being take root in another, though it does not constitute the spiritual realty of either.

\paragraph{Forgetfulness of Being}
True being is beyond both subtle and gross manifestation, but in the process of manifestation modern men forget their real being. Instead, their self-identify lies in the astral body, which is interpret to be the center of interiority. They falsely attribute their essential nature to it, and thus deny the possibility of psychic heredity. Or else, they admit psychic heredity but then conclude that being is totally determined by environmental factors.

So, for all practical purposes man is indeed limited to hereditary and environmental factors. In his forgetfulness of being, he is then subjected to the influences of the cosmic environment unconsciously. He experiences this negatively as a limitation to his being, whereas in reality it is simply a privation or restriction excluding certain possibilities. In sum, he is passive in respect to the environment.

Try, then, to remember. Remember who you are. Your environment, family, body, mind were willed by you at the deepest level to be in conformity with your nature and destiny. The people you are in relationship are there for a reason, because you chose them. A chance encounter, book or idea is intended for you. Be attentive and conscious at all times to their meaning or purpose. They are not a limitation on your freedom, but rather an opening to being.



\flrightit{Posted on 2011-02-06 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{James O'Meara on 2011-02-07 at 18:18 said: }

A good history of the Hermetic roots of the ``Law of Attraction" and a long list of earlier American snake oil concepts: 

Secret Source: The Law of Attraction Is One of Seven Ancient Hermetic Laws: Here Are the Other Six by Adam Parfrey

\url{http://www.amazon.com/Secret-Source-Attraction-Ancient-Hermetic/dp/1934170070/ref=sr\_1\_7?s=books\&ie=UTF8\&qid=1297120616\&sr=1-7}


\hfill

\texttt{Senko on 2012-12-11 at 19:52 said: }

Very nice post Cologero, how each being is determined wether by inheritance, ambience, state of beings or its own peculiar possibilities is one of those points that are never explained totaly explicit and in full by Guenón


\end{sffamily}\end{footnotesize}
