\section{Defending the Defensible}

We have often made the point that the analog of Spiritual Unity in the world is Harmony, not identity of opinion. Just as there are six different Orthodox schools in Hinduism, seemingly at odds with each other, there are, and should be, different schools of thought in the West, which, despite that, are all reflective of Tradition, but from different perspectives. To expect Guenon and Evola to be the same is akin to trying to describe the motion of cars on the highway according to the laws of gravity. Their apparent differences result from thier different perspectives and differing specific interests.

In our time, we often see a split between self-described followers of the one or the other, a split that would have astonished both of them. As we can see from the correspondence between Guenon and Evola\footnote{\url{http://www.gornahoor.net/?p=4489}}, there was not enmity between them, but rather a sense of a common mission. Guenon reviewed several of Evola's books, Evola had several of Guenon's books published in Italy. Guenon even published articles in a Fascist newspaper (\emph{Regime Fascista}). Rather than take sides, we feel it is better to bring out exactly what their views were. A balanced discussion is helpful to that end. However, the goal must be kept in view: to clarify the ideas, not to win an argument. Gornahoor's position is simple: \emph{Those ignorant of Guenon will never understand Evola; those who do not recognize Tradition in Evola, demonstrate their ignorance of Guenon}.

There is much ado about the minor differences between the two writers. For example, in his review of \emph{Revolt against the Modern World}, Guenon opposed Evola's view of Buddhism, and also the notion of a regal initiation. (Keep in mind that it was Guenon who eventually altered his opinion of Buddhism. And he also conceded the existence of a regal initiation in earlier ages. In Christianity the idea of Christ as both Priest and King reflects that ancient tradition, not that Evola would want support coming from that angle!) But in a 400 page book, should we not expect some errors, misjudgments, disagreements, differences of opinion? That does not alter the fundamental value, but gives us latecomers some work to do.

Guenon's overall judgment of Evola' book is this:

\begin{quotex} 
These disagreements should not prevent us from recognizing, as is right, the merit and interest of the work as a whole, and to bring it in a particular way to the attention of all those who are concerned with the ``crisis of the modern world'', and \emph{who think like us} my emphasis that the only efficacious means of rectifying it would consist in a return to the traditional spirit outside of which nothing truly constructive could be validly undertaken.
\end{quotex}

Similarly, Ananda Coomaraswamy had many specific reservations about ``Revolt'', although he published an excerpt and a review in an Indian publication (\emph{The Visva-Bharati Quarterly}, Feb-Apr 1940). Having made his objections, this is his overall judgment:

\begin{quotex}
Nonetheless, this book constitutes a remarkable presentation and exposition of \emph{traditional doctrine} my emphasis and could well serve as an introductory text for the student of anthropology and as a \emph{guide for the Indologist} my emphasis , especially for anyone interested in Hindu mythology and has not understood that, in the words of Evola, ``the passage from mythology to religion constitutes a humanistic decadence.'' The chapter, ``man and woman'' was chosen for the translation because of its clear, intransigent, and --- we can add --- tight peroration of the principles, that are reflected in the institutions and the ideals, such as that of \emph{sati}, that is often no more comprehensible and that certainly are no longer held dear, even as memories by our politicians and reformers who, ``whether by force or consensus, were induced to accept Western models.''
\end{quotex}

There is a lesson here to those who want to excommunicate Evola in the names of Guenon and Coomaraswamy.


\flrightit{Posted on 2011-03-10 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Will on 2011-03-10 at 11:12 said:}

Well said.

Another way of understanding the difference between Guenon and Evola is by means of the Buddhist distinction between Sutra and Tantra. The Sutric approach is characterized by the principle of renunciation. One renounces samsara / becoming / action because it is tinged with suffering, and one seeks nirvana / being / contemplation through a kind of process of elimination. The Tantric approach is based on the principle of transformation. One does not renounce the world, but rather works to transform one's view and one's experience of it. Since, as the Heart Sutra says, nirvana and samsara, emptiness and form, are non-dual, both approaches are valid.

As can be seen from his writings, Evola favored a Tantric approach, both explicitly and implicitly. He wrote about Indian Tantra on numerous occasions, but more importantly, advocated a uniquely Western approach which is arguably true to the principles of Tantra, but which is not a mere western aping of that tradition.

Guenon, on the other hand, was more predisposed to a Sutric approach, as evidenced by his living out his remaining years as a Muslim under Shariah law.

Aidan, in a previous comment, wrote that heroism and individualism are mutually exclusive because the former deals with individuality and the latter transcends it. That is a Sutric perspective, and is valid. But it seems to me that the question for Evola and for those who would adopt his view is, Can one transcend individuality towards divinity through something like a TOTAL FULFILLMENT of individuality?

The Sutric approach seeks to transcend individuality by negating it, thus the practice of humility, turning the other cheek, or `driving all blame into oneself' as in Mahayana Buddhism. The ideal setting for this approach is a monastery.

The Tantric approach -- which in Buddhism is considered more advanced and elitist because it presupposes an experience of emptiness -- does not necessarily follow the same rules. It is a total involvement with the world and a working with all of the energies of life. Evola elaborated his uniquely Western vision of this path in Ride the Tiger, noting that he was writing for the man who is involved in the modern world, but does not belong to it.

It's easy to spot a `heretic' in Sutra because the rules are laid out and produce a great uniformity -- if one isn't a smiling monk, one must be a heretic. However, Tantric practice produces a far more diverse display of characteristics, including what the Tibetans call the `crazy wisdom' tradition of eccentric yogis, and it's much harder to tell who does or doesn't have the goods (or rather, the Platonic Good.)

Heroism is Tantric provided that it is not simply rooted in ego, which is a very real danger, hence the elitist restrictions. While one could argue that perhaps Evola is not quite loud enough in issuing this warning, I think he does demonstrate this understanding.

Guenon called Evola ``the demon of action,'' which I take as a recognition of this complementary position that he held.

\hfill

\texttt{Will on 2011-03-10 at 15:20 said:}

Correction: the above 5th paragraph should read ``heroism and divinity.''

\hfill

\texttt{Sati Shankar on 2011-10-03 at 19:38 said:}

To prevent creeping of any confusion, it becomes necessary here to point out on,

``... .Just as there are six different Orthodox schools in Hinduism, seemingly at odds with each other, there are, and should be, different schools of thought in the West, which, despite that, are all reflective of Tradition, but from different perspectives.''

In fact in the name of tradition and for the sake of it... we some times overlook the very core of the subject matter. Difference between the six systems appear only when we approach superficially... when approached at deeper level, the underlying ``goal'' however is the same.For example,the final goal of life in Hinduism, Buddhism and Jainism is the same... Release from the repeated cycles of rebirth and liberation of the turmoil and suffering.

The spirituality, introspection, monistic idealism, intuition, respect for authority and the strong belief that the truth is to be lived, not merely known, propel a Hindu and Jain towards the goal of attaining Moksha or Mukti and a Buddhist to attain Nirvana.

\hfill

\texttt{Mihai on 2012-05-27 at 16:56 said:}

I am curious : where can I find and read the above mentioned reviews (by Guenon and Coomaraswamy). I cannot seem to locate them anywhere on the internet.

\hfill

\texttt{Cologero on 2012-05-27 at 18:18 said:}

The review by Coomaraswamy was translated and published in the Italian edition of Revolt of the Modern World. I have been unable to locate the journal through my Indian contacts. As much as I dislike doing so (due to the possibility of introducing errors), I may have to translated the review back into English. It is important because it corrects a fundamental error made by Evola on the relationship between the King and the Brahman caste, based on a misreading of a Vedic text.

The review was in The Visva-Bharati Quarterly, vol V, part IV, Feb-Apr, 1940. If you can locate the original, please provide us with a copy.

Guenon published a brief review in Le Voile d'Isis, May, 1934. I don't know if it has been included in any of the English translations of his works.

\hfill

\texttt{Sati Shankar on 2012-05-28 at 00:14 said:}

Dear Mr. Mihai,

You may find brief version of what Coomaraswamy says in his article, in first few pages( and its foundation in his whole essay) of his ``Spiritual Authority and Temporal Power of the Indian Theories of Government''.

\hfill

\texttt{Mihai on 2012-05-28 at 05:26 said:}

I thank you for your replies.

Mr Cologero: I have been unable to locate any of the above mentioned articles. Unfortunetly, Guenon's contributions are hard to come by if they haven't been published in one of the posthumous volumes.

Mr Sati Shankar: I thank you for the recommendation, I'll try to look it up.

\hfill

\texttt{Cologero on 2012-05-28 at 07:37 said:}

Guenon's excellent review of ``Spiritual Authority and Temporal Power of the Indian Theories of Government'' is included in the collection ``Studies in Hinduism''. It clears up any mischief caused by Evola's equivocation on the relationship between the King and priesthood, which has consequences far beyond the merely outward social aspect.

I'll translate the two reviews into English despite the possible errors of a double translation.

\hfill

\texttt{Mihai on 2012-05-28 at 15:59 said:}

The social aspect is quite irrelevant in our state of affairs.

All in all, Evola is a good exponent of Tradition, though he has some flaws which stem from an attachement to his particular view point and inclination.

However, I do believe he has a point regarding the king as ``Pontifex'' being beyond both sacerdotal and warrior, if he unites both of these aspects in himself and is not just the head of the temporal power.

I have yet to clarify these aspects in my mind as well.

\hfill

\texttt{Charlotte on 2012-05-28 at 16:14 said:}

Logres no comments are possible on the other thread, but I did want to say that I don't support the French Revolution and never have, it is a mechanism of the illuminati, it wasn't just a case of the peasants revolting. I also thought it relevant to point out that the Merovingian monarchy of France had already been usurped centuries before by the Franks... .factions in high places lead to unrest lower down the pyramid.

\hfill

\texttt{logres on 2012-05-28 at 17:19 said:}

Yes, the peasants supported the throne, but the ``masses'' very swiftly and joyfully succumbed to the lie that ``all are equal''. Lead by their lusts, they doomed the very France they loved, and themselves. This is what Gornahoor is saying.

\hfill

\texttt{logres on 2012-05-28 at 17:21 said:}

C, if discuss further, let's do it at MoTT. Or the Forum.

\hfill

\texttt{Sati Shankar on 2012-05-28 at 21:03 said:}

Here again a point of emphasis emerges,Who are the ``Ruler'' and the ``Priest'', the ``whole within'', the two are Not two persons, when viewed from within, the two are the same, one decides other implements, we have to look within.One comes out from the other, not the former from the later nor vice-versa,and each on covers the whole. Problem begins when two are considered to be two independents... 

\hfill

\texttt{Sati Shankar on 2012-05-28 at 21:21 said:}

... When viewed as independents, the risk runs in the name of
nationalism or as Tolstoy puts,''Patriotism in its simplest, clearest, and most undoubted meaning is for rulers nothing else but a means of realizing their ambitions and venal ends; for the governed it is a renouncing of human dignity, intelligence, and conscience, and a slavish submission to the rulers. Wherever patriotism is championed, it is preached invariably in that shape. Patriotism is slavery. LEO TOLSTOY, The Open Court, Jul. 16, 1896

\end{sffamily}
\end{footnotesize}

