\section{Eliade on Evola’s Revolt}

\begin{quotex}
I don't believe \textbf{Mircea Eliade's}\footnote{\url{https://ro.wikipedia.org/wiki/Mircea\_Eliade}} review of \textbf{Julius Evola}'s Revolt against the Modern World is available in English. It was originally published in the Romanian journal \textit{Vremea}, VIII, n. 382, Bucarest, 31 March 1935. Eliade would go on to become a world-renowned historian of religion from his position at the University of Chicago. His books are essential reading, although they are not written from the Traditional point of view, but rather intended for a general audience. Nevertheless, he corresponded with Evola and was involved with Tradition and what is now called ``far right'' politics. This should be obvious from the authors he recommends in this review. As we have pointed out, his position is what well-bred men considered sane and normal prior to the French Revolution. As Galileo insisted about the earth, ``yet it moves'', what used to be the far left is now considered sane and normal. Humanity has ``evolved'', except for what remains of well-bred men.

Eliade's review follows. We see he overrated Rosenberg, who is now forgotten, and underrated Evola, who was ahead of his time.
\end{quotex}

This Revolt against the Modern World is the title of Julius Evola's latest book. Julius Evola is an author too little known beyond the Italian border. In Germany, a short time ago, another of his books,
\emph{Heidnischer Imperialismus}, was translated, a modified and, above all, politicized, version of his work \emph{Pagan Imperialism}\footnote{\url{https://tinyurl.com/yam974ra}}. In France, I don't believe that there is anyone besides \textbf{Rene Guenon} who cites or reviews his books. Nevertheless, Evola is one of the most interesting minds of the war WW I generation. Revolt can be compared---for its historical vision and courageous interpretation---to books like Spengler's Decline of the West or Rosenberg's Myth of the 20th Century, even if it will not have by far the same echo of those two books. Everything contributes to Evola's isolation in the field of thought and modern culture: the rigour of his philosophical analysis, his critical spirit, and the courage that he has in supporting a ``traditional'' science that is opposed to fragmented, atomized, secular science.

Evola is ignored by specialists because he goes beyond their mental schemes. He is inaccessible to dilettantes, because he has recourse to a truly prodigious erudition, and at the same time does not make any concession to the presentation of his ideas (it is a way of saying why Evola has no ideas that are his ``own'').

He made his debut with very serious philosophical studies on Idealism. On the other hand, his philosophical formation was recognized even by specialists, and his fundamental work --- Theory and Phenomenology of the Absolute Individual --- is the most serious book on \textbf{Magical Idealism} that had been written up until now. Moving from philosophy to culture and social criticism, Evola always gave proof of a sense of direction and of a unity of a typically Italian attitude. Evola's view is simple: in this view, that no ideology has adopted, he affirms and reaffirms ``traditional'' values. With this term, nevertheless, he means every value created by a civilization that does not make life an end in itself, but considers that human existence is uniquely a means to reach a spiritual, transcendent reality. Evola never exposed more completely and with more fervor what thing this ``traditionalism'' means as much as in Revolt. We recommend this book to all those who wish, if not to go back to old questions, at least to contemplate an explanation of the world and history of a fascinating greatness.

Similar works can be read in very numerous ways: by people ready to accept everything at their own risk and danger, or to refute everything in the same way; but also by persons ready to receive suggestions from wherever they come from, and who are enthusiastic in being able to verify them in any circumstances.

We can say that Evola's work is situated in the cultural lines of Gobineau, Chamberlain, Spengler, Rosenberg. With much more ``seriousness'', however---as paradoxical that this affirmation can seem --- Evola has not forgotten and has not renounced his earlier mentioned studies: he is more ``philosopher'', but rather in a totally unusual way. Magical Idealism brought to its extreme limits is a meal difficult to digest. It is not part of our intentions to bring up this book, that is both anti-Christian and anti-political at the same time, as well as opposed to communists and fascists, that denies a great quantity of living things and values that today crush us, affirming only ``dead'' things and values which no one in our times believes in any longer. Evola does not come under any influence. This is what makes him sympathetic. As to the resonance of his thought in Romania it must be remembered that I published a long article on his work at the end of 1927, while in 1928 I developed a study on his magical philosophy, which remains a manuscript this is apparently lost . Later the only one that he mentioned by name, in our country, was, in 1933, our colleague Vasile Lovinescu\footnote{\url{https://ro.wikipedia.org/wiki/Vasile\_Lovinescu}}.


\flrightit{Posted on 2012-05-30 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{HOO on 2012-05-30 at 22:52 said:}

We do appreciate the effort of Gornahoor \&c. of providing all this material and a number of interesting points.

\hfill

\texttt{Cologero on 2012-05-30 at 23:23 said:}

Hoo, the points of contention I bring up are not those I am obsessed with, but rather those that are in dispute among the best thinkers of Tradition. I doubt such disputes (e.g., regal vs sacerdotal) will be resolved here, but it is necessary to make the distinctions clear. Some men will take one side or the other; some will remain undecided. The point, for now, is moot, since there is no imminent need to take a fixed stand.

\hfill

\texttt{Mihai on 2012-05-31 at 05:26 said:}

Cologero, where did you find this review translated?

I am Romanian myself so I am curious who had access to this (quite rare) stuff.

\hfill

\texttt{Cologero on 2012-05-31 at 05:33 said:}

It can be found in the last Italian edition of Revolt. Obviously, I translated it from the Italian translation of Eliade's original, written in Romanian.

\hfill

\texttt{Avery on 2012-06-01 at 00:19 said:}

Myth of the 20th Century can be found here:

\url{http://www.gnosticliberationfront.com/myth\_of\_the\_20th\_century.htm}

The author is well-read and well-bred, and he is able to command in the language of the era rather than being commanded by it. Definitely the book will offer a bit of insight for anyone who wants to read it. But unlike Spengler, Guenon, and Evola, Rosenberg seems stuck in the metaphysics of his era and fails to transcend it. Here's an example of one of his more dishonest moments:

``However, their numbers were conveniently diminished by the usual expedient of being burnt by the officers of the religion of love. A single Inquisitor persecuted the Waldensian community. For thirteen years, he was repeatedly successful in catching many of them who had made some heretical remark. Prisoners were then tortured on the rack, had their hands cut off, were strangled or burned alive. In spite of this, the archbishop of Embrum had to report to the pope that the Waldensians remained obdurate in their faith.''

This is a sort Christopher Hitchens-style anti-church spiel, basically hating on believers for being too sure of themselves. Rosenberg lacks Nietzsche's ability to stand back from a fight which always involves at least two sides. You would not see this passage in any of the three writers I listed above, and probably Eliade, having read Spengler and Evola, never wrote anything like it either.

\hfill

\texttt{Avery on 2012-06-02 at 11:08 said:}

Interesting quote about this review from ``The International Eliade'' (SUNY Press, 2007):

``As Paola Pisi points out, the review does not correspond exactly to the content of the book: Eliade writes that Evola is `anti-Christian and antipolitical' ... moreover, he misunderstands--as we will see later--the Evolian conception of `Tradition'.''

\hfill

\texttt{Revd Frank Julian Gelli on 2012-06-07 at 13:31 said:}

Not a very good review, considering that it came from an excellent scholar like Eliade. Plenty of eulogies but he is vague and fails to give the real kernel of Evola's argument in `Revolt'. Maybe when he wrote it he was nodding? However, I agree with Avery on Rosenberg's `Myth' -- both the book and its author are best forgotten.

\end{sffamily}
\end{footnotesize}

