\section{The Garden of Metaphysics}

If I were to tell you about my garden, I would mention the two banana trees, one with a small hand just starting. There are the two tomato vines, and pots for basil and Italian parsley. I could mention the several palm trees, the bromeliads, and the bird of paradise just beginning to flower. There are several other flowering plants as well as ground cover.

\paragraph{Seeing Principles}
The one thing I would not do is to debate the issue. I would not try to prove the existence of the bananas nor give an intellectual justification for why I believe in their existence. If you have any doubts, I would simply ask you to visit and see for yourself.

Discovering metaphysical principles uses a similar process. The garden is part of the physical world and is experienced directly through sensory intuition. Metaphysics is about what is beyond the physical and is experienced directly through an intellectual intuition. There is nothing to debate nor is there a need to convince someone, at least as long as we are in the metaphysical realm. Nevertheless, some men may see things more deeply or are more careful observers; this is the source of disputes.

For example, when we talk about the soul and its layers or envelops, that is not a philosophical theory among others, and even less so, is it a problem for psychology. The proper approach is to begin a process of self-observation of one's inner states. In this way, the forces of the parts of the soul —such as eros, thymos, desire, attraction, and so on— can be carefully observed over time. A man can learn to align these forces in their proper order, that is, under the domination of the intellect. He will observe the resistance of the lower forces to such domination. This is his experience of cosmic law, the forces of chaos, and the necessity of participating in the greater battle of logos over chaos.

Those men who have never troubled themselves to such a practice will most likely think this all to be nonsense. They will deny the necessity of the greater battle and will typically embrace the lower forces as somehow “life affirming”. We can imagine a Viking of a thousand years ago, who never left the North land and how incredulous he would be about my description of a banana.

With his self-knowledge, our metaphysician is in a position to understand other men. He may note that there are some fundamental differences in the way men relate to their possibilities; these he categorizes in what are commonly called castes. These are not psychological states and cannot be verified by empirical methods.

He notes that, unlike beasts and inanimate objects, men do not follow the cosmic law, or dharma, by necessity, but freely. Hence, the greater battle interiorly becomes the lesser battle exteriorly. Since men of certain casts of mind do not understand this, they require a spiritual authority as a teacher and maintainer of traditions. The temporal power maintains structures within bounds, offering protection from chaos and formlessness whether arising from without or within.

\paragraph{Applying Principles}
These principles are fundamental to a Traditional world view and are the keys to understand events in the world process, both synchronically and diachronically. While the principles themselves are beyond debate, their application to a given situation is open to discussion. Here is the general approach to take.



Note who or what is the spiritual authority at any moment, or how it has changed over time. The spiritual authority is whatever cannot be publicly doubted or challenged. This is the legitimate task of a lawful spiritual authority, whose job is to maintain spiritual unity of the whole. However, an illegitimate spiritual authority will be enforcing palpabe untruths for the benefit of a subgroup. As \textbf{Guido de Giorgio} pointed out, the true light of God is faceted into a multiplicity through a hierarchy of spiritual beings. Each nation is under the guardianship of certain beings. However, in a state of disorder, an alien spirit may be (mis)guiding a nation. This requires keen discernment. Note that by nation, here, we mean fundamentally a group that is spiritually united before any other considerations, as we have described in the several posts on the Ancient City. These nations can be related vertically in a hierarchal way without destroying their individual characteristics.

Another point to note is who has the temporal power. In a Traditionally organized society, there is a caste devoted to that task, identified through their loyalty, fidelity, courage, and virility. Opposed to that, power is held by lower castes, such as the bourgeoisie or money powers, or else by the proletarian mentality. The latter is easily seen in terms of popular culture, styles of dress, values, religious beliefs, and so on. These can be compared to the greatest exemplars of art, poetry, traditional science, and so on. The complete degradation occurs when outcastes gain some measure of power. In a Traditional society, outcastes would have had no role at all. \textbf{Fustel de Coulanges} documents the role this played in the deterioration of the Greek Ancient City as well as in Rome.

Readers who find themselves interested in and of a mind to participate in contemporary political and cultural movements, have the conceptual tools from Tradition to do so. In looking over the landscape of self-identified conservatives or rightists of whatever stripe, learn to apply these principles. Discern how they view spiritual authority, do they even recognize higher principles, do they understand them. Ask how they view temporal power. Who will hold it? Are these movements composed of men with the right moral fiber? Are they manly in essence or just in outward appearance? Before supporting any such group morally, intellectually, or financially, learn to judge truly. Don't be deceived by a superficial agreement of goals that hide the lack of a common mind. Be wise as serpents.



\flrightit{Posted on 2012-04-01 by Cologero }
