\section{Confronting World Traditions}

Curiously, the anti-modern movement founded on the works of \textbf{Rene Guenon} is itself the product of modernity. Three historical trends converged. First, the West had lost, or was losing, its grounding in its own Tradition, partially for reasons we have recently discussed. Second, there was the confrontation with world culture (from the perspective of the West) that challenged the notion that the West had some monopoly on all verities. Then there was the wild success of the scientific method which, along with the idea of the rights of man, led to project of universalism based on science and liberalism.

Guenon threw his grenade in the midst of those assumptions. Rather than seeing the modern world as on the tip of an incessant progression, he insisted that it is in ``crisis", under the ``reign of quantity". He challenged the idea that the world religions were incoherent and incompatible with each other. Finally, he pointed out that the West had lost something significant in its own self-understanding, even while paying lip service to externalities of its own spiritual traditions.

\paragraph{Grasping other Traditions}
Guenon's ideas arrived at a time when the usual ways of understanding world religions by those in the West were no longer credible. The simplest explanation is that all religions are mere fantasies, although clearly this reductionist approach does not do justice to such complex phenomena. The next is to regard non-Christian religions as ``Satanic", and this is widely held by Evangelical Protestants. This can be taken to extremes as this brief story shows. In South Florida, most nail salons are staffed by Vietnamese for some reason. A woman I knew who attended a megachurch in Ft Lauderdale refused to frequent one such salon because there was a statue of the Buddha displayed in the window.

The opposite view is typical of most New Age groups. For them, there are world teachers, pretty much interchangeable, who brought spiritual teachings at various times. There is little to distinguish them, or relate them to each other. Hence, they are more like those various groups of comic book superheroes.

A close cousin of this view is that a man should choose his tradition as though selecting fruits at the farmer's market. He will choose one based on personal predilections or perhaps social status (some religions are ``cooler" than others in his circles). When this view is combined with the notion that the Western tradition is fundamentally and unavoidedly corrupt and criminal, this will be combined with disdain and derision for the dominant tradition of the West for the past 1500 years.

\paragraph{Natural and Supernatural Man}
A more sophisticated view, adopted by ancient and medieval Christians, is to distinguish between natural religions and the (one) supernatural religion. In this view, they could admit that yogis, Buddhists, and so on, have some sort of spiritual experience that is merely natural, while reserving the highest ``supernatural" experiences to themselves. Given all the material that Gornahoor has made available, this hardly seems plausible. Nevertheless, it may be useful to review the reasons why.

By ``natural", we mean the corporeal and psychic (or soulish) aspects of man, which can be traced to the natural processes (``horizontal") of the manifest world. By supernatural, we mean that part that transcends the manifest world, and hence cannot be explained in scientific terms; this comes from above (``vertical"). Now, it is one thing to make an epistemic distinction between nature and supernature, but quite another to make an ontologic distinction. When Little Miss Muffet was eating her curds and whey, she was no longer drinking milk.

So a man who is only natural, is no longer a man, but is rather an animal, some sort of hominid creature. And I don't mean that in the sense the he is a brute, since animals can also be very appealing, like your dog. But as we recently pointed out, Fr Alois makes a case that man also is a ``spirit-soul", i.e., there is a part of the soul that transcends the natural part of man. In most cases, the spirit-soul is virtual, as man in his current state is unaware of it and it is not an active part of his life.

Nevertheless, and this is important since Thomists today seem to have forgotten this teaching, it is possible to remember one's spirit-soul. In the teaching mentioned above, the supernatural experience is only the result of God's direct action, e.g., in infused contemplation, and, as it was claimed, the pagans had no possibility of that. Fr Alois documented the falsity of that idea, since man qua man has a spirit-soul and can actualize it in various ways, even to the point of developing unusual powers or siddhis.

\paragraph{Transcendental Unity of Religions}
So, at this point, the depth of spiritual insight and metaphysical understanding of the so-called pagans cannot be denied. Of course, there has always been the notion of the ``perennial philosophy", or a common core of wisdom and spirituality across space and time. There have been many attempts to document and catalog the perennial philosophy, culminating in the massive thousand page \emph{Treasury of Traditional Wisdom} compiled by \textbf{Whitall Perry}. One can be convinced by its gravitas alone without reading a word of it. However, there is a fault in this approach, a glaring fault which, when fully grasped, is also a deep insight.

The fault is the temptation to regard it as collection of nice quotable tidbits that can be recalled at parties, i.e., it is merely intellectual and not substantially different from the superhero mob or secret brotherhood of adepts mentioned above. To overcome that, the question should arise of achieving such mystical experiences on one's own, and not just repeating them secondhand. The insight comes when you notice that all the figures quoted followed some tradition or other: they were Christians, Sufis, Hermetists, Buddhists, Taoists, yogis, etc. They were not isolated men coming to spiritual realization alone and apart from any predecessor or spiritual current.

So here Guenon arrives and throws the lifesaver. Yes, there is more to it than just a collection of quotes. Actually, the perennial philosophy is quite developed, with a deep metaphysical teaching that is expressed in slightly different ways in the major traditions. That is part 1. Part 2 is also a ``yes": to achieve the understanding of those teachings, it is necessary to actively participate in one of the traditions that he regards as authentic.

Then you find out that the lifesaver has a leak and won't keep you afloat long enough. The tradition must not only be authentic, it must have maintained an unbroken chain of esoteric teachings to be effective, and in the West, that chain has been broken. So all the saints quoted in the \emph{Treasury of Traditional Wisdom} are to no avail, since their states are no longer accessible to us. Although a few men under Guenon's influence followed him into Islam, there has been no mad rush to do so. Evola is not much help since he rejected the spiritual authority that created the traditional society that existed up until the French revolution; this is his claim, not mine. Yet he also rejected any attempts at a neo-pagan revival. As those who have read his correspondence with Guenon that was published on Gornahoor can see, Evola sought for years, usually among marginal figures in Europe.

So there are some things that make us uneasy. While we don't doubt the metaphysical teachings, it is legitimate to question Guenon's judgment of contingent and historical events. First of all, the transcendent unity of religions sounds too similar to Noam Chomsky's universal grammar. Initially it seems edifying. Unfortunately, by explaining everything, it explains nothing. Neither the ur-tradition nor the universal grammar exists anywhere; they can't even be conceptualized. So, if the traditions did not fall from the sky, they need to be related to each other as part of the world historical process. And if so, they are no longer separate but equal; the traditionalists reject egalitarianism, so we should not suddenly adopt that view in regards to the traditions.

\paragraph{Forgetting, Sleeping, and Death}
Now Guenon's introduction to spirituality was through the French Hermetic tradition, as he was associated with \textbf{Papus}. Although he broke with that movement, he continued to quote from Hermetic writers in his subsequent books. (Do not forget that Evola also wrote a book, and a good book, on Hermetism.) So we are correct to consider it the esoteric tradition in the West. The remaining question, then, is whether or not it is still active, with an unbroken chain. Now one of the reasons Guenon gave for his departure was the Hermetic teaching on reincarnation, which he claims is not traditional. Well, by that standard, contemporary versions of Hinduism and Buddhism are also not traditional, since I don't know of any major teachers in those traditions who do not accept reincarnation as a fact. That does not mean much else.

Guenon, in his letters, did allude to a Hermetic group in Europe with just 12 members; he could not penetrate into it. So just as one black swan disproves the theory that all swans are white, this one group disproves the assertion that the chain of esoterism has been broken in the West. A group like that would be restrictive in order to maintain its purity and efficacy, however, not so restrictive that it would allow itself to die out.

Another event that occurred after Guenon's death is the appearance of writings that do indeed indicate a deep spiritual gnosis; here I have in mind \textbf{Valentin Tomberg} in particular. In brief, we can restate how he sees the traditions as related. Ponder it, and you may see that it is true, even if I have to leave many details out.

He points out that the knowledge of monotheism was an accomplishment, or an end point of spiritual realization, among the pagans such as Plato or Hermes Trismegistus. In the Christian tradition that followed, it is instead a revelation, a dogma, an article of faith (``I believe in one God"), the starting point. For \textbf{Plato}, to know is to remember, so he shows us how to overcome forgetting. This is not unrelated to \textbf{Pope St Gregory}'s teaching about the prophecy of the past\footnote{\url{https://www.meditationsonthetarot.com/the-future-of-prophecy}}.

The \textbf{Buddha} overcame sleep and teaches us the path to do so. Most men are sleepwalking through life and not seeing life as it is. In a fallen world, Buddha saw clearly old age, sickness, and death, things we want do hide from through our distractions and amusements. So the Buddha overcomes sleep.

Christ, the Risen One, overcomes death for us. So, we see such a different, and enlightening, approach. These spiritual teachings are related to each other in a hierarchical way that does justice to them. Plato, the Wise Man, Buddha, the awakened one, and the Christ. That is our path. So to be a Christian today, that is, one connected to the esoteric Tradition, is to follow that same path. So where Plato stopped, that is where we begin.

\emph{We must first become Wise like Plato, then Awake like Buddha, and finally overcome death like Christ}.



\flrightit{Posted on 2013-10-03 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{David on 2013-10-04 at 00:45 said: }

Interesting take. As far as Guenon goes, it is also of note that some of his close contributors and "friends" made progress within freemasonry and that some "traditional" group exist within that order.


\hfill

\texttt{scardanelli on 2013-10-04 at 09:56 said: }

Perhaps he has been discussed here before and I missed it, but I've just begun reading Karl von Eckartshausen's Cloud Upon the Sanctuary, and his description of the Inner Church bears many similarities to Tomberg: that it has existed since the fall of man, that it's members recognize each other through their depth, that this inner church is the truth of the symbolic outer church, and that Christ is the only master. It seems that where most go wrong is searching for the western esoteric tradition externally in some group or order.


\hfill

\texttt{Juan Senko on 2013-10-04 at 15:14 said: }

@Cologero: Excellent post. I hold a very similar opinion. I believe, along with Hani and Reyor that the Holy Sacraments fulfill the role of Initiation Rites. thus, the Western Tradition as far as Catholicism goes is not broken at all. I don't believe this out of intellectual conviction or Faith alone but because so first hand experiences. The Hermetic Tradition would be a parallel current running throught it. So, You have the Church, you have Hermeticism and also Freemasonry, Of course the Western Tradition is not dead! It's very much ALIVE!

I don't know if the group you speak about is the same one and we have different information, but ChaboneauLassay was the head of an Order restricted to 15, something along the lines of ``Order of the Holy Ghost".


\hfill

\texttt{uan Senko on 2013-10-04 at 15:14 said: }

@David: Guenón himself was involved in some Masonic groups that tried to restore the old rituals and landmarks and become operative instead of speculative. Also, you always have the Rectified Scottish Rite.

Oh! and Martinism. Though it is pretty much distorted and full of AMORC type of people Martinism is still a valid initiatory path!

Blessings in Christ and Mary to you all!


\hfill

\texttt{Avery on 2013-10-05 at 15:03 said: }

Guénon located the concept of perennial philosophy in the Indian classics. We know that a similar concept developed in Europe, but we do not hear about it until a much later date, because when India was developing those ideas around A.D. 200, Europe was declaring ``extra ecclesiam nulla salus" and heresy hunting.

Recently I went to church for the first time in 13 years. I liked almost everything I heard but I could not get that problem out of my head. Schuon tells us that ``extra ecclesiam nulla salus" is a false doctrine. I don't believe in his construct but he raises a good point.


\hfill

\texttt{Iñaki Luna on 2013-10-05 at 19:11 said: }

I insist on the fact that René GUENON did search (and research) the remainings of the Great Tradition among several groups in Europe (including Anthroposophy etc. ) but he was never satisfied .

Thus , he HAD to go (on) EASTWARDS … (as well as SCHÖNBERG HAD to create dodecaphonism , for instance …) 

I' ll try to explain this further on later , ok ?


\hfill

\texttt{Cologero on 2013-10-06 at 09:04 said: }

Catching up on comments:

@Iñaki Luna:

Yes, we agree that M. Guenon could not find what he was looking for in Europe, even though he was initiated there by an itinerant painter. So what is your point? That we all relocate to Egypt? Are we sure tradition is alive in the East?

\textbf{I refute you thus}:

The Spirit breatheth where he will; and thou hearest his voice, but thou knowest not whence he cometh, and whither he goeth: so is every one that is born of the Spirit.

So the Spirit is not obliged to be where M. Guenon is looking. True initiation comes in the night: 

@Avery

You bring up a topic that was not mentioned in the post. Nevertheless,

\textbf{I refute you thus}:

\textbf{His True Love}: ``\emph{Avery, my darling, do you love me? Do I make you happy}?"

\textbf{Avery}: ``\emph{Yes, dear, I love you and you make me happy. Nevertheless, there are several other women I've met who could have made me equally happy}."

@Juan Senko

Thank you for being on topic. Yes, I agree with your point and it must certainly be taken seriously. I have made a case for it here. Perhaps Chaboneau Lassay was what Guenon was alluding to. Where there is smoke, there is fire.

@Michel

The only actions of eternal value are the corporal and spiritual works of mercy. That is not something one boasts about, although what I am doing is obvious.

@scardanelli

Good catch about Eckartshausen. Yes, he was influential in Martinism; Tomberg builds on that tradition and is not as radically novel as some seem to believe. As for the rest of your comment, that is the upcoming topic this week for the medtarot group.


\hfill

\texttt{Cassiodorus on 2013-10-07 at 13:32 said: }

This post was excellent Cologero! 

The section on natural and supernatural man gets at the heart of the tension between Christianity and Perennialism. Here's the thing, it seems to me that to reconcile the TUOR with the Abrahamic faiths, one has to regard the historical facts of Revelation in relative and consequently, non-essential terms. If it is not true that ``the supernatural experience is only the result of God's direct action", what exactly is Revelation? 

On the TUOR view, is an Orthodox religion a ``saving mirage", a map that shows the ``path to the summit" and provides ``combustible material" to ignite the slumbering intellect? But, how can one square this off with the central claims of traditional Christianity- that puts supreme importance on a set of ``temporal saving facts"?


\hfill

\texttt{Logres on 2013-10-07 at 23:47 said: }

What if Christianity is a perfection of myth and religion in the highest possible sense? And also the lowest? Christ as both true myth and re-establisher of all lost religions. Would not such a hero, if he existed, be the supreme servant, to reunite to mankind both myth and religion, in his historical trajectory? This seems to have been the view of the antique Europeans, or at least, their intuitive sense. Such a ``fact" would go a long way towards explaining the tensions in the debate.


\hfill

\texttt{Michael on 2013-10-08 at 20:10 said: }

Do you think that Guenon was correct about the necessity of having an intellectual elite that would ease the transition to a traditional society? I think Tomberg has a similar concept in MoT: the difference between the crew of a ship and passengers. 

And is the idea of moving to a traditional society similar to the middle ages still a desirable goal?


\hfill

\texttt{Scardanelli on 2013-10-08 at 21:15 said: }

I think it's a little too easy to misunderstand the phrase ``intellectual elite." We won't be be delivered by journal articles or conferences. I believe what tomberg refers to could more rightly be called the interior church which acts as a ferment or arcanum for the exterior symbolic church and for exoteric culture in general. I think intellectuals of our day, even if they ``lean" towards traditionalism might find it difficult to gain entrance to the interior church. As Eckartshausen says, ``We live in the times of the idolatry of the intellect, we place a common torchlight upon the altar…"


\hfill

\texttt{Cassiodorus on 2013-10-09 at 14:42 said: }

Logres,

I'm inclined to take the view that the Incarnation, when interpreted Traditionally, must be understood as an an absolutely unique event- that is, Jesus Christ is ``ground zero" of the Divine descent. Just as there is hierarchy within the Divine Principle (Nirguna/Saguna Brahman), perhaps the same can be said of the Revelations themselves.


\hfill


\hfill


\end{sffamily}\end{footnotesize}
