\section{The Seed of the Future World}

Recently someone, claiming to be a Hindu priest despite looking European, informed me vaguely that Westerners misunderstand the \emph{Sanatana Dharma}. Since my understanding is heavily colored by the writings of \textbf{Rene Guenon}, we shall rely on his essay on that topic included in the collection \emph{Studies in Hinduism}. We shall use the traditional method in debate of using sentences, so it will be possible to refute, rectify, enhance, or embellish any one of them. We do realize that there is a ``politically correct'' movement afoot to use the term Sanatana Dharma as a synonym or replacement for ``Hinduism'', but since the Hindus themselves cannot agree on that, it merits no further discussion.

On the surface, Guenon agrees with the claim in question since he asserts that the West has no exact equivalent. Although \textbf{Ananda Coom\-aras\-wamy} says that its best approximation is the \emph{Philosophia Perennis} of the Middle Ages, Guenon sets out to show there are some significant differences. However, this is pointless in itself, because our real aim is to use the Sanatana Dharma as the key to bring out the full meaning and implications of the Philosophia Perennis.

In his essay, \emph{The Spirit of India}, Guenon again asserts that true wisdom remains solely in the East. He describes the end of the Kali Yuga and the proper attitude of the new elite:

\begin{quotex}
The world of disorder and error will be destroyed, and by the purifying and regenerative power of Agni all things will be reestablished and restored in the wholeness of their original state, the end of the present cycle being also the beginning of the future cycle. Those who know that it must be so cannot, even at the heart of the worst confusion, lose their immutable serenity. However irksome it be to live in an epoch of trouble and almost general obscurity, they cannot be affected by it deep in themselves, and it is here that we find the strength of the true elite. Undoubtedly, if the darkness should continue to spread more and more, this elite could, even in the East, become reduced to a very small number. \emph{But it is enough that some preserve integrally the true knowledge to be ready, when the ages are completed, to save all that can still be saved from the present world and become the seed of the future world.}
\end{quotex}


There is a little paradox here that Guenon ignores. If the West is furthest along the Kali Yuga, then it will be the first to commence the next cycle, i.e., the West is the ``seed of the future world''. Since the owl of Minerva flies at dusk, the West may well be in possession of the Eternal Wisdom, even if virtually, and Guenon has his own prejudices. Hence, we will have to revisit some ideas of \textbf{Julius Evola}, particularly those that put him at odds with Guenon.

Now, Guenon had the opinion that the West would have to count on aid from the East to restore its traditional spirit. This aid would need a ``point of support in the Western world''; however, he admitted that these possibilities are ``presently difficult to define.'' Perhaps 80 years later we can define them. There have been a series of Indian gurus who have embraced the West, starting with \textbf{Vivekenanda}, then the Self-Realization Fellowship, Transcendental Meditation, and so on. Guenon did not much approve of Vivekenanda, and probably not the others. More recently we know that Sting practices Tantric sex and many women are doing Hatha Yoga exercises. I doubt that they will be restoring the traditional spirit. On the contrary, India is now exporting engineers and managers, the graduates of their prestigious \textbf{Indian Institutes of Technology}, who are all too willing to leave the traditional spirit behind if they are able.

The other choice is Islam, since it has become a daily presence in much of Europe. Although it may be preserving the eternal wisdom, circumcision and the prohibition of alcohol and sausage are obstacles to Europeans. Nevertheless, Islam shows in a vivid way how a traditional society is organized. In Islam, there is no distinct secular and religious spheres since every aspect of daily life is regulated by sharia law. This includes the proper way to eat and pray, down to how to urinate, even how to clean one's private areas. Throughout the day, the Muslim needs to avoid unclean things and perform duties and tasks correctly. That is not so for the contemporary westerner, who is quite pleased with unclean things. It wasn't always the case, since there was a time when a European man sought to avoid temptations and occasions of sin while doing good deeds. The main difference was that the European was not as concerned about corporeal laws, since he tended to spiritualize ideas about cleanliness.

So it seems that Guenon was wrong about the East rushing to set the West spiritually aright, at least willingly. Hence, it behooves us to start defining the possibilities that Guenon was unable to do. One of those possibilities is to use the Sanatana Dharma to clarify aspects of the Philosophia Perennis that have become incomprehensible to modern man. After that, we shall reexamine the differences in spirituality and mentality between East and West as Guenon understands them. Finally, we will turn to Julius Evola and \textbf{Valentin Tomberg} to make that distinction even sharper. Since we here do not suffer from an inferiority complex in regards to the West, nor do we feel alienated from the most recent spiritual tradition of the West, we shall make the case that the Western view is correct and is the seed of the future world despite its current state of disorder. We are convinced that once you think about it a bit, you will also see that it is true. However, that will need to wait until tomorrow.


\flrightit{Posted on 2013-10-29 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{kenneth shaw on 2013-10-30 at 09:23 said:}

Please continue. This is subject is very helpful to me as I try to map out a course for my religoous life.

\hfill

\texttt{Constantine Aetos on 2013-10-30 at 16:27 said:}

Well you were right in your welcome post that ``Others may have an agenda apart from the group''.

Anyway, I would like to add that there is an article by Arthur Versluis titled ``The Political Significances of Western Gurus'' Esotericism, Religion, and Politics (2012). A very interesting article of westerner gurus steeped in eastern thought trying to ``correct'' the west.

\hfill

\texttt{Max on 2013-10-30 at 18:06 said:}

Thank you for this post, I agree fully and have been thinking along the same lines recently. Guenon wrote for the west, that might point toward him also sharing the same ideas brought up here. He probably wrote because he thought he could make a difference -by contribution to the understanding of the west by studying eastern traditions that entered modernity later and therefore was more comprehensible. I have observed that there is a gradual sort of merging of traditions which has interesting implications. It has progressed to the apparent goal of making a kind of synthesis of all traditions that existed, the shepherd gathering his scattered flock. Moving from potentiality of the world in the beginning to fullest realization. Think about it; Christianity inherited from keltic, nordic, roman, judaic, egyptian, greek, and countless others we might not even be aware of just as those traditions themselves probably were syntheses of others that are unknown today. The romans for example incoroporated conquered and subjugated peoples religious views in effect allowing their aspect of the whole to live on and be transmitted. If you look at hinduism it is the same story, and now they contribute to understanding of our inheritance. Those developments are seen in the light of ``the product of the manifestation'', a new seed. The last times are where a complete understanding of all the facets of the manifestation can be reached, a unique opportunity for those living in that moment. Soon we will be dealing only with reality and nothing else left, the highest vantage point at the top of the mountain where we can pass across the rainbow.

\hfill

\texttt{Ash on 2013-10-30 at 18:26 said:}

If we take the ``east'' as including Russia and the eastern Christian countries, would Orthodoxy not present itself as a third solution? Its greater cultural similarity to Western norms and increasing openness to taking in Western traditions, such as examining the forms of worship which were subsumed in the Latin expansion (for better or worse) seem to have made it a successful path for many Westerners to walk, such as James Cutsinger and Fr. Seraphim Rose.

You mentioned Alan Watts a previous post. Do you think that he represented an honest attempt to try and use eastern wisdom to critique the Western decline (which is to say, to some extent trying to grasp the forgotten Traditional principles) or that he was confusing exoteric and esoteric aspects of the east? ``Does It Matter'' appears to make a connection between principles and everyday life, down to clothing and cooking. In that case, the proper order of the higher guiding the lower has been recognized, unlike in yoga classes where fitness and relaxation seem to be the end goals, if not utter New Age nonsense.

\hfill

\texttt{Ken Shaw on 2013-10-30 at 18:52 said:}

In regards to Orthodox Christianity, I've seen Seraphim Rose spoken well of in Perennialist circles

\hfill

\texttt{David on 2013-10-30 at 19:53 said:}

@Ash : I live myself in a Roman Catholic environnement, and I would very like to know as to why Orthodoxy would not be considered Western Religion, in it's ''pure'' form; that is, Medieval Christianity. In other words, Orthodoxy may be more traditionnal, just as Medieval Roman Catholic were (in some regards). They sprang from the same roots, even if language does changes alot and Orthodoxy became more greek than their Western counterpart : can we really say they are alien from western tradition ?

\hfill

\texttt{Ash on 2013-10-30 at 22:02 said:}

David, insofar as both Orthodoxy and Catholicism are manifestations of the true Christian Tradition, I would say they are not alien to each other. Is this the sense in which you consider medieval Catholicism the ``pure'' form of Orthodoxy? This aside, I'm not sure how that argument can be made, as what we now know as Orthodoxy came through Byzantium, which itself was born in the Greek Christian tradition of the Eastern Empire, and is not an offshoot of the Latin Christian tradition, although both have the same root. When I say it is alien to the West, I simply mean that its ``language'' (apotheosis vs salvation by grace, etc) and history differ from the Western experience, not its doctrines. That said, if they were completely akin to each other, we would not still have the tragic schism existing in the Christian world. As a Catholic with Tradition in mind, would you say that you could come into the Orthodox community and consider it as Western as your own tradition? As has often been pointed out on this blog (see the white Hindu in this post), purity of principle in a Traditional stream doesn't always mean that it is suitable for each seeker (although I don't believe this to be the case with Orthodoxy -- if and when I come to an exoteric path myself, it could well be the Orthodox faith).

\hfill

\texttt{JA on 2013-10-31 at 09:42 said:}

after having once been very pro-Orthodox I came to see that it is not suitable for a westerner, seeing as the Orthodox church is very anti-west historically (opposing Charlemagne, the Crusades, they see the Holy Roman Empire as a usurpation, what lost Jerusalem for us was the schismatics making deals with the Mohammedans against the Christian army, etc etc) and also from the dogmatic point of view Orthodoxy is very factional, each local church is frequently at odds with other ethnic and national churches, is that the universal church desired by Our Lord ? no -- democracy does not work, Our Lord designed His Church to be a Monarchy under One Vicar -- the Pope.

Read De Maistre's Le Pape and Soloviev (a Russian)'s Russia and the Universal Church to learn more... ... ... 

\hfill

\texttt{Avery on 2013-10-31 at 10:14 said:}

I, on the contrary, find the Orthodox Church much more palatable than Catholicism. Last Sunday, I sat in on a large Catholic Mass and was blasted with contemporary music and Latino dance-inspired hymns. I didn't think I had preconceptions of how a church should sound, but I was stunned how inappropriate it was. There were Protestants playing ``contemporary Christian music'' in the square outside, and I had the strong impression that they were two peas in a pod. The ROCOR services I've been to are also mostly in the vulgar but maintain an ancient, sacred attitude throughout. This is reflected in the dress I saw: headscarves at ROCOR, lesbian `dos among the Catholics.

The paradox is, if I am church shopping between incompatible faiths anyway, why limit myself to Christianity? If I sense the divine in the ancient, why not just read old books?

\hfill

\texttt{scardanelli on 2013-10-31 at 11:39 said:}

My experience attending mass in a traditional Catholic church celebrating the Latin Mass with priests from the FSSP and a modern mass at a modern church are miles apart. I grew up with the Novus Ordo mass, so there is a bit of nostalgia attached to it, but when one attends the Latin Mass, there is a sense of beauty and presence, and one feels as if he is no longer in an ordinary setting. One is attending a solemn ritual rather than a party. People dress respectfully- the women all wear headscarves and the men wear suits. When I go, I have the sense that I am taking part in something timeless and beautiful. Of course, not all are lucky enough to live near such churches, so I can understand your disillusionment.

I attended the Divine Liturgy at a local Orthodox church several times as well, and found it quite beautiful. I was really attracted by what at the time appeared to be a more developed mysticism. I no longer think that. But in the end, as JA said, it just felt foreign. What is available in the Orthodox church is available in the Catholic church. It is just not as apparent. There are opportunities to immerse oneself in the tradition of the west... one such is going on right now with the MOTT discussion.

\hfill

\texttt{JA on 2013-10-31 at 11:55 said:}

Scardenelli we are 100\& in accord.

I extend an offer to anyone reading this, who has questions about Catholicism, to join me if you are in New York, USA at Holy Innocents Church every Sunday at 10am -- and then tell me if you still think the Mass is vulgar.

I've been to both Latin Masses and Eastern Rite, for a while I went to a Ukrainian church because there was no Latin church I could get to at the time, and -- while this is a personal equation to use Evola's term -- I feel more at home in the Latin rite. The use of instrumental music for one instance appeals to my inner aesthete... ... ..

It's kind of funny that while I may appear as a doctrinaire traditionalist Catholic at times, I actually praise Catholicism more for its aesthetic value to me than for any kind of petty morality, my lifestyle is more that of an Italian Renaissance Catholic.

\hfill

\texttt{David on 2013-10-31 at 14:28 said:}

@Ash : I understand your point, and was not trying to contradict you per se, but to understand. I read as much from Latin fathers that I read from Greek fathers, as much from Roman Catholic than Eastern Orthodox and find myself at home with both of those. Of course, it's not all about books, but rather a way of life and rituals that channel that life; but sometimes I doubt that the two faith are so much apart. I think it's more a problem of degeneration than actual doctrinal difference.

\hfill

\texttt{Cologero on 2013-11-01 at 22:43 said:}

I have missed you dearly, August, and appreciate your insights, but there is really nothing to see here. I have made it clear many times that you cannot choose your religion; however, individualism is well nigh impossible to eradicate in Americans ... here you are free to choose your toothpaste, your church, and even your gender.

Not to single out Avery, but we made this very point in his post about choosing a religion several months ago. However, youth is incorrigible and there may even be a justice in that. He is a victim of what is called ``white privilege'' here, so he has come to expect nothing but the best priests and the most solemn masses as a birthright, so your homely wisdom will most likely fall on deaf ears. Those of us who have been beaten down a bit come to the realization that we get the priests we deserve; this goes double for women and many other things in life that we feel just ``happen'' to us by unknown or even sinister forces.

As for your own unhappy situation, let it serve us all as a reminder that while the joys of victory are ephemeral, the consequences of defeat are painfully long-lasting. Worse of all is an unstable equilibrium based on mutual suspicion.

\hfill

\texttt{Avery on 2013-11-02 at 04:28 said:}

My anchors were cut before I was born. I floated through empty ocean my whole life and only recently became aware of distant shores. I floated by a Buddhist monastery, and I felt no attachment; but a south Indian Brahmin and a Hasidic Jew were pointing me the way to America, and I followed their path. Isn't it odd how such things ``happen''? The word ``happen'' is derived from ``hap'', meaning luck or fortune. It's also cognate with happiness. Faith didn't happen at that Catholic church. It didn't happen at the CCM concert, either. Maybe I didn't have enough humility.

I told my Orthodox friend that he was on the wrong side of the Crusades. He was practically beaming. I started to explain about the Kosovo thing, the one you guys told me about, but he stopped me and said something about representatives of principles, and how you can take them down to the store and exchange them for better ones without affecting the principle itself. A bunch of hooey, if you ask me.

``But those who personify an idea should not, cannot, fall below a certain level for generations; if they do, my friend, the idea suffers too,'' I fired back. I was paraphrasing a book I had read somewhere, by an Italian of the postwar period. Fighting fire with fire!

My Orthodox friend was deep in thought. Finally he said, ``Someone who has actually tasted truth is not contentious for truth. Someone who is considered by people to be zealous for truth has not yet learnt what truth is really like; once he has truly learnt it, he will cease from zealousness on its behalf.''

Something happened!

\hfill

\texttt{Ash on 2013-11-02 at 05:20 said:}

Thank you for your words, August. For my part, my heritage is Westphalian German, which is to say, near 500 years of Protestant heresy. My ancestors have also been Anglicans. Extended relatives create a large Muslim branch of our line in South Africa, due to intermarriage a couple of generations ago. As for me, here on the great Canadian west coast, around me stands the ``tossed salad'' of Canadian multiculturalism and the final break from exoteric religion which comes from being raised by a resolutely nonreligious family.

I don't say these things to complain, but merely to explain the situation many of us face. It's true, religion has its own ``add to cart'' option here and in the United States. When I attended my first Mass at sixteen, on a path that ultimately brought me here, I was in a way taking part in that culture. One hopes that we have all here gone beyond that delusion. The question is, what is ``our'' tradition, the tradition which we have been born into, spiritually if not in terms of our actual family upbringing. If it is simply a question of ancestral faith, then the Roman Church would seem to be the natural home for many of us. But let us not forget that there are a ``second Rome'' and ``third Rome'', which hold fast to the same Christian tradition (assuming it is, as Logres stated, the religion of the Kali Yuga). If it is not by accident that we are born into the current age, may it be possible that it is not an accident that these traditions, which once would have been impossible to access, are now in that Greek church in the suburb? Could that be one of the possibilities which are said to be easier to realize in the Kali Yuga than in earlier ages? If this has false reasoning or interpretation behind it, correction would be appreciated. It might be hard to explain to those noblemen in the Great Beyond, but I imagine that Martel and Charlemagne would have had some interesting conversations with their countryman Guenon about where his path took him.

As for me, I continue my own journey into the metaphysical heritage of the West, which is without a doubt a Catholic heritage. Years ago, I began my first studies into the religions of the world with a prayer that the unknown God would guide me in discerning the true path. Perhaps such a prayer would aid us all better than a shopping cart.

\hfill


\end{sffamily}
\end{footnotesize}

