\section{Initiation and the State of War}

\textbf{Rene Guenon} wrote nothing, as far as I can recall, about his personal spiritual practice. Because he relocated to Egypt, the assumption is that he lived as a Muslim, following all the precepts of Sharia law. This is made clear in a letter written from Cairo to ``L. C.'' on May 17, 1935:

\begin{quotex}
The question of the difficulty of practicing Islamic rites in the countries of Europe is often discussed. The prevalent opinion, which in any case seems most justified to me by the very principles of sharia, is that there can indeed be some exceptions for persons living in non-Islamic countries, their condition being likened to a transitional state or state of war. But it is necessary to add that this concerns only those who remain only at the exoteric point of view. For a realization of an esoteric order, on the contrary, it is necessary not to forget that the observance of rites constitutes the necessary base. Moreover, it is clear that those who want the most must, first of all, and as the preliminary condition, do the least (i.e., observe the rites common to all).
\end{quotex}

We see that Guenon insists that the esoterist must participate in all the exoteric rites that the ordinary faithful are expected to follow. To be concrete, Tradition is not simply a system of thought. A man cannot be a ``traditionalist'' by believing certain metaphysical doctrines or holding particular political opinions. On the contrary, he must actually follow a tradition. That requires societal support to be able to observe the rites common to all.

A day in Islam is structured by calls to prayer, ablutions, avoiding unclean things or activities, and so on. Hence, in 1935 it would have been difficult to do all that in Europe, although I don't know if that is still true in the Europe of today. That was the reason behind Guenon's relocation.

For those left behind in the West, the man of Tradition needs to consider the same issue. For 1500 years the life of a European was also structured by a Tradition. The year involved a complete liturgical cycle, with various celebrations, holidays, festivals, and parades. The day was divided into certain periods of prayer. The rites of the sacraments defined a person's entire life; although ablution was not physical, but spiritual through Penance.

In effect, men of Tradition in the West today are in a ``state of war'' or a state of transition, as Guenon put it, since there is the memory of a recent Tradition yet no way to fully embrace it. Obviously men at war, who are on a journey, may not be able to fully practice the exoteric rites. The question, then, is whether such a transitional state can lead to something. That would require that remnant communities form today so that exoteric rites can be practiced in tandem with the esoteric. Although Guenon eventually became dubious about that prospect, that was a personal opinion not based on any metaphysical principle. In his earliest books, he did think it possible.

However, when Guenon says that the recovery of paganism is no longer possible, that is actually based on metaphysical principles. Early in the Greek cities or in the Roman Kingdom, daily life was indeed fully structured. There were rites involving the hearth, which each home had to have. There were also many rules distinguishing the clean and the unclean. There were ceremonies for various purposes, and so on. Neo-pagan revivals have no way to recover those and artificial reconstructions do not count, since the original religious rites were considered to be divinely ordained. Quoting Nietzsche, psychologizing the gods and goddesses, and creating lists of ``virtues'' may be great fun, but it is hardly what Guenon is describing. Only a new divine revelation can make this a viable option.


\flrightit{Posted on 2014-01-24 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{August on 2014-01-24 at 01:04 said:}

And this is a problem, and the reason certain of us cannot simply dismiss Evola, and his statements, such as the one describing those today who find solutions in religion as (from memory) `unable to stand the tension of the nihilist climate... cases of surrender that hold no interest for us'. Significant words in the ears of some.

A kind of ascesis that sets up barriers between oneself and the world has been, and must always be, a legitimate way for some. If nothing more comes of it, let it be a signal of intent. In some wars, it's good to stand with the regular army. In others, the situation is far murkier, and the solitary ones appear.

Whatever the choice made, complacency is always a mistake; we have no reason to believe that the war is merely local, and who is certain of what awaits him? Though religion, and life, tells of mercy and promotes hope.

\hfill

\texttt{Ash on 2014-01-24 at 03:12 said:}

I was rereading one of Logres' works earlier today and he makes an interesting point in it: the man who plants a garden in the Kali Yuga against all odds has superseded the man who conquers worlds in better times with some odds on his side. In that line, do we overthink the necessity of having to live a religious life the exact way our ancestors would have? Finding the best that one is led to (the FSSP or SSPX traditional masses, the Orthodox parishes, etc) and even re-establishing some of the old traditions in ones' own life or family could work wonders in the Kali Yuga. I've seen the ember days mentioned on this blog before.

As has also been talked about before, those further advanced say that at a certain level we should forget the actual metaphysical goals themselves, at least temporarily, and have present and focused mindfulness on the devotions at hand to assist in the distancing from the ego. Perhaps those around one aren't focusing on metaphysical development as they say rosaries or attend masses, but in that mindset how much does it really matter? Just a thought.

\hfill

\texttt{scardanelli on 2014-01-24 at 12:38 said:}

I have been reading Paul Sedir lately, and he seems to have found a middle path between the esoteric and the lack of societal support for the exoteric. He presents a spirituality of action, of living out the essence of Christianity and the Gospels in everyday life while not necessarily being dependent on occult orders on the one hand or traditional churches on the other (though he certainly supports the Catholic Church and one's involvement in it if this is available to you).

To love ones neighbor, for instance, requires neither order nor church, nor supporting structures in society. Yet to attempt this act reveals the simplicity and complexity of the Christian tradition... What seems like mere sentimentality actually requires a good deal of ``esoteric'' practice and development. To love a man who inspires annoyance or hatred in you requires one to become aware of his negative thoughts, to cease identifying with them, to realize his freedom from these thoughts and to exercise his true will in perfect love.

\hfill

\texttt{Michael on 2014-01-24 at 13:29 said:}

What esoteric rites exist alongside the exoteric rites? I guess that if I have to ask, then they are not esoteric anymore :-) but if anyone can point the way, I would appreciate it.

I try to participate fully in the Catholic Church's liturgical cycle. I realize that the Mass and the post Vatican II calendar leave some things to be desired, but at least it is a legitimate spiritual authority. Meanwhile, I hope for a restoration of parts of the Catholic tradition that were discarded.

@Scardanelli, what writings of Sedir are you reading? Unfortunately, I only have his novel.

\hfill

\texttt{scardanelli on 2014-01-24 at 14:21 said:}

Michael,

Try searching for Sedir at Scribd. There are many works and compilations translated by Zadah Guerin-McCaffery, a member of Les Amities Spirituelles, which apparently is the group he formed after ending his involvement with occultism and Martinism. All of the essays further develop themes he covered in Initiations.

\hfill

\texttt{Michel on 2014-01-24 at 15:19 said:}

In a way, and the realization of this concerns a few people, the problem has never been the existence of the Primordial Tradition. The problem has been a `concealing' of it through the procession of the ages. All other traditions are adaptations of it that are concurrent with the prevailing cyclic conditions. This withdrawal of it, results in a perception of it under various ways, each way congenial to the one perceiving it. Thus one sees it as Islam, another sees it as Christianity and so on. The above article seems to inch at the apparent dissonance between the type of person that can potentially fully realize this and the person that cannot realize the Primordial Tradition . In Islamic esotericism, there are three types of initiates under two primary divisions. The primary divisions are: the murid/mureed and the murad. The mureed is the `normal initiate' the person that experiences a progressive spiritual realization. He is the attracted and God is the attraction. The murad however is different; he is the desired, or the attraction and God is the attracted. God desires him. The result is that he directly, without going through any procession realizes the truth, al-Haqq, even spontaneously. To put it in other words, he is the natural contemplative, the one who's perception of the modulation of the Primordial Tradition is not as intense and in some cases is even lacking as compared to the one who, though being also able to realize this Tradition is however warranted to do so via successive stations or stages.

The other three divisions concern the experience of the Supreme Station and not the manner of ritual activity karma , if any, that results in this experience. The first type of initiate is called the servant. This initiate is basically similar to the mureed, he cannot permanently maintain an experience of Buddhi and can rarely experience Atma, or in western terms, he cannot permanently maintain a state of Nous/Christ-consciousness and rarely does he experience the Father as Niguna Brahaman or in a more direct way, because the spirit/buddhi/nous is the most direct `representative' of the Infinite/atma/Father, it can be said that through knowing it ( Christ) one can know the Infinite (the Father), as through knowing the ray of the sun, one knows the sun, because it (the ray) is merely the sun `projected and limited to the ray'. However, the knowledge of this ray or Christ is not the same as the knowledge of the sun or the Father, because the Father is not limited; even though the spirit/buddhi or the experience of Christ is Absolute, by virtue of it Being the first Principle, it is limited/finite. Thus while to know Christ or to experience satchiananda is to know `all things' it is still a limited state of knowledge, though limited only to itself. The second type of initiate is called the attracted. This can either be a mureed or a murad. Both the former and the latter can assert a permanence in the experience of the Supreme Station. Moreover, they can also ``experience'' the Infinite or the Father. However, when they ``experience'' this, they cannot experience anything else. They do not `come back' as the servant does; they are like the yogins whose bodies may be severed in the most atrocious of ways without them eliciting any reaction. Sometimes, they even withdraw their bodies at the moment of realizing the Supreme Station, it becomes the `Resurrected Body' or the `Transformed trans- meaning beyond, trans-formed `gone beyond form or figure', to `be formless' Body' and `are no more because they have been taken by God' like Enoch or Elijah. The final type of initiate is the Companion. Again, this can either be a murad or mureed, but is mostly a mureed, because he would have to travel through all the stages and know them perfectly; he would undergo a trans-migration , that is a vertical ascension, of being followed by a horizontal expansion of being, or both Satva and Rajas as compared to only Satva. This type of initiate is regarded by the Sufi Masters as the `best' type of initiate. This is because his experience of the Supreme Station is of the Father and while he does ``experience'' this, he still ``experiences'' everything else as distinct, in a way, it's a sort of superimposition of two visions, one Unitive, the other one a vision of multiplicity. He is the one that is able to guide all beings in all worlds to realize the Father in a manner that is congenial to them. Not that the other two types cannot act as guides, but due to the lack of this double vision, they cannot properly do so, for instance, the second type of initiate, the attracted might assume that everyone is like him and hence may sometimes impose `harsh' conditions to those under his guidance, while the first type of initiate due to his `return' to the normal point of view, though possessing Gnosis, if laxity is entertained on his part, he may fall prey to the ego's relentless jaws, or to maya's seductive dance, or to the power of the prince of the air, and hence may mislead those under his guidance. The Companion however possess a vision of all beings, from their respective point of view while at the same time being completely free from maya and being a True Friend of God, possessing His vision as well; he thus is the perfect guide, because he is not privy to a fall and can also raise the fallen to the Supreme Station without `harshness'; with a `kind and gentle spirit'.

The reason I've stated all that is due to an implied relationship that exists between these three types of initiates and the state of the world today, in the establishment of an elite. I have not stated all the characteristics of these different types of initiates, but simply the relevant ones. To me, the connection lies in these three types of men and giving unto them what is rightfully theirs. Due to the extreme obstruction of the Primordial Tradition however, what is needed is the murad or the one that the Primordial Tradition or God desires. This one is scarcely atrophied and is `soft' and is naturally receptive to the presence of the Divine. Thus via an ascesis or sometimes even without one, this one can experience a spontaneous psychic and spiritual regeneration, undergoing both a second and third birth by virtue of mere exposure to the world itself, due to existence of a Spiritual Order that reflects itself cosmically as the drama that unfolds in every macro-cosmic aspect, even those of `ordinary life'; and these births triggered as a result of predisposed identification to the processes that occur outside with the processes that occur inside, or to put in another way, because all objects perceived ``outside'' are nothing other than an extension of the being, the truth is there is no ``outside'',there is only one Reality experienced under different states of being and this under different constrictions of consciousness, which is the sufficient reason for an experience of any state of being. The murad knows this truth innately even if not consciously, ergo by virtue of this softness, or innate disposition, an event as simple as a bird flying to him can suddenly erupt a displacement of consciousness from the individual state of being to the supra-individual state of being. To digress minutely, It seems that Guenon in his `bureaucratic' expositions on the Arcane had this type of initiate in mind, upon insisting that the rituals simply be preserved properly, even if those preserving them do not quite understand what they are doing, or worse even if the rituals do not affect them, so that, when one such initiate (murad) is attracted to this order as an expression of himself, upon exposure to these rituals, his realization would be almost immediate. Evola may not have completely understood this, hence his critique on Guenon's insistence on this `mechanistic' preservation. Ironically, on his own exegesis of the Arcane, by giving practical techniques to the qualified, he acknowledged the preservation of the type of man who would, by virtue of his `softness' be privy to a vertical initiation, devoid of a corporeal order, and hence who would benefit from these practices. He failed to apply the same reasoning to a different situation. To continue, because in the West there no longer exists any genuine expression of the Divine as an Order, ascesis to me seems plausible. The ascetic as a murad who becomes a companion- the third type of initiate- would be more or less, upon congenially experiencing the Supreme Station or the Father, able to convey guidance to the first type of initiate in the primary division, the mureed. The second type of initiate, the murad who becomes ``an attracted''-experiences nothing other than the Truth- need not any intervention, but he cannot properly or even significantly intelligently relay his experience to others that may be able to positively receive them. By `relay', other than basic discursive activity, I of course mean `bring water from the heavens' or `reveal the names of things the basis of Magic ' or by virtue of the harmonious correspondence that exists between the different modalities of reality, show the symbol of the symbolized to his brethren, again show/bring, not `invent', but literally show you the thing as it is from one order to another, given its essential nature. Thus, the only chance the West has, in the period of transition, is the murad who becomes a Companion, to a lesser or greater degree, the `Super Saint' who, being the friend of God would convey His Will to those who also wish to know and adhere to it, though being of a different dharma and hence not credulous to a direct vertical unveiling of God. It is those who Being desired, invoke Him, and bid Him ``come down!'' their hearts always oriented towards Him, never awakening from prostration that He will listen to.

In conclusion, acsesis does not mean `going away from the modern world' Such a description is derogatory, because the modern `world', based on the stupid need for agitation in all forms, characterized by a wanton lust for the excitement of the most base and lowest human impulses, those shared by animals, and in some instances, those that are so low to the point of disgusting these same animals is illusory. Unlike a Traditional Civilization, that is based upon an assertion following an affirmation of the Spiritual Order which, being the first Principle, is the perfect ``existential condition'' and hence in a way, the only thing that properly speaking `Exists' or `Is', reflecting this by a conscious participation in the essences of things and corporeally structuring itself in concordance to these essences from a substantial point of view, the modern world, being literally founded on the collegiate of illusion and indulgence is `not real'. In relation to it, there is nothing to `go away from'. Ascesis is the reception and the kilning of the received to the point of a sudden violent burst in brazen flames of spiritual truth, at times instantaneously and at other times seemingly progressively, the result of which is the experience of the Christ and at times, of the Father. In short, in periods of transition, the thing that can prevent a collective violent shift from one cycle to another, manifested as a cataclysmic event, is first, the existence of the murad who is a Companion, then the existence of the servant. Because the aspect of interest pertains to a collective and social sphere, the attracted, and more specifically, the murad who becomes the attracted initiate, cannot be of much help to anyone but himself. ``The Spirit goes wherever it wishes'', to say it differently, the Spirit will go to those it desires, the murads being desired, upon calling it, get a response.

\hfill

\texttt{X on 2014-01-27 at 23:42 said:}

``A man cannot be a `traditionalist' by believing certain metaphysical doctrines or holding particular political opinions.''

Can one really believe certain metaphysical doctrines? When one begins to ``believe'' it, it is no longer truly metaphysical.

\hfill

\texttt{scardanelli on 2014-01-28 at 08:41 said:}

It seems to me that one's belief or lack thereof is irrelevant. True metaphysical doctrine does not change according to our level of understanding. Truth exists, and we must conform ourselves to it, rather than it to us. At any rate, one cannot know until one believes, and to believe is at least better than merely reading and understanding on a cerebral level.

\hfill


\end{sffamily}
\end{footnotesize}

