\section{The Crown of Creation}

\begin{quotex}
[This] allows us to see the mysterious relationship that exists between Creation and the Incarnation which is its crowning. \flright{\textsc{Rene Guenon}}

\end{quotex}
Among all of Rene Guenon's writings, it is common to overlook the many texts dealing with the recovery of tradition in the West. For a time, that was his primary focus. Guenon taught the Incarnation, the devotion of the Sacred Heart, and the social reign of Christ the King.

I have in mind the first two essays, ``Word and Symbol" and ``The Reform of the Modern Mentality", included in \emph{Symbols of Sacred Science}. These were written in 1926 and published in \emph{Regnabit}, a Catholic journal. At that time, he was still interested in recovering the intellectual resources of the Western Tradition. Guenon makes it perfectly clear that this is not a wasted effort when he writes:

\begin{quotex}
It is certainly wrong to claim that nobody could understand an exposition of pure doctrine. First of all, why wish to restrict oneself to the lowest level on the pretext that this is the level of the greatest number, as if it were always necessary to consider quantity rather than quality? Is this not one consequence of the democratic spirit which is one of the characteristic aspects of the modern mentality? And besides, can they be so sure that so many people would be incapable of understanding? Should we not think that even those who would not understand everything would nevertheless derive some benefit from it, and even more than might be supposed? 

\end{quotex}
\paragraph{Word and Symbol}
The first essay is ``Word and Symbol". Guenon explains why the symbolic form is so important in the transmission of traditional doctrinal teachings:

\begin{quotex}
Symbolism seems to be particularly well adapted to the exigencies of human nature, which is not a purely intellectual nature but requires a sensory basis from which to raise itself to higher spheres. We must take the human make-up as it is. 

\end{quotex}
This makes sense, given the Thomist maxim that all knowledge first begins with the senses. So it would follow that knowledge of spiritual things would use sensual imagery. Obviously, there is the possibility that the symbol replaces actual knowledge of higher spheres. This leads many astray, as they envision spiritual things as the continuation of the physical world. Boris Mouravieff writes this about symbolism:

\begin{quotex}
The understanding of esoteric symbols is the result of revelations given by divine grace, and this demands a progressive opening out, in quality and in strength, of adequate faculties latent in the student. One arrives at this through a tension of the will toward the goal of the search. When it is strong enough and properly oriented, this tension resolves itself in a series of partial revelations that are acquired, and which lead by stages toward those that are given.

The symbolic form is equally good for all, because it helps everyone to understand the truth it represents more or less completely and more or less profoundly, each according to the measure of his own intellectual possibilities. 

\end{quotex}
\paragraph{The Logos and Creation}
Guenon writes that

\begin{quotex}
symbolism has its basis in the very nature of beings and of things, it is in perfect conformity with the laws of that nature, and since natural laws are only an expression and an exteriorization of the divine Will. 

\end{quotex}
Thus, the symbol's origin is higher than the human state. John's Gospel starts: ``In the beginning was the Word." The Logos is both Thought and Word. 

\begin{itemize}
\item In itself, it is the Divine Intellect, the dwelling place of ``possibilities", ideas or essences. 
\item In relation to us, the Logos manifests and expresses itself by Creation, in which the possibilities of manifestation are realized in actual existence. Hence\emph{, the world is like a divine language to those how know how to understand it.} 
\end{itemize}
Here Guenon makes clear that the ``possibilities" he described in \emph{The Multiple States of Being}, reside in the Divine Intellect, or Logos. This puts him in the mainstream of the Western Tradition, from Augustine to Thomas Aquinas.

\paragraph{Incarnation and the Sacred Heart}
The Logos is expressed not only in Creation, but also in revelation. Guenon writes:

\begin{quotex}
The primordial Revelation, which, like Creation, is the work of the Logos, is also incorporated in the symbols which have been transmitted from age to age ever since the origins of humanity. … Moreover, we see in this incorporation into symbols of the non-human tradition a kind of anticipated image, a prefiguration of the incarnation of the Word. \emph{And to a certain extent does this not also allow us to see the mysterious relationship that exists between Creation and the Incarnation which is its crowning}? 

\end{quotex}
The Heart is a universal symbol which takes the specific form of the Sacred Heart in Christianity. The Sacred Heart is the center of the being, not just symbolically, but really. As such, it must occupy a central place in all doctrines issuing from the primordial tradition.

\paragraph{The Social Reign of Christ}
In traditional societies, all social organization must be based on traditional principles. This is true not only in the Eastern traditions, but in the West. In recent posts, we saw that Ancient Rome was based on spiritual principles that regulated daily life. For them, piety was the greatest virtue. Similarly, Guenon points to the Middle Ages as Tradition in the West. Positive law was based on traditional principles of the natural law. In a rare outburst of passion, Guenon announces his political program for the West:

\begin{quotex}
It is high time we reacted against this tendency [i.e., the reversal of the relationship between religion and politics], and in this connection the affirmation of the social Reign of Christ is a particularly opportune manifestation; but to make this a reality, the present-day mentality must be reformed. 

\end{quotex}
The ``opportunity" Guenon mentions is the encyclical \textit{Quas Primas}\footnote{\url{http://w2.vatican.va/content/pius-xi/en/encyclicals/documents/hf_p-xi_enc_11121925_quas-primas.html}} by Pope Pius XI, announcing the Reign of Christ the King, which was promulgated just prior to the publication of Guenon's essay. At that time, Guenon was affiliated with ``The Society for the Intellectual Propagation of the Sacred Heart." He described its task this way:

\begin{quotex}
The reform of the modern mentality, namely the restoration of true intellectuality and of traditional doctrine, is certainly a considerable task; but is that a reason for not undertaking it? It seems to us that such a task constitutes one of the loftiest and most important goals that can be proposed for the activity of a society …

All efforts in this direction will necessarily be oriented toward the Heart of the Incarnate Word, the spiritual Sun and Center of the World, in which are hidden all the treasures of wisdom and science; not of that vain, profane science, but \emph{the true sacred science which, for those who study it in the proper way, opens unsuspected horizons that are truly unlimited}. 

\end{quotex}
\paragraph{Epilogue}
In the intervening years, we see that the ideal of the Reign of Christ the King\footnote{\url{http://www.remnantnewspaper.com/Archives/archive-christ_the_king.htm}} has come to naught. Even in the Vatican, the idea died of neglect among the documents of Vatican II. The modern mentality, founded on an incoherent mixture of rationalism, sentimentalism, and humanitarianism, is more entrenched than ever.

Guenon himself moved on, and lived as a Muslim in Egypt. Nevertheless, it is necessary to keep in mind what he wrote about his ``conversion". It was not a horizontal conversion, i.e., the rejection of one exoteric form and the adoption of a new one. Rather, for him the move was a matter of convenience. The key to understanding this is found in Boccaccio's \emph{Decameron}, which Guenon considered to be an esoteric text.

So it changes nothing essential. What he wrote about the Logos, the Incarnation, the Sacred Heart, and so on, still stands as metaphysically sound. The reform of the modern mentality, which is the Western mentality, is still a noble task. Guenon decided to do that from the outside, as it were. Others will persist in reforming it from the inside.



\flrightit{Posted on 2014-12-31 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Logres on 2014-12-31 at 14:44 said: }

I saw an article yesterday regarding insurance in which someone said ``Society has to be forced to make Society help Society". Isn't this the kind of double-speak that you have to ultimately engage in when you shy away from a traditional Order? You don't want to say ``people have to be made to be good" so you drape it with ``Society". When you deny the One, the Two becomes an endless war of dichotomies and buffering lies. Eventually, ``society" will lose the Two as well, and just go to compost. I need to read more Guenon.


\hfill

\texttt{Matthew Livermore on 2015-01-01 at 10:41 said: }

Can you say any more about how the Decameron is the key to understanding Guenon's conversion?


\hfill

\texttt{Cologero on 2015-01-04 at 10:45 said: }

Re Boccaccio: First day, third story.

Guenon alludes to this story in the essay, ``The Secret Language of Dante."


\hfill

\texttt{August on 2015-01-09 at 15:13 said: }

The modern mentality can only view all reformation in this direction as an oppressive attack. Proper reformation of this mentality is nothing less than its destruction.

The modern mentality knows this too, that is the reason for its instinctive hatred of tradition. It only pretends to endorse freedom of expression as principle, otherwise it would be able to appreciate, for instance, both blasphemous doodles and the angry acts they inspire, each of which can be morbidly entertaining to the neutral observer. But this mentality knows its limits – death presents it with nothing but the prospect of chastisement, so death alone rocks its foundations and makes it cry foul. The modern world lives for death.


\end{sffamily}\end{footnotesize}
