\section{Monkey People and Solar Beings}

\begin{quotex}
I am quite sure now that often, very often, in matters concerning religion and politics a man's reasoning powers are not above the monkey's.
\flright{\textsc{Mark Twain}}
\end{quotex}

\hfill\break
In explaining the mystery of birth\footnote{\url{https://www.gornahoor.net/?p=3972}}, \textbf{Julius Evola} makes it perfectly clear that his entire synthesis depends on metaphysical principles and cannot be explained in terms of biological heredity. His terse exposition needs to be expanded with references to \textbf{Rene Guenon}'s \emph{Multiple States of Being} and the Hermetic teaching on Providence/Will/Destiny. Since this teaching is included in what he calls the ``organic vision of the Germanic-Roman civilization of the Middle Ages'', we shall also rely on the spiritual Tradition of that civilization to clarify his points.

Obviously, he is not offering a logical or scientific proof, although the explanation is rational within the assumptions of Tradition. Ultimately, however, the proof lies in ``remembering'', in bringing into the light of consciousness the awareness of who you are. Those encumbered with a scientific viewpoint have a self-understanding of being the descendent of a monkey; nothing here will convince them otherwise. At best this exposition may awaken dim memories in those who recognize themselves as coming from a higher, transcendent source.

We alluded to this previously in a post on the Theology of the Body\footnote{\url{https://www.gornahoor.net/?p=8093}}. As an heir to the Germanic-Roman religion of the Middle Ages, \textbf{John Paul II} wrestles with similar issues. How am I a man or a woman? JP II treats the Genesis account as a ``true'' myth and tries to ``remember'' that originary state via a phenomenological analysis. He is then compelled to reject evolutionary accounts in favor of a transcendental source; one's sex is not an arbitrary act of God but is essential to one's nature.

Recall the traditional teachings on the two poles of essence (what you are) and existence (that you are). So we can follow the career, let's say, of ``Harry'' from his essence to existence. In what sense does Harry preexist, as Evola claims? Obviously, the state of preexistence means ``not yet existing''. Another way of putting it is that Harry is a ``possibility of manifestation'', i.e., an idea or essence in the Divine Mind. Priority has several meanings, so Evola correctly notes that Harry does not preexist in the temporal sense of priority, but rather in an ontological sense.

As an essence, Harry has various qualities: e.g., he is a male, of a certain spiritual race, with qualities of a specific caste, for example. This is what Evola means by \textbf{karma}, or what Harry is bringing into the world that makes him Harry and not someone else. As such, it is part of his essence. So he will be born with compatible preconceptions as we pointed out in Esoteric Stoicism\footnote{\url{https://www.gornahoor.net/?p=8147}}. That is why the awareness of one's own true being is so important.

God, then, that knows Harry's essence before Harry is born. Harry's birth will come at the intersection of two lines. The vertical line is transcendental and outside of time; that is Harry's essence. The horizontal line represents Destiny, i.e., the World Process, which includes the physical and psychical environment into which Harry will be born.

Guenon points out that Harry's manifestation, or birth, must be compossible with other manifested beings. Hence Harry's birth must coincide with the opportune moment in the World Process. Then he will come into existence in a particular place at a particular time. Note that the categories of being\footnote{\url{https://www.gornahoor.net/?p=8160}} come into play: Harry is a substance, with certain qualities, existing in a place and time, with relations to family and community, and so on.

Note that Harry's possibilities of manifestation are not exhausted just at birth. In his life he will have the opportunity, or actually duty, to continue to manifest all his possibilities. This is his \textbf{dharma}, or the path he needs to follow, not because it is some outside requirement, but rather because it is his own nature.

So in what sense does Harry will himself into existence, as Evola claims? There is a metaphysical principle that something potential cannot cause something actual. Specifically, Harry's preexistence as a possibility of manifestation is only potential. Hence, pace Evola, he cannot will, or cause, himself into existence.

Evola correctly mentions that the Medieval Germanic-Roman religion teaches that God creates the intellectual soul at conception which then is the form of the body. To create means to bring what was potential into actuality. However, Evola neglects to mention that the intellectual soul is not Harry: the Person, the ``I'', the Self of Harry is not the soul in Thomist teaching. The Person of Harry is a constant presence above all the psychical events in the soul.

So in what sense can Evola's claim that Harry chose his birth be defended? His claim that the human Self is the production of a preexisting spiritual entity is rather strange. What could that entity be if it is not the Self of Harry? Since Evola rejects reincarnation, that entity must be the one and only transmigrant\footnote{\url{https://turiya.vidya.hu/konyvtar/pdf/On\%20the\%20One\%20and\%20Only\%20Transmigrant.pdf}} in \textbf{Ananda Coomaraswamy}'s phrase. That can only be God.

In another sense, however, as preexisting in the Divine Mind, Harry's will and God's will are united, since God is simple, without separate parts. Perhaps, then, there is the ``Supreme Identity'' before birth, so Harry willing of his existence is one with God's willing of his existence.

\paragraph{Excursus on Reincarnation}

Since karma and dharma are often associated with reincarnation in the popular mind, it is opportune to explain why this doctrine is untraditional. First of all, it is anti-essentialist. Harry is a substance. So who reincarnates after Harry? Well, it cannot be Harry since his time and place of incarnation were already determined by this essential nature. So, it if it is Dick, then Dick is a possibility of manifestation that is not Harry. There is nothing else to incarnate beyond Harry and Dick. Since reincarnationists believe that one can return as a different sex, or in a different state of life, then one no longer has a given nature and a fortiori no particular dharma to follow.

The other error is that karma gets reduced to the accumulation of accidents in life, and is not an aspect of one's essential being. Dharma shifts from the duty to be faithful to law of one's own nature to obeying some external laws or performing rituals in order to avoid accumulating more bad karma or ``burn off'' earlier karma.


\flrightit{Posted on 2015-07-08 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}

\texttt{Sparrow on 2015-11-18 at 17:30 said:}

While the criticism of karma was on the nose; the argument against reincarnation seems to contradict previously discussed Traditionalist beliefs. If Harry is a mere substance, how can he pass into post-mortem and future states of being?

If he is capable of this, why can't he be equally capable of choosing to be born on earth again in the same state? Given that Harry (from what I understand) is a spirit which precedes the soul and body, why is it not possible for him to enter the human state again with a new soul and body?

Unless the Dalai Lama and multiple other Eastern figures have been playing us for fools this entire time, reincarnation has more traditional support than the idea of multiple states of being. Nations that believed in reincarnation still required people to adhere to the dharma they are born with, even if they believed a woman had been a man in a past life, or vice versa.

I mean no disrespect, but I find this extremely baffling.

\hfill

\texttt{Cologero on 2015-11-18 at 18:11 said:}

Here is Evola's full essay on \textit{Karma and Reincarnation}\footnote{\url{http://www.gornahoor.net/library/KarmaAndReincarnation.pdf}}.

Harry is one particular state of a ``more inclusive'' Being. Guenon claims that the Being cannot take on the same state twice (in this case, the human state). So it is the Being that is Harry in one state, but xxx in another. The human state includes spirit/soul/body. The spirit is not wandering about looking for a soul and body like a hermit crab looking for a shell. Rather, it is the form of the soul and body. Harry qua Harry is not ``choosing'' anything directly, unless he is a realized being.

Coomaraswamy distinguishes transmigration (OK) from reincarnation (not OK), as in the essay ``The One and Only Transmigrant''.

In \textit{Does Death Exist}\footnote{\url{http://www.gornahoor.net/?p=2961}}, Serrano mentions a curious inconsistency about reincarnation.

Your idea of switching sex between incarnations seems baffling.

On the other hand, it is not just the Dalai Lama, but Tomberg, too, accepts reincarnation. Since I have no recollection of prior human incarnations, I have no verification of any of it. Without that, the idea means nothing to me.

\hfill

\texttt{Sparrow on 2016-02-23 at 21:02 said:}

The analogy of the hermit crab, while useful in pointing out the flaws of Cartesian dualism, does not really apply to this case. Given that the spirit is not being presented as something that inhabits the soul or body, but as something that forms them.

The idea that one's sex can change between births is not mine, I am simply recounting what traditions have taught for generations. Evola himself admits that the idea can be found in Plato's works as well as certain doctrines of Islam (Eros and the Mysteries of Love page 150). He did not even mention Jainist tradition, which teaches a woman must be born as a man to attain enlightenment.

But neither do you have recollections of previous states of existence (or else, you and Guenon could describe them). We have actual traditions which support the idea of reincarnation, no proof that the idea of multiple states was ever taught, and Occam's Razor would dictate that reincarnation be the more reputable option.

I have read Serrano's curious observation, and while it is not without its merit, I fail to see how it would disprove reincarnation in favor of Guenon's transmigration, especially in the cases of non-monist traditions such as Buddhism.


\hfill

\end{sffamily}
\end{footnotesize}

