\section{The Kingdoms Fell into Disorder}

Cologero has been channeling me for quite some time now and much of what he takes credit for is really mine. Nevertheless, he recently got a little huffy about it --- or so it seemed --- and he challenged me to write the material directly. Now the last thing I need is for a bunch of curiosity seekers attempting to invoke me. Cologero assured me that very few people read his blog, so I decided to risk it. In particular, he asked me to explain Chapters 18 to 21, regarding the deterioration of a civilization (18 \& 19) and the character of a Taoist sage (20 \& 21).\footnote{See Section \ref{sec:PrincipleAction2021} in this book for \textbf{Julius Evola}`s translation of and commentary on these chapters of the Tao Te Ching.} 

I know I have a reputation for being ``inscrutable'', although --- in my opinion --- that is due more to poor translations than to the difficulty of the material. In other words, it is the Tao that is inscrutable, not me. Unfortunately, any attempt to scrutinize the inscrutable is fraught with issues. Have you ever told a joke to your girlfriend and got nothing back but a blank stare? Eventually, she will protest, ``I don't get it.'' After explaining the joke, she finally claims to ``get it''. Yet, at that point it is no longer a joke, but something else, something markedly inferior. A joke depends on provoking a sudden shift in perceptions.

Likewise, knowing the Tao works the same way. There is a sudden ``aha'' experience, a sublime and rapid change of viewpoint, or what is called ``getting it''. The danger of an explanation is that it may preclude that very aha experience. That is why I wrote:

\begin{quotex}
The (discursive, reasoning) intellect is the source of sorrow

What consistency do dialectical distinctions have?
\end{quotex}

Once you get it, there is no point in explaining, debating, or analyzing it. Right? Or debating a joke? Wrong. I've heard that your civilisation will debate jokes incessantly. That's because you've lost the natural sense of what is funny, and that's not all you've lost. Following the Tao means a return to a natural primordial state. When that state is lost, then the discursive intellect tries to replace what was lost with something artificial and factitious.

In your society, there is a lot of talk about benevolence, justice, and so on. Now there may indeed be acts of benevolence and compassion, but there is simultaneously a lot of hypocrisy. There should be a natural connection with your ancestors and family. But even that is lost, and replaced with ``definitions'' of a family. Do you see what happened? What was natural and spontaneous becomes suddenly a matter for scrutiny and debate. The result is hardly a clarification, since no one is even sure what a family is any longer.

In the primordial state, the chatter about ``humanitarianism and morality'' will disappear, to everyone's benefit. Can anyone who watches your news shows for 10 minutes doubt this? Your partisans claim to have a monopoly on morality within the system, each party accusing the other of evil. Then as a whole, you boast about your high standards vis-à-vis the rest of the world. This is not a critique, but rather an observation since everything is a mixed bag. If you understood good and evil, there would be little discussion of it.

As I pointed out some time ago (really, what has changed since then, despite all the warning?), the mass of people are seeking little more than pleasant diversions. There is no morality in following one's vain desires, tendencies, orientations, and interests. Quite the contrary, although you would never believe it, the way people talk about it. I assure you, however, that there is nothing ``natural'' about them at all. Emphatically not, since the path to true liberation is to become free of such things.

Instead of liberation, those who become aware of those dangers, impose instead a moral code. But that is suppression, not liberation, just adding another layer on top of the already unnatural layer. It leads away from, not to, the natural state.

Following desires, seeking pleasure, defining morality, and rational thought do not lead to the Tao, then what does? I tried, in Chapter 21, to show, by way of personal example, the alternative. I am not moved by feelings, but live in the calm waters below their current. I am at home no matter the outward circumstances. While the world seeks excess, I seek simplicity. Those with superficial vision may consider me to be simple-minded, incompetent, and impractical. I prefer to live in silence than in conversation, to be alone rather than in a crowd. I live like the ocean, wave after wave, without pause. Yet, since I am not battered about by desires, I am a Real I.

The manifest world is the exteriorization of the Tao. Although undifferentiated and unintelligible, it nevertheless contains all things as possibilities. Existing things are contained in the Tao, as well as their essences. The world cannot keep us safe, our safety resides in the Tao. The Tao does not change, yet everything proceeds from it.

How do I know this? Through direct ``super-rational'' intuition. Do you get it now?


\flrightit{Posted on 2017-06-25 by Lao Tzu}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Fred on 2017-06-26 at 06:13 said:}

Fun fact: when Eliphas Levi channeled Apollonius of Tyana the only answer he received was ``death''. I see Lao Tzu is a nicer person.

Also, according to Guenon's ``the spiritist fallacy'' you can't really channel the soul of a dead man, only his `ob' (the word used in the Bible). Sort of a shade, but not his actual self. I don't remember the egyptian word, but it's that thing they tried to preserve inside mummies.

\hfill

\texttt{James on 2017-06-26 at 16:02 said:}

Thanks for a great post.

\hfill

\texttt{THAT on 2017-06-28 at 07:13 said:}

That's true about channeling. Certain kinds of deceased saints and sages, however, exist in states where they may easily interact with people in our world in a recognizable form. One should keep in mind that other beings of the subtle planes may masquerade as something they are not as a deception.

\url{http://www.thejinn.net/ascended_master.htm}

\hfill

\texttt{Aleksandar on 2017-06-28 at 12:50 said:}

Regarding channeling and communication with sages, Goethe once popped up in my dream just to tell me that he achieved some kind of divinity, a ''godhood'' to use his own words. It was a very vivid, unusual dream. I wonder if I was trolled by some astral being in disguise, or fell victim to my own mind's play.

Great text by the way. I'm reading ''The Great Triad'', my first serious introduction to Taoism. And Evola's commentary on ''Tao Te Ching'' somehow slipped me, until now. This site is truly full of hidden gems.

\hfill

\texttt{Arthur Konrad on 2017-06-28 at 15:57 said:}

Two Taoists walk into a mechanized society

\hfill

\texttt{THAT on 2017-06-29 at 06:51 said:}

`Sufism and Taoism' by Izutsu is another useful study that I highly recommend.

\hfill

\end{sffamily}
\end{footnotesize}

