\section{The Primacy of Will}

Profane psychology bases itself on some misconceptions. Its goal is to ``understand'' the reasons for behavior. Its method is to probe hidden desires, attractions, impulses, images, daydreams, memories, and thoughts in order to recover the ``real self''. Those are all perturbations of the mind.

Esoteric psychology is just the opposite. Its technique is not analysis, but rather concentration, meditation, contemplation. According to Hermetic teachings, concentration is the transition of the directing center of the brain from the mind and imagination to the will and morality. Hence, the discursive mind must yield to the will and is therefore insufficient to fully comprehend it.

When the perturbations of the mind attenuate, then consciousness returns to silence, rest, calm, and relaxation. The center of awareness is the Real Self, not the thoughts, fantasies, emotions, sensations, etc., that may arise spontaneously.

The fiction of the primacy of the Mind is easily illustrated. The three centers of the human being engender sensation, emotions, and thoughts. Obviously, sensation operates the quickest, then emotions, and finally thought. For example, touch a hot stove, and the finger pulls away long before the mind is fully aware of what just happened. Emotions are slower than sensations, yet manifest prior to thought.

The opposite is clearly false. For the stove example, it is not the case that the mind determines the stove to be hot, and then commands the finger to move. You can find examples of emotions arising before the discursive mind can understand its significance.

That is why esoteric training proceeds in reverse order. First, one tries to master thoughts, or the intellectual center. Then, the next stage is mastery over the negative emotions. Finally, if it be possible, there is mastery over the body to create a spiritual or glorified body\footnote{\url{http://taylormarshall.com/2012/10/do-you-know-four-properties-of.html}}. At that point, corporeality detaches from physicality, with the following qualities:

\begin{itemize}


\item \textbf{Impassibility}: no longer suffer physical sickness or death


\item \textbf{Clarity}: free from deformity and filled with radiance


\item \textbf{Agility}: can move at the speed of thought


\item \textbf{Subtlety}: can pass through physical objects

\end{itemize}


Finally, we can tie the Will to the virtues. Virtues, like courage, can be quite difficult to define intellectually. Nevertheless, one can be courageous without being able to define it. Moreover, how to act courageously cannot be learned from intellectual study alone. That is our final proof.

\paragraph{The Cosmic Hierarchy}


Profane psychology is committed to naturalism, so it has to assume that thoughts arise out of brain activity. However, when the ``zone of silence'' is achieved, the influx of spiritual forces can be recognized. The diagram of \textbf{Robert Fludd} in The Cosmic Hierarchy\footnote{\url{https://www.meditationsonthetarot.com/the-cosmic-hierarchy}} demonstrates the levels of the hierarchies. These fall into several bands.

\begin{itemize}


\item The sublunary or elemental spirits. These include the earth itself as well as the other elements: fire, air, and water.

\item The planetary spirits.

\item The stars

\item The angelic hierarchy.

\item The Higher mind.

\item The Trinity
\end{itemize}


This is a practical teaching and is meant to be experienced directly. The temptation to create sensory images in the mind, or to expect to ``see'' them with the senses, must be avoided. Rather, they can be experienced as thought forms, or constellations of related thoughts.

\paragraph{Elemental Spirits}


When the mind is attached to or obsessed with these thought forms, it is not free. The True Will is free, but that requires the zone of silence from such thoughts. The visionary \textbf{Anne Catherine Emmerich} claimed that if these spirits could be seen by the senses, the vast number of them would occlude the sky. In practical terms, that means that various thoughts are constantly impinging both on the individual as well as on the human race. The elemental spirits are not benign, yet they are not exactly fallen angels.

Anne Catherine indicates that one such spirit is assigned to each individual:

\begin{quotex}
I have seen that every man receives at his birth two spirits, one good, the other evil. The good one is heavenly by nature and belongs to the lowest hierarchy; the evil one is not a devil, not yet in torments, though deprived of the vision of God.
\end{quotex}


She provides a description of them:

\begin{quotex}
In these spheres dwell the bad spirits who at each man's birth are associated to him by an intimate relation which I clearly understand, which excites my wonder, but which I cannot now explain. They are not lovely and transparent like the angels. They shine it is true, but by an external, unsteady light, as if by reflection. They are either slothful, indolent fanciful, melancholy, or passionate, violent, obstinate, stubborn, or frivolous, etc., a personification of the different passions.
\end{quotex}


Their effect can be recognized:

\begin{quotationx}
These spirits have something sharp, violent, and penetrating in their countenance. They attach themselves with extraordinary tenacity to the human soul as insects to certain odors and plants, rousing in them all kinds of thoughts and desires. They are full of stings, rays, seductive charms. They themselves produce no act, no sin, but they withdraw man from the divine influence, lay him open to the world, intoxicate him with self, bind him, attach him to the earth in many ways.

I have seen that certain inclinations and aversions, certain involuntary antipathies, and especially the disgust we have for certain things, such as insects, reptiles, vermin, etc., have a mysterious signification, since these creatures are images of those sins and passions to which, through their connection with these spirits, we are the most exposed.
\end{quotationx}


The Cthulhu stores of \textbf{H. P. Lovecraft} can be read in this light. The monsters are amoral, with their own aims that are indifferent to human life. It is that indifference that is the most frightening, since evil beings can be recognized and battled. The thought of being eaten by an animal like a shark, alligator, or lion, has always horrified me. Obviously, the animal is neither evil nor immoral. The horror arises from its complete indifference to the qualities of the human being who is nothing but protein to it.

\paragraph{Possession by Elemental Spirits}


\textbf{Valentin Tomberg} describes how possession works\footnote{\url{https://www.meditationsonthetarot.com/possession-by-evil-spirits}} and relates it to phenomena recognized by psychologists.

\begin{quotex}
The objective of Sacred Magic is more than healing, pure and simple. It is the restoration of freedom, including freedom from the stranglehold of doubt, fear, hate, apathy and despair. The ``evil spirits'' that deprive man of his freedom are not at all those of the so-called hierarchies of evil or fallen hierarchies. Neither Satan, nor Belial, nor Lucifer, nor Mephistopheles has ever deprived anyone of his freedom. Temptation is their only weapon and it presupposes the free will of the one tempted. But possession by an ``evil spirit'' has nothing to do with temptation. It is the same thing as with Frankenstein's monster. \emph{A man engenders an elemental and becomes thereby the slave of his own creation.} The demons or evil spirits of the New Testament are today called in psychotherapy obsessive compulsive disorders, phobias, fixed ideas, and so on. They were discovered by contemporary psychiatrists and recognized as real, that is to say, parasitic psychic organisms independent of conscious human will and tending to control it.
\end{quotex}


In alchemical and magical texts, these elemental spirits are salamanders, sylphs, undines, and gnomes corresponding to the elements fire, air, water, and earth. The elemental spirits are in bondage to the elements. As such, they are not free and cannot have the use of reason; they can only manifest in subhuman forms as animals. \textbf{Eliphas Levi} describes their qualities:

\begin{quotex}
Elemental spirits are like children: they torment chiefly those who trouble about them ... they frequently occasion our bizarre or disturbing dreams and produce psychic phenomena . They can manifest no thought other than our own and when we are not thinking they speak to us with the incoherence of dreams. They reproduce good and evil indifferently, for they are without free will and are hence irresponsible... Such creatures are neither damned nor guilty; they are curious and innocent.
\end{quotex}


\paragraph{Planetary Spirits}


The planetary spirits obviously relate to the planets, but for our purposes, are actually the pagan gods. Psychologists will claim that the gods were projections of the human unconscious. However, the case is just the opposite. The gods exist and manifest themselves in consciousness. In Degrees of Existence\footnote{\url{https://gornahoor.net/?p=8618}}, we pointed out how \textbf{Johann Gichtel} and \textbf{Oscar Hinze} demonstrated that the planets correspond to certain centers or chakras in the human being. That is how they can be experienced.

\textbf{Carl Jung} recognized that the gods, which he called archetypes, have an independent existence. However, he did not deign to describe that existence. However, Anne Catherine gives us a more detailed description:

\begin{quotex}
Great order reigns also among the planetary spirits, who are fallen spirits, but not devils. They are very, very different from devils. They go to and fro between the earth and the nine spheres. In one of these spheres they are sad and melancholy; in another, impetuous and violent; in a third, light and giddy; in a fourth, stingy, parsimonious, miserly, etc.

They exert an influence over the whole earth, over every man from his birth, and they form certain orders and associations. In the planets I saw forms resembling plants and trees, but light and unsubstantial, like mushrooms. There are, also, waters on them, some clear as crystal, others muddy and poisonous; and it seemed to me that each planet contains a metal. The spirits make use of fruits adapted to their own nature. Some are an occasion of good, inasmuch as man himself directs their influence to good. Not all the heavenly bodies are inhabited; some are only gardens or storehouses for certain fruits and influences. I see places in which there are souls who, although not Christian, yet led good lives on earth. They are now in uncertainty, feeling that some day or other their lot will change; they are without joy or pain. Like the others, they feed upon certain fruits.
\end{quotex}


That explains the fickleness of the gods and goddesses in European myths. The descriptions of the postmortem states are similar to ideas like Valhalla.

\paragraph{Unfinished Business}


The angelic hierarchy will require a separate post in the near future. The notion that God is fundamentally Will is elucidated on the article on Jacob Boehme\footnote{\url{https://gornahoor.net/?p=8750}}.


\flrightit{Posted on 2018-03-22 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{james on 2018-03-23 at 04:34 said:}

Thanks for another great post.

\hfill

\texttt{Blanchard on 2018-03-23 at 11:21 said:}

I didn't realize Emmerich wrote on this subject. Very illuminating exposition of Elementals and Planetary spirits. Is there any meaningful distinction to be made between Elementals and Demons, or are they the same type of being? If so, where is the distinction to be made?

\hfill

\texttt{geirón on 2018-03-26 at 01:50 said:}

Blanchard, « any assertion to the effect that two things are different implies that they are reductively the same » --- which is to say, Demons and Elementals are the same so far as they have the same qualities and different so far as they have different qualities.

\hfill

\texttt{David on 2018-05-08 at 15:57 said:}

Did you ever realize the diagrams from Symbolism of the Cross?

\hfill

\texttt{Jonathan on 2018-06-07 at 13:07 said:}

«The thought of being eaten by an animal like a shark, alligator, or lion, has always horrified me. Obviously, the animal is neither evil nor immoral. The horror arises from its complete indifference to the qualities of the human being who is nothing but protein to it.»

There is no such complete indifference in the animal kingdom, Sir. Apparently not even in reptiles: I recall that one of the Buddha's disciples had been bitten by a snake; the Buddha told him that this would not have occurred if he had but radiated loving compassion throughout the serpentine kingdom. Evil men may attack you regardless of your sanctity, as the example of Christ and countless martyrs should testify to. Wild beasts, on the contrary, will not attack someone in whom they sense the presence of the spiritual authority of their Lord, and I do not think you have to be a great saint for this principle to apply. However, those in whom the spiritual presence is not immutable may fall into a negative state of fear and so on upon the shock of encountering a dangerous wild animal, such as a bear or tiger, and this might possibly lead to an attack, as one's present spiritual state is of some importance in this regard. A state of complete openness to and trust in God's Grace while fearlessly radiating love will avert any attack from a wild beast. He who practices continually spiritual awareness (for example supported by the invocatory prayer of the heart) whilst in a place where there is a risk of such attacks should have little to fear. For a great master or saint, moreover, perfect sovereignty over wild animals may be possible, as the example of several historical personages should demonstrate (St. Francis is one who immediately comes to mind). This might perhaps have something to do with the Primordial Man's knowledge of the Names of all animals, being their angelic archetypes, which is also their word or vibrating spiritual energy, through which a higher communication with the animal kingdom is possible; and I would add, this is the only essential and inward science of what an animal in reality IS, all the dissective and reductionist pseudo-knowledge of modern science being null and void in comparison. The reason that virtually nobody practice this science is that the intense contemplation required to succeed is better directed to the `one thing needful'; the hour is late, the days are evil, and time is shrinking if we are honest with ourselves. Before concluding, I'd also have to suggest that attack by a wild animal would ultimately be an outward symbolical representation that is synchronous with a corresponding psychic force within the soul that is unleashed beyond control with fateful repercussions. The end that a person meets is never a random occurrence by chance. It speaks a deep language of its own. The animal merely responds and makes himself an instrument of the forces at work.

\end{sffamily}
\end{footnotesize}

