\section{Ascent to the Divine}

\begin{quotex}
What more liberal and more fruitful provision could God have made in regard to Sacred Scriptures than that the same words may be understood in several senses, all of which are sanctioned by the concurring testimony of other passages equally divine.
\flright{\textsc{St Augustine}, \emph{De Doctrina Christiana}}

\begin{verse}
Now I a fourfold vision see,\\ And a fourfold vision is given to me;\\ `Tis fourfold in my supreme delight\\ And threefold in soft Beulah's night\\ And twofold Always. May God us keep\\ From Single vision \& Newton's sleep!
\end{verse}
\flright{\textsc{William Blake}, \emph{Letter to Butts}}
\end{quotex}

The ancient sources all recognized the multivalent interpretation of scriptures. That all changed with the Protestant Reformation. For example, \textbf{Martin} \textbf{Luther} sought the ``single, simple, solid, and stable meaning'' of Scripture. \textbf{John Calvin} seconded the idea: ``the true meaning of Scripture is the natural and obvious meaning.'' This single vision also infected the liberal theologians with their fruitless, century long search for the ``historical Jesus''.

The first level, usually incorrectly named the ``literal interpretation'', is the material, historical space-time interpretation. The esoteric interpretation, on the other hand, deals with the experience in consciousness. Unlike some Gnostic or New Age systems, we don't claim that the esoteric is the ``real'' interpretation. \emph{A fortiori}, however, neither is the material space-time interpretation any more real.

\paragraph{Interiority}

\begin{quotex}
Do not go outside into yourself, the truth lives deep within.
\flright{\textsc{St. Augustine}}
\end{quotex}

On October 1, 2017, Stephen Paddock shot up a concert from a hotel window, killing 58 people and wounding close to one thousand. Those are the facts, so the investigation could have ended right there. Yet, that is unsatisfying because the real questions are ignored: what was the man's motive and what is the meaning of the event.

Hence the FBI took 16 months to conclude the investigation, inconclusively. Note, however, that the FBI did not try to explain the event in terms of the motions of atoms or Newton's laws of force. Nor did they analyze his brain to determine which neurons were firing. Still less did they fabricate a genetic, evolutionary scenario.

Had they done so, everything would have been puzzled and assumed that the FBI was way off-track. Nevertheless ``intelligent'' men today insist that all explanations can be traced back to atoms, neurons, and genes. That is probably because few people bother with self-observation, so they fail to develop the ability to observe their interior life carefully and dispassionately.

The point is that the mere accumulation of facts does not achieve the quest for meaning. The first task is to describe the differences between events in consciousness and events in the world. Just as space and time are preconditions for any experience in the world, so, too, are they required in consciousness. However, the experience is quite different.

Mental space is determined by thoughts and images. Our dreamlife provides obvious examples of this. In a dream, we might envision a scene in a city, which then shifts to a beach scene. That is a change of space. In the material world, time is required to move from the city to the beach, whereas in the dream the change is nearly immediate.

However, in full consciousness, when one is attempting to deliberately shift from one thought to another, say in meditation or concentration exercises, there may be a significant inner time lag. Hence, inner time depends on the intensity of the desire to move from one thought to another.

Of course, this may be difficult to understand if you allow thoughts and images to spontaneously arise in consciousness, while making no effort to monitor and filter such thoughts and images.

\paragraph{Christ the King}

As an example, we can consider the idea of Christ the King that implies that Christ has dominion over the world. Then, at the end of time, He will establish His Kingdom on earth. However, esoterically, this means that the birth of Christ in the soul will take domination interiorly rather the exteriorly. \textbf{Ananda Coomaraswamy} illustrates this point clearly:

\begin{quotex}
The concept of Victory is of the utmost importance in the traditional theory of Kingship. Exoterically it is by an actual or implied victory over others that a King obtains the throne, but esoterically he is the true Victor who subdues his own passions, allying himself with the Self against himself. The heroism expected of the Knight, whether as King or as the Mortal Soul and Outer Man, is then no longer a matter of merely physical courage, but a symbol and evidence of self-conquest and self-knowledge; autonomy, as we have seen, being the outward tally of an inward Self-control. Whoever has thus found Himself is necessarily both fearless and invulnerable.
\flright{\textsc{Ananda Coomaraswamy}, \emph{Spiritual Authority and Temporal Power in the Indian Theory of Government}}
\end{quotex}

\paragraph{Augustine's Ladder}

In \emph{De Quantitate Animae}, Augustine described seven stages in the ascent to the divine. What is striking is the similarity to the same stages in Hermetic teaching. \textbf{Vladimir Solovyov}, in
\emph{Lectures on Divine Humanity}, mentioned in passing the connection
between the writings of \textbf{Hermes Trismegistus} and the Christian Fathers. Since he never fully developed that idea, it is our task to complete his unfinished work.

This relationship, therefore, is quite ancient, and did not suddenly arise in Renaissance Florence. \textbf{Marsilio Ficino} identifies those ancient sources\footnote{\url{https://web.archive.org/web/20230212233109/https://www.meditationsonthetarot.com/heart-of-the-west}}.

These are Augustine's stages, followed by a commentary. It is important to keep in mind that the stages do not represent changes in consciousness of a stable I. On the contrary, each stage means that the sense of I is centred at that stage. The task, therefore, is to move the sense of I upward to higher stages.

\begin{enumerate}

\item \textbf{Animation}: The soul is life-force without consciousness, in common with the plants.

\item \textbf{Sentience}: in common with the animals

\item \textbf{Rational}: but not moral, i.e., worldly wisdom.

\item \textbf{Virtue}: divine wisdom purges the soul.

\item \textbf{Ataraxia}: complete trust in God

\item \textbf{Intuition}: Direct, unmediated experiences of higher truth

\item \textbf{Union with God}
\end{enumerate}

This is a vertical ascent and thus is radically different from horizontal schemas like Ken Wilber's colour system. Such schemas are arbitrary and tendentious.

\paragraph{Natural Life}

The first three stages are aspects of the sublunar sphere. Clearly, they correspond to the three souls described by Aristotle and the Medieval Scholastics: the Nutritive soul (plants), the Sensitive soul (all animals), and the Rational, or intellectual, soul.

As such, they apply to human beings as they are born into the world. Although all people experience animation, sentience, and rationality, everyone tends to be primarily focused in one of the first three states.

Those centred in the first, or sensory stage, is dominated by instinct, movement, thrills, sensation-seeking, sex, food, or the physical body. The fundamental fault in such beings is \textbf{concupiscence}, i.e., excessive or disordered desire.

Those centred in the sentient stage are dominated by their emotions. Feelings guide their lives. When the feeling become negative, such as anger, fear, envy, jealousy, etc., the fundamental fault is
\textbf{malice}.

Those centred in the rational or thinking centre are dominated by their intellectual life, in the sense of discursive thought. At its best, this leads to technological advance, starting from fire, the wheel, agriculture, husbandry, metallurgy, navigation, and so on. Its fundamental fault is \textbf{ignorance}. Of course, ignorance is wide ranging and includes irrationality, logical fallacies, just for starters.

The world, such as it is, is dominated by humans at the first three levels. Although there is some natural goodness, concupiscence, malice, and ignorance foment strife and conflict. There is no way out.

\paragraph{Transitional State}

In traditional societies, transcendent influences in the form of exoteric religions break into the natural world system. Those who are open to such transcendent influences may undergo a moral conversion. In this case, he tries to apply brakes against the natural impulses of the lower levels. This produces an inner conflict that is necessary to develop a strong and independent I.

Natural men and women, when exposed to transcendent influences, may resent that inner conflict, preferring instead to live unencumbered by any inner constraints. They may desire to pursue sexual excess or create intellectual systems totally independent of any transcendent influences. Hence, they assume that the source of the inner check is not transcendent, but rather are imposed by their enemies. That is why neopagans, who prefer to live in the moment, may believe that moral norms are Semitic plots.

The fundamental fault at this level is \textbf{weakness}. One may know what is right to do, yet fails due to a weak will.

\paragraph{Ataraxia}

The state of Ataraxia is achieved when the lower emotions come under control, and the will gets purified. This is a state of perfect calm, tranquility, bodily relaxation, and the silencing in the mind of desires, anxiety, fear, vulgar images, obsessive thoughts. While natural men are plagued by lower and vulgar images, in this state the mind receives images from above in the form of symbols. Now a genuine symbol is not merely a representation of a higher thought, but is itself the thing symbolised. This is related to the art of high culture.

\paragraph{Intuition}

When the mind is cleared of discursive thinking, it gains a direct, intuitive knowledge of spiritual things. This is beyond words and even the symbols of the state of Ataraxia.

\paragraph{Union with God}

This is the state reached when the I totally submits to the Absolute I. This is knowledge of the Whole. It is the Beatific Vision, which perhaps can be reached in this life.


\flrightit{Posted on 2019-02-12 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Max on 2019-02-20 at 11:52 said:}

Mentioning the ``historical Jesus'' was definitely timely since I am currently reading Albert Schweitzer's ``The Quest of the Historical Jesus'' which attempts to summarize all those various ``Lives of Jesus'' that have been penned. However, I am also looking into some Rosicrucian tracts in tandem, which demonstrates a much better understanding, for example, the following gem from the ``Rosa Florescens'' is worth far more than the many books filled with empty speculations on the life of Jesus. A short passage proves to pull a heavier weight than centuries of idle words by countless of men.

``Man can do anything by means of God, who has His dwelling-place within him, and he can understand by and from himself any books such as a learned man might write, as they are nothing more than a memorial or testimony through which man is reminded and convinced of what is within him, for we cannot acquire any understanding from the letters alone but must impart our understanding to them. If that was not the case and the Holy Scripture carried the sense itself within it then there would not be so many sects and groups, because the words would mean the same to everyone. But instead everyone interprets them according to his own understanding, so do not place your understanding in the book therefore but in the spirit, and the spirit within the man, for man must act towards the spirit completely passively and not use his own will but must allow God to work within him freely according to all His convenience and delight, for God, like a father of all arts and suitability, yes the eternal and unique wisdom itself, will also illuminate for him his soul and awareness with the godly heavenly understanding, to such an extent that in this life nothing can be too much for him and in the future life nothing can be against him. For `it is the spirit that quickeneth; the flesh profiteth nothing: the words that I speak unto you, they are spirit, and they are life' John 6:63 Now we must go back and seek out the origin of this text, namely the Spirit, and we must listen to it and learn from it on a silent Sabbath. Out of all the books in the world you can learn without especial difficulty from just one single book, and this book is within you and within all men, both great and inferior, both young and old, both learned and unlettered. But there are few -- yes few indeed! -- who can read it. Yes, there are many very learned people who are forced to deny it within themselves, and so cleave to the dead-letter which lies outside them, and so abandon the Book of Life which, even so, is indited with the finger of God in all men's hearts'' -Rosa Florescens (translation Paul Ferguson)

To climb upward the ladder, the sensitive soul must become attached to the rational soul, since it in turn provides the link and incentivize the nutritive soul. If the transitional state as it is described here demands the cultivation of will, it means that the rational soul must descend to rescue the fallen will from the dominance of the surrounding earth, which it may only do by enlisting sentience as its intermediary. When the latter lays dormant, or is impermeable to higher influences, the will proves fickle on its own. The soul in a fallen state is only virtually one, although in ataraxia we might come to know it so in actuality. From the corrupted mixed state, we must learn to separate what is described as events of the world from interior events since ``the spirit quickeneth; the flesh provideth nothing''. When imagination no longer receives the primary impetus from the body and the world, the sensitive soul that produces the images will shift attention towards the spirit.

There is a teaching of the retrieval of the unfallen state, a state of peace in the ``pure sabbath'':

``And this in brief is the view of the Brotherhood, at least until another time is reached -- not that all men should be equal (for the majority are too stubborn and godless) but that the believers who keep a pure sabbath for God will be like Adam in Paradise, for it is precisely this pure knowledge and peace in Christ that Paradise consists of, things that an old theologian warns us to observe but which are not observed. Indeed, the Rosicrucian promise says that whoever is in this Paradise will not only by quite natural means be able o broadcast his opinion clearly and unambiguously in the practice of the light and in the spirit of the universals, across several hundred miles, but also will happily learn from a book all the arts of the world, and how to attract to himself pearls, gold and precious stones, how to fully understand everything within the micro- and macrocosm, how to secretly learn the thoughts of men, just like Solomon could, how to effortlessly prepare the Blessed Stone for his kin and his neighbours' use and, indeed, to understand the whole of nature naked and revealed.'' -Rosa Florescens

We see from the above in what manner the man in Paradise is described as having an archaic consciousness, many of the claims of which would be regarded as quite fantastic by modern men, but relates to coming to know the mental ``space'' that you bring up, which has other properties than the corporeal space that everybody is familiar with. Here is the same idea again:

``Through sympathy one can `travel' hundreds of thousands of miles by land and sea in a perfectly natural manner without boots or writing-implements, and even if someone is located twelve fathoms below sea-level yet another person can make known to him everything that he wants in a twinkling.'' -Rosa Florescens

\hfill

\texttt{Michael M on 2022-02-13 at 16:42 said:}

Bare-chest and bare-footed, he comes out into the market place;
Daubed with mud and ashes, how broadly he smiles!
There is no need for the miraculous power of the gods,
For he touches, and lo! the dead tress are full in bloom.

\end{sffamily}
\end{footnotesize}

