\section{Chapter 3}

\begin{quotex}
\begin{verse}
By esteeming in excess\\ Conflict is provoked in reaction \\ Emphasizing what is rare to possess\\ Creates guilty desire for appropriation\\ It weakens that which, in things, attracts\\ And the soul will remain calm\\ Thus: the True Man acting as leader\\ Moves without preference and appetites\\ He weakens his impulses the man of desire, the physical I , tempers his inner being his bones \\ Without a exterior knowing and without desires the guides the ten thousand beings\\ He confuses those who know\\ He avoids acting centered in the Self \\ And society will always live free in its order
\end{verse}
\end{quotex}


\paragraph{Commentary}


Some translators have rendered the line ``the \emph{sheng jen} without knowing or desiring to guide the people'' with ``the \emph{sheng jen} ensures that the people have neither knowledge nor will''. In order to ignore the symbolic valence of the respective characters, in a modern Chinese translation the ``without preference or appetites'' then becomes ``the real man empties his heart and fills his belly''!!! Along the same line, there are among sinologists those who have even interpreted the rule contained in these lines as ``stupefying the people'' (the 10 thousand beings) on the one hand and, on the other, as ``repressing the most nobly human in heart and will (\emph{hsin} e \emph{chih}) and coaxing instead the most vulgarly bestial part --- belly and bones (\emph{fu} e \emph{ku}).''

The internal logic of the whole chapter corroborates instead the lesson here adopted. Non-knowing and non-desiring (in the sense of discursive knowing and the desiring of the individual) are the traits repeatedly attributed to the True Man of the \emph{Tao Te Ching} which in him are virtues insofar as they would be the opposite in laymen and the common people, that is, insofar as they would be unattainable (if taken in the positive sense, as virtue).

In other chapters it is said, yes to the utility of keeping the people in ignorance that safeguards their original simplicity (something different from ``stupefying''). But for now, it alludes instead to a propitiatory desaturation of size and a distancing of tensions and disordered appetites (first lines of the chapter). As was said, one thinks that with his sole presence the True Man acts in such a direction over the environment, nurturing an ordered spontaneity in social life.

\paragraph{Chinese text and literal translation}


Chapter 3 (\begin{CJK*}{UTF8}{bsmi}第三章\end{CJK*})
\columnratio{0.4}
\begin{paracol}{2}
\begin{verse}
\begin{CJK*}{UTF8}{bsmi}
不尙賢，\\
使民不爭。\\ 不貴難得之貨，\\ 使民不爲盜。\\ 不見可欲，\\ 使民心不亂。\\ 是以聖人之治，\\ 虛其心，\\ 實其腹，\\ 弱其志，\\ 強其骨。\\ 常使民無知無欲，\\ 使夫智者不敢為也。\\ 為無為，\\ 則無不治。
\end{CJK*}
\end{verse}
\switchcolumn
\begin{verse}
Not to quest for wealth\\ will keep the people from rivalry.\\ Not to prize things that are hard to come by\\ will keep the people from becoming thieves.\\ Not to manifest what you like and desire\\ will make popular sentiments undisturbed.\\ So, in the sages' peaceful and tranquil world,\\ People's state of mind is calm,\\ People live with enough food\\ People's aspirations are lowered,\\ People's physiques are strong;\\ People will be unknowing, undesiring,\\ And even the knowing ones will never dare to act,\\ Action without action.\\ There is nothing not done.
\end{verse}
\end{paracol}

\flrightit{Posted on 2020-03-27 by Lao Tzu}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Geber on 2020-03-28 at 02:06 said:}

``Someone who keeps on going from the shade into the sun and then back into the shade is a fool. A wise man stays permanently in the shade.'' Sri Ramana Maharshi

\end{sffamily}
\end{footnotesize}

