\section{The Human Person}

\begin{quotex}
The general plane is unique, but divides itself in diverse parts, in a way that in everything there are distinct places, some more pleasant, others less so. And souls, also unequal, come to inhabit these distinct places that correspond to their own diversity. In such a way, everything is arranged and the differences in their situations correspond to the inequality of the souls. \flright{\textsc{Plotinus}}

It is not the demon that chooses you, but you yourself chose your demon. You yourself choose the destiny of that life to which you are irremediably connected. \flright{\textsc{Plato}} 

\end{quotex}
The human being is in a unique position in the cosmos, since he participates in multiple states of being. Human incarnation requires three factors. He has matter and form, as do all natural things. Hence, he is the unity of a body and human nature. Animal life is also the unity of a body and the specific essence of the animal. But the human being also has an I, he is a person\footnote{This is not the metaphysical definition of Person as used by Guenon. It is actually the human state.}.

At birth, the I is conditioned, that is, he is born into certain circumstances, and given a particular destiny. Matter is the principle of individuation, so he is born a particular sex, of particular parents, family, nation, civilisation, era, and so on. Those are the biological and historical conditions of life. Yet those conditions are not arbitrary, contingent, nor accidental. Rather, they reflect the transcendent qualities of the person, the I. The Traditional teaching is that the soul is the form of the body. Specifically, that implies that the conditions of the body in manifestation are determined, or even chosen, by the soul.

The person participates in multiple states of being. The highest is the Beatific Vision, then the nine angelic states, the primordial state, and the natural (or fallen) state. The angelic states will be addressed in a future post. For the Beatific Vision, refer to the great mystics like Dante, Meister Eckhart, or John of the Cross.

Ultimately, the I (Atma) is unconditioned, neither manifested nor unmanifested.

\paragraph{Person}
The person manifests in three principles.

\begin{itemize}
\item \textbf{Body}: The body exists. This is the principle of individuation since matter creates individuals. 
\item \textbf{Essence}: Essence pertains to what exists. A squirrel has the essence of a squirrel, a human has human essence. 
\item \textbf{Person}: The person is the one who exists. Animals do not have personhood. Humans and angels do have personhood. 
\end{itemize}
The person is conditioned and individuated. As such, personhood has the following qualities:

\begin{itemize}
\item \textbf{Incommunicability}: Unlike essence, it is not shared. Although you can relate to someone because of a shared human essence, you cannot experience the I of that person. 
\item \textbf{Uniqueness}: The person is unique. 
\item \textbf{Unrepeatability}: The person exists once. That is why a person cannot be in the same state more than once. 
\item \textbf{Indivisibility}: The person cannot be divided. 
\item \textbf{Distinctiveness}: Even twins who share the same essence and even the same DNA, they are each a distinct person. 
\end{itemize}
Other characteristics then follow, such as:

\begin{itemize}
\item Self-knowledge 
\item Freedom of choice 
\end{itemize}
\paragraph{Essence and Personality}
The person has characteristics that he is born with and those which he acquires in the course of life. An abstract I without qualities or conditions would not satisfy the qualities mentioned above. In particular, it would neither be unique nor distinct. Hence, we can further refine the understanding of personhood into qualities one is born with and those which one acquires in the course of life:

\begin{itemize}
\item \textbf{Ontological person}: This is the person as such. It is a gift, what one is born with. It is what makes someone one person but not another. A fortiori, it is transcendent to the various states of being. In particular, it is stable and enduring with respect to the empiriological person. 
\item \textbf{Empiriological person}: This is the ontological person as he manifests himself through his acts, powers, and habits. It has two aspects. 

\begin{itemize}
\item \textbf{Personality}: The expression of the person through nonmoral acts. It is cultivated by the formation of physical, intellectual, and volitional habits. 
\item \textbf{Character}: The expression of the person through moral acts. Character is formed by the acquisition of moral virtues. 
\end{itemize}
\end{itemize}
The ontological person is always ``one", although it can take on a number of subpersonalities. Usually, a person is unaware when he switches from one subpersonality to another. In extreme cases, they may not even be aware of each other.

In our deranged state, the empiriological person is not in harmony with the ontological person. The person has a destiny to fulfil. When that is misunderstood, the acts of the person will be in conflict with his essence, resulting in imbalance, unhappiness, and other problems.

In general, strong personalities are esteemed more highly than men of character. This is not necessarily bad, since the strong make the world and the weak often mask themselves as men of character. The unbalanced person has limited ability to judge another, so psychopaths are often mistake as strong.

\paragraph{Destiny}
The person is born incomplete, with unfulfilled possibilities. One's personal task is to actualize those possibilities through the development of personality and character.

Besides traditional sources, the attributes of the person and the definitions of the ontological and empiriological persons are taken from \textit{The Doctrines of Genesis 1-11} by Rev. Victor P Warkulwiz, M.S.S.

\flrightit{Posted on 2020-08-30 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Auld Wat on 2020-08-30 at 18:09 said: }

Thank you for the timely post! I've been struggling to find employment recently. It has however given me time to reflect on my character and my personality. Saints who suffered far more than I ever have, Rose of Lima for example, are testimony of the actualisation of their possibilities through prayer, fasting and alms giving. Hopefully my efforts will not go unnoticed.


\end{sffamily}\end{footnotesize}
