\section{The Creator has a Master Plan}

A long-time supporter of Gornahoor recently posed this question:

\begin{quotex}
Cologero, why god created lucifer, knowing that one day he would turn his back on him. God really likes to see the circus on fire?
\end{quotex}

Rather than provide a vapid response, I decided to go straight to the source. So I DMed Him and waited several days for a response. He woke me up in the middle of a pleasant dream, but, despite my grogginess, I forced myself to awaken. This is the exchange that resulted. (My questions are in \textbf{bold}; the responses are in a normal font.)

\textbf{Before I get to the purpose of this chat, I'd like to clarify something. If you are being itself, how can there be beings other than you? Our best esoterists and philosophers come to this conclusion.}

You like that? Yes, it is my crowning achievement. You just need to realize who or what you really are. Then the mystery dissipates somewhat without, however, ceasing to be a mystery!

\textbf{Let's cut to the chase then. The question arose in regard to evil. If you are Being, then how does an evil entity like Lucifer come to exist?}

Let me start at the beginning. I had in mind a great idea for a world. I started with general concepts and worked down to more specific ideas. I created angels to share these ideas; they became excited about the concepts.

\textbf{I see. That is what we call the Divine Intellect, Universal Intellect, Buddhi, and so on. That is where the ideas or essences reside in a sense. Did we get that right?}

That is more or less on track. But you really need to contemplate them, to ``sink'' into them, as it were. Just verbal knowledge will prevent you from the fullest experience of it.

I you think about it, you will see that it is true.

\textbf{It's nice to contemplate the Master Plan, but that is not yet a world. What was the next step?}

It is not yet a World because it is still just Thought. I needed a way to project it, to experience it directly. So I had to design sense experience. It is just too overwhelming for a single sense. Together, they constituted the passive experience of the world.

In addition, for what I had in mind, I needed to create a way for creatures to interact with the world. It turns out that I only needed a few basic structures which could vary enormously. You call these homologous structures.

\textbf{More great ideas, but still not yet a world.}

That is because I needed a canvas. An artist, who has structures and colors in mind, still requires a canvas in order to visualize his creation and share it with other.

That is why I created matter, which is a total opposite to me. I am pure actuality, uncaused, unmoved, and complete. Matter, on the other hand, is pure potentiality, in other words, nothing in particular. It has the possibility of being but not the actuality, at least not until I project my ideas onto it. So, you see that a nothing becomes something without any self-diminishment on my part.

\textbf{So why do we see multiple copies of an idea?}

Good question! There is only one idea. Once you understand the essence of a Rooster, there is not need for multiple ideas of the Rooster. So I created mathematics, which contains the idea as well as copies. You can understand the idea of a Triangle, and you can also conceive of multiple triangles in the mind.

But you cannot conceive of multiple roosters; for that they need to be materialized. So mathematics is the dividing line between the ideas, or essences, and existence in matter. You understand that matter contains the principle of individuation.

So I created space and time in order to experience the ideas sensually. I see the Rooster in its perfect wholeness, all at once, from its conception to death. You see the same thing unfolding in space and time. Space and time are the analogs of the two traditional branches of maths: geometry and arithmetic.

That can be confusing, I know, when you don't grasp the big picture.

\textbf{I still don't understand why beings are imperfect if the idea is perfect.}

If I made the perfect rooster, it would necessarily be static; it could not change and develop. If I made every rooster exactly alike, they would just be the same rooster. So roosters follow a statistical distribution, e.g., a standard, Poisson, or even a Pareto distribution. The Ancients did not understand that.

In that way, I added a little variety and uncertainty into the world process.

\textbf{You've explained the rooster. What about people?}

Yes, yes. Let's get to that. The angels are purely spiritual beings; roosters, and all the other animals and plants are material beings. What if I could create a being that was both spiritual and material? A being who could grasp the Ideas in spirit, but also the physical world through the senses. That was a breakthrough.

\textbf{Presumably Adam was that being?}

Yes. I had to create the proper corporeal conditions first. Then a suitable animal body had to be prepared. I chose Adam from among them and breathed the Spirit into him. Adam the realized how different he was from the animals he saw in the world. He could no longer feel at home among them, even if some of them had the same outward appearance. So I created a suitable companion for him and gave them a pleasant place to live.

\textbf{The companion was Eve and the place was Eden. What was that experience like for him?}

I gave him what every guy thinks he wants: free and tasty food like an always open buffet. He had meaningful work like tinkering around the garden. And then he had a beautiful naked woman wandering around and available.

\textbf{But why the temptation, if everything was going so well?}

You do understand that Adam's original consciousness, in my ``image and likeness', has no duality. It is all good, Adam experienced no lack. Yet Adam was placed in the middle, so there have to be two forces at play: one upward toward Spirit and the other downward toward the physical world.

The third force equilibrates those two forces; this is precisely that nondual consciousness that I breathed into him. As long as Adam was consciously aware of the spiritual world, he would be OK.

\textbf{But it didn't turn out OK}

As I said, there is the possibility of evil, but not necessarily its actuality. The thing with Lucifer was supposed to be a joke. How did Adam ever fall for such a cheesy temptation? The talking serpent, the attractive fruit, etc.; he should have seen through it.

I didn't teach him ``good and evil''. And how did he even know what ``death'' meant, unless he learned it from watching the animals. But their experience is different. Physical death is just part of the material world. The real issue is spiritual death, which is what Adam experienced. He then judged the world on his own terms: good and bad, easy or difficult, and so on and on.

\textbf{So here we are! What's next?}

There is an entire historical drama about that, which need not be repeated in a DM. Your task is to describe that pure Edenic state of consciousness.

\textbf{Why don't you just tell me?}

It can't be told, it must be experienced. There is the joy of discovery. On Easter Sunday, don't you hide the eggs from the children, or do you leave them out in a bowl? It is the search that is meaningful.

\textbf{One final question. Which came first, the chicken or the egg?}

Sorry, you're breaking up ... I have to go

\url{https://www.youtube.com/watch?v=MUouOwmtVEY} 

\flrightit{Posted on 2021-04-11 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{info on 2021-04-11 at 23:51 said:}

If perfection is static. How is the non-static living God perfect?

\hfill

\texttt{Cologero on 2021-04-12 at 00:09 said:}

Dear lack of info: the topic was the perfect Rooster. I suggest you find the perfect Rooster and report back to us.

Or are you suggesting that God changes? That would be quite an unfortunate opinion to hold.

\hfill

\texttt{info on 2021-04-12 at 09:12 said:}

@Cologero

The premise that perfection is static. But life isn't static means that it doesn't quite work as a definition. God is perfect yet he acts as befits a Living God.

He is remains the same yet he isn't static either.

\hfill

\texttt{Cologero on 2021-04-12 at 09:26 said:}

Get lost. You are here just to debate the meaning of words with me? And miss the effing point.\\ Dictionary definition of static: ``Having no motion''

God is the ``unmoved mover''.

Meditate on that for six months or so. And consider yourself banned from commenting in the meantime.

\hfill

\texttt{MP on 2021-04-13 at 08:06 said:}

Leon possessed a powerful gospel inspired voice. His work on the (redacted) linked tune is probably a bit much for most, recorded as it was during Pharaoh's rather chaotic `free' style, under the influence of John Coltrane and Sun Ra.

Alternately, Thomas' vocal work on Sanders' tune, Oh Lord Let Me Do No Wrong, is first rate, and would be a good place for a novice to start.

For a while Leon associated with Santana's spiritually oriented jazzish pop group--the time Carlos was hanging with John McLaughlin (during the latter's first Mahavishnu Orchestra phase). With `fitness' guru Chinmoy Ghose. Back then it was important for musicians to follow gurus. Not sure that remains the case, but I'm out of the music scene so don't really know how it is, today.

\hfill

\texttt{Cologero on 2021-04-13 at 13:34 said:}

Wow, MP, I did not suspect that there are any jazz fans left, especially now when any saccharin instrumental music is called ``jazz''.

For obvious reasons, I did not link to the full version with the long solo at the end.

I considered the live version of the song by Santana and McLaughlin. But their version is hardly reverent enough. It seems more self-serving just to show how `with it' they are.

\hfill

\texttt{Angolmois on 2021-04-21 at 09:05 said:}

Lucifer himself is not evil, but the luciferic principle makes evil possible since he is the principle of mind that makes possible both the ascension and descension of man by making possible the discernment between good and evil, and some unfortunate souls choose evil consciously even after they know good and evil (possibly because of a-causal reasons), becoming ``a disaster for the human race''.

Lucifer and Christ are not opposed, since Lucifer is the bearer of light and Christ the light itself. The crucified savior is Lucifer/Ego. Jesus called himself ``the bright morning star'' which is a direct reference to Lucifer. The battle between good and evil is fought within man's soul but one cannot blame Lucifer for making it possible, since without him man would be ``perfect but unconscious and without merit'', a mere useless instrument in the hands of the Gods.

Evil exists within humanity and we have all sinned somewhere along the path, because it is the school of good and evil in which man must ultimately triumph and choose the good willingly.It is man who is to blame for the ``sin of pride'', and this is what changes Lucifer into a Satan. The real blame is on man conscious of his choices.

In the Finnish epic Kalevala, Ilmarinen (literally ``airy'', again a reference to Lucifer) is the forger of the vault of Heaven and the axis mundi that holds existence in one piece, Son of Morning and the Nightfall (Venus) and he is also the helper of Wäinämöinen, who is literally an old covenant Christ figure (Melkhisedek in one of his guises).

\end{sffamily}
\end{footnotesize}

