\section{In Defense of Descartes}

\begin{quotex}
I think therefore I am.
\flright{\textsc{Rene Descartes}}

There is something godlike, then, in personhood, or better said, in a person, whoever he or she may be. Only one needs to bear in mind that the person is not in fact the human individual, as we commonly suppose. Individuality, one might say, is no more than a mask woven, as it were, of attributes which belong in truth, not to the person as such, but to his nature.
\flright{\textsc{Wolfgang Smith}, \emph{Theistic Evolution}}
\end{quotex}

When I was a boy, around 10 or 11, my parents would drop me off at the Library of the Boston Museum of Science, while they went shopping at the nearby Lechmere Sales. My favorite books were about Astronomy and Mathematics. I used to copy the names of distant stars and galaxies into a notebook. Among my favorite books were the four volumes of the series, \emph{The World of Mathematics}, by \textbf{James R Newman}. Its mathematical symbols and diagrams were the entrance to a mysterious inner world, defined by its apodictic proofs, unlike the uncertainty and confusion of the material world. Moreover, the blue hardcover version was a delight to hold, unlike my paperback edition.

One thing that struck me about Descartes, and I've remembered it my entire life, was that he was allowed to lie in bed in the morning. He retained that habit throughout his life as he believed it was ``conducive to intellectual profit and comfort.'' I've never had that luxury until the last dozen years or so. Since I am now blessed with insomnia, I awaken early in the morning. I just simply lie still with nothing but my thoughts and prayers. Every now and then a revelation comes to me and a post is born.

\paragraph{Descartes and the Poet}
Descartes was, for me, the epitome of the logical mind, who inextricably linked philosophy and mathematics together. Years later, I learned that Descartes was among the creators of the modern mind and implicitly an enemy of Tradition.

That view was challenged a year ago, when the poet \textbf{Charles Peguy} asserted that Rene Descartes was among the three greatest philosophers (along with Plato and Henri Bergson). That has been an irritant to me ever since, not unlike a grain of sand inside an oyster.

It dawned on me, during a morning repose, that the poet speaks from Intuition, not rational thinking. So that is how Descartes must be approached.

\paragraph{Descartes and Intuition}

Although the superiority of Intuition over rational argumentation has always been an essential aspect of traditional metaphysics, it is not always the case in practice. In this view, Descartes did not create a ``new'' philosophy to replace the old. Rather, he was describing his own consciousness. He admits as much in his meditations, since he proposes some self-evident truths that can simply be ``seen'', which is what is meant by intellectual intuition.

\paragraph{Modern Philosophy}

First of all, we need to dispatch the notion that Descartes somehow ``created'' modern philosophy, or better, the ``modern mind''. Obviously, neither your next-door neighbor nor any of your friends used to think like a medieval, but suddenly got ``woke'' after reading Descartes. The more likely scenario is that he was describing his inner experience through self-observation. What he observed in this way was different from the self-understanding of his medieval predecessors. Universities teach the history of philosophy or the history of ideas, but they don't teach the history of consciousness. It is a mistake to believe that consciousness is static from age to age; rather consciousness changes over time. That is why and how an earlier age becomes opaque to a later one.

He claims that there are two substances that make up the human: extension and thought, because the body and mind are distinct. It would better to say matter and thought, since extension is not considered a property of manner; hence it belongs to thought.

The Thomist objection is that man is indeed a unified being, and thus cannot be composed of two distinct substances. That's a good point but the solution is hardly better, which is that the human being is a substance with two properties, body and soul. But the body dies and the soul lives on. So the postmortem soul is only an incomplete man. Although the complete teaching is out in the open, it is seldom noticed.

The essence of a human being is to be human, an animal, a living being, and a corporeal object. These manifest as a body, a living soul, an animal soul, and an intellectual soul. So the original ``soul'' is not so simple. Let's take Jacques Maritain's example of Peter. Peter is composed of matter and essence, specifically human essence, and everything that implies. An essence is in the Divine Intellect and can only be known by the intellect, not by the senses. Descartes does not stray too from that, since he agrees that only the intellect can understand.

\paragraph{The Person}

But what about Peter? Where is he? According to Maritain, a substance has a suppositum or subject (of change), which, in Peter's case, is a Person. There is nothing in the essence of ``human nature'' that requires a person, something Maritain does not notice. Rather, he assumes that the individuated Peter has a self. Rene Guenon describes this confusion:

\begin{quotex}
The personality, metaphysically speaking, has nothing in common with what modern philosophers so often call `the human person', which is, in fact, nothing but the individuality pure and simple; besides, it is this alone and not the personality which can strictly be called human.
\flright{\textsc{René Guenon}, \emph{Man and his becoming according to the Vedanta}}
\end{quotex}

Maritain is therefore more modern than he believes, although he does admit that the Person Peter is also an essence in the Divine Intellect, so Peter is more than just a human substance. Now we are in a position to better understand Descartes. Through his method of introspection, which is not so unlike that of Ramana Maharshi, he realizes that he is the ``I'', neither body nor soul, but a Self that transcends them both. So both Descartes and Maritain are knocking on the door, behind which, Guenon was already there:

\begin{quotex}
The personality, let us insist once more, belongs essentially to the order of principles in the strictest sense of the word, that is, to the universal order; it cannot therefore be considered from any point of view except that of pure metaphysics, which has precisely the Universal for its domain.
\flright{\textsc{René Guenon}, \emph{Man and his becoming according to the Vedanta}}
\end{quotex}

\paragraph{States of Being}

Certain consequences follow from this view. The first is that the ``real'' Peter is the personality, the Self in the strict sense. This transcends the ego of the human substance.

The second consequence solves the debate about the body and soul as substances or essences. The body and the various souls that comprise the human being are just states of being of the Self. So death does not affect the real Peter, because it is one of Peter's possible states of being.


\flrightit{Posted on 2021-09-15 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Karl on 2021-09-16 at 06:03 said:}

Ed Feser had a recent interesting blog post on common misconceptions of Cartesian matter that ties in with some of the above.
\url{http://edwardfeser.blogspot.com/2021/09/make-believe-matter.html}

\hfill

\texttt{Cologero on 2021-09-16 at 12:53 said:}

Thanks, that was timely.

But I will point out where Feser is mistaken, not because I think so, but because he does not take into account a fundamental point of medieval scholasticism. See \textit{Phenomenology of the Medieval Mind}\footnote{\url{https://www.gornahoor.net/?p=3559}}.

``Common sense'' used to mean the faculty of the soul to put together all the data experienced by the senses into a coherent representation of the world in the imagination. That means that the ``colors'' are not ``Out there, right now'', but rather they exist in the soul life as the representation of the physical world.

This is analogous to the way that the Intellect knows the essence of what is ``out there''. That is, we don't ``see'' the fox out there, we grasp the fox by understanding its essence.

BTW, this teaching was common to the Middle Ages, as even Islamic metaphysicians had the same understanding. They still do, since that tradition is still alive. Feser is trying, but he has lost full connection with it.

I have a post in the pipeline that will explain how the world comes to be represented in the soul.


\hfill

\end{sffamily}
\end{footnotesize}

