\section{Adam’s Choice}

\begin{quotex}
The IMAGINATION ... I consider either as primary, or secondary. The primary IMAGINATION I hold to be the living Power and prime Agent of all human Perception, and as a repetition in the finite mind of the eternal act of creation in the infinite I AM.
\flright{\textsc{Samuel Coleridge}, \emph{Biographia Literatura}}

What would be needed is a world where socialization and specialization would no longer rip away from each soul its individuality, its spontaneous perception of the life of things, and the religious sense of the beauty of beings. It would be a world where Love would have to precede all Knowledge, a world where the sense of death would be only nostalgia for resurrection.
\flright{\textsc{Henry Corbin}, \emph{Eternal Sophia}}

I was a hidden Treasure. I desired to be known; that is why I created creatures.
\flright{\textsc{Hadith}}
\end{quotex}

Adam, the Primordial Man, was just below the angels, yet could contemplate God's word. Through Intuition, he could see God's plan, but somewhat imperfectly. But Adam was the only human and God wanted more to share in His plan. Thus He proposed to let Adam incarnate in a body of clay. Adam balked and was resistant to the idea. Since God needed Adam's permission, He offered Adam a preview of what Earth life would be like.

\paragraph{The Plan for Adam}

God showed Adam the Divine Soul which gave him the gift of Imagination. He filled Adam's Imagination with sensual imagery derived from the five outer senses of hearing, sight, touch, taste, and smell. Suddenly the ideas known to Adam in the Divine Intellect took on a new aspect. Not only did he know the ideas, but he could now experience them in his mind.

Adam was astounded to experience the ideas, not only spiritually in his intellect, but also in his soul. The world of nature was no longer an abstract thought, he could now experience it. The fields, forests, seas, mountains, flowers, and fauna were displayed in a kaleidoscope of color, the multitudinous sounds of the Earth pounded his soul.

Adam was intrigued especially when he watched the animals move about. God explained that Adam would receive the same powers. He would be given not only the means of locomotion but the power of grasping, in a fashion much superior to that of animals. He would be given the gift of speech; this would allow him to not just hear the Word but also speak the word. Most important was the gift of creation. He would be able to create in the world, in imitation of God. He would till the fields, explore the forests, sail the seas, climb the mountains, and make use of the flora and fauna. Even better, Adam would be able to create beings like himself through reproduction.

\paragraph{The Angels Revolt}

The angels, with astonishment, listened intently to God's Word to Adam. Gazing deeply into God's plan, they protested that Adam and his progeny, with their power of choice, would only create mischief and shed blood.

God said to them, ``You do not see deeply enough. You are appointed to guide humanity and serve them.'' But some angels resented Adam's gift, and refused to serve men. Having devised their own plan, many of them rebelled against God and were banished from the Heavenly Realms. Instead of being of help to men, they would become tempters.

\paragraph{The Creation of Eve}

``How can I create beings like myself?'' asked Adam.

``I will first split you in two,'' God replied.

Adam then fell into the causal state of deep sleep while God split him. When he awoke, God showed Eve to Adam. Adam was overwhelmed at the experience of his soul. He knew the Idea of Beauty, but the experience of Eve's beauty was overwhelming. His love for Eve rivalled his intellectual love for God.

God revealed to them that they are to be together for eternity as their lives are forever entangled. Even in their darkest moments, in times of despair, when they seem separated, they are together and must seek each other. From their spiritual marriage, they will find their true Self.

For the nonce, Adam and Even lived in the Garden that God had prepared for them. There was abundance and ease, and they remained aware of the presence of God.

\paragraph{Warning to Adam}

God gave a warning to Adam. ``Eat and do as you will, but do not eat from the Tree of the Knowledge of Good and Evil. There will come a time when you will be tempted, but remember my command. Otherwise, you will forget me. Life apart from Me is death.''

``I will never forget you, my Lord. I accept your command but the warning is unnecessary,'' promised Adam.

\paragraph{Adam's Fall}

Unbeknown to them, Satan, the leader of the rebel angels, was waiting for them in the Garden. Satan took advantage of their desire for a sensual experience. Satan convinced their soul, Eve, that the taste of the fruit of the forbidden tree provided great pleasure.

For the first time, Adam experienced doubt and fell into darkness. He forgot his origin and his promise. Together they ate, but their delight at sensual experience released them from their innocence.

They fell totally into earth bodies, which became their new center of awareness. They lived among the animals, and now had to struggle for their nourishment. They became alienated not only from God, but from each other.

They must find each other, as God told them, so that through their spiritual marriage a new Self will be born.


\flrightit{Posted on 2021-08-11 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{S.P. on 2021-08-18 at 13:18 said:}

In spiritual love the passion of the flesh is sublimated and, in a way, consumed by the very fire of its love; Eventually, not only is the body conceived as a transparent expression of the soul, ennobled and embellished by its almost Apollonian light, but the passion is spiritualized until it transcends the very desire of the body, like a burning love, a kind of ``melted fire'' . In the sacred union man desires only God.

Once you understand it, not as rational information, but as that understanding that is living experience, then life reveals itself. This is a path that every man should discover, even if most will not finish it in this lifetime. For my part, I appreciate these posts.


\hfill

\end{sffamily}
\end{footnotesize}

