\section{Principle and Action 22}

\begin{quotex}
We continue today the translation of Evola's 1959 edition and commentary on the \emph{Tao Te Ching}, published under the title \emph{Il Libro del Principio e della sua Azione} The Book of the Principle and its Action .
\end{quotex}

\paragraph{Chapter 22}
\begin{verse}
To be whole in the fragment\\ Straight in the curved\\ Full in the void\\ Intact in corruption.\\ Gathered, one succeeds\\ Scattered, one fails\\ Therefore: the True Man preserves the original One\\ He thus becomes a model for the world\\ He does not put himself in the light, and so he shines\\ By not affirming himself, he imposes himself\\ Free of complacency, he becomes great\\ Because he does not contend\\ No one can attack him.\\ ``To be whole in the fragment'',\\ This ancient saying,\\ Could it be meaningless?\\ No: in the Way it is the truth.
\end{verse}

\paragraph{Commentary}

The saying from the first line, repeated in the third-to-last --- \emph{ch'u tse ch'üan} --- can be interpreted both in a normative sense (``to be whole in the fragment'') and in an ontological sense (the wholeness of the fragment). Ontologically, it means that in the Way the fragment (the part) is the whole (just as the curved is the straight, the empty is the full, etc.). It is the idea of the ``partial perfections'' that, in their own way, reflect the Principle.

It must not be understood (as various translators have done) that the fragment will become the whole, etc., materially, through transformation or development; rather, it already is the whole, if seen in the light of the Way.

On the other hand, in the normative sense, that maxim (``to be whole in the fragment'') refers to a specific Far Eastern concept of perfection: to carry out each thing precisely and perfectly, with a natural concentration and without effort. Then, whatever activity is performed, one is in the Way.

Cf. in Chuang-tzu (XIX, 3): ``Unify the intentions; have only one, which merges with vital energy'' (cf. XIX, 10, 12). There is a connection between this idea and the one already explained, of the ``original simplicity,'' even in its ``magical'' aspect.

\paragraph{Chinese text and literal translation}
Chapter 22 (\begin{CJK*}{UTF8}{bsmi}第二十二章\end{CJK*})\columnratio{.4}
\begin{paracol}{2}
\begin{verse}
\begin{CJK*}{UTF8}{bsmi}
曲則全，\\ 枉則直，\\ 窪則盈，\\ 敝則新，\\ 少則得，\\ 多則惑。

是以聖人抱一為天下式。

不自見，故明；\\ 不自是，故彰；\\ 不自伐，故有功；\\ 不自矜，故長。

夫唯不爭，故天下莫能與之爭。

古之所謂「曲則全」者，豈虛言哉？誠全而歸之。
\end{CJK*}
\end{verse}\switchcolumn
\begin{verse}
Bent, therefore whole.\\ Crooked, therefore straight.\\ Hollowed, therefore full.\\ Worn, therefore new.\\ Few, therefore gain.\\ Many, therefore confused.

Therefore, the Sage embraces the One and becomes the model for the world.

He does not display himself --- therefore he is seen clearly.\\ He does not affirm himself , therefore he shines.\\ He does not boast, therefore he accomplishes.\\ He does not pride himself, therefore he endures.

It is precisely because he does not contend,\\ that none in the world can contend with him.

The ancients who said, ``Bent, therefore whole'',\\ were these empty words?\\ No, they are truly whole and return to it.
\end{verse} 
\end{paracol}

\flrightit{Posted on 2025-04-13 by Lao Tzu}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Juliano on 2025-04-14 at 10:35 said:}

``To carry out each thing precisely and perfectly, with a natural concentration and without effort'' -- precisely the First Arcanum from Tomberg. I wonder if more people here is working with the recent book ``The Taoist Alchemy of Wang Liping'' from Nathan Brine, since Charles used to recommend ``Opening the Dragon Gate'' and Brine was authorized by Wang Liping to expose his inner alchemy teachings. I'm working with book one.

\hfill

\texttt{Tannheuser on 2025-04-14 at 22:14 said:}

Exactly, Juliano. You'll also notice that Chuang-tzu's ``Unify the intentions; have only one, which merges with vital energy'' is identical to Tomberg's ``moral purification of the will'' (also in Letter 1), or Cologero's favorite saying that ``purity of heart is to will one thing''.

I haven't read either of the books you mention, although perhaps I will soon. You might be interested that Joscelyn Godwin is publishing a translation of Matgioi's ``La Voi Metaphysique'' early next year:
\url{https://www.innertraditions.com/the-primordial-tradition-of-ancient-china}

\hfill

\texttt{Juliano on 2025-04-15 at 08:06 said:}

Thanks, Tannheuser, I will certainly read Matgioi's book when it is available. Also beware of the new translation of the brazilian masterpiece \textit{Grande Sertão: Veredas} (The Devil to Pay in the Backlands, now translated as Vastlands: The Crossing) by the genius João Guimarães Rosa. It is a masterful synthesis of Christian, taoist and alchemichal transformation in literary form. The old transaltion is quite bad, but this one looks good. It should be available next year.

\url{https://pug-lynx-fgnj.squarespace.com/news/alison-entrekins-new-guimares-rosa-translation-sold-in-pair-of-auctions}

\hfill

\texttt{Ginette on 2025-05-15 at 13:54 said:}

Hello, great article! Could you tell me the source of the literal translation? Is it a book? Thanks.

\hfill

\texttt{Tannheuser on 2025-05-18 at 15:51 said:}

For the literal translation, I am using a number of sources (for example, \url{https://en.wikisource.org/wiki/Translation:Tao_Te_Ching} ), and applying some of my own limited knowledge to clean it up, also. I'm doing my best in this case since I am uncertain of which source Charles was using --- the main value is having the original Chinese characters available here for those that can read them -- as you can see, there are many different directions once can take when translating these to English. Evola himself published a very different translation and commentary on the Tao in the 1920s that was more focused on interpreting it through the lens of his philosophy of Magical Idealism, which Gornahoor will make available in the future.

\end{sffamily}
\end{footnotesize}

