\section{Week 10: Lead us not into Temptation}

Humanity is faced with three temptations, just as Jesus was tempted in the wilderness.

\begin{enumerate}
\item Materialism, or the temptation to turn stones into bread. This is the understanding of the universe as Being, not the universe of the Good or Love. Dead matter somehow creates life and consciousness, not to mention spiritual experience. In popular films and TV shows, we see computerized robots turn into sentient beings. 
\item The desire to plunge into the subconscious as the source of spiritual life. This ignores all higher forces, the superconscious and divine revelation as the true source of religion. 
\item The denial of causality attributes events to chance or to random variation. This absolves human beings from the responsibility of their own future. 
\end{enumerate}

The chakras, or lotus flowers, lie at the foundation of the human being. Temptation is the wrong unfolding of the lotus flowers. Christ's transfiguration illumines the chakras. The I AM sayings correspond to different chakra.


The Old Testament describes the six days of creation culminating in the Old Adam.

The New Testament begins with the description of the birth and growth of the new humanity, the New Adam.


\paragraph{Day 1: The Temptation in Paradise}

\subparagraph{First Fall}
\textbf{I AM:} the resurrection and the life

\textbf{Reading:} Genesis chapter 3

\textbf{Chakra:} Crown Chakra


The temptation in Paradise was the fall into death. The 8-petalled lotus, the Crown chakra, is Union with God. But in Eden this union was destroyed and the Crown chakra represented death.

The temptation first came to Eve, the psychic and astral human element. The intellectual element, or the ``I" then succumbed to the temptation. Hence, egoism entered into the stream of humanity.


\subparagraph{Transfiguration}

\begin{quotex}
Jesus said: I am the resurrection, and the life: he that believeth in me, though he were dead, yet shall he live. (John 11:25)


And it came to pass about eight days after these sayings, he took Peter and John and James, and went up into a mountain to pray. And as he prayed, the fashion of his countenance was altered, and his raiment was white and glistering. And, behold, there talked with him two men, which were Moses and Elias: Who appeared in glory, and spake of his decease which he should accomplish at Jerusalem. But Peter and they that were with him were heavy with sleep: and when they were awake, they saw his glory, and the two men that stood with him.


And it came to pass, as they departed from him, Peter said unto Jesus, Master, it is good for us to be here: and let us make three tabernacles; one for thee, and one for Moses, and one for Elias: not knowing what he said. While he thus spake, there came a cloud, and overshadowed them: and they feared as they entered into the cloud. And there came a voice out of the cloud, saying, This is my beloved Son: hear him. (Luke 9:28-36) 

\end{quotex}
Christ radiated Divine Love into human nature on Mount Tabor.


\paragraph{Day 2: The Sons of God}

\subparagraph{Second Fall}

\textbf{I AM:} the light of the world

\textbf{Reading:} Genesis chapter 6:1-6


\begin{quotex}
And it came to pass, when men began to multiply on the face of the earth, and daughters were born unto them. That the sons of God saw the daughters of men that they were attractive; and they took them wives of all which they chose. And the Lord said, My spirit shall not always strive with man, for that he also is flesh: yet his days shall be an hundred and twenty years. There were giants in the earth in those days; and also after that, when the sons of God came in unto the daughters of men, and they bare children to them, the same became mighty men which were of old, men of renown. And God saw that the wickedness of man was great in the earth, and that every imagination of the thoughts of his heart was only evil continually. and it repented the Lord that he had made man on the earth, and it grieved him at his heart. 

\end{quotex}

\textbf{Chakra:} Ajna Chakra, the two-petalled lotus flower.


The sons of God were the righteous ones. They were attracted to the daughters of men, i.e., the profane who were strangers to God. The effect is that they lost the double vision, i.e., they did not see the vertical dimension, but only the horizontal, which came from the outside. In other words, they judged solely by appearances, and not by spiritual qualities or essences.

The distorted the two-petaled lotus flower of the head, or Ajna, chakra.

The distorted Ajna chakra, the seat of the intellect, no longer reflects the light, but lives in darkness and tries to understand the world without reference to higher things.


\subparagraph{Transfiguration}

\begin{quotex}
Then spake Jesus again unto them, saying, I am the light of the world: he that followeth me shall not walk in darkness, but shall have the light of life. (John 8:12) 

\end{quotex}
The Master opens up the Ajna chakra to the light of the fire principle of Love, so that the intellect can reflect it. Vertical vision from above is restored so that the essential can be distinguished from the inessential.


\paragraph{Day 3: The Tower of Babel}

\subparagraph{Third Fall}

\textbf{I AM:} the good shepherd

\textbf{Reading:} Genesis 11:1-9


\begin{quotex}
And the whole earth was of one language, and of one speech. And it came to pass, as they journeyed from the east, that they found a plain in the land of Shinar; and they dwelt there. And they said one to another, Go to, let us make brick, and burn them thoroughly. And they had brick for stone, and slime had they for morter. And they said, Go to, let us build us a city and a tower, whose top may reach unto heaven; and let us make us a name, lest we be scattered abroad upon the face of the whole earth.

And the Lord came down to see the city and the tower, which the children of men builded. And the Lord said, Behold, the people is one, and they have all one language; and this they begin to do: and now nothing will be restrained from them, which they have imagined to do. Go to, let us go down, and there confound their language, that they may not understand one another's speech. So the Lord scattered them abroad from thence upon the face of all the earth: and they left off to build the city.

Therefore is the name of it called Babel; because the Lord did there confound the language of all the earth: and from thence did the Lord scatter them abroad upon the face of all the earth. 

\end{quotex}

\textbf{Chakra:} Throat Chakra


The 16-petalled Throat Chakra is associated with speech. The hubris of the human beings resulted in its destruction. Hence, The universal language of humanity became confounded into multiple languages.


The primal, total breathing of humanity corresponded to the original, universal language of humanity. Speech is the formation of the breath into sound. The loss of the single universal human language after the building of the tower of Babel signified the disappearance of the primal, total breathing, which became individual and arbitrary. And the Whitsun miracle, where the apostles spoke a single language, which was understood by human beings from different lands as if it were their own, was a resurrection of the primal breathing and the original language of humanity.

The original language spoken prior to the building of the tower of Babel was also total. It was more than a mere communication of information. Rather, it was a kind of thought transmission (telepathy), which had the effect of awakening insight, instead of simply informing. It served to awaken insight and understanding, rather than explaining things. It communicated an intimation of things. Not that the intimations were indefinite, having multiple meanings that had to be explained.


No, they were like rays or bolts of lightning bearing light and clarity within them. Both the speaker and the listener partook of this light, which came from the depths of the heart and penetrated into the depths of the heart— a kind of ``heart magic." This was also so with the speaking of the apostles at Whitsun, which was experienced by all those from different countries as their mother tongue. For what could be closer to each human being than the language of the heart?

At the root of the building of the tower of Babel is the collective will of ``lower selves" to achieve the replacing of the ``true Self of the celestial hierarchies and God with a superstructure of universal significance fabricated through this will.

The building of the tower of Babel had as its effect the ``thunderbolt" of the ``descent of the Lord", who ``confused their language" — that of the builders — and ``scattered them abroad over the face of all the earth", so that they would no longer understand one another's language.

He who rebels against his ``higher Self will no longer live under the law of the vertical but. rather under that of the horizontal, i.e. he will be ``a fugitive and a wanderer on the earth." (Lazarus come forth)


\subparagraph{Transfiguration}

\begin{quotex}
I am the good shepherd, and know my sheep, and am known of mine. (John 10:14)

\end{quotex}
Christ heals the Throat Chakra when all the sheep from different folds hear his voice and understand.

And other sheep I have, which are not of this fold: them also I must bring, and they shall hear my voice; and there shall be one fold, and one shepherd.

All those who were confused and scattered are joined together.

This was confirmed at Whitsun when the disciples all understood each other's language.


\begin{quotex}
At Pentecost his disciples realised the universal and eternal truth of the identity of all individual ``selves" in the higher universal Self—which is manifested by the fact that their speech became that of the higher Self of each person who listened to them. (MotT)

\end{quotex}

\paragraph{Day 4: Changing Stones into Bread}

\subparagraph{Fourth Fall}

\textbf{I AM:} the bread of life

\textbf{Reading:} Luke 4:1-4, Matthew 4:1-4


\begin{quotex}
And Jesus being full of the Holy Ghost returned from Jordan, and was led by the Spirit into the wilderness, Being forty days tempted of the devil. And in those days he did eat nothing: and when they were ended, he afterward hungered.  And the devil said unto him, If thou be the Son of God, command this stone that it be made bread. And Jesus answered him, saying, It is written, That man shall not live by bread alone, but by every word of God. 

\end{quotex}

\textbf{Chakra:} Heart Chakra, the twelve-petaled lotus flower


The ``heart" that we have in mind here is not that of emotion and the faculty of being passionate that one generally understands by ``heart". This centre is the most human of all the centres of ``lotus flowers"; it is the chakra of love.

At the center of the seven chakras, it is the bridge between Heaven above and the Earth below.

\subparagraph{Temptation}

The temptation to turn stones into bread manifests in humanity as the temptation to explain:


\begin{itemize}
\item quality by means of quantity 
\item the living by means of the dead or mechanical 
\item to fashion qualitative ``living" things out of quantitative mechanical ones 
\end{itemize}

The temptation to be nourished by ``stones", i.e., dead things, destroyed the heart chakra.


\subparagraph{Transfiguration}

Chakra work consisted in the attempt to awaken the chakras as they are.

Our task now is the Christianisation of all the centres, i.e. their transformation in conformity with their divine-human prototypes.

The work of Christianisation of the human organisation, i.e. the transformation of the human being into a man of heart, is accomplished in the inner life of man, the lotus-flowers being only the field where the effects of this purely inner work are manifested.

Rather than by stones, we need to be nourished by the bread of life.

\paragraph{Day 5: Descending into the Unconscious}

\subparagraph{Fifth Fall}

\textbf{I AM:} the door.

\textbf{Reading:} Luke 4:9-13, Matthew 4:5-7, Luke 9:28-36, John 10:7-16

\begin{quotex}
And he brought him to Jerusalem, and set him on a pinnacle of the temple, and said unto him, If thou be the Son of God, cast thyself down from hence: For it is written, He shall give his angels charge over thee, to keep thee: And in their hands they shall bear thee up, lest at any time thou dash thy foot against a stone. And Jesus answering said unto him, It is said, Thou shalt not tempt the Lord thy God. And when the devil had ended all the temptation, he departed from him for a season. 

\end{quotex}

\textbf{Chakra:} Solar Plexus Chakra, the ten-petaled lotus flower


The three lower chakras are organs of horizontal perception, what St. Theresa called ``imaginary vision".

As the ten-petaled lotus underlies the senses, it is also the chakra of science, which studies the world revealed to the senses.


\subparagraph{Temptation}

The second temptation in the wilderness is to succumb to the instinctual life, that is, to plummet down from the spiritual heights to the unconscious.


\subparagraph{Transfiguration}

Light drives out darkness. This simple truth is the practical key to the problem of how to combat demons. A demon perceived, i.e. on whom the light of consciousness is thrown, is already a demon rendered impotent.


\begin{quotationx}
And when he went forth to land, there met him out of the city a certain man, which had devils long time, and ware no clothes, neither abode in any house, but in the tombs. When he saw Jesus, he cried out, and fell down before him, and with a loud voice said, What have I to do with thee, Jesus, thou Son of God most high? I beseech thee, torment me not. (For he had commanded the unclean spirit to come out of the man. For oftentimes it had caught him: and he was kept bound with chains and in fetters; and he brake the bands, and was driven of the devil into the wilderness.)

And Jesus asked him, saying, What is thy name? And he said, Legion: because many devils were entered into him. And they besought him that he would not command them to go out into the deep. And there was there a herd of many swine feeding on the mountain: and they besought him that he would suffer them to enter into them. And he suffered them. Then went the devils out of the man, and entered into the swine: and the herd ran violently down a steep place into the lake, and were choked.

When they that fed them saw what was done, they fled, and went and told it in the city and in the country. Then they went out to see what was done; and came to Jesus, and found the man, out of whom the devils were departed, sitting at the feet of Jesus, clothed, and in his right mind: and they were afraid. They also which saw it told them by what means he that was possessed of the devils was healed. 

\end{quotationx}
Here we see that Christ not only drives out the demons but makes them visible in the herd of swine. That is, the demons had dwelled in the unconscious, but exposing them to the Light of Christ brought the man to his right mind.


In MotT, Tomberg explains:


\begin{quotex}
Modern depth psychology has discovered and put into practice the therapeutic principle of bringing unconscious complexes to the light of consciousness. Because—so it affirms—the light of consciousness renders the obsessional complex not only visible but also impotent. This important discovery of modern psychology is in complete accord with the spiritual reality of the ``struggle" of the celestial hierarchies against evil. Because this ``struggle", also, amounts to their presence alone, i.e. to bringing evil to the light of day. 

\end{quotex}

\paragraph{Day 6: The Will to Power}

\subparagraph{Sixth Fall}


\textbf{I AM:} the way, the truth and the life.

\textbf{Reading:} Luke 4:5-8, Matthew 4:8-10, Luke 9:28-36, John 14:6


\begin{quotex}
And the devil, taking him up into an high mountain, shewed unto him all the kingdoms of the world in a moment of time. And the devil said unto him, All this power will I give thee, and the glory of them: for that is delivered unto me; and to whomsoever I will I give it. If thou therefore wilt worship me, all shall be thine. And Jesus answered and said unto him, Get thee behind me, Satan: for it is written, Thou shalt worship the Lord thy God, and him only shalt thou serve. 

\end{quotex}

\textbf{Chakra:} Sacral Chakra, the six-petaled lotus flower

The chakra is the center of Harmony and Health.

Under the influence of the Heart Chakra, the six-petalled lotus will become that of holiness, i.e. harmony between spirit, soul and body. One of the powers obtained from meditating on this chakra is eloquence and clarity, in words and thoughts.


\subparagraph{Temptation}

The harmony of spirit, soul, and body is also the harmony of the physical, astral, and spiritual world. The temptation to power in the physical, material world destroys the sacral chakra by separating the being from the higher worlds.


\subparagraph{Transfiguration}

Christ's words heal this chakra: I AM


\begin{itemize}
\item The Way: Heals the Body 
\item The Truth: Heals the Soul 
\item The Life: Heals the Spirit 
\end{itemize}

This saying results in inner harmony. In the future, as the spiritual penetrates more deeply into the physical world, humanity will become healthier.


\paragraph{Day 7: The Night of Gethsemane}

\subparagraph{Seventh Fall}

\textbf{I AM:} the true vine.


\textbf{Reading:} Luke 22:39-46, Matthew 26:36-46, Mark 14:32-42 Luke 9:28-36, John 15:1-9


\begin{quotex}
And they said, Lord, behold, here are two swords. And he said unto them, It is enough. And he came out, and went, as he was wont, to the mount of Olives; and his disciples also followed him. And when he was at the place, he said unto them, Pray that ye enter not into temptation. And he was withdrawn from them about a stone's cast, and kneeled down, and prayed, Saying, Father, if thou be willing, remove this cup from me: nevertheless, not my will, but thine, be done. And there appeared an angel unto him from heaven, strengthening him. And being in an agony he prayed more earnestly: and his sweat was as it were great drops of blood falling down to the ground. And when he rose up from prayer, and was come to his disciples, he found them sleeping for sorrow, And said unto them, Why sleep ye? rise and pray, lest ye enter into temptation. 

\end{quotex}

\textbf{Chakra:} Root Chakra, the four-petaled lotus flower

The center of creative force


\subparagraph{Temptation}

The apostles slept while Jesus prayed. They descended into the unconscious in which was the center of their being. The root chakra wants to be autonomous.

Unchastity is nothing other than the autonomy of carnal desire, so that the wholeness of the human being— in spirit, soul and body—is ruined. Sexual desire is only one aspect of love —the aspect reflected by that part of the physical and psychic organism which is the special domain of the ``four-petalled lotus"—and it constitutes only one-seventh of the human psychic-physical organism. (from MotT)


\subparagraph{Transfiguration}

The human being is fundamentally what his heart is. When he is awakened, he is centered in the heart. The man of heart is a man who has regained his humanity. The three lower chakras will be aligned to the heart.


Then the creative force of the four-petalled lotus will then serve as a source of energy and inexhaustible elan.

\flrightit{Posted by Cologero}
