\section{Week 11: Deliver Us from Evil}


\textbf{Meditation:} The 7 Words from the Cross

\textbf{Crucifixion:} The humanization, or awakening of the chakras, such as they are

\textbf{Illumination:} The Christianisation of the chakras


Among the effects of this illumination is \textit{certainty}: (from MotT)

\begin{itemize}
\item The three higher lotuses are those of infused certainty or ``imperceptible light", and these are the principal instruments of ``intellectual vision" or trans-subjective revelation. 
\item The three lower lotuses are those of certainty of experience at first hand: they render us ``eye-witnesses" to invisible things. They manifest them to us in ``perceptible light" as forms, movements, colours, sounds and breath — concrete and objective — although incorporeal in the sense of the physical world. 
\item The central centre, the heart or twelve-petalled lotus, gives us the certainty of authentic faith.
\end{itemize}

\textbf{Evil:} The chakras are not in harmony. Thus they work independently and the lower chakras try to dominate consciousness and the higher chakras are unaware of their transcendence source. The human being is fundamentally what his heart is, and the heart is not oriented to the divine. 


The great work of spiritual alchemy is the transmutation of the substances (``metals") of the other lotuses into the substance of the heart (``gold").


\paragraph{Day 1: I commend my spirit}

\textbf{Words:} Father, into Thy hands, I commend my Spirit.

\textbf{Reading:} Luke 23:46, Mark 15:37


\begin{quotex}
And when Jesus had cried with a loud voice, he said, \textit{Father, into thy hands I commend my spirit}: and having said thus, he gave up the ghost. 

\end{quotex}
\textbf{Chakra:} Crown Chakra, the eight-petaled lotus flower

The eight-petaled lotus is the centre of wisdom and is used by the Spirit.

Undeveloped, it is abstract.

When the Heart is developed, it will become full of warmth, like the fire of Pentecost.

To the extent that concentration in the crown chakra takes place, vital activity diminishes — at first in the lower region of the organism, i.e. the genital and intestinal region, then in the region of the stomach, and lastly in the central region in the neighbourhood of the heart. At the moment when the concentration of vitality at the crown centre is complete, the heart and also the circulatory and respiratory system cease their activity — \textbf{this is the moment of death.}


\paragraph{Day 2: Why hast thou forsaken me}

\textbf{Words:} My God, my god, Why hast thou forsaken me?

\textbf{Reading:} Matthew 27:45-46, Mark 15:34


\begin{quotex}
Now from the sixth hour there was darkness over all the land unto the ninth hour. And about the ninth hour Jesus cried with a loud voice, saying, Eli, Eli, lama sabachthani? that is to say, \textit{My God, my God, why hast thou forsaken me?}

\end{quotex}

\textbf{Chakra:} Ajna Chakra, the two-petalled lotus flower

The two-petalled lotus is the centre of intellectual initiative and concentration.

Undeveloped, the two-petalled lotus remains at the disposal of the freedom of human arbitrariness.

The intellectual initiative of the two-petalled lotus will become ``compassion-filled insight" into the world when it is taken hold of by the Spirit united to the human spirit.

But when he is abandoned and forsaken, the human stands alone with human arbitrariness.


\begin{quotex}
He wished publicly to declare the great suffering of His Passion. Wherefore these words: ``My God, why hast Thou abandoned Me?" are not words of one who accuses, or who reproaches, or who complains, but, as I have said, they are the words of One who declares the greatness of His suffering for the best of reasons, and at the most opportune of moments. \flright{\textsc{Robert Bellarmine}}

\end{quotex}

\paragraph{Day 3: I thirst}

\textbf{Words:} I thirst

\textbf{Reading:} John 19:28-29, Psalm 69:21


\begin{quotex}
After this, Jesus knowing that all things were now accomplished, that the scripture might be fulfilled, saith, I thirst. Now there was set a vessel full of vinegar: and they filled a sponge with vinegar, and put it upon hyssop, and put it to his mouth.

They gave me also gall for my meat; and in my thirst they gave me vinegar to drink. 

\end{quotex}
\textbf{Chakra:} Throat Chakra, the sixteen-petalled lotus flower

The sixteen-petalled lotus is the centre of the creative word.

The creative word of the sixteen-petalled lotus will become magical: it will have the faculty of illumining, consoling and healing.


\begin{quotex}
That will be the result of the fulfillment of the scripture. Now, the essence of vertical respiration is the alternation between prayer and benediction or grace. These two elements of vertical respiration manifest themselves in all domains of the inner life —mind, heart and will. Thus a relevant problem for the mind, which is not due to curiosity or intellectual collectionism, but rather to the thirst for truth, is fundamentally a prayer. And the illumination by which it may be followed is the corresponding benediction or grace. True suffering, also, is fundamentally always a prayer. And the consolation, peace and joy which can follow are the effects of the benediction or grace corresponding to it. 

\flright{\textsc{from MotT}}

\end{quotex}

\paragraph{Day 4: Today shalt thou be with me in paradise}
\textbf{Words:} Verily I say unto thee, today shalt thou be with me in paradise

\textbf{Reading:} Luke 23:41-43


\begin{quotex}
We receive the due reward of our deeds: but this man hath done nothing amiss. And he said unto Jesus, \textit{Lord, remember me when thou comest into thy kingdom.} And Jesus said unto him, \textit{Verily I say unto thee, today shalt thou be with me in paradise.}

\end{quotex}

\textbf{Chakra:} Heart Chakra, the twelve-petalled lotus flower

\begin{quotex}
The foundation of the heart is love. The heart lives only when it loves. It is then like the sun.

Chastity is the state of the human being where the twelve-petalled lotus is awakened and becomes the sun of the microcosmic ``planetary system". The three lotus centres situated below it begin then to function in conformity with the life of the heart (the twelve-petalled lotus), i.e. ``according to solar law".

The ``heart" that we have in mind here is not that of emotion and the faculty of being passionate that one generally understands by ``heart". It is the middle centre of the seven centres of man's psychic and vital constitution. It is the " twelve-petalled lotus". This centre is the most human of all the centres of ``lotus flowers". The twelve-petalled lotus (the heart centre) is that of love.

The ``heart" is the centre or lotus of love — where, truth to tell, it is no longer a question of ``above" or ``outside" or even of ``above" and ``below", because love abolishes all distances and all distinctions of space (even distinctions of spiritual space) and has the power to make all things present. It is thus that God is present in a heart glowing with love.

The heart or twelve-petalled lotus, gives us the certainty of authentic faith, which was born in the ``fire of Emmaus". \flright{\textsc{from MotT}}

\end{quotex}
Each thief had a choice: ``Love existence, and you have chosen heaven; hate it, and there you have chosen hell."

The good thief was centred in the Heart.

The magic of the five wounds operates by the presence of the reality of the superhuman spiritual world by means of the wounds, and accomplishes the transmutation from the states of ``limbo" and ``purgatory" to the state of union with the Divine, or ``paradise".


\paragraph{Day 5: Father forgive them}

\textbf{Words:} Father, forgive them; for they know not what they do.

\textbf{Reading:} Luke 23:32-34


\begin{quotex}
And there were also two other malefactors, led with him to be put to death. And when they were come to the place, which is called Calvary, there they crucified him, and the malefactors, one on the right hand, and the other on the left. Then said Jesus, \textit{Father, forgive them; for they know not what they do.}

\end{quotex}

\textbf{Chakra:} Solar Plexus Chakra, the ten-petalled lotus flower

The ten-petalled lotus is the centre of science. Science tends to be self-contained without reference to anything transcendental. This must be forgiven.

When oriented to the heart chakra, the ten-petalled lotus will then become conscience, i.e. the servant of God and neighbor.


\paragraph{Day 6: Behold thy son! Behold thy mother}

\textbf{Words:} Woman, behold thy son! Behold thy mother.

\textbf{Reading:} John 19:26-27


\begin{quotex}
When Jesus therefore saw his mother, and the disciple standing by, whom he loved, he saith unto his mother, \textit{Woman, behold thy son!} Then saith he to the disciple, \textit{Behold thy mother!} And from that hour that disciple took her unto his own home.

\end{quotex}
\textbf{Chakra:} Sacral Chakra, the six-petalled lotus flower

The six-petalled lotus is the centre of harmony and health.

Under the influence of the heart, the six-petalled lotus will become that of holiness, i.e. harmony between spirit, soul and body.


\paragraph{Day 7: It is consummated}

\textbf{Words:} It is consummated.

\textbf{Reading:} John 19:30


\begin{quotex}
Jesus therefore, when he had taken the vinegar, said: \textit{It is consummated}: and bowing his head, he gave up the ghost. 

\end{quotex}
\textbf{Chakra:} Root Chakra, the four-petalled lotus flower

The four-petalled lotus is the centre of creative force.

Under the influence of the heart, the creative force of the four-petalled lotus will then serve as a source of energy and inexhaustible elan.

But that requires that the soul understands what it means that the great work of salvation has been  accomplished. Otherwise, they live without love and repentance.

The three beasts are neutralized:

\begin{itemize}
\item The pig or ignorance 
\item The snake or malice 
\item The rooster or concupiscence 
\end{itemize}

\begin{quotex}
Then I spoke My last word: ``It is finished." Yes, it was accomplished, the great work of salvation. But how much benefit has it been for the many thousands and thousands of souls, even when, through My death and mediation, they also were delivered from the original sin (but did not follow Me in spirit and in deeds)? Heaven was opened for them, but through their sinfulness, living without love and repentance, they drew again eternal damnation to themselves.

Children of men, as My first and last word I say to you again in the full earnest of My love:

Repent. Return to your Lord and God in word and deed. Stop your usury and remember your poor brothers who beg for your mercy in vain. Remember the widows and orphans. And speak righteously to those who are immature.

For it is written: ``With the measure with which you measure, so will you also be measured." Let the fate of former generations be a warning to you. As long as they remained faithful to God they were successful and happy, but when they began only to trust themselves God turned away from those nations, and whole kingdoms were wiped out from the face of the Earth. \flright{\textit{Revealed to Jakob Lorber in 1863.}}
\end{quotex}

\flrightit{Posted by cologero}
