\section{Week 6: Thy Will be done}

Meditation on God's blessings and judgment.

\begin{tabular}{ll}\toprule
\textbf{Blessing} &
\textbf{Judgment}\\\midrule
The seeds in the Garden &
The seeds of the Earth\\
Man shall not be alone &
Cain must wander alone\\
The four rivers of the Garden &
The Flood\\
Adam names the animals &
The Tower of Babel: God confuses human languages\\\bottomrule
\end{tabular}
 

\subsection*{Day 1: First blessing of God}
The blessing of eating the seeds in the Garden, i.e., spiritual nourishment, changes into the curse of feeding on plant
matter

\begin{quotationx}
The third day of creation is the “day” of procreation, of the mystery of seed and of growth. The third day of creation
is the generation of the seed principle, of the principle of potential formative force becoming actualized and bringing
to visible realization its own inner, invisible shape. The third day of creation is the coming into being of the seed
principle in the world, that is, not only of the plant-world manifested to us as plants and trees, but also such
“trees” growing in paradise as “the tree of knowledge of good and evil” and the “tree of life”. \begin{flushright} \emph{Covenant of
the Heart}\end{flushright}

\end{quotationx}

\textsc{Mt 7:16-18:} You will know them by their fruits. Are grapes gathered from thorns, or figs from thistles? So, every sound tree bears
good fruit, but the bad tree bears evil fruit. A sound tree cannot bear evil fruit, nor can a bad tree Dear good fruit.

\textbf{First Blessing of God: Genesis 1:26-31}

And he said: Let us make man to our image and likeness: and let him have dominion over the fishes of the sea, and the
fowls of the air, and the beasts, and the whole earth, and every creeping creature that moveth upon the earth. And God
created man to his own image: to the image of God he created him: male and female he created them. And God blessed
them, saying: Increase and multiply, and fill the earth, and subdue it, and rule over the fishes of the sea, and the
fowls of the air, and all living creatures that move upon the earth. And God said: Behold I have given you every herb
bearing seed upon the earth, and all trees that have in themselves seed of their own kind, to be your meat: And to all
beasts of the earth, and to every fowl of the air, and to all that move upon the earth, and wherein there is life, that
they may have to feed upon. And it was so done. And God saw all the things that he had made, and they were very good.
And the evening and morning were the sixth day.

\textbf{First Judgment of God: Genesis 3:18}

Thorns and thistles shall it bring forth to thee; and thou shalt eat the herbs of the earth.

\subsection*{Day 2: Second blessing of God}
\begin{quotationx}
Revolt, possession, and substitution of the fabricated for the revealed — with these three sins
there correspond three “falls” and effects entailed by them. Cain, who killed his brother Abel, became an exile
— he became a wanderer. \begin{flushright} \emph{Meditations on the Tarot. Letter XVI: The Tower of Destruction}\end{flushright}

\end{quotationx}

\textbf{Second Blessing of God: Genesis 1:27, 2:18-25}

God gave the human being Adam another human being, Eve, so that man would not be alone.

\begin{quotationx}
He who finds silence in the solitude of concentration without effort, is never alone. He never bears alone the weights
that he has to carry; the forces of heaven, the forces from on high, are there taking part from now on. \begin{flushright}
\emph{Meditations on the Tarot. Letter I: The Magician}\end{flushright}

\end{quotationx}

\textbf{Second Judgment of God: Genesis 4:1-16}

Cain murders another human being, so Cain must wander alone again.

\subsection*{Day 3: Third blessing of God}
\begin{quotationx}
In the world there are therefore two different kinds of arriving at a conviction: one can be illumined by the serene
clarity of contemplation, or one can be swept away by an electrifying flood of passionate arguments aiming at a desired
end. The faith of the illuminated is full of tolerance, patience and calm steadfastness — like
crystal; the faith of those who are swept away is, in contrast, fanatical, agitated and aggressive
— in order to live it needs conquests without end. because it is conquest alone which keeps it
alive. The faith of those who are swept away is greedy for success, this being its reason for existence, its criterion
and its motivating force. \begin{flushright} \emph{Meditations on the Tarot. Letter XI: Force}\end{flushright}

\end{quotationx}

\textbf{Third Blessing of God: Genesis 2:4-17}

God gave to the human being the Garden of Eden and its four rivers, or ether streams, as a dwelling place.

\begin{quotationx}
Such are the “trees” of the garden that we cultivate and maintain, i.e., the mysteries of union —
mystical, gnostic, magical and Hermetic — of that which is below with that which is above. For
mysticism, gnosis, magic and Hermetic science are the four branches of the “river” which flows out of our garden of
Eden “to water”—the “river” which “divided and became four rivers”. \begin{flushright} \emph{Meditations on the Tarot. Letter XVI: The Tower of Destruction}\end{flushright}

\end{quotationx}

\textbf{Third Judgment of God: Genesis 6:11-22, 7:20-24}

In anger, God sends the Flood in order to destroy the corrupted human being.

\subsection*{Day 4: Fourth blessing of God}
\textbf{Fourth Blessing of God: Genesis 2:19-20}

God gives human beings the word: nature was just as human beings named it.

\begin{quotationx}
Let us consider the following passage in the Bible, where it says: “And out of the ground the Lord God (Elohim) formed
every beast of the field and every bird of the air; and He brought them to Adam to see what he would name them; and
whatever Adam called every living creature, that was its name”. This does not mean that the human being invented
designations for the beings of the animal kingdom, and even less that he classified them according to genera and
species — along the lines of the system of Linnaeus — but rather that he
received and fulfilled the divine mandate to determine the vocations or missions relative to the human being of the
living creatures hierarchically subordinate to him. For, in the language of the Bible, “name” means “mission” or
“activity of being”, and “naming” is the magical act of determining the vocation or mission of a being. \begin{flushright}
\emph{Covenant of the Heart}\end{flushright}

\end{quotationx}
\textbf{Fourth Judgment of God: Genesis 11:1-9}

God confuses human language (Tower of Babel).

\begin{quotationx}
At the root of the building of the tower of Babel is the collective will of “lower selves” to achieve the replacing of
the “true Self of the celestial hierarchies and God with a superstructure of universal significance fabricated through
this will. The building of the tower of Babel had as its effect the “thunderbolt” of the “descent of the Lord”, who
“confused their language”— that of the builders —and “scattered them abroad
over the face of all the earth” so that they would no longer understand one another's language.
\begin{flushright} \emph{Meditations on the Tarot. Letter XVI: The Tower of Destruction}\end{flushright}

\end{quotationx}

\flright{\small\textit{Posted on 2023-03-19 by Cologero}}