\section{Week 5: Thy Kingdom Come}

The awakened will of the human being calls upon Christ for penance. The stages of the Passion are compensation for human
guilt.

The Transgressions of man as prefiguration of Christ's sufferings in the passion:

\begin{tabular}{ll}\toprule
\textbf{Transgressions of Man} &
\textbf{Stages of the passion}\\\midrule
The \textbf{Fall} of man into sin &
Washing of the feet\\
The curse against \textbf{Eve} &
Scourging\\
The curse against \textbf{Adam} &
Crowning with thorns\\
The fratricide between \textbf{Cain} and \textbf{Abel} &
Carrying of the cross\\
\textbf{Cain} becomes homeless &
Crucifixion\\
Redemption of sins through the \textbf{Flood} &
Entombment\\
The experience of the end of the \textbf{Flood} &
Resurrection\\\bottomrule
\end{tabular}

\begin{quotationx}
Christian meditation pursues the aim of deepening the two divine revelations: holy scripture and the creation, but it
does so above all with a view to awakening a more complete consciousness and appreciation of Jesus
Christ's work of redemption. For this reason it culminates in the contemplation of the seven
stages of the Passion: the washing of the feet, the scourging, the crowning with the crown of thorns, the way of the
cross, the crucifixion, the laying in the tomb, and the resurrection. \begin{flushright} \emph{Meditations on the Tarot. Letter XXI: The Fool}\end{flushright}

\end{quotationx}

\subsection*{Day 1: First stage of the passion: Washing of the feet}
The washing of the Feet is the \emph{humiliation} through which the human being must pass because he submitted to the
temptation “to be as gods”.

\begin{quotationx}
True progress, true evolution, is the advance of beings from life under one law to life under another law, i.e., the
structural change of beings. It is thus that the law “an eye for an eye, and a tooth for a tooth” is in the process of
being gradually replaced by the law of forgiveness. It is thus again that the law “the weak serve the strong, the
people serve the king, the disciple serves the master” will one day give way to the law shown by the Master through the
act of the \textbf{Washing of the Feet}. According to this higher law, it is the strong who serve the weak, the king
who serves the people, the master who serves the disciple — just as it is in heaven, where Angels
serve human beings. Archangels serve Angels and men. Principalities serve Archangels, Angels and human beings, and so
on. And God? He serves all beings without exception. 
\begin{flushright} \textit{Meditations on the Tarot. Letter IX: The Hermit}\end{flushright}

\end{quotationx}

\textbf{Washing of the Feet (John 13:1-20)}

Before the festival day of the pasch, Jesus knowing that his hour was come, that he should pass out of this world to the
Father: having loved his own who were in the world, he loved them unto the end. And when supper was done, (the devil
having now put into the heart of Judas Iscariot, the son of Simon, to betray him). Knowing that the Father had given
him all things into his hands, and that he came from God, and goeth to God; He riseth from supper, and layeth aside his
garments, and having taken a towel, girded himself. After that, he putteth water into a basin, and began to wash the
feet of the disciples, and to wipe them with the towel wherewith he was girded. He cometh therefore to Simon Peter. And
Peter saith to him: Lord, dost thou wash my feet?

Jesus answered, and said to him: What I do thou knowest not now; but thou shalt know hereafter. Peter saith to him: Thou
shalt never wash my feet. Jesus answered him: If I wash thee not, thou shalt have no part with me. Simon Peter saith to
him: Lord, not only my feet, but also my hands and my head. Jesus saith to him: He that is washed, needeth not but to
wash his feet, but is clean wholly. And you are clean, but not all. For he knew who he was that would betray him;
therefore he said: You are not all clean. Then after he had washed their feet, and taken his garments, being set down
again, he said to them: Know you what I have done to you? You call me Master, and Lord; and you say well, for so I am.
\textbf{If then I being your Lord and Master, have washed your feet; you also ought to wash one
another's feet. For I have given you an example, that as I have done to you, so you do also.}

Amen, amen I say to you: The servant is not greater than his lord; neither is the apostle greater than he that sent him.
If you know these things, you shall be blessed if you do them. I speak not of you all: I know whom I have chosen. But
that the scripture may be fulfilled: He that eateth bread with me, shall lift up his heel against me. At present I tell
you, before it come to pass: that when it shall come to pass, you may believe that I am he. Amen, amen I say to you, he
that receiveth whomsoever I send, receiveth me; and he that receiveth me, receiveth him that sent me.

\textbf{Temptation in Paradise (Gen 3:1-12)}

Now the serpent was more subtle than any of the beasts of the earth which the Lord God had made. And he said to the
woman: Why hath God commanded you, that you should not eat of every tree of paradise? And the woman answered him,
saying: Of the fruit of the trees that are in paradise we do eat: But of the fruit of the tree which is in the midst of
paradise, God hath commanded us that we should not eat; and that we should not touch it, lest perhaps we die.
\textbf{And the serpent said to the woman: No, you shall not die the death. For God doth know that in what day soever
you shall eat thereof, your eyes shall be opened: and you shall be as Gods, knowing good and evil.}

And the woman saw that the tree was good to eat, and fair to the eyes, and delightful to behold: and she took of the
fruit thereof, and did eat, and gave to her husband who did eat. And the eyes of them both were opened: and when they
perceived themselves to be naked, they sewed together fig leaves, and made themselves aprons. And when they heard the
voice of the Lord God walking in paradise at the afternoon air, Adam and his wife hid themselves from the face of the
Lord God, amidst the trees of paradise. And the Lord God called Adam, and said to him: Where art thou? And he said: I
heard thy voice in paradise; and I was afraid, because I was naked, and I hid myself. And he said to him: And who hath
told thee that thou wast naked, but that thou hast eaten of the tree whereof I commanded thee that thou shouldst not
eat? And Adam said: The woman, whom thou gavest me to be my companion, gave me of the tree, and I did eat.

\subsection*{Day 2: Second stage of the passion: The scourging}
The scourging is to endure, to bear pain without resisting. The bearing of pain was handed to Eve as a curse.

\textbf{Scourging (John 19:1); Striking on the check (John 18:22, John 19:3)}

Then therefore, Pilate took Jesus, and scourged him.

When he had said these things, one of the servants standing by, gave Jesus a blow, saying: Answerest thou the high
priest so?

They came to him, and said: Hail, king of the Jews; and they gave him blows.

\textbf{The curse against Eve (Gen 3:16)}

To the woman also he said: I will multiply thy sorrows, and thy conceptions: in sorrow shalt thou bring forth children,
and thou shalt be under thy husband's power, and he shall have dominion over thee.

\subsection*{Day 3: Third stage of the passion: The crowning with thorns}
Toilsome labor is Adam's curse. Earthly man must deal with thistles and thorns.

\begin{quotationx}
Every crown is essentially a crown of thorns. Not only is it heavy, but also it calls for a painful restraint with
regard to the thought and free or arbitrary imagination of the personality. It certainly emits rays outwards, but these
same rays become thorns for the personality within. They play the role of nails piercing and crucifying each thought or
image of the personal imagination. \begin{flushright} \textit{Meditations on the Tarot. Letter IV: The Emperor} \end{flushright}

\end{quotationx}

\textbf{Crowning with Thorns (John 19:2)}

The soldiers platting a crown of thorns, put it upon his head; and they put on him a purple garment.

\begin{quotationx}
The “thorns” of the crown centre function as the “nails” of objectivity, which give conscience to thought. It is thanks
to them that thought has not become wholly emancipated and as arbitrary, for example, as the imagination is. Thought as
such is, in spite of all, the organ of truth, not of illusion. \begin{flushright} \textit{Meditations on the Tarot. Letter V: The Pope}\end{flushright}

\end{quotationx}

\textbf{The curse on Adam (Gen 3:17-19)}

And to Adam he said: Because thou hast hearkened to the voice of thy wife, and hast eaten of the tree, whereof I
commanded thee that thou shouldst not eat, cursed is the earth in thy work; with labour and toil shalt thou eat thereof
all the days of thy life. Thorns and thistles shall it bring forth to thee; and thou shalt eat the herbs of the earth.
In the sweat of thy face shalt thou eat bread till thou return to the earth, out of which thou wast taken: for dust
thou art, and into dust thou shalt return.

\subsection*{Day 4: Fourth stage of the passion: Carrying the cross}
Humanity must pass through the carrying of the cross after Cain's murder of Abel. Together they
form a cross:

\begin{itemize}
\item \textbf{Abel}. The vertical stream, through the ascending and accepted sacrifice. 
\item \textbf{Cain}. The horizontal stream. He created the line extending in breadth. 
\end{itemize}

\begin{quotationx}
The cause of all wars and revolutions — in a word, of all violence — is always
the same: the negation of hierarchy. This cause is found already, germinally, at such a lofty level as that of the
communal act of worship of the same God by two brothers — this is the staggering revelation of the
story of Cain and Abel. And as murders, wars and revolutions continue, the story of Cain and Abel remains ever valid
and relevant. Being always valid and relevant throughout the passage of centuries, this is a myth and, moreover, a myth
of the first order. \begin{flushright} \textit{Meditations on the Tarot. Letter I: The Magician}\end{flushright}

\end{quotationx}

\textbf{Carrying of the cross (John 19:17, Matthew 27:32, Mark 15:21, Luke 23:26)}

And bearing his own cross, he went forth to that place which is called Calvary, but in Hebrew Golgotha. And going out,
they found a man of Cyrene, named Simon: him they forced to take up his cross. And they forced one Simon a Cyrenian who
passed by, coming out of the country, the father of Alexander and of Rufus, to take up his cross. And as they led him
away, they laid hold of one Simon of Cyrene, coming from the country; and they laid the cross on him to carry after
Jesus.

\textbf{Cain and Abel (Genesis 4:1-10)}

And Adam knew Eve his wife: who conceived and brought forth Cain, saying: I have gotten a man through God. And again she
brought forth his brother Abel. And Abel was a shepherd, and Cain a husbandman. And it came to pass after many days,
that Cain offered, of the fruits of the earth, gifts to the Lord. Abel also offered of the firstlings of his flock, and
of their fat: and the Lord had respect to Abel, and to his offerings. But to Cain and his offerings he had no respect:
and Cain was exceedingly angry, and his countenance fell. And the Lord said to him: Why art thou angry? and why is thy
countenance fallen? If thou do well, shalt thou not receive? but if ill, shall not sin forthwith be present at the
door? but the lust thereof shall be under thee, and thou shalt have dominion over it. And Cain said to Abel his
brother: Let us go forth abroad. And when they were in the field, Cain rose up against his brother Abel, and slew him.
And the Lord said to Cain: Where is thy brother Abel? And he answered, I know not: am I my
brother's keeper? And he said to him: What hast thou done? the voice of thy
brother's blood crieth to me from the earth.

\subsection*{Day 5: Fifth stage of the passion: The crucifixion}
The Crucifixion is (but not only) the state of belonging nowhere, neither in heaven nor on earth. Cain is the eternal
wanderer. The two thieves are symbols of joining together of the two streams (vertical and horizontal).

\begin{quotationx}
The unconscious, instead of shocking intelligence, allies itself with it, penetrates it, and becomes luminous within it.
But this takes place only after the more-or-less long and painful experience of the crucifixion of consciousness on the
cross formed by the pair of opposites: subjectivity—objectivity (vertical), and the pair of
opposites: intelligence—unconscious wisdom (horizontal). \begin{flushright} \textit{Meditations on the Tarot. Letter XVIII: The Moon}\end{flushright} 

\end{quotationx}

\textbf{The Crucifixion (John 19:17-22, Matthew 27:31-38, Mark 15:20-32, Luke 23:26-43)}

\begin{quotationx}
In the times when the papacy was combined with worldly power the situation of inner crucifixion, and of “inner swooning”
which it produces, was often too heavy a burden inwardly for individual popes, as human beings. Thus there arose
attempts  impossible though this might be at “flight from the cross”: into politics, court life, the enjoyment of art
and culture, and the intoxications of sex. \begin{flushright} \textit{Covenant of the Heart}\end{flushright}

\end{quotationx}

And bearing his own cross, he went forth to that place which is called Calvary, but in Hebrew Golgotha. Where they
crucified him, and with him two others, one on each side, and Jesus in the midst. And Pilate wrote a title also, and he
put it upon the cross. And the writing was: JESUS OF NAZARETH, THE KING OF THE JEWS. This title therefore many of the
Jews did read: because the place where Jesus was crucified was nigh to the city: and it was written in Hebrew, in
Greek, and in Latin. Then the chief priests of the Jews said to Pilate: Write not, The King of the Jews; but that he
said, I am the King of the Jews. Pilate answered: What I have written, I have written.

And after they had mocked him, they took off the cloak from him, and put on him his own garments, and led him away to
crucify him. And going out, they found a man of Cyrene, named Simon: him they forced to take up his cross. And they
came to the place that is called Golgotha, which is the place of Calvary. And they gave him wine to drink mingled with
gall. And when he had tasted, he would not drink. And after they had crucified him, they divided his garments, casting
lots; that it might be fulfilled which was spoken by the prophet, saying: They divided my garments among them; and upon
my vesture they cast lots. And they sat and watched him. And they put over his head his cause written: THIS IS JESUS
THE KING OF THE JEWS. Then were crucified with him two thieves: one on the right hand, and one on the left.

And after they had mocked him, they took off the purple from him, and put his own garments on him, and they led him out
to crucify him. And they forced one Simon a Cyrenian who passed by, coming out of the country, the father of Alexander
and of Rufus, to take up his cross. And they bring him into the place called Golgotha, which being interpreted is, The
place of Calvary. And they gave him to drink wine mingled with myrrh; but he took it not. And crucifying him, they
divided his garments, casting lots upon them, what every man should take. And it was the third hour, and they crucified
him. And the inscription of his cause was written over: THE KING OF THE JEWS. And with him they crucify two thieves;
the one on his right hand, and the other on his left. And the scripture was fulfilled, which saith: And with the wicked
he was reputed. And they that passed by blasphemed him, wagging their heads, and saying: Vah, thou that destroys the
temple of God, and in three days builds it up again; Save thyself, coming down from the cross. In like manner also the
chief priests mocking, said with the scribes one to another: He saved others; himself he cannot save. Let Christ the
king of Israel come down now from the cross, that we may see and believe. And they that were crucified with him reviled
him.

And as they led him away, they laid hold of one Simon of Cyrene, coming from the country; and they laid the cross on him
to carry after Jesus. And there followed him a great multitude of people, and of women, who bewailed and lamented him.
But Jesus turning to them, said: Daughters of Jerusalem, weep not over me; but weep for yourselves, and for your
children. For behold, the days shall come, wherein they will say: Blessed are the barren, and the wombs that have not
borne, and the paps that have not given suck.

Then shall they begin to say to the mountains: Fall upon us; and to the hills: Cover us. For if in the green wood they
do these things, what shall be done in the dry? And there were also two other malefactors led with him to be put to
death. And when they were come to the place which is called Calvary, they crucified him there; and the robbers, one on
the right hand, and the other on the left. And Jesus said: Father, forgive them, for they know not what they do. But
they, dividing his garments, cast lots. And the people stood beholding, and the rulers with them derided him, saying:
He saved others; let him save himself, if he be Christ, the elect of God. And the soldiers also mocked him, coming to
him, and offering him vinegar, and saying: If thou be the king of the Jews, save thyself. And there was also a
superscription written over him in letters of Greek, and Latin, and Hebrew: THIS IS THE KING OF THE JEWS. And one of
those robbers who were hanged, blasphemed him, saying: If thou be Christ, save thyself and us. But the other answering,
rebuked him, saying: Neither dost thou fear God, seeing thou art condemned under the same condemnation? And we indeed
justly, for we receive the due reward of our deeds; but this man hath done no evil. And he said to Jesus: Lord,
remember me when thou shalt come into thy kingdom. And Jesus said to him: Amen I say to thee, this day thou shalt be
with me in paradise.

\textbf{The curse on Cain (Genesis 4:8-16)}

And Cain said to Abel his brother: Let us go forth abroad. And when they were in the field, Cain rose up against his
brother Abel, and slew him. And the Lord said to Cain: Where is thy brother Abel? And he answered, I know not: am I my
brother's keeper? And he said to him: What hast thou done? the voice of thy
brother's blood cries to me from the earth. Now, therefore, cursed shalt thou be upon the earth,
which hath opened her mouth and received the blood of thy brother at thy hand. When thou shalt till it, it shall not
yield to thee its fruit: a fugitive and a vagabond shalt thou be upon the earth. And Cain said to the Lord: My iniquity
is greater than that I may deserve pardon. Behold thou dost cast me out this day from the face of the earth, and I
shall be hidden from thy face, and I shall be a vagabond and a fugitive on the earth: every one, therefore, that finds
me, shall kill me. And the Lord said to him: No, it shall not be so: but whosoever shall kill Cain, shall be punished
sevenfold. And the Lord set a mark upon Cain, that whosoever found him should not kill him. And Cain went out from the
face of the Lord, and dwelt as a fugitive on the earth, at the east side of Eden.

\subsection*{Day 6: Sixth stage of the passion: Entombment}
\textbf{Noah} protected all living things until there was again the possibility for new life. Then they re-emerged.
Meditate on the image of the Ark.

\begin{quotationx}
Just as the Emerald Table saved the essence of ancient wisdom, and just as the Tarot saved the essence of mediaeval
wisdom, across the deluges which occurred in the time that separates us from them, may the essence of modern wisdom be
saved in a spiritual “\textbf{Noah's ark}” from the deluge which is going to come, and may it
thereby be transmitted to the future, just as the essence of ancient wisdom and that of mediaeval wisdom has been
transmitted to us by means of the Emerald Table and the Major Arcana of the Tarot. The tradition of Hermeticism
blossomed in the past and must live in the future. \begin{flushright} \textit{Meditations on the Tarot. Letter XVII: The Star}\end{flushright}

\end{quotationx}

\textbf{Laying in the tomb (John 19:38-42, Matthew 27:57-61, Mark 15:42-47, Luke 23:50-56)}

And after these things, Joseph of Arimathea (because he was a disciple of Jesus, but secretly for fear of the Jews)
besought Pilate that he might take away the body of Jesus. And Pilate gave leave. He came therefore, and took the body
of Jesus. And Nicodemus also came, (he who at the first came to Jesus by night,) bringing a mixture of myrrh and aloes,
about an hundred pound weight. They took therefore the body of Jesus, and bound it in linen cloths, with the spices, as
the manner of the Jews is to bury. Now there was in the place where he was crucified, a garden; and in the garden a new
sepulchre, wherein no man yet had been laid. There, therefore, because of the parasceve of the Jews, they laid Jesus,
because the sepulchre was nigh at hand.

And when it was evening, there came a certain rich man of Arimathea, named Joseph, who also himself was a disciple of
Jesus. He went to Pilate, and asked the body of Jesus. Then Pilate commanded that the body should be delivered. And
Joseph taking the body, wrapped it up in a clean linen cloth. And laid it in his own new monument, which he had hewed
out in a rock. And he rolled a great stone to the door of the monument, and went his way. And there was there Mary
Magdalen, and the other Mary sitting over against the sepulchre.

And when evening was now come, (because it was the Parasceve, that is, the day before the sabbath,) Joseph of Arimathea,
a noble counsellor, who was also himself looking for the kingdom of God, came and went in boldly to Pilate, and begged
the body of Jesus. But Pilate wondered that he should be already dead. And sending for the centurion, he asked him if
he were already dead. And when he had understood it by the centurion, he gave the body to Joseph. And Joseph buying
fine linen, and taking him down, wrapped him up in the fine linen, and laid him in a sepulchre which was hewed out of a
rock. And he rolled a stone to the door of the sepulchre. And Mary Magdalen, and Mary the mother of Joseph, beheld
where he was laid.

And behold there was a man named Joseph, who was a counsellor, a good and just man, (The same had not consented to their
counsel and doings;) of Arimathea, a city of Judea; who also himself looked for the kingdom of God. This man went to
Pilate, and begged the body of Jesus. And taking him down, he wrapped him in fine linen, and laid him in a sepulchre
that was hewed in stone, wherein never yet any man had been laid. And it was the day of the Parasceve, and the sabbath
drew on. And the women that were come with him from Galilee, following after, saw the sepulchre, and how his body was
laid. And returning, they prepared spices and ointments; and on the sabbath day they rested, according to the
commandment.

\textbf{The Flood (Genesis 6:5-22, 7:1-23)}

And God seeing that the wickedness of men was great on the earth, and that all the thought of their heart was bent upon
evil at all times, It repented him that he had made man on the earth. And being touched inwardly with sorrow of heart,
He said: I will destroy man, whom I have created, from the face of the earth, from man even to beasts, from the
creeping thing even to the fowls of the air, for it repenteth me that I have made them. But Noah found grace before the
Lord. These are the generations of Noah: Noah was a just and perfect man in his generations, he walked with God. And he
begot three sons, Sem, Cham, and Japheth.

And the earth was corrupted before God, and was filled with iniquity. And when God had seen that the earth was corrupted
(for all flesh had corrupted its way upon the earth,) He said to Noah: The end of all flesh is come before me, the
earth is filled with iniquity through them, and I will destroy them with the earth. Make thee an ark of timber planks:
thou shalt make little rooms in the ark, and thou shalt pitch it within and without. And thus shalt thou make it: The
length of the ark shall be three hundred cubits: the breadth of it fifty cubits, and the height of it thirty cubits.
Thou shalt make a window in the ark, and in a cubit shalt thou finish the top of it: and the door of the ark thou shalt
set in the side: with lower, middle chambers, and third stories shalt thou make it. Behold I will bring the waters of a
great flood upon the earth, to destroy all flesh, wherein is the breath of life, under heaven. All things that are in
the earth shall be consumed. And I will establish my covenant with thee, and thou shalt enter into the ark, thou and
thy sons, and thy wife, and the wives of thy sons with thee. And of every living creature of all flesh, thou shalt
bring two of a sort into the ark, that they may live with thee: of the male sex, and the female. Of fowls according to
their kind, and of beasts in their kind, and of every thing that creepeth on the earth according to its kind; two of
every sort shall go in with thee, that they may live. Thou shalt take unto thee of all food that may be eaten, and thou
shalt lay it up with thee: and it shall be food for thee and them. And Noah did all things which God commanded him.

And the Lord said to him: Go in thou and all thy house into the ark: for thee I have seen just before me in this
generation. Of all clean beasts take seven and seven, the male and the female. But of the beasts that are unclean two
and two, the male and the female. Of the fowls also of the air seven and seven, the male and the female: that seed may
be saved upon the face of the whole earth. For yet a while, and after seven days, I will rain upon the earth forty days
and forty nights; and I will destroy every substance that I have made, from the face of the earth. And Noah did all
things which the Lord had commanded him. And he was six hundred years old, when the waters of the flood overflowed the
earth. And Noah went in and his sons, his wife and the wives of his sons with him into the ark, because of the waters
of the flood. And of beasts clean and unclean, and of fowls, and of every thing that moveth upon the earth, Two and two
went in to Noe into the ark, male and female, as the Lord had commanded Noah.

And after the seven days were passed, the waters of the flood overflowed the earth. In the six hundredth year of the
life of Noah, in the second month, in the seventeenth day of the month, all the fountains of the great deep were broken
up, and the flood gates of heaven were opened: And the rain fell upon the earth forty days and forty nights. In the
selfsame day Noah, and Shem, and Ham, and Japheth his sons: his wife, and the three wives of his sons with them, went
into the ark: They and every beast according to its kind, and all the cattle in their kind, and every thing that moveth
upon the earth according to its kind, and every fowl according to its kind, all birds, and all that fly, Went in to
Noah into the ark, two and two of all flesh, wherein was the breath of life. And they that went in, went in male and
female of all flesh, as God had commanded him: and the Lord shut him in on the outside.

And the flood was forty days upon the earth, and the waters increased, and lifted up the ark on high from the earth. For
they overflowed exceedingly: and filled all on the face of the earth: and the ark was carried upon the waters. And the
waters prevailed beyond measure upon the earth: and all the high mountains under the whole heaven were covered. The
water was fifteen cubits higher than the mountains which it covered. And all flesh was destroyed that moved upon the
earth, both of fowl, and of cattle, and of beasts, and of all creeping things that creep upon the earth: and all men.
And all things wherein there is the breath of life on the earth, died. And he destroyed all the substance that was upon
the earth, from man even to beast, and the creeping things and fowls of the air: and they were destroyed from the
earth: and Noe only remained, and they that were with him in the ark.

\subsection*{Day 7: Seventh stage of the passion: The resurrection}
The seventh stage is the Resurrection which came when everything re-emerged from Noah. The rainbow is the symbol of
reconciliation with God.

\begin{quotationx}
The concept, idea, and ideal of resurrection is different from the concept, idea, and ideal of eternal life, as a
resting in God, where the soul returns to the “Father's house” and remains there for all eternity.
Resurrection is not merely the end of the old, but actually means the beginning of a new world-creation,  with “a new
heaven and a new earth”. It is not the soul's rest in eternity, not the eternal rest of Nirvana,
but an active cooperation with God in a renewed world, healed of the consequences of the Fall into sin. The “new
Jerusalem” is a new world-order related to the old one in the same way as the resurrected man is related to mortal man.
The resurrection may be conceived of as the reappearance and active participation of the total human being
— as spirit, soul, and body — in the domain of the “world-in-­progress”. As
an idea, the resurrection is the realisation that God is divinely generous, that he does not take away again what he
has once given and granted, but that his gifts — existence, consciousness, freedom, and creative
activity — are valid for all eternity. \begin{flushright} \textit{Covenant of the Heart}\end{flushright}

\end{quotationx}

\textbf{The Resurrection (John 20:1-10, Matthew 28:1-10, Mark 16:1-11, Luke 24:1-12)}

And on the first day of the week, Mary Magdalen cometh early, when it was yet dark, unto the sepulchre; and she saw the
stone taken away from the sepulchre. She ran, therefore, and cometh to Simon Peter, and to the other disciple whom
Jesus loved, and saith to them: They have taken away the Lord out of the sepulchre, and we know not where they have
laid him. Peter therefore went out, and that other disciple, and they came to the sepulchre. And they both ran
together, and that other disciple did outrun Peter, and came first to the sepulchre. And when he stooped down, he saw
the linen cloths lying; but yet he went not in. Then cometh Simon Peter, following him, and went into the sepulchre,
and saw the linen cloths lying, And the napkin that had been about his head, not lying with the linen cloths, but
apart, wrapped up into one place. Then that other disciple also went in, who came first to the sepulchre: and he saw,
and believed. For as yet they knew not the scripture, that he must rise again from the dead. The disciples therefore
departed again to their home.

And in the end of the sabbath, when it began to dawn towards the first day of the week, came Mary Magdalen and the other
Mary, to see the sepulchre. And behold there was a great earthquake. For an angel of the Lord descended from heaven,
and coming, rolled back the stone, and sat upon it. And his countenance was as lightning, and his raiment as snow. And
for fear of him, the guards were struck with terror, and became as dead men. And the angel answering, said to the
women: Fear not you; for I know that you seek Jesus who was crucified. He is not here, for he is risen, as he said.
Come, and see the place where the Lord was laid. And going quickly, tell ye his disciples that he is risen: and behold
he will go before you into Galilee; there you shall see him. Lo, I have foretold it to you. And they went out quickly
from the sepulchre with fear and great joy, running to tell his disciples. And behold Jesus met them, saying: All hail.
But they came up and took hold of his feet, and adored him. Then Jesus said to them: Fear not. Go, tell my brethren
that they go into Galilee, there they shall see me.

And when the sabbath was past, Mary Magdalen, and Mary the mother of James, and Salome, bought sweet spices, that
coming, they might anoint Jesus. And very early in the morning, the first day of the week, they come to the sepulchre,
the sun being now risen. And they said one to another: Who shall roll us back the stone from the door of the sepulchre?
And looking, they saw the stone rolled back. For it was very great. And entering into the sepulchre, they saw a young
man sitting on the right side, clothed with a white robe: and they were astonished. Who saith to them: Be not
affrighted; you seek Jesus of Nazareth, who was crucified: he is risen, he is not here, behold the place where they
laid him. But go, tell his disciples and Peter that he goeth before you into Galilee; there you shall see him, as he
told you.  But they going out, fled from the sepulchre. For a trembling and fear had seized them: and they said nothing
to any man; for they were afraid. But he rising early the first day of the week, appeared first to Mary Magdalen, out
of whom he had cast seven devils. She went and told them that had been with him, who were mourning and weeping. And
they hearing that he was alive, and had been seen by her, did not believe.

And on the first day of the week, very early in the morning, they came to the sepulchre, bringing the spices which they
had prepared. And they found the stone rolled back from the sepulchre. And going in, they found not the body of the
Lord Jesus. And it came to pass, as they were astonished in their mind at this, behold, two men stood by them, in
shining apparel. And as they were afraid, and bowed down their countenance towards the ground, they said unto them: Why
seek you the living with the dead? He is not here, but is risen. Remember how he spoke unto you, when he was in
Galilee, Saying: The Son of man must be delivered into the hands of sinful men, and be crucified, and the third day
rise again. And they remembered his words. And going back from the sepulchre, they told all these things to the eleven,
and to all the rest. And it was Mary Magdalen, and Joanna, and Mary of James, and the other women that were with them,
who told these things to the apostles. And these words seemed to them as idle tales; and they did not believe them. But
Peter rising up, ran to the sepulchre, and stooping down, he saw the linen cloths laid by themselves; and went away
wondering in himself at that which was come to pass.

\textbf{The End of the Flood (Genesis 8:24-33)}

Therefore in the six hundredth and first year, the first month, the first day of the month, the waters were lessened
upon the earth, and Noe opening the covering of the ark, looked, and saw that the face of the earth was dried. In the
second month, the seven and twentieth day of the month, the earth was dried. And God spoke to Noah, saying:

Go out of the ark, thou and thy wife, thy sons, and the wives of thy sons with thee. All living things that are with
thee of all flesh, as well in fowls as in beasts, and all creeping things that creep upon the earth, bring out with
thee, and go ye upon the earth: increase and multiply upon it.

So Noah went out, he and his sons: his wife, and the wives of his sons with him. And all living things, and cattle, and
creeping things that creep upon the earth, according to their kinds, went out of the ark. And Noah built an altar unto
the Lord: and taking of all cattle and fowls that were clean, offered holocausts upon the altar. And the Lord smelled a
sweet saviour, and said:

I will no more curse the earth for the sake of man: for the imagination and thought of man's heart
are prone to evil from his youth: therefore I will no more destroy every living soul as I have done. All the days of
the earth, seedtime and harvest, cold and heat, summer and winter, night and day, shall not cease.

\textbf{The Rainbow, symbol of the reconciliation with God (Genesis 9:12-17)}

And God said:

This is the sign of the covenant which I give between me and you, and to every living soul that is with you, for
perpetual generations. I will set my bow in the clouds, and it shall be the sign of a covenant between me, and between
the earth. And when I shall cover the sky with clouds, my bow shall appear in the clouds: And I will remember my
covenant with you, and with every living soul that beareth flesh: and there shall no more be waters of a flood to
destroy all flesh.

And the bow shall be in the clouds, and I shall see it, and shall remember the everlasting covenant, that was made
between God and every living soul of all flesh which is upon the earth.

And God said to Noah: This shall be the sign of the covenant which I have established between me and all flesh upon the
earth.

\flright{\small\textit{Posted on 2023-03-12 by Cologero}}