\section{Week 8: Give us this day our daily bread} 

\textsc{Task for the week}: Meditate on the good communion, be wary of the bad communion.

\textsc{Reading for the week}: John 13-17

\subsection*{Day 1: Spiritual-physical nourishment}
\textbf{The Good Communion}

And God blessed them, saying: Increase and multiply, and fill the earth, and subdue it, and rule over the fishes of the
sea, and the fowls of the air, and all living creatures that move upon the earth.  And God said: Behold I have given
you every herb bearing seed upon the earth, and all trees that have in themselves seed of their own kind, to be your
meat: And to all beasts of the earth, and to every fowl of the air, and to all that move upon the earth, and wherein
there is life, that they may have to feed upon. And it was so done. 

Genesis 1:28-30

Through spiritual-physical nourishment (the seed forces), Adam connected himself with all the good forces. He mastered
the active forces.

Seed forces arise from the root chakra on the Tree of Life.

\begin{quotationx}
Chastity is the state of the human being where the centre named in occidental esotericism as the “twelve-petalled lotus”
is awakened and becomes the sun of the microcosmic “planetary system”. The three lotus-centres situated below it (the
ten-petalled, the six-petalled, and the four-petalled) begin then to function in conformity with the life of the heart
(the twelve-petalled lotus), i.e., “according to solar law”. \begin{flushright} \emph{Meditations on the Tarot. Letter V: The Pope}\end{flushright}

\end{quotationx}
The four-petalled lotus is that of creative force.

\textbf{The Evil Communion}

The evil seed-force was forbidden. That was the origin of the good and evil communions.  It destroyed the petals of the
root chakra, or life forces. This is the forbidden fruit.

\begin{quotationx}
Concerning the centre from which the current necessary for “Simonian” levitation is made to emanate: it is that of the
“four-petalled lotus” (\emph{muladhara chakra}), where the “serpent power” (\emph{kundalini}) —
latent electrical force — is found. Now, this “serpent power” can be awoken and directed either
above (yoga) or below and outside (arbitrary magic). \begin{flushright} \emph{Meditations on the Tarot. Letter XII: The Hanged Man}\end{flushright}

\end{quotationx}

\subsection*{Day 2: Abraham's Initiation}
\textbf{The Good Communion}

And Sarai said to Abram: Thou dost unjustly with me: I gave my handmaid into thy bosom, and she perceiving herself to be
with child, despiseth me. The Lord judge between me and thee. (Genesis 16:5)

Abram was confronted with the choice between good and evil communion.

Abram = sublime father; Abraham = father of a multitude.

But Melchizedek the king of Salem, bringing forth bread and wine, for he was the priest of the most high God, Blessed
him, and said: Blessed be Abram by the most high God, who created heaven and earth. 

Genesis 14:18-19

Melchizedek = king of righteousness; Salem = king of peace. Abraham was initiated with bread and wine.

For this Melchizedek was king of Salem, priest of the most high God, who met Abraham returning from the slaughter of the
kings, and blessed him: To whom also Abraham divided the tithes of all: who first indeed by interpretation, is king of
justice: and then also king of Salem, that is, king of peace: Without father, without mother, without genealogy, having
neither beginning of days nor end of life, but likened unto the Son of God, continueth a priest forever. Now consider
how great this man is, to whom also Abraham the patriarch gave tithes out of the principal things. 

Hebrews 7:1-4

\begin{quotationx}
Primordial water penetrated by divine breath is the essence of blood; breath reflected by the water is light; the
rhythmic alternation from absorption of the breath by water to its reflection by it is respiration. Light is the day,
blood is the night, and respiration is plenitude (Salem). Melchizedek, king of Salem, priest of the Most High God is
therefore appointed to plenitude, to respiration, whilst the anointed king, guardian of the throne of David, or the
Emperor, is appointed to the day. \begin{flushright} \emph{Meditations on the Tarot. Letter V: The Pope}\end{flushright}

\end{quotationx}

\textbf{The Evil Communion}

And the king of Sodom said to Abram: Give me the persons, and the rest take to thyself. And he answered him: I lift up
my hand to the Lord God the most high, the possessor of heaven and earth, that from the very woof thread unto the shoe
latchet, I will not take of any things that are thine, lest thou say I have enriched Abram: Except such things as the
young men have eaten, and the shares of the men that came with me, Aner, Escol, and Mambre: these shall take their
shares. (Genesis 14:21-24)

Abraham refused the evil offer to become king of Sodom and Gomorrah.

The king of Sodom offers Abram bread (the provision for the two cities) and stone (the conquered cities of the enemy kings).

\begin{quotationx}
If ten righteous men had been found in Sodom and Gomorrah, God would have spared these cities. \begin{flushright} \emph{Meditations on
the Tarot. Letter X: The Wheel of Fortune}\end{flushright}

\end{quotationx}

\subsection*{Day 3: Moses feeds his people}
\textsc{The Good Communion}:
It is a good communion when Moses feeds his people with Manna.

\begin{multicols}{2}\small
And the Lord said to Moses: Behold I will rain bread from heaven for you: let the people go forth, and gather what is
sufficient for every day: that I may prove them whether they will walk in my law, or not.  But the sixth day let them
provide for to bring in: and let it be double to that they were wont to gather every day. \flright{\itshape Exodus 16:4-5}

So it came to pass in the evening, that quails coming up, covered the camp: and in the morning, a dew lay round about
the camp. And when it had covered the face of the earth, it appeared in the wilderness small, and as it were beaten
with a pestle, like unto the hoar frost on the ground. And when the children of Israel saw it, they said one to
another: Manhu! which signifieth: What is this! for they knew not what it was. And Moses said to them: This is the
bread, which the Lord hath given you to eat. This is the word, that the Lord hath commanded: Let everyone gather of it
as much as is enough to eat: a gomor for every man, according to the number of your souls that dwell in a tent, so
shall you take of it. And the children of Israel did so: and they gathered, one more, another less. And they measured
by the measure of a gomor: neither had he more that had gathered more: nor did he find less that had provided less: but
everyone had gathered, according to what they were able to eat. \flright{\itshape Exodus 16:13-18}

They said therefore to him: What sign therefore dost thou shew, that we may see, and may believe thee? What dost thou
work? Our fathers did eat manna in the desert, as it is written: He gave them bread from heaven to eat. Then Jesus said
to them: Amen, amen I say to you; Moses gave you not bread from heaven, but my Father giveth you the true bread from
heaven. For the bread of God is that which cometh down from heaven, and giveth life to the world. They said therefore
unto him: Lord, give us always this bread. And Jesus said to them: \textbf{I am the bread of life}: he that cometh to
me shall not hunger: and he that believeth in me shall never thirst. \flright{\itshape John 6:30-35}
\end{multicols}

\begin{quotationx}
St. John of the Cross understands the way of purification through the inner desert as a path for mature souls who are
strong enough to bear the tediousness, stillness, and loneliness of the “night of the spirit.” Nevertheless, these
souls also have need of being strengthened, encouraged, and cheered up. It is one of the experiences that every human
being who treads the “desert path” knows, that at night, in the condition of sleep, something happens that gives him
again and again new strength to endure and not to despair. The night, which seems to be just as desert-like, unmoving,
and dark as the “night of the spirit” experienced during the waking consciousness of day, in time transforms itself
into a giver of strength of soul and courage of spirit for the human being. It is as if something received as a kind of
aftereffect from the heavenly choirs of the spiritual hierarchies was actively giving new life and strength to the soul
(and often to the body, too) of the awakened wanderer. The result is renewed courage for life and the temporary
disappearing of any lack of hope. This strengthening is not due to a dream or any kind of instruction during dreaming,
but purely and simply to the condition resulting during the night. Despondency, or even despair, simply disappears by
itself, and one is reinvigorated and strengthened to continue the “path through the desert.” It is not arbitrary to
compare this aftereffect of the night with the miraculous nourishment of \textbf{manna}, to see here an analogy with
this feeding of the chosen people in the desert during their wandering through the wilderness. \begin{flushright} \emph{Covenant of
the Heart}\end{flushright}

\end{quotationx}

\textsc{The Evil Communion}:
Evil communion is the worship of the golden calf and yearning for the fleshpots of Egypt.

The golden calf is imitated in the worship of those who attend to external life without awakening to life, i.e., without
making an effort to acquire exact knowledge of the laws of life.

Now, there is the Word, and there are egregores before whom humanity bows down; there is revelation of divine truth, and
the manifestation of the will of human beings; there is the cult of God, and that of idols made by man. Is it not a
diagnosis and prognosis of the whole history of the human race that at the same time that Moses received the revelation
of the Word at the summit of the mountain, the people at the foot of the mountain made and worshipped a golden calf?

\begin{quotationx}
The Word and idols, revealed truth and “ideological superstructures” of the human will, operate simultaneously in the
history of the human race. Has there been a single century when the servants of the Word have not had to confront the
worshippers of idols, egregores? \begin{flushright} \emph{Meditations on the Tarot. Letter XV: The Devil}\end{flushright}

\end{quotationx}

Those at the foot of the mountain collected contributions of gold jewelry and made from them the so-called golden calf,
the idol of a golden bull. With respect to this archetypal phenomenon of “falling away,” it is not a question merely of
the victory of inclination—preferring the sense perceptible and material to what is supersensible
and purely moral—but of something deeper and more significant. It is actually a matter of an
insurrection of the collective will of the people asserting itself against the aristocratic, hierarchical order that
Moses represented.

The turning away from the God who was revealed and proclaimed, in favor of the self-chosen and created god is, as it is
depicted in the Bible, the original phenomenon of all stages and forms of falling away from the truths of revelation in
favor of the collective will of the people, which usually comes to expression as the demand for being in tune with the
“spirit of the times.”

The “golden calf” was made and took the place of the dogma of the God of Being. The “golden calf” came about not through
doubt in the revelation proclaimed through Moses, but to make it plausible.

\begin{quotationx}
When Yahweh wanted to destroy the people of Israel because of their turning away from Him and on account of their
worship of the golden calf at the foot of Mt. Sinai, Moses asked that he, instead, be blotted out from the book of life
rather than that the people of Israel be destroyed. Thus he attained pardon for the people of Israel. In his
sacrificial willingness a deeper understanding of the divine expressed itself. \begin{flushright} \emph{Covenant of the Heart}\end{flushright}

\end{quotationx}

\begin{multicols}{2}\small
And the people seeing that Moses delayed to come down from the mount, gathering together against Aaron, said: Arise,
make us gods, that may go before us: for as to this Moses, the man that brought us out of the land of Egypt, we know
not what has befallen him. And Aaron said to them: Take the golden earrings from the ears of your wives, and your sons
and daughters, and bring them to me. And the people did what he had commanded, bringing the earrings to Aaron. And when
he had received them, he fashioned them by founders' work, and made of them a molten calf. And
they said: These are thy gods, O Israel, that have brought thee out of the land of Egypt. And when Aaron saw this, he
built an altar before it, and made proclamation by a crier's voice, saying: Tomorrow is the
solemnity of the Lord. And rising in the morning, they offered holocausts, and peace victims, and the people sat down
to eat, and drink, and they rose up to play. And the Lord spoke to Moses, saying: Go, get thee down: thy people, which
thou hast brought out of the land of Egypt, hath sinned. They have quickly strayed from the way which thou didst shew
them: and they have made to themselves a molten calf, and have adored it, and sacrificing victims to it, have said:
These are thy gods, O Israel, that have brought thee out of the land of Egypt. And again the Lord said to Moses: See
that this people is stiffnecked: Let me alone, that my wrath may be kindled against them, and that I may destroy them,
and I will make of thee a great nation. \flright{\itshape Exodus 32:1-10}
\end{multicols}

\subsubsection*{The fleshpots of Egypt}

The exodus out of Egypt and the wandering in the desert preceded the revelation on Mt. Sinai. For Egypt was the epitome
of all kinds of worship of the elements of existence in space (sun, moon, and stars) and in time (fertility, the power
of procreation, life and death, natural evolution), i.e., those very things that work as necessities of nature and
represent what is coercive about worldly existence.

\begin{quotationx}
Egypt was the “house of bondage” because of the worship of the compelling necessities of existence, the “gods” of
material being. The exodus of the Israelites out of Egypt was for this reason an unprecedented revolutionary event: a
multitude of human beings wanted to go forth into the desert in order to worship there the God who is not to be found
in material existence. \begin{flushright} \emph{Covenant of the Heart}\end{flushright}


And they chode with Moses, and said: Give us water, that we may drink. And Moses answered them: Why chide you with me?
Wherefore do you tempt the Lord? So the people were thirsty there for want of water, and murmured against Moses,
saying: Why didst thou make us go forth out of Egypt, to kill us and our children, and our beasts with thirst? 

\flright{\itshape Exodus 17:2-3}
\end{quotationx}

\subsection*{Day 4: The Feeding of the 5000}
The feeding of the 5000 with five barley loaves and two fishes.

\begin{multicols}{2}\small
After these things Jesus went over the sea of Galilee, which is that of Tiberias. And a great multitude followed him,
because they saw the miracles which he did on them that were diseased. Jesus therefore went up into a mountain, and
there he sat with his disciples. Now the pasch, the festival day of the Jews, was near at hand. When Jesus therefore
had lifted up his eyes, and seen that a very great multitude cometh to him, he said to Philip: Whence shall we buy
bread, that these may eat? And this he said to try him; for he himself knew what he would do. Philip answered him: Two
hundred pennyworth of bread is not sufficient for them, that every one may take a little. One of his disciples, Andrew,
the brother of Simon Peter, saith to him:

There is a boy here that hath five barley loaves, and two fishes; but what are these among so many?

Then Jesus said: Make the men sit down.

Now there was much grass in the place. The men therefore sat down, in number about five thousand. And Jesus took the
loaves: and when he had given thanks, he distributed to them that were set down. In like manner also of the fishes, as
much as they would. And when they were filled, he said to his disciples: Gather up the fragments that remain, lest they
be lost. They gathered up therefore, and filled twelve baskets with the fragments of the five barley loaves, which
remained over and above to them that had eaten. Now those men, when they had seen what a miracle Jesus had done, said:
This is of a truth the prophet, that is to come into the world. Jesus therefore, when he knew that they would come to
take him by force, and make him king, fled again into the mountain himself alone. \flright{\itshape John 6:1-15}
\end{multicols}

\textsc{The Good Communion}:
Drives that spring from love bring about healings. Love is present only where the human being is able to act out of
intuition consciously acquired.

\begin{quotationx}
The seven miracles in John's Gospel did not have the serpent as their agent, nor was the brain the
instrument of their accomplishment, nor was cerebral intellectuality the source of their initiative. The agent here is
the dove, i.e., the Spirit which is above the brain, above the head, and which \emph{descends} upon the head and
remains there — the Spirit which \emph{transcends} cerebral intellectuality. This Spirit is the
source of initiative and, simultaneously, is the agent and instrument of divine or sacred magic. \begin{flushright} \emph{Meditations
on the Tarot. Letter X: The Wheel of Fortune}\end{flushright}

\end{quotationx}

As the Sun — raying out light, warmth, and life — “nourishes” all beings and
unites them in a “community of nourishment,” so Jesus Christ functioned at the feeding of the five thousand as the
“nourishment giving center” for the five thousand. He carried out during the short time of the “feeding” what the sun
effects in the course of the year, when it brings about a “multiplication of bread” through the sprouting, growth, and
ripening of corn.

The disciples received and passed on the blest (eucharistic) bread, just as the moon receives and passes on the light of
the sun in a dimmed, toned-down form.

This mediating effect of the moon, which transforms the boundless, streaming strength of the sun such that it becomes
more individually acceptable, can also be understood in relation to the experience of the sacrament of holy communion
received at the altar.

\textsc{The Evil Communion}:
Feedings that take place without Christ fulfil the purposes of Ahriman. They are feeding out of the heart of Satan and
strengthening the I via currents of the fallen chakras. The feeding always enters into a man when a destruction of the
seed is undertaken out of passion. Ahriman enters their hearts, haunting the life-sphere and destroying the life-forces
in the life-centre of the heart.

\subsection*{Day 5: The Last Supper}
\textsc{The Good Communion}: The Last Supper.

Scriptural readings: Matthew 26:20-30, Mark 14:17-26, Luke 22:7-23

\begin{quotationx}
Teresa Neumann lived in our time at Konnarsreuth (Bavaria) solely from Holy Communion for decades; St. Catherine of
Sienna lived nine years from Holy Communion alone; St. Lidvina of Schiedam (near Rotterdam, Holland) likewise lived for
many years exclusively from Holy Communion —to cite only the cases that are well-verified.

This is the significance of the words: “Man shall not live by bread alone but by every word that proceeds from the mouth
of God.” Here is its principal implication: as the law of evolution, the law of the serpent, comprises the struggle for
existence and as “bread” or food is the principal factor in the struggle for existence, the fact of the entry of grace
into human history since Jesus Christ signifies at the same time the possibility of gradually abolishing the struggle
for existence. \begin{flushright} \emph{Meditations on the Tarot. Letter VI: The Lover}\end{flushright}

\end{quotationx}
Christ's last words at the institution of the holy sacrament at the Last Supper: “Do this in memory
of me” point toward the sacraments, too, as being a reenlivening in the present of what happened in the past. In the
holy sacrament at the altar, memory becomes an act of the divine magic of transubstantiation, an act relating to the
real (not just remembered) presence of the body and blood of the Redeemer. What once took place, takes place now in the
present. In the sacrament, memory does not become a journey into the past, but instead brings the past into present, an
evocation that summons something out of the realm of forgetting, sleep, and death.

\begin{quotationx}
The words: “Do this in memory of me” actually mean: “Do this, so that I may be present.” \begin{flushright} \emph{Covenant of the
Heart}\end{flushright}

\end{quotationx}

\textsc{The Evil Communion:} The temptation of Jesus in the wilderness.

Scriptural Reading: Matthew 4:1-11

The temptation to Materialism:

\begin{quotationx}
And when he had fasted forty days and forty nights, afterwards he was hungry. And the tempter coming said to him: If
thou be the Son of God, command that these stones be made bread. Who answered and said: It is written, Not in bread
alone doth man live, but in every word that proceedeth from the mouth of God. \flright{\itshape Matthew 4:2-4}
\end{quotationx}

\subsubsection*{Bible Passages}
\begin{multicols}{2}\small
But when it was evening, he sat down with his twelve disciples. And whilst they were eating, he said: Amen I say to you,
that one of you is about to betray me. And they being very much troubled, began everyone to say: Is it I, Lord? But he
answering, said: He that dippeth his hand with me in the dish, he shall betray me. The Son of man indeed goeth, as it
is written of him: but woe to that man by whom the Son of man shall be betrayed: it were better for him, if that man
had not been born. And Judas that betrayed him, answering, said: Is it I, Rabbi? He saith to him: Thou hast said it.
And whilst they were at supper, Jesus took bread, and blessed, and broke: and gave to his disciples, and said: Take ye,
and eat. This is my body. And taking the chalice, he gave thanks, and gave to them, saying: Drink ye all of this.  For
this is my blood of the new testament, which shall be shed for many unto remission of sins. And I say to you, I will
not drink from henceforth of this fruit of the vine, until that day when I shall drink it with you new in the kingdom
of my Father. And a hymn being said, they went out unto mount Olivet. \flright{\itshape Matthew 26:20-30}

And when evening was come, he cometh with the twelve. And when they were at table and eating, Jesus saith: Amen I say to
you, one of you that eateth with me shall betray me. But they began to be sorrowful, and to say to him one by one: Is
it I?  Who saith to them: One of the twelve, who dippeth with me his hand in the dish. And the Son of man indeed goeth,
as it is written of him: but woe to that man by whom the Son of man shall be betrayed. It were better for him, if that
man had not been born. And whilst they were eating, Jesus took bread; and blessing, broke, and gave to them, and said:
Take ye. This is my body. And having taken the chalice, giving thanks, he gave it to them. And they all drank of it.
And he said to them: This is my blood of the new testament, which shall be shed for many. Amen I say to you, that I
will drink no more of the fruit of the vine, until that day when I shall drink it new in the kingdom of God. And when
they had said an hymn, they went forth to the mount of Olives. \flright{\itshape Mark 14:17-26}

And the day of the unleavened bread came, on which it was necessary that the pasch should be killed. And he sent Peter
and John, saying: Go, and prepare for us the pasch, that we may eat. But they said: Where wilt thou that we prepare?
And he said to them: Behold, as you go into the city, there shall meet you a man carrying a pitcher of water: follow
him into the house where he entereth in. And you shall say to the goodman of the house: The master saith to thee, Where
is the guest chamber, where I may eat the pasch with my disciples? And he will shew you a large dining room, furnished;
and there prepare. And they going, found as he had said to them, and made ready the pasch. And when the hour was come,
he sat down, and the twelve apostles with him. And he said to them: With desire I have desired to eat this pasch with
you, before I suffer. For I say to you, that from this time I will not eat it, till it be fulfilled in the kingdom of
God. And having taken the chalice, he gave thanks, and said: Take, and divide it among you: For I say to you, that I
will not drink of the fruit of the vine, till the kingdom of God come. And taking bread, he gave thanks, and brake; and
gave to them, saying: This is my body, which is given for you. Do this for a commemoration of me. In like manner the
chalice also, after he had supped, saying: This is the chalice, the new testament in my blood, which shall be shed for
you. But yet behold, the hand of him that betrayeth me is with me on the table. And the Son of man indeed goeth,
according to that which is determined: but yet, woe to that man by whom he shall be betrayed. And they began to inquire
among themselves, which of them it was that should do this thing. \flright{\itshape Luke 22:7-23}

Then Jesus was led by the spirit into the desert, to be tempted by the devil. And when he had fasted forty days and
forty nights, afterwards he was hungry. And the tempter coming said to him: If thou be the Son of God, command that
these stones be made bread. Who answered and said: It is written, Not in bread alone doth man live, but in every word
that proceedeth from the mouth of God. Then the devil took him up into the holy city, and set him upon the pinnacle of
the temple, And said to him: If thou be the Son of God, cast thyself down, for it is written: That he hath given his
angels charge over thee, and in their hands shall they bear thee up, lest perhaps thou dash thy foot against a stone.
Jesus said to him: It is written again: Thou shalt not tempt the Lord thy God. Again the devil took him up into a very
high mountain, and shewed him all the kingdoms of the world, and the glory of them, And said to him: All these will I
give thee, if falling down thou wilt adore me. Then Jesus saith to him: Begone, Satan: for it is written, The Lord thy
God shalt thou adore, and him only shalt thou serve. Then the devil left him; and behold angels came and ministered to
him. \flright{\itshape Matthew 4:1-11}
\end{multicols}

\subsection*{Day 6: Instituting the Eucharist}
Meditation on Luke 22:7-38.

\textbf{The Good Communion: The Eucharist}

\begin{quotationx}
The Eucharist is the whole of Christianity; and through it Christianity has become living magic. Since Jesus there are
still sorcerers, (but) there are no more mages. \begin{flushright} \textsc{Josephin Peladan}\end{flushright}

\end{quotationx}

On the day of the unleavened bread, Jesus and the disciples had a meal in a large, furnished dining room.

The Jesus established the Eucharist with the bread and wine.

\textbf{The Evil Communion: betrayal, strife, and denial}

Although participated in the first Eucharist, there was an evil side.

\begin{itemize}
\item \textsc{Betrayal}: Jesus predicted Judas' betrayal 
\item \textsc{Strife}: The disciples bickered among themselves for positions of power and authority 
\item \textsc{Denial}: Jesus predicted Peter's denial 
\end{itemize}

\subsubsection*{Bible passages: The Last Supper}
\begin{multicols}{2}\small
And the day of the unleavened bread came, on which it was necessary that the pasch should be killed. And he sent Peter
and John, saying: Go, and prepare for us the pasch, that we may eat. But they said: Where wilt thou that we prepare?
And he said to them: Behold, as you go into the city, there shall meet you a man carrying a pitcher of water: follow
him into the house where he entereth in. And you shall say to the goodman of the house: The master saith to thee, Where
is the guest chamber, where I may eat the pasch with my disciples? And he will shew you a large dining room, furnished;
and there prepare. And they going, found as he had said to them, and made ready the pasch. And when the hour was come,
he sat down, and the twelve apostles with him. And he said to them: With desire I have desired to eat this pasch with
you, before I suffer. For I say to you, that from this time I will not eat it, till it be fulfilled in the kingdom of
God.

And having taken the chalice, he gave thanks, and said: Take, and divide it among you: For I say to you, that I will not
drink of the fruit of the vine, till the kingdom of God come. And taking bread, he gave thanks, and brake; and gave to
them, saying: This is my body, which is given for you. Do this for a commemoration of me. In like manner the chalice
also, after he had supped, saying: This is the chalice, the new testament in my blood, which shall be shed for you.

But yet behold, the hand of him that betrayeth me is with me on the table. And the Son of man indeed goeth, according to
that which is determined: but yet, woe to that man by whom he shall be betrayed.  And they began to inquire among
themselves, which of them it was that should do this thing. And there was also a strife amongst them, which of them
should seem to be the greater. And he said to them: The kings of the Gentiles lord it over them; and they that have
power over them, are called beneficent. But you not so: but he that is the greater among you, let him become as the
younger; and he that is the leader, as he that serveth. For which is greater, he that sitteth at table, or he that
serveth? Is it not he that sitteth at table? But I am in the midst of you, as he that serveth: And you are they who
have continued with me in my temptations:

And I dispose to you, as my Father hath disposed to me, a kingdom; that you may eat and drink at my table, in my
kingdom: and may sit upon thrones, judging the twelve tribes of Israel. And the Lord said: Simon, Simon, behold Satan
hath desired to have you, that he may sift you as wheat: But I have prayed for thee, that thy faith fail not: and thou,
being once converted, confirm thy brethren. Who said to him: Lord, I am ready to go with thee, both into prison, and to
death. And he said: I say to thee, Peter, the cock shall not crow this day, till thou thrice deniest that thou knowest
me. And he said to them: When I sent you without purse, and scrip, and shoes, did you want anything? But they said:
Nothing. Then said he unto them: But now he that hath a purse, let him take it, and likewise a scrip; and he that hath
not, let him sell his coat, and buy a sword. For I say to you, that this that is written must yet be fulfilled in me:
And with the wicked was he reckoned. For the things concerning me have an end. But they said: Lord, behold here are two
swords. And he said to them, It is enough. \flright{\itshape Luke 22:7-38}
\end{multicols}

\subsubsection*{The Temptations in the Desert}
\begin{multicols}{2}\small
And Jesus being full of the Holy Ghost, returned from the Jordan, and was led by the Spirit into the desert, for the
space of forty days; and was tempted by the devil. And he ate nothing in those days; and when they were ended, he was
hungry.

And the devil said to him: If thou be the Son of God, say to this stone that it be made bread. And Jesus answered him:
It is written, that Man liveth not by bread alone, but by every word of God.

And the devil led him into a high mountain, and shewed him all the kingdoms of the world in a moment of time; and he
said to him: To thee will I give all this power, and the glory of them; for to me they are delivered, and to whom I
will, I give them. If thou therefore wilt adore before me, all shall be thine. And Jesus answering said to him: It is
written: Thou shalt adore the Lord thy God, and him only shalt thou serve.

And he brought him to Jerusalem, and set him on a pinnacle of the temple, and he said to him: If thou be the Son of God,
cast thyself from hence. For it is written, that He hath given his angels charge over thee, that they keep thee. And
that in their hands they shall bear thee up, lest perhaps thou dash thy foot against a stone. And Jesus answering, said
to him: It is said: Thou shalt not tempt the Lord thy God.

And all the temptation being ended, the devil departed from him for a time.

\flright{\itshape Luke 4:1-13}
\end{multicols}

\subsection*{Day 7: The Road to Emmaus}
The state of the risen Jesus Christ is the goal and the hope of the path of destiny of humanity. It is the most perfect
ideal of which one could ever think or dream. For it unites the most far reaching hopes of the noblest ideals of this
world with the highest and deepest ideals of the world beyond.

\textbf{The Good Communion}

And it happened that, while he was with them at table, he took bread, said the blessing, broke it, and gave it to them.
With that their eyes were opened and they recognized him, but he vanished from their sight. (Luke 24:30-31)

\begin{quotationx}
It was always through a sign that the Risen One let himself be known. For the form and appearance of the Risen One
differed from Jesus of Nazareth through the fact of being timeless and ageless. His countenance was simply the
expression of his being—of his spirit and his soul—and could only be
recognised by those who through the sheath of his bodily self had known something of his true being, his soul, and his
spirit. And the signs by which the Risen One let himself be known were such as to give a pointer and an indication
toward their personal earlier experience and knowledge of the soul-spiritual being of Jesus. They were reminded of
their earlier experience and knowledge of the being of Jesus Christ. \begin{flushright} \emph{Covenant of the Heart}\end{flushright}

\end{quotationx}
The Risen One did not appear in the likeness of Jesus, as those nearest to him had known him immediately before the
Crucifixion and earlier; that is, not as the resuscitated or reawakened Jesus who had been crucified, nor the Jesus who
had been baptised in the Jordan—but Jesus in a new form. For this reason, those people who had
known him recognised him only by some intimate sign.

Thus, Mary Magdalena, who at first took him for a gardener, recognised him only by the manner in which he pronounced her
name “Mary.” Thomas recognised him when the Risen One showed him the marks of his wounds. \emph{The two disciples from
Emmaus knew him in the breaking of bread}.

\begin{quotationx}
The heart perceives diverse presences as impressions and nuances of spiritual warmth. It is thus that the hearts of the
two disciples going to Emmaus recognised the One who went on the way with them before their eyes and their
understanding did, and who said to one another after their eyes opened and they recognised him: “Did not our hearts
burn within us while he talked to us on the road, while he opened to us the scriptures?” The heart burning in diverse
ways — this is the kind of “vision” and spiritual knowledge which is proper to the heart. \begin{flushright}
\emph{Meditations on the Tarot. Letter XII: The Hanged Man}\end{flushright}

\end{quotationx}

\textbf{The Evil Communion}

\begin{quotationx}
Tradition lives not thanks to organisations, but rather in spite of them. One should content oneself purely and simply
with friendship in order to preserve the life of a tradition; it is not necessary to entrust it to the care of the
embalmers and mummifiers par excellence that organisations are, except for the one founded by Jesus Christ. \begin{flushright}
\emph{Meditations on the Tarot. Letter XX: Judgment}\end{flushright}

\end{quotationx}
The mummification of the Pharaohs in ancient Egypt.

\begin{quotationx}
There will come a time when it will be seen that in vain have the Egyptians honoured the deity with heartfelt piety and
assiduous service; and all our holy worship will be found bootless and ineffectual. For the gods will return from earth
to heaven; Egypt will be forsaken, and the land which was once the home of religion will be left desolate, bereft of
the presence of its deities. This land and region will be filled with foreigners. . . In that day will our most holy
land, this land of shrines and temples, be filled with funerals and corpses. \begin{flushright} \textsc{Hermes Trismegistus:} \emph{Asclepius}\end{flushright} 

\end{quotationx}
\textbf{Bible Passage}

\begin{multicols}{2}\small
And behold, two of them went, the same day, to a town which was sixty furlongs from Jerusalem, named Emmaus. And they
talked together of all these things which had happened. And it came to pass, that while they talked and reasoned with
themselves, Jesus himself also drawing near, went with them. But their eyes were held, that they should not know him.
And he said to them: What are these discourses that you hold one with another as you walk, and are sad?

And the one of them, whose name was Cleophas, answering, said to him: Art thou only a stranger to Jerusalem, and hast
not known the things that have been done there in these days? To whom he said: What things? And they said: Concerning
Jesus of Nazareth, who was a prophet, mighty in work and word before God and all the people; and how our chief priests
and princes delivered him to be condemned to death, and crucified him. But we hoped, that it was he that should have
redeemed Israel: and now besides all this, today is the third day since these things were done.

Yea and certain women also of our company affrighted us, who before it was light, were at the sepulchre. And not finding
his body, came, saying, that they had also seen a vision of angels, who say that he is alive. And some of our people
went to the sepulchre, and found it so as the women had said, but him they found not.

Then he said to them: O foolish, and slow of heart to believe in all things which the prophets have spoken. Ought not
Christ to have suffered these things, and so to enter into his glory? And beginning at Moses and all the prophets, he
expounded to them in all the scriptures, the things that were concerning him. And they drew nigh to the town, whither
they were going: and he made as though he would go farther.

But they constrained him; saying: Stay with us, because it is towards evening, and the day is now far spent. And he went
in with them. And it came to pass, whilst he was at table with them, he took bread, and blessed, and brake, and gave to
them. And their eyes were opened, and they knew him: and he vanished out of their sight. And they said one to the
other: Was not our heart burning within us, whilst he spoke in this way, and opened to us the scriptures?

And rising up, the same hour, they went back to Jerusalem: and they found the eleven gathered together, and those that
were staying with them, saying: The Lord is risen indeed, and hath appeared to Simon. And they told what things were
done in the way; and how they knew him in the breaking of the bread. \flright{\itshape Luke 24:13-35}
\end{multicols}

\flright{\small\textit{Posted on 2023-04-02 by Cologero}}