\section{Knowledge and Love}

Originally published in 2006.

Jack and Jill were playing and fetching water one day, when Jack had this sudden realization:

``I love you'', he said to Jill.

This is the essence of metaphysical intuition, not in the sense of a gut feeling or hunch, but of a direct and positive realization. Direct because it is not mediated --- it is an immediate realization. Positive because it is absolutely certain knowledge. Note, too, that this example demonstrates that knowledge = being. Jack's knowledge of his love is identical with a change of his being from ``not loving'' to ``loving''.

Notice how the ``modern'' mind looks at knowledge:
{\small
\begin{longtable}{p{.3\textwidth}p{.6\textwidth}}
	\toprule
	I think I love you. &
	This is Jack the philosopher. It's an interesting theory, but Jill is not impressed.\\
	
	I believe I love you. &
	Jack may have reasonable grounds for his faith, but Jill has her doubts.\\
	
	I get this funny feeling in my stomach when I'm with you. &
	No, Jack, that feeling is not your stomach. Don't confuse lust with love.\\
	
	My E-17 neuron in the left cortex has released 7 mg of endorphins. &
	The materialist. What Jack calls ``love'' is simply an electrochemical event in his brain. ``Love'' is folk psychology and such talk will give way to scientific descriptions just as astrology has yielded to astronomy.\\
	
	You remind me of my mother. &
	Does anyone still read Freud? Did anyone truly believe it?\\
	
	You look like sound genetic material. &
	Of course, the biologist tells us the selfish genes want to reproduce, even if only half of them get a chance. ``Actually, Jill, Judy over on yonder hill has good genes, too, but her father won't let me near her.'' And Jack came tumbling down. Wonder why?\\
	
	I don't want a power relationship. &
	``Sensitive'' Jack has been reading postmodern literature on ``love'' as a disguise for patriarchal oppression of women. This is what we mean by ``mediated''.\\
	\bottomrule
\end{longtable}}

\flrightit{Posted on 2011-04-11 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Zmaj on 2011-04-12 at 00:58 said:}

One more:

``I love you Jill, by the will of my will'' -- Jack the magus. Will it last?

\end{sffamily}
\end{footnotesize}

