\section{Worldviews, Representations and Reality}

A few years ago we wrote Privation\footnote{\url{https://gornahoor.net/?p=1815}}. Now is the time to revisit that idea, by expanding it with some of \textbf{Hermann Keyserling}'s conceptions. As we pointed out then, the world is a projection of the I. This can be made clearer from Keyserling's claim that ``the representation creates reality, not vice versa.'' (We will use the terms ``representation'' and ``worldview'' somewhat synonymously, although the latter has wider application.)

If that sounds like a wild claim to you, then consider its opposite: how exactly does reality create a representation of itself in the mind or consciousness? Although most people naively consider that to be obviously true, including the soi-disant ``brights'', a moment of reflection should convince you otherwise.

This is the question that Kant tried to answer. Shankara's answer\footnote{\url{https://gornahoor.net/?p=7326}} is that the mind fabricates an illusory, which we then take to be a true model, of the world. As we recently pointed out, \textbf{Johann Fichte}'s philosophy has a general similarity to the Advaita Vedanta see \textbf{Bhagavan Das} ; hence, we can use his insights to explain this idea more fully. \textbf{Bryan Magee}, in his book on Schopenhauer, summarizes Fichte's philosophy like this:

\begin{quotex} 
Fichte saw the whole empirical world as deriving from the subject and taking the precise and detailed form it does out of logical necessity. However, since the world derives from out conception of it, and not the other way about, our conception of the world must itself be a free creation on our part---`free' in this context meaning precisely that it has not empirical derivation.
\end{quotex}

Furthermore, this applies equally to the human, moral world---which is our main interest---as well as the physical, empirical world. Hence, we will discuss each in turn.

\paragraph{Science as a Creative Activity}

In this section, I will be employing some basic symbolic logic statements, for those trained to understand them. If that gives you \emph{agita}, I believe the text by itself should convey the main point.

A scientific theory can predict the occurrence of a set of facts; this used to be called ``saving the appearances.'' Symbolically,

$T \Rightarrow F$

where $T$ is the theory and $F$ represents the set of facts, $F^1, F^2$, ... that it predicts.

In words, the theory logically implies the occurrence of the facts. The more facts it can predict, the more useful it is. However, the theory itself can never be proved to be ``true''. That is because there is no logical, computable, deterministic, mechanical, or empirical method to go from the facts F to the theory T. Since the contrapositive of a proposition is equivalent to the proposition, the most we can say is that the theory can be falsified if a predicted fact turns out not to be true. That is,

$\sim F^n \Rightarrow \sim T$

Hence, in practice, there can be no ``computer program'', for example, that will generate true theories based on a collection of facts. So what then is a ``theory'' if it is not empirical and not derivable from anything empirical? A specific example will illustrate this.

For more than 200 years, educated people accepted \textbf{Isaac Newton}'s theory of universal gravity as an absolute truth since it accounted for the movement of the planets and gravitational attraction on a human scale. However, \textbf{Albert Einstein}'s theory eventually disproved Newton's since it predicted some facts that the latter theory could not. If Newton's theory was not a true law of physics, and since it does not logically follow from the facts, then what exactly is it? The only answer is that it was a creative and free construction of the human mind, just as Fichte described; in other words, it is an ideal.

Since the ideal (or theory) implies the facts, we are forced to the conclusion that the real derives from the ideal, and not the other way around. Once we include the technological methods that derive from the theory, we see that this is even more true.

\paragraph{The Moral World}

Fichte's other insight was that not just the physical world, but also the moral world, derive from the subject. Magee summarizes this position:

\begin{quotex}
Both worlds derive from free activity on the part of the noumenal self. This self can thus be said to create both worlds: and this provides us with the key to the nature of total reality. ... For morality to be possible there must be choice, and for choice to be a possibility for me, it is necessary that something should exist other than my self. Similarly, for moral action to be a possibility for me there needs to be some challenge, something that exists in opposition to my self, or at least something that is a potential obstruction to my activity. So if I am to be a moral being at all it is necessary that there should be a world which is not me, a world of objects which can obstruct me.
\end{quotex}

In analogy to the physical world, the representation of the world, or worldview, will determine the phenomenal world out of its logical necessity. The physical world appears more restrictive in that there are few effective theories or representations of it. This does not include the possibility of miracles, siddhis, magic, or other supernormal powers to affect the physical world. At the least, the notion that the representation creates reality does not preclude that possibility.

When it comes to the body, the situation is even more convincing. No one today doubts the possibility of psychosomatic phenomena, so our thoughts affect our state of health to a large degree. In the human, moral realm, our world representation is the only thing that matters. People have a conception of the human world, and will then act in accordance with that conception. This will create a world that conforms to that conception. When groups of people have different conceptions, conflicts will result. The dominant worldview, then, can only be the result of the Will to Power.

This competition of worldviews is thus the battlefield for moral action. There is a lot to be found in print analyzing this battlefield, some of it well written. However, such commentaries seldom make the move from the real to the ideal. Specifically, to repeat, the facts do not create the worldview, but the worldview creates the facts. Hence, the real issue is to understand whence these worldviews arise. This worldviews are actually illusions; they seem to account for the facts, yet there are facts that would falsify the worldview. However, worldviews are very difficult to discard, once established in the mind. Unlike mistakes in our understanding of the physical world, which are very quickly noticeable, mistakes about the human world take much longer to show their full effect, oftentimes even over a few generations.

The source of worldviews arises from inner states; there are various possible combinations of psychological forces that give rise to specific worldviews. This is not the place to discuss that, but details will come out over time.

In preparation for his battle, \textbf{Arjuna} had to undergo a process of spiritual development. Only then will the true nature of the battle be understood. Arjuna would be stuck in illusion until God-realization would reveal the true nature of creation. This, then, needs to be our first step.


\flrightit{Posted on 2014-08-04 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Anthony on 2014-08-06 at 14:38 said:}

I'm still confused. Naive realist would argue that creating a Theory that makes a Prediction ``all apples are cubical'' does not make it be so in Reality (consult nearest apple for Fact). Here T ? P and R ? F ? $\sim$ P ? $\sim$ T, i.e. Reality falsifies the Theory, which is unable to shape Reality. What's wrong with this? How does here the Real derive from the Ideal?

The process of ideas -> tendencies -> acts -> facts as explained somewhere on TraditionInAction pages was more intuitive to me.

\hfill

\texttt{Cologero on 2014-08-06 at 15:21 said:}

Yes, Anthony, that does sound confused. What is a theory and how does it derive from reality?

\hfill

\texttt{Anthony on 2014-08-07 at 01:15 said:}

Theory is a model approximating reality. For example, Metzinger robots generate a model from set of facts

\url{https://www.youtube.com/watch?v=hh20JMEI-FI}
Here is all creative part reduced to mere computation.

PS! Those question marks in my previous comment should read as implications.

\hfill

\texttt{Cologero on 2014-08-07 at 08:57 said:}

I'm afraid, Anthony, that you have understood nothing. You don't know how to engage in a dialog, you don't understand the laws of logic, you don't know what the word computable means, and you don't know what a model is. So, you are on moderation until you can address those issues. For your benefit, note these points:

1) In a dialog, you select a specific point that was made and comment on it. Don't expect me to look at irrelevant links.

2) The is no law of logic that proves a premise from a true conclusion. Specifically, if p => q, and we assert that q is true, then it tells us nothing about p.

3) A computable process is an algorithm that will complete in a finite number of steps. There is no such algorithm to go from q to p in the example above.

4) ``Approximation'' is another word for ``false''. We already made that point. Viz., although Newton's theory gave good empirical results, it is a false theory. So what? And Newton did not use a computer, so where did his theory come from?

5) A model requires two things: the model and the thing modeled. For example, when I was a boy I used to build scale models of aircraft and motor vehicles (I guess this is no longer a popular hobby). I could compare the model to the real thing, understand the scale, and determine how realistic the model was. You say a theory is a ``model''. But if you knew ``reality'' as the thing modeled, you wouldn't need the model, would you?

Apparently your point is that there is no true knowledge, only various mathematical ``approximations'' that seem to predict phenomena in limited circumstances. And you haven't even addressed the human world. What is your theory of human life and society? Can your robots describe a soul or tell us its destiny? I believe you stumbled upon the wrong web site.


\hfill

\end{sffamily}
\end{footnotesize}

