\section{The Locust Conspiracy}

In which we explore the outer reaches of scientific knowledge including psychopathy, time travel, entropy, quantum physics, conspiracy theories, and miracles.

\paragraph{Definitions}
An \textbf{opinion} is a proposition that is neither demonstrably false nor self-contradictory. Hence, it is the least reliable form of knowledge; it may be falsified in the future. Emphatically, an opinion is not any thought that pops into your mind.

A \textbf{belief} is a proposition that is actionable, that is, it will lead to action in the world. If I believe it will rain on Saturday, then I won't pack the picnic basket. That is how you can tell if someone really believes what he claims.

\textbf{Not Even Wrong} is an argument that is neither correct nor incorrect because its premises are so off base or the argument is confused. For example, Amalric of Bene in the 12th century was saved from the charge of heresy because the investigators determined that his views were pure lunacy.

The opposite of a false proposition is a true proposition. The opposite of a ``not even wrong" proposition is still false.

\textbf{Psychopaths} and \textbf{Sociopaths} have a lack of conscience yet can appear charming to the unwary. Since they are manipulative, they can do very well in achieving power in politics or business. It amazes me how few people can recognize a psychopath. Haven't they ever wondered how some people, usually not as competent or intelligent as they are, manage to rise above them?

It is pointless to argue with a psychopath, since they get a thrill out of irritating the lesser beings below them. They even boast about it. Also, the charge of hypocrisy against them is ineffective. Quite the contrary. They revel in ``getting away with it" and even love their hypocrisy to be made public. They often reveal themselves with a wry smile, usually at inappropriate moments.

A \textbf{conspiracy theory} is an explanation that relies on a secretive cabal of sinister and powerful groups working together to achieve a result, often over the course of generations. Obviously, it is hard to prove without being one of the insiders. The irrational belief in a conspiracy theory is considered to be a mental illness: viz., \textbf{illusory pattern perception}. the Wikipedia article \textit{List of conspiracy theories}\footnote{\url{https://en.wikipedia.org/wiki/List_of_conspiracy_theories}} includes the belief in a \textbf{white racist patriarchy} to be one such theory.

The \textbf{Open Conspiracy}\footnote{\url{https://forcingchange.wordpress.com/2012/01/16/advancing-the-open-conspiracy-h-g-wells-and-the-world-state/}}, envisaged by H G Wells, offers a better explanation. Its tenets are being implemented today, but not by a secret cabal, but rather openly by people and groups implementing the policies independently. The Biblical explanation should be sufficient:

\begin{quotex}
The locusts have no king yet they fly in formation. (Proverbs 30:27) 

\end{quotex}
Keep in mind that there is nothing great about locusts.

\paragraph{The Missing W}
In one of his books, Richard Dawkins created a thought experiment in which he showed how the phrase ``METHINKS IT IS LIKE A WEASEL" could be generated by tossing some tiles and selecting those that would lead to the phrase. But suppose the letter W was missing from the tiles; then, no number of tosses would create the phrase. The same notion can be applied to the evolution of the human being. Humans could not have ``evolved" from matter unless that there was that possibility. That is, the biological human is the W tile, what we have been calling a possibility of manifestation. The goal has to be baked into the entire world process.

\textbf{Reductionism} is the idea that more complex systems can be explained by simpler systems. That requires that the elements of the simpler system interact in a certain way to create the complex system. But that interaction is the very definition of the complex system and was not predicted from the simple system.

Theories of everything or the big bang theory have predicted nothing significant. They cannot even predict the daily activity of a small city.

\paragraph{Social Studies}
Actually, life in the city can be understood to a large extent, if you really try. But the effort will make you few friends.

\paragraph{The Laws of Physics}
There are no laws of physics that need to be obeyed. As Albert Einstein admitted in the Evolution of Physics, a theory is the free creation of the human mind. Newton's theory of gravitation was his creation, yet is certainly false. Einstein's theory of gravity is Einstein's gravitation; better than Newton's, but most physicists don't believe it is really true.

A sonnet follows the convention of 14 lines with a particular rhyming pattern. A scientific theory has its own convention: it has to ``save the appearances", i.e., it explains the phenomena we experience. It is not falsified, that is, no phenomena had refuted it. Ideally, it should predict future phenomena. Relying on ``chance" hardly qualifies as an explanation.

\paragraph{The Theory of Evolution}
Neo-Darwinism = Darwinism + genes. That is the proper name, not the ``Theory of Evolution", because nothing evolves. One species disappears, another appears. A proper theory will explain in detail the entire process. Neo-Darwinism may be true, but not as it is currently understood.

It is like attending a seminar on achieving financial independence, and learning that the secret is to play the slot machines at the casino. Certainly, one of the attendees might hit the jackpot at some point, but that hardly makes the advice very sound.

\paragraph{The Tunnel}
Imagine that every time you drove your automobile through a tunnel, you didn't know where you would
come out on the other side, because you would create a diffusion pattern. You should be grateful that you don't follow the ``laws of physics".

\paragraph{Time Travel}
The equations of physics work in both time directions. The equation is the same whether the particle is moving forward in time or backward in time, so it is hard to tell which is which. I am pretty sure that I could tell which direction I am traveling. For example, were I going backward in time 20 years, I would be much healthier. So much for that law of physics.

\paragraph{Unscrambling an Egg}

The one exception is the second ``law" of thermodynamics, which states the entropy is always increasing. It means that disorder is increasing, so things can't be getting ``better and better". So we should be able to determine the direction of time by measuring entropy.

However, at least one physicist has a better explanation: what we call order and disorder are simply different states of the system. So if I break an egg and scramble it, the ``egg in the shell" and the ``scrambled egg in the pan" are just two different states of the egg. You may have noticed that no two scrambled eggs look exactly alike. You can scramble eggs, probably daily for the next hundred years, and no two will be exactly alike. Nevertheless, the scrambled state is much more likely than the egg-in-the-shell state.

The shell state always a possibility, and from the standpoint of physics, neither state is more privileged than the other. Just as the particle in motion may be going backward in time, the scrambled egg could return to the shell state without violating any law of physics.

\paragraph{Miracles}
This gives us a better understanding of miracles. A miracle is the manifestation of a not very likely state of matter. We are deluded into thinking that our common experience is somehow necessary, whereas it is merely the more likely scenario. It is not contrary to any ``laws" of physics for a system to be in one state rather than another.

For example, it is unlikely that you will win the lottery, but it is very likely that someone will win. The winner was fortunate to be in the most unlikely state of being.

\paragraph{Objective Reality}

For some reason, we think that our experience of the world out there, right now, is ``objective" while our inner life is ``subjective". The opposite is the case. Our knowledge of the external world is a matter of opinion, subject to change. We create a representation of the world, which may be called the ``collective conscious". Whatever doesn't fit is our unconscious; e.g., black holes, the centre of the earth.

On the other hand, we are certain of our inner experiences, at least our conscious experience, even if much of our inner life is unconscious.

\flrightit{Posted on 2020-09-02 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{J. R. R. Tolkien on 2020-09-02 at 03:58 said: }

Cologero.

This regards something other than the topic here, Do you know anything about the order created by Peladan that is active in this very day?

Namely, the Kabbalistic Order of the Rose-Cross. Do you know if they have a valid initiation? It seems te me they are frauds… But they claim that they have rites that are primordial. The order continued after Peladan with Papus and de Guaita, but then the clergy of a Gnostic church sect became their leaders, today's leader is a 33 degree Freemason who promates Thomas Jefferson's Bible that exclude the divinity of Christ, and removes all miracles, reducing the Christ to a spiritual advisor. The new leader of theirs also promotes humanism in his site. Maybe I am wrong about them being frauds?


\hfill

\texttt{Cologero on 2020-09-02 at 05:11 said: }

Why don't you just spend the \$299 for your initiation and let us know?


\hfill

\texttt{J. R. R. Tolkien on 2020-09-03 at 00:52 said: }

I will do, and once I am initiated, I will make my own order to be able to initiate others for a cheap price, because that's just too much money for the poor people who can't afford it. :((


\hfill

\texttt{J. R. R. Tolkien on 2020-09-03 at 01:00 said: }

``Meyrink was certainly in the current of very numerous traditional ideas, especially of a Jewish source; but, not to mention that this presupposes precisely an initiation" Letter VII to Evola, Guénon

According to Guénon in an article here, to have traditional ideas presupposes initiation? 

1. The person has traditional ideas in a period where people are tending against it.

2. Therefore this person has received a proper initiation.

Wouldn't this mean that initation doesn't require attachment to a true tradition, if it's just a matter of ideas? Is Guénon in opposition to what he wrote before, or is this my incomprehension?


\hfill

\texttt{burningfireeye on 2020-09-03 at 11:22 said: }

Since you talked about psychopaths here, I wanted to think about what it would be for a psycopath to become a Christian.
I know a Christian apologist who is himself a psychopath, his name is David Wood, and here's a
video of him explaining what it is to be a Christian psychopath: \url{https://www.youtube.com/watch?v=isxA4qvLUwg}

What David did not to say is that psychopaths do not recognize feelings towards evil doings, but they certainly recognize chaos, they recognize that blind yearning towards more and more material things can never satisfy the will, even the ``pleasure" of killing people is chaotic, they can realize that if everyone was like them, there would be no more order, and this is one of the things that cause anxiety in them, another thing that causes horror in them is nihilism, the world of chaos, the world that makes no sense. They can recognize that we are ultimately in need of order, and they can recognize God as the supreme order, the supreme Good, and Beauty, and in this they can realize that it is important not to harm others, so that they too can be in harmony with the cosmos. They are able to control their actions, but they don't have feelings like normal people have, which means they can be responsible for an evil action, they know what is an evil action as a normal person, they know it leads to chaos and insanity, but they don't feel guilt in the same way, which then makes it harder for them. 

Now modernity especially is a period where people are very sentimental and compassionate, whereas in others periods, right reason was the rule rather than emotion, so they can understand that killing a being is evil because life has an intrinsic value, and will not do it because of right reason, and not because of emotion, this they would be able to assimilate and understand. Normal people also have evil thoughts in their minds all the time, they are capable of doing many terrible things, the only thing stopping them is fear, or because it is ``disgusting". Everyone can be disciplined to learning to do good and avoiding evil, profound change of conscience is possible, also supernaturally the person can be changed by forces that are not merely natural, if they seek the Above.


\hfill

\texttt{burningfireeye on 2020-09-03 at 11:29 said: }

The person can be changed through subtle forces of nature that are not merely natural. As is the case of acupuncture, and other forms of medicine.


\hfill

\texttt{burningfireeye on 2020-09-04 at 00:17 said: }

Again, it is very easy for a psychopath to become egoistical, to become liars and deceivers, because they don't have any bad feelings towards others, their conscience does not accuse them. But it is not true that they are driven towards evil, or that it is natural for them to feel pleasure in killing, this is an acquired pleasure. Again, normal people have thoughts of killing others and probably would kill if no one was around, and if it wasn't so dirty and smelly, so it is the same, they have also these thoughts, but without the remorse, without any weights in their conscience.

But they are guilty of their actions, the maxim ``do not unto others what you would not want done to yourself" is obvious, they know that when they commit these actions against others they wouldn't want it done against themselves, and they know they are disrupting the common order and good, that's why they can repent their sins like David Wood, they would rather believe in God and in the supreme order of the universe than to believe in chaos and misery, that when they abandon that Eternal Reason, things become meaningless.


\hfill

\texttt{burningfireeye on 2020-09-04 at 00:31 said: }

It is also possible for a normal person to become as a psychopath, which is a sociopath. The world is guided by forces beyond the physical realm, and the psychology of today cannot explain much of the psychological phenomena. In the world of today, people are leaning that chaos is natural, they learn that the unnatural is healthy, and that they are an evolved form, they are in battle against nature, in this world it would seem impossible for a psychopath to become a Christian, because given that it is extremely hard for a normal person to accept Christianity which is being shown as against progress and evolution, it would be impossible for the one born with the dispositions, but that is not exactly true, in the first place, it is very difficult to diagnose someone as being a psychopath, just as it is difficult to diagnose any normal pathologies for people with normal intelligence, that is because the forces that act in the human being are obviously not just physical, but for a person to understand that is very hard, because the scientists don't want anyone to think in any way different from the way they themselves think, it is the danger of excommunication that emprisons many people's conscience.


\end{sffamily}\end{footnotesize}