\section{Good Vibrations}

When your crazy new age friend tells you that everything is vibration, she is not wrong. But when your hard headed scientific friend tells you that ``miracles'' are impossible because they ``break'' the laws of physics, he is likely wrong.

Planck's equation shows that energy and frequency are directly related; higher frequencies carry more energy. Einstein's famous equation $E=mc^2$ also shows the equivalence of energy and mass. According to quantum mechanics, particles exhibit wave behavior, so that mass is also associated with a wavelength. These are called matter waves\footnote{\url{https://en.wikipedia.org/wiki/Matter\_wave}}.

These waves are easily observed in small particles like electrons. They have even been observed in some large molecules. There are not two sets of laws of physics. If wave-particle duality is true at the quantum level, it is true for the larger objects of our everyday experience.

Phenomena like bilocation, walking through walls, and disappearing do NOT break any laws of physics, as the attached video shows. If we don't witness them often, that is because the components of large objects do not vibrate coherently with each other. When I was a boy, I was fascinated by this topic and used to calculate the frequencies of everyday objects, including humans. I did not take into account that there is no physical definition of a human beings, unlike subatomic particles and even molecules.

So the liver has a frequency that is different from that of the heart, etc., not to mention all the free molecules. The point is that a human being is a mixture of vibrations, so that quantum phenomena cannot occur. That does not mean that is impossible, but just that it is unlikely. A human body is just a possible state of matter, and the body that walked through a wall is another state.

Unless you think that the laws of physics do not apply everywhere, large scale quantum effects are possible. If you think that they have never been witnessed, that is because you deny the study of anomalous events. However, the Cosmos is very large and very old, not to mention that there may be a great many of them. We don't know everything that may have happened.

Metaphysically, there is no distinction between the possible and the real; if it is possible, then it has or will happen. Suppose, just suppose, that a being has control over his corporeal state, so that it vibrates at a single frequency, much like a laser. Then quantum effects can occur.

\url{https://www.youtube.com/watch?v=rlTYHRDuxrY}

\flrightit{Posted on 2022-12-06 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{veritycacciatrice on 2022-12-07 at 09:31 said:}

``Metaphysically, there is no distinction between the possible and the real; if it is possible, then it has or will happen. Suppose, just suppose, that a being has control over his corporeal state, so that it vibrates at a single frequency, much like a laser. Then quantum effects can occur.''

Let's suppose your boomer friend has thrown down yet another trad.physics gauntlet and you don't have time to address every single point, though you would dearly love to, because you can.

One has control over their corporeal state to a far greater degree than average Man, exoteric church, state or medicine would accept, and in fact it is precisely with alignment of `AVM' that such control is realised.

We're all friends here, we all know AVM is a charge which, when applied through properly resistant purifying processes ignites the chrism.

Plasma physics answers more questions than modern scientism can even coherently ask. Why so resistant to an electrically unifiable/unified perspective?

\url{https://www.youtube.com/watch?v=mYtXSVOV60w}


\hfill

\texttt{Cologero on 2022-12-07 at 22:56 said:}

\begin{quotex}\footnotesize
I saw don Genaro disappear in front of my very eyes as if a wind had swished him away. His leap or whatever he had done had had such a profound effect on me that I felt as if his movement had ripped something in my intestines. 
\flright{\textsc{Carlos Castaneda}}
\end{quotex}

\hfill

\texttt{De Bosit on 2022-12-08 at 07:32 said:}

``Metaphysically, there is no distinction between the possible and the real;''

*taps random spot on desk* just had to nuke one of my prison towns. Okay do I have enough karma kredit now?

Ok so, the thing is it's a double edged sword. You can imagine yourself with lots of power and stuff, but at the same time you ignore maybe that you are unconcious cattle being driven to the slaughter, dreaming of being human, just one hard knock on the door and you wake up, grunting.

Thankfully we can transpose scenarios on less dangerous outcomes, like if I imagine throwing thunderbolts, I could to keep this statement true just take credit for some anonymous thunderbults, or just assume it's true.

When actors play a character very well, they lose touch with reality. What we see on the screen, they actually ARE. Then they have to go home and Tony Soprano has to pretend to be a nice guy again. No wonder he had issues.

Just imagine the angel beaucracy needed to answer all the uesless prayers for gimme gimme gimme and stuff. It has to be made real on some plane or level. If you pray for gold, you might get golden gummy bears, and be content that you got an answer, or remain stubborn until you find yourself enslaved in a gold mine, close to the gold, but never owning it.

\hfill

\texttt{Dimitri on 2022-12-08 at 12:03 said:}

Is it worth studying physics as someone interested in Tradition, Cologero? If so, where would you recommend starting? Wolfgang Smith, perhaps? Any thoughts on the book Quantum Reality by Nick Herbert? Gornahoor appears to be sympathetic to Intelligent Design as well. What do you make of that?

I strongly get the impression that due to the mendacious nature of the the scientific establishment and their apparent commitment to materialistic scientism that there is much more to these matters than what they tell the general public. They project far more confidence and certainty than is warranted. Decent article on this:

\url{https://apexsnotes.substack.com/p/dont-trust-the-science}

I'm also planning on getting into Rupert Sheldrake's material soon.

\hfill

\texttt{Kaukomieli on 2022-12-08 at 12:43 said:}

If there exists a what could be called a buddhic field, then the person attempting superhuman feats must realize that beforehand and then put the other ``bodies'' in complete harmony and unity; reason must be intact, emotions calmed, desires released, and the etheric body and sthula sharira put to serve a single goal of moving the whole being in the infinite field of consciousness apparently outside the limited, diminutive corporeal being. Before that Captain Farebanks does not succeed even if he tried a million times.

\hfill

\texttt{Cologero on 2022-12-08 at 19:59 said:}

@Dimitri. I am interested in science because I was trained in science. As some recent shows have shown, physics deals with the ``gross state'', which is a small part of total reality. Within that domain, it have been somewhat effective. My goal is to show that there is nothing in Tradition that is inconsistent with physics. To do science, you need to start at the beginning. It is long, difficult, and requires commitment. I won't get involved with various conspiracies. I owe my life to medical technology and effective pharmaceuticals. If drugs fail from time to time, so be it. That's life.

\hfill

\texttt{Dimitri on 2022-12-09 at 15:14 said:}

Based on my limited knowledge I've seen that Einstein's physics and quantum physics are not exactly compatible. W. Smith rejects the former. Others seek a unifying theory. What do you make of Einstein?

\hfill

\texttt{Kaukomieli on 2022-12-10 at 13:08 said:}

@Dimitri:

``What do you make of Einstein?''

His mathematical metaphycis take us to a Kabbalistic world where ``spirit and matter'' are bridged to make a dimensional apparatus of a numerical and purely quantitative kind. It is truly a sad thing that the divide of Aryan and Judaic physics -- and conceptions, politics, art and cultural differences, and economic measures -- couldn't be bridged without the atomic bomb and the cruelest war in history. What could be more convincing than the atomic bomb?


\hfill

\end{sffamily}
\end{footnotesize}

