\section{Dialectics for Dummies}

\begin{quotex}
The dialectician is a kind of buffoon: one laughs at him, one does not take him seriously.
\flright{\textsc{Friedrich Nietzsche}, \emph{Twilight of the Idols}}
\end{quotex}

The Enlightenment ideal of ``Free Enquiry'', as opposed to the dogmatism of the Middle Ages, was supposed to bring in a new era of human freedom and knowledge. Rational minds would debate in the ``marketplace of ideas'' and the world would flock to the best idea with the same enthusiasm as they might flock to the better mousetrap.

It has been remarkably effective in the physical realm, but a mixed blessing in the human realm. Instead of the proverbial ``world peace'', free enquiry often has had a corrosive effect on societies by destroying their spiritual foundations without providing a more secure replacement.

Now that every conceivable idea has been discovered and analyzed, there is nowhere else to go. Discussions have deteriorated to kayfabe debates\footnote{\url{https://en.wikipedia.org/wiki/Kayfabe}}; onlookers watch the debaters, cheering on their favorite personality with no regard to facts or reason.

Those more skillful in dialectics have memorized the arguments and counter-arguments several layers in, much like the way chess players memorize a dozen moves for each possible opening move. Only a few are capable of carrying on the debate beyond that point.

No consequential debate has ever been fully resolved. Yes, some may appear so for a season, but not for eternity.

\paragraph{Getting Paid}

At one time, I was considering becoming a corporate speaker. I had developed a local reputation as an interesting and compelling speaker so I wanted to take it to the next level. I therefore attended seminars at the Florida Speakers' Association. They certainly did not teach me to develop excellence in dialectics and rational thinking. Here are the secrets that professional communicators revealed to me; keep them in mind if you aspire to attract 100K subscribers to your youtube account.

\subparagraph{Stories}

Apart from mathematicians, no one remembers arguments. Stories, however stick in people's minds well after the speech. The story may be funny, or better, it should arouse anger, contempt, or pity. Symbols can serve the same purpose.

\subparagraph{Hominess}

One speaker admitted that his speeches had become too polished, sophisticated, and perfectly delivered. He learned to deliberately put mistakes into his talks or stumble every now and then. In that way, he discovered, he was able to relate to his audience more effectively.

\subparagraph{Comedy}

I vividly remember one very successful speaker at the Q\&A session following his talk. A wanna-be asked him, ``Does your talk need to be funny?''

He responded quickly, ``Only if you want to be paid!''

Keep that in mind the next time you prepare a talk. I've learned that if you are too subtle, then many people cannot tell if you are being funny or serious.

\paragraph{Solemn Contest}

\begin{quotex}
Philosophy on the nature of Plato would define it more as an erotic competition as a training and internalization of the ancient agonistic gymnastics and their prerequisites... What finally grew out of this philosophical Plato's erotic? A new art form of the Greek \emph{agon} Greek: solemn contest , the dialectic.
\flright{\textsc{Nietzsche}}
\end{quotex}

This is truer than you think. Rather than a disinterested search for truth through dialog, dialectics has become a new form of contest. Professional debaters, e.g., between a believer and an atheist, can earn five figure fees for their ``debates'', whose results are known in advance. That is why Las Vegas never sets odds on such debates.

\subparagraph{Joe}

Take Joe, for example, who was once a friend of mine on fb. He was a doctrinaire libertarian and a supercorrect Catholic, who was willing to take on anyone in a debate; he was sure to destroy his antagonist.

I never engaged him directly, but my Aikido style frustrated him so he unfriended me everywhere. I ran into him on another forum a year ago in which he announced that he had become, depressingly, a nihilist and no longer held any convictions. I suggested that he hold a mock debate between the two versions of himself. That is when he blocked me.

\subparagraph{Richard}

Richard claims to have a very high IQ. He wrote that he wants to preface his posts with: ``This is what I think, take me on!'' He admits that he takes pleasure in provoking others with his arguments. Everyone wisely avoids a direct confrontation.

\paragraph{No Other Choice}

\begin{quotex}
One chooses dialectic only when one has no other means. One knows that one arouses mistrust with it, that it is not very persuasive. Nothing is easier to wipe than a dialectical effect: the experience of every meeting, where it speaks, is proof of that it can only be self-defense, in the hands of such as no longer have other weapons.
\flright{\textsc{Nietzsche}}
\end{quotex}

The weak resort to dialectics whenever they lack real power. This tactic is used in political debates. The weaker side resorts to arguments and facts. They point out inconsistencies in their opponent's policies; the clamor about the hypocrisy in their opponents' behaviors. This is all for naught. If you want to win, you may have to push the issue, and let the other side resort to arguments. As Fat Tony says,

\begin{quotex}
Suckers try to win arguments, nonsuckers try to win. 
\flright{\textsc{Nicholas Taleb}}
\end{quotex}

\paragraph{Conclusion}

Whenever I describe a jam that I'm in, my sister naively suggests, ``Why don't you just say to him ... '', and you can fill in the blanks. Obviously, she has never had to deal with tough men. In such situations, there needs to be a carrot and a stick, even if only implicit and subtle. Only then can negotiations begin.

\url{https://www.youtube.com/watch?v=8wGS29G-gSs} 

\flrightit{Posted on 2022-02-08 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Balrog Booger on 2022-02-09 at 12:41 said:}

The writer Vox Day is an interesting case study of this phenomenon. He tries to outline a explicit difference between rhetoric and dialectic, and he prides himself on being above and impervious to rhetoric. After experiencing many of the problems with formal debates, he now refuses them, but instead of maintaining a dignified aloofness from debate, he just calls anyone who contradicts him stupid on his blog. He also spends a lot of time calling out what he calls ``gamma males'' for behaving exactly how he does.

An example he just posted yesterday:

\begin{quote}
You're flat-out wrong, Curt Doolittle . And I'm a lot smarter than you are, so you really should think twice before trying to ``correct'' your intellectual superiors.
\end{quote}

Not that he's all bad, by any means. I still frequently read his site, despite the silliness of his ego posting.

\hfill

\texttt{Santiago on 2022-02-09 at 19:41 said:}

mfw when truth literally sustains your world but nobody cares

\hfill

\texttt{Heiko on 2022-02-10 at 04:29 said:}

Oh yes, Vox Day and Curt Doolittle in one comment. I too once spent too much time engaging into such unfruitful discussions. If you change your approach to trying to understand the other side, asking questions etc. it has a way better effect on the other person and you yourself. Trying to answer the question, they might see how stupid their thoughts and arguments are.

\end{sffamily}
\end{footnotesize}

