\section{The Flowering of European Civilisation}

\label{sec:FloweringEuropean}

To the profound comprehension of this law of the intellectual generation of ideas, are due the marvels of Catholic civilisation. To that wonderful civilisation is due all that we admire and all that we see. Its theologians, even considered humanly, put to the blush modern and ancient philosophers; her doctors excite wonder by the immensity of their science; its historians by their generalising and comprehensive views, cast those of antiquity into the shade. St Augustine's ``City of God" is, even today, the most profound book of history which genius, illuminated by the rays of Catholicity, has presented to the astonished eyes of men. The acts of her Councils, leaving aside the divine inspiration, are the most finished monuments of human prudence. The Canonical, excel in wisdom the Roman, and the feudal, laws. Who is before St Thomas in science, St Augustine in genius, Bossuet in majesty, St Paul in power? Who is greater as a poet than Dante? Who is equal to Shakespeare? Who surpasses Calderon? Who, like Raphael, infused life and inspiration into the canvas? 

Place people in sight of the pyramids of Egypt, and they will tell you, ``Here has passed a grand and barbarous civilisation." Place them in sight of the Grecian statues and temples, and they will tell you, ``Here has passed a graceful, ephemeral, and brilliant civilisation." Place them in sight of a Roman monument, and they will tell you, ``Here has passed a great people." Place them in sight of a cathedral, and on beholding such majesty united to such beauty, such grandeur to such taste, such grace to such delicacy, such severe unity to such rich variety, such measure to such boldness, such heaviness in the stones, with such suavity in their outlines, and such wonderful harmony between silence and light, shade and colour, they will tell you, 

Here has passed the greatest people of history, and the most astounding of human civilisations: that people must have taken grandeur from the Egyptian, brilliancy from the Greek, strength from the Roman, and, beyond the strength, the brilliancy, and grandeur, something more valuable than grandeur, strength, and brilliancy—immortality and perfection. 

\flright{\textsc{Donoso Cortés}, \textit{Essays on Catholicism, Liberalism and Socialism}}

\flrightit{Posted on 2009-10-14 by Cologero }
