\section{Scientific Clarity and Assurance}

\begin{quotex}
\textit{Ce que nos pères ont fait par coutume et par sentiment, le poursuivre nous-mêmes avec l'assurance et la netteté scientifiques par raison et par volonté.}

What our fathers did through custom and feeling, we ourselves pursue it through reason and will, with the assurance and clarity of science. \flright{\textsc{Charles Maurras}}

\end{quotex}
With this quote, we see the dominating influence of \textbf{Auguste Comte}`s system of positivism on Maurras' entire political program. The other dominant influence was \textbf{Joseph de Maistre}, with his love of Tradition structured by the Church and the Monarchy. However, unlike Maistre, Maurras did not have Maistre's indomitable Catholic faith to justify his worldview. Nevertheless, Maurras accepted Comte's claim to have discovered the true scientific basis for de Maistre's views.

Of the two great ``fathers" of a scientific sociology in the 19th century — Marx and Comte — Marx has won out handily and Comte is nearly forgotten. So we are left with a view of science that allegedly supports the forces of revolution. That this ``science" is ideological, is all too obvious. Comte, on the other hand, rejected both religion and ideology, in favour of a purer science, which was simiply the exposition of laws concerning what can be seen or touched. As such, it avoided both metaphyscal pre-suppositions (e.g., materialism) and reductionism. In place of the latter, Comte proposed a hierarchy of sciences, even leaving room for a scientific ethics.

Nevertheless, Comte gets little support primarily because self-described religious conservatives have no interest in supporting Comte's atheism or Maurras' paganism. On the other side, secular rightists believe they can corral Darwin to their own ends, and accept its materialism and reduction as ``rational", often adopting a form of zoological racism. These secularists also oppose the role of religion in society, which they regard as mere superstition, whereas Comte and Maurras accept religion as a natural conservative and stabilizing force. What religion accomplished through ``custom and sentiment", the positivists seek the same ends through ``reason and will".



\flrightit{Posted on 2009-10-20 by Cologero }
