\section{Rape of the Sabine Women}

At the founding of Rome, Romulus was favoured by the gods, though his city was impoverished and heirless. There are three qualities that correspond to the three functions or castes, as stated in Table \ref{tab:QualitiesCastes}:

\begin{table}[h]\centering\small
\begin{tabular}{rp{.3\textwidth}cc}\toprule
\textbf{Quality} & \textbf{Meaning} & \textbf{God} & \textbf{Caste} \\\midrule
Di & The favor of the gods & Jupiter & Brahman \\ Virtus & Courage or manly energy & Mars & Kshatriya \\ Opes & Resources, property means of action, fertility, and propagation & Quirinus & Vaishya \\\bottomrule
\end{tabular}
\caption{Qualities vs. castes}
\label{tab:QualitiesCastes}
\end{table}

\paragraph{Romulus}


Romulus claimed to be the son of Mars and was a favorite of Jupiter. As High Priest, Romulus founded Rome with a ritual at the auspicious moment. He and his young followers were strong and brave, or, as they would have said, ``virtuous''. Nevertheless, they were impoverished and without women. To remedy these deficiencies, Romulus sent the following message to the neighboring cities:

\begin{quotex}
Cities ... arise from the lowest beginnings. As time goes on, those which are aided by their own valor and by the favour of the gods achieve great power and renown. They said they were well assured that Rome's origin had been blessed with the favour of the gods, and that valor would not be lacking; their neighbours should not be reluctant to mingle their stock and their blood with the Romans, who were as truly men as they were.
\end{quotex}


\paragraph{Titus Tatius}


Titus Tatius, king of the Sabines, scorned Romulus' offer. The Sabines were wealthy and complacent, and could not discern any positive qualities or accomplishments among the Romans. Seeing the Romans as a ragtag collection of Trojan immigrants and a criminal element, the Sabines scorned that offer, although several of their young women seemed attracted to the Romans.

\paragraph{The Abduction}


The Romans then abducted many of the women during a festival; this was not really a ``rape'' as the women seemed to go somewhat willingly and then accepted life as Roman wives. Titus Tatius attacked, gained entry into the city by a bribe, and pushed the Romans back. In response, Romulus resorted to prayer:

\begin{quotex}
O Jupiter, it was thy omen that directed me when I laid here on the Palatine the first foundations of my city... . Deliver the Romans from their terror and stay. I here vow to thee, O Jupiter, a temple, to be a memorial to our descendants from the City saved by thy present help.
\end{quotex}


The prayer was answered and made known to Romulus, who then commanded his troops:

\begin{quotex}
Romans, Jupiter Optimus Maximus commands us to stand and renew the fight.
\end{quotex}


With that, the Romans held their ground and drove the Sabines away. When the battle became stalemated, the Sabine women intervened, making the point that they ought not be forced to choose between their husbands and their fathers. The two tribes reconciled and the Sabines were absorbed into Rome as the third function or caste, which completed the Roman city.

\paragraph{Conclusion}


From this story, we see illustrated several Traditional points. First of all, Romulus offers a sacrifice to Jupiter and then discerns his Will. But, having discerned Jupiter's Will, Romulus is not passive; \emph{virtus}, or power, is necessary to bring that will into manifestation. This incident makes clear the relation between contemplation and action, the priest and the warrior, although Romulus himself plays both roles.

In contrast, Titus Tatius depends on his gold, material wealth, bribes and trickery, as befits the Vaishya. The Sabines as a race of the Soul or race of the Spirit are now forgotten, although as the race of the body, they live on as an element of the miscegenated Roman race.

There is nothing inapplicable to today. Even for Christians, divine predilection and predestination are dogmas, although you don't hear much about them today. God will favor some over others, in particular, those who offer prayer and sacrifice. Clearly, the God who ``loves all his children equally'' is a recent innovation. Predestination means that those ``saved'' will be able to obtain everything they need for that end. Universalism as a serious teaching is a recent innovation.

Finally, those who rely only on their ``opulence'' and wealth have already lost their spirits and souls. Not valuing fertility and propagation, they can see in the abduction of the Sabine women only the desire for fellatio in the back seat of a chariot or the need to fill administrative posts in the growing bureaucracy. Their genes may very well persist, although not necessarily in the form they envision.

Reference: \emph{Archaic Roman Religion}, Vol I, by Georges Dumezil


\flrightit{Posted on 2011-04-05 by Cologero}

