\section{Caste and Social Order}

\paragraph{Spiritual Authority and Temporal Power}


As was shown in the Doctrine of Castes\footnote{See Section \ref{sec:DoctrineCastes} in this book.}, a person's caste is the reflection of his nature, hence of his function in society. These functions, then, form a hierarchical relationship among each other. The first such relationship is that between \textbf{Spiritual Authority}, held by the Brahmins, and \textbf{Temporal Power}, held by the Kshatriyas. This differs from a theocracy, since the Brahmins hold no political power, but would serve instead as advisors to the Kshatriyas.

So, from one perspective, the Brahmins are superior to the Kshatriyas, yet from another perspective, the principles of spiritual authority, or wisdom, and temporal power, or strength, are two aspects of a more fundamental principle. Corresponding to these two castes are the roles of \textbf{Pope} (or High Priest) and \textbf{Emperor} (or King). When the fundamental principle is forgotten, the two roles are held by different men. Yet knowledge and action cannot be separated like that, just as the head that directs cannot be separated from the body that acts. Thus the two roles by right should be joined in a Philosopher King.

Guenon points out that in the story of the Magi-Kings bearing gifts, or in the idea of Christ being both Priest and King, the West was once aware of this unifying principle.

\paragraph{Brahmins}


The social function of the Brahmins, or priesthood, is the conservation and transmission of traditional doctrine. Individually, the Brahmin is oriented toward achieving transhuman states, or the Transcendental Man. He follows the path of intellection, or Jana Yoga.

\paragraph{Kshatriyas}


The Kshatriyas have two functions:

\begin{enumerate}


\item Administrative and judicial directed toward the maintenance of internal order

\item Military directed toward protecting the social organization from external threats
\end{enumerate}


On the individual level, the Kshatriya is oriented toward the realization of all his possibilities in the human state, or the True Man\footnote{\url{http://www.gornahoor.net/?p=1039}}. (A dim remembrance of this is in the motto of the US Army: ``Be all that you can be''.) He follows the path of Bhakti Yoga, and is more emotive. Thus the Kshatriyas need to get ``pumped up'' with inspirational rhetoric and grand ideas, or else with anger against opponents.

\paragraph{Synarchy}


In \textbf{Synarchy}, as described in Campanella's \textit{City of the Sun}\footnote{\url{http://www.gornahoor.net/?p=518}}, and developed more fully by \textbf{Saint-Yves d'Alveydre}, a third function is added, under the authority of the Vaishya caste. This is all economic activity of agriculture, trade, manufacture, medicine, and so on. Of course, this is subordinated to the spiritual authority and temporal order.

\paragraph{Servile Class}


The Shudras, or servile class, is in some sense close to the Brahmin caste since it is the unconscious reflection of spiritual authority. Provided there is no false resentment resulting from outside agitation, the unconscious instincts of this caste correspond to the conscious intuitions of the Brahmins. If the teaching function is done well, what the Brahmin knows is identical to what the shudra believes by means of the myths and rites.

\paragraph{Degeneration of the Castes}


This harmonious social order is the reflection of the cosmological Unity of Being. The degeneration of castes disturbs this harmony. The Kshatriyas may be unjust, thus creating internal disorder, or they may fail to protect against external threats. Note that these threats are not necessarily military, since they may come in other forms.

The Vaishyas may make economic gain the sole end, so they put their efforts into speculation rather than productive activity. They reject the spiritual teachings on usury, just prices and just wages. Through manipulation of capital they may concentrate wealth in a few hands rather than in a wider distribution. (This is not \emph{post facto} ``wealth redistribution''.)

The other castes may begin to spurn or mock the shudras, creating resentment. Their \emph{nous}, or intelligence, is only virtual, so they follow the superior castes, which should have actual intelligence (as understood metaphysically). Once this arrangement is broken down, the erotic and thumotic drives lack any constraint.


\flrightit{Posted on 2011-02-11 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Ted on 2011-02-12 at 12:13 said:}

Magnificent article The best I have ever read on Gornahoor thus far.

\hfill

\texttt{Sati Shankar on 2011-02-13 at 01:12 said:}

In the course of human civilization,was entrusted the responsibility of what is called material and spiritual to the Kings,(the rulers) and the Priests respectively, but neither could rise to the expected hight.The false ego super-dominated the Kings, the rulers. It has been observed in every age, that most of the preachings by thinkers to correct this anomaly went in vain.

`Such was the disgust of Plato,leading him to declare that no solution would emerge ``till philosophers become Kingsin this world.'' The Sages of ancient India solved this enigma by urging a conjunction of the king with the priest, as out of this union would emerge a government based on Dharma which would result in the happiness of the subjects.'

The priest, i.e., The Divine and the King, the ruler,are within... and the whole of this science has to do with a victory over the powers of perception and action.Only he can rule over others whi rules over himself.

Cast in Dharma signified the inherent nature and capability on the individuals, and NOT giving any weight to birth and material inheritance, story of Yajnavalkya in Kathaupanishad is enough to show it,

But in the course of time this intent was overtaken by the birth and material inheritance,and the false ego, the force behind it have been so strong, blinded the intellect

Cast, as it is interpreted by ``modern'' literate thinkers is not the cast that was originally suggested.

We still have, more or less, the same pattern but with new nomeneclatures, we have clarks, sweepers, executives, presidents... . all presuming specific qualities and capabilities in respective persons without giving attention to their birth inheritances... but defy and fail to give due credits to the tradition we have been and still are and the future generations shall be in.

\hfill

\texttt{Aidan on 2011-02-15 at 13:05 said:}

This is an excellent article. May I ask whether you meant to say: ``On the individual level, the KSHATRIYA is oriented toward the realization of all his possibilities in the human state''? The Brahmin, as you state, is oriented towards the realisation of the possibilities inherent within the suprahuman state and therefore has his ground and starting-point above the level upon which the Kshatriya operates. Full realisation of all human possibilities would not begin to touch that with which the Brahmin is concerned.

I would also add that disruption of an otherwise harmonious social order may also occur much more directly through the corruption of members of the Brahmin caste. Guenon hardly mentions it, and you do not, but it is surely the most fundamental disorder. A Brahmin whose personal nature, concerns and behaviour were of an order lower than that of his caste would have a disastrous effect upon all the other orders and, hence, upon his entire society. When such a disordered being holds a key position and fails to discharge his very precise responsibilities, the consequences cascade downward and reverberate in the most profound ways. Consider how often this has occurred. Consider how many have turned away from true religion because of it. It is easy to blame usurping Kshatryias but it is this alone that allows the subsequent wreckage to occur. If the Brahmins did not depart from their proper order, how could later, inferior sources of disorder gain purchase in a society in which suprahuman possibilities are daily made manifest to all?

\hfill

\texttt{kadambari on 2011-02-16 at 11:04 said:}

That is true the corruption of the Brahmins leads to the corruption of the entire order, but corruption of rulers is also important. Take for example the Rajput Kshatriyas in India. Intially they fought against the Muslims with bravery. Then they quarrelled amongst themselves and stooped so low as to begin to give their daughters to the Muslims who were Turko-Mongols of a slave dynasty, and forgot their original ideals of bravery and sacrifice for the land. However intelligent Brahmin advisors might be, if you have a stupid King then all proves pointless.

So there is a delicate balance between Kingship and Priesthood... 

\hfill

\end{sffamily}
\end{footnotesize}

