\section{Against the Neopagans}

On the Internet, there is an essay, attributed to Julius Evola, going by the title of ``Against the Neopagans.'' Since it has been so widely disseminated, it is impossible to know its source or the motivation behind it. In fact, it is the translation into English from the German translation of three chapters in \emph{Sintesi di dottrina della razza}. After Italy's ill-fated alliance with Germany in the late 30s, Mussolini felt compelled to come up with a racial policy to match that of Germany; previously, neither racism nor anti-semitism had been part of Fascist ideology.

Since Evola had been critical of German racial theories, Mussolini authorized the translation of Evola's book into German \emph{Grundrisse des Faschistichen Rassenlehre} as representing the Fascist position on race. I've compared the two versions and it seems the translation was not exact, but rather tendentious. Understandably, it leaves out passages that would only be of interest to Italians. What is more interesting, however, is that it omits passages that portrayed German thinkers in a negative light. For example, these three paragraphs are omitted:

\begin{quotationx}
Regarding the anti-ascetic prejudice of neo-paganism , we already said: neo-paganism redoubles the incomprehension demonstrated by Nietzsche. The Aryan would not have known asceticism in the normal way: his true mysticism would have been on this side: he would never have thought about a supernatural fulfillment of the personality.

Superstitions, residues of the dark ages or of Estruscan magic, lies and instruments for the tactic of temporal dominion of the clergy selling indulgences: that how they regarded everything that is a sacrament, rite or supernatural power. This shows they do not know that all of the life of ancient, Aryan, and specifically pagan Roman, civilization always had a ritual character, the rite accompanying every form of collective and individual life, not in the sense of an empty ceremony, but as the instrument of a real connection between human and supernatural forces. Opposed to that, H S Chamberlain attributed to the Aryan spirit the conquests proper to the so-called free thought and modern profane science.

Naturally, when one believes that Lutheranism represented an awakening of the spirit of the Nordic race see \emph{Foundations of the 19th Century} instead of representing, as it in fact did, the inducement for a further spiritual involution of the race and to a large extent its semitization: \emph{incomprehension can only be added to more incomprehension}. Thus there was something ingenuous, as Guenon rightly noted, in the scandal that Protestantism felt when faced with the claim of infallibility that, in the order of transcendental knowledge (in the West, we say: in matters of ``faith'') was accepted pacifically by ancient Aryan civilizations not for one man alone, as in Catholicism, but \textbf{for every member belonging legitimately to the brahman, that is, the ``solar caste'' of spiritual leaders}.
\end{quotationx}


\flrightit{Posted on 2011-03-26 by Cologero}
