\section{Idiot’s Guide to the Modern World}

\begin{quotex}
Man can be truly free and himself only when he maintains the centre of his being on a metaphysical plane. When he detaches himself from such a plane and focuses on practical goals, on temporal achievements, and, in general, on whatever was the domain only of lower castes taken in themselves, he abdicates, disintegrates, opens himself up to subterranean forces, whose tool he is soon destined to become, unless he takes account of them.
\flright{\textsc{Julius Evola}, \emph{Nascita ed Essenza del Mito Moderno}}
\end{quotex}


This is intended as an interlude between the \textit{Tools of the Occult War}\footnote{See Section \ref{sec:TacticsOccult} in this book.}, which delineates the mental constructs leading to the modern world, and the soon to be translated \textit{Birth and Essence of the Modern Myth}\footnote{See Section \ref{sec:BirthEssenceModernMyth} in this book.}, which focuses on the myths propping up the modern world and the presumed role of a future elite. By ``idiot'', we do not mean any personal insult, but rather its traditional meaning as someone who cannot follow a logical argument. Specifically, this term applies to those who call themselves ``Traditionalists'' of some sort, although they pay no attention to first principles or the corollaries derived from them.

Hence, the bearers of the modern world are not necessarily idiots. They may even agree on the general trends leading to the modern world, although their judgment of it is positive. First principles cannot be proven by science or philosophy, but can only be known and understood through an intellectual intuition. Those who cannot see those principles can nevertheless be logical; their logic will take them somewhere else.

To be perfectly clear, Gornahoor accepts one master of Tradition, the man Evola called the ``Master of Modern Times'' and \textbf{Guido de Giorgio} called the only master of the 20\textsuperscript{th} century: viz., \textbf{Rene Guenon}. Evola cannot be understood other than in the light of Rene Guenon, although I have known many men who try to do without that light. De Giorgio's mystical vision incorporates all the elements of Guenon, from Taoism, Vedantism, Sufism, to the metaphysical poetry of Dante. Clearly, anyone who deviates from Guenon's judgments needs to be very clear about the reasons for that.

For the point of this topic, Evola specifically refers to Guenon as the background thinker for understanding the tools of the Occult War. Specifically, the modern world can be understood as the progressive degeneration of castes, starting from the Medieval era. This means that the Western civilization of the Middle Ages must be regarded as a Traditional civilization. Those who deny this, regarding it instead as a long deviation, find themselves unable to account for the modern world without resorting to absurdities and historical inaccuracies.

Once this is conceded, the next factor is to understand the meaning of ``caste''. This cannot be understood in psychological or genetic terms. There is no gene for it, no Darwinian explanation for it, no personality test for it. The members of a caste share a certain nature and vocation; they share a manner of thinking and specific mode of existence. In our time, few people know which caste they should belong to. The schema from Table \ref{tab:SchemaCastes}, adapted from Evola, may be helpful:

\begin{table}[h]\small\centering
\begin{tabular}{cc}\toprule
Spiritual leaders & Solar and metaphysical \\
Warrior aristocracy & Secularized warrior \\
Bourgeoisie & Burghers, merchants, humanists \\ 
Workers & Plebeian and collectivistic \\\bottomrule
\end{tabular}
\caption{Schema of castes.}
\label{tab:SchemaCastes}
\end{table}

Both Evola and Guenon understand the decline of the West as the passing of power from a higher caste to a lower caste; Evola asserts that this is a historical fact, not a matter of opinion. Before continuing, it must be clearly understood that this has nothing to do with labels, the attachment to which would indicate an inability to understand essences. In particular, it is not a question of Catholicism vs Paganism. There are high and low caste pagans, as well as high and low caste Catholics. As Gornahoor has taken pains to point out, members of a caste have more in common with each other across denominational boundaries than within.

That is why Gornahoor sees continuity between the ancient and the medieval. Some may object to this, but their reasons are illogical; in particular, they place themselves outside the bounds of Tradition as understood by Guenon and Evola. There is no harm in repeating this historically verifiable and metaphysically valid process of degeneration.

In the \textbf{Middle Ages}, there stood a supernatural and supranational spiritual authority. As such, it possessed authority over the spiritual life of Europe and indirect power over the political and economic spheres. This indirect power acted more as an influence, as it did not take the place of the proper roles of the aristocratic and bourgeois castes over the political and economic spheres respectively.

The existence of a spiritual elite is undeniable; Guenon points out \textbf{Bernard de Clairvaux}, \textbf{Dante}, \textbf{Thomas Aquinas}.

Specifically, Dante was in an initiatic order and the Divine Comedy is a guidebook to spiritual initiation according to Guenon, who was familiar with the primary spiritual texts of the world's major traditions. \textbf{Ananda Coomaraswamy} called Dante the greatest of Europeans.

Those who cannot see this reveal themselves to be of a lower caste.

The Reformation, French Revolution, and the Bolshevik Revolution were the primary events in the successive degeneration of castes. Until now, when there is a soft revolution of the so-called ``fifth caste'', that is, of those who would be outside the Traditional caste structure. Note how this schema, understood not just by Guenon and Evola, but also by the Catholic counter-revolutionaries\footnote{\url{http://www.pliniocorreadeoliveira.info/UK_RCR.pdf}}, explains political movements.

The Left is the party of revolution, i.e., of the degeneration of castes, and the right seeks to maintain that proper ordering of castes. Compared to communism and socialism, capitalism is to the right, since it is the system of the bourgeoisie who oppose the revolution of the proletariat. However, to the original right of the Throne and Altar, bourgeois capitalism and liberalism represent left-wing, not conservative movements. Hence, any true Right will stand for spiritual authority as primary. It is not Gornahoor's task to predict which spiritual authority that will be. But any right wing movement needs to address that question first and foremost, or risk becoming a mere tool.


\flrightit{Posted on 2012-05-14 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Egemen on 2012-05-15 at 02:13 said:}

That's a very clear definiton of the modern view of ``The Castes'', in my opinion. Recession and disintegration of the tradition or as known name ``Kali Yuga'' is progressing over all civilizations without seperation of religions and nations... i am glad to follow your movement, ``Gornahoor'', and i want to support you if i can. if you seek any translations from turkish or arabic to english; of articles that are particularly about Guenon's interests; i will be help about them... 

\hfill

\texttt{Angolmois on 2012-05-15 at 13:50 said:}

This makes very clear the fact that spiritual realization should always be the first and primary objective both individually and collectively, and why it should be the thing emphasized by all truly traditional and right-wing movements, and not the futile battle of fighting over particularities and petty opinions. It also makes clear the fact that bourgeois capitalism and liberalism and ``conservatism'' are nothing but futile attempts of trying to hold back the last and final wave of the kali-yuga, and in the end, few of its very instruments of ``horizontal'' devastation.

\hfill

\texttt{logres on 2012-05-15 at 23:46 said:}

Here's a journal that understands what Gornahoor sees, but without the metaphysically deeper vision to explain why this is so:

\url{http://distributistreview.com/mag/2012/05/the-contradictions-of-liberals-and-conservatives/}

I would think the next thrust would be to pit the bourgeoise liberals/conservatives against each other in a battle for the reins of power in America. When the Taliban knew America was coming to Afghanistan (which we kindly announced to them months in advance), their first action was to assasinate the ``Lion of Panjeer'' -- the only man of honor (and known to be such) who could provide a legitimate authority to bring Afghanistan peace and order. Are we supporting and looking for those who are natural aristoi? Wasn't this one of the reasons Guenon fled from France -- that he was threatened as to his life, by fellow ``occultists''?

\hfill

\texttt{Angolmois on 2012-05-16 at 01:20 said:}

The Afghanistan murder was most likely done or paid by the agents of the American-Israel world order by the CIA, since the drug sales are one of the major reason for the existence and uprising of the puppet government there.

I think Guénon was attacked by the agents of counter-tradition first and foremost, and perhaps his ``fellow occultists'' were the puppets of these. Freemasonry and other shady things comes to mind immediately.

\hfill

\texttt{logres on 2012-05-16 at 12:18 said:}

If that is true, then the arm of Mordor is grown long indeed. I considered the event to be a masterstroke of dark policy, \& a warning to those who may be the subject of such in the future. ``the children of this world are wiser in their generation than the children of light''

\hfill

\texttt{Charlotte on 2012-05-20 at 10:19 said:}

Afghanistan... ..tragedy -- such a rich country, such amazing history, profitable centre of America's illegal drug cartel, say no more.

\hfill

\end{sffamily}
\end{footnotesize}

