\section{The Tools of the Occult War}

\begin{quotationx}
In this article from the January 1938 issue of \textit{La Vita Italiana}, Julius Evola describes his understanding of the occult war and its tactics. This should clear up two widely spread misconceptions.

\begin{itemize}

\item The first misconception is that the occult war has something to do with so-called occult or magical practices. To the contrary, it is not a question of men employing higher, or transhuman, forces to achieve their aim, but rather of those forces using men, with or against their will, whether they are conscious of it or not, to achieve aims apart from the desires of men.

\item The second misconception is that Evola is an adherent of conspiracy theories. Conspiracy theories restrict themselves to a two dimensional understanding of history, where some small group of men control events for their own ends. Evola is referring to something much deeper.
\end{itemize}

The rest of the article speaks for itself and it is surprising how, today, such a view is barely comprehensible. This is despite the work of \textbf{Wilhelm Wundt}, who made a similar point from a purely secular point of view. Yet the rapidity of the way public opinion can be swayed ought to give one pause.

Whatever Evola says about the occult war can only be understood by keeping in mind Guenon's writings of the anti-tradition, counter-tradition, and related topics. It may be surprising to some that Evola mentioned the Catholics as holding a similar understanding to his. That is understandable to those who only know the Vatican II church. But older writings do indeed describe an occult or spiritual war; the older Church and Evola even agree on the specific historical organizations or agents of the anti-tradition. We will, at some future date, reproduce some of those writings.
\end{quotationx}
\paragraph{Part I}
The occult war is when the forces of worldwide subversion lead from behind the scenes, adopting means that almost always elude ordinary methods of investigation. The notion of an occult war belongs to a tridimensional vision of history, to a history considered not according to the two dimensional surface of apparent causes, events, and leaders, but rather one depending on the third dimension of depth, the hidden dimension, which retrieves the decisive forces and influences which are often not even ascribable to the simple human element, whether individual or collective.

Having used the adjective ``hidden'', it is necessary to make its importance precise, indicating that it is not a question of an obscure and irrational substrate of scientific history, of something that could stand in the known part of it, but in the same relationship by which the subconscious stands to the consciousness of waking in the individual. This analogy is certainly exact, if considered from the point of view of those who, in this tridimensional view of events, appear not as the
\emph{subjects}, but rather of the \emph{objects} of history: in their acting and thinking, they do not take into account the true influences which they obey and the true ends that are accomplished through them. They have their true centre more in the subconscious and unconscious than in their waking consciousness, yet, in this respect, we can certainly say that it is in the subconscious that the most decisive movements of the occult war develop.

But from the point of view of the true \emph{subjects} of history, things are quite different, and we can speak, in this regard, not of the subconscious or the unconscious, but we must think of super-intelligent forces, forces that know perfectly what they want and the most appropriate means to realize what they want. To suppose that this substrate of history is accounted for by the ``irrational'', by ``life'', by ``becoming'', or by whatever else of those confused entities invented by modern historicism at the moment in which they thought they went beyond so-called scientific historiography, is among the most dangerous errors. We believe, on the contrary, that we see in that one of the suggestions prudently spread in certain circles typical of the aims of certain movements of the occult war in modern times.

We will explain ourselves more clearly in what follows. For now, it is enough to establish this point: we must not allow the third dimension of history to vanish in the fog of abstract and philosophical views, but we must consider it inhabited and occupied by quite precise ``intelligences'', whose expression and closest mouthpieces were almost always certain secret societies or organizations, but which must not at all totally express the intelligences, but just having their beginning and end in them.

Examining the various conceptions of history widespread in the West, perhaps it is the Catholic conception, out of all of them, that can claim views of the same type. It is typical of such conceptions, in fact, to see history not as a mechanism of natural, political, economic, or social causes, but as the unfolding of a type of plan (the ``providential'' plan) opposed by adverse forces and their historical emissaries, forces that assume either the moralistic designations of the ``forces of evil'' or the religious one of anti-Christian forces or forces of the Antichrist. One such conception can already furnish a foundation when it is transposed from the teleological plan to a practical plan and if it makes its general principle through an interpretation deepened by events. In the second place, when it is generalized, establishing its susceptibility to still be of value in non-Christian civilizations, for which it will be well, in this regard and with reference to human institutions, to speak generically of forces of tradition and forces of the anti-Tradition, the forces of the spiritual hierarchy and revolutionary subversion, the forces of the cosmos and of chaos. In this particular case, the anti-Traditional forces will also appear naturally as anti-Catholic forces, insofar as, in the Catholic tradition, they are intended to combat the Traditional spirit in general, to destroy the basis of every hierarchy and to ruin their historical representatives.

In our time, we will never be able to insist enough on the necessity of penetrating a view of the type that must not be considered as one among the many possible ``speculations'' of philosophy, but rather as an indispensable organ for true action. We are pleased to reproduce here a passage of the famous Protocols, to which we had previously called attention.

\begin{quotex}
The purely brute mind of the gentiles is incapable of use for analysis and observation, and still more for the foreseeing whither a certain manner of setting a question may tend. In this difference in capacity for thought between the gentiles and ourselves may be clearly discerned the seal of our position as the Chosen People and of our higher quality of humanness, in contradistinction to the brute mind of the gentiles. Their eyes are open, but see nothing before them and do not invent (except perhaps, material things).
\end{quotex}

It is a truthfully alarming thing to think in how many cases such a judgment about the gentiles---who would be, this this special reference which makes Israel the principle agent of worldwide subversion, the non-Hebrews, but who can also be considered, in general, as those who we already called the ``objects of history''---has still today a profound content of truth. The point of view of the greater part of the so-called ``men of action'', compared to those of the hidden adversaries, can certainly be said to be babies: they concentrate all of the strength on what is directly graspable with the hands, and are incapable of establishing the relationships of cause and effect beyond the range of an extremely limited zone and of a nature almost always grossly material. Moreover, there is a fundamental lack of principles, since they almost always believe that the point of view of action cannot be ``dogmatically'' subordinated to principles, but must be related to the ``unavoidable necessities of the moment''.

Unfortunately, this is still the level typical of most counter-revolutionary forces. It is a situation that can be compared to that of someone who finds himself perfectly equipped for the high mountain, with boots, sack, cord, and pick and then finds himself, after happily passing many tests, facing an expanse of water, at a lake or a sea, throws himself into it enthusiastically, believing that his equipment will still help him and lead him forward. Instead, it will only be of use to make him sink more rapidly to the bottom.

The occult war is exercised on that which, with an image drawn from positive science, we will call the ``imponderables'', or imponderable quantities. It is adopted fairly often to produce nearly insensible changes, from which sizable effects will proceed slowly but fatally. It almost never acts with forces opposed to other forces, but rather through an opportune direction of it, so as to lead to the desired effects. What Wundt called the heterogony of ends\footnote{\url{https://en.wikipedia.org/wiki/Heterogony_of_ends}}, plays a fundamental part in it: it is precisely about acting, while some forces or men believe they are willing and producing a given effect, and it alone, in actuality, they will produce or prepare something else quite different from it, thus uncovering an influence and an intelligence that transcends them. It is for this reason that, on another occasion in \emph{La Vita Italiana} , we said that the examination of the ``differential'' existing between the will and the event, between the principles or programs and the true consequences of it in history, offers the most valuable material for whoever truly wants to account for the true forces which history itself has obeyed.

But in this article we want to move beyond the general, and speak of a few of the most frequently used means, in our time, of the hidden powers of worldwide subversion to cover their action, to prevent that of their adversaries, and to continue to exercise their influence.

It is incumbent to point out that the idea here was often given by observations scattered in the various works of \textbf{Rene Guenon}. Guenon, in our opinion, has a nearly unique competence in all contemporary Western literature for similar types of problems.


\paragraph{Part II}

\begin{quotationx}
These are the first four of the eight tools of the occult war identified by \textbf{Julius Evola}. For readers unfamiliar with the philosophical currents of Evola's time, ``positive'' is usually called ``scientific'' today, at least in English. He also criticizes the philosophies of life (e.g., certain Germans, perhaps even Nietzsche as ``life-affirming''), becoming (Bergson, Whitehead), and absolute spirit (Hegel, Gentile).

As other manifestations of subversive forces, he goes on to mention a dilettantish ``traditionalism'' (which is rampant today) and even zoological racism (the ``hipster racism'' or ``race realism'' of today). Then there are all the New Age movements, which have only multiplied in the past 75 years. They hide the true Tradition and can be identified by their preoccupation with strange and unusual phenomena, rather than true spirituality; Evola calls them demonic. We can add a certain fascination with the dark and the ugly, that some today want to identify with Tradition.

This is one topic that I hope will lead to many comments, discussing, identifying, and clarifying these tools of the subversive forces.
\end{quotationx}


\paragraph{The Positivist Suggestion}


It is necessary to get used to thinking that the so-called ``positive'' mode of doing history and of considering events is somewhat less a spontaneous product and prejudice typical of a very limited mentality, than a \emph{suggestion} spread methodically in modern culture by anti-traditional forces, with the purpose of ``covering'' their action. Whoever believes that history is simply made by men and determined exclusively by economic, political, and social factors sees nothing else. And it is exactly such a world that ``sees nothing else'' that is needed by whoever wants to act clandestinely. A civilization dominated by the ``positivistic'' prejudice offers the ideal condition for an action developing along the ``third dimension'': and this is exactly the case of the current civilization.

Previously, we mentioned besides ``positivistic'' interpretations, those having as their basis abstract ideas of becoming'', of ``life'', or even of the ``Spirit''. In that we can consider a case of the application of a second tool of the occult war, which is:

\paragraph{The Tactic of the Surrogates}


It is followed, every time that the danger of an ``awakening'' is manifested and the energies seek to turn beyond the plan dominated by ideas, in the operation of which the occult game of the forces of subversion can easily take place. In the case now mentioned, these confused ideas of ``becoming'', ``life'', ``absolute Spirit'', etc., in fact, are only a decoy thrown at those no longer satisfied with positivistic patterns, and those whose gaze does not bring itself where it should. From the fog of such ideas, the field of action remains so well covered, as by the dark night of ``positivistic'' agnosticism.

The ``tactic of the surrogates'' develops, however, in a much more characteristic and effective way under the form of:

\paragraph{Tactic of Falsifying Substitution}

This is what we are talking about. When the effects of destructive action, appearing even on the material plane and giving rise to clearly visible phenomena, exhibit a degree hot enough to provoke a reaction, this reaction confusedly seeks symbols and myths to serve as the base for an engine of reconstruction, and can also find it legitimate and consistent with its goal. It is almost always about symbols and points of reference that, in their historical aspect, remaining in the traditional past, recalled from atavistic depths through the forces of an organism that feels itself threatened by final destruction. In these cases, the occult war does not take place with a direct action, i.e., not by opposing such symbols, but ensures that they do not circulate and are not worth falsifying or counterfeiting, inasmuch as the reaction either remains limited, or diverted, or even overturned in an opposite direction, on which they naturally move on to apply the same influences present in the evil against which they wanted to act and in the decadence, from which they wanted to rise up from and believed in good faith to be rising up from.

This tactic can develop in diverse domains, in the political as well as in the spiritual camp, and seems today to have a totally special efficiency. It is worth looking at an example.

The most typical is constituted by ``traditionalism''. The vague aspiration of a return to tradition, i.e., to a hierarchical and qualitative order centered in a transcendent spirituality and in an elite of its qualified and legitimate representatives, is diverted to in the form of ``traditionalism'', i.e., of vague assumptions of a dilettantish, or militant, character that confuses Tradition with custom and the routine, that leads into a narrow-minded particularism, that stops at given, handed-down forms, and knows nothing of their spirit. In such a way, they present an easy game to their adversaries and provide a good decoy to those who, through the false goal of anti-traditionalism, bring their true blow against tradition and foment new developments of revolutionary subversion. So the reaction remains contained and the maneuver leads successfully to prearranged goals. What, in certain extremist circles, the idea of race threatens to become because of some assumptions in terms of zoological materialism, as so even the Nordic-Aryan idea and even the pagan idea, come back to a conjuncture of the type. It is about ideas that, if they were understood and represented in accordance with their true nature and their ancient spiritual and traditional content, could certainly carry out a creative and healing action within the limits of some peoples of the North. As they appear instead in the circles already indicated, first of all because of a lack of principles, then because of a profound incomprehension and deformation of the ancient ``Aryan'' world, and finally through the tyranny of immediate interests and irrational impulses, make us suspect the effect of a distorting action of the type indicated. In relation to the new Italy, one should make every effort that something of the type does not occur in certain re-evocations of Romanity. The idea of ancient Rome certainly exists among those who today could have the greatest reconstructive efficacy: it is necessary to see that this efficacy is not paralyzed by means of the reconstituted facsimiles or by mere rhetorical resurrections, or by academic, erudite, archeological and even simply juridical, political assumptions, to which naturally elude completely the central strength and soul of the Roman symbol.

\paragraph{Tactic of Reversal}


The secret forces of worldwide subversion knew perfectly that the fundamental premise of order to be destroyed was the supernatural element, spirit conceived not as a philosophical abstraction, but rather as a higher reality and as the means of a process of transcendent integration of the personality. The limited influence that, in such a regard, the Catholic tradition could still have (here we cannot stop to analyze the means to come to that conclusion), the subversive forces had to try in every way to ensure that every new contemporary aspiration toward the supernatural, developing outside of the dominant religion, became distorted and paralyzed. The whole of what can be called neo-spiritualism, with allusion to the various theosophical, spiritist, Orientalizing, and ``occultistic'' sects can be considered the effect of one such maneuver. In all that, again, it is a question almost exclusively of deformation, counterfeits, and distortions of spiritual doctrine, due to the least qualified persons, who in this respect, seem often to be subjected to a truly demonic influence.

The consequent outcome is twofold. First of all, it provided an easy way of extending the general discredit into which those counterfeits had to fall, to the doctrines in themselves, with which they were practically put in condition of no longer damaging. We can mention that, in recent times, the maneuver has even employed political factors: and there are some circles that, in good faith, are able to see only ``Masonry'' (something that was made the synonym of revolutionary, anti-Christianity or international subversion) wherever one speaks of secret associations of esoteric symbols, even when such references lead back to ancient traditions, whose original spirit and aristocratic nature everyone should note.

In the second place, because of poorly understood theories and false practices, they are joined to channel all the forces that, constricted by the suffocating stranglehold of materialism and yearning for something extra-natural, not toward the supernatural, but rather toward the subnatural, i.e., the subconscious, mediumship, and, in general, toward a world that, far from representing an integration of the personality, can only propitiate its collapse and disintegration, as much by making of it the tool of some of the darkest ``influences''. So with neo-spiritualism the danger constituted by true spirituality was contained and reduced to an almost inessential minimum.

\paragraph{Part III}


\begin{quotex}
Tools 5 and 6 of the Occult War. In the Ricochet, Evola warns against causing divisions for particularistic reasons, particularly against another Tradition. Unfortunately, this is not uncommon today. He warns against scapegoating a group or organization for all problems. The focus must always remain on the plan of the subversive powers; focusing on the alleged agents will distract from the fundamental task.
\end{quotex}


\paragraph{Tactic of the Ricochet Attack}


It consists in provoking one of the forces to be torn down because of their still traditional character to undertake an action, that initially was aimed against a similar force and apparently of value to strengthen and increase the first force, but that then ricochets against it, dragging it down into the same ruins that it had propitiated against the other. By means of timely infiltrations, the secret forces of worldwide subversion often incite the representatives of a tradition to defend it through the discrediting another: whoever is subject to such a maneuver does not notice that to attack the tradition in the person of another people or civilization, for particularistic reason or contingent and inferior interests, means to be condemned to see itself attacked, sooner or later, in their own home, through a ricochet motion. The forces of worldwide subversion very much count on this tactic, and therefore make every effort to put every higher idea under the subjection of particularistic interests, pride, ambitions, and proletarian tendencies. They know that this is the best means to destroy every unity and true solidarity and to propitiate a state of division, in which their game will be extremely easy.

In this case, on can also allow every Machiavellian utilization of revolutionary forces back in. Some statesmen of scant discernment believed that propitiating or helping the revolution in the nations unfavorable to them was the best means to sustain and defend their true interests. Without realizing it, they prepared the exact opposite result, while they believed they served the revolution like an instrument, they themselves became tools of the revolution, which had to reach them after having passed over the body of their adversaries. We can say that the entire history of modern times, above all starting from the French Revolution, was the theatre of a subversion that tragically worked in such a way. We can never insist enough, in such regard, on the fact that only the absolute, ascetic, unfaltering adherence to an idea can serve to safeguard against the forces of the occult war. Wherever this standard is abandoned and follows contingent factors, the front of the resistance is broken and, with it, the possibility of a true autonomy.

We do not want to fail to mention a final appearance of an similar maneuver: the ideology characteristic of the ``principle of nationality'', the ``self-determinism of the people'', and imperialism, proclaimed by the Allies to mobilize every force against the Central Powers and then to beat them World War I , is exactly the best among the ideologies adapted to propitiate a general revolt of the nations of color against the great European powers and to overturn every supremacy from them.

\paragraph{Tactic of the Scapegoat}


When the hidden forces of worldwide subversion avoid the danger of being unmasked in some of their aspects, they see to it that all the attention of their adversaries is directed and concentrated on some elements, that only in part, or only subordinately, can be considered as responsible for their evildoing. All the reaction is then unloaded on such elements, becoming the true scapegoats. And the hidden forces remain free to pursue their game, their adversaries believing they have at this point identified the enemy and do not have to look for others. We repeatedly admonish certain anti-Semitic extremists to remain well attentive, so that, in their seeing the Jew everywhere, they do not end up by making themselves victims of a trap of that type. The same admonition could be applied equally to those who instead see everywhere Masonry or Protestantism, and so on, analogous processes being also verified in many other areas. It is necessary to guard against every unilateralism, never forgetting the complex plan of the occult forces which we must combat.

\paragraph{Part IV}

\begin{quotationx}
The final segment of \textbf{Julius Evola}`s essay. The final tactic, the \textbf{Tactic of the Substitution of Principle to the Person}, is very important, as it describes the creation of the modern world from the Medieval period. As a corollary, it defines what can be meant by the \textbf{Right} in the relative sense: it is whatever resists the dissolution of a stage, and the \textbf{authentic Right}, in its only true meaning, is whatever works toward the stage prior to the successive degeneration of the castes.

Evola concludes with the necessity of recognizing and indicating the tools used in the occult war. Only then can a man be called a \emph{counter-revolutionary}.
\end{quotationx}


\paragraph{Tactic of Dilution}


This is a particular case of the ``tactic of substitution''. To understand it, it is necessary to start from the idea, that the process that led us up to our current general crisis, had distant origins and various phases and that in each one of such phases---corresponding to specific forms of Civilization, State, ethics, etc.---the crisis was already present even in a less acute, more dilute, potential form rather than actual. The theory of ``progress'', or evolutionism, can be considered as another of the suggestions hurled by the subversive hidden forces into the general environment, in order to distract anyone from looking from the perspective of the origins and always pushing well beyond the process of falling, under the thrust of false mirages, and, above all, of that of the conquests of the technical-mechanical civilization. However, the hard experience of tragic events in recent times saw to it that such a hypnosis was no longer able to conserve its efficacy and was no longer able to prevent many spirits from recognizing that the presumed course toward progress was a course toward the abyss and that a going back was imposed. The hidden front then put into notion every means to prevent any radicalism. One of the preferred methods was, above all, putting a prejudice against ``anachronism'' into circulation, against ``that which is no longer adapted for the times'', and then directing the forces that returned to look toward the origins, toward one or the other of those preceding phases, in which the crisis and evil were found in forms less advanced and, for that very reason, less visible. And even this trap had success. The leaders of worldwide subversion know naturally that, having come to that point, the danger is eliminated, because the whole was reduced to soon retrace the same path and to push, at the end of it, to dissolution even these forces that were awakened and that wanted to react and pull back.

We could also indicate many historical examples of this tactic, both recently and further in the past Above all, the political Leaders of current anti-subversive movements should not lose it from view. Let us say in all frankness---since in such a field a lack of frankness is certainly equivalent to a fault---there are aspects of contemporary nationalism that would merit being tested thoroughly. Anyone can note, in fact, the anti-traditional, anti-hierarchical, revolutionary character that, in the past, compared to the preceding ecumene of imperial and feudal civilization, had the notion of nation. Now, it is the nation that today predominantly challenges to combat the extreme forms of the crisis and subversion, represented by various red internationals communist parties . There is therefore the necessity to formulate a concept of the nation different from that which makes of it a stage of the path that had to lead to what today we must combat. Several years ago we provided the necessary points of reference to confront and resolve this problem, and to distinguish two nationalisms \emph{La Vita Italiana}, Mar, 1931 . We must return to our essay.

\paragraph{Tactic of the Substitution of Principle to the Person}


It is undeniable that, in many respects, the decadence of traditional institutions occurs through the decadence of their leaders and representatives. But this is not the only cause: in order to come to a true dissolution and involution, it is necessary that the tactic of the substitution of principle to the person be added to it, a new tool of the occult war, which works in the following way. When the representative of a given principle appears unworthy, a process against the principle in itself is set in motion in the person of this representative, or the process is extended from the person to the principle. They do not come to the conclusion that that representative does not measure up to the principle and must therefore be substituted with another person who can really represent it, but instead they go on to say that the principle itself is false or harmful and must be substituted with another principle.

In how many cases does the attack against this or that degenerate, vain, or vicious aristocrat get translated into the attack against the very aristocratic principle in itself and into a tool of demagogy? And perhaps the subversive and heretical action of Luther, having as his excuse the corruption of the representative of the Roman Church, had a different meaning? Again, history is rich with episodes of the type, equivalent to many other moments of the worldwide subversion. Let us remember that the ancient hierarchical system was constituted by a spiritual authority, under which stood the warrior aristocracy, then a bourgeois caste and then the working masses. The fall of this traditional system, in many of its aspects, had as its instrument the tactic here indicated. The warrior aristocracy rose up against the decadent representatives of the pure spiritual authority, not through substituting them with another, more worthy representatives of the same principle, but by drawing from that the pretense to be emancipated or lay claim for itself the supreme authority. In a following phase, i.e., in a following civilization, the Third Estate rises up against the degenerate warrior aristocracy not because it ceded the place to the true aristocrats, but to seize power for itself. In yet another phase, the process against capitalism and the bourgeoisie (i.e., against the Third Estate) in its destructive and negative aspects does not aim for a corresponding reform, but is only the excuse for the revolt of the proletariat and the usurpation of power---descending progressively from step to step---straight through to the party of the proletariat (bolshevism, communism).

\paragraph{Conclusion}


The fact that, for reasons of space, we only indicated some examples and we were concerned above all with the exposition of the principles, not to prevent---we hope---seeing what vast fields of application these principles can have and to that valuable results it could reach those who, adequately equipped, can apply them methodically in any field. Since one can assert without uncertainty that there is not a single area in which the occult battle of the forces of worldwide subversion are not in any developed way; and indeed these areas that would seem more unrelated to events of the type are those that, by principle, should be considered with the greatest wariness.

We repeat one more time that it is not a question of ``philosophical positions'', but of things much more real and that there is no leader or combatant at the front of the counter-subversion and the tradition that can be called mature and at the peak of its true tasks, before having developed in itself the faculty of seeing clearly in this subterranean world of cause and of indicating quickly the action of this or that invisible tool which our enemy does not cease to adopt in every epoch that, like the current one, sounding the prelude to the moment of final decision and the final battle of all, a cycle of civilization.


\flrightit{Posted on 2012-05-07 by Aeneas}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Lyon on 2018-05-07 at 14:54 said:}

Hmmm... Does this possibly dovetail with the work of E Michael Jones, specifically in his work titled, \textit{The Jewish Revolutionary Spirit and Its Impact on World History}? Jones, in that work, describing the comparatively small portion of the iceberg that can be seen above the surface.

I've only recently been introduced to his writings via Fidelity Press and his culturewars.com website. Admittedly, both Evola and Jones are approaching our predicament from differing vantage points.

As for Evola, the only work of his I've read in depth (... studied and ruminated) was The Hermetic Tradition, Symbols and Teachings of the Royal Art.

My understanding of both Jones and Evola are still in formation. Any general guidance and cautions on either is appreciated.

\hfill

\texttt{fractalkhaosstrikesonceagain on 2018-05-09 at 17:48 said:}

As usual with your Blog GornaHoor: Thanks a lot Cologero for OPENING OUR EYES : You help us to face our real(s) foe(s) ... 

\hfill

\texttt{jalexandermaximilian on 2020-08-18 at 21:08 said:}

I've been thinking lately, about the past decade or so, of these occult forces, which use men, men who appear as powerful but are really just pawns of these forces. It seems now more than ever ``leaders'' are not in control of their own thoughts and actions It seems the world of men is being propelled toward some finale. Can it continue in this manner without a finale, or will ot continue unto a whimper? This sort of understanding is why now I eschew all U.S. polytricks, the men you see are not in control of anything... even themselves. They may think they are. But their works and goals are all for another Thing. I seek God. Then wonder if I am my own Master? From where does even thought arise?

Lyon: Jews would merely be another group of pawns serving another more powerful force, they are certainly hardheaded and organized, but that seems to me to be a case of an external mind using them to its own goal with great success.

\hfill

\texttt{veritycacciatrice on 2022-01-09 at 05:34 said:}

``For now, it is enough to establish this point: we must not allow the third dimension of history to vanish in the fog of abstract and philosophical views, but we must consider it inhabited and occupied by quite precise ``intelligences'', whose expression and closest mouthpieces were almost always certain secret societies or organizations, but which must not at all totally express the intelligences, but just having their beginning and end in them.''

I witnessed this first hand less than a fortnight ago.

My son was doing some study on Garibaldi in Italian and he suggested I watch a video he'd seen on it (more for the comedy styling than the content).

I was following along and was surprised to see the most bloody battle location in the second War of Independence chosen due to its name.

\hfill

\texttt{Charlotte on 2012-05-09 at 07:53 said:}

Yes, subversion -- `black magic' works by subverting and perverting the true spiritual forces, in no small part through `bad practice'. Depending on the individual and their potential and actual power, this can indeed be disastrous. The abomination of the desolation, the breaching of the Holy of Holies. the rend must be fixed. The ancient forces tend to react badly -- out of character, out of bed the `wrong way' -- when needlessly awoken (mediumship), roused for the wrong reasons. And yes, it is all too easy to be deceived outside the known framework, the `safe zone, amounting to entering a labyrinth without the guiding thread which one follows in order to return. The severance of the silver cord is no laughing matter, one requires rescue!

\hfill

\texttt{logres on 2012-05-09 at 09:23 said:}

Evola is unusually clear and prescient, and not so wordy, in these passages. Many thanks! I am interested in his idea that the Catholic Church (at that point, and perhaps still) could be made a bulwark against certain tendencies and currents. I fear that we may have moved beyond such a point, irremediably, although certainly the question cannot be decided on an empirical basis. My instinct tells me that the Church could still be awakened, that it slumbers, and that it is a sleeping giant. Who will awaken her?

\hfill

\texttt{logres on 2012-05-09 at 09:52 said:}

Tactic number 3, falsifying substitution. Classic example of this would be the resurgent Confederacy of the southern states in 1865 -- they thought they were fighting for altar and hearth -- in reality, they were being used to provoke a frightful advance of the occult power, by obscuring the goal in the midst of a provoked reaction. I suppose 1938 Germany might be an even better example. Each time this ``false start'' occurs, it cements the overall process.

\hfill

\texttt{Charlotte on 2012-05-09 at 12:38 said:}

Oddly enough I was compelled yesterday to revist one of Rudolf Steiner's stranger books, \textit{Secret Brotherhoods and the Mystery of the Human Double}, in particular the lectures relating to Individual Spirit Beings and the Constant Foundation of the Universe.

I know there are no big Steiner fans here (unless they're hiding!), but in the present context he is very helpful and penetrates even further into the `nature of the beast', as it were. I just transcribed a large portion of this work on Alchemical Weddings:

\url{http://alchemical-weddings.com/alchemical-weddings/the-constant-formation-of-the-universe}


One thing I've ommitted is the constant talk of `post Atlantean epochs', be they fifth, sixth or whatever, as I've never really made sense of them, perhaps I don't like the labels.

Here is the end of the extract, which I think you might enjoy:

What lies in store for humanity is that cosmic forces will be won for humanity in a dual way that is wrong or a single way that is right. This will bring a genuine renewal to astrology, which is atavistic in its old form -- a form in which it cannot continue to exist. Those who know about the cosmos will struggle against each other... .for the remainder of earthly existence the dead will collaborate with us to a high degree.

The moment you penetrate fully into all the secrets of the double, you will find yousrelf able to assess the value of brains that are merely lumps which you have placed on a scale, lumps which can be brought to life solely by the double.

You must come to realise that all the wishy-washy talk of `cosmic religosity' or `how powerful the yearning for it is', and so on is nothing other than a means of spreading fog over these matters which ought actually to be entering the world in clarity and which, above all, ought only to be carried in clarity into human life as practical moral and ethical impulses.

It generally sheds light on the nature of the occult war as it is manifesting now, very helpful I thought, along with Evola's texts. When occultists manage to agree you know you're onto something!

Also coincidentally, the recent talk of `Saturn' made me think, firstly, that he was eaten by his children so we should really focus more on IoPitar? Also I was contacted by a group related to the Philosophers of Nature, so of course the memory of `touching saturn' was revived. Interesting synchronicities from a personal point of view that may also resonate with others here.

\hfill

\texttt{logres on 2012-05-09 at 14:34 said:}

Steiner was tainted with evolutionary worldviews, but ``knew'' better. That statement of his about religious ``feeling'' demonstrates such. Steiner was Goethe's heir, so he shared Goethe's failings. Just as there is a danger that the individual could ``fail'' and collapse in the postmortem state, so there is a cosmic danger that this world could implode, in more than a physical manner. Christians can ``rest'' in Christ's sure sacrifice, but this may entail complications which are unnecessary, not to mention enormous suffering.

\hfill

\texttt{Charlotte on 2012-05-09 at 14:54 said:}

well I think it's tricky to find any occultist (human after all) who doesn't have failings, though? Evola, Guenon etc are routinely attacked for being `fascist' or whatever -- I think it's partly a hazard of the job and a consequence of the times they lived in. Even VT had, in my opinion, unfair ideas to do with shamanism, so I can't even agree with him on everything. Lots of wisdom to be learned from them all but we must always take care to show discernment and think for ourselves... also to be heedful of the fact that we are at a different stage in history, esoteric or otherwise. Nevertheless, I do think Steiner's insight into the occult war are very illuminating in the present context, bringing a bit more spirt to the dryness of Evola, which I nonetheless found useful.

\hfill

\texttt{Angolmois on 2012-05-09 at 15:10 said:}

Steiner may be helpful as you say Charlotte, but in reality the only reason why Evola is ``dry'' is because he doesn't stray into unnecessary or fantastic areas in his analysis of the occult war.

You are certainly right about thinking for ourselves. Many esotericists have a hidden need, sometimes even a vice, to rest upon the shoulders of this or that authority.

\hfill

\texttt{Charlotte on 2012-05-09 at 15:27 said:}

I wonder how we should see the present wars around the world in light of all this... .

\hfill

\texttt{logres on 2012-05-09 at 22:25 said:}

Charlotte, thanks for the quote, which I read in full at your site. I agree that Steiner is useful, just trying to place him in full context. As Cologero says, we learn from and (hopefully) in some sense, go beyond (in the sense of seeing their failings more clearly).

\hfill

\texttt{Charlotte on 2012-05-11 at 12:36 said:}

I think we have to define `fantastic' as it's a matter of opinion depending upon experience. From a rational perspective much of what Evola says is right, and yet I personally see a lot more depth in Steiner's analysis, which speaks to personal experience more concretely. Something is only `fantastic' if you don't believe it's possible, but there are not the usual rules and boundaries in the `other worlds', of time and space, for one, which are pretty enormous limiting factors on the physical plane. The rules `on the other side' are different but the playing is just as hard. I agree, however, that one must learn to read Steiner in a particular way and not take everything to heart or on board, simply to scan for what resonates.

\hfill
\texttt{logres on 2012-05-10 at 08:34 said:}

Ricochet -- Luther thinks he is re-masculinizing the Church, but actually, the destruction of intellect \& tradition leads to sentimentality. A warning... to me.

\hfill

\texttt{Angolmois on 2012-05-11 at 03:39 said:}

... but actually, the destruction of intellect \& tradition leads to sentimentality.''

And it leads also to protestant capitalism, bouregous liberalism, and finally to ``kulturkampf'' and to the destruction of monasticism in the 19th century. Quite a ricohcet.

\hfill

\texttt{Charlotte on 2012-05-11 at 04:40 said:}

so many actions and reactions in the course of history -- eventually the balance will be found

\hfill

\texttt{Avery on 2012-05-10 at 23:52 said:}

Substituting a principle for a person is a very interesting metaphor. It's really a process of the lower ranks being instructed that their superiors do not have special right but are only employing a general principle which can be understood by anyone, which leads them to think that the principle is more important than the person. They rid themselves of what their perceive as a useless middleman, only to discover that they are lost without a dedicated superior to guide them.

\hfill

\texttt{Charlotte on 2012-05-11 at 06:06 said:}

`superior'... .hmmm... ..people have different qualities, in the ideal scenario they will take it in turns to lead depending on what is required -- just like a team of long-distance walkers will take it in turns to keep the pace. This should of course happen organically and intuitive in a utopian sense, but in reality there tend to be infights and intrigues if there is no rigid plan to follow. It also supposes the basic equality of peoples, which I do believe in principle but in practice it doesn't always show through. Often people just need a chance to shine and being trapped in a rigidly hierarchical `management structure' doesn't necessarily allow this. Nowadays people tend to be discouraged from thinking for themselves, one of the reasons I'm sure why alcohol is so widely prevalent, cheap and encouraged, whereas cannabis, for instance -- something Americans could once pay their taxes with, and which it was compulsory for farmers to grow in England -- is outlawed and demonised. Such is the fate of anything that interferes with the profit potential and control mechanisms of large corporations. I realise you guys probably hate this kind of talk and possibly I was missing the point of your comment, but I still felt moved to say this. Also it's true lol!

\hfill

\texttt{Cologero on 2012-05-11 at 08:38 said:}

Sure, Charlotte, just as you allow your neighbours to ``take turns'' fixing your car, or providing you with healthcare.

Alcohol and marijuana are so popular because they are the great equalizers and levelers. In their ``normal state'', men differ considerably from each other, in their intelligence, wisdom, sanctity, ability ... in every conceivable way. But every drunk and every stoned-out person is exactly alike.

\hfill

\texttt{Charlotte on 2012-05-11 at 12:01 said:}

ah well, too much of anything is a bad thing, but I'm not one for stereotyping. I do find it intensely ironice that an extremely useful plant is demonised whereas super strength alcohol and tobacco (a totally non-beneficial product) are considered acceptable. You obviously don't know the history behind `wacky backy', but I can promise you it was once considered so useful that to grow it was considered helping the war effort. And no, that doesn't mean I spent all day `stoning'. Apart from anything else I hate tobacco. As for the pros and cons of mind altering substances we could go on all day about this if we were of the HPB, Crowley or Gurdjieff school -- as it happens I tend to think there are far more cons because abuse happens far more often than use. However, the Pythia of Delphi and others like her might well disagree, as might the rainforest shamens or other representatives of ancient tradition. Nothing new age about it. Unlike LSD, which was created by the Tavistock Institute to create mind-blown hippies and other enemies. I hate to have to point it out, but probably even the Essenes partook of holy moly, according to recent evidence that has detected cannabis oil around their settlements -- the medicinal benefits of this plant are beyond dispute, it's prejudice that makes people blind to it. As it happens I don't keep a car, it's bad for the environment, but when I did the neighbour did help fix it more than once. Healthcare in England is free for all and we like it that way... .

\hfill

\texttt{Charlotte on 2012-05-11 at 12:04 said:}

actually, the one thing you CAN say about stoners is they're total pacifists and adept at minding their own business, whereas alcoholics are invariably violent towards self and others.

\hfill

\texttt{Charlotte on 2012-05-11 at 12:12 said:}

the reason alcohol and marijuana are popular is NOT because they're levellers, it's because in the case of the former it lowers inhibitions -- good for social interraction -- and stimulates the brain to produce dopamine, which is highly pleasureable. It is also very addictive, which leads to repeated use. THC and cannabidiol on the other hand are known neuroprotective antioxidants but the effects of smoking or ingesting it range from total euphoria, feelings of `universal connectivity', increased empathy and other things that sound dreadful to some folk, to paranoia, nausea and delusions. Strong stuff. But I digress, let me get back to the study books.

\hfill

\texttt{Cologero on 2012-05-11 at 19:20 said:}

You just stereotyped them, and proved my point. Drunks are all alike and stoners are all alike ... great levelers.

\hfill

\texttt{Charlotte on 2012-05-12 at 09:38 said:}

what rubbish, I did no such thing!! I read Evola experimented a lot with hallucinogens -- does that mean he's `just like them' ?! :-)

\hfill

\texttt{Logres on 2012-05-13 at 00:30 said:}

``... and always pushing well beyond the process of falling, ''

I didn't understand the syntax here... 

The overall thought that Satan leads men back in circles is a brilliant one. Tea Party? Reformation?

\hfill

\texttt{Cologero on 2012-05-13 at 08:23 said:}

I'm open to suggestions to make it read more clearly ... the word is spingere: push, urge, drive. The idea is that the theory of progress is driving the process of the fall forward.

... by the way, recent high-level announcements and the instant worldwide reaction, seemingly coordinated, seem to demonstrate this tactic in action. So who can anticipate the next ``thrust''?

\hfill

\texttt{Graham on 2012-05-15 at 23:26 said:}

Re: the substitution of the principle to the person, there is a variant. It's interesting how modern popes and clergymen apparently recognize their own lack of spiritual authority, leading them to `tone down' the principles they are supposed to represent. For instance, John Paul II rejecting the papal tiara was significant. Stated otherwise: conscious of their own lack of `charismatic' authority, they seek to present the church as a bourgeois `traditional' and `bureaucratic' authority, something they feel at ease with.

\hfill


\end{sffamily}
\end{footnotesize}

