\section{Tactics of the Occult War}

\begin{quotex}
\emph{Be wise as Lucifer and as innocent as the Holy Spirit.}\end{quotex}

The tactics of the \textbf{Occult War} describe the third dimension of history, that of depth, beyond the dimensions of space and time by which we conventionally view things. Yet, those who can only understand in terms of facts, material causes, and visible leaders still think at the level of brutes. Even worse, many men, wishing to be seen as intelligent, deliberately cultivate this view, thereby blinding themselves to any deeper understanding.

Let no one be deceived into believing that this is a simple matter. Even among those who desire to, or claim to, follow Tradition, there are many who fall victim to the tactics of the occult war, including some whom I have known personally for some time and would have expected a bit more insight.

This War, at its most fundamental layer, as the battle between the forces of Order and the forces of Chaos. This necessarily assumes, without question or dispute, that there is a Cosmic Order, which we have been calling \textbf{Logos} and obviously a force behind that Logos, which we have called \textbf{Providence}, in accordance with Guenon and other Traditional and Hermetic authors. Hence, two requirements are imperative for anyone who wishes to consciously engage in this War.

\begin{enumerate}

\item He must understand the Cosmic Order

\item He must align his Will with the forces of that Order
\end{enumerate}

The dimension of depth is the province of the Spirit, apart from and beyond any physical or material considerations. Obviously, then, to begin such an engagement, the spiritual warrior must first of all wage that battle in his own consciousness. Specifically, he must understand his own inner constitution, above all, the various layers of his soul life. He must strive to bring them into harmony with the Cosmic Order by subduing the lower layers to the higher. He must learn to harness the erotic and thumotic impulses that, left to themselves, lead to a process of inner integration.

To understand Gornahoor, readers must keep this foremost in mind. On the one hand, we have striven to describe the Cosmic Order in many different ways, apart from any a priori ideological commitment or allegiance to any particular movement or institution. We also hope to inspire others to take up arms in this battle in the way they see fit. In addition, one must also understand the forces of Chaos. Those forces employs several tactics, which we summarize below. We have already written about some of them in detail, others we shall take up as the need arises.

\paragraph{Scientific Suggestion}

This tactic involves the suggestion that science and the commitment to the world of facticity is the only way to understand the world. For example, we have described Francis Bacon's jettisoning of the knowledge of formal and final causes\footnote{\url{https://gornahoor.net/?p=3468}} as the basis for science, to the point where an educated man no longer knows what they are and think it fantastic when they are explained to him.

\paragraph{Replacement}

When men are no longer satisfied with science, they may turn to abstruse philosophical or ideological explanations. Such philosophical systems are so complex and subtle that their adherents spend several years, usually in their youth, to fully grasp them. This commitment and one-pointedness prevent them from reaching the dimension of depth. We see this today, unfortunately, even among those who aspire to Tradition, but commit to philosophic ideas, often of French origin, a strange counterweight to Guenon.

\paragraph{Counterfeits}

When the effects of Chaos are too obvious to ignore, some may turn to Tradition, but more often a counterfeit of Tradition. For example, we see today those who commit to a form of neopaganism that never existed as a Tradition anywhere, any time. These counterfeits take up the visual symbols and residues of earlier Traditions, but lack the understanding of their spiritual foundations.

\paragraph{Inversion}

This tactic begins by attacking spiritual intuition and gnosis, replacing them by psychological considerations. The logic and rationality of metaphysics is abandoned in favour of irrational and subconscious forces. Hence, there is the rise of all sorts of neospiritualities, extolled even in some New Right circles, that reject the actual Traditions, which they consider too rigid and limiting. Just today, some woman tried to convince me that masturbation is the key to discovering God. It is certainly a democratic and egalitarian notion, although you would think that enlightenment would be universal by now.

\paragraph{Ricochet}

This occurs when the representatives of one Tradition oppose themselves to other Traditions. Gornahoor has taken pains to point out how Traditions agree at the level of depths and has not opposed any real Traditions. Yet there are many who attack the dominant and most recent Tradition of the West, whether out of ignorance or malice, for a fleeting political advantage. In the attempt to attract committed adherents, they end up on the same side as their enemies. For example, neopagan groups use the tactics of their declared enemies to discredit the Medieval Tradition, thereby undermining Tradition itself. Th Ricochet is the best way to destroy every unity and solidarity.

\paragraph{Dilution}
The forces of subversion replace an idea that seems too radical at a given time, with a more diluted version that seems more acceptable.

\paragraph{Misidentification of Principle with its Representatives}

This has been a very effective tactic of the forces of subversion, since the mass of men are nominalists, meaning they cannot even understand principles but only individual things. Thus if some leader or member or adherent of a Tradition has betrayed his principles, men come to reject the principles themselves. This is all too common, even those I know personally who claim to be ``Traditionalists''.

\paragraph{Replacing Infiltrations}

Unfortunately, organizations and societies whose original impetus was to preserve and promote some aspect of Tradition degrade over time. The leaders, as well as members, of such groups become ignorant of their founding spirits, and instead grow accustomed to the material benefits and unearned prestige that accrue to such leadership. Such organizations are easily discredited along with the principles they had been created to uphold.

\paragraph{Simplification}

Instead of making the effort to grasp complex forces and multifarious phenomena, men are too easily attracted to overly simplistic explanations. For example, the modern ideals of liberalism, socialism, scientism, and rationalism are the outcome of several historical antecedents among which are the Renaissance, the Reformation, French Revolution, Marxism.

Clearly, that is replacing a spiritual cause with a material one. Then there are those who are part of such subversive movements, yet beyond all logic, still expect a different result. Some see a superficial commitment to certain issues, and then assume they are natural allies. However, at the level of depth, nothing could be further from the truth. The most egregious example that bypasses 1500 years of Western history in a one sentence slogan, attribute all the modern deviations listed above to an alleged secularization of the Medieval Tradition. Yes, I have heard that from educated men, as I'm sure many readers have.

Based on writings of different authors.


\flrightit{Posted on 2012-01-25 by Cologero}

