\section{The Modernity of Hobbes}

\begin{quotex}
\textbf{Julius Evola}`s review of Carl Schmitt's monograph on Thomas Hobbes originally appeared in Lo Stato, January 1939. Together, they expose the mentality of the modern mind in the political realm. This article will appear in two parts. This part is Evola's review of Schmitt's essay. Given the length of this piece, and its depth, I will provide my own commentary at a later time. Its relevance to the failure of the ``right'' and ``conservatism'' to gain any traction in the modern world will become obvious.
\end{quotex}


According to \textbf{Carl Schmitt}, three thinkers, after a long period of ignorance of the meaning and value of their doctrine, due to the dominance of the rationalistic mentality in vogue up until recently, seem now ready for a new valuation. The first is \textbf{Vico}. The second is \textbf{Machiavelli}. The last is \textbf{Thomas Hobbes}.

We recognize today in Vico the precursor of a new historical method, the exploration of the world of origins and ``myth'', conceived as a fundamental element in the primordial life of nations. The period of new European political activism has liberated Machiavelli from unilateral and discredited widespread interpretations and recognized in him the adversary and overcomer of moralistic lies and political cant, the defender of the style proper to a ``heroic realism''. As for Hobbes, Carl Schmitt himself intended to rehabilitate and to clarify the true meaning of his political doctrine in a very recent monograph titled \emph{Der Leviathan in der Staatslehre des Thomas Hobbes} (1938), a monograph so rich with interesting and also urgent points, to merit a special examination.

The political doctrine of Hobbes, as is well-known, is centered in the symbol of the Leviathan, reminiscent of the Bible. It speaks of the Leviathan for the first time in chapters 40 and 41 of the \emph{Book of Job}, where it is described as the most powerful and indomitable of the beasts of the sea. It is certain that, from this, his adversaries drew a precious inducement to jump on it, discrediting and maligning the Hobbesian doctrine, describing it as that in which the State becomes an omnipotent beast, a monster that devours and crushes everything. But in truth things, in respect to the Leviathan, are quite different.

Leaving aside, for the moment, the symbolic aspect, and limiting ourselves to the theoretic, in the Leviathan Hobbes sought to depict an almost mystical type of political totalitarianism, the State as a being that is simultaneously man and god, machine and beast, and that claims for itself the supreme authority and supreme law. Thus, as it is depicted in the cover of the original English edition of Hobbes' book, the Leviathan had none of the traits that could recall the biblical marine monster. It appears in it instead as a majestic human figure, crowned, holding the sword in his right hand and the staff in the left, i.e., the symbols of the double power, the temporal and religious. The union of these two powers is a fundamental point of Hobbes' doctrine. It would have been a single thing at its origins: essentially due to the Hebraic-Christian conceptions this unity would have been lost. It must be restored. The State, for Hobbes, must return to being a \emph{deus mortalis} mortal god , something ``transcendent''. But already here the point that invalidates all of the Hobbesian doctrine is betrayed, and that had to cause the end of the same historical Leviathan: as Schmitt notes, Hobbes thought that the distinction of the two powers was foreign to pre-Christian conceptions, something, hitherto, correct: ``since for him, religion was a part of the political'', something incorrect. The opposite is in fact true: the foundation of the totalitarianism exhibited by all the great traditional political unities was in the fact that the political was a part of the religion, and not vice versa. And it is not about a subtlety, but a fundamental divergence: as we will soon have to see.

According to Hobbes' political sociology, the originary state is the state of nature and specifically that of \emph{homo homini lupus}\footnote{Man is wolf to his fellow man} and \emph{bellum omnium contra omnes}\footnote{War of all against all}. From the anxiety at this stage and the need for security, peace, and order, the State draws its origin and, in truth, in an almost contractual way. There is no doubt that, in this regard, Hobbes belongs already to the ``modern world'' and is in severed antitheses with the preceding, medieval conception. Schmitt makes this point clearly: ``The sovereign is not the \emph{defensor pacis} Defender of Peace\footnote{\url{http://en.wikipedia.org/wiki/Defensor\_pacis}} as a peace relating to God; he is the \emph{creator pacis} creator of peace, only an earthly peace. The process is the inverse of the ``divine right'': since the power of the State is omnipotent, it assumes a divine character only for this reason. But its omnipotence has an origin totally other than divine: it is a human work and is actualized through a contract stipulated by men.'' Schmitt continues:

\begin{quotex}
The decisive element here is that such a contract does not reflect, as in the medieval conception, the reality of a collective entity created by God, a natural preexisting order: on the contrary, the State, that order and collective entity, is the result of the intellect and creative human power and comes to life only through the contract. This contract is conceived in a completely individualistic way. Individuals, as atomic entities, encounter each other in their anxiety, until the light of the intellect penetrates them, and then a ``consensus'' arises in terms of a general and unconditioned submission in the face of the supreme power.
\end{quotex}


It is thus that in the armed and crowned figure of the Leviathan his body is represented as composed of an infinite multitude. We witness, here, an almost magical transformation (we will see that this attribute can assume in this the entirety of its significance): the absolute rises from the relative. Individuals conjure up the Leviathan, recognizing in it the only way for salvation, security, peace: but with that they conjure something transcendent, something that is greater than their sum and the individuality of each contracting party, something, in respect to which they, hypothetically speaking, no long have any power.

Hence, the image of the Leviathan as ``machine'': \emph{machine magna, machine machinarum} machine of machines . For the mentality of Hobbes' time, still far from the technico-industrial era, the machine represented something mystical, not to say, magical: a human construction, it seemed nevertheless to possess its own life, expressing itself in its function. For which Hobbes' Leviathan appears as ``machine'' as well as God, man, and beast. Hobbes, as Schmitt points out, is the first to have systematically exposed the idea of the State conceived as a \emph{magnum artificium} great artifice , created by men and technically perfect, as ``a machine'' having in itself, i.e., in its function, its ``truth'' and its ``right''. We reach, by that path, the Hobbesian formula expressing the extreme, but purely logical consequence of all this theory: \emph{auctoritas, non veritas} authority, not the truth . It is the formula --- it should be said, ``exquisitely modern'' --- of irrationalism and political pragmatism.

Schmitt attempts, to tell the truth, a mitigation, saying that at the base of Hobbes' view is the fact the he ``no longer poses any distinction between authority and truth, making the \emph{summa auctoritas} highest authority from the \emph{summa potestas} highest power '', where that formula would express only the norm of a way of technical and realistic thinking, neutral in regards to ``values'' and ``truth'', separating the metaphysical and religious notion of truth from the idea of command and political function, to confer on the latter a primary and undisputed autonomy. But how then not to see that the Leviathan would become only the symbol of temporal hypostatized power and that each of his pretenses to rule not only the sword, but also the staff, to encompass in itself not only the fortress, but also the temple (that is also found in the figuration) would be usurpatory and the abuse of power?

In order to understand a theory, such as Hobbes', in its true significance, it must be considered not in itself, in the abstract, but in its operation in its historical ``place''. At the time in which Hobbes lived the crisis of the type of State we call ``traditional'' was more than visible and that the Middle Ages had known, to a certain extent. Here authority drew its absoluteness from the pretense of coming from above, of reflecting ``truth'' itself and putting in \emph{fides} faith the fundamental cement of every political unity, it shaped this very unity to the elements of an essentially spiritual, and mediately religious, nature. In Hobbes' time the animating force for a State of such a type had already failed: its structure survived, i.e., a State that sill conserved the absoluteness and the pretense of the supreme authority, previously conferred on it from reference to principles of a super-political order: metaphysical or a least religious.

Leviathan is the attempt to save and preserve these surviving forms by grafting on them a new principle of life. In a situation in which there no longer existed men capable of reawakening in individuals the heroic capacity for obedience and recognition on the base of super-individual and super-material principles, we look for a new myth and a new foundation: the contractual one with the related theory of the ``state of nature''. They seek, i.e., to persuade individuals that the supreme authority can be accepted regardless of the ``truth'', for its intrinsic utility in regards to exigencies, now, of a materialist and anti-heroic nature, like those that are at the base of the needs for security, peace, and order on the plane of purely human and contingent reality. It is such that, while the Leviathan is presented as a ``machine'', i.e., only a human construction, and its body, its substance, appears now constituted by the mass, except that it conserves some traits typical of the ancient traditional State; it claims for itself the unity of the two powers. The State is divine, although the ``divine right of kings'' is denied. Divine---``more than merely secular'', according to the ancient English formula, \emph{God's Lieutenant}---is for Hobbes the same sovereign, although his authority is now destitute of every foundation from above and the replacement by an acknowledgment on the part of individuals on the basis of reason and convenience. An amphibious form, therefore, and, in its depths, contradictory.

Can we prescind, in regards to the Leviathan, from this historical placement? And then what must we think about this ambiguity? What import has the very contractualistic myth have in it?

It is clear that, once the State is secularized, the Leviathan becomes a colossus with feet of clay. Its absoluteness can still linger for a certain period, by inertia. Through the conditions of its life, it then goes on to have a fatal distinction, reproducing fundamentally the distinction between the two powers that the Leviathan had tried to surpass. And the distinction between the modern external world, in which, for various reasons, by chance or other, the individual subordinates himself and recognizes the ``divinity'' of the State, and a type of inner forum, in which the individual feels himself free and inviolable. This distinction is announced already with the ``\emph{cuius regio, eius religio}'' whose realm, his religion\footnote{\url{http://en.wikipedia.org/wiki/Cuius\_regio,\_eius\_religio}} of the Reformation, religion becomes thereby a politically conditioned fact; it was exacerbated in the period of Baroque absolutism, reached its critical stage with the so-called theory of the ``positive'' State or State ``by right''. But with this condition necessary to conserve its ``neutrality'' and its absoluteness independent of any ``truth'' whatsoever, the Leviathan began to dig its own grave. And what Schmitt himself thinks in the political realm, without however stressing that it is a question of the result of a fatal logic. He says,

\begin{quotex}
at the moment in which we admit the distinction between interiority and exteriority, the superiority of the former over the latter, therefore even that of the private sphere over the public sphere, it is already decisive. An external power, as far as integrally, specifically, and even loyally recognized as only a public and exterior power remains nevertheless something empty and dispirited. Such a terrestrial god has only the appearance, the \emph{simulacra}, of divinity. There is nothing divine that can be imposed from the outside. \emph{Non externa cogunt deos} Nothing divine lets itself be externally enforced.
\end{quotex}


In these words the essence of the process is understood that, by stages, has led to the individualistic, liberalistic, and rationalistic erosion of the true State. Having come this far, when even the State conserves leviathan-like traits of an absolute power, transcending the right and the strength of every individual, it is deprived of true autonomy, it becomes fatally an instrument of what Schmitt calls the ``indirect power''. Characteristic of such a power, according to Schmitt, is the distribution of the relations between political command and risk, between power and responsibility, between protection and obedience: it is a question of a masked, and absolutely indirect, act that gathers all the advantages of power without explaining to anyone the risks naturally inherent in it. We easily recognize, in this precision, the nature typical of the forces that had, and have, the real power in the ``agnostic'' States of an apparently democratic of liberal type. The State becomes in effect a neutral machine that obeys the game of the ``indirect power'' more strongly.

Such a type of State therefore came into crisis, and new political forms arose, characterized by an abrupt aversion for every form of ``indirect power'' (up to the anti-Masonic and anti-Semitic country of the ``single party'' State), by a new love for political totalitarianism, for pure authority. By way of these new forms, what must one think, as in a certain measure seems to be the idea of Schmitt, of a ``currentness'' or ``modernity'' of Thomas Hobbes' political myth?


\begin{quotationx}
After reviewing \textbf{Carl Schmitt}`s monograph, \textbf{Julius Evola} moves on to his own unique interpretation. At least, I am unaware of a similar interpretation. Rather than viewing Leviathan as the beginning of the modern political world, he sees it as the residue of the Traditional world, the remnant of a civilization that has lost its connection to transcendent principles and is left only with its forms of authority.

Here, drawing from some ideas of \textbf{Donoso Cortes}, Evola sees the possibility of a recovery of Tradition from the stage of the Leviathan. Facing Europe in revolution, Donoso postulated the arrival of a ``man on a white horse'', a dictator who would restore order and tradition. Of course, Donoso had in mind a Catholic state, but that doesn't alter the essence of Evola's view.
\end{quotationx}


After touching on all that, it is rather awkward to respond to a similar question: in particular, why, if we must respond in the affirmative, would it be necessary to see if the correspondence, in the end, is resolved in a valorization of Hobbes, or in grounds for suspicion of other aspects of the new political forms? This point merits a brief examination.

For that, it is necessary to state beforehand that the connection of the Leviathan with the myth of the contract must be considered absolutely inessential and contingent. Schmitt demonstrates a sensibility of everything particular when he states that when an author makes use of an image, like the Leviathan, he enters a domain where it is no longer a question of verbal formulas, dialectical abstractions, or mere ``values'', but instead they act and dominate forces and powers, thrones and agencies, which reminds us of these words from J. G. Hamann: ``There is just one step from transcendental ideas to demonology''. It is about something ``elementary'': the substance of the Leviathan is essentially irrational, and when individualistic and utilitarian motives can also serve, in a preliminary phase, to lead toward a similar form of State that, in its actualization, calls to life something else, attests to an evocation, a really magical deed, no longer having anything to do with the rationality and the utility of individuals.

From a collective whole something transcendent can come to life, susceptible of manifestation in a center of pure irresistible power: power that, only as such, is truth and authority: the center the vortex. This is Leviathan.

We have already come to know this Leviathan as a phase of an involutive process: it is the survival of the ancient traditional State, whose absoluteness moved from the level of \emph{veritas} to that of simple \emph{auctoritas}. That, therefore, is in a retrospective vision. Let us now consider, instead, the possible return of the Leviathan to political developments beyond the acute phase of the crisis, i.e., beyond the complete individualistic breakup of the State. It should always depict a sovereignty deriving from the galvanized collective, therefore from forces of a prepersonal and irrational nature. It is not the collectivism of Marxism and bolshevism, since these movements, at least in their ``myth'' (and here, regarding ``evocations'', the myth is not to be disregarded), oppose every form of sovereignty and are centered in a completely materialistic and uniformistic vision of the world. Instead, it would deal with every anti-individualistic and anti-rationalistic movement, that profess again a conception of the State as ``\emph{dio mortale}'', that fuels this symbol with potent forces of faith and sacrifice, but unfortunately without surpassing the stage that Keyserling would call ``telluric'', i.e., the stage at which authority suffices and truth is superfluous; in which myths, and not true principles, are the best instrument to capture and organize collective forces; in which the miracle of an exceptional personality, of a ``man of destiny'' saturated with the ``numinous'', and not a pure ``divine right'', founds and legitimates sovereignty and command and confers a transcendent character to the idea of the State.

Seeing things this way, we can speak of a certain modernity of the Hobbesian Leviathan. But even in its new appearance it can claim so little, as in its emergence in the descending, involutive process to the character of a stable, definitive form. In the descent, once it reached the Leviathan from the true and traditional State, it was not possible to stop it, i.e., not to descend still even lower. In the new processes, we will necessarily find ourselves, sooner or later, facing an alternative. Either the new totalitarian movements will not succeed in surpassing their purely activist and vitalistic irrational phase: and now it will be difficult for the new Leviathans to be able to completely secure itself in the face of ``indirect powers''---with which they no longer mean the irresponsibility of parties or some secret associations, but influences of a deeper elusive nature, always capable of infiltrating wherever they do not find the way barred by the presence of authentic principles and in a firm, abiding truth: and such forces do not have as a goal the maintenance and the intensification of worldwide anti-traditional subversion. Or else, we will enter into a new phase in which the Leviathan, so to speak, will become the body formed to make possible the incarnation and the manifestation of a principle and a higher order: with that, the collectivistic and irrational aspect of the principle of totalitarianism and authority will be surpassed and will again implement a type of truly spiritual and traditional hierarchical organization.

We can already point out that an analysis of the purely symbolic aspect of the Leviathan would nearly lead to the same result. Once the affinity of the Leviathan with the dragon the other mythical characters is ascertained, Schmitt tries to demonstrate that symbols of the type would have negative and evil meanings especially in Semitic civilizations, while in others the dragon often appeared among the regal, imperial, and metaphysical symbols. That is not completely exact or, better said, this interpretation must be subordinated to another more general and essential one. Not only for the dragon or the serpent, but for many other ancient symbols, they assumed two opposed meanings, in relation to two different periods. When a power, in its positive, superior spiritual origin, falls and through regression becomes the tool of obscure, ``elementary'', and ``demonic'' influences, in the symbolism of the various peoples, the symbols, once unknown, come to acquire a negative quality, while the positive ones that were originally proper, are transferred to other symbols, alluding generally to a restorative function. That also fits exactly the Leviathan and confirms its ambiguous, double nature. And this duplicity, according to everything we said, reflects an alternative that will be decisive for the ultimate result of the forces of worldwide revolution.


\flrightit{Posted on 2012-07-05 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Golgonooza on 2012-07-06 at 07:01 said:}

Very interesting article, thank you.

\hfill

\texttt{Mihai on 2012-07-09 at 04:25 said:}

This description of the ``leviathan'' corresponds very well with Plato's tyrant, of which he gives an ample description in the ``Republic''. As the ``leviathan'', the tyrant is an inverted image of the ``philosopher-king'', the latter being an image of superior principles, with knowledge of higher realities and laws, while the tyrant is an incarnation of the collective, inferior strata of the human psyche, being raised to power to represent the animalistic desires of the masses, in himself, the tyrant being the sum of ignorance.

In Christian terms, such a rule can be called ``the reign of the Antichrist'', though not necessarily in an absolute sense.


\hfill

\end{sffamily}
\end{footnotesize}

