\section{Monotheism, Totalitarianism, and other strange Memes}

A meme is an idea that spreads from mind to mind, uncritically, much like the way a virus infects the body. Just as a virus is usually deleterious, if not fatal, to the host organism, so also is the meme to clear and critical thinking. Because of some alleged or fantasized connection or compatibility between the new right/neopagan movement and Tradition based on memes, we find it once again to address them. These are:

\begin{itemize}

\item The illogic of mythos

\item The relationship between monotheism and totalitarianism

\item The origin of liberalism

\item Volkisch arguments
\end{itemize}

Our primary goal is to clearly demonstrate the radical incompatibility between Tradition and the New Right because the latter movement claims an ``inspiration'' from the former. We will make that absolutely clear to any but the obtuse. This exercise is not to prove that one or the other false, but just to show they are logically and historically incompatible. Since the new right is primarily a political movement, we are not necessarily attacking their goals or tactics. They simply are not our goals or tactics.

\paragraph{Illogic of Mythos}

\textbf{Alain de Benoist} has openly stated his method is to prefer Mythos over Logos, as his inspiration is \textbf{Nietzsche} who regarded Reason as merely in instrument for the Will to Power. Hence, in this sense of the word, a myth is akin to a meme; it is to be accepted because it gives ``life'' or power, not because it is true. Nevertheless, new rightists argue as though they are saying something true; rather, they are using the power of rhetoric to persuade others to a point of view. That is why new right web sites are so concerned with graphics or appearing ``hip'', as that is how you penetrate the mass mind. The conclusion is what counts, not the logical process to get there.

\paragraph{Monotheism and Totalitarianism}

This has got to be one of the strangest memes, articulated by Benoist in \textit{Monotheism vs. Polytheism}\footnote{\url{http://home.alphalink.com.au/~radnat/debenoist/alain10.html}}, with a reach further than I would expect. Julian Assange ended his interview with the Muslim cleric Sayyid Nasrallah\footnote{\url{http://rt.com/news/assange-hezbollah-nasrallah-syria-263/}} by asking him about it. Listen to the response if you like (circa 25:00). Benoist's essay suffers from four major defects:

\begin{itemize}

\item Belief and Action

\item Misunderstanding of Monotheism

\item Myth of tolerance

\item The Value of Totalitarianism
\end{itemize}

\subparagraph{Belief and Action}

It is true that a strongly held belief leads to action. For example, if I believe the weather will be fair this weekend, I may make plans for a picnic on Saturday. However, my belief does not cause the picnic, because the weather may change and ruin my picnic. But that is Benoist's position. He claims that the belief in monotheism results in totalitarianism. His corollary, which doesn't even follow, is that belief in polytheism results in toleration. He simply assumes the preference of toleration, the ultimate liberal value, to totalitarianism. Not even considering the question of the truth of the two options, Benoist himself believes neither in mono- nor polytheism, since they are equally mythos, or unproveable and irrefutable perspectives. Hence, the two theisms act as shibboleths\footnote{\url{https://gornahoor.net/?p=3736}} to identify friend and foe. Polytheists are friends and are allowed in the new right circle of trust, otherwise, not, since they are masks for totalitarianism.

\subparagraph{Misunderstanding of Monotheism}

Benoist's understanding of monotheism is absurd, stooping to the most primitive level. Monotheism\footnote{\url{https://gornahoor.net/?p=3344}}, understood in the classical sense, is part of Tradition. Its roots in the West do not lie in the Old Testament, but rather in the constant Western tradition. \textbf{Plato} and \textbf{Aristotle}, each accepting an ultimate principle, recapitulate more ancient traditions, reaching back to the Vedas. \textbf{Plotinus}, who was a major source of doctrine for Evola, claimed that everything arose from the One. The Medieval understanding of God, as Being, continues that same thread. It is hardly worth commenting on Benoist's thesis, given its clearly false premises.

\subparagraph{Myth of Tolerance}

The next claim is the polytheism leads to tolerance, a claim that is absurd metaphysically and false historically. If polytheism means anything at all, it means that there are multiple independent powers, not subject to some unique and supreme power (i.e., monotheism). How, then, is it possible for them to almost act in harmony? Obviously, it is not at all possible, so these powers will come into conflict with each other with nothing higher to resolve any such conflicts. Hence, Poseidon is a jealous god and resents the gift given to Zeus. Men are subject to the arbitrary and inconsistent whims of gods and goddesses, one offering blessings and the other putting up obstacles, or take up different sides in a conflict.

Historically, polytheism did not lead to mutual tolerance. This is clearly pointed out by \textbf{Fustel de Coulanges} in his Ancient City. When two cities came into conflict, they did not tolerate each other. Either the weaker became subservient to the stronger and was forced to acknowledge the latter's gods, or else it was utterly destroyed, men, women, children and livestock.

\subparagraph{The Value of Totalitarianism}

This is the final refutation of Benoist. Rather than a negative, this is essential to the Traditional world view, as Evola asserts\footnote{\url{https://gornahoor.net/?p=4035}}. For the Traditional civilization, every aspect of life was guided by spiritual principles and cosmic law. Is the Law of Gravity totalitarian because everything that goes up must come down? Perhaps so, but that is in the very nature of things. Analogously, if there truly is a cosmic law, then all of human life is subject to it. Paradoxically, this leads to tolerance. In traditional civilizations, the cosmic law adapts to different cultural conditions, manifesting as positive law. If two such civilizations should confront each other and recognize the cosmic law mutually followed, they can distinguish between contingent adaptations and necessary principles. This reliance on common principles will allow them to resolve their differences. (Obviously, this doesn't happen in conflicts between strictly exoteric entities.)

\subsection*{Liberalism and the Folk}

\paragraph{Liberalism}

\begin{quotex}
As to tolerance, it is surprising how far removed from the equity and prudence of the Church are those who profess what is called Liberalism. For, in allowing that boundless license of which We have spoken, they exceed all limits and end at last by making no apparent distinction between truth and error, honesty and dishonesty.
\flright{\textsc{Pope Leo XIII}, \emph{On Human Liberty}}
\end{quotex}


It is not our intent here to debate the irrefutable facts of history, but rather the spiritual roots of liberalism. Intolerance, which is not equivalent to totalitarianism, is a fact of life. Any healthy social body will resist influences that threaten it, including spiritual as well as material threats. Christians from time to time forced conversions, contrary to its alleged principles. Pagans tortured Christians, not to mention waged all out wars between each other. It is hardly a novelty and it is absurd to think it will never happen again. Ruling powers almost always use force to maintain themselves. When an American politician claims that Assad is ``killing his own people'', I would like to ask him how many of his own people did Abraham Lincoln kill? Such actions are always felt justified by someone.

Benoist writes that Christianity and liberalism are spent forces and hopes for a new paganism to fill the vacuum. Perhaps it will, but I think we have demonstrated that ancient paganism was not quite what he envisions.

The meme we dispute is the New Right claim that liberalism is secularized Christianity. Benoist, given his early involvement with Maurras' thought, certainly knows the long standing opposition between the two points of view. To combat that meme, we took the trouble to edit and format Donoso's book on the topic\footnote{\url{https://gornahoor.net/library/CortesEssays.pdf}}. Leo XIII is closer to understanding liberalism than is Benoist. The former recognizes that in liberalism, there is no difference between truth and error, right and wrong ... they all assert and demand equal rights. But that is precisely the Nietzschean view espoused by Benoist: there is no truth, there are only perspectives.

The ethics of liberalism are not at all those of Catholicism, at least when it was Traditional. Instead, we claim, with much more justification than Benoist's thesis, that it is actually the vision of Charles Maurras that is secularized Catholicism. If we accept Evola's claim that the historico-material world is a reflection of spiritual reality, then a secularized version of that would be to take the reflection itself as the sole reality without reference to its spiritual underpinning.

That is precisely what \textbf{Charles Maurras}, following the lead of \textbf{Auguste Comte}, did. Positivism in their hands proposed a political and social program that was compatible with, if not identical to, what any Catholic government would have implemented. Benoist certainly knows this, but neglects to mention it; his sycophants are ignorant of it. But Benoist rejected Maurras along with the latter's counter-revolutionary program. What he would replace it with is far from clear; however, it is necessarily not the counter revolution.

We don't know the purpose of the memes spread by Benoist, other than to stifle logical thinking and replace it with easily assimilable slogans. We suggest that those on the right who are uncomfortable with theological language and reject the supernatural put aside Benoist's books for a time. Check out Comte and Maurras; perhaps you will find more useful ideas.

\paragraph{Volkisch Movements}


The volkisch meme persists in some quarters; this is only a concern when they claim to be speaking for Tradition. In Tradition, the folk are lunar and passive. While there may be some sort of folk soul, it is changeable over time, just as is their material and biological constitution. As passive, the folk are absorbers of ideas, not creators. Ancient peoples believed they were founded by a divine being, usually an ancestor, who formed them, promulgated their laws, taught them their rites, and gave them a spiritual vision. This was subsequently developed by their priests, poets, and others from their spiritual and political elite.

From the Traditional perspective, then, we are not so concerned about what the folk believed, but rather about what the elite understood. We have heard recently commentators claim that ``our people'' are Baptists, and others say ``our people'' are Odinsts; they are speaking of the same people. That is a sentimental attitude, but lacking in spiritual depth. Lasting change requires a real conversion; then there will be a vision without which the people will perish.


\flrightit{Posted on 2012-04-25 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{logres on 2012-04-26 at 08:20 said:}

Saladin always comes to mind, as an exemplar of somebody with whom the Crusaders could wage a war that was not a modern war of extermination, but a holy war by both sides, acknowledged as such.

So -- ``If two such civilizations should confront each other and recognize the cosmic law mutually followed, they can distinguish between contingent adaptations and necessary principles. This reliance on common principles will allow them to resolve their differences. (Obviously, this doesn't happen in conflicts between strictly exoteric entities.)''

The last point is well-taken; when the esoteric cores go, leaving the husks behind, the World Wars are on the way. Conversely, one sees someone like Dinesh D'Souza suggesting that Christianity has more in common with Islam than liberalism does, a suggestion which provoked a huge amount of anger from one of my progressive friends (who I think was frightened by the spectre of Christians and Muslims turning on their political masters together).

\url{http://en.wikipedia.org/wiki/Dinesh_D\%27Souza}

\hfill

\texttt{Cologero on 2012-04-26 at 08:54 said:}

Not just progressives, but neocons. I find it curious that in opposing Islam, neo-conservative commentators always seem to take the liberal position.

\hfill

\texttt{golgonooza on 2012-04-27 at 11:46 said:}

The very process of secularization as commonly thought of as a `rolling back' of superstition to create a space for the neutral and rational public sphere has been called into question by radical orthodoxy. It claims that liberalism has been born out of this belief in the neutrality of the secular, when in fact the secular is actually a pseudo-theological creation, replacing the sacred with different ultimate concerns.

For instance Neo-liberalism rests on the idea that political subjects need protecting from each other, from which the notion of rights comes. In this pseudo-theology the fundamental nature of the world is violence and power.

This just replaces the Augustinian view of the world as fundamentally harmonious -- violence appears in this view as an intrusion into that order, and humans don't need protecting from each other, but reconciling with each other.

\hfill

\texttt{Cologero on 2012-04-27 at 20:20 said:}

Good points, golgonooza. Pace the New Right, it reinforces in a different way that liberalism is not the logical outcome of Tradition, but rather a pseudo-Tradition, that arose from a rejection or reversal of fundamental principles.

Liberalism denies the existence of truth with the naive assumption it will spontaneously arise in a ``neutral and rational public sphere'' through debate and conversation. Instead, it became the battleground for different perspectives, subject only to the Will to Power. Those with the most power came to dominate the so-called neutral public sphere. Nietzsche's thought is the logical outcome of the liberalism: he praises the will to power to embrace and promote perspectives rather than to arrive at Truth. The latter project, he calls decadent.

Nietzsche expected out of the ruins of liberalism that a noble pagan superman would arise. Instead, we ended up with the hermeneutics of suspicion. If the will to power is the ultimate reality, then of course power and violence is the result, both of which Nietzsche praised. But there is no \textbf{superman} to dominate and enforce his will over the rest. Instead we have a society of the petty minded, suspicious of each other, and battling over their respective material and sensual interests.

\hfill

\end{sffamily}
\end{footnotesize}

