\section{Identity and Sameness}

\begin{quotex}
The reason man carries his head in the skies is because he does not find here anywhere to rest his head. And why would he look here to rest his head? Doesn't he cling to unity? And can unity find its rest in the order of mixed things?
\flright{\textsc{Louis-Claude de Saint-Martin}, \emph{Man of Desire}}
\end{quotex}

On more than a few occasions, I have stumbled upon the claim that there are neo-pagan groups in Russia or the Ukraine who are combining \textbf{Julius Evola} with \textbf{Alice Bailey}. That is quite curious since I cannot conceive how that would work. Unfortunately, I have found no details, at least in English, even after a few inquiries.

Bailey was a Theosophist who produced a massive output of occult literature, although she claimed to be no more than the amanuensis for a reclusive Tibetan \textbf{Djwhal Khul}. However, no such Tibetan sages were flushed out when the Chinese communists invaded Tibet nor is the Bailey opus remotely close to any of the Tibetan literature I have read. In his commentary on the Voice of the Silence, \textbf{Aleister Crowley} suggested that \textbf{Helena Blavatsky} humorously claimed that work was dictated by a Tibetan; presumably, he was hinting at the origin of his own Aiwass. Although Bailey does not seem to have had much of a sense of humor, we will still attribute the writings to her.

For those unfamiliar with her, she is a prime influence on the New Age thing. Her organization is an NGO at the United Nations and manages the meditation room. That makes her the target of conspiracy theorists, particularly given her teachings on one world, one religion, one race. But that is the future and in current circumstances she claims the Aryan race is the superior one and she comes across as somewhat anti-Semitic in some place; perhaps that is her appeal to the Russians.

In the interests of full disclosure, I must admit that I was under her influence some time ago. Her books came in a hypnotic austere blue and were full of occult teachings and fascinating diagrams. Adding her to other currents popular at that time, I tried to create my own personal spiritual tradition. Although I have long since given away her books, who knows what has stuck in hidden places in the mind. Before dealing with the causes of the modern world as she saw them, I'll start with a little review.

\paragraph{Ideas Have Consequences}

\textbf{Richard Weaver} wrote the book, \emph{Ideas Have Consequences}, that everyone should read. Thought shapes behaviour so incorrect beliefs will lead to inappropriate actions. Even if it takes several generations, all the ramifications of a germinal idea will eventually try to manifest in the order of mixed things. Although the rare geniuses of a new system of thought understand what they are doing, their immediate followers typically do not. They ignore the hidden assumptions behind the system and then also fail to see all the consequences of a given idea. That is why Gornahoor tediously examines those assumptions so the consequences will make sense, as they are absolutely necessary.

For example, this fellow at the Libertarian Alliance\footnote{\url{http://libertarianalliance.wordpress.com/2013/10/05/julius-evola-2/}} reads Evola as a preparation for something called the Traditional Britain Group. His impression was quite negative since he is the intellectual heir of ``Bacon-Locke-Newton-Hume-Darwin''. We recognize those names as the creators of the modern world, not traditional Britain, so that project seems stillborn.

Bacon rejected formal causes, so all that exists for him is the order of mixed things. Locke was the prophet of ``toleration'', so Britain now tolerates things that traditional Britain would have considered abnormal and unhealthy, if not insane. Hume denied causation and objective morality. Finally, Darwin claimed we cannot understand life since everything depends on random events and the survival of the fittest. That is modern Britain. There are no objective moral standards; no one sees the causal relationship between policy and results; things just ``happen'' and are to be accepted; the survival of the fittest means how one can get the most with the minimal output of time, energy, and resources.

As the saying goes, a broken clock is correct twice a day. Even if a man holds a correct view from time to time, that does not mean there is a necessary and logical connection between his thought and such a view. If I choose 17 in a roulette game, and the number wins, it is an arbitrary and accidental event. Perhaps I have some compatible friends who always bet red. We get a streak of reds and we think we are doing something right; a black number then shows up and we feel defeated. A rival group across the table always selects black. So we enter into a mock battle with them, winning some, losing some, all the while believing we are on some epic adventure. However, our final destination after many, many rolls will be back where we started.

Meanwhile, the House always wins, since it knows the mathematical principles of betting on the roll of the ball. If we do not abide by certain principles, our actions will be random. Our only recourse is the will to power and the survival of the fittest. Few people will admit to this, as true as it is. If there are no certain principles, there are only opinions. Then whose opinions will prevail? There is no way to decide it intellectually; the will to power can be quite devious in its operation, sometimes disguising itself in sweet moralistic terms or by making promises that are impossible to keep.

The bearers of the ideas of the modern consensus reality are quite powerful. Their opponents assume that the world they created will eventually fall apart or else result in difficult and unacceptable conditions. At that point, the masses will turn to them and their opinions. Unless they are driven by transcendental principles, that victory will be short-lived. It would only mean that the ball finally fell on a red number.

\paragraph{Actions Produce Results}

\emph{Pace} Hume, actions also have consequences, even foreseeable consequences. For example, if France was formed by 40 Catholic kings, then the overthrow of that spiritual and temporal authority will have to result in a France that is different from what it was prior to that. Many do not like the results they see, and it may take generations for all the possibilities of an historical upheaval to be actualized. The sane solution would be to support the status quo ante, but the identitarian groups I hear about want no part of that. Actually, just like the revolutionaries, they also oppose the status quo ante, usually with harsh, and unfortunately uninformed, criticism. It is hardly clear what then is the source of identity; it is certainly beyond doubt that race or nation in any biological sense is failing to provide the necessary bonds (note the word necessary; that does not preclude that some choose such bonds on the basis of personal whim).

This is the situation today of European identity groups. Their minor victories, say in Greece or Hungary, are no more than a brief run of reds on the roulette wheel. Perhaps, Hungary is an exception since its leaders do seem to be authentically Traditional, unlike most groups that are rallying around the New Right.

Another unfortunate case is ``Southern Nationalism''. However, their situation is the result of long past decisions. Keeping any moral judgments in brackets for the moment, African slavery was a form of involuntary immigration of an alien people into their land. Now they have buyer's remorse and object to voluntary immigration and seem displeased that the blacks in their midst have a different outlook on life. The common objection is that ``their'' particular ancestors did not own slaves. However, if you are a nationalist, you accept what comes with it; you cannot become an individualist temporarily when it suits your purpose.

\paragraph{War and Identity}

So, finally we can look at Alice Bailey from her book \textit{Esoteric Healing}. She understands that events happen beyond anyone's conscious will. However, for her, the changes that led to the modern world are a ``good'' thing. Identitarians should consider some of these ideas carefully, since it shows what they are up against. I wonder how the Russian pagans would consider the following passages.

\begin{quotex}
It might be truly said, brother of mine, that the rapid unfoldment of this heart lotus is one of the reasons why the world war she regards WWI and WWII as one long war could not be avoided. In one sense, it was a necessary happening (given the blind selfishness of the bulk of humanity) because it had become necessary to do away with all the old forms of government, of religion and of the crystallised social order. Humanity has now reached a point of group awareness and of group interplay of a deeply spiritual kind, and new forms were required through which this new spirit could function more adequately.
\end{quotex}

That is not an uncommon view today. It is impossible to convince someone of its opposite by any intellectual arguments alone. This is the interesting passage with too much to comment on:

\begin{quotex} 
In the Aryan race, which now dominates the planet, you have the first true appearance outwardly on the physical plane of the basic synthesis of humanity which will be---in its better form---so strikingly significant of the next major root race, the sixth. Inter-marriage between nations and races, the fusion of bloods for hundreds of years---due to migration, travel, education and mental unity---has led to there being no really pure racial types today. This is far more certainly the case than the most enlightened think, if the long, long history of mankind is considered. Sexual intercourse knows no impenetrable barriers, and people today have in them all the strains and the blood of all the races, and this (as a result of the world war, 1914-1945) will be increasingly the case. This development is definitely a part of the divine plan, no matter how undesirable it may appear to those who idealise purity of relationship, or how ruthless its application is at the present moment. Something intended is being brought about and it cannot be avoided. The urge to mate becomes peculiarly strong when men are removed from their familiar settings and experience the novelty of complete loneliness, when the normal inhibitions and customs imposed by family relationships and national standards are removed, when danger of death is constantly faced and the larger value submerges the lesser values and the usual conventional attitudes, and when the physical organism has been trained and brought by scientific treatment and heavy feeding to the height of physical efficiency. I am speaking in terms of physical effectiveness and not of mental efficiency, which may or may not parallel the former.
\end{quotex}

Her Tibetan source is very matter of fact, and makes no moral judgment. The book was published in the early 1950s, so it seems to predict things that were not commonly believed at this time. Those who praise the marshal spirit may want to consider the unforeseen consequences of war. This next passage reminds me that in the 1930s, Italian soldiers were bringing back Ethiopian wives from their African adventure; Mussolini had to pass a law against it, but that is the price of Empire.

\begin{quotex}
The armies of the world are everywhere and are spread over every country; racial transmigration is a universal factor, both from the angle of military necessity and from the plight of the civilians who find themselves in the path of war. This movement of millions of men everywhere is one of the paramount factors which will condition the new civilisation, and its importance is based upon the fact that in twenty-five years' time men and women will be a hybrid race whose fathers and mothers will be of every imaginable nation; white fathers will have had physical relation with women of every Asiatic or African origin, thus producing a fusion of blood which---if recognised and rightly handled and developed, from the educational angle and with understanding---will express in embryo the nature of the sixth root race, and which will be in fact HUMANITY without any racial or national barriers, with no so-called pure blood and exclusive castes, and with a new and virile sense of life because of the infusion of stronger stocks with the weaker or worn-out types and of the newer racial strains with the older and more developed. I hold no brief for the manner in which this is being brought about. It could have happened without war and through a conviction that all men are equal and human, and that the mixture of races would solve many problems; war, however, has hastened the process and the soldiers of all the armies of the world are having physical relations with women of all races, all civilisations and all colours. This must, whether regarded as right or wrong according to the code of ethics and standards of the observer, produce an entirely new situation with which the world of the future will have to cope; it must inevitably break down national prejudices and racial barriers---the first producing more effect than the latter during the initial stages. Inevitably a more homogeneous humanity will appear during the changes of the next one hundred years. Many attitudes and many customary reactions which today hold sway will vanish, and types and qualities and characteristics for which we have as yet no precedent will appear upon a large scale.
\end{quotex}

To oppose this is like opposing the weather. ``Conservatives'' can do nothing about it and I don't see that identitarians have made due preparations:

\begin{quotex}
Whether the conservative and the so-called strictly ``moral'' people dislike this worldwide happening has no bearing on the case. It has happened and is happening daily and will materially bring about far reaching changes. These inter-racial and mixed relationships have always happened upon a small and individual scale; they are now happening on a large scale. For the results of this due preparation must be made.
\end{quotex}


\flrightit{Posted on 2013-10-11 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{X on 2013-10-11 at 05:00 said:}

Mr. Salvo, I should say that sometimes you possess some genuine righteousness and true sanity.

\hfill

\texttt{Cologero on 2013-10-11 at 08:53 said:}

for the Hungarian viewpoint:
\textit{László Virág -- The Criteria of The Right: The Struggle Between The Right \& The Left Regarded From A Metahistorical Viewpoint.}\footnote{\url{http://www.youtube.com/watch?v=LG4r3-4UGYI}}

\hfill

\texttt{Cologero on 2013-10-11 at 09:05 said:}

``no one who holds materialistic or atheist views can be regarded as a real rightist'' --- Laszlo Virag

\hfill

\texttt{Thorsten on 2013-10-11 at 11:07 said:}

``no one who holds materialistic or atheist views can be regarded as a real rightist''

For this reason I always found the National Socialist critique of modernity's and Judaism's alleged materialism hypocritical and amusing as it indicates the degree of philosophical and metaphysical confusion endemic to that movement.

As for the prospects of preserving European ethnicities and races those things would naturally find a resolution of sorts in the context of a traditionally ordered hierarchical society. If there is any connection between physical race on the one hand and character or race of the spirit on the other, the races would find their just stratification within a society in which character and spiritual vocation serve as the primary organizing criteria for this society. This combined with the revision of immigration laws is the best that can be hoped for in terms of achieving both the preservation of European mankind and a civilization of a traditional mould. I think that the most troublesome alien elements would leave of their own free will in such as scenario.

\hfill

\texttt{Constantine Aetos on 2013-10-11 at 14:51 said:}

Don't forget birth and proper child rearing is another element in preserving European civilization.

By the way some interesting articles written by a religious Jew on Europe's suicide:

\url{http://www.atimes.com/atimes/Front_Page/KK17Aa01.html}

\url{http://www.atimes.com/atimes/Front_Page/FJ05Aa01.html}

\hfill

\texttt{Ash on 2013-10-13 at 01:49 said:}

The views expounded in these texts seem in line with Spengler's views on race. He has an interesting phrasing in one of his works where he says that ``what matters is not a pure Stock, but a strong one, which has a Race within it.'' This upends the National Socialist view on race as being a purely biological phenomenon. Rather, race is a phenomenon which, while it requires the ``use'' of the bonds and psychology of the biological ethnicity or stock, ultimately transcends them. Evola makes this same distinction in his works. For those concerned with moving into a Traditional current and passing it on, the key would be to form relationships and families bearing this in mind. Many of those Ethiopian women were likely Orthodox Christians who lived in the reign of the Emperor Haile Selassie. The question is, is that preferable for the soldier who may have instead married an Italian woman whose children became part of the generation of '68?

Were I the wise Guru of an identitarian circle, I suppose my advice would be this: ``Recognize the source of the Tradition from which your identity -- regional, ethnic, and European -- stems. Study that Tradition, immerse yourself in it. Secondary is the work of rebuilding European and Western identity, like seeds growing from the ashes of a burned out forest. This is only possible by accepting that we are, to use another's term, working as `subalterns.' Build your communities and plant those seeds -- literal and metaphorical -- and true progress will be being made daily.''

\end{sffamily}
\end{footnotesize}

