\section{The Opposite of a Fact}

After the passage of a sufficient interval of time, I have taken it upon myself to address some issues that brought unwanted attention to this blog. Specifically, it involves the unexamined assumption that there is a sort of natural affinity between the tenets of Tradition and various identitarian movements, which go under the label of ``New Right''. Although there are some points in common, there are nevertheless significant differences in the understanding of the world and the task at hand. First of all, taking a cue from Rene Guenon, let us be clear about the basics of Tradition.

\textbf{The first tenet} is that a mutual understanding between the world's religious traditions is possible because in their depth they are expressions of certain metaphysical principles. This is not at all the same as syncretism or indifferentism, since each religious tradition has its place and it, rather than any corporeal or bio-physical considerations, is the source of the identity of a people. Nor is it a call to replace an exoteric tradition with something more abstract, higher, or ``sophisticated''; faith is virtual gnosis, or else it is mere opinion, so it cannot be casually discarded.

\textbf{The second tenet} is that there is a cyclic process of decadence and that the West is, unfortunately, at the vanguard of that decadence. The immediate cause is the rejection or forgetting of its spiritual Tradition. The consequences of this forgetfulness are legion and have been recounted \emph{ad nauseum} by this point. However, the cure proposed by Tradition is quite different from that of the new right. According to the former, the West requires a healing through spiritual regeneration. This will begin with an awakening of a few to the full scope of the problem; they will inaugurate the recovery of tradition which will eventually gather influence.

The new right, on the other hand, is either irreligious or, at best, rejects outright its most recent Tradition. For them, the problem is ``bio-cultural'', that is, limited to the race of the body and the race of the soul. That is an impossible gulf to cross, since it leaves transcendence out of consideration. It sees the threat on the horizontal plane, particularly the alleged population replacement by alien peoples, among other grievances; it can only see symptoms and misses the true cause. To the extent it claims to be ``metapolitical'', it can only hope to persuade through ``infinite discussion'', first, the current culture makers and then, by extension, the larger public. Thus it relies materially on the democratic methods it opposes formally.

Hence, Gornahoor, with a worldwide audience, cannot be limited to Euro-centrism. Of course, our primary aim is the healing of the West so we investigate other traditions in order to revive our own. The end result, whenever it may occur, is not at all Euro-centric, since it affects the entire world. From the perspective of Tradition, the fundamental problem is not the encroachment of aliens into the West, whatever inconveniences that may bring, but rather the deformation of non-European cultures through the deviance and decadence of contemporary Western forms of life.

\paragraph{Natural Nationalism}

We recently pointed out what Metternich called the natural feeling of nationalism. Every society has the natural right to preserve its bio-cultural integrity. Yet, culture is the outer expression of the spirit of a people; if that spirit is deformed, it will show up as cultural distortions. It is not enough to save the body if the soul is lost. As a fish rots from the head, the challenge of bio-cultural integrity will come from the top. Since we dealt with these phenomena in the discussion about Vilfredo Pareto\footnote{\url{http://www.gornahoor.net/?p=6353}}, there is no point to do it again here.

There is another factor in play, however. The nobility did not fully share that sentiment, at least not in the same way. Julius Evola makes this clear when he rejects Hitler's claim that ``Being a street sweeper in the Reich should be considered a greater honour than being the king of a foreign country.'' Specifically, the nobility led their lives apart from the commoners. They would choose a spouse from the nobility of other nations rather than a commoner of their own ethnicity. In effect, they became a race apart, even to the extent that the people had rulers who were not even of the same ethnicity. This led to the idea of ``high culture'' as opposed to the folk cultures of the people.

Unfortunately, the cosmopolitanism of the nobility often led to the contempt of the masses they were supposed to be ruling. This feeling persists today in the ruling elite, even though no longer noble; they often have more in common with their peers of other nations than in their own citizens. Even on the right, there is too much contempt expressed about the ``mass man''.

Here is where the spiritual elite are supposed to mediate things, as Vladimir Solovyov pointed out\footnote{\url{http://www.gornahoor.net/?p=5323}}. The spiritual leaders give a voice and a consciousness to the folk, which lifts them out of the mass.

I can bring up a point here about some objections to the Venner post that I did not understand, since many saw it as a personal attack. I was told it was the use of the word ``vulgar''. Now its primary meaning\footnote{\url{http://www.merriam-webster.com/dictionary/vulgar}} is ``of or relating to the common people''. Because of the faux contempt of the common people, this word has picked up unsavory connotations, particularly in English (in French, ``vulgarisation'' simply means ``popularization'').

\paragraph{Facts and Opinions}

This brings me to the point, and it involves something distasteful. However, I have become persuaded that it is necessary, since, once comments have been archived, they become authoritative unless refuted. The opposite of a fact is not a mistake, which we are always happy to correct, but rather an unsupported personal opinion that pretends to be a fact. Hence we are compelled to address some of those pseudo-facts. For my source, I am relying on New Culture, New Right by \textbf{Michael O'Meara}. Unless and until the publisher or distributor sends out an errata sheet with the book, we shall consider it the authority on all things ``new right''.

\textbf{Assertion}: ``Venner was not, strictly speaking, a New Right author.''

O'Meara writes (p 17):

\begin{quotex}
After an adolescent stint in the Action Francaise, the precocious Alain de Benoist participated in Venner's and Mabire's Euope-Action, the first right-wing group to embrace issues anticipating the cultural and identitarian concerns that would hereafter define the European New Right.
\end{quotex}

\textbf{Fact}: Venner was ``strictly speaking'' one of the ideological founders of the European New Right.

There was an objection to our characterization of fundamental views of the new right. O'Meara mentions five currents that form the new right (p 29):

\begin{enumerate}

\item The anti-modern traditionalism of Rene Guenon and Julius Evola

\item Communitarian or volkisch currents and ``European nationalism''

\item Neo-paganism

\item Postmodernism

\item Science
\end{enumerate}

Someone objected to our characterization of the new right as neo-pagan; you can believe the comment writer, or you can believe O'Meara and Gornahoor. Your call.

There was also an objection to our characterization of the new right as rooted in biological elements. It is difficult to interpret point (2) in any other way. Also, point (5) is specifically oriented, according to O'Meara, ``to the life sciences and their genetic, eugenic, and ideological implications.''

There was an objection to the alleged influence of tradition on the new right, again in direct contradiction to O'Meara. But here is where the confusion comes in. The new right has absorbed Guenon's anti-modern critique, but not his ``traditionalism'' as such. This explains the one-dimensional understanding of those whose understanding of Tradition is limited to new right authors.

Although there are some points of contact between the two worldviews, there is also a gap. Tradition is looking for a spiritual as opposed to a bio-cultural renewal. Since most of our new right readers have abandoned us, this will probably be our last post on this topic for a while. There is too much misunderstanding, but as we have shown, it is not on our part.


\flrightit{Posted on 2013-06-10 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Ash on 2013-06-10 at 23:58 said:}

First, I must thank you for taking the time to write this article. More and more, there seems to be an influx of people using the world Traditionalist to their own ends. Among others is a certain podcast I see recently launched. Spouting lines from Evola, abhoring the decline of the white race, bashing Pussy Riot, throwing your support behind any State that happens to have problems with Israel and America, and listening to martial, neoclassical, and folk metal do not constitute Tradition (doesn't discount them either, I personally love folk metal). The core of the Traditionalist school is outlined well here; the core of Tradition itself, the meaning and reality reflected by that word, is a way of living and an experiential knowledge called Gnosis. Without that fundamental aspect, the rest doesn't matter in the slightest. Thank you also for the comments on nobility. The close relationship between Emperor Frederick II Hohenstauffen and his Muslim counterparts is something which knee-jerk anti-Islam groups like the EDL, or frankly even many within the Identitaire/Identiaere movement, cannot begin to understand in their current ideological framework.

On the complex nature of the movement which grew from the new right, a few words should be said. The nouvelle droite, of course, was neopagan in nature. However, from that has grown, apart from the body of intellectual writings of de Benoist, Venner, and others, a legion of offspring: the Identity movement in France and Germany, others like Casa Pound in other European countries, the German equivalent Neue Rechte, which publishes through outlets like Junge Freiheit and upholds a more Christian core, and so on. Where the New Right conflicts with Tradition, it is of course wrong. But the New Right is by its own admission an intellectual movement, and therefore, it can be changed and improved as required. De Benoist himself was influenced by Maurras in his youth; perhaps this long line of intellectual forefathers can be remembered today. Indeed, with blogs like this and others, it's already happening.

It's true that the New Right concerns itself with bio-cultural issues. But while the spiritual reality may be highest in the chain of being, without culture the chain would be broken. So if those of us influenced by their ideas remember first principles and make sure that the ``cultural'' critiques we develop are informed by them, the currently tense relationship might begin to flower.

\hfill

\texttt{Avery Morrow on 2013-06-11 at 12:03 said:}

Two out of four responses now are complaining that Tradition does not ``solve the problem'' of living in the world that the New Right proposes to solve, and I address both of them.

What is an ``action in the world'' but a public action? In the traditional sense of the Romans, Greeks, pre-1789 Europeans, Easterners, and Evola, political action that derives from private interest is not properly speaking public, and attempts to veil one's private interests as the public interest are disingenuous and a betrayal of the public sphere. Even Hannah Arendt, who was a big booster of democracy, understood that democracy \emph{destroyed} all the fundamental precursors of the former public sphere. Our actions are no longer public but social, aimed at gaining acceptance in a populist mass society. And those behaviors that society has seized from the realm of custom have had their energies redirected into ``social progress''. Reactionary politics, even ``think tank'' political theory, in the late modern world is like pissing into the wind.

Benoist does not understand this because there is little evidence he understands the social principles of Rome at all, much less the \emph{sophia perennis} which informed them and is the subject of this blog. His manifesto is \emph{On Being a Pagan}, which carries all the insightful ancient wisdom of noted philologists like Starhawk and Silver RavenWolf. As a set of first principles Benoist's neopaganism is a poor substitute for Catholicism.

Benoist wants awareness of ``classical democracy', but he doesn't even understand what it meant to be a citizen of Athens or Rome, as far as I grasped from \emph{The Problem of Democracy} and his interviews. The social principles of an age when citizenship was a privilege granted to a small group of freeborn families are not a model for a public sphere that has been uprooted and replaced with the whims of mass society.

We are long past the days when mainstream political thought might blow in the reactionary direction. World War II is done and gone, and we are all stupider than our forefathers who died in that war. If and when a reactionary wind does start to blow, it will not be a sweet or calm wind. The preparation needed to guide future generations is spiritual, because populist politics cannot be used to scare mass society straight.

\hfill

\texttt{Michel on 2013-06-11 at 13:14 said:}

This is a very interesting website; it however seems predominantly directed to White Europeans, with the aims, by my conjecture, of rejuvenating, propagating and sustaining by various expositions which would result in reactions on the part of the addressed, a `Western Traditional State of Mind'. That probably explains why `Christian mysticism' is referred to with such reverence. A truly noble cause, but unrealistic in terms of its goals, at least by virtue of the elements contained herein (perhaps it has `secret operations' of a more `radical' nature, or perhaps the members really are engaged in spiritual activities rooted in pure Unity that would constitute a true suggestive action which would hopefully lead modern men back to a proper Orientation). Whenever i meet so called `Traditionalists' (forgive the modern influence of systematically labeling everything), I am always shocked by their ignorance concerning `worldly affairs' that relate to the `counter-initiatic elite' as well as their passivity and general state of complacency. But perhaps i mistake disinterest for ignorance. At any rate, i say this because it relates to my initial statement of the highly improbable outcome that this blog is hoping for in relation to the decadent state of the modern West. Firstly, the actions are wholly `verbal' (and here i repeat, going by the material herein and the resulting speculations they arouse); discursive. Secondly, they depict a frame of mind that is scarcely informed on matters relating to the enemy(`counter-initiation') and an overwhelmingly large amount of information relating to its puppets (products of Western education who nevertheless have been brazened as `intellectuals'). Perhaps a little information regarding the activities of the `satanic' elite would help increase the magnitude of the actions entailed herein. It should be noted that while everything here seems to remain wholly speculative, the counter-initiates are actually involved in actions that bring about detrimental results on mankind, results that indirectly spark the many debates entertained on this blog and on other `Traditionalist forums'. Forgive the critique, but i am frustrated by extremely complacent people who think that by discussing a certain topic which by modern standards is `bold', `strange' or even `mystical', they are going to bring about any change, if this change is intended on a `collective' scale. If the bar to be surpassed is the chaotic rumble engendered by the modern world, then there really is nothing that you are doing to change anything, insofar as you wish to influence `other people'; this modern deviation that seems to be acting as the `bar' to be surpassed is being constantly engineered and manufactured by the most `diabolic' group in the world today, the counter-initiation, and hence all `verbal' responses in light of this engineering manifested as the state of existence in the West, can only be reactionary and hence utterly useless. It is like being ill, reading a few books that describe (vaguely- Guenon only used the Traditions as an illustration to `awaken' the modern man, i doubt his works even count as anything `preliminary' with regards to an Initiatic Order, they are not even Religious, thus they are very basic, and their `uniqueness' only shows the deepened state of confusion that the modern mind is in, they offer no `spiritual method'; if a man hides behind ten thousand lies, connected to each other by numerous threads, the snap of one scissor renders the entire mesh-work desolate, but that does not make the scissor as powerful as an axe; the Truth was simply presented to the modern man in a way that his deviated mind could understand, the `Basic Traditional Man' was well aware of these things that strike so many as `novel') the illness and strangely claiming to have defeated the disease, without recognizing and destroying its source. Gnosis and Wisdom are the answer, and these are `interior' and any `exterior' action can only be `vivified' with `Spiritual Power' directed from the `vantage point' of these two. But speaking of these things with a lingering modern mentality that is characterized by being overly `scholarly' and `speculative' won't do anything and only places one on a level subject to the very `suggestive influences' that the devil's saints are constantly propagating. I ask, does anyone here know the state that the counter-initiates have reached in regards to the propagation of the New World Order? Does anyone know that the very `scholars' and `philosophers' are but fragile puppets at the hands of a group of people that understand and are in `possession' of certain aspects of Reality that, in spite of being wholly intermediary (psychic), have devastating effects even on the Western elite insofar as they remain in their passive state? Christ said that in the end, the deception will be so great that there exists a possibility of this elite falling under its power. In his books `The Crisis of the Modern World' and `The Reign of Quantity and the Signs of the Times', Guenon warns against the adoption of complacent attitudes while thinking that mere `erudition' constitutes a true active reform that has effects... does sitting around for hours talking about how philosopher y was wrong in manner x when contrasted with philosopher z produce any important effects? The friends of Satan know the utter stupidity of letting such activities engulf one's entire enterprise, especially if this enterprise is directed towards stopping them. Anyways, i know I've angered some people, but i am right. In finishing, can anyone reliably, as an example, tell me of some activities they know that relate not to a puppet, but to a puppet master and what they are doing to stop or mitigate it, if this is their goal? I would have a large amount of respect for such an individual, because, being able to, if it is the Will of God, `fight' the counter-initiation, their level of Realization is truly Spiritual, for one cannot fight the counter-initiates from even a psychic `vantage point', if you've been there lately, you can clearly see their enormous power... it is only from the Unity of the formless that one, if Willed by God, can hope to actually do anything; and that is the truth, at the Center, one is one with the Great Architect, having found the lost Word, one is able to Realize the Universal Plan for beings in Being. I know this comment doesn't directly adhere to the elements dealt with in the preceding article, but i feel it does adhere to the `zeitgeist' of `Traditionalists'.

\hfill

\texttt{Jason-Adam on 2013-06-11 at 13:22 said:}

I submit this proposal to differentiate the true man of Tradition from the false : ask yourself, what is your purpose in life -- is it to reach enlightenment or is it to change the world ? If you want to reach enlightenment you must abandon all attachments to the things of this world which are transitory and falsehoods. The world is hell and we must escape from hell to truly live life. Attachment to politics is another craving we must escape from. Remember that the the ideal state is one that is organised in a manner that is best suited for individuals to achieve enlightenment, in other words the aim of traditionalist politics is to overcome politics with spirituality. I welcome all comments.

\hfill

\texttt{Graham on 2013-06-11 at 21:17 said:}

I agree with you as well, August, on the most practical approach to a restoration. You've described the approach I've chosen. However, if I may, the reason it's an unpopular approach is not that it seems boring (or not only that), but more fundamentally, that it's arduous. I'm sure there's no need to go into details.

August, what role should members of the vaisya caste play in restoring Tradition? How can vaisyas participate, beyond the general approach on which we agree? I'm interested in your thoughts.

\hfill

\texttt{Ash on 2013-06-11 at 22:45 said:}

By fulfilling the role of the Vaisya as best as possible. Take your principles into account in business and work, marry and raise a family, practice to the fullest extent the householder tradition as it manifests in your faith, and use what public role you may have to defend truth and your community. Alain Danielou's book ``Virtue, Success, Pleasure, Liberation'' would probably be a fine work to consult.

I'd note on the ``naming your caste'' practice though, that I think it's a bit over-thought. One's caste is innate. It will manifest in ones life, whatever one does; that being the case, it seems a bit redundant to overthink ``what caste am I'', as there seems to much danger in trying to live in strict accordance with one caste one imagines himself to be, and thus unintentionally stunting ones true potential.

\hfill

\texttt{Cologero on 2013-06-11 at 23:59 said:}

``Just to be exact, I will note that the cure was proposed by Guenon, not `Tradition', and its eventuality is not a foregone conclusion. The Fates might decide that a restoration of the West is not necessary.''

Yes, August, the ``restoration of the West'' is but one option. Guenon actually proposed, or foresaw, a few possible outcomes as we have mentioned several times: \textit{The Prophecies of Rene Guenon}\footnote{\url{http://www.gornahoor.net/?p=35}}. Your proposal is what we have in mind, since the greater spiritual combat is what is necessary. We have encouraged readers to bear sons; I would prefer to hear news of such births to long discussions. We have never advocated a revolution, but rather the opposite of a revolution. We certainly do not write for the average Westerner and prefer that they move on.

\hfill

\texttt{Timotheus Lutz on 2013-06-16 at 16:48 said:}

According to Alain de Benoist, Evola and `Traditionalism' have had only minor influences on the New Right:

TELOS: . . . what is the significance of Julius Evola for the European New Right in general and you in particular? . . . 

BENOIST: Why Evola? Among all the authors who have been used or cited by the New Right he is certainly not the most important. Evola belongs to the school of integral traditionalism, along with people such as Rene Guenon, Guido De Giorgio or Frithjof Schuon. The New Right does not belong to that school. As for myself, I have never been under his ``influence.'' The authors who have had the greatest influence on me are Louis Dumont and/or Arthur Koestler. I have also criticized the belief in an original Tradition (in the singular and with a big T), which is dearly a myth. But it is also true that Evola published interesting and valuable works about ethics, spirituality and philosophy of religions. His work is vast and one does not have to accept all of it. . . .

\url{http://home.alphalink.com.au/\~radnat/debenoist/alain15.html}

\hfill

\texttt{Ash on 2013-06-16 at 23:06 said:}

Thank you for the compliment, it's still developing but I'm hoping the content proves useful.

In terms of how Vaisyas can pursue esoteric paths, what about societies like the Freemasons (in their original form) or Sufi societies in countries like Turkey? Many of these are examples of people who can be characterized as Vaisya pursuing the esoteric cores of their faiths with greater zeal. There certainly is a precedent for it, in my view. Again, Danielou's book is a fantastic source for the ``householder'' tradition.

\hfill

\texttt{Cologero on 2013-06-17 at 19:39 said:}

That's our point, Mr. Lutz: why Evola and why not Mary Poppins, if the interlocutor had to pick an influence at random. Perhaps because Evola himself was a supporter of GRECE in its nascent stage? But that is precisely the point we made: the new right took some interesting things from Evola, but rejected the elements that were specific to Tradition.

Perhaps you could post that on a web site that presumes a close connection between the two and report back with the results.

\hfill

\texttt{John Morgan on 2013-06-18 at 17:01 said:}

Ash, I think you have written the most balanced, measured and intelligent contribution to this debate thus far.

\hfill

\texttt{John Morgan on 2013-06-18 at 18:28 said:}

For someone who wants to be a pure and true traditionalist, it seems to me that apoliteia is the only road to take. Which is why I decided that I am not a doctrinaire traditionalist. It may end up being futile, or not, and it may just be my own weakness, or not, but I cannot simply content myself with watching the world decay around me and not attempt to do something about it. As I have said all along, the New Right is not a traditionalist movement in the strict sense of Guenon and Evola, nor has it ever claimed to be such. However, the points of overlap are clear. One may not think that this is sufficient, or one may disagree with their approach or methods, but in the absence of any other movement that is attempting to address these problems, it remains the only endeavor in this direction. For those for whom all political thought and action at the present time is a useless gesture, I don't see why it's necessary to engage with political issues at all.

\hfill

\texttt{Ash on 2013-06-19 at 00:11 said:}

I'd say that this holds true for the Brahmin caste, for a lack of better words, although even this seems to be a bit flexible. The temporal power of the Pope, championed by de Maistre, one of the principle influences for Gornahoor, is most certainly not a good example of apoliteia. For the other castes, however, this seems to be false. Is the kshatriya not the ruler and warrior? Even the merchant and worker castes were active in the life of the village and township, if at guild and parish level rather than throne and altar, during the medieval period which Gornahoor does hold as an example of a traditional age. Even if the age of kali yuga demands different practices, it seems that this means simply that the work that the ``politically active'' man does would take a different form. That doesn't necessitate joining some group; banding your neighbours together to stop Monsanto from taking over traditional community structures and seeds would be a fine example and in line with the principles espoused here.

\hfill

\end{sffamily}
\end{footnotesize}

