\section{The Constitution of a Traditional Society: Warriors}

\subsection*{The Role of the Warrior Caste}

\begin{quotex}
This is part I on the role of the Warrior caste in a Traditional Society. Here Guido De Giorgio describes the role of the warrior caste and its relationship to the priestly caste. He then relates that to the Great War and and the Lesser War.
\end{quotex}


In the active life, they are the bearers of power and therefore constitute the second caste of traditional society whose duty is the maintenance of activity guided by virtue. We say power only in the arena of the active life because in the contemplative life, true power is manifested in its highest reality, from the invisible to the visible, from the divine to the human, and thus represents the supreme authority that belongs to the priestly caste. Power in the active life is realized by giving a sacred character to every manifestation; hierarchically organized, the Warriors sacrifice themselves in order to remove the purely contingent aspect from human activity and to make it a kind of need that is accepted by consciously sacrificing themselves for the fulfillment of divine reality. They are the hosts of the earth and choose the most difficult path in it, keeping the sanctity of the intent unbroken in the pang of the most intense activity by virtue of a continual ascesis that is a purification and a preparation for glory. Since the Warriors reach glory that makes them the victors of the time in perpetuity that is immortalized for ages and their name is preserved as the constant example of sacrifice. They choose to be victims to affirm the victory of man over death even where death reigns: voluntary and active victims, arbiters of their destiny, aware of the transience of the flesh and the perpetuity of their example. As everything pivots on the contemplative life for the Priests, so for the Warriors everything is based on the active life because they constitute its summit and its law. In fact, the active cycle reflects the contemplative, and as the death of the human is the prelude necessary to the fruition of higher states, so the Warrior lives in order to die in the awareness that everything is vanity except victory over death, which is also the seal of glory.

They obey a purely interior discipline that purifies the passions, exalts them, and directs them toward a single end, the affirmation of power in the intensive deployment of a force that acts materially, but has its origin in the world of Rhythms. As the spiritual and the noetic dominate in the Priests, in the Warriors the psychic dominates, the hidden network of Rhythms that heighten human possibilities creating the heroic turmoil which, like a flame, nurtures the great warrior virtues.

No one, more that the Warrior, is averse to Silence, to the world of the Spirit, because he voluntarily chose the active life in its most violent form to set his strength against it, to defeat it with his own violence, exalting his hidden energies and carrying the Rhythms to such a high tonality that it encroaches on the Forms and bends them integrally, surpassing them. We observe that this overflowing of strength cannot be absolutely effective without a true and proper ascesis, a self-denial in every moment for self- assertion only in victories, unless this assertion has the aspect of rhetorical jubilation, but rather it is like the blossoming of now equalized warrior virtues, recognized in the flower of all the activity precisely because surpassed and triumphant. While all other men compromise, they are the intransigents, those who make the obstacle the prize of their strength, for which the cement is the sanctification of man: they know that only death can placate them, so in confronting it, they escape it, and in escaping it, they confront it because it comes only when the peak of power is reached in the peak of abandonment.

In this sense, peace is their war and war is their peace. They are appointed by the Priests who sanction them the sacred character of their activity with the rite, which is uniquely directed to self-sacrifice as an offering of all of the active life on the threshold of contemplation, a sacrificial purge of the existence that is continued even after death, and even after death. While the contemplative death that is faced by the Priests can neither be seen nor understood by the masses of men who, as we said, believe but do not know, they do get a glimpse of, but do not understand, the degrees of integrative knowledge, while the death faced by the Warriors is visible to all; it reflects, in an active place, the former and is carried out, analogically like the former, with a sacrifice whose value for men is as much higher as it is more apparent. But while this sacrifice is fixed plastically for the masses that relive it in its most conspicuous phases, its secret rhythm and intimate character elude them. In the Warrior the active life asserts itself by exalting itself and in the fierce heroic heat, it denies itself: in fact, the force that animates the Warrior is love in its most destructive capacity and devotion in its most constructive form: love and devotion constitute the warrior cycle which is that of triumphant virility. While knowledge dominates for the Priests, for the Warriors, love dominates, because all their strength is a type of offering in a continuous dissatisfaction that is satisfied only with death. While they cannot know the divine world, they love it and tend to it, climbing over human limitations with a constant overcoming of Rhythms over Form, of the psychic over the organic that it sustains, it increases a hundredfold the resistance to the obstacle and the force that surpass it, establishing a formidable aura of secret flashes that the ancients symbolized with the real, but invisible, divine assistance of the heroes.

The Forms, in heroic action, undergo a greater or lesser inflection in accordance with the intensity of the Rhythms that make them jump, by encroaching on them, replacing them, fringing them, so to say, with an imprecision that makes them more indefinable, while the Warrior himself is dragged, aware only of the sanctify of his sacrifice, in an always more overwhelming turmoil that is crowned, while filling up with death. But we can say here that love defeats death as glory defeats time, not in the absolute sense that is realized only through the knowledge and in the knowledge which dissolves every duality in the integral fulfillment of the divine.

It is necessary now to speak of the great and the small war to make precise the two fields of active contemplation and heroic activity.

The Great War is the development of the being in the surpassing of human conditions that are first confronted and then surpassed, destroyed, transformed, and resolved only in the divine sphere: the enemy to defeat is the man who must be struck centrally in himself: the field of action is his own heart that must be emptied of every dross: victory is that of truth over ignorance, of divine reality over cosmic and human illusion that is dispersed with the realizing knowledge just as the fog breaks up in the sun without residue, and that which disappears had never existed in reality. Victory in the Great War is Divine Solitude, the decisive summit of every ascesis where nothing remains except pure aseity and what in Him is absolute ineffability. The starting point of the Great War is the non-duality of the human and the divine: the destination is divine unity: but since what vanishes is not a reality but only ignorance, once this is dispersed, only what always was remains, that cannot be said of the transient, but only and uniquely of the eternal.

Now this victory, which is the only real and definitive victory of Life over death, which, for simplification, we say call the resolute peak, while in the reality of its development it is manifested as a difficult complexity of states, it requires such a concentrated force, a task so resolute of the whole being who must be transfigured in a succession of progressive realizations, from being able to be produced only in externally favorable conditions, not that these are the necessary principle, but practically they become so. Hence, the isolation of the Temple for the priest, of the remote place for the ascetic, the hermitage and finally an ensemble of conditions that favor and facilitate this which is the highest and the holiest of all rites, the most heroic of all actions, and the most perfect of all tasks, the killing of man in the Law of God, the real abandonment of Rhythms and Forms dissolving in the allness of Silence. Now evidently whatever type of action is contrary to a similar achievement, and every rhythm must be recomposed, every whirlwind assuaged for the achievement of the Great Peace that is the true Victory of the Great War. The enemies to defeat in fact are so numerous, so terrifying, the dangers to avoid, the states to revive so repetitive and successive, that only the integral cessation of every activity can make victory possible. It is about surpassing the human and the cosmic in all their forms and in all their rhythms, to remake the creative process in the reverse path, to lead to where God formed man with His breath, to be able to initiate the paradisiacal cycle. This is the model barely noticed for the Great War that constitutes the traditional secret, the sacred deposit held in trust by those who know, i.e., the Priests.

One can understand from that how much this caste surpasses all the others and is not, almost the truth, a caste, but a true and proper supercaste, because of its absolute preeminence. Thus its supreme authority, its consecrating, legitimating power, the dependence of the three castes on it, the justification of their existence, because nothing that is accomplished by men has value in reality if it is not sealed by the divine spirit.

Now the Warriors themselves are consecrated by the Priests, nor can they not be, without ceasing to belong to their caste since the war that they fight must represent a reduced version of the Great War in order to be legitimate, it must lose, i.e., the external character and conquer a deep meaning that justifies its attainment. The Warriors therefore submit themselves to an inner discipline, to a true and proper ascesis that will consist in their depassionalization so that by killing, they know how to kill themselves, by defeating, to defeat one themselves self, by considering the enemies as victims, in the sacred sense, and not as flesh from the slaughterhouse, by respecting them as themselves, by passing over the carnage with the love that redeems, over the contaminations with the purity that justifies, over death, and over atrocity of death with the consciousness that nothing can die because nothing can be born, only the eternal existing in its inaccessible reality.

\subsection*{The Devotion of a Warrior}

\begin{quotex}
It seems to me that \textbf{Guido De Giorgio} and \textbf{Julius Evola} agree on the role of the Warrior caste, with some fundamental disagreements. The first obviously is that Evola fails to understand the relationship of the warrior caste to the priestly class. The second is more difficult to express. Evola is ``cold'', i.e., for him, the warrior is cold and detached. We could say that for De Giorgio, the warrior is hot: he loves, he is devoted, he is faithful. For Evola, the warrior fights as it is in his nature. For De Giorgio, too, it is in his nature, but he fights for his family, his land, his Temple, his lord. We can note here how the notion of bhakti yoga has been sentimentalized and the understanding of devotion trivialized. There is something else, more subtle, to consider, but since it is not part of De Giorgio's purpose, I'll save it for another time.
\end{quotex}


The Small War acquires a deep ascetic significance and imposes on the caste, only to which it is reserved, not slight surpassings that have, as we said, love and devotion for its foundation. The Warrior is guided by his love for the Divine Principle that is close to him, almost accessible and even distinct, in which he centers himself when the heroic peak happens and the confluence of Rhythms generates the supreme passion, that of dedication to God of the carnage: then a super-action of the normal human faculties operates, that adapt, raising himself in the highest development of the consuming fire and the Warrior becomes the great sacrificer, the
\emph{victimarius\footnote{\url{http://referenceworks.brillonline.com/entries/brill-s-new-pauly/victimarius-e12204070?s.num=0}}},

and the baptism of blood is a catharsis that renews him, freeing him with his violence from human brutality, redeeming him as in a purifying basin Ex 30:18 that cleanses horror with piety at every death. Piety in fact pushes him to sacrifice of himself and of the enemies in each one of which he catches sight of his own effigy; in the clash of a fight, he seeks to beat back the resurgent duality, the terrible ``thou'', the shadow of himself, the eye that looks askance at him, the hand and the arm that threaten him, the love in which his own love is mirrored, the ``other'' who blocks his vision of the Supreme Unity. He kills and gets killed in the name of God, in the name of God he defends his lord and his land since his weapons are blessed and he is invested just as his lord is, and his country is the place fixed by God for the conquest of heaven, the support, the base of his ascension, and he must defend it in order to keep God's law in the world, as long as the enemy does not deprive him of God by snatching his land, home, temple. His caste is the guardian of sovereign power in which the virtues are combined, that when entirely followed, lead man to the Edenic fruition, restoring his eternal and original country to him, whose necessary starting point is this land. The king and the country are made sacred by the radiant vision of divine Glory since in loving it, the warrior loves God, and in protecting it, he defends God's possession.

Note that the realization by means of warrior ascesis is always indirect, while the priestly ascesis is direct or final: the former is accomplished through the world of Rhythms, the latter in the Silence, the former remains in duality, the latter reaching unity from non-duality. This difference is essential and eliminates whatever arbitrary reduction that would subvert the traditional order. The warrior caste obeys its own law that is justified only through a higher law which, originating in the divine, extends it. In fact, if the Primordial Tradition is taken as the starting point, the castes are reduced to just one, that of the Priests, nor is it put there for others because in this stage only the contemplative way exists which excludes any action: there is no caste in reality in the Primordial Tradition in which \emph{human unity} is realized in its most absolute expression. The King here is the high priest and one can speak of regality only in the sense of cognitive absoluteness. But in successive traditional forms, the distribution of human possibilities is necessarily carried out, because we are already in a greater state of decadence and complexity through which life assumes a internal and external double aspect; in this stage, the priestly class assumes an absolute position that puts it back in the Primordial Tradition, that of the Warriors is internal in respect to the purely exterior Workers but, considering the whole tradition in its triple aspect of Silence, Rhythm, and Forms which correspond to the Priests, Warriors, and Workers, the second caste occupies an intermediate position of equilibrium between the two extremes and serves to maintain the traditional unity because its attribute is power.

Representing the traditional boundary circularly like a fortified city, we will say that the center, the Temple, corresponds to the priestly caste, the periphery, i.e., the ramparts, to the Warrior Caste, while the workers remain in the middle. Considering such relationship, the Warrior caste occupies the most dangerous place in the traditional defense, the one most exposed and, if one understands the value of this depiction in relation with what has been said, the Warriors are the lords of Rhythms and the art that is most appropriate to them is Magic.

Whoever is capable of reflection will understand without difficulty how, this caste living in the paroxysm of action, must necessarily be involved in the wave of hidden forces that are unleashed especially when the rhythm of the active life is intense, so that the warrior ascesis has as its principle goal the knowledge of the world of Rhythms, of the laws that govern it, and the defense from the traps of the field of shadows. That appears more clearly to those who consider with attention life and the work of the great conquerors where one easily sees how materially infinitesimal elements have engendered inevitable imbalances. What is true for every man is doubly so for the Warriors who live intensely in order to confront the most terrible enemies, those who are invisibly close to us and of which the visible enemies represent a type of material duplicate.

Speaking of this caste we implicitly asserted the necessity of war, but not less implicitly we asserted the need that it is uniquely entrusted to the true and proper Warriors that they have the legitimate use of it. Now as one does not become a priest, so one does not become a warrior, and the contamination of castes has unfortunately produced the current degradation since, with democratic leveling, introductions of every type have taken place, distorting the traditional order and engendering the fall of a truly civil society. A return to normality worked with discernment, progression, and measure would permit the restoration of the sole and true hierarchy in the attributions of caste conformed to the nature of man, to his possibilities, and to the effective development of his more specifically productive activities.

The warrior is the ascetic of the active life and hits discipline is purely interior: whoever cannot kill himself will never be able to kill or, better said, in killing he will profane life and death because his cause is not holy. Now a man kills himself by his own hand or with the sanctified arm in a way that the enemy is visible and close, otherwise a true assassination is accomplished and not an act of purification. If war, as we said, is justified only as the symbol of the Great War, it is inconceivable that the battle be not focused, faithful, open, frontal, because only by knowing oneself can one dominate himself and only in seeing his own enemy can he be surpassed without the battle and the carnage having the bestial character of the battle and anonymous, unholy, and sacrilegious carnage. If one is capable of developing all the aspects of the question when placed in its true light, one understands that war must be made by warriors and not by the entire democratically leveled people and violently carried back in great conflicts to the state of savage desperation: it follows from it that war must be normally conducted man to man and, in its most typical expression, brought back to the preliminary dispute, to the duel.

We said that the Warrior caste is founded on a dualistic base that the Latin term \emph{duellum}, the ancient form of \emph{bellum}, expresses with perfect obviousness: war is the expression of this fundamental duality of which love and hate---the latter reducible, if you reflect on it well, to love---represent two extremes: in the event a traditional restoration is produced, which is difficult but not impossible, and which, in every way, only modern devirilization can gravely obstruct with its two most typical expressions, pessimism and skepticism, war should be led back to its fixed normalizing function, i.e., to exist permanently for the single caste of the Warriors. The duel that so called civil society of the West, barbarous in reality, repudiated with satanic thoughtlessness while they turn current conflicts into a hellish game of machines that work the foolish and most ungodly destruction, would be the essential condition of the return to normality in case it was led back to its natural, spontaneous, i.e., effective, form and not to that modern parody that stands to the duel as the courtier stands to the knight.

If normal arms were only in the hands of the Warriors, the duel would take back its traditional value and would constitute the customary standard training of the caste, developing the sense of honor that has now faded in the heart of men. The duel with cold weapons\footnote{\url{http://en.wikipedia.org/wiki/Cold_weapon}} \emph{arma bianca} is the most noble, the purest, the highest expression of the warrior caste because it establishes a reciprocity of force, honor, and love in two men confronting each other, who are saved by killing each other and sanctify themselves by being killed. One observes that duality remains before and after: before life against life, after life in the face of death. And this duality is precisely the permanent indication of the warrior caste and is the basis of its purpose. War in a traditional society is the amplification of the duel, but other factors intervene: family, country, lord. It should have led back to normality, limited to the only caste that can do it without involving entire peoples, working that chaos that compromises the very existence of a civilization. Brought back to clarity, to the brutality of the ancient wars, although circumscribed and delimited, it should be, if not permanent at least frequent, giving in that way the acceleration necessary to the rhythm of life that makes it deeper and more fertile for revelation.

Modern wars---and for us, this ``modernity'' lasts for centuries---are the products of the democratic degeneration that has leveled humanity substituting for the Warrior caste a fictitious hierarchy in times of peace, even chaotic in times of conflict when every citizen must become what he never was, i.e., a soldier. The soldier obeys external discipline while the warrior obeys the inner: the warrior is always a warrior, the soldier can occasionally become a warrior through force, not by choice or by natural inclination. But there is a more important consideration. The Warrior is the refusal of sentimentality and among the ancients it did not exist: only modern degeneration has created through the monstrosity of democratic conflicts the romanticism of war, because in these conflicts in which all the people participate, there are only mediocre warriors, while the mass is constituted by a type of human gelatin that trembles with pathetic oscillations at every twisting that violates his nature. Instead, war reduced to lesser proportions, but with a more intense and more frequent rhythm, because it is entrusted only to men and not to machines, to Warriors and not to other castes, would be an indispensable element of a full and fecund life if led back to the traditional type, i.e., sanctified by the rectitude of causes that would give rise to it. Its delimitation would constitute a permanent and normal base of the activity to paralyze by no means all the people, but to intensify the rhythm of the active life with a continual contribution of strong and decisive reflexes.

\subsection*{Conclusion}
\begin{quotex}
This is the final segment on the role of the Warrior caste in a Traditional Society. Here we see that the point of war is to establish peace. The respective spheres of influence of the priests and warriors are made clear, although caste conflicts can still break out for the reasons specified. \textbf{Guido De Giorgio} then relates the roles of the castes to contemporary events circa 1940 : when priests are no longer the bearers of the holy science, there is a state of perpetual war and democratic leveling, then spiritual ruin results, as man is left to his own devices.
\end{quotex}


The return to normality restores the people's strength established of the Warrior caste which abides by love and whose sacrifice, different from that of the Priests, comprises an ascent of psychic energies oriented in the direction of heroic glorification. As action is subordinated to contemplation, so the warrior is subordinated to the priest, without however this subordination equalizing or confusing the two domains that are autonomous and separated and which constitute the spiritual authority and temporal power. These two jurisdictions are as cleanly separated and distinct as eternity is from time, therefore the two domains are unmistakable through the impossibility or reducing the first to the second or of extending the second to make it correspond to the first, since the eternal cannot be contained in the transient, nor can the transient cease to be such and to coincide with the eternal. This only consideration suffices to show that the temporal regiment is independent of the spiritual without however opposing it in any way, but rather in parallel turning to itself up to what is the maintenance of traditional unity that contains the both together and harmonizes them.

If in the primitive phase of perfection, with life being only contemplative, we can acknowledge a true unity of development oriented toward purely spiritual ends, this is not possible in the subsequent traditional forms where action appears immediately next to contemplation, temporal activity next to the spiritual, from which arises the separation of the two domains.

But this separation is necessary for the purity of the ends to which activity and contemplation tend: a fusion or an overpowering in part by one of these jurisdictions would constitute a true anomaly because it would confuse the divine with the human, the spiritual with the temporal, the sacred with the profane without reaching any profitable result. When we spoke of subordination we meant it only hierarchically, i.e., from the unique, traditional point of view that embraces and contains the contemplative life and the active life: in this sense one cannot escape anyone that the temporal is subordinated to the spiritual as and in the same measure that the human is subordinated to the divine from which it derives, and the existence and the justification of existence.

The Priest consecrates the Warrior and arranges that his action returns to the traditional sphere, but in his sphere the Warrior is autonomous and he neither can nor must collide with the Priest since his activity does not oppose the spiritual activity carried out in the bosom of the first caste. In a traditional society every dispute between the temporal and the spiritual is resolved in the unity the encompasses them; if, in fact, there is conflict, this must uniquely be ascribed to the temporary deviation of one or two domains due to essentially individual, and therefore unimportant, factors.

As long as the Priest remains priest and the Warrior, warrior, no conflict is possible because each one of them will understand he belongs to a sphere whose activity is clearly separated and distinct: this knowledge, this awareness, must guarantee the harmony between the temporal and the spiritual. But whenever, through complex circumstances that are not always easily discernible, a serious conflict between the two powers may happen, it is necessary to come back to the principles that safeguard traditional unity and then they will be resolved hierarchically in the seat of pure spirituality to rectify the deviations and annul the deflections that can be produced by one side or the other, with leading the two castes and therefore the two powers each back to its own domain.

The elimination of the purely individual factors will produce the placation of the dispute, because unfortunately these will take place only through the surpassing of certain personalities who respectively break the law and the norm which regulate activity and establish the limits of the two castes. The relationship between the two powers is very delicate and the development of Europe shows how often the disputes between the spiritual and the temporal have assumed immense proportions up to the point of becoming outright hostility. There is a reason for all that: the imperfect realization of the traditional unity, that must be ascribed to the defection of the priestly class which can always make its authority matter, with well understood spiritual means, provided they don't distance themselves from sacred knowledge and in this, it only demands the defense of the highest values of the spirit. The temporal cannot undermine and diminish the eternal, and therefore the temporal power, in any way whatsoever and by whatever means, will not be able to harm the spiritual authority which will always know how, when it wishes, to withdraw from the world and allow it to deteriorate because it is deprived of divine help. Reflect very attentively on that in order to understand what is happening in Europe and in the world.

Nevertheless the bearers of sacred knowledge must at all cost maintain contact with traditional principles and undergo whatever duress, submit to any violation, in order to continue to save men by getting them back in touch with the eternal. The spiritual authority asserts itself with quite other means than those with which the temporal power asserts itself and we think that, in case of conflict, through oppressive or destructive that is the deviation and the overpowering of the temporal power, the spiritual authority will always have the upper hand, if it obstinately remains in its domain and does not descend into open battle with the temporal on the terrain of activity properly assigned that is not absolutely of its sphere.

We think instead that as much as it will stay in its domain, so much more effective will its own work be because therefore, as was said with the greatest absoluteness, it assumes its authority from God and not from the world, and its strength, nourishing itself fervently on the divine, will not be able to triumph over hidden dangers, incomprehensions, suppressions, and apparent collapse. Because if in addition the deviation of the temporal power assumes such proportions by becoming absolutely anti-traditional and sacrilegious and the Warriors, forgetting themselves, insulting the principle of love, which is the foundation of their existence, were exceeding their deplorable actions, compressive of every traditional spirituality, know that nothing can touch, profane, and be admitted to God's Truth that is not God himself and that also the ruin of the world would leave perpetually together the Silence or where the divine mystery unfolds.

But we are convinced that the dispute between the two powers is almost always due to a remote deviation in the very heart of traditional unity and above all in the priestly caste whose members have often forgotten what is their only and great strength: sanctity. Whenever the spiritual authority, without any concession to the profane, takes up its true function, the maintenance of God's truth in the world for its salvation, and the Priests, who are its bearers, may not fail in their task which is contemplation and not action, effective realizing knowledge and not pure external, literal science of holy things, the traditional unity would harmoniously include the two distinct powers, but in parallel tending toward a single end, the return of man to God, through time, into eternity.

The temporal power in fact, of which the Warriors are the bearers, offers as terms the realization of the divine through time, in a concert of all the faculties tuned by their own equilibrium. Until now, in regard to this caste, we have spoken of war, but it is now necessary to consider the positive aspect of war, i.e., peace. War is an imbalance that produces a balance which in its turn is broken through the intermittent character of human affairs: but if one considers that in analogy with the spirit whose domain is eternity, war appears as a renewal, a purification, radical emendation, followed by the achievement of a state of absolute and permanent pacification that represents the fruition of beatitude.

As much as equilibrium is superior to disequilibrium, so peace is superior to war of which it is the crowning achievement. No consideration of aesthetic order may contaminate the vision of the truth in the seat of pure spirituality. The Warrior yearns for peace since his war is holy only if it leads to a permanent and definitive peace. War represents the germinative suffering, and peace, the flower and the coronal fruit. If intermittence is a type of destiny of human and temporal affairs, it must never be considered as a cyclic law on the part of the Warriors and the Leader, meaning to say by that, that each war must be completed for the end of the disequilibrium and the realization of peace. That intermittence exists, but it does not make law, this is the foundation of the right and healthy development of temporal power. If, as we said, war is in analogy with the Great War that represents the permanent conquest of the divine, it must tend to peace and this must be the true and holy goal of the Warriors: only a soft and superficial esthetic, ignorant of the profound truths of certain analogies and the symbolic relationship between the temporal and the eternal can consider war as the end in itself, a spastic state, fascinating for its own anomaly; only when true peace is not realized, can the nostalgia of war be felt, this very dubious sentiment of order because it belongs to the purely sentimental sphere.

In order to prevent this anomaly, it would be necessary to avoid, above all else, the wars that are not decisive, considering them as imperfect actions, true and proper permanent crises that contribute to the crumbling of traditional societies.

What has happened and is happening in Europe must be considered under this light and then one will understand the gravity of the current situation in which, through the confusion of castes, through consenting to an absolutely bestial democratic leveling, the detachment of the human from the divine takes place and therefore the very end of man is prepared. When we speak of the end, we mean the true and proper end that is certainly not that of the flesh, as every person gifted with the most elementary good sense can comprehend, but the spiritual ruin, the collapse of those bridge that connect the world to the Superworld, of which follows the instantaneous formation of the hidden ways that connect the earth to the darkness of the Underworld. If humanity does not want to become an immense and sterile hell, it is necessary that it brings back the great traditional ways by which only will new bases be found for new developments.

The Warriors must participate with the Priests for the restoration of order because these two castes are the bases of traditional unity: from their concordant action everything can be made smooth, the one being the bearers of sacred knowledge, i.e., of wisdom, the other the judges of power; the one, if you will, the club bearers, the others the fasces bearers. The harmony of the spiritual and the temporal can rebalance the world against all pessimisms and skepticism so acute and stupidly brayers of the current times; in no mode will be realized a permanent arrangement of the world without the cooperation of the spiritual authority and the temporal power because the return to normality must be effectuated from above to below, from the higher to the lower, and not the other way around.

For centuries the masses flounder in bestial tumults and will still flounder as long as the Priests and the Warriors permit it: that if these two castes feel the enormous weight of their responsibility and have the awareness of their respective domains and their true obligations, no one could prevent the restoration of traditional unity, the new equilibrium that would reopen the great passageway to eternity for the glory of God in heaven and the peace of men on earth.

\flrightit{Posted on 2013-03-04 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Saladin on 2013-03-04 at 22:19 said:}

Wow! That was intense! Cologero, you mentioned in your earlier post that comments have declined by 50\% since you started posting translations of de Giorgio. If everyone is like me, it may be due to being awestruck after reading materials like this. Thank you for making these translations available!

\hfill

\texttt{Logres on 2013-03-05 at 12:45 said:}

Seconded -- I believe there is not much to add, and the meditation may take a lifetime. The Trinity is partially ordered after caste, with the Father as the archetype, or Person without which Godhood does not exist, the Son as the executor or warrior (this is more clear in Revelation), and the Spirit as the silent partner who upholds everything. A case could be made for switching Spirit and Son, depending on how you view it. If God has caste-like orderings, what hope do mortals have to escape from it?

This was a beautiful translation.

\hfill

\texttt{Pickman on 2013-03-05 at 18:07 said:}

Ever since reading that brief introduction to De Giorgio in Evolas' ``Road of Cinnabar'' (under ``My Explorations of Origin and Tradition'') I have hungered for De Giorgio translations of ``outraged and aggressive letters, strewn with brilliant insights''. My praise extends to Cologero for his selfless task in making these works a possibility.

I get that sense of magic once more when reading this mans work, whose absolutist spirit is most congenial to my own character.

\hfill

\texttt{David on 2013-03-05 at 18:42 said:}

Di Giorgio is a genius, and those translation are a gift from the Holy Spirit. Thank you.

\hfill

\texttt{Jason-Adam on 2013-03-06 at 01:04 said:}

Personally, I find little reason to comment on De Giorgio's writings because they are so full of wisdom I have nothing more to add than to just nod my head and learn from him.

\hfill

\texttt{Max on 2016-04-01 at 07:14 said:}

Gandalf in a powerful scene that was not included in the final movie adaption of ``The Return of the King'':

``A thousand years this city has stood. Now at the whim of a madman it will fall. And the White Tree, the tree of the king, will never bloom again.''

``Why are they still guarding it?''

``They guard it because they have hope. A faint and fading hope that one day it will flower. That a king will come and this city will be as it once was before it fell into decay. The old wisdom borne out of the West was forsaken. Kings made tombs more splendid than the houses of the living and counted the old names of their descent dearer than the names of their sons. Childless lords sat in aged halls musing on heraldry or in high, cold towers asking questions of the stars. And so the people of Gondor fell into ruin. The line of kings failed. The White Tree withered. The rule of Gondor was given over to lesser men.''

In Tolkien's world the race of Men represents the warrior caste, and the Elves the priestly caste. As the end of the world approaches, many men become corrupted by their own mortality which would have otherwise been a source of great heroism and glory, and the Elves do not wish to take action but prefers to withdraw and isolate themselves contemplating what is eternal as if no external events concerns them. The Elves come to believe that the worlds destiny is inevitable and that resisting it is pointless, however man at the brink of destruction, by remembering the meaning of death as a gift of men, sees that there is nothing to loose and everything to win, with the fate of men's souls unknown even to the wisest. There are no shortages of battles, and even the ``living dead'' -- believing in nothing -- out of the despair of what they have become, are bound to answer the call of the Emperor where he to return. A role of the still remaining Elves includes awakening the King to his quest and to help provide the sign and means by which to conquer, reforging the broken sword that they have safe kept throughout the ages.

\hfill

\texttt{Aryayana on 2016-04-07 at 15:19 said:}

I think there are some good scenes dealing with the relationship between the First (spiritual elite) and Second (warrior nobility) castes in the Japanese movie `Zen' about Master Dogen. See, for example, the part starting from 1:28:24 and lasting until Dogen returns to his temple.

\url{https://youtu.be/8s5Uz8XSc-Y}


Notice how the Zen master's non-acting spiritual power prevails over the Shogun. The warrior-leader had the power even to kill the master if he so desired, but in the end it is the latter who wields the purest authority and who, indeed, is the true master of the former in terms of the essential hierarchy, and who may save him when he goes astray.

\hfill

\texttt{J on 2018-05-11 at 20:09 said:}

What would a Warrior in today's world do, given that there's (seemingly) no legitimate path for them to fulfill their destiny, and they can't just withdraw from the world like a Priest? Asking for a friend.

\hfill

\texttt{Avery Morrow on 2013-03-07 at 11:12 said:}

`` The brutality, although circumscribed and delimited, of the ancient wars, it should be, if not permanent at least frequent... ''

Guido De Giorgio seems to be gunning for Evola's role as ``greatest comic book supervillain of all time''. Too bad he looks more like Richard Feynman in the few photos available of him online.

\hfill

\texttt{Cologero on 2013-03-07 at 13:03 said:}

Yes, and it may discourage some warriors if they think they could be called out on a duel at any moment. Do you think his justification is incorrect? (``giving that acceleration necessary to the rhythm of life that makes it deeper and more fertile for revelation'') In other words, to live without the thought or awareness of death: does that also hide the rhythm of life and diminish the possibility of revelation? He is getting at something ... he may be right or wrong.

Does it relate to this:
\textit{Noble Savages}\footnote{\url{http://www.nytimes.com/2013/02/19/science/napoleon-chagnons-war-stories-in-the-amazon-and-at-home.html}}.

\hfill

\texttt{graham on 2013-03-08 at 14:45 said:}

As a woodworking apprentice nowadays, I'm looking forward to the part on workers and forms, it may be helpful in the task of discernment.

\hfill

\texttt{Logres on 2013-03-08 at 21:58 said:}

There is an interesting tale, here, about ``Deadwood Lighter'', a Cherokee medicine man who used dream magic to win against a deadly raid from a neighboring tribe. Would this be an instance of magic employed by a warrior culture (even though these tribes had ``degenerated'')?\\
\url{http://en.wikipedia.org/wiki/Townsend,_Tennessee}


\hfill

\texttt{Avery Morrow on 2013-03-08 at 23:58 said:}

Of course, the crux of his argument is that modern, conscripted life gives a different meaning to war than traditional gentleman's duel, and it's hard to argue. But to use the concept of \emph{war} itself instead of valor and a readiness to duel more generally, and then putting it in the imperative form, can't help but arouse amusement.

\hfill

\texttt{Cologero on 2013-03-09 at 01:23 said:}

The final segment will make things clearer, I presume. De Giorgio is not promoting war for its own sake, unlike others.

\hfill

\texttt{Caleb Cooper on 2013-03-09 at 11:25 said:}

I tend to favor an expansive use of the term magic (I perturbed my RCIA deacon by insisting that the Mass was a magic ritual and the Eucharist Blood Magic), but this sounds more like the guardian Archangel of the community initiating communication with it's representative before heaven. Magic to me seems to imply some intent or desire in the operant setting subtle forces in motion.

\hfill

\texttt{Jason-Adam on 2013-03-10 at 21:00 said:}

I am interested to know how de Giorgio would have explained the conflicts between the Popes and the Emperors in the Middle Ages... ... ... .

also, if the spiritual and temporal realms are autonomous of each other, does that mean the two swords (spiritual and temporal power) that Boniface VIII claimed the church had is a wrong idea ?

\hfill

\texttt{Cologero on 2013-03-10 at 21:58 said:}

Didn't De Giorgio explain the conflicts in this segment? He certainly alluded to them. The causes may be the weakness of the Priests who no longer are the bearers of sacred knowledge, or occasionally individualistic or egoistic factors.

De Giorgio is not refuting Boniface VIII. He emphasizes that the Warriors are subordinated to and consecrated by the priests; that is how they derive their power. The doctrine of the two swords does not mean that the priests are actively engaged in temporal affairs, just the opposite. In \emph{Unam Sanctam}, Boniface VIII wrote:

\begin{quotex}
Both, therefore, are in the power of the Church, that is to say, the spiritual and the material sword, but the former is to be administered for the Church but the latter by the Church; the former in the hands of the priest; the \textbf{latter by the hands of kings and soldiers, but at the will and sufferance of the priest}. ... one sword ought to be subordinated to the other and temporal authority, subjected to spiritual power
\end{quotex}


So the temporal power is indirect, not directly controlled by the spiritual authority. But bring this up in the chapter on the Capo.

The question is does Vatican II reject Unam Sanctam, which it certainly seems to have done.

\hfill

\texttt{Jason-Adam on 2013-03-10 at 23:00 said:}

Ah, you've cleared some things up -- so far De Giorgio's thoughts are in quite exact agreement with the teachings of the Church. I eagerly await the chapter on the Capo as I think it's the most important one in the book.

I am currently trying to understand the decline of Christendom -- we all know the events of history -- but I am more interested in the whys... ... so essentially the decline began in the Church, I suppose de Giorgio means, which is why Boniface was not able to control Phillipe ? but what caused the qulaity of the priests to descend from their former level ?

\hfill

\texttt{Cologero on 2013-03-10 at 23:32 said:}

De Giorgio gives his interesting take on it in one very telling paragraph (beginning with ``The elimination of the purely individual factors ... ''). Certain personalities (e.g. Phillipe) will bring in individual factors. He ends with a cryptic claim, if anyone has any thoughts on it.

To wit, the priestly class may choose to defect, since the priests should always be able to control the temporal power through spiritual means. Obviously, actions in time cannot affect the spiritual (``gates of hell''), which can also be found in Guenon. That is why the latter objects to Evola's raising of action to the same level as contemplation. So the priests may ``withdraw from the world and allow it to deteriorate''. De Giorgio concludes with this:

``Reflect very attentively on that in order to understand what is happening in Europe and in the world.''

How do you deal with that, Jason-Adam? Is the Church self-destructing as the secular media believes, or is there a deliberate withdrawal? Didn't Benedict XVI says there will be a contraction in the church? And De Giorgio mentioned the ``apparent collapse''; is that true or is the collapse real? And who are those ``solitary ascetics''?

\hfill

\texttt{Jason-Adam on 2013-03-11 at 00:06 said:}

If I understand him correctly, De Giorgio is saying that the spiritual centre has not degenerated but is voluntarily withdrawn from the world allowing the tides of history to take their courses ?

The solitary ascetics are the priests... ... whose true nature is hidden... ... ..

\hfill

\texttt{Cologero on 2013-03-11 at 00:59 said:}

That is the way I read him. Of course, many structures have indeed degenerated, but no material action can destroy the spiritual center. It may be forgotten by many, but it can never be lost. Some day I'll have to tell the story of the book's publication. The manuscript was found several years after his death. The original editor had to use a pseudonym, although the book seems pretty tame today.

\hfill

\texttt{Synodius on 2013-03-11 at 06:45 said:}

Interesting, I just finished reading Rama Coomaraswamy's Destruction of the Christian Tradition where he mentions a quote of Paul VI about `auto-demotition' and the whole book favors the version of destruction. Also it seems they are not only dismantling the Church but actively pushing for NWO. Doesn't look `apparent' to me, but what do I know? Can you please provide some context to what Benedict said about contraction?

\hfill

\texttt{Saladin on 2013-03-13 at 22:39 said:}

The Spiritual Center(s) never degenerate. It is only castes (and before all others the Priestly caste) that degenerate. There is little doubt that the priestly caste in Christendom has declined to the point of non-existence. Therefore it is only the solitary ascetics who carry the flame and enlighten humanity (albeit in a subliminal manner) in this darkest of ages.

And, no! The solitary ascetics are not priests, although priests can be solitary ascetics. Guenon, De Giorgio, ... are fine examples of solitary ascetics but there are many more whom we have never heard of!

\hfill

\texttt{scardanelli on 2013-03-14 at 13:42 said:}

This seems also to relate to the post on Evola and the intitiatic centers of Tibet falling to the communists. The church retracts materially in the face of the modern world just as the Tibetan Lamas would not act to safeguard sacred centers against communist invasion. Yet perhaps what seems like a loss only appears so from the perspective of action, for `` t he spiritual authority asserts itself with quite other means than those with which the temporal power asserts itself.''

Perhaps this retraction is what leaves so many dumbfounded and bemoaning the lack of a spiritual path in the west. They are looking without, rather than within. As De Giorgio states, the bridges to the superhuman may be crumbling, but the way remains open for those with the eyes to see.

So what looks like a retreat is actually a safeguarding, an attempt to ``maintain contact with traditional principles and undergo whatever duress, submit to any violation, in order to continue to save men by getting them back in touch with the eternal.''

If I am not misunderstanding Tomberg in this regard, would this be a retraction of the church of St. Peter in order to safeguard the church of St. John?

\hfill

\texttt{Max on 2015-02-16 at 15:35 said:}

De Giorgio states that what provides meaning and legitimacy to war is peace. If the situation is such as that can not be achieved the only viable option is withdrawal to prepare and await the conditions that allow victory since we can not aim for anything other than the Greatest Peace. The Romans never fought any battles if the signs were not auspicious -corresponding to the will of the gods. Continuing the thought of ``the roman'': when is the tiger tired? The Augurs silently gazed into the heavens.

When anti-reality fought reality it became increasingly unreal while reality stayed the same. The separation widened until un-reality could no longer recognize what is real or not, starting to fight its own shadow. For unreality every sky became cloudy, but Reality will make itself known like a lightning from the eternal clear sky.

\end{sffamily}
\end{footnotesize}

