\section{Myths, Symbols, and Lies}

Coincidentally, in today's email stream a rabid denunciation of Plato's \emph{Republic} appeared on a Thomism list. The writer attributed to Plato what he regarded as the tyranny of the modern world's political system.  He conflated the self-appointed political and economic elite with the guardians, with the mass media as the tool of mind control.

Of course, what he is describing is not the ideal state, but rather the final processes of the state's decline from democracy to tyranny, just as Plato predicted. Actually, Socrates was describing a much different system, rule by the best, or aristocracy. In the Republic, the guardians of the state were educated to overcome all personal self-interest. Through training in mathematics, gymnastics, and dialectics, they eventually would come to know the truth of the good.

This, then, is the first crucial question: Is it possible to know the truth of the good? If it is possible, then those who know would have the right and duty to rule — not for their own benefit, but rather for the common good of the city. The denial of that possibility is simply nihilism: the belief that no one can know the truth or the good. This is the actual position of the modern mind. Unwilling to let the best rule, the modern man simply assumes he knows the truth of the good, even without having undergone the arduous education like the guardians. For him, the good is ``what is good for me" and who but himself would know that?

The follow-up question is, ``how is it possible for those who know to actually rule?" For this, Socrates' response is simply Fortuna, or dumb luck. In the case of the Western world, \textbf{Rene Guenon} was somewhat pessimistic. He proposed that a spiritual elite would need to arise in order to lead back to a traditional civilization, although he thought that possibility rather unlikely.

\textbf{Boris Mouravieff}, on the other hand, was more sanguine about that possibility. Relying on his interpretation of Christian revelation (following a particular thread of Russian religious thinking), he presumed that the era of the Holy Spirit would lead to the rule of an elite of virtuous men and women. He points out that the Knight was accepted as the ruler in the Middle Ages because of his physical prowess. In the modern age, the men of science, technology, and commerce became the new elite, due to their superior rational knowledge. In the age to come, the spiritual elite will likewise be recognized. He summarizes:

\begin{quotex}
In the Cycle to come, the new characteristic of the elite man will be his ability to distinguish, spontaneously, without warrant or demonstrating proof, the true from the false. Such a man might also be endowed with the spiritual gifts of which St Paul speaks. It is obvious that those who add faculties of this nature to existing culture will automatically reach the head of human society. Their commanding authority will be accepted, in the same way that the authority of the intellectual was accepted when he substituted himself for the knight, and for the same reason: an obvious superiority. 

\end{quotex}
This view will be criticized as ``elitism" or ``Gnosticism" by those who don't understand. The two alternatives are clearly impossible, to wit, either democratism (knowledge of the truth of the good is available to everyone), or nihilism (such knowledge is not possible).

\paragraph{The Holy City}
The philosopher, the scientist, and the theologian do not create anything, but rather study what already is. The scientist studies nature, or the not-I, the theologian studies revelation, and the philosopher contemplates what is. Hence, they cannot lead us to the ultimate answer.

So while the philosopher contemplates the ideal city in thought, that is not the self-understanding of any actual city. In \emph{The City and Man}, \textbf{Leo Straus} concludes with praise for \textbf{Fustel de Coulanges'} \emph{Ancient City}. Readers will note that that book was also highly praised by Julius Evola. Strauss calls that city the ``Holy City", and he refers to the ``dark side" of the city, i.e., the side that is resistant to philosophical and scientific penetration. The Holy City was founded by a divine or semi-divine figure and its life revolved around religious rites, veneration of ancestors, and so on. None of this fits into the rational or scientific view of things.

In an essay on the myths of the origin or Rome, \textbf{Julius Evola}, relying on Bachofen, says that myth ``is a positive testimony to the spirit of a people." Spirit makes the city, not rationalism nor the material circumstances. Hence, his method is to understood the birth of the state through its myths and symbols, as well as the people who founded it. This is one of the sources of his ideas about races of the soul and spirit as more important than the race of the body.

The twin founders of Rome, Romulus and Remus, were the semi-divine offspring of Mars and the Vestal Virgin Rhea Silvia. At birth, they were left to die in the Tiber River. Since this is a common mythological theme (among them, Moses), Evola explains the symbolism of water this way:

\begin{quotex}
the waters, traditionally, have always depicted the current of time, i.e., the basic element of mortal, unstable, contingent, passionate, fleeting life. Taken from the waters and transported from the waters is the weak man. Saved from the waters, or capable of standing on the waters, of not sinking in the waters—is the seer or hero, the ascetic or the prophet. In the myth of the origins of Rome this symbol must therefore characterize again the ``divine" element of the founders of Rome, their, so to speak, supernatural dignity. 

\end{quotex}
So it is not the philosopher, and \emph{a fortiori} not the technocrat or scientist, who founds the city, but rather the hero, the seer, the ascetic, the prophet. These can be related to Evola's three stages of development.

\begin{enumerate}
\item The ordinary man is taken up in the flow of the waters of the world of becoming. He lives spontaneously, not consciously. Mouravieff uses the image of the ``confluence" of A, or worldly, profane influences. 
\item Some men rise above this and study their life as the other. They are the scientists, philosophers, and so on. 
\item This is the stage of the awareness of the I. Such men have mastery, or will. They are the heroes and rishis. 
\end{enumerate}
The next sections can be regarded as appendices. They are extended quotes on the prevalence, or even necessity, of lies in the world, and the meaning of symbol and myth.

\paragraph{World Ruled by Lies}
In Volume I of \emph{Gnosis}, Boris Mouravieff makes this remarkable claim.

\begin{quotex}
We live in a world ruled by lies. Lying and stealing are the dominant elements of human character whatever the race, creed or caste. Whoever says that this is not true simply tells another lie. Man lies because in a world ruled by lies it is not possible for him to do otherwise. To all that has already been said, one must add the following peculiarity which at first looks paradoxical; that the progress of this civilization, which is the fruit of an intellectual culture, considerably increases the need for lying.

If we refer to the different aspects of this subject, an analysis of lying permits us to distinguish the following modes:

\end{quotex}
\begin{itemize}
\item \begin{quotex}
Lying to others 
\end{quotex}
\item \begin{quotex}
Lying to oneself 
\end{quotex}
\item \begin{quotex}
Useful lies 
\end{quotex}
\item \begin{quotex}
Useless lies 
\end{quotex}
\end{itemize}
\begin{quotex}
To these classic cases of lying, one must add two particular cases:

\end{quotex}
\begin{itemize}
\item \begin{quotex}
Hypocrisy: the pretense of virtue, of praiseworthy sentiments, with the intent to deceive persons of good faith 
\end{quotex}
\item \begin{quotex}
The integral lie: this characterizes that person who, from a habit of lying and cheating on every occasion 
\end{quotex}
\end{itemize}
\paragraph{Symbol and Myth}
The following passages are from \emph{Myth, Religion, \& Mother Right} by \textbf{J J Bachofen}.

\begin{quotex}
Myths became images and shadows of higher ideas, and by their mysterious character inculcated a profounder veneration. \flright{\textsc{Plutarch}}

Myth is the exegesis of the symbol. It unfolds in a series of outwardly connected actions what the symbol embodies in a unity.

The higher meditations inspired by the riddle of death, the grief and consolation, the hope and fear, the foreboding and joyful anticipation, are expressed only in art. There is a profound reason for this. Human language is too feeble to convey all the thoughts aroused by the alternation of life and death and the sublime hopes of the initiate. Only the symbol and the related myth can meet this higher need.

The symbol plucks all the strings of the human spirit at once; speech is compelled to take up a single thought at a time. The symbol strikes its roots in the most secret depths of the soul; language skims over the surface of the understanding like a soft breeze. They symbol aims inward; language outward. Only the symbol can combine the most disparate elements into a unitary impression.

Words make the infinite finite, symbols carry the spirit beyond the finite world of becoming into the realm of infinite being. Intimating the ineffable, they are mysterious as all religion by its very nature must be, a silent discourse appropriate to the quiet of the tomb, beyond the read of mockery and doubt, those unripe fruits of wisdom. Therein lies the mysterious dignity of the symbol … therein lies the spell of the mythical representations, which show us the great deeds of the primordial age in the muted light of distant melancholy recollection 

\end{quotex}


\flrightit{Posted on 2014-12-04 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{andros on 2014-12-05 at 04:59 said: }

`The integral lie: this characterizes that person who, from a habit of lying and cheating on every occasion,' … is completed by … `ends by believing his own lies and thus loses all sense of truth.' (Chapter XVII, 2)


\hfill

\texttt{Max on 2014-12-07 at 06:36 said: }

For a state to be organic it must naturally be formed in an organic way. That does not involve thinking about ideal states -because then what. Asking people to come to the new state that does not even exist? It is not possible to construct something organic because then it just ends up as another machine, not a living being.

When the conditions are right, a child is conceived and born. The real causes for a new life coming into existence is nevertheless transcendent to the physical act. Thinking about ideal babies will not make any, neither do good parents think of children in those terms. Rather, being in itself is proof of perfection. We love that which is for what it is because it is and does our best to help it realize its true potential.

An organic state is also born and unfolds its truth over time. It is not created by man as a ready-made thought-construct that just awaits implementation. 

The Holy City is first of all in higher worlds. The spiritual origin of an organic state becomes concealed and explained as myth, understood according to each ones capabilities. A living organic state is superior to any thoughts (regardless of how many worldly faults could be found with it) just by the fact of being here, as is Tomberg's view of the church. 

Churches are based on the basilica, which was a Roman courthouse and before that a ``royal hall": the original meaning of the Greek word. This is so because a church is the royal hall of the King of Kings, where divine law and justice radiates out to the lands and people. At earlier times the basilica was also a place of economic activity and human law but that was put an end to with the advent of Christianity in an effort to separate the sacred from profane.


\hfill

\texttt{Michael on 2019-12-04 at 17:34 said: }

Interesting that lies and myth seem to vibrate along similar senses in experience. However, lies we create and myths speak to something more and are not created by ``I", as the quote makes out. When developing awareness in moments of lies or when spoken to, it seems like there is an experience when being lied to, a sort of ``sense" that even if we aren't sure is there resounds in our experience. Perhaps this sense can be developed in such a way to know false from truth, allowing one to further appreciate the depths of that connection when symbols are present and understand lies when they are made.


\end{sffamily}\end{footnotesize}
