\section{Europe in Dormition}

Dormition is a state that is neither life nor death. The choice must be made ... will it be proactive and creative or passive and reactive?

\paragraph{Venner, Maurras and Jünger}

I would like to translate some of \textbf{Dominique Venner}'s blog posts, except his literary heirs promise to go after ``pirates'' of his works. Instead, I will review an interview published in November 2011 under the title ``Europe in Dormition\footnote{\url{http://www.dominiquevenner.fr/2011/12/entretien-avec-louis-montarnal-publie-dans-laction-francaise-2000-n-2827-du-3-au-16-novembre-2011/}}''. Dormition indicates a state that is neither dead nor alive. It referred to \textbf{Mary}'s ``falling asleep'', since she could not die. Later, \textbf{King Arthur} was said to be in dormition in \textbf{Avalon}.

When asked, Venner claims that Europe fell asleep after the two world wars. The result was that world power was transferred from Europe to the USA and USSR. More than power, Europe lost faith in itself and was eaten away by a moral crisis and feelings of guilt. There is no connection, however, between the loss of power and the moral crisis, and the subsequent ``dormition''. Europe's wounds were self-inflicted and the dormition, which he fails to see, was due to the destruction of the last remnants of the Holy Roman Empire. In the time period from Charles the Great to Charles I, there was no moral crisis, there was no guilt, and Europe was wide awake.

Nevertheless, at least in 2011 Venner was hopeful. He pointed to the resurgence of civilizations long thought dead, such as China, India, South America, or Islam. The same could happen to Europe. He claimed that massive immigration could kill Europe. On the other hand, the culture shock could perhaps reawaken Europeans to their true identity (whatever it is). This is happening now, ``behind the scenes'', as it were, in popular culture although not yet in the political sphere.

Venner went on to discuss two of his major influences: \textbf{Charles Maurras} and \textbf{Ernst Jünger}. He admired Maurras' hard-nosed character and his courage in the face of ordeals. After his trip to Athens in 1898, Maurras returned a convinced pagan, an inspiring thought to Venner. Venner looked to Homer as the foundation of European civilization: nature as the base, excellence as the principle, and beauty as the horizon.

Those values he received from Maurras who, however, was never confused about the true French identity which was created by the forty Catholic kings of France. While the foundation is important, the structure arising from it is what counts.

Jünger was authenticated by his life, in Venner's eyes. Venner was moved by Jünger's high spirituality formed in the forests and nature. He quotes St. Bernard:

\begin{quotex}
You will find more in the forests than in books. The trees will teach you things that no master will tell you.
\end{quotex}

That is the spirituality of Jünger's French and Gaulish ancestors, which is what Venner terms ``tradition'', which ``develops in us without our knowing it.''

Venner never mentioned that both Maurras and Jünger converted to Catholicism before their death. It is hard to believe that, in such thoughtful men, the conversion was motivated by weakness or ignorance. Most likely, it is the correct conclusion of old men who have known nature, excellence, and beauty, yet were still seeking for their fulfillment. Ironically, in Venner's words, in those men tradition slowly developed within them. Venner makes the choice starkly clear: conversion or suicide. Mock me if you will, I am just trying to save you many years of lost time, if not suicide.

\paragraph{His Favorite Ideas}

Charles Maurras, wrote the following as a pagan. I do not repeat this to convert any pagans, but rather to set the standard that they will need to surpass.

\begin{quotex}
All my favorite ideas---order, tradition, discipline, hierarchy, authority, continuity, unity, work, family, corporation, decentralization, autonomy, organization of workers---had been preserved and perfected by Catholicism.
\end{quotex}

\paragraph{The Moral Conception of the World}

The intellectual contemplation of nature, excellence, and beauty, as taught to us by the Greeks, is our foundation. However, European civilization advanced because it was \emph{creative}, something constructed, not merely contemplated. Hence, to awake from ``dormition'' is to once again begin to create. In the conclusion of \emph{The Theory of Mind as Pure Act}, \textbf{Giovanni Gentile} explains:

\begin{quotationx}
Idealism is anti-intellectualistic and, in this sense, profoundly Christian, if we take Christianity as meaning the intrinsically moral conception of the world. This moral conception is one which is entirely alien to India and to Greece even in their greatest speculative efforts. The philosophy of India ends in asceticism, in the suppression of the passions, in the extirpation of desire and every root of the human incentive to work, in nirvana. Its ideal, therefore, is the simple negation of the real in which morality realizes itself, human personality.

And in Greek philosophy the highest ethical word it can pronounce is Justice. Justice renders to each his own and therefore preserves the natural order, but it can neither create nor construct a new world. Greek philosophy, therefore, cannot express the essential virtue of mind which is its creative nature, it must produce the good which it cannot find confronting. How could Greek philosophy understand the moral nature of mind seeing that its world was not mind but nature?

Nature need not be material, it might be ideal but it is what the mind contemplates, not what it makes. \textbf{Greek morality ends in the Stoical doctrine of suicide}, a doctrine consistent with its immanent tendency to an intellectualistic conception of a reality in which the subject has no worth. Christianity, on the other hand, discovers the reality which is not until it creates itself, and is what it \emph{creates}. It cannot be treated like the Greek philosopher's world, already in existence and waiting to be known til the philosopher is ready to contemplate it, when he has drawn aside as it were, when, as Aristotle would say, all the wants of his life are appeased and life is as it were complete.

It is a reality which waits for us to construct, a reality which is truly even now love and will, because it is the inward effort of the soul, its living process, not its ideal and external model.

It is man himself who rises above humanity and becomes God. And even God is no longer a reality who already is, but the God who is begotten in us and is ourselves insofar as we with our whole being rise to Him. Here mind is no longer intellect but will. The world is no longer what is known but what is made: and therefore not only can we begin to conceive the mind as freedom or moral activity, but the world, the whole world of the Christian, is freed and redeemed.
\end{quotationx}


\flrightit{Posted on 2014-01-09 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}

\texttt{Michel on 2014-01-09 at 16:21 said:}

``This is happening now, ``behind the scenes'', as it were, in popular culture although not yet in the political sphere.''

Really? You mean that the public is undergoing some sort of re-awakening to true European identity? I thought all that new-age, new-atheism,new-right stuff was just bardo, the sort of shimmering glimmers of wandering influences from different epochs, slowly fading away, flicking them here and there, or that it was the result of suggestion from the counter-initiation in prepration for their establishment of the parody of genuine Spirituality.

Or perhaps being more `positive' it was a sort of minor pre-figuration of the new Golden Age. After all, the sun will rise from the West (though with globalization, what does `west' mean? Is it really Europe, or is it possible that it just indicates the furthest extent of exhaustion in terms of possibility, from whence, due to the intense rapidity of this development, everything will suddenly be `suspended' and become motionless, before the wheel starts turning again?) .

\hfill

\texttt{Cologero on 2014-01-09 at 18:06 said:}

At least, Michel, that is what Venner believes. Perhaps this short piece by an American philosopher in Paris confirms it: \textit{Does Immigration Mean `France Is Over'?}\footnote{\url{http://opinionator.blogs.nytimes.com/2014/01/05/does-immigration-mean-france-is-over/?\_r=0}} It is indeed curious that the American is disappointed by the remnants of French identity.

\hfill

\texttt{Ash on 2014-01-09 at 21:30 said:}

The Stoic doctrine of suicide can be summed as being that suicide is permissible when it is impossible to live virtuously. Cato would be an example, as he took his life when he felt that it was no longer possible to serve Rome by continuing his war or reconciling with Caesar. Refusing to return would be dishonourable. In Gentile's terms, what would the Christian do in such a case? I ask because this is why Venner chose the path he did, as I understand his last writings. Whether he was correct in this is beside the point; if he believed that there was no more honour to be gained through living, what should he have done? Turn his focus inward through Catholic practice and seek to change reality that way instead of the non-cooperation which his suicide effectively was?

As an aside, this post is quite prescient with the continuing appearance of cracks between the neopagan elements of the New Right who interpret Tradition in Venner's terms and those who acknowledge the Tradition which we discuss here. I think Cologero once mentioned the possibility of Pagans who knew and accepted the mission of Christianity in the world, but I've yet to see many.

\hfill

\texttt{Constantine Aetos on 2014-01-10 at 02:22 said:}

``The intellectual contemplation of nature, excellence, and beauty, as taught to us by the Greeks, is our foundation. However, European civilization advanced because it was creative, something constructed, not merely contemplated.''

Would the reconstruction of Europe be done by man only or aided by God?

\hfill

\texttt{mohenko on 2014-01-10 at 13:52 said:}

What an inspirational piece, thank you.

\hfill

\texttt{Cologero on 2014-01-10 at 15:40 said:}

``Would the reconstruction of Europe be done by man only or aided by God?''

How would you know the difference, Mr. Aetos?

\hfill

\texttt{Michel on 2014-01-10 at 15:43 said:}

I thank you for the link Cologero and for the post Inaki. Of course, such topics, when taken from an extremely constricted point of view lose any attraction they might have had to me. The one thing I see is that there seems to be an encroachment on the opposites of true values within a Traditional Society so that when taken within a inverted context, they become the foundations for a near-upside down society. There would be the need for some sort of tire-like structure that further stratifies or simply makes the already existing stratification of the biological races more apparent. For one, this awakening is not based on the thing that gave Europe, as a `collective unit' its identity, viz Catholicism, but rather from loosely jointed `paganism' (for example, Astaru or Asatru or whatever) in a chaotic demetrian guise. On the one hand, if a so called `spiritual dimension' is to be considered, the only thing that interests popular culture is interaction with as I said, and as you stated in one of your articles, bardo. On the other hand, in relation to something most people can relate to, as you also stated, this is further given weight by a loosely understood and external unifying factor, viz biological race, and as for example, indicated by the author of the post on Immigration and European `culture'.

The people that Venner sees/saw as `going back to their roots' are indifferent to anyone who happens to have a white or pink skin pigmentation (whereby white non-Frenchmen are treated as expatriates and black or brown people are seen as parasitic sub-human carriers of strange diseases). That's it. In the end, its just a collegiate of fascination with psychic elements that correspond to dead bodies and corpses in the gross world and with a stupid justification for this fascination based on skin color. This is hardly a re-awakening to true identity. If anything, certain counter-initiatic ambitions, such as the establishment of a superior biological race and the demise of those that they deem inferior comes to mind when such attitudes are encountered. For instance, in the Georgia Guidestones, one of the aims or `commandments' for the new world is the reduction of the current earth's population by about 92.8\%, so that only 500,000 people are left alive. Most probably, all of them white, or `Aryan' as Hitler would have loved. Moreso, certain murals at the Denver International Airport show the death of both black and brown races and the hailing of what appears to be a bravarian boy. This is indicated by the motionless corpse of a black girl and that of an native American Indian. The entire airport is also designed to look like a giant swastika, interpreted in the Nazi sense. So in the end, I'm led to believe that this `nostalgia' is just bardo, or the reaction to suggestion by the counter-initiatic forces, as they slowly continue to reveal themselves and popularize their values.

At any rate, as is obvious, a true resurgence must be shown via the existence of genuine Dcotrinal Priniples, upon which disemminated practice would be based. These Doctrinal Principles would not be based on `convention' neither would they be based on discursive analysis, even of the highest kind, they would rather be based on Transcendence, and would be rooted within the essential way that things are fashioned in accordance to their Divine Principles, asserting their Transcendent Unity (beyond even gods or goddesses, even though they correspond to supra-individual states of consciousness). Thus any theories or attitudes that take as their foundation elements based on duality (and the lowest kinds in this case, disentangled psychic influences and gross physical qualities) must be either out of genuine stupidity, or worse, out of reaction to satanic forces. Don't you at least see that while Genuine Tradition relays its Doctrine from the Top down, this one tries to mimic the same in a near inverted sense? That ultimately, the things that define the manner of being of creatures, their dharmma or predisposition are the things that are most external to them, viz gross characteristics? And that it is these things that should be considered when trying to `rediscover identity' ?

\hfill

\texttt{Michel on 2014-01-10 at 15:48 said:}

sorry, I meant 500 million people, not 500,000 people.

\hfill

\texttt{August on 2014-01-10 at 20:03 said:}

``Thus any theories or attitudes that take as their foundation elements based on duality (and the lowest kinds in this case, disentangled psychic influences and gross physical qualities) must be either out of genuine stupidity, or worse, out of reaction to satanic forces.''

Or, a third possibility, attempted resistance to dissolution of the remaining psychic elements that still persist, consciously, in these people.

\hfill

\texttt{Michel on 2014-01-11 at 10:26 said:}

Yes August, that could also be the case.

\hfill

\texttt{Constantine Aetos on 2014-01-11 at 16:05 said:}

@Cologero

Look at it Abu Dhabi and look at Isfahan, Iran. I highly doubt Emir Zayed was inspired by Allah to build skyscrapers.

\hfill

\texttt{Cologero on 2014-01-12 at 23:56 said:}

Although I have never visited that part of the World, I see what you mean from various images.

So Constantine, you seem to know God's will, so what more do you expect? The rest, then, depends on man and theosis, i.e., aligning his will with God's. I'm afraid there are many distractions along the way, but it should not be impossible, even in something basic such as the architecture you point out.

\hfill

\texttt{Logres on 2014-01-13 at 15:06 said:}

Michel, I'm glad someone else finds the Denver airport murals disturbing (other than conspiracy theorists, who are still closer than most to the truth). It's quite possible that liberalism as a whole is being used to foster a new tripartite society of ethnic status and color, although I think they may not have made up their mind if they want it to be multi-cult, or just white at the top. My guess is that the real hard core elements are mostly white ``supers'' in the counter-initiation. That is, they are probably interested in things like ``genetic'' weapons. That said, they are willing to let others in, at a price. Remember, it's just like the real thing, except evil. I think white people in America could do worse than stop worrying about being ``white'', and start learning Gaelic, German, Dutch, or whatever their ancestors real language was. If we are to be movement of quality, it doesn't make sense to buy into generic whiteness -- that way lies powder-puff store-bought bread like Ideal. To speak personally, I'd rather acknowledge, for instance that I myself am a hybrid of German-Celt, and try to discern clearly the elements of my own soul, including learning those languages, or at least a lullaby or two, remembering my ancestors, and recapturing what is good and can be remembered from the past. The rest will take care of itself, if I work very hard. My wife has a little bit of Indian heritage (many Southerners do), and there's something of value there, as well. Someone near here just took my kids and I on a falcon hunt with Harris hawks. It was just squirrels we were after this time of year, but such activities help me stay out of the always present attraction to far-right populist movements, which will no doubt have their day. Uranus and Saturn are conjuncting (I am told) around 2030, and this is supposedly favorable to right wing movements. In the end, it's up to the ``I'' to change the world, not a mass movement. ``The poor, you will always have with you'' -- so too, the masses. In the USA, you will eventually see a fragmentation of regions, with the Southwest being Anglo-Hispanic, the Deep South continuing with an African influence, and the West Coast having a lot of settlers from Asia. I don't think they can hold it all together, without very dangerous repercussions. ``Light federalism'', with regional differences, might work for awhile maybe indefinitely.

\hfill

\texttt{Michel on 2014-01-15 at 12:54 said:}

Hello Mr. Logres, it's good that your children are given a chance to interact with the freshness of nature in these atrophied days. Anyways, about the airport, among other ``questionable'' monuments, we should note that the people behind them churn out billions of dollars for their construction. Billions. I doubt that, with that kind of money involved (at least from their point of view), they just randomly chose murals, or architectural shapes without giving some serious thought into it. If I can recall, I think that Denver International Airport, costed \$4.1-\$4.6 billion dollars. No one is joking with that kind of money and commitment. The apocalyptic horse at the entrance, the Hitler wearing a gas-mask, the various Freemasonic elements (used in a malefic way) all tell a story that is slowly unraveling, a story that these people take very seriously, other than the use of magical sigils for literal hypnotic processes, these inverted art forms are a way for ``those with the eyes to see''. On that note, I still wonder why other than Guenon and the re-introducers of Metaphysics, no one takes the existence of such groups seriously, nor the most simple implications of their actions on society.

\hfill

\end{sffamily}
\end{footnotesize}

