\section{Shadows and Reflections: The Lunar Race}

\textbf{Julius Evola} begins the topic of the Lunar Races with \textit{The Demetrian Race}\footnote{See Section \ref{sec:DemetrianRace} in this book.}. First, let's clarify Evola's method and its relationship to Mother Right by \textbf{J J Bachofen}.

Bachofen's book is a diachronic study of various cultures over time: their origin and evolution. Each one is guided by a different spiritual type; Bachofen is heavily dependent on the use of myth. Evola converts that to a synchronic study. That is, he recognizes each type as active and manifesting at the same time. The myths, then, describe the characteristics of various types of people, which Evola identifies as the ``races of the spirit''. As such, the specific historical and pre-historical manifestations of societies of a given type are not the major concern, except for purposes of illustration.

The Demetrian race does not recognize transcendent spirituality. Its focus is on the earth, matter. It lacks a spiritual center, so is prone to pantheism and universalism. Its concern is with the laws of nature, agricultural cycles, and cosmic harmony of the ``we are all one'' type. Since, for it, everything is ``natural'', it fails to recognize anything as deviant.

The Lunar Race is not self-effulgent. That is, it can only reflect the light from an external source. This is characteristic of the intellectual type who can only study or contemplate nature. This is also the Hermetic teaching. \textbf{Valentin Tomberg} identifies the 18th Arcanum, ``The Moon'' with ``materialistic intellectuality''. This then eclipses the ``creative light of consciousness'' for human intelligence.

In Arcanum XIX, The Sun, Tomberg goes on to say that the ``task of human intelligence is to liberate itself from the magical enchantment which separates it from spontaneous wisdom ... to arrive at
\emph{intuition}.'' This intuition is the spontaneous wisdom of the heart, as Evola also points out.

You can take this as a warning, if you like, against incessant chatter and infinite discussion. What is necessary is a higher realization, not to waste your time among the ``\textbf{shadows of ideas and things}''.


\flrightit{Posted on 2015-06-07 by Aeneas}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{aegishjalmer on 2015-06-08 at 03:15 said:}

Thank you, this is something I needed to hear

\end{sffamily}
\end{footnotesize}

