\section{The Country of the Blind}

\begin{quotex}
The time has long gone by when anyone who claims the title of philosopher can think of religion as a superfluity for the educated and an `opiate for the masses'. It is the only known explosive in the economy of that delicate internal-combustion engine, the human mind. Peoples rich in religious energy can overcome all obstacles and attain any height in the scale of civilization. Peoples that have reached the top of a hill by the wise use of religious energy may then decide to do without it; they can still move, but they can only move downhill, and when they come to the bottom of the hill they stop. \flright{\textsc{Robin Collingwood}}

\end{quotex}
R G Collingwood was the last of the British idealist philosophers (save a recent mini-revival). Although greatly influenced by the Italian idealists, including Giovanni Gentile, he found himself on the opposing side of World War II. That involves a conflict of the man as philosopher and as citizen. In the former role, he needs to develop an objective political science; while in the latter role, he needs to justify his own nation.

British idealism is genteelly liberal, while Italian idealism in general, not just in Gentile and Julius Evola, is nothing of the sort (apart perhaps from Benedetto Croce). Nevertheless, Collingwood does make some interesting points in political science. In particular, we will focus on the three laws of politics that he identified. (cf. \emph{Essays in Political Philosophy}.)

\paragraph{The Blind Leading the Blind}
The short story, \emph{The Country of the Blind}, by \textbf{H G Wells}, is about a lost South American tribe in which everyone is born blind. A mountaineer accidently discovers their city. He imagines that, with his sight, he would become their ruler. Unfortunately for him, the villagers do not believe in his ``alleged" sense of sight. They think he is crazy and his eyes are a cause of his disease. The consequences of this will be brought up at the end of this essay.

\paragraph{Civilization}
First of all, political science, as a human science, has a \emph{telos}, unlike the physical sciences. So if you are a warrior for civilization—whether by pen or sword–, it is necessary to define exactly what that entails. Collingwood says that civilization includes three conditions:

\begin{itemize}
\item \textbf{Law and order}. This is a condition of society in which men live according to definite rules, known to all, and enforced by the political power. 
\item \textbf{Prosperity}. This is a condition of society in which a man acquires the things he needs for sustenance and comfort not by taking them from others but by earning them. 
\item \textbf{Peace}. This is the condition of society in which the arbiter of human affairs is not violence but agreement. 
\end{itemize}
At this point, Collingwood describes WWII as a battle between civilization and its destruction by the Axis powers. Of course, there is more to be added to these brief summaries, but these three conditions may be useful to determine the state of a civilization. This is an exercise for the reader.

\paragraph{Society and Community}
A number of human beings living together constitute a \textbf{community}. A \textbf{society} is a kind of community with origins from Ancient Rome. Under Roman law, a society (\emph{societas}) is defined as a number of partners who must be adult, free, male, Roman citizens acting together in the pursuit of aims advantageous to all. That is, they are members through their own free will. As Evola points out, a king wants loyalty freely given by free subjects.

A community, on the other hand, lacks some of these features. In particular, actions are no longer always governed by freedom, but often by constraint. Many examples of such constraint can be found, from the subtle to the gross. Collingwood defines it thus:

\begin{quotex}
The essence of constraint is that a man should be brought into a condition of emotional disturbance in which he cannot make the free decision. The disturbance may be of many kinds; for example, it may be a state of fear. 

\end{quotex}
\paragraph{First Law of Politics}
The First Law of Politics is that \emph{in a community there are two classes: the rulers and the ruled}. Collingwood points out that \textbf{Abraham Lincoln}'s ideal of a ``government of the people, etc." did not include women, children or non-citizens, however it may be re-interpreted today.

\paragraph{Second Law of Politics}
The Second Law of Politics states that \emph{the division between rulers and ruled is permeable}. That is, there may be movement back and forth between the two divisions.

\paragraph{Third Law of Politics}
This is the more interesting law. It states that \emph{the characteristics displayed by the ruler will tend to be imitated by those over whom he rules}. Specifically, in a democracy the ruler will be chosen by the feeling of affinity rather than on any particular qualities of a candidate. The issues are too complex even for educated citizens. 

I grew up in a small town outside of Boston which still had town meetings. The issues that came up for votes, however, were along the lines of how many new snow plows should the town buy this year, not existential questions. That is, it was more like a society with a common aim.

However, to be an informed citizen requires knowledge of political philosophy, bioethics, economics, geopolitical strategy, and so on. That knowledge is rare, especially among the rulers. Rather, they focus more on the second law and spend their energy in maintaining or gaining power. The ruler reflects the soul qualities of a community as the images indicate.

\paragraph{Refrain}
By the third law, the ruler is like the ruled, although he excels in some quality. Harkening back to the Wells' story, this quality must be something that the community recognizes. Hence, since a democratic community is not composed of rational and intelligent men, they cannot recognize those qualities in a ruler. Rather, they gravitate toward qualities related to appearance and style; it is not what he says, but how he says it that matters. So we get unintelligence, of which there are two kinds.

\begin{itemize}
\item \textbf{Negative unintelligence}. A person in this frame of mind, trying to grasp something, grasps nothing. He comes away from all mental effort empty-handed. 
\item \textbf{Creative unintelligence}. Rather than ``nothing", this type of unintelligence creates fantasies and chimeras. It fabricates imaginary causes for things. This is more intractable since it is a simulation of real intelligence. 
\end{itemize}
Hence, there are two ways to be a \textbf{fool}.

\begin{itemize}
\item There is being \emph{foolish to stupidity} in which the mind grasps nothing. This may happen, for example, at a weekend seminar on transcendental idealism. 
\item Then there is being \emph{foolish to craziness}. In this case the mind creates illusions or hallucinations about what it is trying to understand. 
\end{itemize}
The stupid fool knows his weakness and may try to overcome it. Unfortunately, the crazy fool believes he is being intelligent, so he has no motive to change. The crazy fool has as much initiative over his fellow man as does the intelligent man. In some ways more, since he does not have the reticence of the intelligent man.

\paragraph{Postscript}
Collingwood believes he is describing the Axis powers as the fools. In his opinion, Britain, as the peak of Protestant Christianity, represents the peak of civilization. Germany, he claims, was the last European nation to embrace Christianity, so it is still pagan at its core. Even Roman Christianity in Italy owes much to ancient Roman paganism.

Britain won that battle, but may eventually lose the war; the verdict is still out. Britain is hardly a Christian nation anymore. If his understanding is correct, Great Britain is heading downhill, despite all appearances. If it eventually stops at the bottom, then Collingwood's reputation will return, but this time as a prophet, not as a philosopher.



\flrightit{Posted on 2015-03-25 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Mark Citadel on 2015-03-25 at 13:29 said: }

Having experienced the degeneration of a few countries, Britain being one of them, I can tell you that it may be one of the first to go into a period of total delf-destruction. The people cannot even stomach populist parties like UKIP because they're so `extreme'.

There is an adherence to a rival religion known as `British values' which are explicitly anti-Traditional and enshrined into the legal and education systems. It's not even as if there is a significant hold-out vestige of Tradition. The Kali Yuga picked the bones clean with this one.


\hfill

\texttt{Max on 2015-03-25 at 15:05 said: }

The current condition of ``civilization": laws are arbitrary, changing all the time, so numerous no one knows them and enforced selectively. ``Earning" something is most often synonymous with stealing it lawfully and unopposed. Violence and conflict are artificially engendered and engineered.

How to move in a positive direction.

1. Most seems incapable of either being ruled or to rule. The ordered man knows both and will rule himself.

2. An organic hierarchy is not static. A new nobility will be earned and self selected.

3. In Germanic languages the words for stillness, peace, freedom and friendship have the same root. A society will be formed by affinity.


\end{sffamily}\end{footnotesize}
