\section{Spiritual Reinvigoration}

\begin{quotex}
Across the centuries, it is the drowsiness of the disciples that opens up possibilities for the power of the Evil One. Such drowsiness deadens the soul, so that it remains undisturbed by the power of the Evil One at work in the world and by all the injustice and suffering ravaging the earth. In its state of numbness, the soul prefers not to see all this; it is easily persuaded that things cannot be so bad, so as to continue in the self-satisfaction of its own comfortable existence. Yet this deadening of souls, this lack of vigilance regarding both God's closeness and the looming forces of darkness, is what gives the Evil One power in the world.
\flright{\textsc{Joseph Ratzinger}, \emph{Jesus of Nazareth: Holy Week}}

The primary source of beings, he is also the principle of cohesion and harmony; he makes of the created world a COSMOS, an ordered universe, by giving it a meaning and a value, by orienting it toward a goal.
\flright{\textsc{Huby}}
\end{quotex}

Last weekend I listened to a long interview on a new right site criticizing ``Christianity''. There was nothing new in it, just the same repetitious, and mostly false, claims. On the other hand, I suppose he looks around and sees nothing compelling in the religion of his ancestors, and projects that back into time. Unfortunately, he has no understanding of metaphysics, so most of his objections are wide of the mark.

So I want to address some of his concerns at a deeper level. But first of all, let us be clear about the meaning of ``religion''. It is not first of all about hard-to-believe stories. In this regard, we rely on \textbf{Rene Guenon}'s description of the fundamental points of religion and tradition in order to evaluate the Traditional Christian religion. It can also serve as a set of guidelines for anyone trying to design a new religion on his kitchen table. Since most of these issues have been dealt with in other essays on this blog, I will only sketch them out here.

\paragraph{The Elements of Religion}

Guenon identifies three properties necessary for a religion. These are:

\begin{itemize}

\item Doctrine or dogma

\item Moral law

\item Cult or worship
\end{itemize}

\paragraph{Doctrine}

Dogma or doctrine represents the intellectual element of religion. Guenon has elaborated a fundamental metaphysic in several of his books. Hence, a true tradition is in conformance with that metaphysic. Now religious dogmas can be restated as metaphysical doctrine, but the opposite is not the case. That is because dogmas take form in specific historical situations.

For example, the dogma of monotheism is understood doctrinally as representing the primary principle of Being. Closer to our task, Ratzinger elucidates the Christological dogmas defined at Nicaea and Chalcedon by bringing out their deeper meaning in respect to our understanding of what it is to be a person, or the relationship between the human and divine will.

\paragraph{Moral Law}

A religion is based around a moral law which is its social element. If the created world is a ``cosmos'', it has an inherent order. Unlike a rock which has no choice about following the cosmic order, a person can deviate from it. The moral law guides us into being in conformance with that law. Somewhat paradoxically, this leads us to freedom, since the alleged ``freedom'' in acting against the law turns out to be bondage to lower forces.

Antinomianism is rife in Christian thinking today. ``Love'' is understood as the absence of moral law, since that is seen as ``judgmental'' or ``intolerant''. This sort of ``love'' is the love of servitude. As \textbf{Nikolai Berdyaev} points out:

\begin{quotex}
Man is in a state of servitude. He frequently does not notice he is a slave, and sometimes he loves it. It would be a mistake to think that the average man loves freedom. A still greater mistake would be to suppose that freedom is an easy thing.
\end{quotex}

That is the danger of antinomianism. As Jesus said, ``If you love me, you will keep my commandments.''

\paragraph{Cult and Worship}

The religious rites tie the intellectual and social elements together. All rites have a symbolic character, and so can also be understood in their metaphysical sense. Moreover, rites are also efficacious as conduits of a spiritual influence.

\paragraph{Exoterism and Esoterism}

Guenon points out that a tradition has an exoteric or outer aspect as well as an esoteric or inner aspect. Guenon writes:

\begin{quotex}
Exoterism, comprising the more elementary and easily understandable part of the teaching, which was consequently more readily brought within everybody's reach, is the only aspect to be expressed through the writings that have come down to us in a more or less complete form.

Esoterism, being more profound and of a higher order, addressed itself as such only to regular disciples of the school who were specially prepared to receive it, and was the subject of a purely oral teaching.
\end{quotex}

Nevertheless, they do not represent two different teachings and cannot be contradictory. Oftentimes, the esoteric understanding is misunderstood by the super-correct exoteric adherents, who consider a threat and a different religion. On the other hand, esoterists can become prideful and forget their connection to the exoteric teaching. The burden is on the esoterist to avoid causing scandal. That is why they should participate in the rites, even when they are understood on a deeper level, and avoid trying to force their ideas on the exoterists.

As an example, Ratzinger points out that Jesus is ``observant'', i.e., exoteric: he celebrates the Jewish feasts, prays in the Temples, recites the psalms. Yet he also provides the inner meaning to exoteric ideas such as the purity regulations, the inner meaning of the Ten Commandments, and the Temple. Jesus followed temple, prayed psalms. But he also brought something new, i.e., the inner meaning of the commandments, or the temple.

\paragraph{Symbols, Myths, and Science}

The interviewee also complained about the stories that he could not believe. That is not an uncommon problem, first of all because in the current state of degeneration there is little or no self-awareness as a spiritual being in many people. Nevertheless, this could still be a problem for seekers. Hence, we have pointed out Guenon's and Evola's understanding of myth, even translating some of Evola's essays on the myths of the founding of Rome.

Likewise, we've pointed out the deeper meaning in the infancy narratives of Jesus. Guenon also insisted that there could not be any conflict between positive science and metaphysic properly understood. That is why we posted an essay on the story of Adam and Eve. Moreover, we've used the examples of the positivists August Comte and Charles Maurras to show that there is no necessary disconnection between science and the natural moral law as such.

\paragraph{Cause of Degeneration}

Another common objection is that ``Christianity'' somehow ``caused'' the degeneration of the European people. In other words, a healthy minded people somehow grasps onto an idea that then causes their decline. That is an obvious absurdity, confounding cause and effect. Rather, a degenerate people, having lost sight of their spirituality, will then latch onto inferior ideas. Evola, in \emph{Sintesi}, understands this, although in a rather confused way.

\begin{quotex}
We can continue to call the Germanic peoples of the invasions of the Roman Empire ``barbaric''; barbaric, however, not in relation to degenerate Roman civilization, in which those people appeared, but in relation to a higher stage from which those races had fallen. Among the proofs of such an interior degeneration, or spiritual obscurity, is the relative facility with which such races converted to Christianity and then to Protestantism. For that reason, the Germanic peoples in the first centuries from the fall of the Western Roman Empire to Charlemagne, were not able to oppose anything of importance in the spiritual domain to the crepuscular forms of Romanity.
\end{quotex}

That is, he claims, the Germanics had already degenerated by the time of their invasion of the Roman Empire; hence they readily converted to Christianity. Oddly enough, in his view this degeneration lasted until Charlemagne, the driving force behind Christian Europe. So the inner logic does not support Evola's contention. It is true that the Germanics had fallen from a higher stage, but it was Christianity that then raised them back up, to form the foundation for the Medieval civilization.

\paragraph{Alleged Semitic Influence}

Another objection is that Christianity carried Semitic ideas into the European mind. Once again we turn to the thoughtful pagan Julius Evola. In this passage, he points out that the ancient Hebrews were spiritually sound.

\begin{quotex}
It is not the ancient Hebraic, messianic idea, but its degeneration and materialization that is the true focal point of the subversive forces turned to the final destruction of our civilization and to a Satanic dominion over all the forces of the earth. In its original sacred form, prior to the period of the prophets (that indicated the first mystical and democratic fall of Israel's ancient tradition), the idea of the Messiah had many traits in common with familiar conceptions and ideals of essentially Aryan civilizations, from which, moreover, the Hebrews more than once borrowed many elements.
\flright{\textsc{Julius Evola}, \emph{Transformazione del Regnum}}
\end{quotex}

He sees the idea of the expectation of the return of the ``Messiah'' as equivalent to the return to a ``Golden Age''. When the Messianic ideal degrades into a merely material form, devoid of its spiritual roots, then there is subversion. This is more or less what some have called the ``immanentizing the eschaton''. So here we have the justification for the Nine Worthies\footnote{\url{https://www.gornahoor.net/?p=331}}.

\paragraph{Conversion of Iceland}

A final point is the idea that Christianity was somehow forced on the pagans. Apart from some isolated cases, that was not what generally happened. As an example, we can point to the story of the conversion of Iceland. First of all, we cannot deny the courage of the Vikings, their ferocity in battle, fearlessness in the face of death, and their love of their fate.

Nevertheless, there was a certain amount of thuggery, as the Eddas reveal. They would often raid unprotected monasteries on the coast of England, even extracting ransoms from Paris. By the standards of the Saxon \textbf{Robin Collingwood}\footnote{\url{https://www.gornahoor.net/?p=7992}}, this disqualified them as a civilization.
Moreover, there was the acceptance of human sacrifice, e.g., of a slave when her master died, or of a boy killed to guard a buried treasure until its owner died. Does that make you squeamish? If so, then perhaps you have absorbed Christian sensibilities. You don't see neopagan revivalists promote such customs. But back to Iceland.

Around the year 1000 AD, Christianity had already appeared so the population became divided. The \emph{Saga of Burnt Njal} tells the story of a debate between the Christians and the Pagans. Each party declared its own laws; thus the Christians refused to be under the laws of the Pagans, and vice versa. The Christian leader went to the pagan priest \textbf{Thorgeir} to determine what the law should be.

Thorgeir spent the day in meditation. On the following day, the Christians and Pagans met again to hear his decision. Thorgeir explained that there could not be two conflicting laws and asked for pledges from both sides to abide by his decision.

In short, Thorgeir proclaimed that the Christian law would prevail, so that idol worship, exposing children to perish, and eating horseflesh sic would henceforth be prohibited. Furthermore, the holy days of Christianity would all be observed.

The Christians could have easily been suppressed, but the superiority of Christian civilization was recognized. Nowadays, that is denied, even by most Christians who believe in the separation of Church and State. It is forgotten that Christ is also the lawgiver, so that the social Reign of Christ\footnote{\url{https://www.gornahoor.net/?p=7768}} is the norm.

\paragraph{Reinvigoration}

Religion in the West today has lost its spiritual focus and is often little more than a humanitarianism or bland moralism. There are ideas that should be a living presence, not empty phrases. We have addressed several of these: the Sky Father, Cosmic Christ, Divine Sophia, Angels
\& Demons, theosis, spiritual warfare, the need for awakening. There are necessary efforts required for this reinvigoration. This applies, \emph{a fortiori}, to neopagan revivalists if they aspire to anything more than a superficial change.


\flrightit{Posted on 2015-04-02 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Mark Citadel on 2015-04-03 at 16:23 said:}

A really excellent piece. I'm going to have to post a link to it. As well as the challenge presented by the secular right of both Nietzschean and techno-commercialist schools, we also have to deal with the challenge posed by neo-Pagan movements such as the `New Right'. While these groups may possess admirable insight and correct political doctrines, they are missing many of the pieces to the larger jigsaw that will be a successful and ultimately triumphant Reaction.

Have a blessed Easter, Cologero

\hfill

\texttt{Mark Citadel on 2015-04-03 at 16:47 said:}

Excellent article. Linked to it. It is critical that we meet the challenge of Pagan as well as secular rightists. It is not enough to have a few pieces of the greater jigsaw if we wish to see an eventually triumphant Reaction.

Have a blessed Easter, Cologero

\hfill

\texttt{francismercuri on 2015-04-05 at 01:24 said:}

Concerning the alleged Semitic roots of Christianity: it is of interest that while there are those, as you point out, representing something of a ``right-wing of Western civilization'' if you will, who bemourn the influence; with the American popular ``right'', we find the other extreme--those who, calling themselves Christian, idolize every sort of idealization of Jewish religion and history, to the point of having something of a ``Jewish envy''. And, both seem to miss the boat.

Christianity is not, as these Christians (and there atheist opponents agree), a ``reformation'' of ``Judaism'' (or the Old Testament covenant), nor even a prolongation of it--it is something entirely different, whose relation with the OT involves two main points:

1) That it took the OT ``forms'', and installed them with new, and apparently contradictory meanings.

2) It developed the physical strata from Adam/Eve to Joseph/Mary in which to, ``in the fullness of time'', allow the Christ to take flesh, after which, the previous ``mantle'' was done away with, and of no more import.

Bulgakov points out, and contradicts those who consider the OT God as representing the ``Age of the Father'', because he argues that with the Mosaic ``I am that I am'', something of an exclusivity is involved, without suggesting a Father/Son relationship. Then, Jesus comes on the scene with something strange, the notion of God as ``Abba'', ``Our Father''. For Bolgakov, the ``Age of the Father'' is represented better by paganism than OT ``Judaism'', although this seems best balanced with Berdyaev's observations regarding the involvement of Jews in the history of revelation.

Of those extreme, literalist, Judeaophile ``Christians'', their logic has led to another consequence, which is mass support of political Zionism, and gross anti-Islamism. These ``dispensationalists'' consider that the ultimate aim of history must involve the rebuilding of the physical temple of Jerusalem, and restoration of primitive Jewish sacrifice, thus conditioning their support of this or that ``war''. But, from a Christian point of view, this is not only nonsense, but demonic. As we said, Christianity is something ``different'' and ``new''--as Paul said, blood sacrifice is not only no longer a valid element of tradition, it is, when offered, an offering only to the infernal kings and princes. This is the ``Abomination of desolation''.

And, so with Christ, a new law and new order came to being, cutting asunder anything of the previous covenant, while revealing a very new God.

There was need for ``salvation'', along with existence of free will; while neither the blood sacrifice of old, or man fallen, left to ``evolution'', could accomplish the act as such. Thus a new man had to be, partaking of both possibilities, while being neither alone. It is both ``this'' and ``that'', while being neither ``that'' nor ``this''. In Christ arisen as first born of the dead, Nietzche's formula is reversed, for ``That which should be is; and that which is should be''.

\hfill

\texttt{Jacob on 2015-04-05 at 17:10 said:}

It seems to me (though I guess it is arguable) that Christianity still has those three things Guenon says a religion needs. So it is not broken, but just needs redirection.

I think it'd be correct to say Christianity is the fulfillment of all Traditional paths and not only Judaism. The comment a few posts ago about Christ being a priest in the order of Melchezidek as opposed to the order of Aaron has really stuck with me.

\hfill

\texttt{francismercuri on 2015-04-05 at 18:22 said:}

Jacob: In addition to Paul's Letter to the Hebrews (7), and Psalm 110, note the Roman Catholic Eucharistic Prayer I (sadly, seldom used these days)... .

``Look with favor on these offerings and accept them as once you accepted the gifts of your servant Abel, the sacrifice of Abraham, our father in faith, and the bread and wine offered by your priest Melchisedech''.

\url{http://www.catholicliturgy.com/index.cfm/fuseaction/text/index/4/subindex/67/contentindex/22/start/9}

\hfill

\texttt{Logres on 2015-04-05 at 19:35 said:}

If I am remembering correctly, Evola's dealing with Nietzsche in \textit{Revolt Against the Modern World} was both more sensitive and appreciative, and more dismissive, of him, than the New Right would be comfortable with.

\hfill

\texttt{Pickman on 2015-04-06 at 15:10 said:}

The problem still remains with the question of origins. No matter how far down the line we are, the point of ownership for the Western Indo-European (Aryan if you wish) Caucasian will still be problematic.

It was this initial schism of origins that had a profound effect on Evola and much of the Right in general. To the extent that it can be solved at all, still remains indeterminable despite these passages.

\hfill

\texttt{Pickman on 2015-04-06 at 15:13 said:}

The ``alleged'' Semitic influence, is still a noteworthy variable to contemplate. It is not alleged, but a fact. How we deal with this has become a matter of mental gymnastics.

\hfill

\texttt{Logres on 2015-04-06 at 23:19 said:}

The alleged influence on what object? The influence had already entered the barbarian peoples (which explains, for instance, why when you read about the Gallic Wars, Caesar comes off as a good deal more even handed and just than most of the Celtic tribal leaders, like Indutiomarus). Are you talking about the West, Christianity, or perhaps the Aryan peoples? I agree there are a lot of mental gymnastics masquerading in the real world, but it's certainly fair to point out that the ``nice'' kid who's hanging with the bad crowd can't really blame their ``influence'', since bad company corrupts good morals. There was already an inner correspondence, or ``elective affinity'' to begin with. Maybe the Semites were smart to dump their moral and mental refuse on such willing test subjects; it would be hard to say from the perspective of the New Right. It's certainly been a winning strategy so far.

\hfill

\texttt{Cologero on 2015-04-08 at 01:12 said:}

The ``problem'' exists in your own mind, Pickman. Apart from the ``right in general'', which has absolutely no interest for us, Evola identified the origin. Bal Tilak provided evidence from the Vedas and Herman Wirth attempted a more scientific proof. In addition, there were Teodoro Poesche, Karl Penka, and Ludwig Wilser, in case you want to do your homework. Since you've shown no evidence of being familiar with any of these ideas, your one sentence dismissal is a waste of the disk space that I am paying for.

No issue of import has ever been resolved through philosophical demonstration or scientific proof. Hence Evola ultimately points to Fabret d'Olivet who was privy to traditional and esoteric sources. Now you may choose to live your life like Schroedinger's Cat in a state of mental indeterminacy, but a man has to make a decision one way or the other.

\hfill

\texttt{Justin on 2015-05-12 at 12:34 said:}

An excellent piece, Cologero. Your treatment of Exoterism and Esoterism brought to mind the late great Stratford Caldecott's description of Christianity as ``a kind of Esoteric Judaism turned inside out and offered to the world.''

\url{http://thechristianmysteries.blogspot.co.uk/2011/11/is-christianity-esoteric.html}

\end{sffamily}
\end{footnotesize}

