\section{Which Way, Western Civilization}

\begin{quotex}
We believe France has not benefited fully from the energy, drive, and ideas of its minorities.

We will track intangible measures of success --- a growing sense of belonging, for example, among young French minorities, and a burgeoning hope that they, too, can represent their country at home, and abroad, even one day at the pinnacle of French public life, \emph{as president of the Republic}.

19 January 2010 Cable from \textbf{Ambassador Charles H. Rivkin} to \textbf{Secretary of State Hilary Clinton}
\end{quotex}

\paragraph{The Basis of Western Civilization}

\begin{quotex}
Armin Mohler takes as the basis of the vision of the world the opposition existing between two general conceptions, which he calls \emph{linear} and \emph{cyclic}. According to the former, history is development, novelty, evolution and tends to a final end that justifies it; this is the conception characteristic of the various progressive currents, but Christianity as well insofar as it gravitates toward an ``end of times''. The latter conception is based instead of the idea of the ``eternal return'', of the repetition of the same forms; that would be the basal view of the ``conservative revolution''. In our opinion, the contraposition in these terms is not well formulated. If anything, we should speak of historicism and antihistoricism, of ``civilizations of being'' and of ``civilizations of becoming''. It is not a question of expecting the return of the same forms, but rather of believing that fundamental values never change, of recognizing a \emph{normative} order containing \emph{a priori} and \emph{ab initio} all the principles, without which a civilization and a normal society are inconceivable.
\flright{\textsc{Julius Evola}, \emph{Rassegna Italiana, Rome, June 1952}}
\end{quotex}

There has been a lot of loose talk among the paid intelligentsia this week about the need to ``save'' Western civilization. The claim is that it is under threat from a group known as ISIL. Now the hundreds of casualties in Paris last Friday is a horrific crime, but hardly an ``existential'' threat to France as some have claimed. There is no danger of regime change in France or the breakdown of her political and economic structures. ISIL has no navy, no air force, and perhaps 30 thousand soldiers. Chicago, USA, has more gang members that than, and more murders this year than Paris will have.

Even the New York Times predicts that the impact on Parisian life and economy will be temporary and of short duration\footnote{\url{http://www.nytimes.com/2015/11/17/business/international/france-shootings-economy.html}}. So, if there is in fact an existential threat, the danger is within. Perhaps there is a sense of unease because that menace still lies below conscious awareness.

In any case, this is the time for reflection on exactly what is this Western civilization that is worth preserving? What exactly are those enduring values, the normative order, and the a priori principles? This is more than a way of life, the smell of fresh bread from the bakeries, a croissant in the morning, a glass of wine for dinner, and so on. It is more than museums, art, and all the other artifacts of modern life. None of this has been addressed. Most importantly, does Western civilization depend on the Westerners who created it, or is it more like an empty shell that can be inhabited by anyone without changing its essential nature?

\paragraph{Phases of Western Civilization}

We can look at the phases of Western civilization and perhaps find some answers there.

\paragraph{Primordial State}

\begin{quotex}
According to Roman tradition, the Golden Age of Saturn was a primeval era of purity and simplicity---virtues that made Rome great. Saturn, originally an Italian god of agriculture, was deposed by his son Jupiter, king of the gods. Thereupon, Saturn fled to Latium, where he became its king, establishing a society that lacked weapons, money, walled cities, and all similar corrupting influences. During this era the fruits of the soil were gained without toil.
\flright{\textsc{Alfred J. Andrea}, \emph{The Human Record}}
\end{quotex}

The Romans had their own Garden of Eden story. The dream of this primordial state enthralls many today, especially those who fancy themselves the best and the brightest. Who do you recognize proclaiming the following values?

\begin{itemize}

\item No weapons: No guns, no war

\item No money: everyone will share in the goods of the community

\item No walled cities: Open borders and free migrations

\item Fruits without toil: free medical care, free college tuition, free food, free cell phones
\end{itemize}

Unfortunately, Saturn is also the scythe bearer who used it to sunder the relationship between Heaven and Earth. Hence, the reign of Saturn is a human creation, ultimately closer to a nightmare than to a dream.

\paragraph{Roman Ecumene and Ancient Athens}

Unfortunately, Jupiter prevailed over Saturn. Rome was founded by Trojan descendants who maintained Greek civilization. The Roman Ecumene was based around the Mediterranean, including North Africa and parts of the Near East, so it wasn't quite European. It excelled at administration, military affairs, and engineering. When it decayed from within, the Northern barbarians occupied the shell. However, they could not administer the far flung empire, nor could they maintain the carefully engineered roads and aqueducts.

\paragraph{Christendom}

Christendom replaced the Roman Ecumene. As learning and order were slowly restored, the Holy Roman Empire became the model for Western civilization proper. The African and Asian parts were lost, and it began to extend throughout Europe. In many ways, it was the interiorization of its pagan predecessor. It made public the idea of the one God, which previously had been known only to Sages or those in various mystery religions.

\paragraph{Enlightenment}

What people mean today by ``Western civilization'' is actually the Enlightenment civilization with more or fewer features from the pagan and Christian past. However, this is not ``Western'' at all, since it claims to be universal. Rather, it is a global culture that took root first in the West. It has ideas of universal human rights, supported on empirical science rather than revelation, as in Christendom, or thinking, as in the pagan times. Conflicts arise due to the archaic remnants of the past still remaining. However, once you accept enlightenment principles, the a priori and ab initio principles give way.

\paragraph{The Future of Intelligence}

Each era has its ideal of the intelligent man. The Philosopher was the model of the educated man for the ancient pagans, and the Theologian, for the Medievals. Of course, it is the Scientist who is supreme in the Enlightenment culture. The Philosopher, the Theologian, and the Scientist are always in battles for the intellectual high ground.

The fight is hardly fair. As this chartof GRE scores demonstrates\footnote{\url{https://dailynous.com/value-of-philosophy/charts-and-graphs/}}, the hard sciences and mathematics are attracting the most intelligent students. Theologians, or ``religious studies'' students, are not even close. Philosophy is hardly relevant anymore. As such, it is mostly verbal jousting with little relevance to life. It has been supplanted with Sophistry, i.e., those paid to promote certain opinions, and Ideology, or various ``studies'' that have little relation to actuality, but rather to dreams of a return to the rule of Saturn or, more often, Matriarchal and Lunar cultures\footnote{\url{https://www.gornahoor.net/?p=8154}}.

Since the promoters of sophistry and ideology are typically majors in political science, journalism, and the like, this is seldom noticed. They are not at the highest ends of the intelligence grid. It might be embarrassing to mention the education majors.

To get back to the point, in debates between science and religion, the latter is at a severe disadvantage since it lacks the most intelligent spokesmen. Often, the religious adherents are simply embarrassing, intellectually.

So the Enlightenment culture is actually an anti-culture, since it involves the overturning of the values, principles, and order of Western civilization, properly called. This deserves to be overthrown, but not by violence. Rather, a new generation will create it, a special type who can master Philosophy, Theology, and Science. These will be the Artists of a new civilization.

\paragraph{Embassy Paris: Minority Engagement Strategy}

The basis of Enlightenment civilization is shown clearly in this rather interesting cable from Wikileaks: Minority Engagement Strategy\footnote{\url{https://wikileaks.org/plusd/cables/10PARIS58\_a.html}}, which I encourage you to read.

Keep in mind that prior cultures in Europe recognized different kinds of people, and accepted cultural diversity as the norm. \textbf{Joseph de Maistre} wrote this about the French Constitution:

\begin{quotex}
The constitution of 1795, like its predecessors, has been drawn up for Man. Now, there is no such thing in the world as Man. In the course of my life, I have seen Frenchmen, Italians, Russians, etc.; I am even aware, thanks to Montesquieu, that one can be a Persian. But, as for Man, I declare that I have never met him in my life. If he exists, I certainly have no knowledge of him. ... This constitution is capable of being applied to all human communities from China to Geneva. But a constitution which is made for all nations is made for none: it is a pure abstraction, a school exercise whose purpose is to exercise the mind in accordance with a hypothetical ideal, and which ought to be addressed to Man, in the imaginary places which he inhabits... .
\end{quotex}

Nevertheless, both the USA and France believe they know what is best for ``man'' from China to Geneva, although they have opposed views on how to implement it. The French believe in equality and secularization. That is why they make it a point not to consider a citizen's race, ethnicity, or religion. Hence, they don't keep any statistical breakdowns based on those criteria. Any disparity in outcomes is due to the man, not his ethnicity or religiosity.

The USA, on the other hand, counts everything. From income levels, educational attainment, crime and prison population, etc., it keeps the numbers. If the outcomes don't match the population percentages, that becomes a problem to be solved.

Apparently, judging from the ambassador's cable, France's disparate outcomes are likewise problems that the USA feels compelled to solve. Hence, it interferes in France's internal affairs, seeking to further the advancement of the Muslim and other minority populations. The State Department does that through NGOs that it funds\footnote{\url{https://en.wikipedia.org/wiki/National\_Endowment\_for\_Democracy}}. There are four: two for the main USA political parties, one for business interests, and another for labor. These are allegedly independent, although they are staffed by the children and friends of political bigwigs.

The cable shows that the USA is counting France's minority participation in public life. The cable explains many of statements made by President Obama and Secretary Clinton, which otherwise may seem inscrutable. For example, the following statement may explain Obama's ``I told you so'' attitude about recent events:

\begin{quotex}
We believe that if France, over the long run, does not successfully increase opportunity and provide genuine political representation for its minority populations, France could become a weaker, more divided country, perhaps more crisis-prone and inward-looking, and consequently a less capable ally.
\end{quotex}

Surprisingly, or not, the USA considers a Muslim president of France to be a desirable outcome. This is why Obama and Clinton are so careful about how they phrase their statements about Islam. For them, Islam is not the enemy, but actually the end game for France.

ISIL should have had more patience since they have a potential ally in an unexpected place. Instead, they risk being annihilated by Russia.


\flrightit{Posted on 2015-11-19 by Cologero}
