\section{The Hyperborean Race and its Branches}

\begin{quotex}
This is Part 2 section 7 of \emph{Sintesi di dottrina della razza} by \textbf{Julius Evola} in which he discusses the Hyperborean race. The texts that he refers to are \emph{Revolt Against the Modern World} and \emph{The Myth of Blood}; the chapters in Revolt on the Hyperboreans should be read in conjunction with section.

From the corporeal perspective, this can be understood as a specific race of humans. From a spiritual perspective, it can be understood in terms of sacred geography. In the latter case, the Hyperboreans represent a race of people who lived in the light, i.e., in self-effulgent consciousness. Or both interpretations can coincide. 

\end{quotex}
The limit that can be given to our doctrine of the race regarding the exploration of its origins lies at the point when the Hyperborean race had to abandon the Arctic home in successive waves following different paths, because of the freezing that rendered it uninhabitable—it was already pointed out, in previous citations, justification for the idea that the Arctic region became completely ice bound only after a certain period: the memories of that home, preserved in the traditions of all peoples in the form of various myths where it always appears as a ``land of the sun", as an island continent of splendour, as the sacred ground of the God of light, and so on, were already, in this respect, rather eloquent.

Now, at the point of the beginnings of the prehistoric Hyperborean emigrations, the Hyperborean race could be considered, among all of them, the superior one, the super-race, the Olympic race reflecting in its extreme purity the very race of the spirit. By and large, it seems that all the other human stocks existing on the earth in that period, appeared either as ``races of nature", that is, animalistic races, or, as races that had become ``races of nature" through involutions of the preceding racial cycles. In fact, the traditional teachings speak of a civilization or of an Antarctic race that had already fallen into decline in the period of the first Hyperborean emigrations and colonisations, and whose Lemurian residues were represented by sizable groups of black and Malaysian races. Another racial stock, distinct both from the Hyperboreans and from the Antarctic-Lemurians, was that which originally occupied the Eurasian continent (the Finnish-Mongoloid race) as the yellow-brown race and, as the as red-brown race and also again the yellow-brown, occupied both a part of the Americas and Atlantic lands that are now extinct.

It would obviously be absurd to attempt a precise typology of these prehistoric races and of their primordial combinations in regards to exterior characteristics. We must refer to them only to prevent some misunderstandings and to be able to orient ourselves among the ethnic formations of the successive periods. Also, the investigation of skull fossils can tell us very little, both because race is not characterized just by the skull, not even the simple race of the body, because there are reasons to justifiably assert that, for some such races, fossil remains could not be preserved up until our time. The dolicephalous, or elongated skull, combined with a tall and slim figure, blonde hair, clear skin, blue eyes, is, as we noted, characteristic of the most recent descendants of the Nordic races who came directly down from the Arctic regions. But all that does not constitute the final word; even within the limits of the positive order, it is necessary to bring in considerations proper to the study of the race of the second level [race of the soul] in order to get our bearings.

In fact, we already said that the essential element for race is not given by simple corporeal and anthropological characteristics, but by the function and the meaning that they have in the whole of a given human type. Dolicephalous men with a tall and slender build are also found in fact among the black races, and a white coloration and almost blue eyes are found among the Ainu of the Far East and the Malaysian-Polynesian races, naturally having a totally different significance in such races. Nor must we think here only of anomalies or tricks of nature, in certain cases being able to treat of extinct somatic survivals in the proceeding types from races which, in their most remote zenithal period, were able to have similar characters to those which, in the period we are considering, were instead found concentrated in the Nordic-Hyperborean element and here accompanied, up to a relatively recent period, by the significance and by the corresponding inner races.

As to the emigration of the races of Hyperborean origin, having previously written about them in the books previously cited, we can limit ourselves to noting three principle currents. The first took the direction northwest to southeast reaching India and having as its final echoes the Indian race, indo-afghan and indo-brachimorphic in Peters' classification. In Europe, contrary to what you may believe, the traces of such grand currents are less visible or, at least, more confused, because they had a superposition of waves and therefore a composition of successive ethnic layers. In fact, after this current in the northwest-southeast direction (the transversal Nordic-Aryan current), a second current followed the direction West to East, in most of its branches through the routes of the Mediterranean, creating centres that sometimes should be considered even more ancient than those derived from the preceding transversal wave, because in this case it is not always about a forced emigration, but also of a colonization occurring before the destruction or the eventual uninhabitability of the original centres of the civilization of Hyperborean origins. We can call this second current, with the related trunk of races, Ario-Atlantic, or Nordic-Atlantic, or, finally, Atlantico-Occidental. In fact, it originates from an Atlantic land, in which it constituted a centre that, at its origin, was more or less an image of the Hyperborean centre. This land was destroyed by a catastrophe, whose memory likewise is found in the mythologized memory of the traditions of almost all peoples, and then after the waves of the colonizers a true emigration followed.

It is said that the land of Atlantis was familiar at its origin with a type of facsimile of the Hyperborean centre, because the data reaching our time lead us to think that a regression happened, both from the point of view of race and from the point of view of spirituality, in these Nordic stocks who had already ventured towards the south in the most ancient epoch. The mixing with the red-brown aborigines seems, in this regard, to have had a destructive and sizeable role, and if a precise record in Plato's story is not found, in which the union of the ``sons of the gods"—of the Hyperboreans—with the indigenous people was given the blame in the end, that recall what, in other recorded myths, are described as the ``fall" of the heavenly race—of the ``angels" or, again, the sons of the gods, \emph{ben elohim}—who mated, at a particular moment, with the daughters of men (of the inferior races), committing a contamination significantly likened to, by other texts, the sin of sodomy, carnal relations with animals.


\hfill

Further reading

From the positive, scientific perspective, see \emph{The Science of God} where the physicist Gerald Schroeder discusses the existence of pre-Adamic hominoids.

In \emph{The Psychology of Man's Possible Evolution}, Peter Ouspensky mentions the possibility of the existence of pre-Adamic hominoids.

Boris Mouravieff, who claims to be drawing on ancient Western traditions, accepts the existence of a pre-Adamic race in his \emph{Gnosis} trilogy.

The idea of the polar origins of the Hyperborean race was discussed in \emph{The Arctic Home in the Vedas} by \textbf{Bal Tilak} based on myths recorded in the Vedas.

For a recent theological and metaphysical discussion of the existence of humanoid animals, see \textit{Knowing Ape from Adam}\footnote{\url{http://edwardfeser.blogspot.com/2014/12/knowing-ape-from-adam.html}}.



\flrightit{Posted on 2015-03-01 by Aeneas }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{John Bardis on 2011-03-26 at 22:12 said: }

In regard to the reference to Mouravieff and the pre-Adamics, this has to do with the idea that a human has three so-called centers: moving, emotional and intellectual (while animals lack the intellectual center). But, then, humans also have higher emotional and intellectual centers.

According to Mouravieff there are, basically, two species of humans: those with higher centers and those without higher centers. So someone without the higher centers would be an animal with an intellectual center–a very smart animal.

Those without higher centers live only one life, like animals, then they die. People with higher centers reincarnate.

Those without higher centers can be quite attractive physically, as tends to be the case generally with animals. After all what is more beautiful than a horse or a tiger or a hawk? And, in fact, these people without higher centers very often have blue eyes and blond hair, as everybody knows.

Mouravieff, of course, isn't talking about ancient pre-history. He is talking about the situation today. And in fact those without higher centers far out number those with higher centers.


\hfill

\texttt{Cologero on 2011-03-26 at 23:18 said: }

Of course Mouravieff is referring back to ancient pre-history, as in the first chapter of Book 2. M. claims that there were two distinct creations: first pre-adamic humanity and then Adam. M. also refers the the miscegenation of the two races in terms very similar to Evola's:

\begin{quotex}
It is since Adam that man has reeived the capacity to pass form growth to development in his evolutoin, and \emph{only part of the humanity of those times recieved this gift.} The Bible speacks of a long period of coeistence of the first humanity alongside adamic humanity. It later refers to the latter as passing through a perido of recession following mixed marriages which were considered by God as evidence of \emph{great perversity} … the first humanity still retained animal characteristics 

\end{quotex}
I don't recall anywhere that M. described the physical characteristics of pre-Adamic man, so they don't seem to be identifiable with any specific peoples today. In any case, we are primarily interested in the spiritual race, as you described, rather than the corporeal race.

Mr. Bardis, you must travel in rather exalted circles where ``everyone" knows the inner states of their associates. I myself am dubious about that claim. The primary difficulty is that the Adamic man who has the higher states only in potentitial — that is, unactualized — is indistinguishable from the pre-Adamic man who lacks them totally.


\hfill

\texttt{John Bardis on 2011-03-27 at 07:39 said: }

Well, yes, certainly, there is the whole question of ancient pre-history.

And the main problem, of course, that Mouravieff deals with at great length is how Adamic man can actualize his higher states.

But there's a further problem that is of much interest to Mouravieff, and that is, quite simply: What is the place in society of pre-Adamics? I suppose, in terms of caste, they would be the fourth caste. Obviously this caste makes up the vast majority of the population. So, then, there are very serious ethical and social questions involved here.

As to your last point, yes, certainly, that's why I mentioned that the color of eyes or hair or skin has nothing to do with it, at least today. Undoubtedly before the mixing of the races such factors gave a clear indication of the matter.


\hfill

\texttt{Cologero on 2011-03-27 at 11:56 said: }

In chapter 14 of Book 3, M does attempt to address that issue. Yes, the fourth caste, because only the first three castes are ``twice-born" … a phrase used in the Vedas, by Evola, and again by Mouravieff. I brought it up because it is fundamentally the same position as other esoteric teachings … something M himself points out. What he does, though, is to move the question of race totally to the spiritual realm, a direction Evola also traveled, but only halfway.

So the challenge is whether Mouravieff has truly revealed esoteric Christian teachings that had been kept secret. If true, what does it do to the alleged idea of Christian universalism or the idea we are ``all children of God"? I don't see how this could be resolved short of a sufficient number of Adamic men becoming ``twice born"; they would ``see" it.


\hfill

\texttt{John Bardis on 2011-03-27 at 13:52 said: }

There's an implicit argument in Gurdjieff-Ouspenski-Mouravieff that is quite suggestive.

Natural man is either man 1, man 2, or man 3. With spiritual development one gets man 4, man 5, man 6 and man 7 (with man 8 making one brief rather unfortunate appearance).

Supposedly it takes 100 natural men to support the existence of one man 4, 100 men 4 to support the existence of one man 5, 100 men 5 to support the existence of one man 6, and 100 men 6 to support the existence of one man 7.

So, then, one can imagine the redeemed earth has having a population of 101,010,101 adults, consisting of one legitimate ruler of the world and 100 million natural men. So this problem of how to integrate these natural men into society is quite pressing.

But, then, there's the question: Could this planet support a man 8? This would require a population of more than 10 billion adults–and, again, with a full 10 billion of them, almost all of them, being natural men.

Of course I'm not suggesting that the above is the literal, scientific truth. It is just a hopefully suggestive model.


\hfill

\texttt{Cologero on 2011-03-29 at 20:05 said: }

Isn't the situation more dire than you suggest with mere numbers? Unlike Traditional societies, where the spiritual order was recognized and accepted, the natural man of today. in his spiritual democratism, does not even accept the possibility of such an hierarchy that you propose. How then would he ever consent to support it?


\hfill

\texttt{John on 2011-03-30 at 18:33 said: }

According to Mouravieff there are a certain number of souls assigned to our planet. The final redemption of the earth will take place when all the souls can incarnate at the same time.

The final redemption is also associated with the coming again in glory of one who Gurdjieff calls a man \#8. 

This suggests that there are 100 million souls assigned to our planet. This means that our planet should be able to sustain the existence of 10 billion ``natural men". This, of course, isn't possible without the development of technology.

This also has to do with the question of the Pope as opposed to the councils. If the Church isn't big enough to sustain a man \#7, then the men \#6 would be involved in a council. Or if the Church were big enough to support several men \#7, then this would call for a council. But if there are 100 men \#6, then a man \#7 will rule. And if there are 100 men \#7, then Man \#8 will come in glory to rule the world.

This has to do with the resurrection of the body. This has to do with the kingdom of heaven on earth.

What ``natural man" will support is the least thing to be concerned about. It is odd, by the way, that as animals become extinct, or as the population of animals decreases, the population of humanity increases. Is this just a coincidence?


\hfill

\texttt{Cologero on 2015-03-03 at 21:20 said: }

Obviously, despite the many responses to the poll question, this is not very interesting. I am aware of one outside comment: \textit{The Sons of God}\footnote{\url{https://evolacomments.wordpress.com/2015/03/03/man-is-the-son-of-god/}}.


\hfill

\texttt{Marcel K. on 2015-03-04 at 08:31 said: }

I have read the German translation by the man himself several times over the last few years. Lately I have been questioning the `white nationalist views I have held in the past, which makes this the most problematic of Evola's writings for me, due to the heavy influence of Nordicist thought which I consider to be pseudoscientific at best. Evola emphasizes that the second and third level of racial classification are more important than the first, but states that there is a significant correlation between the corporeal race and the races of the soul and the spirit. Like Ludwig Clauß he believed that a solar spirit is correlated with a `Nordid' anthropological type, which he considered to be the original Indo-European type.

Today we know that the `Nordid' type developed as an adaption to the harsh climate of northern Europe, and that the bearers of Indo-European culture were of a different type which, for all we know, might not have been all that European in the racially modern sense. I think that in order to solve the problem of race, Evola's simplistic theories concerning the homeland of the Indo-Europeans and their anthropological type have to be rejected. The primary significance of the Hyperborean race and the Urheimat is that of a myth, but I believe it would be wrong to base this myth on a faulty anthropology like John and Ted did in their comments.

However I think it is too early to dismiss the notion that there exists a mysterious physical type in which the solar spirit is very common. Perhaps with modern genetic and anthropological evidence we will eventually be able to trace back Indo-European civilization to a distinct physical type. I wouldn't rule out the possibility that there is a modern group of people who retained this type, or in whom it reemerged. Consequently we might be able to predict where a revival of tradition is most likely to occur. But to my mind it is just as likely that the focus on physical race is entirely misguided, and that a connection between the race of the body and the race of a spirit does not exist.

This comment is very rambling and I apologize if it misses the mark.


\hfill

\texttt{Cologero on 2015-03-04 at 09:00 said: }

Actually, Marcel, your comment is right on the mark. That is why we have been reluctant to publish the full translation, despite its being nearly complete. Such a discourse on race and the Hebrews will not be understood in today's climate. Evola himself, in his later books, toned down the rhetoric he used in the 30s and 40s; not surprisingly, they are the most ``popular" (viz., Ruins, Tiger). I don't think Evola is so much simplistic, rather that he has an utter disdain for positive science. His methodology of using myths to reconstruct the past is opposed to ``objective" scientific methods.

The deeper issue, which is Evola's real motivation, is to oppose universalism, individualism, rationalism, and evolutionism. I think there are more effective ways to do that. For example, Gornahoor has been promoting a typology based on the various possibilities of inner dispositions, i.e., based on one's dominant center of sensation, emotion, or thought. This serves the purpose that the notion of ``spiritual race" was intended to serve while being independent of physical race. Moreover, it is tied into the traditional notion of the interior states of man, so it is verifiable … not scientifically, but esoterically.


\hfill

\texttt{William Neville on 2015-03-04 at 10:27 said: }

While I agree that publishing such material would probably open up a good deal of misunderstanding and futile argument, one cannot disregard the meaning of physical race altogether. Inner states trump their outer manifestations, yet the symbolism of the outer race still points to something higher, even if that something higher no longer exists within the person bearing the physical mark of it. Though a juvenile biological racism is to be much avoided, I don't think it is a mistake that people have recognized something special in the fair northern races for thousands of years

For example, let us look at what Pope St. Gregory the Great said about the Angles (the antecedents of modern day Anglo-Saxons), according to the Venerable Bede. We can see that he acknowledges the supernatural significance of their fair skin and light hair, etc. while recognizing that it did not (at the time) reflect their individual inner state in terms of salvation and enlightenment:

``I must here relate a story, handed down to us by the tradition of our forebears, which explains Gregory's deep desire for the salvation of our nation. We are told that one day some merchants who had recently arrived in Rome displayed their many wares in the marketplace. Among the crowd who thronged to buy was Gregory [not yet elected pope], who saw among other merchandise some boys exposed for sale. These had fair complexions, fine cut features, and beautiful hair. Looking at them with interest, he enquired from what country and what part of the world they came. ``They come from the island of Britain," he was told, ``where all the people have this appearance." He then asked whether the islanders were Christians, or whether they were still ignorant heathens. ``They are pagans," he was informed. ``Alas!" said Gregory with a heartfelt sigh, ``how sad that such bright faced folk are still in the grasp of the Author of darkness, and that such graceful features conceal minds void of God's grace! What is the name of this race?" ``They are called Angles," he was told. ``That is appropriate," he said, ``for they have angelic faces, and it is right that they should become joint heirs with the angels in heaven. And what is the name of the province from which they have been brought?" ``Deira," was the answer. ``Good. They shall indeed be rescued de ira— from wrath—and called to the mercy of Christ. And what is the name of their king?" ``Aelle," he was told. ``Then," said Gregory, making play on the name, ``it is right that their land should echo the praise of God our Creator in the word Alleluia."


\hfill

\texttt{Lu Dongbin on 2015-03-04 at 15:15 said: }

Regarding the racial type of the original Indo-Europeans, I am not inclined to agree that the bearers of Indo-European culture were not racially ``Nordid" or at least ``Nordish." The evidence is overwhelmingly in favor of them being Nordic racially due to the fact that the Nordid type is present in all IE groups to some degree, as far West as Ireland and as far East as the Tocharians/Da Yuezhi, whose mummies discovered in Xinjiang confirm this fact.

Furthermore, it is known that this type pops up among Pashtuns, Hunza, Pamiris, Kalash, Kashmiris, Persians, Kurds, and so on even today, albeit rarely. The Brahmin castes of India still have the highest incidences of Europid features in relation to the rest of the population and ancient Greco-Roman literature is replete with various figures who possessed Nordic racial features, such as Achilles, Alexander the Great, Sulla, etc. Furthermore, thunder gods like Thor and Indra are often described as copper-bearded, to say nothing of other gods like Apollo, etc.

In my view, arguing that the original Aryans or Indo-Europeans were not racially Nordic or Nordish is more due to the modern backlash against the regimes which propagated that view in the 1930s and 40s, political correctness, modern Indian nationalism, and so on. It's very obvious what type they were, which is why most of the early European investigators prior to any politicization of the issue held that view.

The above says nothing about the spiritual race I realize, but nonetheless felt it worthwhile to add. While the Hyperborean seat also has a spiritual dimension, I think the memory of such a seat in the various mythologies of IE people as well as the data presented in Tilak's ``Arctic Home in the Vedas" leads to convincing evidence that it was also a historic reality.


\hfill

\texttt{Max on 2015-03-04 at 16:18 said: }

If the Tradition began in the North, it is reasonable that it should return towards the end. The mission of the middle ages could be seen as a transmission of the torch from west and northwards. According to the ancient sources it seems that the polar region had a pleasurable temperate climate which is obviously not the case today and i am unsure of what geological process causes that and if it will reverse someday. It could be interesting to compare the mythological sources to what the natural sciences know of the ice ages.

Concerning the movements out of the Arctic during the big freeze, there is the notion of ``ice-age refuges": places somehow sheltered from the worst conditions. It is likely that not everyone relocated to southern lands but that some populations survived maybe isolated and surrounded by glaciers, giving rice to genetic bottle necks in northern populations. New findings are coming to light, it was previously thought that the whole of Scandinavia was covered by several kilometers of ice but that view is turning out to be false. Solid ground existed and some species of plants and animals were never extinct in the northern regions of Eurasia and consequently did not have to recolonize from the south. Exactly where they survived however is more uncertain. In any case, we know that the northern human racial characteristics could only develop in extreme conditions and in isolation, probably in a very small group of people.

For a race to actually live in the conditions after the descent of darkness gives ``inner light" a whole new meaning so i find ideas like the ``personal sun" interesting in that regard. Just imagine spending maybe months and years in deep meditation, almost hibernation, with the Sun all the while preserved internally like a sacred memory. In norse mythology the primordial cow Audhumbla licked the ice and from there emerged the first Man while her life-giving milk ran in rivers to the four corners of the world. If thinking about what could possibly sustain humans in such harsh conditions milk is an answer. Compare a map of worldwide lactose-tolerance with the current theories on cattle domestication in the middle east to see if there is any inconsistencies. The Nordic race is the only one to be completely adapted to a life together with animals without having to kill them for meat. This should be kept in mind when meditating on Man's original task on earth as guardians and protectors of the animals and forests. Peoples of the germanic and nordic countries to this day feel strongly for animals even though it has been sentimentalized into irrational forms lately.

In earlier times it was said that we all have to eat or be eaten, life is sustained by death, the world is a circle and so on. We now know that is not altogether true. There is another possibility of true freedom, goodness and beauty by cooperation for the benefit of everyone. I read somewhere that the greatest act of charity is towards those we do not expect anything in return from.

Just because i had Aguéli on my mind recently i would like to mention that he was responsible for preventing Spanish-style bullfighting in France by shooting and wounding the matador with a revolver. He also smoked opium, dressed in strange costumes, partook in anarchistic conspiracies, painted modern art, traveled through the orient and translated esoteric arabic treatises on ``the people of blame". Well, at least it is refreshing to see intellectuality and uncompromising action united in one person considering how uncommon that is.

If it is possible for humanity to inherit original sin from one man, the merit of the forefathers should also be equally possible to inherit. It means that the lineage and in the end race really matters. Even if several races or families have a common origin, not all of them may have preserved that heritage equally. In the case of a king, he is the highest representative of the nation or the city because he is the fullest expression of their nature. The head of the senior line that in the purest and most undiluted way preserve the original impulse and wholiness and by his presence will allow his subjects to be whole as well, directing their action higher.

In old Aryan society, worse than death was exclusion -becoming an outlaw. Failure to live by the law means life outside the law, astonishingly moral and simple. Our present society however idealizes criminals and outcasts. Relegation to the forest is not necessarily a punishment since the forest is apart from outlaws inhabited by solitary mystics and holy men. When looking for the way forward, and some are not even aware that there is anything to look for, it is easy to miss the forest for the trees. The truth is we have been judged to the forest passage, an exile, and the first step is a change of perception: ``We are already in the forest".

``The real issue is that the great majority of people do not want freedom, are actually afraid of it. One must be free in order to become free, because freedom is existence – it is above all a conscious consent to existence, and the desire, perceived as a personal destiny, to make it reality. At this point man is free, and this world filled with oppression and oppressive agents can only serve to make his freedom visible in all its splendor, just as a great mass of primary rock produces crystals through its high pressure." -Ernst Jünger, The Forest Passage


\hfill

\texttt{Aeneas on 2015-03-15 at 19:55 said:}

They (Evola, Guenon) write of the increasing density of men, more or less, equivalent to the distinction of pneumatic, psychic, and hylic men. As such, the Hyperboreans cannot be identified strictly by DNA (the hylic/material level) nor even by archeological clues, which actually would reflect the soul level. The gap between the man whose self-understanding is as an I with a will compared to the man who regards himself as a “computer made of meat” is enormous.

Hence, identification can only be made by “remembering”. This is a truth that is not found, but “willed” as Scaligero points out. Those who do remember will need to find each other.

\hfill

\texttt{Max on 2015-10-05 at 12:34 said: }

George MacDonald wrote a children's novel on the theme of Hyperborea called ``At the Back of the North Wind".

`` `You see, Diamond,' said North Wind, `it is very difficult for me to get you to the back of the north wind, for that country lies in the very north itself, and of course I can't blow northwards.'"

No road leads to that mysterious country. When approaching the magnetic north, compasses presumably become useless. Young Diamond in the story however finds his way. There is a fitting proverb: ``The sailor does not ask for tailwind. He learns to sail". A sailor makes use of what circumstances he happens to be in to reach his destination. Knowing how to sail the winds does not exclude faith as some people seem to think. It is not possible to avoid the north wind by attempting to steer around, one must pass through.

When the child arrives at the threshold to the land beyond the north wind, he finds it surrounded by a barrier of ice and guarded by North Wind, now more resembling a great figure outside an Egyptian temple. He still recognizes her and walks towards her, having a feeling of being swallowed up in whiteness.

We do not get to know much about this country since it is difficult to talk about, but the author mentions a Durante who also knew it. Nevertheless, from the stories about Diamond's life upon returning to the world, we can see some of the fruits of his journey such as:

Incapable of lying

Not affected by fear

Seeing the good in humans and events

Awareness of an underlying melody to life

Becoming an inspired poet

Intuitively understanding animals and little children

```The cabbies call him God's baby,' she whispered. `He's not right in the head, you know. A tile loose.'"


\hfill

\texttt{Aleksandar Gueric on 2015-10-06 at 05:27 said: }

I should reread Tilak's "Arctic Home". There is a Serbian edition with a very good preface by late Dragoš Kalaji.

As someone who never had acces into any kind of secret, esoteric knowledge or gnosis, my opinions and thoughts on the subject of Hyperborea and other "lost civilizations" remain in a domain of exoteric musings and intuitive pedestrian epistemology.

To me, Hyperborea (Yperboreia) is a Sacred, Polar Center manifested on all levels of existence, from terrestrial, material one all the way up to the Celestial or Cosmic one (Cosmic North of my personal mythos). But, over the ages of spiritual degradation, lower materializations vanished behind the hermetically sealed gates, remaining 'open'. perhaps only to a very few Initiates (who must acend through a pillar of Light because '.neither by ship nor on foot would you find the marvellous road to the assembly of the Hyperboreans’’). The most profane interpretation, the materialistic one, is echoed in fairy tales, existing through such indelicate currents like, for example, Ray Harryhausen's 'Sinbad and the Eye of the Tiger’’. Although, viewed from high esoteric planes, profane and childlish, this mythopoeic capsule in a form of a cheap fantasy movie can plant a seed into child's consciousness which is still very much '.Golden Age’’-like, initiating him/her into a sacred mystery. But in the eye of a spiritually atrophied, mechanized individual, this medium remains an empthy shell. This is why it's important for our children to read mythology. Or watch Ray Harryhausen's movies, or read Conan the Barbarian. Or all of the above.

Hyperborea – Lemuria axis often sparked my imagination because I felt in it a sort of primordial struggle, a 'battle'. between Good and Evil, Light and Darkness. Strangely enough, Harryhausen's Sinbad visits Lemuria in the 'Golden Voyage of Sinbad’, finding an exotic land covered in jungle and quasi Hindu temples and palaces, already aged and ruined and inhabited by primitive aghori looking tribes who practiced human sacrifice to a grotesque Centaur god. And surprise, surprise, another deity worshipped deep in their caverns is – Kali. In our own reality, living in the Kali yuga, seems that Lunar Lemurian influence is on a rise, treatening to flood entire world, including the North, with it's waves of darkness.

Perhaps David Icke isn't completely crazy and his reptilians are simply spiritual Lemurians who became a rulling elite after the removal of what was left of a true European solar nobility? (a la reptiloid Jews, just look at Soros, Barbara Spectre or Nataniel Rotshschild and tell me if they look completely human?) Icke, being unable to interpret the true depth of this mystery, simply tells us instead a materialistic story based on ufo's and alien memes.

There is also a dichotomy, perhaps, between Black Sun and Golden Sun, with the former one having a dominant influence in our world, at least in the Kali yuga. Maybe because it represents a hidden, secret Sun? National Socialists heavily used Black Sun and it seems to me that it's somehow connected with the Antarctic, more than the Arctic, which is very interesting if true.

Theosophists considered South Pole a source of all lethal astral and cosmic forces, while the North is the fountain of all beneficient ones, thus being related to the practices of right hand and left hand path. There is some confusion about what is today LHP, since, living in the 'evil' age, being antinomian would mean being 'good’. In that sense, Black Sun use must be some kind of a 'Ride the Tiger'. practice. I am just guessing.


\hfill


\end{sffamily}\end{footnotesize}
