\section{The Group of Aryan Races}

\begin{quotationx}
In this section from \emph{Sintesi di dottrina della razza}, \textbf{Julius Evola} describes the third migration from the Hyperborean regions. This was the incursion of the Germanic tribes into the Roman Empire. He points out that they were a degenerate remnant, and were unable to continue the civilization of the Roman Empire. It took several centuries for European civilization to re-emerge, beginning with \textbf{Charlemagne}.

Evola then sketches out the various branches that followed from the Hyperborean migrations. As we see, his understanding of ``race'' is more in line with the UNESCO understanding of a race as an ``ethnic group''. As such, the Hyperboreans, or ``Aryans'', include the inhabitants of the Indian sub-continent, ancient Egyptians, Persians, and even Amerindians and Eskimos.

By ``race of the first level'', Evola refers to the race of the body, which is how most people conceive race. Of course, there are also two higher levels: the race of the soul and the race of the spirit. He rejects so-called ``scientific'' racism, since it deals strictly with the race of the body. For Evola, genetics cannot explain the soul and spirit of the various races.

As justification, he refers to the philosopher Emile Boutroux\footnote{\url{https://en.wikipedia.org/wiki/\%C3\%89mile_Boutroux}}, who denies that even the physical world is deterministic. This type of philosophizing is virtually unknown today, but was a dominant perspective among French and Italian philosophers at the time.

\emph{A fortiori}, man, as a free spirit, can hardly be fixed and determined by his body or genetic makeup, despite certain tendencies or habits. Any attempt by ``science'' to ``fix'' a given race to predetermined characteristics or behaviors is ultimately deleterious.
\end{quotationx}

The most recent of all is the emigration of the third voyage which followed the north-south direction. Some Nordic lineages had previously traversed this direction in prehistoric times; they are those, for example, who gave rise to the Doric-Achaean civilization and brought the cult of the Hyperborean Apollo to Greece. The last wave was that of the so-called ``migration of the people'' which happened at the fall of the Roman Empire and corresponds to the race of the Nordic-Germanic type. In this respect, we must make a very important observation. Those races that spread in the north-south direction descend more directly from the Hyperborean lineage who last left the Arctic regions. For that reason, they often display, from the viewpoint of the race of the body, a greater purity and conformity to the original type, having had fewer possibilities of encountering different races. The same however cannot be said from the viewpoint of their inner race and traditions. Their self-maintenance, further from the sister races, in the conditions of a climate which had become particularly harsh and unfavorable could cause in them a certain materialization, a one-sided development of certain physical qualities and especially of character, courage, resistance, tenacity and inventiveness, having, however as its counterpart, an atrophy of the truly spiritual side.

That is seen previously among the Spartans: to a greater extent, however, in the Germanic peoples of the invasions that we may continue to call ``barbaric''; barbaric, however, not in relation to degenerate Roman civilization, in which those people appeared, but in relation to a higher stage, from which those races had fallen by that point. Among the proofs of such an interior degeneration, or spiritual darkening, is the relative facility with which such races converted to Christianity and then to Protestantism. For that reason, the Germanic peoples in the first centuries after the fall of the Western Roman Empire, up to Charlemagne, were unable to oppose anything of importance in the spiritual domain to the crepuscular forms of Romanity. They were enchanted by the exterior splendor of such forms and easily fell victim to Byzantinism, they were unable to reanimate much of Nordic-Aryan existence, in spite of everything, in the Mediterranean world, that by means of a faith corrupted, in more than one aspect, by Semitic-Southern racial influences as soon as they later on gave form to the Holy Roman Empire under the Catholic sign.

Thus even some of the German racial theorists, like Gunther, had to recognize that, while trying to reconstruct the vision of the world and the type of spirituality characteristic of the Nordic race, had to refer less to the testimony included in the traditions of the Germanic peoples of the period of the invasions -- fragmentary testimonies, often altered by foreign or decadent influences in the form of popular superstitions or folklore -- than to the superior spiritual forms characteristic of ancient Rome, ancient Greece, Persia, and India, i.e., of the civilizations derived from the two first waves.

For the combination of the races and traditions generated by these three currents, the first transverse (Ario-Nordic lineage), the next horizontal (Nordic-Atlantic or Nordic-Occidental lineage), the last vertical (the Ario-Germanic lineage) we can apply, not so much for true conformity, but rather on the basis of a usage that has become current, the term ``Aryan''. In order to take into consideration the races defined by the most noted and recognized of racial theories of the first level, we can say that the trunk of the Aryan race, having at its root the primordial Hyperborean race, is differentiated in the following way. There is first of all, as the blond race, the branch called Nordic in the strict sense, that some differentiate into Teutonic-Nordic, Dalic-Falic, and Finno-Nordic sub-branches. The same lineage in its mixing with the aboriginal Sarmatian populations then gave rise to the so-called East-Europid and East-Baltic types. All these human groups, from the point of view of the race of the body, as was noted, preserve great fidelity or purity in respect to what can be presumed to have been the primordial Nordic, or Hyperborean, type.

In the second place, we must consider the races that are most different in relation to the original type, whether in the sense of its phenotype, or form, which the same dispositions and the same hereditary genes caused under the action of a different environment, or of mix-mutations, i.e., mutations produced by more accentuated mixing. It is a matter of predominantly brown types, of smaller stature, in which the dolicephalous is not the rule or is not too pronounced. We mention, using the terminology that is more in vogue, the so-called race of the man of the west, the Atlantic race that, as Fischer defined, is already different from it, then the Mediterranean race, from which, in its turn according to Peters, the variety of the Euro-African man, or African-Mediterranean is distinguished, where the dark component has greater prominence. Sergi's classification, according to whom these last two types more or less coincide, must be rejected and, from the point of view of practical racial theory, especially the Italian, is among the most dangerous. Equally dubious is calling the Mediterranean race, as Peters does, the Pelasgic: in conformity with the meaning that the word had in Greek civilization, it is necessary to consider the Pelasgic type, in a certain way, in itself, especially in terms of the result of a degeneration of some very ancient Atlantic-Aryan stocks established in the Mediterranean before the appearance of the Hellenes. Particularly, from the point of view of the race of the soul, this meaning of the ``Pelasgics'' is justified, among whom the ancient Etruscan people also are included.

In a certain way by itself is the Dinaric race because while it is very close to the Nordic type in certain aspects, in others it shows characteristics in common with the Armenoid and desert race and, like that which some racial theorists define properly as the Alpine or the Vosgic race, is prevalently shown brachycephalous: the sign of crossing that happened along other directions. The Aryan race of the east has, again, distinct characteristics, both physical and mental, through which it is distanced substantially from the Nordic type.

There is nothing contradictory, from the traditional point of view, in assuming in the doctrine of race of the first level the clarifications that the various authors make in relation to the physical and, in part, mental, characteristics of all these branches of Aryan humanity. Only it is not necessary to create too many illusions on the importance of all that, in the sense of establishing rigid limits. Thus, although neither white nor blonde, the superior races of Iran and India, and, although not white, many ancient Egyptian types can moreover be included in the Aryan family. Not only: authors like Wirth and Kadner, who sought to utilize recent studies on blood groups for racial research, were led to believe that some North American Amerindian and Eskimo stocks were closer to the primordial Nordic type, than the greater part of the Indo-European Aryan races just noted. And in this kind of investigations, for example, it indicates moreover that the primordial Nordic blood in Italy has a percentage close to that of England, and certainly higher than that of the Germanic peoples. It is therefore necessary to not be fixated on rigid schemas and to think that, except for rather rare cases, the ``form'' of the original super-race, more or less latent, obstructed, overpowered, or exhausted, subsists in the depths of all these human types. Under certain conditions, it can reappear as predominant and shape a given type from itself, which proves them equivalent, even where it is least suspected, that is, where the previous races according to the schematic and static conception of race, would have instead made the appearance of a type of race, let us suppose, Mediterranean, Indo-Afghani, or Baltic-Oriental, seem probable.

\textbf{Emile Boutroux}, in one of his classic works \textit{On the contingency of the laws of nature} has compared the regularity of phenomena that permits relatively exact scientific predictions, to the course of a river, whose waters follow a riverbed that however it itself hallowed out, so that they can modify it and in exceptional circumstances even abandon it. He considered natural laws as a sort of ``habit'' of things: what originally was also able to be a free act, in repeating itself, automatizes and mechanizes itself and ends up appearing to be a necessity. If that is valid for the laws of so-called inanimate, physical nature, it is of greater validity in respect to races. The proper description of racial theory of the first level, that, in certain authors, leads up to distinctions of an almost scholastic pedantry, fits in the measure to which the forces of races, so to speak, follow a type of law of inertia that automatizes them and fixes them in the state in which, at a given moment, they find themselves being. In these conditions the models drawn for each of the Aryan race fit, determinisms are verified, and sciences observe. At the moment in which the most original forces begin to advance, these constructions demonstrate their relativity and an excessive attention conceded to racial theories of the first level and its ``scientific'' results can even end up bringing damage to the creative and evocative task of race theory.


\flrightit{Posted on 2015-04-15 by Aeneas}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Mark Citadel on 2015-04-15 at 09:47 said:}

Often with these writings on the Hyperborean Race and its various branches, I feel the need to break everything up and draw diagrams. The run-on sentences get a little too hard to follow. Its one of the troubles I had with a few of Evola's chapters. Excuse me while I go and decipher what is no doubt enlightening information that there is no way I comprehended with just one read.

\hfill

\texttt{Aeneas on 2015-04-15 at 12:43 said:}

Please let us know of the most confusing passages. Evola does not write very pretty Italian. We've already had to break up several longer sentences into shorter ones in the translation, but perhaps that was not enough.

\hfill

\texttt{Bill on 2015-04-15 at 13:34 said:}

Wouldn't it be more accurate to say that Evola doesn't reject scientific racism, but recognizes it as subordinate to the higher aspects of race?

\hfill

\texttt{Aeneas on 2015-04-15 at 17:05 said:}

Bill, if you can provide some exact quotes to support your ``more accurate'' understanding, then we can discuss. But first of all, you'll have to explain what you mean by ``subordinate''. You may also want to comment on the paragraph about Emile Boutroux.

\hfill

\texttt{obscure on 2015-04-15 at 20:38 said:}

Let us just say that the distinction between `phenotype' and `genotype' is, from the highest point of view, superficial since both `types' are phenomenal. The gene is just as much a symbol as the facial structure, but it possesses a certain technical value with regard to material analysis and the overall program of techno-science. If one starts from this point of view, the real discourse begins with this question: What is the intelligible essence `behind' the symbol? In other words, how does one really `substantiate' the phenomena?

The materialistic racist errs especially in his individualistic viewpoint. For individuals are always understood within some holistic structure. I couldn't study a race unless I studied them as a nation and not just a nation constituted in terms of `culture' (or what the anthropologists mean by `culture'), but in spiritual terms. Ultimately, this leads to the understanding of a people as a spiritual corporation. If one goes far in this direction then the type of `evolutionary psychology' which amazes the multitude and the cliques will be seen in its full vapidity. But, to take such an intellectual path requires a certain individual decision which is not in the least bit suitable for all.

\hfill

\texttt{Mike on 2015-04-15 at 21:06 said:}

Thank you so much for this.

Since you were kind enough to offer some clarifications, I am having a difficult time with this passage:

``... that proves the corresponding even where it is least suspected, that is, where the antecedent, according to the schematic and static conception of race, would have made instead the appearance of a type of race, let us suppose, Mediterranean, Indo-Afghani, or Baltic-Oriental, seem probable.''

What is the ``corresponding'' and the ``antecedent'' that Evola is referring to?

And so I know I'm on the right track, in the first half of this passage (sentence), Evola is stating that for Aryans, a return to exact Hyperborean form (race of the second and third levels) is impossible to them, since they have all, to varying degrees, parted from the Hyperboreans in terms of race of the first level. Am I correct in this understanding?

\hfill

\texttt{obscure on 2015-04-15 at 22:00 said:}

And to Cologero,

Thank you for publishing all of these items.

\hfill

\texttt{Aeneas on 2015-04-16 at 00:09 said:}

Mike, I tried to clarify that passage a little bit with a re-translation. Curiously, that entire passage is missing from the German translation and replaced with a totally different one.

I believe he is saying the opposite: specifically, the qualities of the original race can reappear, often unexpectedly. We are all trying to pull too much out of one section. Evola will discuss Mendelian genetics (whose scientific validity he doesn't deny) as well the the problems with crossings, and so on.

\hfill

\texttt{Mark Citadel on 2015-04-16 at 14:37 said:}

``They were enchanted by the exterior splendor of such forms and easily fell victim to Byzantinism, they were unable to reanimate much of Nordic-Aryan existence, in spite of everything, in the Mediterranean world, that by means of a faith corrupted, in more than one aspect, by Semitic-Southern racial influences as soon as they later on gave form to the Holy Roman Empire under the Catholic sign.''

You really have to pick this apart to find out what is actually being said. I'm sure everyone appreciates all the effort you put into translation. This is invaluable stuff.

\hfill

\texttt{Max on 2015-04-19 at 13:52 said:}

The Nordic-Germanic wave were unable to continue civilization writes Evola, however their intent was to keep civilization alive in face of roman degeneration and they managed to restore the monarchy with new northern blood. Evola says they descend more directly from the polar center than the remnants of the earlier migrations. They were physically pure, but ``asleep'' from long periods of dark isolation. The material conditions took its toll, a certain heaviness in the inner life of the nordics. They are sometimes prone to materialistic cults of the body or external phenomena, often a type of spiritual lethargy can be seen as counterpart to an outer activity. It is then a question of illuminating the shadow cast on the north.

Evola does not equal Hyperborean to exclusively Indo-european, so it makes one wonder if the Indo-european speakers should actually be regarded as part of the same wave to a degree, or if a classification according to language is irrelevant. In the case they are seen as the third wave spread out over several millennia, the chronology should move back in time to make the origin of the earlier waves prior to the end of the ice-age. The Germanics represent the most northen population and consequently exhibit some tendencies developed to an extreme. One of the characterisitics is a high degree of individuality, whose negative shadow manifested in northern protestantism. Individuality could be seen as a fault or virtue depending on the expression it takes and we should strive to bring out the positive quality in each ``character flaw'' while remedying its negative. This particular flaw actually seems to be highly developed in Evola himself, for better or worse. In a positive expression, rather than just denying authority, it could for example take the form of an inviolable accountability before God above else, the northern ideal of every man being his own King, or a healthy disregard for false authorities.

There is definitely an anarchic tendency in the northern soul, explored by Jünger for example, but it is coupled to a strong social conformity advantagous in sparsely populated areas. The modern world brings out the negative side of both, so what gives rise to a democratic leveling by ``conforming'' should instead be realized in a higher state by humility, a certain concealment in a calm and ordered behaviour and distaste for expressing every whim arising from below. A negative development would be to repress all impulses becoming almost mechanical as is a Germanic stereotype.

Many ethtnicities and physical races contains the ``higher'' race latently, and contrary to earlier periods it does not express itself ``naturally'' and integrally. We have to distinguish various influences acting upon us and uncover some that have been obfuscated. This is a practice in ordering ourselves. We can not simpy accept everything as ``a part of me'' undiscriminately as is the modern habit, but have to consciously explore our ``parts'' and choose the higher over the lower. It means that race is not something we just posess passively but have to acquire through effort and introspection. It would not be surprising if an inner change shows in outer appearance too, not necessarily in hard facts and measurements but as the subtle expression that is impossible to calculate or catalogize. It is said that initiates tend to recognize each other, supposedly not due to hair colour or the shape of the nose though.

I am wondering if the classical civilizations of the ancient world could actually be attributed to a ``recombination'' of different Hyperborean voyages sharing a common past to then reunify after being separated for a long period. In any case, the Germanics were those stubborn enough not to migrate much south, but stay and gradually be surrounded by darkness rather than subject themselves to the violent revolutions of the south; it has to be admitted as some form of asceticism.

\end{sffamily}
\end{footnotesize}

