\section{The Demetrian Race}

\label{sec:DemetrianRace}

The other ``spiritual races'' of the cycle, to which our contemporaries belong, have as their premise a division and a separation of the two elements of ``spirituality'' and ``virility'', and also ``transcendence'' and ``humanity'', that are synthetically united in the solar race. First of all, we point to the lunar race or Demetrian race. As to the analogous relationship, while the solar element is that which has in itself its own light and, in general, its own principle, the sun being, in such regard, the center of a given planetary system, the lunar element is that which instead receives or draws its light and principle from another. The sense of spiritual centrality is therefore lost in the lunar race, whether through degeneration (the moon as the extinguished sun), or through passive mixing with ``telluric'' type races of other cycles, that have degraded its original solar quality. The moon, as Bachofen points out, was also called the ``celestial earth'' by the ancients. There is therefore a sublimation of the law of the earth, the fate that is manifested under the forms of cosmic harmony and law of nature. Lunar man no longer feels himself to be the active center of spiritual reality: he is not even this reality, but rather the one who contemplates it, who studies its laws, who surpasses material action with contemplation and ``tellurism'', but still has not reached spiritual action.

The adjective ``Demetrian'', that we also give to this race, refers to a diffused, pantheistic, less dominating spirituality that is imbued with the sense of cosmic naturalistic laws and by a sacredness placed essentially under the feminine symbol: a spirituality that was precisely typical of the ancient Demetrian cults.

By extension, the priestly man is lunar in comparison to the regal man, he is the man who in the face of the spirit behaves like a normal woman facing a man, i.e., with a sense of submission and devotion. It is then interesting to note that the ancient traditions associated what today would be called cerebrality or intellectuality with the moon, instead associating the heart with the sun and attributing higher forms of knowledge to it. The intellectual is indeed also the lunar type, the man of passive reflection who, as the word indicates, moves only among ``reflections'', among the shadows of ideas and of things.

Thus the lunar race has varied characteristics. In the political field, wherever there is a division between the temporal and spiritual powers, the lunar spirit inevitably arises. The lunar ruler receives the highest consecration of his power from someone else, from a different non-regal priestly caste. In general, the lunar man has spiritually feminine traits. He lacks the feeling of centrality. In correspondence to the races of the body, he has the predominant nature of the Demetrian race which is that of the nations we called Atlantic-Occidental, in their prehistoric forms that led up to the Pelasgic, Minoan-Mycenaean, or Etruscan civilizations and to its further revival, Pythagoreanism. Such a race represents a decay of the hyperborean spirituality which appeared suddenly in the Atlantean regions through processes of action and reaction, causing a series of other changes. Lunar elements can moreover be observed in the race called the man of the East (Alpine-Oriental) by some race theorists---the psycho-anthropology of Claus designates this race as the ``race of evasion''---der Enthebungsmenschen---something that corresponds visibly to a characteristic of the lunar man.


\flrightit{Posted on 2015-06-07 by Aeneas}

