\section{How Not to Save Europe}

\begin{quotex}
This perhaps shows more clearly than anything else that deviation has, so to speak, only to be developed and allowed to pursue its course to the end in order finally to lead to subversion properly so called, for when that which is most inferior seeks to imitate and make a counterfeit of superior and transcendent principles, then is the time when real subversion can justly be spoken of.
\flright{\textsc{Rene Guenon}, \emph{The Reign of Quantity}}
\end{quotex}

From time to time, I get notices from readers about items that they believe are of personal interest to me. They may be books, requests to review articles, or links to web site or talks. Most recently, someone sent me a link to what he called ``New Righters debate Christianity''. The idea density of that description is quite low. I asked for a specific reason to listen to that, but never received a reply.

Obviously, anyone who understands Gornahoor would know that we have nothing in common with the so-called ``New Right'' and we disdain debates. This lack of understanding also demonstrates why it is pointless to go on ad infinitum. There are hard limits to human understanding. Nevertheless, with time running out on the number of posts, I want to make our position clear once again. But first, a few comments on the debate.

I had outlined this piece without listening to the ``debate'' because the topic was poorly framed. However, I subsequently looked up \textbf{Jonas De Geer} and see that he is a serious man, so I intend no disrespect to him. Mr Johnson's point reduces to this: ``my unicorn is prettier than your mule.'' His imaginary racial solidarity, imaginary Europe, and imaginary neo-pagan religion are beyond criticism, unlike something that actually exists, or has actually existed.

There is the obvious confusion of ``Europe'' as the name for a geographic continent; that is merely a conventional (artificial) designation. However, when we go looking for the ``real'' Europe, that is, Europe as it exists in consciousness, as a self-identification, we run into a problem. When \textbf{Hilaire Belloc} claims, ``Europe is the faith and the faith is Europe'', that is what he means. There was a self-identification of oneself as being in Christendom, effectively much of what is now called ``Europe''. Beyond that, there were heathens, i.e., those outside of Europe.

For the Romans, the Empire was the whole world, the Ecumene. Beyond the Ecumene, there were only barbarians, i.e., those outside of civilization. The Roman Empire included a good chunk of Europe, but no one thought of himself as a European, but rather as a ``Roman''. Even the Germanic tribes that overran Rome, considered themselves to be Roman, not European. They formed the ``Holy Roman Empire'' in its various configurations, not the Holy European Empire.

No one answered the question about when the Swedes identified themselves as European. Did the Vikings consider the English and the French as racial brothers when they were terrorizing those nations? The question is risible. They only became part of ``Europe'' after their conversion to Christianity. There are several other quibbles, some of which we mention in passing, both to clear the air and to point out Mr Johnson's confusions: Notre Dame Cathedral in Paris was confiscated by the State and is not under the control of the Church.

Then there are some points about anger. \textbf{St John Chrysostom} pointed out:

\begin{quotex}
He who is not angry, whereas he has cause to be, sins. For unreasonable patience is the hotbed of many vices, it fosters negligence, and incites not only the wicked but even the good to do wrong.
\end{quotex}

\textbf{Thomas Aquinas} defines wrath as ``the spiritual strength to attack the repugnant''. In this case, it is not an irrational emotion set aflame by demons, but rather an inner power that is in the service of the good and the true. \textbf{Peter Chojnowski} in his booklet \textit{Flesh of my Flesh} has a chapter on this topic, particularly as it relates to men in the modern world. Because of the ideals of ``tolerance'' and ``niceness'',

\begin{quotex}
anger is kept from its normal release in the rectification of that which is disordered. When normal release in acts of ordering are forbidden due to a legal and juridical preoccupation with rights and tolerance, you have personal and social explosion waiting to happen. ... to prevent a man from expressing in any way his repulsion to the disorderly, the perverted, the obnoxious, the dishonorable, is to invite and even ensure the engendering of psychological frustration which can only manifest itself in violent rage.
\end{quotex}

So back to the debate itself. Whatever the stated purpose was at the beginning, it quickly devolved to answering this question: ``How can Europe---or ``the White Race'', which is for some reason considered synonymous---be saved. Perhaps that is a burning issue in the confines of that small room, but we are immediately struck by a larger issue. To wit,

\paragraph{Europeans have no desire to be saved}

They are especially opposed to being saved by anything like the New Right, which is considered racist, nationalist, and fascist. Many tens of millions of Europeans were killed in the 20\textsuperscript{th} century to combat those ideologies. Despite occasional outbursts of enthusiasm, European nationalist parties have no traction at the current time. Hence, they refuse to take the actions the New Right considers necessary. Hence, they are forced to embrace the following dubious premise.

\paragraph{Beliefs lead to action}
Specifically, the belief they hold is the reason the Europeans refuse to take a certain action. So in order to provoke the desired action, it is first necessary to hold the belief that would necessitate that action. Hence, the task is changed from convincing Europeans to take action to convincing them to embrace a desired belief, or worldview. But then the New Right encounters another obstacle: the European mind is not a blank slate, but is already filled with beliefs, opinions, and worldviews of various sorts. Hence, Europeans must somehow be convinced that their beliefs are false. Herein lies the next obstacle:

\paragraph{Europeans are happy as they are}

Europeans are prosperous. The top 10 happiest nations\footnote{\url{http://www.telegraph.co.uk/travel/destinations/europe/11560153/Which-is-the-worlds-happiest-country.html}} on earth are European. So how welcome is either Mr Johnson's jeremiad of racial annihilation to the average Joe European?

Clearly, the debate has no point, other than to discredit Christianity. If Traditional Christianity represents ``superior and transcendent principles'', then Mr Johnson is proposing its counterfeit. That is the force of subversion as described by Guenon in the epigraph. Mr Johnson is proposing a new pagan god, the white race, based on no principle higher than biology and no morality other than its promotion. Unfortunately, biology has not been able to provide any stable definition of race. As we recently read, Julius Evola includes many peoples under the ``Aryan'' umbrella, far beyond the immediate confines of Europe.

There are two inconvenient facts that were mentioned in passing, but deserve being emphasized.

\begin{enumerate}

\item When Europe was Catholic, it didn't need saving

\item Now that Europe is \emph{de facto}, if not \emph{de jure}, non-Christian, it does need saving.
\end{enumerate}

For example, during time (1), there would be no issue with birth rates as there is for time (2). Now you can trust and hope that Mr Johnson's future option 3 unicorn will raise birth rates to replacement levels or you can try to identify the real solution.

There was no explanation given to the change in worldview. Mr Johnson's assumption is the post-modern assumption that beliefs and worldviews are arbitrary and that people are malleable, able to adopt and discard them at will. He makes no rational argument for a contemporary European to discard his current liberal worldview and start believing in unicorns. The truth of the matter is the opposite of what is assumed. It is not that a faulty worldview will lead to degeneration, but rather that a degenerate people will adopt a faulty worldview.

Part 2 will reframe the debate in terms of Tradition.


\flrightit{Posted on 2015-04-26 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{X on 2015-04-27 at 02:02 said:}

Great post, Mr. Salvo, haven't read such an article which I fully agree internally for a long time on internet.

``The Roman Empire included a good chunk of Europe, but no one thought of himself as a European, but rather as a `Roman'. Even the Germanic tribes that overran Rome, considered themselves to be Roman, not European. They formed the `Holy Roman Empire' in its various configurations, not the Holy European Empire.''

I specially like this paragraph.

\hfill

\texttt{Max on 2015-04-27 at 08:14 said:}

It is good to see that a few souls still know the meaning of the Golden Cross flying on a clear sky blue background even as others have forgotten a long time ago. Those who came before us recognized this sign appearing in the sky above and made it their own by heart.

To Europeans it can be quite amusing if Americans having abandoned Europe goes on about paganism regardless of having almost no connection to it, and it becomes even more absurd if they come here to lecture us on it. There is nothing wrong in exploring ones history, but this conception is a fantasy. We can understand if people loosing contact with their roots tries to cling on to the most external aspects, but merely being of European descent is no merit. Even a son of the king would have to prove his nobility in action or loose his privileged status. An irony is that in the unrealistic and hypothetical case Europeans returned to some pagan past, Americans would hardly be welcome anymore.

An honourable thing for a pagan would be to challenge his opponent for a duel to determine who has the favour of the gods. Johnson admits that his pagan vision is nothing but modernism with some biological racial awareness added, managing to insult his pre-christian ancestors too. De Geer asserts that a true religion can not be artificially constructed and then throws the challenge: who are willing to die for a belief in ``superior DNA''? So there you have it, neo-pagans, see and learn, but be assured though that the real ``pagans'' of old did not sacrifice themselves for that.

\hfill

\texttt{Iron Man on 2015-04-27 at 08:54 said:}

These debates are indeed quite pointless because of the failed assumption that the West could actually be either Christian or pagan in an industrialist and capitalist society. Tradition and capitalism are wholly irreconcilable! Many of the New Right have pointed this out in their somewhat interesting books (Benoist, for one is worth reading). The battle, if you will, is not about pagan vs. Christian, but a reaction to what exists by that name in the world today. Unfortunately, Capitalism has replaced Salvation and Gnosis. Catholicism and the Vatican have, since WW2, merged with the Anglo-American empire (which is in many ways the new Roman Empire) to wreak havoc throughout the globe for the purpose of world domination by a band of imperialistic oligarchs. The Anglo-American empire is the most nationalistic, fascist, racist, atheistic, and destructive civilization in the history of man (see William Blum's book, Killing Hope), whilst Europe is at its most shameful, complacent, and weak period, existing only as vassals to the new Rome and the holy dollar. But the East is just as corrupted by the capitalist idea, and it is not from the East that the Westerner will be awakened. We have long reached the point (beginning in the early 20th century) when neither the West nor the East can be saved from its folly. Man in its complete and original sense no longer exists, and probably never will again, at least not on a large scale.

\hfill

\texttt{laughingtorch on 2015-04-27 at 10:10 said:}

``Whatever the stated purpose was at the beginning, it quickly devolved to answering this question: ``How can Europe---or ``the White Race'', which is for some reason considered synonymous---be saved. ''

I've tried to have this conversation about races within Europe.

And I agree with X, the article is very spot on. Can't wait to read it again.

\hfill

\texttt{Alistair Fraser on 2015-04-27 at 17:48 said:}

Cologero -- another exceptional piece of commentary in such a lean writing style, i am extracting wisdom from your crafted works , highly appreciated .

``Beliefs lead to action.

Specifically, the belief they hold is the reason the Europeans refuse to take a certain action. So in order to provoke the desired action, it is first necessary to hold the belief that would necessitate that action. Hence, the task is changed from convincing Europeans to take action to convincing them to embrace a desired belief, or worldview. But then the New Right encounters another obstacle: the European mind is not a blank slate, but is already filled with beliefs, opinions, and worldviews of various sorts. Hence, Europeans must somehow be convinced that their beliefs are false. Herein lies the next obstacle: ''

Observing ``europeans'' and others in my own homelands , allows me to make the following observations and confirms why their beliefs do not lead to any actions that may alter the political or philosophical status quo , there are 2 classes , the ones still scrambling on the rat-trap materialistic ladder of monetary achievement from birthright or otherwise , and the ones that are soma-cosed from our new unofficial pharma-religion which has posted its pharmacy outlets in every local neighborhood with delivery vans for those that cant make the trip for their daily doze and most of the uk population are now rattling with pills, eyes wide shut marching into wherever they may be led

\hfill

\texttt{Jacob on 2015-04-27 at 20:12 said:}

I usually don't speak this candidly (and if my comment is divisive just delete it), but After actually reading Benoist recently, I'd say The New Right are not of the right at all. There seems to be no belief in anything objective at all, it was like Foucault for nationalists.

'' It is not that a faulty worldview will lead to degeneration, but rather that a degenerate people will adopt a faulty worldview.''

I've seen this explained in the last few posts (including Evola himself say it), but I'm having a hard time reconciling that to the subtle ruling the dense. I might just be overthinking it and drawing a false dichotomy though.

Looking forward to your next post.

\hfill

\texttt{Lu Dongbin on 2015-04-28 at 15:24 said:}

I agree with most of the post, but if the faith is Europe and Europe is the faith, does that mean that if all native Europeans were ethnically replaced by African Catholics from the Congo, Europe would still be Europe? How about Arab or Chinese Catholics? I wonder if non-Europeans would feel similar, i.e. if all of Korea were ethnically replaced by white Europeans who continued the traditions of Korean Confucianism, Buddhism, and shamanism, do you believe they would think Korea would still be Korea?

Also, does historical change play a role? The pagan Norse Vikings may not have been considered part of Europe by the Christians of the same era, but since Europe now does possess a largely common historical cultural and religious background, can those elements (such as religion) now be changed and the common identity still remain?

\hfill

\texttt{Cologero on 2015-04-28 at 17:22 said:}

Where do we begin, Lu Dongbin? Common identity is more than a landmass and more than DNA. So exactly what is ``common'' or constant over time?

\hfill

\texttt{obscure on 2015-04-28 at 21:18 said:}

Lu Dongbin,

Of course, the form of a nation is a unified set of abstract principles which inform the possible material structures of that nation (the arrangements of entities and their modes in space and time). DNA relations (or anything phenomenally concrete for that matter) exist in the first place due to a certain perpetual obedience to form from one generation to the next. If you prefer, I could use the language of Nietzsche: `Physiological value judgments.' Now, these `judgments' clearly transcend the concrete phenomena which manifests them. Even the totality of phenomena is of no use if one does not go beyond the brute recognition of phenomenal arrangements.in one's thinking. One must, at the very least, have a sense for aesthetics.

The central error in your criticism lies in its grounding in a merely `factual' perspective, but reality is no mere arrangement of facts. It is a real whole-thing. The word `replace' as you are using it does not signify anything. That is to say, these imaginary population swaps express a kind of total non-entity. For example, look at the reality of actual `ethnic replacement' (what a stale term): Does it not simply consist of a certain mode or set of modes which actually signify nothing but general disobedience to form; the negation of principles? We must consider that the temporal procession of things is indeed TEMPORAL and NOT a merely spatial arrangement of facts, a techno-symbolic synthesis of data, a momentarily meaningful agglomeration of `atoms and accidents', etc.

Problems of this order are purely verbal and ought to disappear with the least exercise of serious thinking.

\hfill

\texttt{obscure on 2015-04-28 at 22:01 said:}

Jacob,

``I've seen this explained in the last few posts (including Evola himself say it), but I'm having a hard time reconciling that to the subtle ruling the dense. I might just be overthinking it and drawing a false dichotomy though.''

Possibilities utterly precede any particular manifestation. If a people becomes degenerate in one particular temporal procession then indefinite possibilities of degeneracy already precede, through their mere ontological status, whatever particular mode of degeneracy actually becomes manifest. Manifestation is a certain contraction or restriction on otherwise self-subsisting possibilities.

Even those possibilities suitable for a particular degenerate people, insofar as they are not pure privations, are `above' any particular manifestion. For, the positive side of degeneracy is in actuality simply this: That the lowest orders of activity are pursued in such a way as to exclude recognition of high and noble things. Even `materialism' is a positive worldview, although it necessarily involves and presupposes privation in order to be expressed in its `purest' form.

\hfill

\end{sffamily}
\end{footnotesize}

