\section{A Priori and Ab Initio}

\begin{quotex}
\textbf{Armin Mohler} takes as the basis of the vision of the world the opposition existing between two general conceptions, which he calls \emph{linear} and \emph{cyclic}. According to the former, history is development, novelty, evolution and tends to a final end that justifies it; this is the conception characteristic of the various progressive currents, but Christianity as well insofar as it gravitates toward an ``end of times". The latter conception is based instead of the idea of the ``eternal return", of the repetition of the same forms; that would be the basal view of the ``conservative revolution". In our opinion, the contraposition in these terms is not well formulated. If anything, we should speak of historicism and antihistoricism, of ``civilizations of being" and of ``civilizations of becoming". It is not a question of expecting the return of the same forms, but rather of believing that fundamental values never change, of recognizing a \emph{normative} order containing \emph{a priori} and \emph{ab initio }all the principles, without which a civilization and a normal society are inconceivable. \flright{\textsc{Julius Evola}, A Failed Revolution\footnote{\url{https://gornahoor.wordpress.com/2015/12/06/a-failed-revolution/}}, \emph{Rassegna Italiana}, June 1952} 

\end{quotex}
Looking back at what is now called the German ``conservative revolution" several decades after, Evola calls it the ``failed revolution". He identifies the simplistic arguments—literally, ``chump change" polemics—which do not ``take reality into account". Half a century later, in our time, we see the same chump change formulas, the same logical impasses, being repeated by the various incarnations of the ``right". Specifically, the debates center around specific forms rather than the principles behind those forms. When the discussions break down, then novelties are introduced: there is a ``new" right, an ``alternative" right, and so on; in other words, the focus remains in ``becoming" rather than in ``being". A normative order is never defined; rather it is debated, perhaps with a great deal of intensity, yet without resolution. So what comes first and at the beginning?

\paragraph{A Priori}
First of all, the world of being and the world of becoming must be distinguished. Thus, the first chapter of \emph{Revolt Against the Modern World} begins with this:

\begin{quotex}
In order to understand both the spirit of Tradition and its antithesis, modern civilization, it is necessary to begin with the fundamental doctrine of the two natures. According to this doctrine there is a physical order of things and a metaphysical one; there is a mortal nature and an immortal one; there is the superior realm of being and the inferior realm of becoming. Generally speaking, there is a visible and tangible dimension and, prior to and beyond it, an invisible and intangible dimension that is the support, source, and true life of the former. 

\end{quotex}
\textbf{Rene Guenon} makes a distinction between Being and Existence; the former is the principle of the latter. The same doctrine is formulated in Buddhism as samsara and nirvana. In Taoism, as Heaven and Earth, or Yang and Yin.

Now this is not a metaphysical dualism, or as Evola put it, ``There is dualism and there is dualism." That is, there are not two independent opposing principles, as Manichaeism and other Gnostic religions believed. Rather, the realm of Being is the ``support, source, and true life" of the realm of Becoming. Hence, it is impossible to understand the realm of Becoming without first understanding the realm of Being.

For many, if not most, this comes across as ``life denying". This is more noticeable in the exoteric teaching of this doctrine, which seems to value ``Heaven" more highly than life on earth. That can lead to a certain passivity. Yet, its opposite is to live a life enthralled to the glamour of the world. That life values the phantasmagoria of passing phenomena, which is itself ``life denying."

\paragraph{Ab Initio}
To find out what there was from the beginning, we can turn to the revelation of the Gospels. Mark begins \emph{in media res}. Matthew begins with Abraham and Luke with Adam. However, John begins at the beginning, with the Logos. Since John is the beloved disciple\footnote{\url{https://gornahoor.net/?p=8195}}, this is the best place to start. Hence, the Hermetists adhere to the Church of John\footnote{\url{https://gornahoor.net/?p=16881}}.

Now in the beginning the Logos was with God. Since God is the principle of Being, there can be no distinction between Being and the Logos, or said differently, creativity and order are at the heart of Being. To be clear, Order is not inferior to Being, nor is there an eternal conflict between Order and chaos.

Hence, Evola can affirm that there is a normative order to Being, even without fully accepting the Revelation that grounds that order. The Left\footnote{\url{http://www.gornahoor.net/?p=5709}} is fully aware of that order, since its purpose is to overturn it at every opportunity. The various, sundry, new, alternative, and ersatz Rights, on the other hand, all too often are only dimly aware of that normative order.

There is no point in repeating the lineaments of that order, as it has been done often enough. This is just a reminder of the criteria to use when entering into a discussion or debate:

\begin{enumerate}
\item Is the question related to the realm of Being or of Becoming? 
\item Is the goal to repeat past forms or to manifest eternal values? 
\item Are the values proposed eternal or situational? 
\item Is it a matter of recognizing the normative order or of undermining it? 
\end{enumerate}


\flrightit{Posted on 2015-12-07 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Alistair Fraser on 2015-12-08 at 17:49 said: }

@Cologero = A super article above and also the ascesis linked one, your such a treasure of expression


\hfill

\texttt{Max on 2016-01-13 at 16:19 said: }

Interesting how you say that the left is more aware of the normative order than the right. They may not understand it well, but in acts and the actual outcome of events they always manages to be destructive in one way or another. Someone who just resolved to do everything in the other way as compared to the left would perhaps hit closer to the mark than the various alternative rights, although it is not fail-proof and still passive. 

It requires a lot more to be right than to be left, for the latter it is simply a matter of furthering opposition and division, which anyone can do, while the right requires integration. The advancement of the agenda of the left has generally nothing to do with skill on their part, but is rather opportunistic, without much of an evil grand plan at work as some attempts to see it. If the left appears to act in unison, it is not because they are united, or have a common mind, but because they fall down the same slope.

I think there were an earlier allusion that leftists seems to have some sort of vague idea of the primordial state, if yet flawed, however their actions leads in the opposite direction. Why is there such a dicrepancy between what they say and what they do? It must mean that they have been fooled, that they are inwardly divided, but they do not notice that or avoids it in every way.

In some persons there is a rare drive for coherency, which is among the most beautiful things there is, and it requires a real open mindedness and willingness to abandon the lesser for the greater in a reach for higher and higher horizons.


\end{sffamily}\end{footnotesize}
