\section{Mayday, Mayday}

\begin{quotex}
High-minded words of formal meaning are used to arouse passion, prejudice and sentimentality in favor of disguised real aims. This method cannot serve the truth, yet throughout history and down to the present it is consistently used to deceive people in the interests of the mighty.
\flright{\textsc{James Burnham}, \emph{The Machiavellians}}

The cause of all wars and revolutions --- in a word, of all violence --- is always the same: the negation of hierarchy.
\flright{\textsc{Valentin Tomberg}}

The primary causal agent is always an idea become person, with a will that pursues determinate ends---a cognizant will that has a program to realize, a concrete thought, effective in history. ... The few, in so far as they were the conscience and the will of an epoch, were the agents of history. They recognized the forces that were available, and employed that which was really the only active and effective force available to them---their own will.
\flright{\textsc{Giovanni Gentile}}
\end{quotex}


\paragraph{Being and Becoming}


There two worlds of Being and Becoming, or the Civilizations of Space and Time, have corresponding forms of knowing. For the former, there is the ``supreme knowing'', or \emph{paravidya}, which is direct knowledge of the highest reality. In contrast, \emph{apravidya}, is the lower knowledge of appearances. Although Rene Guenon seems to disdain what he calls ``profane science'', as apravidya it has its proper place, provided it is subordinated to the higher knowledge.

When Being is forgotten, there is only becoming. Knowledge of becoming alone is illusory because the ego superimposes its own interpretation onto the real. A real science will dispel such illusory interpretations in order to penetrate to the real. For practice, we will try to apply such teachings onto the ongoing presidential election in the USA. Therefore, we will focus on political science. Table \ref{tab:SciencesCastes} lists the sciences specific to the four traditional castes.

\begin{table}[h]\small\centering
\begin{tabular}{*{3}{p{.3\textwidth}}}\toprule
\textbf{Caste} & \textbf{Science} &  \\\midrule

Spiritual Authority & Theology & Cosmic and Natural Law \\

Temporal Power & Political Science &
\begin{minipage}[t]{\linewidth}\raggedright

Laws of Power.

Ideal: Justice

Deformation: Lust for power
\end{minipage} \\

Productive Class & Economics &
\begin{minipage}[t]{\linewidth}\raggedright

Laws of Economic Activity

Ideal: prosperity and fairness

Deformation: greed\strut
\end{minipage} \\

Servile Labor & Sociology & 
\begin{minipage}[t]{\linewidth}\raggedright Laws of Social Behavior

Ideal: the higher castes provide the means of salvation, security, and prosperity to the people.

Deformation: Excessive sensation in the forms of unrestrained sex, drugs, and entertainment.
\end{minipage} \\\bottomrule

\end{tabular}
\caption{Science for the castes}
\label{tab:SciencesCastes}
\end{table}

First of all, it is important to note that this schema is not arbitrary, since hierarchy is reflection of the celestial order. Hence, this is the order of Being, eternally present, beyond time; it is the very opposite of a nostalgia for the past. Forgetfulness of being results in judging things from the viewpoint of becoming: hence, instead of a theophany, what are seen in hierarchy are only privations, defects, and contingent historical events.

There is no doubt about the notion of hierarchy, since it stands at the base of all political divisions of mankind. One group, the reactionaries, favors the manifest existent of the hierarchy while the revolutionaries, to one degree or another, seek to overthrow them. Besides the castes, there are relationships between the sexes and the peoples of the world. The latter are not eternal. Of course, there are counterfeit hierarchies that are not based on eternal principles.

\paragraph{Formal and Real Meaning}


We see from the table that man is subjected to laws from the various levels of being. From below, he is under biological laws, chemical, and physical laws. He may even be under demonic laws. From above, he is subject to social, economic, and political forces, not to neglect the natural law.

Now Theology and Political Science have been regarded in the West as the highest of the sciences: theology, for the Medieval schoolmen, and the knowledge of politics as the highest aim of the Greek philosophers. Nevertheless, apart from economics, which is math intensive, the social sciences today tend to attract the least capable students. The result is that political science is underdeveloped with deleterious effects on political discourse.

\textbf{James Burnham} in \emph{The Machiavellians} develops a theory of power based on several recent thinkers influenced by Machiavelli: \textbf{Gaetano Mosca}, \textbf{Georges Sorel}, \textbf{Robert Michels} and \textbf{Vilfredo Pareto}. Burnham makes an important distinction between the formal and real meanings of political rhetoric. The formal meaning is an ideological, idealistic, or moral illusory superimposition that appeals to the sentiments and passions of people. As such, it disguises the real aims, which are those of power.

When factions debate ideological issues, the issue of who exactly is seeking power becomes obscured. Now the aims of the political system are security against outsiders and justice within. Nevertheless, the political leaders are still subjected to the laws of power, since few men are just. Hence, on the one hand the ideal state aims to establish security and justice, while on the other hand, the leaders need to retain power against enemies without and interlopers within. That is why, as we shall see, the philosopher-king does not rule.

The Machiavellians approach political science not as the abstract vision of the ideal state, but rather as the study of how people actually behave when seeking and exercising power. Hence, they try to penetrate the formal meaning of rhetoric to expose the real meaning. Their findings are too rich to explain here, but we can just point out that lying, duplicity, violence, and so on are inevitable concomitants to rule.

Hence, we should not be at all surprised that politicians lie as a matter of course. The strange thing is that the populace cares so little about such lies. Of course, that is likely because the people themselves are liars and consider it quite normal. Then the ideologues actually believe the formal rhetoric and are willing to accept any means they think will lead to the desired ends. Probably, too, many of the politicians are self-deluded and believe in their own formal rhetoric.

Formal rhetoric is often just a sentence, a phrase, or a myth that captures the attention of voters despite its intellectual vacuity. Examples are ``we are a nation of immigrants'', ``black lives matter'', ``women make 79 cents on the dollar'', and so on. There are just enough emotional stirrings that counter any subtleties or exceptions to such generalizations.

The exercise of power depends on force. In the ``advanced'' societies, social force alone may suffice. Next would be financial pressure. Hence, extreme violence, gulags, and so on are seldom required. Nevertheless, the prisons are full, and as we saw in Waco a decade ago, the state is perfectly willing to annihilate a group like the Branch Davidians when its authority is mocked.

So the philosopher king is in a bind. In order to establish justice, he must hold power. In order to retain power he may have to act unjustly.

\paragraph{Levels of Being}


This is why so called ``debates'' about ``the issues'' are so ineffective. I have never seen a talking head on TV suddenly change his opinion after hearing out his opponent. Those living in Maya only grasp their own superimposition, or interpretation, of reality, rather than reality itself.

Idle chatter about this or that movement or ideology, or how one system of thought ``caused'' another. This is ludicrous since a thought in itself has no causal efficacy. An idea gains power only when a person with a will entertains it. So the real, and important, question is why a person would believe one idea rather than another. It is related to the person's level of being, since he cannot understand any idea that comes from a higher level.

So it cannot be liberalism, or conservatives, etc., that is the problem. Rather the world declines due to the degeneration in level of being. Higher things are forgotten.

\paragraph{Bullet Points}


\textbf{Hillary Clinton, the Dominatrix}: this is how she deals with men who stray from the approved program: We came, we saw, he died\footnote{\url{https://www.youtube.com/watch?v=Fgcd1ghag5Y}}.

\textbf{Ted Cruz and Theocracy:} The talking heads know nothing about Evangelicals and their disputes. I recently spoke to a New Yorker, a book editor, quite intelligent and educated; she admitted she had never spoken to an Evangelical. Most evangelicals are of the ``left behind'' type, expecting to escape from the world. Cruz, however is a ``dominionist'' who believes that the world will soon come under Christ's dominion in seven areas (which they call mountains): religion, government, business, media, education, entertainment, and family. He should have been more upfront about the real motivation for his campaign.

\textbf{The Conservative Mind}: The talking heads and politicians all claim to be the ``true conservatives''. I heard one such fellow tonight defend a position that no conservative used to hold. They should start with \textbf{Russell Kirk} and move on from there.

\textbf{Naïve Journalism:} It was revealed that the Obama administration ``played'' the mostly young journalists who were covering the Iran nuclear deal. Of course they are ignorant and even naïve, deluded by Obama's formal, rather than real, meaning. A check of declared college majors by SAT score will show that the poorer students go into journalism and communication, so we should not be surprised. Their elders are not really any more intelligent.

\textbf{Climate change and the rain dance}: Vilfredo Pareto identifies many ``residues'' in human societies; these are cross-cultural constants, although they will be disguised in different forms. One is the idea of controlling the weather through rain dances, rituals, prayers, etc. For the modern man, this manifests as the belief in ``climate change''. Although this may not explain the science of global warming, it certainly explains the rabidly emotional and evangelical attitude of its main proponents.

\textbf{Elizabeth Warren Talks Tough}: The Democratic senator from Massachusetts denounced Trump as ``sexist, racist, and xenophobic''. From the context, we can assume that she regards those as moral defects, although I can't find them in any traditional (read ``authentic'') manual of moral theology. These accusations are intended to destroy a man; in comparison, Trump's many insults are quite tame.

\textbf{Political Exposed Person:} A US corporation is not allowed to do business with a family member or close associate of a Politically Exposed Person. For example, a company like Goldman Sachs, hoping to do business in Rio, could not pay consulting or speaking fees to the wife of Brazil's Defense Minister. A ``quid pro quo'' does not have to be proven. Apparently, it is OK for a foreign entity to bribe a US official. Or at least, no one is seriously discussing it.

\textbf{Well-informed voters}: To understand current events requires a knowledge of history, geology, politics, science, ethics, and so on. That combination is quite rare. The ideal is that the most intelligent and educated should be put in positions of influence

\textbf{The Women of Fox:} Fox news is supposed to keep us informed, but their reporters hardly qualify. Their dress code is interesting: no exposed cleavage at all, but the women are expected to show a lot of thigh. Must be good for ratings.

\textbf{Cuckoo conservatives on Glenn Beck}. I caught a few minutes of Beck interviewing someone last week. The poor fellow was frantic, accusing Trump of being a dictator and inviting assassination attempts. No wonder educated people regard conservatives as nearly insane.

\textbf{Identity politics}: Although politicians pander to various demographic groups, there is little doubt about the results. Racial and ethnic groups vote as a mass, so not more than a few percentage points are actually in play. No one is discussing how Hillary will ``go after'' the white male vote, since she plans to win without them. Just who do they want to run this country?

\textbf{The Ultimate Sacrifice:} This is usually considered giving up one's life for his country. Sometimes, however, the king must be willing to go to Hell for his people. That is the ultimate sacrifice.

\textbf{Personhood:} In the USA, dogs and corporations are regarded as persons, but the fetus is not.

\textbf{Term limits:} The Roman Consuls used to be limited to a term of one year. Term limits would solve many political problems.

\textbf{Jains}: The Jains in India are a fascinating phenomenon. Their spiritual leaders are very strict ascetics, but the ordinary adherents are expected to engage in material pursuits. They are, on the whole, quite successful in business and the professions, despite their minority status. They could serve as the model for disaffected Westerners, since a mass movement of any sort seems unlikely.

\textbf{Guilds and Corporations}: Medieval guilds were the economic foundation of the age. It was a cooperative effort between the owners and the servile class. Workers were protected with a living wage and the guilds would provide insurance to the member's families. An apprentice could travel from place to place and be accepted everywhere. The guilds regulated competition and prices.

These developed in some places to ``corporations'', one for each economic sector. Thus, a corporation is not a private business concern as defined in the USA. Rather, it involves the cooperation of the political power, the productive class, and workers, all working within the parameters set by the spiritual authority.

\textbf{No comment}: Bed, bath, and beyond recently changed the name of their ``votive candles'' to ``filled candles''.

\textbf{Sins that cry out to heaven for vengeance}: These are allegedly the worst sins: Murder (including abortion), sodomy, oppression of the poor, and defrauding workers of their just wages. Two for the Democrats and two for the Republicans.


\flrightit{Posted on 2016-05-10 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Mark Citadel on 2016-05-11 at 17:40 said:}

This was an enjoyable foray into contemporary politics, and I completely agree with your assessment of these various facets. I must confess, I once was a listener to Glenn Beck, though was always more of a Mark Levin fan. Ah, the days when that kind of politics held intense interest for me. Conservatism must die to clear the path for Reaction in the coming vacuum.

On an unrelated note, thanks for linking to my essay on Evola's racial theory. I hope I didn't make any glaring errors.

\hfill

\texttt{Cologero on 2016-05-15 at 07:45 said:}

@Mark Citadel, at a time when disparate voices are all claiming to be the ``true conservative'', it is opportune to re-establish its meaning. To be intellectually vigorous and consistent, it can only mean following these two streams of thought:

\begin{itemize}


\item Edmund Burke, in the English speaking world.

\item Joseph de Maistre, on the continent
\end{itemize}


By this standard, we are all liberals.

\end{sffamily}
\end{footnotesize}

