\section{The Eagle and the Vultures}

Originally published in 1939, \textbf{Massimo Scaligero} never allowed \emph{La Razza di Roma} (The Race of Rome) to be republished. The following selection from that book describes the spiritual basis for the founding of Rome. The following points highlight the gulf that separates the mentality and spirituality of the ancient Romans from moderns in the West today. Here are the main points from the text:

\begin{itemize}

\item Everything in Rome comes from a sacred, supernal sense of life.

\item The rite then discloses a universalistic value assigning celebrations of nature, mystical and esoteric meanings, queries the celestial forces for the choice of location and evocations to demonic powers, the genius of the people, to the chthonic and uranic gods, for the fortune of the city that is going to arise.

\item Suffice it to say that, in the same way that a modern with mechanical operations and procedures becomes master of distance, the rite gives form to matter and organizes its own outer life, so the ancient Roman Empire, through the ritual technique made perfect thanks to the regal-priesthood union that entailed the action of a self-conscious, ``solar'' will, and the body of a dynamic, mediating, ``lunar'' force, established an ascending contact with the magnetic forces of the universe and acted psychically through them.

\item There existed a science of such action: unlike the mechanical sciences placing all men on the same levels (since mechanical means can be operated both by the learned and the ignorant), it required a spiritual dignity that was not for everyone; it demanded the presence of psychical qualities, in a dynamic and exceptional sense, and hence connected to a higher morality that was not at all different from that of the mystic, the priest. two races.

\item In driving the trench, he lifts the plow where he wants the gates, so that it does not touch the ground. Then he raises the walls of the city limits, following the line of the trench, and outside, hugging the walls, dig the trench of the circumvallation.

\item The same force of the rite ensures that the founders are liberated from the ``impiety'' of leaving the land of their forefathers and promotes a new link between soul and matter, between the spirit of the race and the land chosen for the foundation.

\item What is necessary to emphasize is that the divine element constitutes the essential part of the birth of Rome, it is only the religious aspect of the mastery of events, of inevitability, obtained through the possession of transcendent energy that was familiar to the ancient Initiate, as the control and mastery of physical energies is to the contemporary engineer and mechanic.
\end{itemize}

Hence, \emph{pace} the moderns, race is based on spiritual foundations, rather than biological. Thus the twins Romulus and Remus can be said to belong to different races. Rites have a real power, derived from their connection to supernal forces. They have to be done with precision, although the techniques are hidden and reserved to the few.

When men leave their homeland, that is an act of impiety against their ancestors and so it needs to be rectified. The Urbe is defined by its walls.

The complete text is found at The Foundation of Rome\footnote{\url{https://gornahoor.net/?p=14988}}.


\flrightit{Posted on 2016-10-29 by Aeneas}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{- on 2016-10-30 at 20:51 said:}

``When men leave their homeland, that is an act of impiety against their ancestors and so it needs to be rectified.''

Is it then the duty of traditionally oriented white Americans, Australians, Argentines, etc, to move back to their ancestral homelands in Europe? (and conversely the same for Asians, Africans, etc.)

\hfill

\texttt{Cologero on 2016-10-30 at 22:19 said:}

@Hyphen, did you bother to read the text? The complete sentence is:

\begin{quote}\footnotesize
The same force of the rite ensures that the founders are liberated from the ``impiety'' of leaving the land of their forefathers and promotes a new link between soul and matter, between the spirit of the race and the land chosen for the foundation.
\end{quote}

First of all, it was your ancestors who committed the impiety, not you. Second, the solution proposed was not to return, but to perform the rite that would free them from the karma of the impiety.

The result, then, is a ``new link'' between the spirit of the race (of the ancestors) and the new homeland. That could be your ``duty''.

Was there not a necessity involved in leaving the original land? Did not Aeneas have to escape from Troy?

\hfill

\texttt{Max on 2017-01-21 at 09:31 said:}

I came across a commentary on the act of founding taken from ``A Thousand Plateaus'':

``For sublime deeds like the foundation of a city or the fabrication of a golem, one draws a circle, or better yet walks in a circle as in a children's dance, combining rhythmic vowels and consonants that corresponds to the interior forces of creation as to the differentiated parts of an organism. A mistake in speed, rhythm, or harmony would be catastrophic because it would bring back the forces of chaos, destroying both creator and creation.

Finally, one opens in the circle a crack, opens it all the way, lets someone in, calls someone, or else goes out oneself, launches forth. One opens the circle not on the side where the old forces of chaos press against it but in another region, one created by the circle itself. As though the circle tended on its own to open onto a future, as a function of the working forces it shelters. This time, it is in order to join with the forces of the future, cosmic forces. One launches forth, hazards an improvisation. But to improvise is to join with the World, or meld wit it. One ventures from home on the thread of a tune.''

The encircling is a necessary protection against chaos. It creates a new space where the carefully gathered elements can germinate and launch forth in a new tune unfolding through time, ``giving shape and direction to events''.

The current ``new speak'' makes discrimination an evil. The ability to discriminate is actually the result of higher faculties and we even have to do it in order to live and thrive. One cannot love without discriminating. The word Paradise of Iranian origin means Enclosure.

The ancient Greek ``octave species'' are named after kingdoms living according to different rhythms. To tear down all barriers makes music impossible and produces nothing but meaningless noise and eventually death.

Vultures are birds that feed on death:

\url{https://en.wikipedia.org/wiki/Sky\_burial}
The modern environment should at least be apt for learning how to die considering there are no shortages of scavengers around. Breaking with ones ancestors and homeland is like cutting up the body into pieces, dissipating the accumulated forces.


\hfill

\end{sffamily}
\end{footnotesize}

