\section{Odin Sightings in Europe}

I recently heard from a young Swede --- let's call him Loki to protect his identity. Although we had corresponded in the past, in a very superficial way, we never met in person. Despite his attempt to sound calm and controlled, even over the phone I could discern a certain agitation in his voice. Reluctant to discuss any details, he insisted we meet in person. He suggested a gay bar in Wilton Manors since --- as he correctly pointed out --- no one would recognize us there. I was not so sanguine, but nevertheless reluctantly agreed.

As I suspected, Loki informed me that he was a member of a religious cult that has been practicing the old ways without interruption from the dawns of time. But I did not expect what came next. With unconcealed excitement, he told me that recently there have been a significant number of Odin sightings in Sweden! Of course, I enquired about the reason for that.

Loki went on for a couple of hours, but I need to condense it here in order to get to the surprising main point. He told me that the gods have been angry ever since the introduction into their land of the new way of the Nazarene. Even worse, they had become weak due to the paucity of worshipers and the omissions of the sacrifices the god's required. Nevertheless, they worked covertly to overturn the new religion. Initially, they worked through the promotion of the Reformation, then gradually by discrediting the Nazarene.

In the latter case, they made an occult alliance with the traditional enemies of Christianity. This was greatly facilitated by the technological advances of the past 100 years. By means of this alliance, they were then able to promote ideals and values through television, movies, and other media that were antithetical to the teachings and values of the Nazarene sect. They would be promoted as being liberating and ``life affirming.''

Loki continued. About 60 years ago, he said, the gothar (priests) sensed an opportunity. They covertly worked to return wild boar into Sweden, which they were able to use for sacrifices. The volvur (seers, prophetesses) in his group affirmed that the gods were pleased with those sacrifices and were becoming stronger. Nevertheless, the effects were not exactly as they had hoped for.

Although the influence of the Nazarene had certainly waned, new interlopers of a worse type have been invading his land, ravishing the women, consuming resources, and displeasing the gods even more. This led to a rift in their group. They resolved to consult with three experts around the world in order to resolve the impasse, of which they considered me one.

I was quite unprepared for what I heard next! Loki indicated that their volvur were split on the next step. One faction --- the one that Loki supported --- wanted to perform a human sacrifice in order to gain the full favors of the gods. The other faction was opposed --- and he had to spit the words out through his teeth, ``Because they have become squeamish under the influence of the Nazarene!''

Moreover, Loki asserted that only the nine best pagans would be a suitable sacrifice, and he asked me for a list of candidates, due to my many contacts around the world. I tried to dissuade him. ``The Aztecs, at least, sacrificed their enemies, not the best among them.'' Unfazed, he stared at me and asked again for the list. In a moment of malicious weakness, I considered giving him the names of some annoying neo-pagans whom I know; but I quickly realized that would be immoral, since he had asked for the ``best'', and these fellows are certainly not the best.

\paragraph{The Best of the Pagans}

Plato tells us that ultimate reality is the Good, and the world is a reflection of the Ideas, not the whims of the gods. Aristotle tells us that there is One Mover of the world, not different divine factions warring with each other. There are four causes active in the world; animal sacrifices are not a causal factor. The Stoics taught the life of Reason to understand the world, not the consultation of prophetesses. The pursuit of Virtue is what pleases the gods, not human sacrifices.

Then, after a few moments consideration, I assured Loki that the ``best of the pagans'' have already been sacrificed. The world he despises is therefore the expected result of their sacrifice, so any homeopathic remedy of another such sacrifice will amount to nothing.

True prophecy is the ability to discern the likely natural consequences of one's acts. Of course, an evil man wills those consequences deliberately. There are other ways to commit evil. There is gross negligence, which is to act in reckless disregard for the ultimate consequence. A man may act that way for immediate gain, regardless of foreseeable consequences. The perpetrator is still culpable for the results of his negligence. Then, obviously, some people are simply too stupid to understand the long-term consequences of their acts. Or they may have a high time preference that makes them psychologically unable to act in any other way.

\paragraph{Philosophizing with a Hammer}

In the USA, there have been several recent events in which soi-disant conservative thinkers have been prevented from speaking at universities. They then scream for the ``right to free speech'', as though anyone hearing their allegedly ``good ideas'' will immediately embrace then. That puts them in the subordinate position of begging the power holders to ``let'' them speak. That simply exhibits their utter powerlessness in the places that matter.

Ultimately, the real power holders are deliberately willing the world, so there can be no question of negligence. Why would they be interested in ``free speech'' when the ends have already been chosen? In any case, Friedrich Nietzsche exposed the false notion of the ``good idea'', showing it to be a mask, a camouflage, for the ``will to power''. If free speech cannot lead to the truth, then it can lead only to confusion and conflict.

So I suggested to Loki that he seek power, not more sacrifice. To do so, he must paradoxically become more ``convincing''. His gods must retreat, since the conflicts they arouse are the cause, not the solution.

\paragraph{Confabulation}

In a dementing society in decline\footnote{\url{http://charltonteaching.blogspot.com/2013/06/the-dementing-society.html}}, confabulation is used to explain events. When memory is impaired, i.e., a society forgets its source, its nature, its destiny, it fabricates stories to explain events. There then arise the obvious candidates for cause of the ``decline of the West'': the ``Jews'', Christians, the Reformation, etc.; most recently, I've heard that it is the \emph{filioque} clause. Confabulators are ``generally very confident about their recollections, despite contradictory evidence''. In that case, you need a hammer, not an argument.

\paragraph{Choosing One's Own Death}

When Dominique Venner shot himself at Notre Dame\footnote{\url{https://gornahoor.net/?p=6475}} a few years ago, the pagans praised him, even though they seemed to fail to understand the conflict that drove him to it. A man, they said, has the right to choose the time and manner of his own death. If that be true, then why would not a nation, a civilization, have the same right? Then we can stop looking for a confabulated cause, and put the blame on the lack of will. A real priest, a follower of the Nazarene, explains what is happening in several European countries.

\begin{quotex}
Politics is the art of what can be done based on what is. The first reality to take into account is the ``biological'' reality. A country whose population is stagnating, diminishing, or aging, creates a vacuum for younger, more active, poorer peoples. The second reason is a corollary of the first. A country that no longer has children is a country that has lost confidence in itself, its culture, its history and its values. It is plagued with ``cosmopolitanism'' meaning, not so much a generous welcome of others, but rather the stagnation which preludes death. The immigrants sense that, in this depressed country, they can keep their own customs while benefiting from the local wealth, for the natives no longer have a zest for life and camouflage this death wish beneath a false notion of welcome and sharing.
\flright{\textsc{Fr. Gregory Celier, SSPX}, \textit{Angelus Magazine}, May 2017}
\end{quotex}

It is too late to win an argument with Anubis as Death approaches. Instead, you must hope for redemption.


\flrightit{Posted on 2017-05-28 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}

\texttt{James on 2017-05-29 at 10:31 said:}

Thanks for another great post.

\hfill

\texttt{THAT on 2017-06-05 at 19:20 said:}

Don't they even know that the most powerful human sacrifice is ``a male child of perfect innocence and high intelligence'', as Mr. Crowley put it? Big plus if the sacrifice is of one's own blood, preferably one's first-born, as we know from the old pagan Carthaginian nobility and others. Heartless and inhuman, but is our politically correct abortion of innumerable babies today any less barbaric, one has to ask? These are the modern sacrifices that feed demons in the hope for worldly ease.

Entertaining post, by the way.

\hfill

\texttt{White Sharia on 2017-06-05 at 23:09 said:}

Clearly Odin has been angry ever since his peoples forgot the primordial Nordic monotheism.

\hfill

\texttt{Boreas on 2017-06-23 at 03:35 said:}

Odhinn and Christ are not enemies. Odhinn prepared the way for Christ, even if any his or Christ's followers would never understand it.

\hfill

\texttt{Aleksandar on 2017-06-23 at 17:48 said:}

I've always considered Odin an archetype, a state of consciousness, not a deity per se in polytheistic sense, as Thor or Tyr are. Analogies with Christ and echoes of primordial Hyperborean Monotheism were always interesting to me. Not just the crucifixion but also having a title All-Father, appearing as Trinity and giving birth to Baldr, Nordic Christ/Saoshyant and successor to the throne. Unfortunately, neo-pagans refuse to recognize the Tradition in favor of larping. Individuals and pagan gurus like Varg Vikernes do more harm than good, in that sense. Best of pagans were Monotheists, culminating in Plato's metaphysics. No wonder that Plato and Aristotle appear on Christian frescoes in some monasteries.

Then again, maybe these sightings (if genuine and not under the influence of mead) aren't so different from experiencing a ''guardian angel'' or higher self. Years ago in my more heathen youth when I had some interest in runes and Odin (and doing some invocations of Odin), a coin with letters ''Odin'' appeared out of nowhere in my wallet and I was sure it was a ''sign''. Of course, it was only a Russian coin and ''Odin'' simply stood for 1. Still, it was strange to find it in my wallet, far from Russia. Even stranger, same coin disappeared same way it appeared. In other words, seek and you will find what you look for.

Sightings and visions across the Europe are the result of spiritual hunger. But from turning away from Christ, they envision localized forms and empty shells. One Norwegian, after telling him that I'm religious and believe in Christ, said that most of Norwegians would laugh in my face. In best case, some of them would think of Christ as some proto-sjw. But then, they have no problem imagining a red-bearded, hammer wielding, goat riding god.

In my opinion, reintroduction of initiatic, active Christian orders would fix a lot of issues, turning back to the ascetic pathos and warrior-monk tradition. People need activity, action, a purpose, as an expression of faith. Christianity needs a revitalization and restoration or else spiritually dead population will soon disappear from the face of Earth.

\hfill

\texttt{Fred on 2017-06-25 at 04:47 said:}

Odin is a trickster God, one day he bless you to be the greatest warrior, the next he makes you die because you are needed at ragnarok.

\end{sffamily}
\end{footnotesize}

