\section{Mongolian Beliefs}

\begin{quotex}
This is a chapter from The Middle Ages by \textbf{Karl Heyer}.
\end{quotex}

\textbf{Attila} was known to his people as the ``Heavenly Son'' and the Chinese Emperor was worshipped as the ``Son of Heaven''. Out of the same ancient Atlantean mood and attitude to the world Genghis Khan also saw himself as the servant of the ``Ever Blue Sky'', and this ``Ever Blue Sky'', surely, is none other than a representation of the ``Great Spirit''.

\textbf{Genghis Khan} harked back consciously to Atilla, calling him his ancestor and deriving from him the legitimacy for his own rulership claims. Known initially as Temujin, he called in 1206 a great council of all the tribes who were either subject to him or with whom he was allied. There the shaman Gökchu-Teb‐Tengri, the son of another shaman, Muslik, who seventeen years previously had foreseen that Temujin was destined to become Khan, declared that the ``Ever Blue Sky'' had now commanded him to announce to the Mongolians that Temujin had been chosen to rule over all peoples and bear the name Genghis Khan\footnote{The designation ``Mongols of the Blue Heavens'' expressly links the Mongolians with the majesty of the ``Ever blue sky'' or the ``Great Spirit'' whose tool Genghis Khan saw himself to be. The attribute ``blue sky'' also contains other connotations when applied to the Mongolians. It expresses the attitude of soul of these tribes by implying that they bore within them a mood generated by the blue of the heavens, a feeling of reverence towards the infinite spaces of the universe which gather up the human being in the ancient Atlantean sense, making him lose his individuality and putting impulses for action into his passive soul.}

He was acclaimed ruler of the world, the ``ruler blessed by the Spirit of Light'', as one historian called him, a most significant designation in view of the Luciferic grounding of his mission. Perhaps either Munlik or Gökchu was the very priest mentioned by Rudolf Steiner as having provided Genghis Khan with his decisive impetus.

Now Genghis Khan proclaimed that he would lead the Mongolians to rule the world, calling his people the Mongols of the ``Blue of Heaven'', a name that awakened their national pride. Whatever their clan, thenceforth ``those who lived in yurts'' felt uplifted by calling themselves ``Mongolians''. His every deed manifested ``the commands of the sky'' and at the behest of the ``Ever Blue Sky'' he set out to conquer the world.

The life of ancient Atlantis is very evident in the way the Mongolians lived without individuality in an entirely collective consciousness within the ties of blood in family, clan and tribe. It is the same ``compactness'' already mentioned in connection with the Huns, and is diametrically opposed to any kind of western consciousness. The whole of society was founded on kinship.

Genghis Khan established what in modern terms might be called a ``fundamental law of the state'' so that his realm might exist for all eternity. This ``Yassa'' was founded on the tribal feeling and patriarchal social structure of the nomads. Crimes committed against kinspeople and against tribal discipline were heavily punished. This group consciousness lacking in any individuality was obviously the basis for the decadent visionary consciousness which Rudolf Steiner said was to have been spread over the West by the invasions of the Mongolian hordes. One expression of this atavistic consciousness was the Mongolian's picture language which appears to have moved from image to image rather than communicating in concepts.

Another trait of the Mongolians, their proverbial filthiness, fits in with this kind of consciousness. Centuries earlier than Genghis Khan the Chinese had already been calling their neighbours in Mongolia, Manchuria and Turkestan ``the filthy ones'' or ``those who stink''. It was treated with religious reverence and any urge to be clean was severely restricted by ritual prohibitions.

The ``Yassa'' contained the sentence: ``All is pure, nothing that exists is dirty'', and threatened the death penalty for those who bathed in rivers or washed their clothes. Furs had to be worn until they fell to pieces. There is a connection between physical cleanliness and the conceptual clarity of an alert consciousness: an atavistic, dreamy, visionary consciousness is abetted by squalor.

All this points to the mood of soul manifested by the Mongolian races of that time, an eastern mirroring of Atlantean traits closely bound up with the Luciferic element. This was the more inward aspect, and it revealed the essence of the impulse with which they strove to blanket the world.

It is matched, however, by an equally strong other side to their nature, and it is this outer side that is most often described. It might be termed their ``genius for organisation'', which came into play in the way they planned and carried out their wars as well as in the organisation of the state. With the utmost discipline the concentrated power of the whole nation was made to serve the intentions of the central military and political leadership. On a grand scale a military and political strategy was developed which owed its success to that unquestioning and concentrated discipline.

When these two sides are taken together it becomes obvious that the polarity on which the phenomenon of the Mongolian was founded was exceptionally far -- reaching. Luciferic and Ahrimanic impulses complemented one another in a unique union. Instead of being kept apart and suspended in balance by a human middle element they were welded together to form an irresistible force.

This in essence appears to be the phenomenon of Mongolianism, and seen in this light it confirms an experience often met with in history, as well as in life, which is that the strongest attacks of Lucifer and Ahriman, especially in human consciousness, are achieved when these powers work in concert. The successes of the Mongolians were made possible by the precise military and political organisation and discipline with which they served an intoxicating Luciferic idea, and by the inspiring tempest of the Luciferic impulse reaching out from the past for which they provided support in superb military and state organisation.

Despite the terror which these successes still strike into our hearts, we cannot but remain utterly astonished at the power with which they were achieved. Genghis Khan conquered the greatest empire ever known on earth. His Impulse was destructive and its terrible traces are indelibly recorded in history by rivers of blood and millionfold murder.

The Mongolians of Genghis Khan in the narrowest sense were nomadic peoples stemming from Mongolia. In the wider sense, however, the term ``Mongolian'' includes all peoples of Mongolian descent, especially the Chinese and Japanese. These, especially the Chinese, looked back to a very ancient, highly sophisticated culture. Like that of the nomads it, too, harked back to the echoes of Atlantis.

The strongest planetary effect on the Mongolian race as a whole comes from Mars. Just as the Malayans are the Venus race, the Negro peoples the Mercury race, the Red Indians the Saturn race, and the Europeans or Aryans the Jupiter race, so are the Mongolians the Mars race, by which is meant that it originally came into existence when certain Mars beings exercised their influence in the blood stream of these peoples. In this sense Rudolf Steiner characterized the Mongolians as having become hardened in their blood aspect, that aspect that gives external expression to the ego.

Among the cultured Mongolian races, the Mars character of the past is especially apparent in the Japanese. Their noble warrior caste of the Samurai was founded on the ideals and virtues of chivalry culminating in the cult of the sword, of which it used to be said:

\begin{quotex}
The sword is the soul of the Samurai; if it is rusty the soul is tarnished; if it is bright, the soul of the Samurai is also bright.
\end{quotex}

Even the name of the Samurai might be said to echo the Hebrew name of the Archangel who accompanied the age of Mars: Samael.

Above all, however, the Mars character at its most typical is to be found among the Turanian nomads and the Mongolians of Mongolia itself. Fighting amongst each other these nomadic tribes wondered all over Mongolia. `How could they have lived in peace'' we read in one of the newer books about Genghis Khan, ``when pastures were growing ever more scarce and there was nothing beyond the border either''?

Restricting each other's freedom of movement on all sides, the hordes fought incessantly. They were driven to roam, but now they were forced to fight in order to do so. Or it could equally well be said that they roamed in order to fight. War for them was a part of life, indeed it was the very reason for existing. It was the normal state, whereas peace represented a disagreeable crisis. Some of the nomadic tribes of the Urals and the Altai mountains used but one expression for ``to wander'' and ``to cut down with the sword''! It would be hard to find a better description of the archetypal Mars nature of the Mongolians.


\flrightit{Posted on 2022-08-11 by Mu Nianci}

