\section{Land of the Twice-born}

\begin{quotex}
Man has to love the past in order to remember it.
\flright{\textsc{Valentin Tomberg}}
\end{quotex}


Those who see in Gornahoor nothing but an apology for New Age occultism are severely in error. While there are many claiming to expose alleged secret conspiracies, they offer nothing positive. Gornahoor, on the other hand, is revealing to those who can understand the hidden roots of Western civilization. Hermetism, as revealed by Ficino, is nourished by Hermes, Orpheus, Pythagoras, Plato, Plotinus, among others. We have mentioned other sources of Western thought\footnote{\url{https://www.gornahoor.net/?p=556}}. These represent the foundations of Western and European thought. The full history was known as recently as Donoso Cortes\footnote{\url{https://www.gornahoor.net/?p=223}} and Vladimir Solovyov in the 19th century, and later by the recoverers of Tradition\footnote{\url{https://www.gornahoor.net/?p=735}}. It would be to our detriment to forget our past.

Europe was populated by the twice-born, millennia before Jesus' nocturnal visit to Nicodemus where he fully revealed the second birth of water and spirit. This makes Europe a spiritual reality and its true peoples, persons of intelligence, imagination and will. Those who have not experienced the second birth will continue to see nothing in a man but a biological entity who can be manipulated by thought and emotion. But worlds are not created by books, arguments or debate that can only appeal to the head. Rather worldviews\footnote{See Section \ref{sec:VincitOmniaVeritas} in this book.} must be grasped by the heart, the true center of a man's being. That is, a genuine worldview is an inspiration, born of the Spirit with its perfect reflection in the still waters of a people attuned to the transcendent.

\flrightit{Posted on 2010-11-23 by Cologero}

