\section{The Wise Man rules the Elements}

In \emph{Transcendental Magic}, \textbf{Eliphas Levi} quotes from a ``Hebrew manuscript of the 16th century'' regarding the powers and privileges of the Mage. They are grouped in three septenaries, with a conclusion, corresponding to the 22 Hebrew letters of the alphabet.

\paragraph{First Septenary}

\begin{itemize}

\item He beholds God face to face, without dying, and converses familiarly with the seven genii who command the entire celestial army.

\item He is above all griefs and all fears.

\item He reigns with all heaven and is served by all hell.

\item He can neither be surprised by misfortune nor overwhelmed by disasters, nor can he be conquered by his enemies.

\item He knows the reason of the past, present and future.

\item He possesses the secret of the resurrection of the dead and the key of immortality.
\end{itemize}

\paragraph{Second Septenary}

He enjoys the privilege ... \begin{itemize}

\item To find the Philosopher's Stone\footnote{\url{http://en.wikipedia.org/wiki/Philosopher's\_stone}}

\item To possess the Elixer of Life\footnote{\url{http://en.wikipedia.org/wiki/Elixir\_of\_life}}

\item To know the laws of perpetual motion\footnote{\url{http://en.wikipedia.org/wiki/Perpetual\_motion}} and to square the circle\footnote{\url{http://en.wikipedia.org/wiki/Square\_the\_circle}}

\item To transmute base metals into gold

\item To subdue ferocious animals and to paralyse and charm serpents

\item To possess the Ars Notoria\footnote{\url{http://en.wikipedia.org/wiki/Ars\_Notoria}}

\item To speak learnedly on all subjects without preparation or study
\end{itemize}

\paragraph{Third Septenary}

The seven least powers are ... 

\begin{itemize}

\item To know at a glance the deep things of the souls of men and the mysteries of the hearts of women

\item To force Nature to make him free at his pleasure

\item To foresee all future events which do no depend on a superior free will or an indiscernible cause

\item To give the most efficacious consolations and the most wholesome counsels

\item To triumph over adversities

\item To conquer love and hate

\item To have the secret of wealth, to be always its master and never its slave. To enjoy even poverty and never become abject or miserable.
\end{itemize}

\paragraph{Last but not Least}

The Mage rules the elements, stills tempests, cures the diseased by his touch and raises the dead.


\flrightit{Posted on 2010-09-10 by Cologero}

