\section{Gods and Angels}

\begin{quotex}
As a man of science, I cannot be any other than a pantheist; as an artist, it is impossible for me to be other than a polytheist; as for being human—that is also taken care of. \flright{\textsc{Goethe}}

Goethe has a threefold inner attitude towards existence. He was \emph{wholly} a scientist and had thus to see in the whole of nature the revelation of the Deity. He was \emph{wholly} an artist and had to recognize the individual qualities of the Divine. And he was wholly human… Although not directly mentioned, the third element in his spiritual makeup was the ideal of Christ. \flright{\textsc{Valentin Tomberg}, \emph{Inner Development}}

\end{quotex}
There has been a movement of late to promote polytheism as the authentic tradition of the West. This is justified by three arguments:

\begin{itemize}
\item \textbf{Etiology}. Polytheism is allegedly associated with Europeans and monotheism with Semites. 
\item \textbf{Instrumentalism}. Polytheism leads the believer to a certain set of values judged to be superior to values of believers in monotheism. 
\item \textbf{Tradition}. Polytheism is the authentic Tradition of the West. 
\end{itemize}
None of these arguments addresses the question of the truth of polytheism, nor do they define it properly. The proponents of the first two arguments can be said to be invincibly ignorant, since they are naïve about metaphysics and don't fully grasp the significance of their position.

However, the proponents of Tradition are deliberately ignorant, since there is no support in Guenon, nor even in Evola, to justify their position. ``Tradition" does not mean a return to the forms of the past, something that, in any case, is quite impossible.

In his essay \emph{Monotheism and Angelology} (from which we quote below), Rene Guenon exposes the nature of the error that gives rise to polytheism. First of all, 

\begin{quotex}
polytheism consists in admitting the existence of a plurality of totally independent principles, whereas these are secondary aspects of the supreme principle. This lack of understanding is the result of a failure to understand those traditional truths that refer to the divine aspects or attributes.

While individuals may believe this error, in the general case, it reflects the degeneration of a traditional from about to disappear. No tradition whatsoever could ever be polytheist. In particular, any theory of the evolution from polytheism to monotheism [e.g., that of Comte], is the reversal of the normal order of things.

All genuine tradition is essentially monotheistic, that is, it affirms above all the oneness of the Supreme Principle — or more strictly speaking, ``non-duality" — from which everything is derived and on which it entirely depends. Let's be clear that polytheism and monotheism apply to the religious plane, and Principle is the corresponding idea at the metaphysical level.

Nevertheless, monotheism is not incompatible with a multiplicity of divine attribute and even with angels, understood as representing or expressing certain divine attributes within the order of supra-formal manifestation. 

\end{quotex}
Angels exist in all traditions, and the Devas of Hinduism correspond to the angels of the Semitic religions. There is no question of ``borrowing" from one religion to another. ``Angels" are not mere ideas or irrational beliefs.

Angels are supra-formal states of being, and they are actually higher states of Being. Understood in this sense, we could even regard the Norse, Greek and Roman gods and goddesses as angelic beings, provided we look beyond the sentimental portrayal of the angels in Christian art. But in no way are they independent principles; they are valid to the extent that they reflect the attributes of the One principle.

\textbf{CONCLUSION}: The re-emergence of polytheism is not a reflection of a revolt against the modern world, but rather of an acquiescence to it. \emph{It is a symptom of the degeneration of the modern world.}



\flrightit{Posted on 2010-06-27 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Francis A. Mercuri on 2010-06-28 at 00:11 said: }

Excellent review Charles.

To all of this, can be added Guenon's additional arguments contra an absolute ``polytheism" (presented in a highly consolidated manner as follows):
\begin{enumerate}
\item The Metaphysical Absolute as such, although infinite, can not truly be spoken of, since it transcends all limitations. Any attempt at defining the Absolute, immediately generates a limit, and the conceived ``Absolute" ceases to be so. However, it is somewhat feasible to regard the Absolute in terms of containing all ``Possibilities", or ``Total Possibility", which are in tandem, the Possibilities of Being, and the Possibilities of Non-Being. 

\item Being is the Principle of manifestation. As it is the Principle, and not manifestation itself, Being is never manifested, and remains always ``outside" of manifestation. Yet, Being is responsible for manifestation, when it acts in a ``determining", and ``organizing" fashion upon the possibilities situated within Non-Being. But, where then ``is" Being?

\item Non-Being is the Principle (or pole) of non-manifestation, containing within itself obviously, all Possibilities that remain un-manifest, in ``equilibrium". Since Being, although the pole of manifestation, never manifests, as the Principle of manifestation, Being is ultimately ``contained" within Non-Being; all of which returns both Principles to the Absolute, or ``Total Possibility", in an ultimately ``non-dual" manner. 

\item Based on 1,2, and 3, it is ``impossible" that a plurality of equal ``Absolutes" could exist in competition among each other, and therefore any ``polytheism" must ultimately, to avoid contradiction, come to regard its various ``deities", as equivalent to the so-many ``determinations" arranged by Being; and although such ``attributes" might eventually, when retraced, lead back to their origin in ``Total Possibility", to argue that any one of them, limited as they are, by their very constitution, can be equated with ``God", or ``The Absolute", is to blatantly reject, or ignore, any ``notion" of an Absolute as ``metaphysical Infinity".
\end{enumerate}

\hfill


\end{sffamily}\end{footnotesize}
