\section{The Heart of the West}

\label{sec:HeartWest}

\begin{quotex}
Why do most Christians not remember the past? Because they \emph{do not love} the past. One has to \emph{love} the pagan past.
\flright{\textsc{Valentin Tomberg}}
\end{quotex}


\textbf{Marsilio Ficino}, writing in 1471, provided the following list of the founders of the \textbf{Hermetic Tradition}. By building on earlier thinkers, this progression produced a body of profound and sublime thought that has served as the background of Western intellectual right up until modern times. Instead of relying on some arbitrary modernist canon, those seeking to understand or recreate the West as a spiritual entity ought to start with this list. The goal is not to learn what to think but rather how to think, specifically, how to align one's mind with the spirit of the Heart of the West.

There are some points about this list:

\begin{itemize}


\item Egypt is the ultimate source of Western civilization, not India.

\item Scholars may reject the schema because of the lack of a paper trail. This is not unexpected, since Hermetic teachings were passed on orally and in systems of symbols.

\item Extant writings represent a school started by one of the figures, although gathered, edited and modified by later students.

\item There are those who can only see occult or New Age ideas in these thinkers. This only shows how far we have drifted from the intellectual roots of the West.

\item The beginnings of Hermetism was geographically located in what later became the Roman Empire.
\end{itemize}


\paragraph{Hermes Trismegistus}


An Egyptian priest, Hermes was the founder of Hermetic philosophy and science. Early Christians considered Hermes to be a prophet of Christianity. They recognized in him the existence of a Primordial Tradition whose doctrines are incorporated in all true religions.

\paragraph{Orpheus}


He brought music, poetry and literature to the Greeks, all distinctive to the West. He also brought them knowledge of the mystery religions.

\paragraph{Pythagoras}


Pythagoras was a Greek who studied in Egypt and then founded an esoteric school in Greece. Mathematics served as the base for his philosophy. The use of number in esoteric symbolism is characteristic of the Western races.

\paragraph{Philolaus}


Philolaus continued the work of Pythagoras through his use of mathematics and music. He recognized the existence of ideal forms which define the material world.

\paragraph{Plato}


Plato developed the Pythagorean philosophy in great detail. He is the most influential of Western philosophy and Christianity is, as Whitehead famously said, ``Platonism for the masses.'' Despite the great deal of written material available, Plato's most important teachings were given orally in the Academy.

\paragraph{Apollonius}


Apollonius was a Pythagorean philosopher, teacher and miracle worker.

\paragraph{Plotinus}


Plotinus was an Egyptian philosopher who developed the more esoteric aspects of Platonism. Although a pagan, his philosophy was adopted by some of the most important Christian theologians. He spent his life in Rome, thus completing the circle from Egypt to Rome.


\flrightit{Posted on 2010-12-13 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Will on 2010-12-14 at 08:28 said:}

Excellent post. For years, scholars and archeologists dismissed the idea that Hermetism actually came from Egypt, believing instead that either the Greeks or the Renaissance Italians like Ficino simply made it up. Then, someone discovered a piece of Egyptian writing containing the phrase ``Thoth Great Great Great.'' Thoth of course is the Egyptian name for Hermes, and `Thrice-Great' is his title in Hermetism.

While I agree that the West historically goes back to Egypt and not India, it seems to me that Indian doctrine is very important for our future. Our holy language in the West is Greek, and to a lesser extent Latin. The holy language of India is Sanskrit. Beginning in the 1800s or so, it began to dawn on people that these languages all have a common root. To me, this linguistic connection is emblematic of a deeper connection. Also, Guenon seemed to regard the Vedic tradition, up to and including Advaita Vedanta, as the gold standard against which to measure all other traditions or metaphysical systems.

\hfill

\texttt{Ernest on 2010-12-14 at 09:32 said:}

I have begun to compose and work through a Western tradition reading list. The others are on there, but what would you recommend in terms of reading for Orpheus, Philolaus and Apollonius?

\hfill

\texttt{GF on 2010-12-14 at 10:55 said:}

Yes, gentlemen, we have a difficult job ahead of us.

I've been attending Mass; among the religious, let-alone the atheists, the material we have to work with is discouraging. Opinionated, irreverent, and soft, is how I would describe them. They are interested in a `lunar' and democratic ``God is my friend'' spirituality.

I must ask, Cologero, what is your relationship to the Church? Do you attend Mass? -- Novus Ordo or Tridentine? Do you spend time with the other Church-goers? Whenever I go I'm hounded by old ladies, and rather silly teens inviting me to their youth group. I tried going to the men's group, thinking it would be somewhat better -- at least more amenable to serious ideas -- but they weren't. A discussion on how to prepare spiritually for Christmas went very badly; I tried to introduce the notions of rhythmic mental prayer, fasting before mass, and practicing a bit of chastity, but they seemed to feel that actually imitating Christ or the Saints would be somehow arrogant -- their way of saying ``eh, I'd rather not''. They wound up talking about making a small donation to an international charity. The fact that they also recently participated in prayers at a nearby Synagogue left me feeling disgusted. I'm ready to grapple with the real world, and try to heal it, but this feels like a waste of time.

\hfill

\texttt{EXIT on 2010-12-14 at 12:07 said:}

GF, this is exactly what I've been saying. Using Christianity as the basis for a renewed western tradition is good only in theory but in practice you stand alone. Trying to insert metaphysical or spiritual ideas into the average Christian is a practice of smashing one's head against the wall. Their interest only includes petty doctrinarian moralism and sentimental historical absurdities. The Church, which is at the heart of the corruption of Christianity, dangles Jesus before the masses and then demands that no one become like him spiritually, only morally, while at the same time, the Vatican displays no such morals itself, molesting children and laundering money for the mafia and crooked businessmen.

As for the hermetic tradition, it all comes down to accessibility. Since there is no order present today which has preserved a valid chain of hermetic initiation (modern occultist groups don't count) then how can one carry on a tradition which no longer exists? So far, all those who talk about this love for the past have never put forth a solution other than some Evolian mysticism which excludes initiation and has gone absolutely nowhere.

\hfill

\texttt{Cologero on 2010-12-14 at 12:40 said:}

It's gone nowhere if we only look at quantity.

\hfill

\texttt{Cologero on 2010-12-14 at 12:47 said:}

For Orpheus, start with Fabre d'Olivet, in particular, the ``Hermeneutic Interpretation of the Origin of the Social State of Man''. There is an ancient \textit{biography of Appolonius}\footnote{\url{http://en.wikipedia.org/wiki/Life\_of\_Apollonius\_of\_Tyana}} available.

\hfill

\texttt{Cologero on 2010-12-14 at 13:05 said:}

Welcome to the Kali Yuga, GF, no one is going to do it for you. Religion is now reduced to moralism and sentimentalism, as Guenon pointed out. There are two resources available for Christian Hermetists.

One is Vladimir Solovyov's ``Lectures on Divine Humanity'', which is a metaphysical reconstruction of Christian dogma. In an intriguing footnote he writes:

\begin{quotex}\footnotesize
Although the close inner connection between Alexandrian theosophy and the Christian doctrine is one of the firmly estalished theses of Western scholarship, for one reason or another, this perfectly correct thesis does not enjoy common acknowledgment in our theological literature. Therefore, I consider it necessary to devote to this question a special appendix, where I will touch upon the significance of the native Egyptian theosophy (the revelations of Hermes Trismegistus) ... 
\end{quotex}


Unfortunately, that appendix never appeared. That is why Gornahoor is taking up this task.

The other resource is Valentin Tomberg's ``Meditations on the Tarot'' which, despite its faults, is a worthy attempt to reintegrate the Hermetic tradition into Church theology and practice. It is good for its spiritual exercises and its radical reinterpretation of Christian concepts; it is weaker in its understanding of metaphysics.

Both these books will gain traction and attention, but will never become the majority opinion. There are good and necessary reasons for this.

As far as participation in the Mass, Tomberg considers the sacraments to be valid and effective (at least for the Tridentine mass). But since that is in the realm of mystery, it cannot be proved or disproved on the Internet. That remains your personal decision.

\hfill

\texttt{Cologero on 2010-12-14 at 13:20 said:}

Perhaps the Vedanta is the gold standard for metaphysics, but that is like complaining that your neighbour's wife is more attractive than your own.

\hfill

\texttt{Will on 2010-12-14 at 17:31 said:}

Cologero,

The notion that Christianity is ``ours'' (the West's) and that the Vedic tradition is ``theirs'' (India's) is somewhat relative. Two thousand years ago, paganism was ours and Christianity was someone else's. But things change over time. Also, if we accept the idea of a Primordial Tradition, or Ficino's notion of a Prisca Theologia, then what is important is the shining through of primordial metaphysical truth. I'm not saying that Westerners shouldn't be Christians, or that we should become Buddhists or Hindus, but the fact is that Christianity is dying demographically while Eastern traditions are growing here. There are a lot of problems with this, which I've written about on this site, but it is happening nonetheless. I don't know what this will mean for the future, but it seems to me one possibility is that these Eastern traditions will become Western in a manner similar to the way Christianity became Western.

\hfill

\texttt{Cologero on 2010-12-14 at 18:35 said:}

Unfortunately, Will, you are so eager to make your point that you miss the point. In a post titled ``The Heart of the West'', I referred to seven metaphysicians, none of whom were Christian. Nevertheless, you somehow conclude that the point is that Christianity is ``ours''.

The real point --- and I do hope some of our readers got it --- is that the West does indeed have a ``prisca theologia''. If anyone chooses to look elsewhere, go right ahead.

\hfill

\texttt{Will on 2010-12-14 at 19:08 said:}

Mea culpa. I was reading all the comments about Catholicism and confused the post with the comments.

\hfill

\end{sffamily}
\end{footnotesize}

