\section{And God Created Woman}

I've been reading \textit{The Doctrines of Genesis 1-11: A Compendium and Defense of Traditional Catholic Theology on Origins}\footnote{\url{http://www.iuniverse.com/bookstore/BookDetail.aspx?BookId=SKU-000090351}} by Rev. Victor Warkulwiz, a Catholic priest with a Ph.D. in physics. His intent is to outline a Catholic theology of creation based on a literal reading of Genesis. Since the Church has never dogmatically defined that interpretation, it is unclear why this book has come out now. Fr. Wakulwiz's method is to create a reasonable doubt in the currently accepted scientific cosmological narrative. While we can agree that the totalizing claim of science to explain everything is unjustified, within more narrow frames of reference it has great explanatory power. To be credible, those who propose a fundamentalist interpretation of Genesis will eventually need to move beyond negative critique and create a positive scientific theory that accords with such a reading. Similarly, they would need to reconcile the Biblical history with what we know of recorded history of other peoples.

Nevertheless, his exposition does bring up many issues of metaphysical and esoteric interest on sexuality, race, and the nature of the soul, all of which we hope to explore in a series of posts. Unfortunately, in these areas he falls short; blinded by science, he fails to follow up on these metaphysical issues which are much more fruitful and relevant than the literal reading.

\paragraph{Adam's Rib}
Adam went into a deep sleep before Eve was created from his rib. Some of the Fathers considered this a form of anesthesia to prevent his feeling pain. This is unsatisfying, especially considering that in the primordial state the subjective experience of pain is not possible. It may be better to regard this sleep as a state of unconsciousness, so that we can see Eve as a projection of Adam from the subconscious. St. Ephrem of Syria is close to this idea when he writes:

\begin{quotex}
Moses said, `Male and female he created them,' to make known that Eve was already inside Adam... she was not only in his body with him but also in soul and spirit with him.
\end{quotex}

Fr. Warkulwiz makes a reasonable inference that all of Eve's chromosomes were received from Adam. Hence, they were identical twins except for the Y chromosome; Eve received a duplicate of Adam's X chromosome. Conversely, it would not have been possible to create Adam from Eve. This scenario shows that man is sui generis, not descended from animals. Otherwise, Adam and Eve would be part of the mutated litters of some other animal species. Without necessarily accepting the biological explanation of the event, we can summarize some esoteric principles:

\begin{itemize}

\item Adam was created in a state of wholeness

\item He awakens from a deep state of sleep to discover Eve

\item Eve is his near identical twin, in body, soul, and spirit
\end{itemize}

\paragraph{Mysterium Coniunctionis}

Obviously this brings up the idea of the Mysterium Coniunctionis, which is far from the concern of the book. Nevertheless, this explanation provides a fruitful and theologically correct background for the commentaries of the alchemists, \textbf{Plato}, or \textbf{Carl Jung}, on soul mates or mystical marriage. Even \textbf{Thomas Aquinas} is alleged to have stopped work on the \emph{Summa} to complete an alchemical text on this very topic (Aurora Consurgens). The uniqueness of the soulmate is consistent with the revelation of the esoteric doctrine of `Polar Beings' by Boris Mouravieff. There is even a neo-Christian preacher named Robert Thieme\footnote{\url{https://www.goodreads.com/book/show/15822361-right-man-right-woman}} who drew out this doctrine, which he called `Right man, right woman' from the Genesis account.

\paragraph{Image and Privation}

So we see a series of transformations. Initially, Adam is created in the image and likeness of God, that is, God is reflected in Adam. Then Eve is created from Adam, and Adam is reflected in Eve. Finally, the ideal creation is re-created by Adam-Eve as a result of their deformed consciousness.

This is a process of successive privations. There is the intentional privation\footnote{\url{https://gornahoor.net/?p=17095}} of God, to make a place for Adam. Then there is the privation of Adam to create Eve. Finally, the world is experienced as privation. Initially, the world was under the dominion of Adam, but with the Fall, that dominion comes only with effort, toil, and pain\footnote{\url{https://gornahoor.net/?p=1374}}.

\url{https://www.youtube.com/watch?v=iGNlIyEGKPg}

\flrightit{Posted on 2011-07-11 by Cologero}

