\section{Innate Wisdom and Collective Initiation}

Confucius taught that there are two kinds of sage, those who are so by birth and those who are so through their own efforts. The Sage or True Man is the highest stage of the exoteric path, but the first stage of the esoteric path.

\begin{table}[h]\centering\small
\begin{tabular}{ccc}\toprule
\textbf{Exoteric Path} & \textbf{Stage} & \textbf{Esoteric Path} \\\midrule
Sage & Stage 3 & Man of Tao \\

Man of Learning & Stage 2 & Man of Talent \\

Man of Letters & Stage 1 & Sage \\\bottomrule
\end{tabular}
\end{table}

\paragraph{Natural Sage}

What is of interest to us here, in a time and place where true initiation is difficult if not impossible to attain, is the case of the man born a Sage. Such a one does not need an outer Guru, as he has an ``inner Guru''. His realization may be only virtual and not actual, but contact with certain ideas may awaken him to the remembrance of his true identity.

\paragraph{Collective Initiation}

In group work, the group acts as the \emph{upaguru}, that is, a catalyst that can activate the inner Guru. By observing and understanding not only oneself, but also the others in the group, the participant comes to understand how to actualize all his possibilities. If he is unable to see directly some such possibilities in himself, he can nevertheless see them in his fellow initiates; this will help him recognize those qualities in himself.

\paragraph{Exoteric}

The method is to recapitulate the stages of initiation. To be a \textbf{man of letters} means to have read broadly and experienced much.
Such a man is familiar with literature, history, science, maths, and philosophy. To be a \textbf{man of learning} is to know this material in depth. Thus we would engage in a careful study the metaphysical works of Guenon. This will be the intellectual preparation required to understand what is going on in the inner work.

\paragraph{Esoteric}

This learning is done simultaneously with group and independent spiritual work. We are only concerned with the group work. As this is experiential, it really cannot be described publicly. However, the goal is Hermetic initiation, being born again of the Spirit and Water.

\paragraph{Activities}

It is hoped that an organized body of knowledge will come out of this work. Perhaps a wiki of topics covered in the exoteric studies. A definite requirement is the creation of symbols and diagrams based on the abstract geometrical descriptions used by Guenon. Longer term, members should go out an engage the world as philosophers, artists, politicians, businessmen, and all other gainful employment.

\paragraph{Obstacles}

As Guenon point out, in our age there will be obstacles to any attainment. But bear in mind that we ourselves have chosen these obstacles and hence the means to overcome them. I had hoped to be able to have a weekly group discussion of the metaphysical books. However, those interested are in widely disparate time zones, making any choice of a time inconvenient for some. We will need two separate discussion groups.

References (all by Rene Guenon):

\textit{Innate Wisdom \& Acquired Wisdom}

\textit{Guru \& Upaguru}

\textit{True Man and Transcendent Man}


\flrightit{Posted on 2011-02-19 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}

\texttt{Matthew Anderson on 2015-05-22 at 14:14 said:}

Interesting. How important is it to widen the base or breadth of knowledge? I would assume that a degree of breadth is necessary to balance depth. Breadth without depth would be meaningless, like being a human encyclopedia or a robot. But what about depth without breadth? I suppose it wouldn't be very useful. And how does height differ from depth? (or does it?)

\hfill

\texttt{James Yen on 2019-03-22 at 15:03 said:}

You may be interested in Chögyam Trungpa's ideas of the ``blessed'' and ``direct'' tulku. Reginald ``Reggie'' A. Ray also elucidates some of Trungpa's ideas in his books. Basically a ``blessed tulku'' is a bodhisattva with some residual amount of progress which an environment catalyzes in him in his next birth, he then continues his practice. A ``direct tulku'' is someone who is born enlightened, essentially.


\hfill

\end{sffamily}
\end{footnotesize}

