\section{Gnosis, properly so called}

\begin{quotex}
What is now called the Christian religion already existed among the ancients, and was not lacking at the very beginnings of the human race. When Christ appeared in the flesh, the true religion already in existence received the name of Christian.
\flright{\textsc{Augustine of Hippo}}
\end{quotex}

Augustine was involved both in Gnosticism and Paganism before his final conversion to Christianity, so there is reason to consider his claim with some seriousness. However, to do so will require a revaluation of the three ancient movements; this is an attempt to make that plausible.

It is known that the early Christians divided themselves into three levels: the Catechumens, the Faithful, and the Perfect. The Catechumens were not yet baptized; after that, they joined the ranks of the Faithful, or believers. The Perfect were the knowers. This internal division corresponds to the external division into pagans, Christians, and Gnostics as shown in the following schema.

\begin{table}[h]\small\centering
\begin{tabular}{ccc}\toprule
Internal & External & Mode of knowing \\\midrule
Catechumen & Pagan & Reason \\

Faithful & Christian & Faith \\

Perfect & Gnostic & Wisdom \\\bottomrule
\end{tabular}
\end{table}
We know of the existence of the third level within Christianity through the writings of the Alexandrians, especially Clement. There are those who deny the existence of a separate esoteric doctrine within Christianity. In a sense this is correct, since the doctrine is the same; nevertheless, the understanding can be much different. There used to be four levels of understanding of biblical writings, of which the historical was the simplest and least important. However, in our time the obsession with a literal and historical interpretation of the Bible --- which is common among Fundamentalists, Modernists, and even anti-Christian --- occludes that earlier understanding. I read a book a few years ago that attempted to discredit Christianity by pointing out its symbolic nature and ``borrowings'' from pagan teachings; this is risible, since it does not refute, bur instead, demonstrates our point.

Very little is known today about the pagan mystery cults. We can assume that many of them were simply drinking clubs, as is the case with many ``secret societies'' today. Although it frustrates the professors, the actual teachings of the authentic cults are irrelevant. We know that they would have taught the Logos and cosmic order, the meaning of gnosis, and the means to attain it through the three trials. \textbf{Rudolf Steiner} in \textit{Christianity as Mystical Fact}, claims that the pagan mysteries were absorbed into esoteric Christianity, and even uses Augustine to make his point. He writes:

\begin{quotex}
The \emph{cognition} of the eternal which the ancient Mysteries withheld from the multitudes, when presented in this way by Christian thought and feeling, became an \emph{article of faith} which by its very nature was related to something unattainable by mere knowledge.
\end{quotex}

Thus, what the knower knows is presented as articles of faith, or dogmas, to be believed by those neither ready for nor capable of gnosis. \textbf{Julius Evola} alludes to that in The Hermetic Tradition, when he claims that if faith is lived intensely and coherently, then it can create the conditions for liberation at the moment of death.

Now Evola also writes of ``the derivation of the hermetic tradition from a primordial royal and heroic vein'' and its ``concealment'' within Christianity. Obviously concealment means nothing other than esoteric.

So the question arises as to the fate of the various movements that we categorize as Gnosticism and why they had to be suppressed. Arising also in Alexandria with many points in common with the genuine Gnosticism, many problems developed. As we see in the various ``esoteric orders'' of today, there is more interest in the external trappings of the grades rather than in the internal transformations that are the real meanings thereof. Hence, there were many who would claim to rank among the perfect or the knowers without having the necessary qualifications. They would misunderstand and misinterpret the genuine teachings and apply them in rather strange ways.

In effect, rather than an esoteric order within society, they created competing ideologies and belief systems. As we have mentioned repeatedly, there is no way to resolve and harmonize belief systems. Thus, a belief system based on genuine gnosis needs to be protected. This is the intent of the adage that the Sage does not disturb the peace of the faithful whereas the so-called Gnostics did indeed do so.

Thus, the Gnostics became a revolutionary force. The net result was that gnosis was separated from faith. The religious and political leaders seldom rose above the initial impetus of the faith doctrine, and the genuine gnostics were lumped together with the more bizarre elements within the various Gnostic movements, and had either to continue in obscurity or else to write using heavily symbolic language.

\paragraph{The Philosopher’s Heaven}

\begin{quotex}
Man, once awakened to a will to create, wants to bypass the limits of the circles of necessity and master the power residing in Fire.
\flright{\textit{Corpus Hermeticum}}
\end{quotex}


In Chapter 12 of The Hermetic Tradition, Julius Evola interprets the statement in the Zohar:

``The visible is the reflection of the invisible'' to mean ``Earth is the visible part of Heaven''. Yet, as \textbf{Jacques Lacarriere} writes in The Gnostics:

\begin{quotex}
Man, like the universe, is a failed creation, a lamentable imitation, the mere semblance of a man, a counterfeit man.
\end{quotex}


Although the Gnostics adapted much of the esoteric teachings of the early centuries of our era, they made a subtle error. The Marxist Lacarriere expresses it:

\begin{quotex}
It seems a simple, obvious, and irrefutable fact that today, as in the time of the Gnostics, the alienation of man is global; it is also true that the economic, social, and political causes of alienation should be removed first.
\end{quotex}


Thus, he could recommend ``modifying the means of production, transforming the nature of economic exchanges, and the distribution of wealth'' as the task of the Gnostic, falsely so-called.

This puts him firmly on the side of revolution and we see this form of spirituality resuscitated in our own time in the various New Age movements. Instead of the futile and erroneous Gnostic path of liberation by throwing off any and all external and conventional structures, Evola quotes \textbf{Jacob Boehme}:

\begin{quotex}
Paradise is still in this world, but man is very far from it, so long as he fails to regenerate himself.
\end{quotex}


Evola explains that it is ``the power of Earth in man is that which forces on him, via the physical body, the materialistic vision of the world.''

This ``petrification of the spiritual world created by the bodily senses'' is conditioned by the ``dualistic law of I-Not-I'' which is ``the main obstacle to modern understanding of the traditional sciences.'' Evola writes that the solution is to become aware of the Air, Water, and Fire elements which ``are other existential states, other modalities of consciousness, quite separate from the body, that can transpose all the principles of things according to their non-corporeal nature.'' In contrast the Gnostics believed:

\begin{quotex}
That which weighs us down, makes us heavy, and sends us to sleep, is the cloacal matrix, this borborian matter from which we were extracted; above all, it is this basic flaw in our very structures that renders us incapable of assuming our predestined mission.
\end{quotex}


In opposition to Evola, the Gnostics took a realist position, postulating that there is some entity that created \emph{flaws} in our structure. Evola, on the other hand, sees matter, oppressive structures, and so on, as \emph{privations}, created by ourselves spontaneously and unconsciously. Our task then is to elevate our consciousness and develop the Will to overcome such privations.

Unable to transcend the I-not-I perception, the Gnostics postulated Yahweh as an ``evil'' demiurge, opposed to the True, Transcendent God. Of course, this is a metaphysical impossibility. The perceived evil is a privation, in this case voluntary\footnote{\url{https://www.meditationsonthetarot.com/the-void-and-the-fullness-are-indissolubly-bound}}. The Traditional esoteric teaching, too, recognizes in Yahweh a being akin to the Demiurge, although with the assigned task of maintaining the laws and structures of the world.


Further reading:

\textsc{Jacques Lacarriere}, \textit{The Gnostics}

\textsc{Julius Evola}, \textit{The Hermetic Tradition}

\textsc{George Heart}, \textit{Christianity: Dogmatic Faith \& Gnostic Vivifying Knowledge}


\flrightit{Posted on 2011-05-05 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{James O'Meara on 2011-05-06 at 11:52 said:}

``Very little is known today about the pagan mystery cults. We can assume that many of them were simply drinking clubs, as is the case with many `secret societies' today. ''

Well, some may have been `simply' drinking clubs, but the drinking club as such was an important part of Classical culture viz, Symposium as well as the origin of the Christian `love feasts' and churches see the recent book From Symposium to Eucharist .

Michael Hoffman has a collection of material, including both academic sources as well as his own speculations, here:

\url{http://www.egodeath.com/WineAndSacredMeals.htm}

\hfill

\texttt{James O'Meara on 2011-05-06 at 11:59 said:}

``So the question arises as to the fate of the various movements that we categorize as Gnosticism and why they had to be suppressed.''

The whole point of having freedom of thought is that some bad apples will appear. Why assume that ALL Gnostic movements ``had to be suppressed'' or that the orthodox were the blameless ones who ``had to'' do it. I prefer Freke and Gandy's theory: the NT is the exoteric, psychic story given the catechumens, Gnosticism represented the inner, pnueumatic doctrine. The outer churches such as Rome were cut off from the inner members in Jerusalem by the revolt and diaspora, leaving them in ignorance of the real doctrine. Rising to power in Rome, then set about exterminating the ``heretical'' Gnostics Pope vs. Eckhardt writ large . End of story.

\hfill

\texttt{Cologero on 2011-05-07 at 00:09 said:}

I thought that is what we wrote, James, and I apologize for not making it clear. However, we find the ``inner members'' in Alexandria, not Jerusalem. The Gnosticisms have not been totally repressed and can be found again at New Age gatherings.

\hfill

\texttt{logres on 2011-05-16 at 23:52 said:}

Mr. O'Meara, the division of the Church into Gnostics \& Orthodox probably did violence to both parties. Arguably more to the ecclesiastical party innerly, and arguably more to the esotericists in the physical/outer sense. Wouldn't (ideally) the two co-exist without being identical?

\hfill

\texttt{James O'Meara on 2011-05-17 at 13:16 said:}

logres,

Ideally, yes. As Guenon points out though it seems to have escaped H. Smith and the other Schuonians the exo/eso division is exclusively Judeo-Christian. In China there are two distinct traditions, with different aims; in India, a sort of `natural' exotericism, where anything can be heard, but only the `right' people can understand. In these lands, metaphysics operates freely, without interference from social conformity and `the good of the state.' Even in Islam, Guenon insisted that Sufism was `Islamic esotericism' not `Esoteric Islam'; metaphysics using the local religion's dress, rather than the pretense of being its `inner level' and subsequent persecution.

\hfill

\texttt{logres on 2011-05-18 at 21:56 said:}

It seems to me that the exoteric Church wants (or thinks it wants) esotericism ``within'' it, defined as safely within proscribed boundaries and subject to scrutiny, which (of course) would mean it not being esoteric by nature (properly) at all. The aim itself (or motive) is correct, but the standard and goal are delusionary -- hence the disconnect and misunderstandings. Many like myself were initially confused and wanted the same thing, except with a proper goal added. I am convinced now that I am ignorant of anything but what it should ``not'' be, and have a general feeling for regrouping and a direction or tendency. These discussions have been very helpful.

\hfill

\texttt{Count Cagliostro on 2011-05-12 at 18:47 said:}

``Earth is the visible part of Heaven``

Sometimes difficult to believe in, but still a wonderful quote!

\hfill

\texttt{logres on 2011-05-14 at 16:39 said:}

Thank you, this actually helps -- the ``Forbidden Religion'' tract by J.A. Aragon is really well written and done, and very convincing rhetorically, but that is precisely why I need such metaphysical grounding. Without that, Gnosticism is the ultimate ``authenticity'' and probably the growth of that movement in the past was due to precisely such shaky premises -- namely, that of a decaying Semitic worldview influenced by popularized mystery religion.

\hfill

\texttt{Matt on 2011-05-30 at 17:40 said:}

I have been reading Mouravieff's Gnosis and relating this post to that writing, I suppose one could say that the various Gnostic groups failed to see what Moravieff noted (assuming I read it correctly) that the laws of the cosmos were in place for the legitimate reason of putting a limit on the destructive aspect of Time, to ensure that the creation endures (while the initate works to transcend the laws, not destroy them), whereas the Gnostics saw them as inherently evil.

\hfill

\end{sffamily}
\end{footnotesize}

