\section{Theosophy and Gnosis}

A chapter in the book Freedom and the Spirit by \textbf{Nicolas Berdyaev} deals with the Theosophical movements vis-à-vis authentic gnosis. \textbf{Valentin Tomberg} writes that Berdyaev is one of those who ``show in their works a progress which is very advanced in substantially bringing together intelligence and the intuition of faith.'' Berdyaev was also respected by Evola for his ``unsuccessful'' attempt to recuperate Christianity for our time; they probably agreed on the freedom, creativity, and absoluteness of the person.

Berdyaev begins by pointing out that there already existed an ``authentic Christian theosophy'', among whom he included: \textbf{Clement of Alexandria}, \textbf{Origen}, \textbf{Dionysus the Areopagite}, \textbf{Meister Eckhart}, \textbf{Jacob Boehme}, \textbf{Franz Xavier von Baader}, and \textbf{Vladimir Solovyov}. To them he adds the pagans \textbf{Heraclitus}, \textbf{Plato}, and
\textbf{Plotinus}. In this sense ``theosophy'' means knowledge of the divine. \textbf{Helena Blavatsky}'s Theosophical movement, on the other hand, is unrelated to that tradition. Berdyaev describes her and other's writings:

\begin{quotex}
Contemporary theosophical writings are devoid of divine and creative inspiration. They betray no talent; they are frankly boring; they are, moreover, written in a style which would be more appropriate to a manual of mineralogy or geography; in fact they are almost impossible to read!
\end{quotex}

By ``Theosophy'', Berdyaev is also including the anthroposophy of \textbf{Rudolf Steiner}, whom he finds more ``interesting'' (not truer).
He criticizes the intricacies of its cosmological system, particularly since ``God'' cannot even be found in that system. Theosophy is a child of its time, being a form of spiritualized naturalism and evolutionism that were---and are---in vogue. It makes no demands as the spiritual world is attained automatically through evolution.

\paragraph{Theosophy as Symptom}

Theosophy is philosophically naïve and its adherents unsuited to high culture. We could say that the movement abounds with over-educated proles who thus seek a simple spiritual teaching that is allegedly based on science and the secret teachings of the ages. This is proved by the fact that there has arisen little in the way of high culture and incisive social commentary from theosophical quarters. The one exception may be Tomberg, but his magnum opus was created after he abandoned that movement.

Nevertheless, despite the ``very low level of theosophical literature'' and ``charlatanism'', it must be taken seriously when understood as a symptom of the crisis of Christianity and science, which may no longer speak to contemporary man. Unfortunately, it does not challenge the modern world, but rather adopts its values as a sign of evolution. Thus, it will ultimately fail to satisfy the ``deep spiritual longing'' which depend on religious faith and authentic mysticism.

\paragraph{Secret and Mystery}

Ostensibly claiming to be an esoteric teaching, Theosophy really turns it into a form of exoterism. Its esoterism offers a ``secret rather than a mystery''. A secret is information, conceptual thought, that can be told by one person to another; there is a natural human fascination with knowing a secret that others do not know. However, a mystery cannot be communicated that way, it is ineffable. It depends on one's own experience and intuition; the way can be pointed out, but each man must reach the destination on his own, depending on his own spiritual level.

Of course, Christianity also recognizes the distinction between the exoteric and esoteric. Berdyaev writes:

\begin{quotex}
There is both a deeper and a more superficial understanding of Christianity. Esoterism in Christianity almost coincides with mysticism. Christian mystics have been truly esoteric. But the contemplation to which they devoted themselves, though beyond the reach of simple Christians, has nothing of the enigmatic about it. To understand them completely it is only necessary to possess an experience similar to theirs.
\end{quotex}

It is not clear to Berdyaev exactly what the esoterism of Theosophy consists of. He points out the profound difference between a Blavatsky or Steiner and a \textbf{Jacob Boehme} or \textbf{Louis Claude de Saint-Martin}. He claims that ``when pagan polytheism flourished, monotheism was regarded as an esoteric truth to be hidden from the masses who were unable to reach such a high level.'' But he asks:

\begin{quotex}
What still remains to be discovered is what constitutes esoterism in an age of confusion like the present one, which is without any unique, integral, and predominant belief.
\end{quotex}

\paragraph{The Conception of Man}

There is a radical and profound difference in the way Man is understood in Christianity and in Theosophy. Of the former Berdyaev writes:

\begin{quotex}
Christianity is anthropocentric and anthroposophic in the truest sense of the words. According to Christianity, man is the highest order of being and is superior even to the angelic hierarchy. The Son of God became incarnate in a man and not in an angel. ... he is created in the image and likeness of God, he is not a product of cosmic evolution nor is he re-absorbed by it; he is not the child of nature. ... The human species cannot be eclipsed by a new race, whether of supermen, angels, or demons.
\end{quotex}

There is much more, but the point is clear. There is no evolution beyond man, which is the implication of the previous series of posts on involution and evolution. In particular:

\begin{enumerate}

\item There will be no evolution into higher races, with new ``organs'', such as Steiner taught

\item There is no ``superman'' coming as Nietzsche believed

\item There will be no transhuman based on technology
\end{enumerate}

There is no secret, but there is a mystery. The open secret is that we know who man is and what he must do to be ``saved'' or ``liberated''. The mystery then is to actually follow that path. This is because:

\begin{quotex}
The Christian conception of man is hierarchical and not evolutionist. Man is not a transitory fragment of the cosmos, a mere step in its evolution; he is superior to the cosmos, independent of its infinity, and in principle embraces it completely.
\end{quotex}

The Theosophical understanding is that man is a temporary and fleeting manifestation in an indefinitely long cosmic process. He comes into being and is dissolved again.

\paragraph{Pagan Gnosticism}

Berdyaev points out that in Theosophy man is but the play of cosmic forces, formed by the hierarchy of spirits superior to man. This part is unclear to me, given the genuine teachings of Dionysus. I believe the point is that the angels are free and do not block communion with God, while the theosophical hierarchy of spirits limits man's freedom. But the next point he makes is very telling:

\begin{quotex}
Theosophy re-establishes the old demonolatry and man remains subject to the genii.
\end{quotex}

Evola and Berdyaev part ways on this point. For Evola, the daemons, genii, manes, Laertes are an essential part of pagan teachings. For Berdyaev, however, they are the reminders of the time when man was unfree and bound to such spirits. Thus he can conclude:

\begin{quotex}
The freedom of the human spirit achieved by Christianity is no more and we are confronted with a return to ancient, semi-Christian, semi-pagan, Gnosticism.
\end{quotex}

Berdyaev then makes this observation:

\begin{quotex}
Theosophy and anthroposophy of every variety deny personality, and combat this principle in the interests of a sort of cosmic communion. They complicate the problem by failing to observe the distinction between personality and individuality. For Christian thought personality is a spiritual category while individuality is a biological one.
\end{quotex}

Both Evola and Berdyaev are in perfect agreement with that conclusion. Readers will have to decide about the role of the daemons in man's life.

\paragraph{Excursus on Various Traditions}

At this point, I want to bring in a discussion that Berdyaev did not deem necessary, although his thought has consequences for the answer. While Tradition sees hierarchy and inequality everywhere, it is strangely egalitarian when it comes to the various traditions. That is, Rene Guenon will say that each tradition is identical and the choice is arbitrary. Normally, a man follows the tradition of his people, but more advanced souls will choose one based on their personal needs. Furthermore, those very advanced will be beyond any particular tradition, as Evola believed himself to be.

On the one hand, a metaphysical teaching must be complete. On the other, it is impossible not to notice that traditions follow each other in time. Could we not regard this in the same way as the involution of man? That is, there is one main axis which represents the ``perfect'' tradition. The various traditions are attempts to manifest it, some closer some further away. There are futile lateral branches that are in reality counter or anti-traditions.

\paragraph{Excursus on Reincarnation}
In personal conversations, I have noticed that there is a great deal of confusion about the question of reincarnation in relation to tradition. Let us be completely clear at the outset. This is not a question to be decided by whether or not such and such a tradition ``teaches'' reincarnation. That question is irrelevant and its only purpose would be to distinguish false from true traditions. No, the only way to deal with this issue is to understand the metaphysics of it. Once it is clearly understood that reincarnation is simply not possible, there can be no more doubt. No one today who understands math and science will bother trying to square the circle or build a perpetual motion machine. Although Guenon, Coomaraswamy, and Evola have critiqued the idea of crude reincarnation, it is still wise to consider Berdyaev's point of view. He explains, in regards to the various theosophical systems:

\begin{quotex}
Man is reincarnated and thus loses his image, the substratum of personality ... the unique and integral, the ontological core of things, can be found nowhere. Personality comes together and then dissolves, only to reappear in other personalities. ... Man possesses a physical body which corresponds to the mineral kingdom, an etheric body which corresponds to that of plants, an astral body which corresponds to that of animals, and a spiritual Ego which has an affinity with God. All these constituent parts are then dissolved and so personality itself disappears.
\end{quotex}

Berdyaev compares this to those Russian Easter eggs, each one containing a smaller version within itself. While traditional teachings also understand the various soul elements in man, they nevertheless constitute a unity. Hence, you cannot take the vegetal soul of one being and arbitrarily combine it with the intellectual soul of another, like some children's game of Mr. Potato Head. As we recently pointed out, the spirit, or rational-intellectual soul, of man is attracted to a given genotype and historical-social situation. It would be absurd for the soul to incarnate at a different time into a completely different genotype and situation. This is what Guenon means by his claim that a being can be in the human state just one time.

It is this unity of personality that Berdyaev emphasizes: man as body, soul, person, an integral whole not an arbitrary construction.

Part II will deal with the failures of the Christian churches to teach an authentic gnosis in our time and what the solution might be.


\flrightit{Posted on 2013-04-24 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Mihai on 2013-04-25 at 04:09 said:}

This post is outstanding. I don't have the time nor the disposition right now to discuss the many profound and extremely points brought here.

I will have to engage in a deeper study of Berdyaev's work, as I believe that he might be a great source to balance some of the views more problematic views of the traditionalists.

\hfill

\texttt{Jason-Adam on 2013-04-25 at 13:45 said:}

Indeed this post is quite outstanding.

Let me say that the concept of hierarchically arranged traditions can reconcile the notions of perennial philosophy with the fact that there is no salvation outside the Church.

\hfill

\texttt{Synodius on 2013-04-26 at 10:15 said:}

I'm trying to understand mechanics of incarnation from the traditional perspective, and don't know how to view the nature of the spirit. I'll try to summarize how I see it, correct me please if I am wrong.

The spirit is the most universal part of man, but it is still subject to some limitations until it realizes the supreme identity. Until then (although time is taken only analogically here) it enters some degree of existence where it clothes itself with an individual soul corresponding to its specific nature and the soul then in turn clothes itself with a corresponding body. Individual faculties somehow provide means needed by the spirit for achieving the liberation. This can be done while still in the body or after death while still in the same degree of existence, although preserving individual consciousness after death is a kind of achievement which is not automatic. If this is not achieved, at the end of the cycle the being is reabsorbed and thrown into another degree of existence in the new one, into circumstances that are only in some way analogous to that in the previous incarnation but necessarily different (as the whole degree of existence is different), again determined by the nature of the spirit. Was this nature in some way changed of updated by the acts in previous life? Or only in some cases, because acts that modify just corporality for example, cannot touch the spirit?

There is no trace of individual memory from the life in different degree, so the individual consciousness has only one shot to be around during the liberation. If after death the individual consciousness is preserved can it enter some body by force so to speak? If the soul, body and circumstances change, is there some margin of change within which these can be still linked exclusively to the same spirit? And what about the change in the spirit itself? Are these questions wrongly posed? I've read Evola's Awakening, now going through Multiple states \& S.of the Cross and trying to make some sense of it... 

\hfill

\texttt{August on 2013-04-27 at 14:31 said:}

Interesting, relevant, and also disturbing questions.

``... re-emerging into circumstances that are only in some way analogous to that in the previous incarnation but necessarily different (as the whole degree of existence is different), again determined by the nature of the spirit.''

Moreover, if one's incarnation is and can only be a reflection of the `color' of his spirit, and we have attained the centrality and metaphysical opportunity of a human birth, how great is the risk that we lose this centrality in our next round of existence (assume no total liberation, and the active refusal to accept religious preservation of the human state via salvation)? Recall Guenon's ominous paragraph about the danger of being reborn as the equivalent of an animal or vegetable!

That should give us all something to chew over, especially if the human state really is so superior to other beings'; yet this view is primarily theological, and would be strange if transposed to the infinitude of Possibility, given that there is nothing inherently impossible about the idea of an indefinite number of states of being with more freedom, and hence superiority, than the human.

\hfill

\texttt{Caleb Cooper on 2013-04-28 at 11:13 said:}

Human superiority is based on his spiritual nature; i.e. he contains the full image of God, which he derives from his prototype, Adam, before the Fall. Now granted that down here for nearly all humans this remains an unrealized potential, because the likeness of God has been distorted, and we are in present fact inferior to the spiritual hierarchies.

To flesh this out a bit, God can exercise his creative power through all the infinite realms of creation. Man is the closest being to God in that the full image within him gives him this freedom as well. Most angels have very narrow ranges in which they are content to act as go-betweens, transmitting God's powers to.

They aren't very aware or interested of things outside their purview. There are many places ``where angels fear to tread''; in my inner work on myself and others I've noticed that when you're in a negative place where the vibrations are very low and dense, it can be very hard to invoke the aid of angels. Tomberg expresses this is as them being so pure that if they tried to force themselves to descend to an atmosphere where the devil is strongly present, the atmosphere would not support them; i.e. there would not be enough celestial oxygen and they would ``faint'', lose consciousness and float back up to where they can breathe. In these cases one can often still invoke the help of the Saints, showing the superiority of man in his ability to travel and be active where angels fear to tread.

We should note the incident recorded in the Islamic tradition where the most glorious of the Cherubim, Lucifer, was brought by God to the newly created Adam, and told to bow before him. Because of his pride however Lucifer refused, setting in motion his fall and sowing the great enmity between the devil and man.

God created man for a greater purpose than any other order of beings; to be His vicar in the created realms, the one entrusted to tend and cultivate God's garden. Adam can be seen exercising this prerogative in the naming the animals which gave them their missions. With the fall man lost his position as God's regent, but Christ reversed this fall and restored the image and likeness of man, ascending back to the position of God's representative to all the realms which has always been our birthright to reclaim if we wish.

TL:DR The truly human state understood as its essential spiritual core is the full image of God and thus superior to even the angels.

\hfill

\texttt{Cologero on 2013-05-05 at 15:28 said:}

Synodius and August: I have considered adding to these comments, but decided the only alternatives are Silence or a long discourse. The discourse is unnecessary since many have previously addressed these questions in a deep way, and you fellows have read them. But as our recent translations have shown, it is insufficient to consider these questions solely on a rational basis. However, there is a post planned for this week that will address this topic in part. I'll keep these thoughts in mind.

Shankara mentions the opportunity of the human birth, but adds some conditions as we shall see. Although Guenon has remarked that there is nothing special about the human state, given indefinite possibilities, it is not clear what he means. Even in his Islam, the angels superior states had to bow down to Adam. So, despite --- or perhaps because of --- its difficulties, it offers possibilities that other states don't. As for ``salvation'', Shankara points out that both the celestial (or heaven) and the hell body, do not offer these possibilities. Actually, they are extensions of the human state.

As was pointed out, acts cannot ``change'' the spirit, which is transcendent to them. However, there can presumably be changes from the human state to another; that is why the teaching of a literal reincarnation makes no sense. If you return again as a human, then what could possibly have been gained? But the spirit has various possibilities of manifestation, including those of the human state. So it make not be so much that the spirit ``changed'', but rather than it had to undergo a necessary step in order to reach a higher state. Obviously, this all ends in a flash with final liberation.

\hfill

\texttt{August on 2013-05-10 at 06:55 said:}

FWIW, here is my reply to Caleb Cooper (delayed by comment system problems), the aim of which is to propose horizons beyond the human that lack nothing of the potential for complete realization that humans enjoy here:

Yes, good notes Mr Cooper, especially about the ability of the Saints to intercede in realms where even Angels will not. By activating the potential of human centrality in this realm, and realizing it, one may gain access to the sole `portal' between the worlds, the `Universal Axis', if you like. We have learned the general names of such men, common between some traditions, as Primordial Man (one who realizes centrality of the human), and Universal Man (one who is absolutely liberated in this life), both of whom are of course more than mere natural men.

But beyond theology, and hence beyond our human state of being, every world is centered around the Universal Axis (=omnipresent God), and so there ought to be an indefinite number of states with beings that are situated at the center of their respective universe, with all the potential that this implies, and who are therefore in an equivalent position to we humans in this state. They too can pursue complete Deliverance, no less than we. Moreover, their state may be characterized by less pitfalls, less grossness, or be shared with fewer other beings; any number of other conditions that render their world and state relatively freer and perhaps superior to ours might apply.

Indeed, the whole idea of an initiatic hierarchy implies an indefinite series of states that are superior to our point of departure here, such that one who realizes his human potential partially (by going beyond the human), but not fully in the Great Liberation, may be reborn in a state where his centrality with respect to the Axis is preserved, and he is somehow better placed than in his earthly life. He may already have some experience of this state during his work in this life. And if we have had many traditions in our particular world, we can be assured that traditions of a kind are to be found in superior worlds, ever calling us home. But at some point of development, the experience of a given being is likely to converge with that of others, despite the point of departure; and the Universal Man need have no concern for superior states, because he has reached perfection beyond all states of being, and he `will not be born again'.

TL;DR Any state of being that is a full expression of the potential of its world will be centered around the nucleus of that world, which nucleus is penetrated by the Universal Axis, allowing that central being potentially direct access to union with God (Being) and Deliverance. Such cases necessarily involve transcendence of the particular state in which the being commenced the work. Man may be potentially superior to Angels, given their inherent limits, but not necessarily to other central states of being beyond this world, all of which can theoretically be used to attain the final goal directly.

Those who have not approached metaphysical premises outside of a religious framework, as pure concepts, may find these explanations unclear. There are no contradictions with theology, mainly a different vocabulary and broader scope. Guenon's `Multiple States of the Being' is a primary source.

Nevertheless, life is a mystery, and there is much room for prayer at this stage, that the mercy of God will bless us with the light that we need on our path of return, just as He has found fit to show us this world.

\hfill

\texttt{Synodius on 2013-05-15 at 21:42 said:}

If I understand correctly the geometrical representation from Symbolism of the Cross it is not the human state (or any other state) as such that is important, it is only its central point that is of metaphysical importance, the point corresponding to the True Man, which makes this state the starting point for the realization of the Universal Man. The being at this central point is superior in a way even to a being in a higher state but who has still not attained the center of its own state, that would explain why angels have to bow before a Man (in the center), but this superiority would not be exclusive for the human state because this can also be said about any state in relation to its respective higher states and so Guenon can say that the human state is not in any way special. This would also mean that theoretically by reincarnation into the same state but with better chances for reaching the center you could gain something while on the other hand by moving a level up to a higher state doesn't necessarily mean an improvement from the point of view of total liberation, because you start analogically in the same distance from the center as you were located in the previous state in the moment of death which is identical with the birth to the other state to which you were transported by the will of Heaven. This moment of death is not the same as the death of the body and various teachings about post mortem events (toll houses, tibetan book of the dead... ) deal only with what follows bodily death but precedes death/birth to another state and what determines its outcome; and are not to be taken literally because the womb to which one gravitate according to eastern conception would be on a completely different plane and past sins which are emphasised in the west are just something like habitual gravitational fields which make it hard to remain detached in the moment of this death. I still have a lot of questions but this is a sketch of how I imagine guenonian `reincarnation' might work... 

\hfill

\texttt{Synodius on 2013-05-16 at 08:06 said:}

Maybe I am trying to read too much into the Guenon's geometric representation which really impressed me but I felt the need to add more detail into it to get a more complete idea of how it works in a mechanistic sense, which might not be its purpose. I think it's message is really powerful: by locating the whole psychic realm into the same plane as just an extension of the human state it puts things into the right perspective and prevents seeing any kind of psychic curiosity as `spiritual', and by emphasizing the indefinite multitude of the states it makes one see the relativity and contingency of all conceptions that are too human. While I appreciate all this and was never a fan of the usual idea of reincarnation (because of all the stupidity that usually goes with it, e.g. I learned recently that a woman teaching taiji in the park in my area claims that she learned it in China where she lived 500 years ago although the form she teaches is a short Chen style form which is around 70y old) is it really impossible in this model? The soul could under certain conditions (not necessarily as a rule) continue in a different body which could have a benefit of not having to pass through a `hard reset' by the birth into a completely different state. The possibility of a more conscious continuation would make the methods that claim to help with this transition more interesting. Also the cases of demonic possessions, if they are not taken only psychologically, would suggest the possibility of breaking the exclusive link of one spirit/one soul/one body. Also, where would you locate the spirit in the geometric representation? Only at the central points along the Axis? This way the spirit would always be transcendent with respect to all contingencies represented by the horizontal line or plane, but then how to imagine the differences between the spirits of different beings? This is something which led me to ask maybe inappropriately about the possible change in the spirit, about some kind of spiritual substance that undergoes change. I spent too much time using some kind of jungian psychological container for everything higher than matter and now reading Guenon I am discovering that there is much more to this distinction of body/soul/spirit that I learned as a catholic kid. And while I understand the differences between intellectual and merely mental and imaginative faculties that Guenon stresses so often, I still have troubles to put it all into a coherent model. I know, rational systematization can become an obstacle to experiential knowledge, but maybe someone can relate more details to that geometric representation without distorting it

\hfill

\texttt{August on 2013-05-16 at 10:26 said:}

Hi Synodius. I have given thought to similar strong questions, and offer some more views here. As always, we are dealing with considerations derived from a true premise, in an attempt to explore the boundaries of possibility through logical application of `consequentials'. I heartily invite anyone to challenge what I have set out by pointing out potential flaws or misapplications of logic, from first principles to conclusions. Let us not abide by error!

---

``... the womb to which one gravitate according to eastern conception would be on a completely different plane... ''

You could well be using the term `womb' in a figurative sense, but let me emphasize that there ought to be states of being where one is not born in the progenitive sense.

``... is earthly reincarnation really impossible in this model? The soul could under certain conditions (not necessarily as a rule) continue in a different body which could have a benefit of not having to pass through a `hard reset' by the birth into a completely different state.''

Firstly, let's clarify that by `soul' we are referring to an individual's particular subtle elements, and not some higher faculty. Now, assuming that by `different body' you mean being born again after physical death into a newly conceived body, how could this constitute continuity unless the infant was born with the knowledge that the previous human had accumulated up to the point of death? This does not happen. As for `migrating' into an already grown body, what does that say about the soul already connected to that body? Such a case would be some sort of demonic machination, and not what you are referring to.

Now, to establish why earthly reincarnation really is an impossibility under that model, we need consider questions of time. Time, as a succession of events, is a particular condition of this state of being. As such, it is illusory, and cannot be transposed to the Absolute. In actuality, we (our Spirit) are all that is, and has always been; there is no time in the Spirit, only an absolute metaphysical moment, that all states serve to veil. Our Spirit is simultaneously participating in all possible states, in a moment that we shall call the Eternal Present. Our advantage on earth as humans is that we have the intellective faculty that allows us to conceive of this fact (that we are Spirit in human form) in an abstract way, and work toward directing the energies contained in the re-presentation of this world to our journey home by direct means, so that (ideally) we do not meet with another existence and its particular, veiled modes of awareness. The tragedy is that the very faculty and world that allows us to do this must be consumed and destroyed by this heroic battle, together with an indefinite number of other lives that we might have become fixed in (assuming Deliverance)! Yet, in victory, we attain eternity, because the course of our whole time on earth was only an instant from the perspective of our Spirit, and it is an instant that plays forever. It was the act that out Spirit projected once into eternity, and your life is the way the Eternal Present looks and must needs look through the prism of this state of being. The fact that we get to see it, and conceive of its time, history, and future development, does not change the primary fact that our experience of it is taking place within the instant of the everpresent Spirit.

Imagine a powder. When it is burnt by fire, it creates red smoke. When it is dissolved in water, red fluid. The powder is your Spirit, the fire and water states of being. Earthly reincarnation would be akin to taking this powder and seeing it become green when burnt by the same fire, or yellow when dissolved in the same water, randomly. This could only happen if the powder changed, in which case it is not the eternal Spirit.

Now, does any of this negate the possibility of some Master Adept returning, actively and voluntarily, to earth, say by appearing spontaneously in some desolate meadow, in human form but without any human needs, to reminisce about that time on earth when he was subject to the illusion of time? To see how his old home in this state looks now that he has gone `beyond Being'!? I can't exactly conceive of how going beyond form would affect the possibilities of form -- how objective is the existence of the world of forms to one who goes beyond them?

Importantly, if our Spirit is characteristically reflected on earth as a wise human, alongside other Spirits that are briefly here as tulips, paprikas, and cockatoos, why exactly would we incongruently manifest in another state as an equivalent of these non-intellectual beings? Guenon openly suggested this was the case, but it seems that this is merely so because it is possible. Truly a mysterious universe where one's Spirit incorporates the possibility of effectively subjecting itself to a great and wise life in one state, whilst simultaneously being a lowly worm in another, and hanging the Sword of Damocles over its own head, that it might become centered in awareness of a dull state and all its grim attendants! Are we to take comfort in the idea that a man, composed as he is of parts, can choose to recoil from his lower nature and avoid ever having to fix his intellection amongst the existential possibilities of that dark part of his Spirit? Or maybe there is some consolation hidden in these lesser states.

In any case, Deliverance is not a question of `nothingness', in the sense of everything ceasing after one has experienced life (this is a horrible thought!). Deliverance seems like a full and complete repossession of the Great Unity, making it one with your Self. I do not see how this could happen without implying the eradication of anything that is not Self or willed by Self, such that you are anything and everything thereafter (where `things' apply). And this brings us to your next question.

``... how to imagine the differences between the spirits of different beings?''

Yes, let us try to clarify somehow. Here I would emphasize the importance of Non-Being. If we stopped at Being, it would seem that all spirits simply converge in God, and that any resulting differences are merely His choices expressed on contingent planes. In this case it would be odd to talk of Self in a transcendent sense at all, there would just be many contingent beings with relative selves and a God at the center representing an absolute limit. But in Non-Being there is a `space', if you like, wherein is an Absoluteness that is totally effaced, such that it can house non-converging differences that are capable of total expansion without limiting one another. Hence, apophatically, the Infinite. Pure Being is something like a prototype of all possibilities of Being, wherein differences can manifest indefinitely, but are all permutations of Being, their common element.

The Spirit should be conceived as a reflection of the Great Unity, combining both Being and Non-Being.

``Also, where would you locate the spirit in the geometric representation? Only at the central points along the Axis?''

In the three-dimensional cross representation, the spirit passes through the center, but is not actually there, and it emanates everywhere. Think of the physical sun as a ball, at the center of which is a port opening onto a hypothetical world of pure light; at this center we have the highest concentration of light in our world, but its rays permeate our whole world to greater or lesser extents, and through a greater or lesser variety of obstacles.

\hfill

\texttt{Cologero on 2013-05-22 at 23:00 said:}

Synodius, in the Spiritist Fallacy Guenon gives a clearer account, albeit in passing. ``Trans''-migration means the transition to a different state. This is unlike the popular notion of reincarnation which can only be a ``migration''. He also points out the possibilities of psychic inheritance, which may in some instances take several generations to recur. Again, this psychic inheritance does not prove ``reincarnation'' although it may ``feel'' that way to those who are sensitive to it.

\hfill

\texttt{Sigurd on 2017-08-02 at 21:33 said:}

Removing some latent Theosophical ideas has been one of the primary obstacles I've faced in approaching the subject of tradition. Even in reading contemporary works that may be helpful in explicating the lived experience of traditional societies, it is often difficult to ascertain elements of another order mingled in. An example that comes to mind is Jeremy Naydler's Temple of the Cosmos, where he gives a brilliant introduction to Egyptian tradition, but makes a point that the way of experiencing the cosmos ``sympathetically,'' having a symbolical sensitivity to cosmic events, and seeing thought-forms and emotions as ``other'' and as beings, has been, more-or-less, superseded in the natural evolution of consciousness, seemingly suggesting that a sort of ``coenaesthesis'' is the ``bridge'' to the next stage of human consciousness.

On this topic, perhaps I can request a response (or referred to an appropriate place to edify myself) on the topic of the astral or ``etheric'' body. It is to my understanding that the contemporary notion of the subtle body as a `projectable,' spectral-like doppelganger of the corporeal body finds its roots in Theosophism. The etheric body in a traditional sense (including all those variations that envelope Neoplatonism, Hinduism, etc.) is a sort of vehicle or ochema of the Self that links the Self to the conditions of the body/somatic vehicle. It is thus a modality of consciousness, in contrast to a projection of a sort of prefigured, animated ``double,'' though the latter is an ability one acquires in the course of initiation.

Is this, then, the appropriate understanding of the subtle form? To where ought one categorize such phenomena as contemporary ``astral projection'' or projection of a double that is considered the subtle form by some occultists and spiritualists? Would much appreciate a response, Cologero.

\end{sffamily}
\end{footnotesize}

