\section{Benedictines and Lectio Divina}

\subsection*{1. Introduction : the Voice of God}

When we speak of the Bible, we speak of the Christ-Logos as the recapitulation of the Word of God as it fills the universe as a sign of God's will, inscribed in the order of all things. It is therefore necessary to meditate on what was passed on to us by the Scriptures and the Tradition, both in the Guenonian sense of the word, but also as the Christian Tradition who wrote the Second Testament of God. In prayer and meditation we speak to God; in the Scriptures it is He who speaks to us. It is in this dialogue that we can understand the perfect Silence of His presence. We have to liberate the senses, to look out for Unity, the source of the Being in a relation of the Other; we have to get to the Self by fixing ourselves on God. The Word of God is a Word of Life and Truth (Mt 13, 19). In Hebrew the word speaking is ``dabar” which also signify ``revealing things”, to ``shed light” on them, to ``act in a profound manner upon them”; to speak is to act. Just as in Genesis, God speak, and things are. In the truest sense of language, the speech does what it says, and says what it does. The Ancients, like Philo, believed that Scripture and Nature were interconnected: to understand one is to understand the other, and to understand one you have to look at the other. In reading the Bible, we open ourselves to that understanding of the universal. The universal gives way to the particular, just as the Celestial Jerusalem can be traced (and will descend from the sky) on Earth. Of course, the existence of the earthly Jerusalem is to attain the Celestial One, to come out of the secret of things.

In this brief text, I will look specifically at one way of understanding the text, Lectio Divina, and how (and why) it is applied by a specific order of Christian monks, the Benedictine.

\subsection*{2. Main: Learning the Scriptures. Rule of St Benedict}

\paragraph{Silence}
The first rule of the monk is Silence. This silence is well defined as their motto is \textit{Silence yourself and listen, }as can be seen around their monasteries. Outward silence may come to mind first, but as you go through their monastery, you will see that it is not so. They use Gregorian chant and other ancient music for their offices; they work and need to speak to properly do it. They even take some times to chat between themselves. But the silence that is true to them is the inner silence. In prayer, in walking, in Oratio (i.e. mental prayer, see below), etc. they all fix themselves on God and respect the Silence. The first word of the rules of St Benedict is ``Listen…”. The lower silence (exterior) is always linked with the higher silence (inner one).

\paragraph{Obedience, stability and conversion}
Their three vows are the three vows of monasticism in general. Obedience to the superior as obedience to God. They respect authority as it is a direct hierarchical representation of the divine hierarchy (see St Denys) in the heavens. For them, the Abbot acts as Christ within the community. There is a strong sense of self-disinterest in this form of obedience, as they try to get rid of their self-will to attain the Will of God. The other vow is stability, linked with the practical establishment of a sense of enduring fidelity to one own monastery; but also it is a stability that is within God, an inner stability (linked with silence). Lastly conversion for them is the path of converting part of themselves through purification to a higher spiritual connection with God through Christ; through Him in imitation and in consciousness.

\paragraph{Inner life}
Their inner life is developed through the shedding of the false self, the human will and the vice within themselves. They work through prayer, offices, chants, rosary, labor and Lectio Divina. I will expend on those last two later on. As Cassien puts it, the asceticism is the center of the monk's life to purify his heart. It is done through humility and prayer. There must be a desire for what he calls the ``Real Science”. The goal is to be in complete control and concentration at each moment and purify the thoughts at each moment. And that is done through the Scriptures, or through the praying of the Scriptures (for the Fathers, it was necessary to know the Bible in one owns memory, so that with time we speak through it, with it; so that with time, our interior teachers would be Moses, Abraham, St Peter, etc. in everyday life).

\subsection*{Ora Est Labora (Work is Prayer)}

\paragraph{Ora}
Prayer is the center of the monk's life. They gather to pray, and they return in their individual cell to pray. They pray when they work, and they work as they pray. Praying is working, and working is praying. They develop humility and virtues through incessant praying, and they sanctify their lives through this purification. Their whole life – their whole Being – becomes a prayer offered to God. They participate in the Cosmic law by imitating Christ (good works, faith and spiritual development – ``Luke 5:16 But Jesus often withdrew to lonely places and prayed”, the lonely place of the Being, the Soul) and try to sanctify Creation. They participate in the Absolute within the Relative so that the Relative may become again Absolute; they pursue the way of Christ.

\paragraph{Labora}
They work because they want to be part of the World as they sanctify it, not so much as to enter the world. Of course, as they want to be stable and self-sufficient, there is a macro-micro relation present in their work: on a smaller scale they have to work to satisfy the necessity of living; yet in the upper levels of understanding, they also do it to sanctify their work that others de-sanctify in the modern world.

\paragraph{Ora Et Labora (Prayer and Work)}
Therefore, work and prayer are linked and interconnected. They represent the goal of the monk to go beyond dualism. By praying (contemplation) and working (action), they combine them both and make them look lower than the actual goal of both of them: attaining God. It is their way of explaining the non-dualistic perspective of the contemplative.

We could also point out to an old Christian text:

\begin{quotex}
A wandering mind is made stable by reading, vigil and prayer.

Flaming lust is extinguished by hunger, labor and solitude.

Stirrings of anger are calmed by psalmody, magnanimity and mercifulness.

\end{quotex}
\paragraph{Mary}
And that contemplation is expounded also in their devotion to Mary (as is within the Church in general). Mary is the perfect contemplative. When the Angel came to see her, she listened. Having listened to the Word of God, she shed out her own self (her will) and said ``yes” to the Life that is God; so that His work may be done through her (Luke 1:38 “I am the Lord's servant,” Mary answered. “May your word to me be fulfilled”). There is also the symbolism of the chivalric love with Mary as the Divine Sophia: Love (Divine Love – also note that to really Love, one as to abandon himself in the Other that is loved; that is, we have to abandon ourselves in God) as a way of knowledge. St Bernard wrote on this extensively for those who want to learn about it.

\subsection*{Lectio Divina}

\paragraph{Introduction}
When reading and praying, the only goal is God and His Unity. Every other consequences (happiness, peace, etc.) are just good consequences. Before starting the whole process, it is important to conquer noise and activity in the mind. The reader must regain his Silence. There must be emptiness through disinterest, stopping the flow of thoughts in the mind. What we mean here is that one must control his mind through incessant prayer or meditation before starting. If one is can't aim at this for exterior or interior reasons, he must try to cleanse his mind by reading without stop the Scriptures; Lectio demands a total concentration on the object at hand, that is, the Word of God.

Through grace, we are to let the Spirit (source of the Scriptures) guide us through the process to give life to the Word of God by the Logos. It doesn't depend on us as much as on Divine Will and our reception to it. Those of you who find it difficult to think about reading the Scriptures without being self-willed should think ahead and ask themselves if they are not too proud to ask for help and guidance in reading the Scriptures; therefore the problem lies with lack of silence and disinterest. More so, as the Fathers said, if you only read to find what you want, or if you think you can interpret everything in the Scriptures (as Nature), it is not God or the Self that you love or seek, but yourself.

In all the process, there is to be a bareness and dispossession of the ego, of the self (through humility) against any romantic view of the Self and the Cosmos; the Being must align itself with The One Up High; this elevation of the heart (intellect) toward the divine knowledge must not be done through a subjective and sentimental reading of the text (even if those can happen as consequences of our human nature, and at least influence us in a positive desire for more of the Scriptures, it's act, or God).

The four part (Lectio, Meditatio, Oratio and Contemplatio) are gradual process. At first, one will probably only tackle with the inner silence and Lectio. After some time, he will be able to integrate Meditatio; that being mastered, Oratio; that being mastered, Contemplatio. It is progressive in the act of the Lectio Divina (Lectio to Contemplatio) and through time in terms of practice and devotion. What we don't know, we learn in the Scriptures; what we have learn we meditate, what we have meditated transpire in our acts and Being.

Lectio Divina, contrary to the spiritual reading of the XVIIe century, transcends the dualism between affect and intellect: it is reconciled in the Meditatio/Oratio part, and breached through in the Contemplatio.

Note that in some of the oldest books of monastery, Lectio Divina is strongly attached with a master-disciple scheme. Therefore, this introduction to Lectio Divina can be but incomplete: but at least one will not fall in the hell where the neutral ones are.

In the following description, there will be less and less information, as more one progress, more one progress in his own way. Furthermore, some parts cannot be described in words (contemplation of God (Contemplatio); personal prayer (Oratio))

\paragraph{Lectio}
Lectio is the basis of the process, and it must be taken with utmost respect and reverence for the Word of God. The Lectio should be done at a determined moment (and not at ``lost time” or ``when one has time”). One should have Mt 6:6 in mind when starting the process of Lectio. It is through discipline and effort that the rejection of the modern rhythm can be done to acknowledge one with the Divine Rhythm. Lectio should be done therefore in isolation and silence.

Cassien speaks highly of self-purification before lectio. We should control and purge the basic vices of the Thumos/Eros before using the Nous to read; or nothing will be revealed. Unique prayer of uninterrupted words spoken with the heart is for him the best way to purify oneself. The goal is to become harmonious with the teachings of the Word (that is, the Cosmos).

One should read without discrimination (i.e. choosing what to read): some texts will ``tell nothing” but it is at that moment that God will take on that silence. That being said, some text should be read, but maybe not re-read without end, unless they do speak to one. The goal is that one should not read John 1 at each Lectio, but one should not also read The Book of Numbers through all the year in hope of God speaking to him.

One should accept that some truths are higher than him, and therefore develop humility through that lack of understanding. It is through assiduity, fidelity and grace that one will unlock the secrets; God will speak through his Word, and when the whole of the Scriptures is in one, the whole Cosmos will become the Scriptures and God will speak freely to the one. We have to learn the Scriptures to perfect memorization, to transform us into the Word. Also, one should not forget that some signification can come from reading the Scriptures without being ascribed to them per se. In other words, you should not expect everyone to understand everything the same while reading; neither should you ascribe to every view out there.

One should try to read with a loud voice the text, as to integrate it in his mind. One would be wise to read the text without immediately thinking about it, that way he will not suffer the subjectivity or the ideas that comes with it. There is no thinking about efficacy, technical results or psychological results. This is not a New Age meditation technique to feel good; there should be no attention on the ego-self, we should look at the text as if it was told for the first time to the Cosmos, with the Eyes of God. We have to be part of the text: when we read Genesis 1, we have to be there. We have, through our own speech, to create the world with God. In reading the Scriptures, we actualize its Potentia. One is able to put the Eternal Word in the Eternal Present as if to stay in the Word.

The goal being understood, one should not shun the external help to better understand the text (especially if one does not read Hebrew and/or Koine Greek): hermeneutics, biblical research, cultural/historical analysis, etc. Those methods can help one not to fall to preconceived ideas and therefore to subjectivism. It should not shape the reading, but rather help at non-reading. The objective is still the Lectio, not the method. One reads the Word of God for spiritual unction and Charitas rather than culture and erudition. Yet, to help us understand what the text meant to those who, by Divine Grace, wrote it, we have to be able to immerse ourselves in them, and therefore reach God.

To choose what to read, one could make himself a plan on one or two years spanning the full Scriptural text interconnected with Patristic texts; or one could follow the liturgical year of the Church.

\paragraph{Meditatio}
Meditatio is also called ``Mastico” as to masticate the Word that we received. It should engage the whole of the Being through memorization, actualization, attention and whispering of the passage that was read or retained (if one reads three chapters, it would not be wise to meditate on the three of them at the same time; one should aim for particular to universal: start with the verse that was the most spiritually descriptive, then moves on to meditate on the whole passage, then the whole Bible, then the whole Cosmos). As St-Basil said: God becomes fixed in us through memorization.

Meditatio should be taken as deep reflection and intensive internalization of the Word. It can use different schemes to attain that goal, like particular to universal; analysis of words and sentences; ``mantra”-like reflection on the words; etc. Some teachers also suggest writing down reflection made, to read them afterward and keep them as a teaching that God made through Oneself to oneself.

\paragraph{Oratio}
The Oratio is the mental prayer. As St Augustin said : ``If the text is prayer, pray; if he is mourning, mourn; if he is joy, let yourself be filled with happiness”.

The Word came to us through Scriptures and comes back to God through prayer; just as the Word became Incarnate so that we too could rejoice in God. One should open this part by psalms, chanting, verbal prayer, internal prayer, etc. Fathers speak of ecstatic divine love that should not become the goal in itself. Then it becomes a ``tranquil meeting with God” where the Word goes up like the Spirit toward the sky so that we can see the divine light in the most profound part of the Self. It is to open the door on which the Word-Logos is knocking; to contemplate the Word became Flesh in ourselves.

\paragraph{Contemplatio}
See Jn 11, 20: like Mary, we have to be still and listen. No vision, no ecstasis, no apparitions (II Cor 5:7 \textit{For we live by faith, not by sight}). The veil between the Verb and the Self vanish; duality fall within Christ. There contemplative knowledge can be learned. There is no sensuality, no sentimental approach. Here the Divine Love (charitas) is united to Faith. The way to enter hear is to come back to the Silence of the beginning through the whole process so that every possible thought is rejected in a Divine Silence that God will fill. It is useless to think of God, because one would only think of god (see Eckhart); it is useless to think of God when one is in His presence. The Word is condemnation or salvation: we need to put it in practice, just as Jesus Christ, being incarnated, was the new Adamic Archetype and acted upon the Good and the good. The second one to reach and teach the first.

\subsection*{3. Conclusion}

This outline of the Lectio Divina is to show that some method for contemplation exists and are strong within Christianity, both in the East (Hesychasm was spoken of here at Gornahoor) and in the West. Be it Lectio Divina, Hesychasm, Eastern meditation, Dikhr, etc. the important part to remember is that praxis is necessary. Without some form of inner purification, inner contemplation and inner meditation, what we read and think stays on the speculative level, on a level of minds and thoughts that serves no purpose. We have to reach our Being through Traditional teachings. Furthermore, it is not through vice and sensuality that we reach that sort of enlightenment (I exclude myself from any sort of debate about a Right Hand Path), and therefore the exoteric and esoteric goes hand in hand, and both prayer and asceticism are needed to reach deeper techniques of perfection. The Benedictines may not be a Traditional order per say, but they do represent a Traditional forces in the sense that they possess knowledge and ideals of reaching God. In the very least, they are a depository of saintliness, of Good and Beautiful and True.

As a conclusion I would like to offer a small thought. Huichols shamans in Mexico think that if they stop doing their purification rituals and offering, the whole of the Cosmos will crumble upon itself. In a sense, this is true for every Priest in the world: in an age as dark as ours, every bit of purification and sanctification of the world is important. If the world would become totally closed to God, totally profane, it would indeed crumble upon itself and come back to a stage of total Chaos. Those monks at least sustain the world as they try to unite it with God through themselves.


\flrightit{Posted on 2013-09-29 by David}

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}

\texttt{jc00001 on 2013-09-30 at 13:33 said: }

Excellent article about the practical application of exegesis. I haven’t found anything like this before. A true treasure.

\hfill

\texttt{Matthew on 2013-10-18 at 14:54 said: }

Great article, thanks. If only more in the modern world took this approach to the reading of scripture.

\hfill

\end{sffamily}\end{footnotesize}
