\section{The Infancy Narratives}

This is a review of \emph{The Infancy Narratives} by \textbf{Joseph Ratzinger}. Since the story is, or should be, very familiar, we want to instead focus more on the exegetical methods used by Fr. Ratzinger than on the story itself. He writes that good exegesis involves two stages.

\begin{enumerate}
\item \textbf{Historical component}. What did the authors intend to convey through their text in their own day? 
\item \textbf{Contemporary component}. Is it true? Does it concern me? How so? 
\end{enumerate}
Obviously, spiritual writings do not concern me in the same way as a new tax code might. Unlike a subject like chemistry, for example, spiritual truths are best communicated through myths and symbols. As such, their meaning, their ``concern for me", are not always open to everyone. That is because they describe ``spiritual processes that have given birth to a new people" [Bachofen].

That is true in regard to the narratives surrounding the birth of Jesus. Nevertheless, these narratives also claim to be representing ``the history of events occurring materially on earth". There are two easy, but ultimately unhelpful, solutions (beyond complete rejection).

\begin{itemize}
\item Accept the spiritual teaching but not the material. Christ, then, is one of the world's great teachers. This superficially sounds more ``intellectual", since it avoids the messiness of actual events. 
\item Naively accept the historical aspect without a deeper concern. This is the sentimental solution. Unfortunately, facts are mute and do not speak for themselves. Hence, there is the need for a careful exegesis of the type taught us by Fr. Ratzinger. 
\end{itemize}
We have the model of \textbf{Marius Victorinus}, the neoplatonic philosopher who initially accepted Christianity as an intellectual system, but not as a historical teaching with institutions, and so on. Or perhaps the Inklings, like C. S. Lewis and Tolkein, had a better idea: it is a myth that actually happened. Here, we will take the perspective of Victorinus. The historical component will have to be an individual decision.

In the first part, we highlight Fr. Ratzinger's exegetical points, drawing out their logical consequences, and placing them into a larger perspective.

The second part on the birth of Jesus and the revelation to the shepherds will appear on Christmas Eve.

The final part on the Magi will appear on the feast of the Epiphany.

\paragraph{Salvation}
Perhaps we no longer consciously think in terms of a savior, but at the time of Jesus' birth, the idea of a savior was in the air. This was written about Augustus Caesar, the Emperor at the time:

\begin{quotex}
Providence, which has ordered all things, filled [Augustus] with virtue that he might benefit mankind, sending him as a Saviour … the birthday of the god was the beginning of the good tidings that he brought for the world. From his birth, a new reckoning of time must begin. 

\end{quotex}
Fr. Ratzinger points out that religion and politics were not separated as they are now. Salvation then also included peace, a recurring theme in the Christmas message. Yet, a political solution is not the point. In \emph{Introduction to Christianity}, Fr. Ratzinger describes the situation of man, which has been called ``original sin".

\begin{quotex}
The seat of original sin is to be sought precisely in this collective net that precedes the individual existence as a sort of spiritual datum, not in any biological legacy passed on between otherwise utterly separated individuals. … no man can start from scratch any more, completely unimpaired by history. No one starts off in an unimpaired condition in which he would only need to develop himself freely and lay out his own grounds; everyone lives in a web that is a part of his existence itself. 

\end{quotex}
In other words, we inherit this structure. This is not biological heredity, otherwise it would be a medical problem. That is, a ``physician of the body" could develop a serum, a vaccine, a surgical procedure, or a genetic recoding to ``cure" original sin. So original sin must be a defect of the soul, not directly the body. \textbf{Rene Guenon} states how this is possible, while still being inheritable:

\begin{quotex}
That there is a psychic heredity as well as a physiological heredity is hardly in doubt and is even a fact of common observation. But what few take into account is that at the least it supposes that the parents furnish a psychic seed as well as a biological seed. (\emph{The Psychic Fallacy}, p 176) 

\end{quotex}
So salvation can come only from a \emph{physician of the soul}, a designation for Christ. The world system cannot be cured from within. With the knowledge of good and evil, man learned to separate the ``no" from the ``yes", the false from the true, evil from good, doubt instead of faith. The influences of the world are legion and contradict each other. Only a transcendent influence from above can be the source of salvation. So being enters history. \textbf{Boris Mouravieff} identifies the world system as the ``A" influences and the transcendent as the ``B" influences.

\paragraph{Death and Destiny}
In his analysis of the symbolism of Ocnus the Rope Plaiter found in a tomb, \textbf{J J Bachofen} describes the pagan worldview revealed therein, in these terms:

\begin{quotex}
The thread of death is woven into the web of which every tellurian organism consists. Death is the supreme natural law, the \emph{fatum} of material life, to which the gods themselves bow, which they cannot claim to master. Thus the web of tellurian creation becomes the web of destiny, the thread becomes the carrier of human fate … the loom, carrier of the supreme law of creation written in the stars, was assigned to the uranian deities in their sidereal nature; and, finally, that human life and the entire cosmos were seen as a great web of destiny. 

\end{quotex}
The radicalness of the new message cannot be known unless and until one can place himself in the place of the pagans at that time, fully feeling or recreating in his interiority the weight of this heavy conception of death and destiny. The pagan view, as exemplified, for example, in Plato and Hindus, was entirely moralistic. A person was rewarded or punished after his death in accordance with his acts while alive. This judgment was impersonal and mechanical, the \emph{lex talionis}, ``an eye for an eye". In contrast, the new message is one of life and freedom, not death and destiny. Fr Ratzinger recounts \textbf{Bernard of Clarivaux}'s sermon on the topic:

\begin{quotex}
After the error of our first parents, the whole world was shrouded in darkness, under the dominion of death. Now God seeks to enter the world anew. He knocks at Mary's door. He needs human freedom. The only way he can redeem man, who was created free, is by means of a free ``yes" to his will. In creating freedom, he made himself in a certain sense dependent upon man. His power is tied to the unenforceable ``yes" of a human being". 

\end{quotex}
Thus God is revealed both as Being and as acting in history, transcendent and immanent. Unlike the pagan gods who themselves were subject to death and destiny, God is revealed as the creator of the cosmos. No longer subject to the stars, man becomes free. So free, in fact, that the salvation of the world depends first on Mary's free choice. Only in freedom can there be an escape from the domination of worldly influences. In \emph{Introduction to Christianity}, Fr. Ratzinger writes:

\begin{quotex}
Only by the action of the individual can the transformation of history, the destruction of the \emph{dictatorship of the milieu} come to pass. 

\end{quotex}
This ``dictatorship of the milieu" is the set of A influences.

\paragraph{Genealogy and World History}
Even to the casual reader, Matthew's and Luke's genealogies differ from each other in significant ways. Rather than engage in hypothetical contortions to harmonize them, Fr. Ratzinger makes another choice. Anyone familiar with esoteric texts will recognize that deliberate contradictions indicate that the text is to be understood symbolically, not literally. Hence, Fr. Ratzinger brings out the symbolic meaning of the two accounts.

Matthew starts with Abraham and leads forward to Jesus, in three sets of 14 generations each. Curiously, he points out that the number for the name ``David" is 14, giving force to the idea that Jesus is the heir of David. This is the only place where Kabbalistic numerology is used, hence this can be understood as a tacit endorsement of that method.

So for Matthew, Abraham is the beginning of the history with Jesus as the fulfillment. The first phase of 14 generations leads from Abraham to David; the second phase, a period of decline leading to the Babylonian captivity; and the third, rises again to Jesus. However, Joseph, while the legal father, is not mentioned as the ``begetter" of Jesus, who was instead born of Mary. So legally he belongs to the house of David, but his origin is elsewhere, from above. The idea of vertical and horizontal heredity will be discussed again in the third segment

Luke, on the other hand, starts with Jesus as an adult and then works the genealogy backward to Adam. This indicates the Jesus is a new orientation to world history. Adam transcended animality by being made in the image and likeness of God. Jesus, on the other hand, is the god-man. The god-man transcends man just as fully as Adam transcended the animal. We, then, are offered the opportunity to achieve theosis, to truly become god-like. Thus the unfulfilled goal of both Adam and Plato, to be like the gods, finally becomes possible in world history. This will be brought up again in the third segment.



\flrightit{Posted on 2014-12-15 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Tom Blanchard on 2014-12-24 at 10:45 said: }

Lovely Christmas reading – thank you for this. Looking forward to the second part.


\hfill

\texttt{Michael on 2015-01-02 at 11:01 said: }

Beautiful post. I am curious about your comment on kabbalistic numerology: do you see this as a tacit endorsement of the kabbalah as a whole? I am not sure how much of the kabbalah is unique to kabbalah versus just being part of the Judaism of that time. In other words, I am wondering if you would recommend study of the kabbalah.


\hfill

\texttt{Cologero on 2015-01-04 at 10:52 said: }

Well, I am curious about why Fr. Ratzinger, who is very careful with his words, would mention it. Of course, in Hebrew every word is also a number. Nevertheless, I don't believe that any ``serious" exegete, up until how, has proposed that the numbers have any significance.

As for the study of the kabbalah, you could start with Meditations on the Tarot and determine for yourself if the kabbalistic references add anything to the depth of understanding and spirituality. We can be sure that Fr. Ratzinger was at least familiar with the existence of that work, if not its actual contents.


\hfill

\texttt{Mercurius on 2015-12-21 at 22:28 said: }

This past year, I've been working with/in a very demanding company that tends to suck away anything of ``free time"–thus, at best, due to time strictures, I've only been able to occasionally touch the hems of spiritual Life in the Western liturgical cycle. 

As a result, Advent came and went fast, not allowing much opportunity to reflect, and denying something of a developing ``Christmas Spirit". Yet, this post, was corrective–it compensated, and provided JUST what I could not attain by myself ``this season", restoring my lacking ``Christmas Spirit" to where I'd expect it to be–and, coincidentally proving, or illustrating that, ``Saviors" are necessary

Michael, re ``Kabbalah" in antiquity, and the Ratzinger remarks, I myself (at the least), find the ``Apostolic Gnosis" series of Lea and Bond pretty important exegesis from a Greek wind. 

\url{http://www.amazon.com/The-Apostolic-Gnosis-Materials-Study/dp/1872189695}


\end{sffamily}\end{footnotesize}
