\section{Detraction, Soul Mates and Paranormal Powers}

\paragraph{Detraction}

\begin{quotex}
Detraction is the unjust damaging of another's good name by the revelation of some fault or crime of which that other is really guilty.
\end{quotex}

This would include, for example, the public revelation of a private conversation if the intent was to damage the reputation of one of the parties. In recent days, there has been a very explicit example of such detraction. Yet, this is hardly the first time such an event has been deemed newsworthy. What has always surprised me is the lack of understanding of the role of detraction, which is a very serious moral fault. I have not heard any TV commentator mention the word, nor have I seen a blog mention it, despite the many words written about that event.

Someone pointed out to me that the news media run on detraction. That is, they are always gossiping or ``digging up dirt'' on someone. If that be true, then the media are evil. It is good journalism to uncover some crime or malfeasance in office; not so if the intent is simply to harm someone's reputation if there is no crime or other just reason involved.

\paragraph{Cloning and Soul Mates}

There hasn't been much talk about human cloning lately. Presumably, this will be available only to the wealthy who can afford such a procedure. That makes no sense to me, since the clone does not really benefit the original. Wouldn't it be better to clone one's soul mate, like Adam and Eve? Now you cannot get a man from a woman's DNA, but you can do the opposite. The female clone would have the exact DNA of the man, except that she would have two of his X chromosomes instead of the Y. Isn't that what a man is looking for? I mean a woman who is just like him in every way.

I may start a company to do this. You should buy in now, while the shares are priced low.

\paragraph{Freedom of Speech and Intolerance}

The ideal of freedom of speech has been severely tested recently. For example, there is that owner of a basketball franchise who was punished for his thoughts. In fairness, his ideas are inappropriate in a league that is 90\% black. Condaleeza Rice was hounded out as a speaker at Rutgers; again, her point of view didn't fit into campus culture, if you can call it that. Soi-disant ``conservative'' commentators have defended Rice, but not Sterling.

Freedom of speech means, in practice, that expression of ideas is tolerated. Although seldom made explicit, the corollary of this point of view is that the intolerant are not entitled to such freedom. Although Herbert Marcuse promoted this idea in the 60s, I believe it can be traced back to Karl Popper. In Rice's case, the principle is that the alleged criminal has no right to speak of her crimes. The in-group defines itself in both cases by what it tolerates and what it rejects.

There is a lesson there for the various groups who claim to be ``saving Western civilization'' (or Occidental, if you prefer to sound pompous). I heard one such fellow interviewed this weekend. Unfortunately, he is an atheist. Since when has Western civilization been defined by atheism? There are others who promote various non-traditional lifestyles. These movements are intellectually incoherent as they are based, at best, on quite secondary issues. There needs to be more intolerance in this group, but few have the stomach for that. What then to tolerate? We stick to the principle of looking back to the time when well-bred men knew what was normal and healthy.

\paragraph{Girls Gone Wild}

Carl Jung identified the masculine qualities in a woman's psyche, which he called the ``animus''. Although essential for her psychological development, for the unindividuated woman, the animus can take over the personality. It this case, the woman believes she is being as ``tough'', for example, as a man, although in reality she is just being argumentative and domineering; however, there is a less polite term in common use for the same phenomenon.

Recently, for example, Sarah Palin made the claim that \emph{waterboarding is our baptism for terrorists}. Leaving aside the blasphemy that everyone is ignoring, that is just some ugly caricature of what, she must believe, sounds like a ``tough guy''.

This is yet another example of a really poor metaphor expressed by a public person. There have been worse. I'm sure they must have done poorly in the analogy section of the SAT. Maybe they should take a couple of years off to read poetry and literature, although we assume that the speechwriters have done so.

\paragraph{Dignity in Office}

Politicians have no sense of dignity, presumably because they were elected by the ``people''. For example, Boehner did a poor imitation of a crying baby in a speech. Instead, he could have made an intellectually compelling and coherent case for his point of view. But, yikes, that would only happen in an alternative universe of well-bred men.

I won't even mention the picture of Obama on a bicycle with a silly grin, short pants, and helmet. I am much more careful now about how I go out in public.

\paragraph{Paranormal Powers}

This was a topic in a discussion group several years ago, where we were discussing \emph{Darwin's Dangerous Idea}. The topic of psychic powers came up. I made the point that if psychic powers really existed, then they would be much more common today. Clearly, a population with such powers would have a reproductive advantage over their neighbors, and the relevant gene would have spread.

Of course, that assumes that genetics explains everything about man, which we in fact deny. The existence of paranormal powers, siddhis, mind control, and the like, would therefore refute Darwinian evolution.

\paragraph{Old Gold}

I recently came across a large collection of notes and writings from around 1978 to 1985. I may put some of them online as examples of how thought develops. There were many of the same themes written about here, but I did not have the advantage of having read Rene Guenon at that time.


\flrightit{Posted on 2014-05-05 by Cologero}

