\section{The Herald Angels Sing}

Relying on the best of recent biblical exegesis, Fr. \textbf{Joseph Ratzinger} provides interesting details regarding the actual birth of Jesus in Bethlehem. He avoids the sentimentalism often surrounding the Christmas story, and instead brings out its essential meaning. Interested readers can consult the book. What I prefer to focus on here is the meaning rather than the events, which are familiar to all. In particular, God's revelation via angels brings out important details that are still helpful today. The method of this revelation was different for Mary, Joseph, and the Shepherds. These took the forms of command, dreams, and song, respectively, all forms of communication that bypass the lower intellectual mind and pass directly to higher centers.

We have already discussed the annunciation to Mary. This came in the form of a command, not a logical proof nor a discussion. Such an incredible message could only take hold in a sinless consciousness, aligned already with the will of God, and not beclouded by the “personal equation”. There can be no debate; only the free choice between obedience and rebellion.

\paragraph{The God of the Human Race}
As we saw in the myths of the origins of Rome, the idea of birth of a god-man from the mating of a god with a virgin was not at all unknown in ancient time. Even the ancient Greek cities had their spiritual origin from a man they revered as a god. The god, in all those cases however, was the god of a specific people and bestowed a particular identifying spiritual gift on them. However, as Fr. Ratzinger points out, such myths created a mixed being, a demi-god, whereas Christ was fully God and fully man, without confusion or mixture. Moreover, unlike the timeless quality of a myth, this story is determined at a specific place and time.

So the meaning of the birth of Jesus is different. The Holy Spirit, the father of Jesus, is the “giver of life” itself, not just one aspect of life. He is the God of the human race, not just a particular people. Nevertheless, such a possibility was known to a few people at that time:

\begin{quotex}
As to the god of the human race, a few philosophers had an idea of him; the mysteries of Eleusis might have afforded a glimpse of him to the most intelligent of the initiated; but the vulgar never believed in such a god. \flright{\textsc{ Fustel de Coulanges}, \emph{The Ancient City}} 

\end{quotex}
With the birth of Jesus, this God was revealed to all. The revelation to the vulgar will be described in this segment in regard to the shepherds. The revelation to the initiated will be the topic of the next segment on the Magi.

\paragraph{Horizontal and Vertical Heredity}
While Luke traces the genealogy of Jesus horizontally back to Adam, the first man, John indicates the vertical origin of Jesus as the incarnation of the Logos. Analogously, the birth of each human being is the result of both horizontal and vertical heredity. In horizontal heredity, the physical and psychic characteristics of the ancestors are transmitted to the descendants. In the vertical dimension, the intellectual soul, or individuality, is breathed into the body/soul.

This is explained more fully by \textbf{Valentin Tomberg} in Letter XX of \emph{Meditations on the Tarot}. Horizontal heredity operates by imitation of the ancestors. As Fr. Ratzinger points out, we are born into a “collective net” of our ancestors, tracing back to the very origins of the human race. The unraveling of this net requires the birth of the perfect man. Only then can we experience a second birth, as adopted children of God, which lifts us out of and liberates us from that net. We see this in John:

\begin{quotex}
To all who received him, who believed in his name, he gave power to become children of God, who were born, not of blood, nor of the will of the flesh, nor of the will of man, but of God. \flright{John 1:12-13}
\end{quotex}

\paragraph{Joseph's Role}
Since Joseph's role is often given short shrift, it is useful to focus on what Fr. Ratzinger writes about him:

\begin{quotex}
Whereas the angel “came” to Mary, he merely appears to Joseph in a dream … Once again this shows us an essential quality of the figure of Saint Joseph: his capacity to perceive the divine and his ability to discern. Only a man who is inwardly watchful for the divine, only someone with a real sensitivity for God and his ways, can receive God's message in this way. 

\end{quotex}
Obviously, what was revealed in Joseph's dream was quite difficult to believe. So how was Joseph able to discern God's will, even as revealed in a dream? That is because Joseph was a just, or righteous, man, and a just man is one whose life is lived in and from the word of God. Also, Joseph was “inwardly watchful”, that is, he monitored and guarded his thoughts. He was able to see past the trap of the collective net of human influences to recognize the divine influences that transcended them.

Of course, many people ask God for a “sign” to reveal His will for them. But are they just? Could they really recognize God's message, given their spiritual state. Or do they prejudge the sign so that only something consoling or beneficial is interpreted as a sign? Fr. Ratzinger offers this challenge:

\begin{quotex}
God is constantly regarded as a limitation placed on our freedom, that must be set aside if man is ever to be completely himself. God, with his truth, stands in opposition to man's manifold lies, his self-seeking and his pride. God is love. But love can also be hated when it challenges us to transcend ourselves. It is not a romantic “good feeling”. Redemption is not “wellness”, it is not about basking in self-indulgence; on the contrary, it is liberation from the imprisonment in self-absorption. 

\end{quotex}
The duty of the father is to name the son. This name, “Jesus”, was revealed to Joseph. It means “Yahweh is salvation”. The angel in the dream explains: “He will save his people from their sins.” Since this is the purpose of the incarnation, it is imperative to understand exactly what that means. Fr. Ratzinger explains:

\begin{quotex}
Man is a relational being. And if his first, fundamental relationship is disturbed—his relationship with God—than nothing else can be truly in order. This is where the priority lies in Jesus' message and ministry: before all else, he wants to point man toward the essence of his malady … if you are not healed there, then however many good things you may find, you are not truly healed. 

\end{quotex}
Clearly, it is not simply a matter of “doing good”, as Plato thought, or of avoiding certain behaviors. Rather, sin is the state of not being in proper relationship with God, the Logos.

\paragraph{The Shepherds}
Shepherds were in the fields around the region where Jesus was born, “keeping watch over their flocks by night.” Once again, we see the theme of watchfulness. The shepherds were not just outwardly close to the event, but “they were also inwardly closer to the event, unlike the peacefully sleeping townsfolk. … inwardly they were not far from the God who had become a child.” The watchfulness of the shepherds has become part of the monastic tradition.

When the angel appears to the shepherds, they were filled with fear. Just as the angel Gabriel reassured Mary not to fear, the angel likewise dispels the shepherds' fear. The spirit of fear is the opposite of the spirit of love.

Then a multitude of angels appeared praising God: “Glory to God in the highest, and on earth peace among men with whom he is pleased.”

Fr. Ratzinger makes this fascinating point:

\begin{quotex}
Christianity has always understood that the speech of angels is actually song, in which all the glory of the great joy that they proclaim becomes tangibly present. 

\end{quotex}
As \textbf{Julian Jaynes} documented in \emph{The Origin of Consciousness in the Breakdown of the Bicameral Mind}, speech in the forms of commands, music and poetry bypass the rational mind. Specifically, there is a direct revelation or intuition, without the duality of yes/no, faith/doubt, etc. Even among the Greeks, the muses sang the poetry, so that they literally heard the muses.

So the true experience of angels is not dualistic, and not even visual. The angelic song goes straight to the shepherds' hearts, not their heads.

\flrightit{Posted on 2014-12-24 by Cologero}

* * *

\begin{footnotesize}\begin{sffamily}

\texttt{X on 2014-12-24 at 23:50 said:}

“Clearly, it is not simply a matter of `doing good’, as Plato thought, or of avoiding certain behaviors. Rather, sin is the state of not being in proper relationship with God, the Logos.”

Sin is the state of not having fully internally realized Logos.


\hfill

\texttt{David on 2014-12-26 at 03:14 said:}

This series of article with Joseph Ratzinger are really good. I have alot to think and assimilate. Everything appear so simple yet it is so far reaching; and this yet again is only an illusion of the mind because of its relation to Being. Thank you for all your articles but those one really speaks. Eager to read the last one.


\hfill

\texttt{Q on 2014-12-26 at 08:09 said:}

“The spirit of fear is the opposite of the spirit of love.”

But fear of God is the beginning of wisdom, isn't it? How do you interpret that? I always simply looked on that as meaning that wisdom starts by fearing to tread on the wrong path.


\hfill

\texttt{Scardanelli on 2014-12-26 at 09:49 said:}

Q, I think the type of fear discussed above is one of closing oneself off to influences from above and clinging to the profane. Love on the contrary is opening oneself to the divine. The fear of God as the beginning of wisdom refers to fear as a type of humility before God. This type of fear is a self emptying that must take place if we are to receive wisdom.


\hfill

\texttt{Tom Blanchard on 2014-12-26 at 11:39 said:}

Our priest gave a very similar homily this Christmas Eve, specifying that the inner meaning of the Shepherd is watchfulness, illustrating the interior disposition necessary to hear the speech of the Angels. Quite likely he also took inspiration from Ratzinger's Infancy Narratives…: \url{http://www.anglicanphiladelphia.org/sermon/christmaseve-14.mp3}


\hfill

\texttt{Michael M on 2020-12-24 at 10:21 said:}

Interesting to note that it was a dream that came to St. Joseph, at night, but contrasting to other nightly revelations as specifically a dream vs. Nicodemus' talks, Mary's visitation, the Shepherds meeting and the Magi and the Star. Again he is told later to take the family to Egypt…through a dream, so perhaps we can see that for the Just, the Lord may speak to them there. Joseph was inwardly watchful and his dreams were not inundated with the incorrect tendencies and worries which most will give as the reason not to interpret dreams. Not something I've worked out entirely but given the next year (2021) declared to be a year dedicated to St. Joseph it may be time for these works to be contemplated more.

A small comment on the inner watchfulness of the Shepherds, as the Magi are to be addressed in a later post I will save that part when it comes up to contrast.

From a short lecture by Fulton Sheen, he compared the Shepherds to those that know nothing. I took this to mean a more natural, do nothing, know nothing, unconcerned with the world approach, somewhat `natural' or simple in that they live and subsist on one task which is their flock and develop a type of watchfulness that is scorned by those concerned with material things. Good, honest, and hard working folk, here angels speak to their hearts, for they are not centered in the mind but in the every day, every moment present of life. Song and emotion (pure) are tied together as it moves the heart and can be grasped through the heart outside of time unlike a purely `rational' way of understanding, the full Gestalt of it is different, much like the simple life of a monk which is mentioned above as well.

\end{sffamily}\end{footnotesize}
