\section{Salvation and Evolution}

At the conclusion to the Meditation on the Arcanum of the Star, \textbf{Valentin Tomberg} makes the following appeal:

\begin{quotex}
In our time, therefore, it is a matter of the task of effecting the third step of the evolutionary spiral of the Hermetic tradition --- the third ``recovery'' of the subject of the Emerald Tablet. Our time makes an appeal to the collective efforts of Hermetists of today to make a third summary, which will be for our time what the Tarot was for the Middle Ages and what the Emerald Tablet was for antiquity... This is why a new, modern summary is required, which will be as viable as the Emerald Tablet and as the Major Arcana of the Tarot.
\end{quotex}

With this, Tomberg exhorts us to move past the Tarot and to reformulate Hermetic wisdom in a contemporary way. Nay, it is not him but our Mother who exhorts us to our duty to the future. That is the message of the woman in the Star:

\begin{quotex}
It is she who is the Mother of the Future, and this is why her message confronts us with duty towards the future---the duty of the river of the unbroken Tradition. It is necessary for us to make efforts to comply with this!
\end{quotex}

In this Letter Tomberg relates Salvation to Evolution. Salvation is from the Father and tells us ``\emph{what}'': the Mystery of Salvation through the Son. The Mysteries of the Mother tells us ``\emph{how}'': through biological and spiritual evolution. That is, salvation comes from above, the eternal, the timeless, while evolution is the projection of salvation onto the dimension of time. In the following passage, Tomberg gives us a number of predecessors whose collected efforts teach us the synthesis of Salvation and Evolution.

\begin{quotex}
But let us not forget that this synthesis of today has had its history, and that this is due also to ``labor pains''. It was born after a long series of continuous efforts from century to century:

\begin{itemize}

\item the effort of \textbf{Heraclitus}, the philosopher of the perpetual change of matter;

\item that of the \textbf{Gnostics}, who made the drama of the fall and return of \textbf{Sophia Achamoth} resound in human history;

\item that of \textbf{St. Augustine}, the father of the philosophy of history, who brought to light the twofold current in mankind's history---the ``City of Man'' and the ``City of God'';

\item that of the alchemising Hermetic thinkers who affirmed and re-affirmed untiringly the principle of the transformability of the base into the noble;

\item that of \textbf{Martinez de Pasqually}, who wrote his \emph{Traité de la reintegration des êtres} (``Treatise on the Reintegration of Beings'');

\item that of \textbf{Fabre d'Olivet}, the author of \emph{L'histoire philosophique du genre humain} (``Philosophical History of the Human Race''), showing the dynamic operation of the triangle destiny-freedom-providence in mankind's history;

\item that of \textbf{H. P. Blavatsky}, who added and opposed to Charles Darwin's materialistic evolution a breath-taking vision of the spiritual evolution of the universe;

\item that of \textbf{Rudolf Steiner}, who emphasized that the center of gravity of spiritual-cosmic evolution -- to know Jesus Christ -- is also not far from \textbf{Teilhard de Chardin}`s ``Omega point'');
\end{itemize}

All these efforts have contributed ---in a visible or invisible manner ---to the synthesis of today. They live, all together, in the contemporary synthesis of evolution and salvation, which is the fruit of this collective effort from century to century.
\flright{\textsc{Valentin Tomberg}, \emph{Letter XVII: The Star}, corrected translation}
\end{quotex}

\paragraph{The Philosophical History of the Human Race}

The temptation is to get ``stuck'' on the Tarot, collecting different versions, and so on, whereas our real task is to understand, develop, explain, and expound on the contents of the Letters. To that end, this post will focus on Fabre d'Olivet's contribution to the synthesis in his \emph{Philosophical History of the Human Race}. Need I repeat here that Tomberg regarded this history as the most lucid he had encountered?

Note that his is a ``philosophical'' history, not an exact sequential reproduction as profane historians try to write. Instead, he relies on his powers of Creative Imagination, which Tomberg describes as reading the Akashic record. There are many theories of history, e.g., the Marxist, the Freudian, racial, materialistic, and so on, but Fabre relies instead of his precise understanding of the three Forces that actually motivate the world process, which we previously described in Providence, Will, Destiny\footnote{\url{https://gornahoor.net/?p=8116}}.

\begin{itemize}

\item Destiny

\item Will

\item Providence
\end{itemize}

Corresponding to those macrocosmic forces, there are these three spheres of life:

\begin{itemize}

\item Instinctive (necessities): sensation, instinct, common sense

\item Animistic (passions): sentiment, understanding, reason

\item Intellectual (inspirations): assent, intelligence, sagacity
\end{itemize}

These spheres are not fully developed in all human beings. Fabre describes the evolution of man as the progress from the instinctive life based on sensation and motivated by attraction or fear, through the animistic life, and finally to the fully human life of the Intellect or Intuition. Of course, that is not the same as discursive thought which is still in the animistic sphere.

However, the development of the intellect cannot happen automatically, mechanistically, or deterministically. Fabre d'Olivet explains:

\begin{quotex}
Nothing was made in advance with him, although everything was determined there in principle. Providence, whose work he was, willed that he should develop himself freely and that nothing should be forced in him.
\end{quotex}

\paragraph{The Coming Deluge}

With this esoteric key, Fabre explains the development of Marriage, social and political structures, commerce, justice, and so on. In an upcoming post, we will summarize his main points. The Deluge is coming. This is the loss of Tradition and the forgetfulness of esoteric or Hermetic knowledge, which is leading us backwards toward our animal and instinctive natures. The collective ignorance of the human race regards that as ``progress'', the liberation from social and religious structures that have sustained the human race, bringing Salvation to many.

Signs of the Deluge: Instead of justice, right is established by force. In place of impartial knowledge, there is a battle of opinions, each held with passionate intensity. The Emerald Tablet saved the essence of ancient Wisdom, and the Tarot saved the essence of Medieval Wisdom. The task of preserving modern wisdom requires a new ``Noah's Ark'', whose form is yet to be determined.


\flrightit{Posted on 2016-07-10 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Comment on 2016-07-12 at 17:53 said:}

But ``the psyche of Adam and Eve before the fall, in the terrestrial Paradise of the Garden of Eden . . . was in direct contact with God, since the purity of their lower emotional centre allowed them to maintain permanent contact with the higher emotional centre and through this with the higher intellectual centre.''

(Boris Mouravieff)

How could the Traditional description of the Golden Age above be reconciled with the evolutionist idea presented here? Mouravieff provides the synthesis. He writes that Adamic man ``had a higher emotional centre and a higher intellectual centre. But he still lacked the lower intellectual centre, of which he had a most urgent need for the transformation of matter, and particularly to cultivate the soil''. (I guess we shouldn't expect to find Atlantis anytime soon).

However, when he evolved that lower ``critical mind which accompanied the notion and desire for acquisition'', the consequence was that `` his lower emotional centre hardened. In this way it lost direct contact with the higher emotional centre, that is, with the real I, the contact which made Adam and Eve the children of God.'' This would then be how the Kali Yuga started.

\hfill

\texttt{Mark Citadel on 2016-07-17 at 16:53 said:}

Astounding and thought-provoking. I really must pick up some of Tomberg's stuff. Whatever ark is built to preserve this millenia-compounded wisdom, I would wager it will need to be tougher than any previous esotericism, as we could be looking at something far more earth-shattering in scope and magnitude than had previously been experienced.

The Greater Malthusian Trap, the Olduvai Cliff, we are racing headlong to some kind of rapid, almost instantaneous detonation. What awaits beyond will be only for the eyes of those who have honored the Tradition. The rest will be cut away like so much rotted meat, As always, the Father has given them every opportunity to correct themselves and they have stubbornly refused. They do not understand evolution and they certainly do not understand salvation.


\hfill

\end{sffamily}
\end{footnotesize}

