\section{Dreaming in the Dog Days}

\begin{quotex}
But when the artichoke flowers, and the chirping grasshopper sits in a tree and pours down his shrill song continually from under his wings in the season of wearisome heat, then goats are plumpest and wine sweetest; women are most wanton, but men are feeblest, because Sirius parches head and knees and the skin is dry through heat.
\flright{\textsc{Hesiod}}
\end{quotex}

In the midst of a restless sleep during a stultifying dog day night, following a massive hurricane, I was entertained by a dream. As I pulled into my driveway, I noticed my two closest college friends with a young boy waiting for me. Although they were my friends, they were not friends with each other. On the contrary, each represented a different, mutually incompatible, trajectory of my life. B. had been my friend since grammar school. As such, he represented the conserving element of my past, the drag of inertia. R., on the other hand, was a new friend; together we explored the lifestyle choices of the social upheavals of that era.

Being a trickster, I resolved to deceive them. Pretending not to know them, I casually asked if they had been sent to clear my windshield. To my surprise, they did not recognize me. R. was suspicious and went into my car to look for clues regarding my real identity.

Following him, I found a text and opened it. A secret was revealed to me about the Tarot. I learned I was not in the land where the centre is everywhere. I read the next paragraph, but had trouble understanding it. Multi-coloured balls were dancing around the final word. It was somewhat confusing and I wondered how they could possibly know that to be true. Unfortunately I cannot now recall the word. I suddenly awoke, and the ideas for this essay rushed into my head; I'm afraid I will leave out something essential.

\paragraph{The Crown of Creation}
According to Valentin Tomberg, the Minor Arcana represent an extension to the final Major Arcanum, The World. As such, they represent an ascent, known as Jacob's Ladder\footnote{\url{https://www.meditationsonthetarot.com/psychurgical-practice}}.

\begin{quotex}
The Minor Arcana of the Tarot represent the way of ascent from consciousness belonging to the world of action (the phenomenal world) through the world of formation and the world of creation to the world of emanation.
\end{quotex}

The suit of Pentacles corresponds to Assiah, or the World of Action. This is the mixed, or dualistic world. Not fully God's creation, it is the result of Adam's fall. It is ruled by Satan, the Prince of the World. Tomberg explains:

\begin{quotex}
The world of our experience is the phenomenal manifestation of both the world created by the Word and the evolutionary world of the serpent.
\end{quotex}

This world is the world of images, intellectual constructions, and imagined ideals. However, these are not liberating, but rather keep humanity enthralled. Boris Mouravieff describes it this way:

\begin{quotex}
Dreamlike imagination, which is somnolent, comes from the Absolute III, and is also found to a certain degree in animals: it is a passive force. The latter form of imagination, the `dream of the sleeping serpent', produces a hypnotic effect on man, keeping him in the state in which the vast majority of humans pass their lives.
\end{quotex}

Tomberg agrees:

\begin{quotex}
It is the Serpent who is the prince of this world, and who is the author and director of the purely biological evolution following the Fall.
\end{quotex}

Mouravieff is even clearer on this point.

\begin{quotationx} 

Satan is the Regent of organic life on Earth within the limits of the mental and physical life of the three kingdoms of nature: flora, fauna and man. He dominates them by the action of sexual energy in the four phases of life: conception, birth, growth and finally reproduction. This sexual energy is the dynamic expression of his Love, which is carnal Love, while the spiritual life, which is accessible to man through the mediation of his higher centres, remains outside the sphere of the World of Action.

His task is thus limited and he carries it out, as we said, by the sexual Love which emanates from him, by the force of attraction between the sexes, and by Illusion, the dreams of the sleeping young Serpent. The current of his Love acts like a drug on both man and woman and, in its different forms, keeps them in a hypnotic state from puberty to the end of their days.
\end{quotationx}

\paragraph{The Sephiroth}

For preparation, the first step is to identify the qualities of the Sephiroth as they appear in the World of Action. The loss of ``vertical consciousness'', i.e., direct consciousness of the Divine, colours our experiences. By forgetting the vertical, the World of Action seems to be self-sufficient and independent of any transcendent foundation. Thus, the philosophy of naturalism -- along with its children materialism, positivism, and atheism -- seems a reasonable way to understand this world. Hence, some of the sephiroth might be described this way.

\begin{itemize}

\item \textbf{6 Beauty}. Beauty is sensual and the images of beauty keep us in a dreamlike state. The beauty of nature ties us to the desire for experience. The beauty of people excites the desire for biological evolution, even past the time when that makes any sense.

\item \textbf{5 Justice}. Justice is the restoration of equilibrium. So in a dispute, Justice would resolve the two points of view with reference to the Will of God. In the World of Action, humanity replaces God, so that man determines what is good or evil. Since there is no belief in transcendence, Justice devolves into a power struggle between the parties.

\item \textbf{4 Mercy}. In vertical consciousness, Mercy is oriented toward the supernatural end of man. When that awareness is lost, Mercy is oriented toward man's natural goals.

\item \textbf{3 Intellect}. The Intellect is dominated by Science. The aim of the Intellect is not to know higher things, but rather to provide the understanding to manipulate nature.

\item \textbf{2 Wisdom}. There is no true Wisdom. Rather, cunning, slyness, and artfulness are considered to be the hallmarks of Wisdom.
\end{itemize}

Of course, a meditation on the Minor Arcana would have to develop similar ideas more deeply. The pip cards correspond to the ten sephiroth.

What about the Crown of the World of Action, the Ace of Pentacles? For most of humanity, life in the Assiah is a difficult struggle. It is bounded by toil, disease, pain, and death. Even during the good times, their presence casts a shadow over life, and they seem to impinge at the most inopportune times.

Nevertheless, at any time there are those who seem to have mastered life. They have physical beauty, sex appeal, intelligence, ability, health, fame, and, of course, sufficient fortune so that the cost of anything never needs to be considered. They become the objects of envy. I'm sure you can name many such people, but I'll mention two quite different men, if you consider them objectively: George Clooney and Donald Trump. And there are those who suddenly appear, almost out of nowhere, like Jennifer Lawrence.

Everyone desires that life of leisure, and their images are everywhere. They keep us in a hypnotic state, since very few will ever meet such success. In the Mebes Tarot system\footnote{\url{https://en.wikipedia.org/wiki/G._O._Mebes}}, the Ace of Pentacles represents Wealth, both Inner and Outer. Outer wealth is visible and comprehensible to the masses.

Inner wealth signifies what must be given up in order to climb the ladder. Idle wealth is useless, rather it must be what is paid. Now, the Crown of the World of Action is related to the Tiphereth of the World of Formation. That latter is represented by the 6 of Swords, which symbolises the Science of Success.

\paragraph{The Seven Sleepers}

This morning, I was dreaming about the Seven Sleepers\footnote{\url{https://en.wikipedia.org/wiki/Seven_Sleepers}}. In the dream, I was communicating via thought about how we would know when to wake up together. Then we all woke up.


\flrightit{Posted on 2017-09-20 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{james on 2017-09-21 at 02:49 said:}

Thanks for another great post.

\hfill

\texttt{Arthur Konrad on 2017-09-22 at 19:19 said:}

It appears that a gentleman is only faithfully perceived as such if he does not have to labor for his wealth. It seems that for this reason the aura of a gentleman persistently eludes the self-made men.

Then, just the day prior to reading this, after experiencing unprecedented level of concentration while meditating, I came to suddenly think how superficially subconscious, almost petty the dreams are. Perhaps I'll be reconciled with them

Nature dominates nature, and if the lower nature relents, it might be reintegrated in the world of new, endless possibilities

\hfill

\texttt{Max on 2017-11-26 at 20:34 said:}

One wonders where final words go... 

While thinking about spleen there were some lines by Baudelaire mentioning dog-days.. .

I do not know what it all means.

`Where do the dogs go, you ask, you inattentive men.' They go about their business.

Business meetings, amorous meetings. Through the fog, through the snow, through the mud, in the burning dog-day sun, in the rustling rain they go, they come, they trot, they pass under coaches, impelled by fleas or passion, by need or by duty. Like us, they get up early in the morning and are trying to make a living or chasing after their pleasures.''

I guess some good old ``philospher dogs'' have simply seen too much -- not much impressed by what is on offer by the world any longer.

Most dreams are ridiculous anyway. What is strange is that we believe them without reservation, and how well-crafted they seem. Must be the work of some unknown genius. Perhaps we will meet.

\end{sffamily}
\end{footnotesize}

