\section{Sophia and the Eternal Feminine}

\begin{quotex}
\begin{verse}
Do you remember the roses over white foam,\\
A purple reflection in azure waves?\\
Do you remember the image of a beautiful form,\\
Your confusion, and trembling, and dread?
\end{verse}
\flright{\textsc{Solovyov}, \emph{Das Ewig-Weibliche}}

Das Ewig-Weibliche zieht uns hinan.

\flright{\textsc{Goethe}, \emph{Faust}}

\end{quotex}
It is not a coincidence that, starting sometime in the 19th century, visions of Sophia or Mary became more frequent. We will use \textbf{Vladimir Solovyov} as the expositor of the Sophianic current in the Eastern churches and Fatima as an example from the Roman church.

\paragraph{Solovyov}
Like \textbf{Dante} and \textbf{Hafiz} before him, Vladimir was precocious in having become aware of the Anima at the young age of 10. Fortunately for him — as it was for both Dante and Hafiz — his love for the young girl was not reciprocated. Emotionally devastated by rejection, he had his first of three encounters with Sophia:

\begin{quotex}
He perceived suddenly the blue of the sky all around him and in his own soul; and through this blue he saw the Eternal Womanhood, woven as it were of blue ether, and holding in her hand a blossom form unearthly countries. She nodded in his direction, smiling at him with a radiant smile, and then disappeared in the mist. His soul became blind to all worldly things, and the childish love which caused his jealousy became repugnant to him. \flright{\textsc{N. Lossky}, \emph{History of Russian Philosophy}}

\end{quotex}
Perhaps had Vladimir and the girl become childhood sweethearts, he would have ended up in the bourgeois life of a parish priest. But his rejection of carnal life set the course of his life. In answer to the question, ``Have you ever been in Love", Solovyov responded:

\begin{quotex}
Seriously, once; otherwise, 27 times. 

\end{quotex}
From that, it seems clear that he had multiple flirtations over the course of his life, but that his real love was for Sophia.

Solovyov died of kidney disease in 1900.

I recently rid myself of \#27. I then received a blessing from Mary while meditating on her Coronation in Heaven.

I will probably die of kidney disease if COVID-19 doesn't kill me first. But \#28 might save me.

\paragraph{Path of the Initiate}
There are two esoteric paths:

\begin{enumerate}
\item \textbf{The way of the Monk} requires the renunciation of life experiences. This makes it one-sided. 
\item \textbf{The way of the Initiate} is the more difficult path. The initiate uses life experiences to further his growth. This gives him knowledge of all the forces exerted in the world. Ultimately, it is a solitary path. 
\end{enumerate}
Although monks tried to convince Solovyov to join their monastery, he regarded that path as a form of slavery. He ultimately chose the solitary path of the Initiate. This allowed him to freely choose his own course of study. Eventually, he went to England to study Indian, gnostic, and medieval philosophy. Yet, his work was not merely intellectual and academic; his life was punctuated with spiritual experiences.

In the British Museum, he experienced his second encounter with the Divine Sophia, who told him to go to Egypt. His third encounter took place in the desert, 20 kilometres outside Cairo. His dream life was significant for him. For example, in the dream state, he would envision prophecies and engage in conversations with the dead. Many passages in his notebooks include the fruits of automatic writing.

\paragraph{Solovyov's Predecessors}
Solovyov was much more than a theologian, or philosopher, or even a mystic. It took another Initiate to recognize his real project. Valentin Tomberg made this insightful observation:

\begin{quotex}
[I] had never encountered a work written before the time of Rudolf Steiner that contained such a profound concept of the nature and mission of Jesus Christ, a view presented against the background of cosmic history. … A question now arose: How did Solovyov come to this amazing cognition of the more profound mysteries of the Son and even the divine Trinity as a whole? In studying the works of Solovyov, we find no actual path or method leading to this perceptive knowledge. We can indeed follow his thoughts and thereby be convinced of their truth, but the works of Solovyov offer no explanation as to how we can reach similar new knowledge. \flright{\textsc{Valentin Tomberg}, \emph{Christ and Sophia}}

\end{quotex}
That may have been true of his \emph{published} works. Keep in mind that the Russian religious censors were quite strict. The memory of the fate of Konrad Nordermann and Quirinius Kuhlmann, who brought the ideas of Jacob Boehme to Russia a century before, was not forgotten. They had been severely tortured and burned alive by the authorities.

Hence, Solovyov was careful with his writing. He had to couch the gnostic concept of syzygy in the \emph{Meaning of Love} in more orthodox terms. His promise to write a book on the influence of Hermes Trismegistus on the Church Fathers was never fulfilled. His controversial books were written in French and published in France.

Nevertheless, his notebooks—which were not available during Tomberg's lifetime—provide the full sources of his ideas. For example, in \emph{Theological Principles}, he includes this schema

\begin{table}[h]\centering\small
\begin{tabular}{rcc}\toprule
Neoplatonism and Kabbalah &
Law &
Old Testament\\
Boehme and Swedenborg &
Gospels &
New Testament\\
Schelling and Solovyov &
Freedom &
Eternal Testament\\\bottomrule
\end{tabular}
\end{table}
It is easy enough to recognize in this the idea of the Cistercian \textbf{Joachim de Fiore} on the Three Ages: the Age of the Father, the Age of the Son, and the Age of the Holy Spirit. Solovyov, thus, regarded himself as the fulfillment of the process begun in Neoplatonism and the Kabbalah, continued through Boehme and Swedenborg, then though Schelling. Hence, he brought Absolute Idealism into his system, expanding on Schelling, particularly on the idea of Freedom. In other words, he claimed knowledge of the Absolute. He also developed his ideas of cosmogony and theogony from Schelling; this is what Tomberg was referring to. (These topics will be discussed in an upcoming post.)

So, we do indeed recognize the path leading to Soloviev's knowledge. As a matter of fact, what was only implicit in Solovyov's published works—viz., Neoplatonism, Kabbalah, Boehme, for example—was developed openly by Tomberg himself. He didn't need to concern himself with the censors.

\paragraph{Dialogue with Sophia}
Employing a technique which we have recommended several times, Solovyov engaged in a dialogue with the Anima. The manuscript was titled, \emph{The Sophia}, which Solovyov described as a ``mystical-theosophical-Philosophical-Theurgical-Political" dialogue. Although it was never published in his lifetime, Solovyov regarded it as the foundation of all his later works. It was not simply a rhetorical device, since the manuscript included Gnostic symbolism and contained passages of automatic writing. (Automatic writing comes directly from the subconscious or from a supernatural source. The Enochian language was revealed to John Dee in this way.) Instead, it must be understood as a revelation of Sophia.

\paragraph{Essence and Existence}
In the manuscript, Sophia's dialogue partner is the Philosopher, the ``lover of Sophia". Although the Philosopher instantly recognizes Sophia, she is not simply a projected part of his unconscious; she is an independent being in herself. Of course, Solovyov knows Sophia from his three visions.

Nevertheless, Sophia questions the philosopher's idea of knowledge. All traditional wisdom recognizes the two worlds:

\begin{enumerate}
\item the metaphysical order, the world of Being, the invisible and intangible world, the ideal world 
\item The physical order, the world of Becoming, the visible and tangible world, the real world 
\end{enumerate}
In Orthodox theology, this shows up as the distinction between the essence and energies of God. The energies of God are known through his activities, but His essence is inexhaustible, and cannot be fully known. In the West, the word for energies (erg) is translated as act (actus). So actual grace, for example, is experienced as God's energies. When the philosopher claims knowledge of the Absolute, Sophia questions him. The following snippet is instructive.

\textbf{Sophia}: Do you know me? Do you know with whom you are speaking?

\textbf{Philosopher}: As if I could not know you!

\textbf{S}: You no doubt know me as a phenomenon, that is, insofar as I exist for you or in my external manifestation. You cannot know me as I am in myself, that is, my thoughts and intimate feelings as they are in me and for me. You know them only when they manifest themselves outwardly in the expression of my eyes, in my words, and my gestures. These are only external phenomena and yet …

\textbf{P}: And yet, when I look into the deep azure of your eyes, when I hear the music of your voice, is it outward phenomena of sight and sound that I perceive? My God! I know your thoughts and feelings, and, by your thoughts and feelings, I know your inner being.

\textbf{S}: And this is the way that all beings know each other. You know the interior phenomena by way of the exterior, and, by that, the being—itself, or what one philosopher [Kant] has called the intelligible character [noumenal or thing-in-itself].

Solovyov is describing ``spiritual vision", that is, the ability to discern the spiritual world beyond the world of phenomena. It is the ability to understand the inner state of another person (which in pop culture is mistakenly called ``mind reading").

\paragraph{Metaphysical Principles}
Not surprisingly, Sophia revealed to Solovyov some metaphysical teachings that are compatible with Traditional metaphysics, as in \textbf{Rene Guenon}'s \emph{Multiple States of the Being}, for example. As was just mentioned, there is the distinction between Essence—the principle of unity—and Existence—the principle of multiplicity.

Sophia points out that the principle of the Absolute is not Being, but rather Nonbeing. She points out:

\begin{quotex}
[The Absolute] can be neither the direct object of our outer sensations, not the direct subject of our inner consciousness, for everything in these two categories is what we call being. 

\end{quotex}
The principle of Being cannot be Being, since it would be a being outside of all being. Hence, the Absolute principle is that which has the possibility of being; in other words, it is Infinite possibility. Yet the absolute cannot lack anything, so it must include the opposite of itself. That is,

\begin{quotex}
In order to be truly absolute and infinite, it must also be the principle of relativity and finitude. 

\end{quotex}
At this point, the Philosopher admits that this understanding brings him to a state of mystical enthusiasm.

Sophia explains that is because the absolute principle is Love. In expressing itself as relative and finite, that is a form of self-denial and the affirmation of the other, the very definition of Love.

\paragraph{The Universal Religion}
Sophia revealed the universal religion, which she described as:

\begin{quotex}
The universal religion is the fruit of a large tree, whose roots are formed from primitive Christianity and whose trunk is formed from the religion of the Middle Ages. 

\end{quotex}
Solovyov wrote in the margins:

\begin{quotex}
The universal religion is not only the positive synthesis of all religions, but also the synthesis of religion, philosophy, and science. 

\end{quotex}
As the positive synthesis, Solovyov wants to avoid the notion of a ``mere Christianity", which is a negation. On the other pole, he has to avoid a New Age type syncretism or a vague religion of humanity. This idea was most fully developed in one of his last books, \emph{Russia and the Universal Church}, which had to be published in French, for obvious reasons.

Solovyov denied that the mystical bond of the Roman and Orthodox churches had ever been severed. His biographer explains:

\begin{quotex}
In the light of the ``fire of Truth and Knowledge" that lived in Schelling's works, Soloviev acquired a new awareness of the implications of the Johannine aspect of Christian revelation and mystical experience. With the great German philosopher, he came to recognize that ``the true Church which has its basic foundation in Peter, goes through Paul, finally to become the Church of St. John". \flright{\textsc{Paul M. Allen}, \emph{Vladimir Soloviev: Russian Mystic}}

\end{quotex}
Over the years, Solovyov had become closer to the Roman Catholic church, even taking its sacraments. However, as much as the partisans would like to one side or the other, this could not be considered a ``conversion", since that would make no sense in his understanding.

Rather, he envisioned a Universal Church, that would include both the Catholic and the Eastern churches, although it would be centered in Rome with the Pope as its head. Solovyov had pointed out in his lectures that Reason was ineffective against the passions and the party spirit. Hence, a millennium after the schism, passion and sectarian interests continue to hide the mystical bond from most people. If Reason is useless, the perhaps the spirit of Love, the principle of the Absolute, can end the separation.

\paragraph{Filioque}
A short note on the Filioque is appropriate in the context, since, for the most passionate, it is a fundamental issue. \textbf{Sergei Bulgakov}, one of Solovyov's closest followers, has a long discussion in \emph{The Comforter}, his book on the Holy Spirit. He asserts that the Filioque is not a cause for separation, because its real meaning still needs to be clarified. He wrote:

\begin{quotex}
These dogmatic divergences must not serve as a pretext for mutual accusations of heresy, inasmuch as they can be harmonized or at least considered not to exclude one another. Not imaginary heresy, but the schismatic spirit and false dogmatism sustain that separation of which the ancient church was free. 

\end{quotex}
Unfortunately, the most rabid and partisan spirits seem to find all sorts of bizarre consequences that follow from the Filioque. On the other hand, Bulgakov, who is a more sober and deeper thinker, cannot actually find any meaningful practical impact:

\begin{quotex}
In practice, the two sides West and East, do not differ in their veneration of the Holy Spirit, despite their divergence regarding the procession. It seems highly strange that such a major dogmatic divergence would have no practical impact whereas normally a dogma always has a practical significance, determines the religious life. Even the most extreme zealots of the schismatic spirit have not been able to apply the imaginary dogma to life or to indicate its practical consequences. 

\end{quotex}


References:

\emph{Vladimir Soloviev: Russian Mystic,}\textbf{ Paul M. Allen}

\emph{Divine Sophia: The Wisdom Writings of Vladimir Solovyov}, \textbf{Judith Deutsche Kornblatt}

\emph{History of Russian Philosophy}, \textbf{N. Lossky}

\emph{Christ and Sophia}, \textbf{Valentin Tomberg}



\flrightit{Posted on 2018-07-18 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Arthur Konrad on 2018-07-19 at 06:25 said: }

The way of the monk is not normally an esoteric path, quite the opposite. One-sided renunciation misses the point, which is contemplation of one's desire, which is an inner force, and not the renunciation of physical objects themselves for their own sake, which represents exterior and in itself insignificant morality.

Marian visions are mostly suspicious and subversive. They rest somewhere on the plane of mediumship, that is to say – it is enough that one becomes obsessed with something for his subconscious to oblige him, but that is counter-productive and is a form of spiritual passivity. But more than anything else, it leads one astray of proper contemplation and realization of the subtle vital and creative force symbolized by the eternal feminine, which is again – subtle and profound and not reducible to mere visions and pleasant sensations. ``Dialogues" with various spiritual entities in particular fall in this sensual domain. 

Without aiming to promote partisan discussions, but from this spiritual opus seems to be derived much of what came to be Christian mysticism and typically Christian sensuality (of which Solovyev is a proper foot soldier) and is almost a hallmark of its mode of being. 

Too much speculative, fantastical and allegorical, all in all – Christianity through and through. 

P.S. Remember that Buddha, in his conversation with Mara, did not allow himself to be deceived.


\hfill

\texttt{Cologero on 2018-07-27 at 00:42 said: }

Mr. Konrad, we have been describing in the past few posts the phenomenology of the encounter with the ``eternal feminine". Did not the Buddha, as you described, have to pass through similar experiences? There is no virtue, nor value, in the inability to be tempted at all. That is the way of the monk, who hopes to avoid such temptations in his solitude; that ultimately is ineffective.

Even St. Anthony was tempted; there is nothing particularly Buddhist about it.

Yet, the man who is not deceived will make his choice willingly and consciously, but he still makes the choice. In your account, you left out the notion of sacrifice. See for example the choices made by Zanoni and Mejnor. Zanoni made the sacrifice.

The Buddha, unlike the mass of men, was not deceived by the attraction of mortal life. Nevertheless, a Buddha rejects Nirvana and reincarnates willingly and deliberately.


\hfill

\texttt{Joe on 2020-07-21 at 23:09 said: }

A reference to COVID-19 in 2018–is this a vindication of Marian visions?


\hfill

\texttt{Cologero on 2020-07-21 at 23:29 said: }

Joe, you caught me red-handed re-editing old posts. So it is neither a prophetic vision nor the Mandela effect. But congrats for being such a careful reader.


\end{sffamily}\end{footnotesize}
