\section{Death by a Kiss}

\begin{quotex}
Modus quo rationales animae per archangelum Deo sacrificantur, qui a Cabalistis non exprimitur, non est nisi per separationem animae a corpore, non corporis ab anima nisi per accidens, ut contigit in morte osculi, de quo scribitur praeciosa in conspectu domini mors sanctorum eius. \flright{\textsc{Giovanni Pico della Mirandola}, \emph{Cabalistic Conclusions XI}}

The way in which rational souls sacrifice themselves to God through the archangel, which the Cabalists do not describe, occurs because of the separation of the soul from the body, not of the body from the soul, if not by accident, as seen in death by a kiss, about which is written ``Precious in the sight of the Lord is the death of his faithful." [Psalm 116:15]. [My Translation]

Let him kiss me with the kisses of his mouth: for thy love is better than wine. \flright{\emph{Song of Solomon 1:2}}

Whosoever would [behold God] before they are cleansed and healed, are so stricken by that light of truth that they see in it not only no goodness, but even much evil, and therefore deny to it the name of truth, and with their lusts and miserable pleasures, refusing health, flee away in their darkness, which is even their death. \flright{\textsc{St Augustine}}

\end{quotex}
Before there was a French Hermetic Tradition, there was the Platonic Academy\footnote{\url{https://en.wikipedia.org/wiki/Platonic_Academy_(Florence)}} in Florence. In particular, Giovanni Pico della Mirandola first synthesized Christianity, Hermes Trismegistus, and the Cabala, along with Neoplatonism, Aristotle, Pythagoras, Orpheus, and Zoroaster.

He summarized those teachings in the 900 theses. Some of which were adapted from Latin philosophers and theologians, Arabs, Greeks, Neoplatonism, Chaldeans, Hermes Trismegistus and finally the Cabala. The rest were his own opinions. The latter includes the mystical Cabalistic Conclusion \#11.

Plato said that perfect or undying love required divesting oneself of human life by separating the soul from the body. Yet an understanding of such love requires first the love of beauty in one's youth. In the story of Eurydice and Orpheus, Orpheus was unwilling to part with his life and therefore saw only a shadow of his beloved. Hence, Orpheus failed Plato's test. Pico explains it this way:

\begin{quotex}
Corporeal and sensible beauty can excite in the soul a memory of its intellectual part and cause it to run away from the earthly life towards the eternal, where it is refined into an angel by the flame of love. 

\end{quotex}
Pico understood the three worlds:

\begin{itemize}
\item Angelic or Intelligible 
\item Celestial or Astral 
\item Sublunar (world of darkness) 
\end{itemize}
The sublunar world is known through the senses, but the angelic world is known by the intellectual part of the soul. That is because the angels don't have sense experiences, but know the forms or ideas intuitively. For example, in the sublunar realm we may experience heat through the senses, but the angels know the idea of heat through the intellect, and so on.

Only the world of sensual experience has reality for the common man. He does not experience the ideas directly, which are obscure or are considered to be mental or social constructs. As Augustine points out, the common man clings to his lusts, pleasures, and manufactured drama. The call for spiritual purification would destroy his factitious world, so he resists it in terror. It can only feel like death to him. Yet the Hermetist seeks this out deliberately. Pico describes the experience:

\begin{quotex}
It is possible then through the first death, which is merely the separation of the soul from the body, for the lover to see the beloved celestial Venus and, face to face with her, meditating on her divine image, blissfully nourish his purified eyes; but whoever wishes to possess her still more intimately and, not content with seeing and hearing her, to be worthy of intimate embraces and fervent kisses, must separate himself in total separation from the body through the second death. 

\end{quotex}
In other words, the erotic impulse must be purified of every sensual image in order to reach the beatific vision. He must die before he dies. This Platonic love still idealises the image the beloved, yet the sensual element is left behind.

\textbf{Baldassare Castiglione} asserted that sensual love is an evil, yet it is excusable in the young provided they show gentleness, courtesy, and valour. He writes:

\begin{quotex}
When they are no longer of youthful age, they wholly abandon it, having that sensual desire, as the lowest rung of the ladder by which man climbs unto the true love. 

\end{quotex}
Castiglione regarded the love of the 63-year old Michelangelo for the poetess Vittoria Colonna as an example of Platonic love. We have previously described how the loves of Wagner and Goethe at an advanced age led to periods of great creativity. Of course, Dante regarded Beatrice as his beloved celestial Venus, and was led into the celestial realm.

The kiss represents the union of the Soul with God. Pico was more modest than the Tibetans who used the image of the yab-yum. He explains why:

\begin{quotex}
Take note that the most perfect and intimate union that the lover may have with the celestial beloved is signified by the union of the kiss, because all other coming together or coupling beyond that, as in the case of physical love, is in no way allowed to be used as a metaphor in this holy and most sacred love. And because the learned cabalists maintain that many of the ancient patriarchs died in such an intellectual rapture, you will find in their writings the expression the death of the \emph{binsica}, which in our language means death by a kiss, which is said of Abraham, Isaac, Jacob, Moses, Aaron, Mary , and some others … nor will you read more in their books than that \emph{binsica}, or death by a kiss, occurs when the soul unites itself with those things separated from the earth in an intellectual rapture to such a degree that being lifted out of the body it abandons it entirely. 

\end{quotex}
What then is this binsica, mors osculi, death by a kiss. Ioan Couliano illustrates it with some examples:

\begin{quotex}
It is a terrifying vision of the intelligible world that Pico rediscovers in the fable of Tiresias: because he saw Diana naked, which means nothing other than ideal Beauty, the source of all true wisdom, Tiresias went blind, losing the use of his sensual sight but receiving the gift of prophecy, or incorporeal sight. The same thing happened to Homer in the throes of the inspiration that made him contemplate the mysteries of intellect. And Paul too, after his journey to the third heaven, went blind. 

\end{quotex}
The death by the kiss is therefore the contemplation of the divine intelligences. The body is in a state of catalepsy analogous to sleep paralysis\footnote{\url{https://en.wikipedia.org/wiki/Sleep_paralysis}}. There is still the question of the means to reach this state. We can start with \textbf{Marsilio Ficino}'s understanding of the angelic hierarchy in Table~\ref{tab:AngelicHierarchy}.

\begin{table}[h]\small
\begin{tabularx}{\textwidth}{>{\itshape}rX}
\toprule
Seraphim &
Speculate on the order and providence of God\\
Cherubim &
Speculate on the essence and form of God\\
Thrones &
Speculate, but some descend to work\\
Dominions &
Architects who design what the rest execute\\
Virtues &
Execute, move the heavens, and concur for the working of miracles as God's instruments\\
Powers &
Watch that the order of divine governance is not interrupted and some of them descend to human things\\
Principalities &
Care for public affairs, nations, princes, magistrates\\
Archangels &
Direct the divine cult and look after sacred things\\
Angels &
Look after smaller affairs and take charge of individuals as their guardian angels\\\bottomrule
\end{tabularx}
\caption{The Angelic Hierarchy and their Functions}
\label{tab:AngelicHierarchy}
\end{table}
The angels are not ``seen" in a sensual way, but one can see them incorporeally by ascending through deeper understanding of the higher intellectual centre. By developing spiritual vision, one learns to see the forms or ideas animating the conditions of the sensual world.

The Guardian angel is the prototype of one's own being, or one's more perfect double. Through understanding the archangels and principalities, one learns to discern large groupings. For example, in the sensual realm, one sees people as simply individual beings. However, on a higher level one sees the roles of nations, ethnicities, clans, families, and so on. Yet, that is never static. At the level of the powers, one discerns how everything is manifested in time.

Ultimately, one experiences the world as a reflection of the order and providence of God. Hence, in previous posts, we suggested meditations on Being, the Moral Order, and so on.

Yet there are real dangers there for the Magician. If he has not properly prepared himself through the purification of his will and his intellect, he risks contacting bad angels.


\hfill



References:

H/T \textbf{Charlotte Louise} for alerting me to this Cabalistic Conclusion.

\emph{Eros and Magic in the Renaissance}, by Ioan P Couliano

\emph{Giordano Bruno and the Hermetic Tradition}, by Frances Yates

\emph{The Religion of Beauty in Women}, by Jefferson Fletcher



\flrightit{Posted on 2019-01-23 by Cologero }
