\section{Rest, See, Love, Praise}

\begin{quotex}
After this period God shall rest as on the seventh day, when He shall give us (who shall be the seventh day) rest in Himself... The seventh shall be our Sabbath, which shall be brought to a close, not by an evening, but by the Lord's day, as an eighth and eternal day, consecrated by the resurrection of Christ, and prefiguring the eternal repose not only of the spirit, but also of the body. \emph{There we shall rest and see, see and love, love and praise}. This is what shall be in the end without end. For what other end do we propose to ourselves than to attain to the kingdom of which there is no end?
\flright{\textsc{Saint Augustine}, \emph{City of God}}
\end{quotex}

Augustine represents the time from Eden to the General Resurrection as the days of the week, culminating in the eighth eternal day. In that day of rest, the soul will finally see, it will love what it sees, and love evokes praise.

\paragraph{The Path of Knowledge}

\begin{quotex}
Not all men are in the same condition, and all are not led or disposed to a knowledge of the truth in the same way. For some are brought to a knowledge of the truth by signs and miracles; others are brought more by wisdom. ``The Jews require signs, and the Greeks seek wisdom'' (1 Cor 1:22). And so the Lord, in order to show the path of salvation to all, willed both ways to be open, i.e., the way of signs and the way of wisdom, so that those who would not be brought to the path of salvation by the miracles of the Old and New Testaments, might be brought to a knowledge of the truth by the path of wisdom, as in the prophets and other books of Sacred Scripture.
\flright{\textsc{Thomas Aquinas}, \emph{Commentary on John}}

In this way then, from what has been said, we can understand the matter of this Gospel. For while the other Evangelists treat principally of the mysteries of the humanity of Christ, John, especially and above all, makes known the divinity of Christ in his Gospel.
\flright{\textit{Ibid.}}
\end{quotex}

\paragraph{Something rather than Nothing}

\begin{quotex}
Consider the universe... are we, now, to imagine that its maker first thought it out in detail... and that having thus appointed every item beforehand, he then set about the execution? Such designing was not even possible; how could the plan for a universe come to one that had never looked outward?... And all is made silently, since nothing had part in the making but Being and Idea --- a further reason why creation went without toil.
\flright{\textsc{Plotinus}, \emph{Enneads}}
\end{quotex}

These Ideas exist in the Mind of God, from all eternity, not in stages as a general contractor would construct a house. As Rene Guenon pointed out, some of these Ideas are possibilities of manifestation, i.e., they can and will participate in Being; moreover, by definition, they must needs become manifest. Nothing or the Void is not a possibility of manifestation. Nothingness is unthinkable. Hence, when God withdraws himself, leaving a void, that void will is filled up with creation.

\paragraph{The Meaning of Free Will}

The common misunderstanding of Free Will is the power to choose among alternatives. Since a Free Will is necessarily uncaused, this choice is usually conceived as arbitrary. For example, in one of \textbf{Jean Paul Sartre}'s novels, a character stabs himself in the hand in a café, just to demonstrate his freedom. This is because most people can only understand the \textbf{gnomic will}, by which a person comes to a decision after deliberation.

The \textbf{natural will}, on the other hand, arises from one's very nature. It is free because it is not compelled from some source beyond one's own nature. A fortiori, God's will arises from his very nature or essence. It is not compelled by any creature.

\paragraph{Essence and Personality}

Man is a composition of Being and Becoming, or non-being. He has an essence, eternally known to God. When born, this essence manifests in time and space, as a substance. In the course of life, he acquires accidents, say as represented in Aristotle's categories. These are contingent, in the sense that they are not self-caused but depend on various external circumstances. One's life task is to fully manifest one's essence. That means the development of a unifying principle, or Self, that unites the disported elements of the soul.

Augustine speculated on how these accidental qualities would manifest after the resurrection. To what extent would the resurrection body resemble the original body. Obviously, the new body would need to be recognizable to others, but it would need to be more perfect. Nevertheless, he speculated that the wounds of the martyrs would be visible, yet not ugly. Moreover, nothing of life should be lost.

\paragraph{Being and Hell}

\begin{quotex}
Hell is the state of the soul powerless to come out of itself, absolute self-centredness, dark and evil isolation, i.e. final inability to love.
\flright{\textsc{Nicolas Berdyaev}, \emph{The Destiny of Man}}
\end{quotex}

There was a pagan adage that asserted, ``It is better to die young, and best never to have been born at all.'' Augustine opposes this view with the revealed truth that creation is good, or, said another way, being is better than non-being. The notion of Hell seems to bother many people, so they concoct alternatives such as annihilationism or universal salvation.

Since being is good, a corollary is that being is hell is still better than non-being; this refutes annihilationism. The other objection is that an eternal hell is unjust and not fitting to the ideal of Christ. We know that the Son will come to judge the living and the dead. Note the word ``judge''; the Son acts as the impartial judge, not as an angry prosecutor. The person's whole life will be put on the balance, not any one specific act.

To deny the possibility of hell is to deny free will. As mentioned above, one's true will arises from one's very nature. That is what makes the judgment just; one's state of being is revealed in his thoughts and deeds. The notion that even the fallen angels will be redeemed is far-fetched. Unlike a material being, who combines being and non-being, a spiritual being cannot change. The decisions of the fallen angels were instantaneous and immutable. The attitude is that, ``It is better to reign in hell than serve in heaven.'' So, it is absurd to think that the demons will be coerced, against their will, to ``serve'' in heaven.

Moreover, ``eternity'' does not mean everlasting, or infinite duration in time. That is, eternity refers to ``quality'' and everlastingness to ``quantity''. It is the quality of postmortem experiences that matter, not the duration. This is consistent with the state of the soul.
\textbf{Valentin Tomberg} describes this difference:

\begin{quotex}
This subjective state of soul is neither long nor short---it is as intense as eternity is. Similarly, the blessedness that a saint experiences in the vision of God is as intense as eternity ---although it could not so last, since someone present at the ecstasy of a saint would time it as a few minutes. The ``region'' of eternity is that of intensity, which surpasses the measures of quantity that we employ in time and space. ``Eternity'' is not a duration of infinite length; it is the ``intensity of quality'' which, if compared with time and thus translated into the language of quantity, is comparable with an infinite duration.
\flright{\textit{Meditations on the Tarot}}
\end{quotex}

\paragraph{Union with God}

\begin{quotex}
No one has ever seen God: the only Son, who is in the bosom of the Father, he has made him known.
\flright{\textsc{John 1:18}}
\end{quotex}

The viewpoint of the Vedanta and the Neoplatonists is that the essence is absorbed into the Absolute without, however, the contingent or conditioned aspects of one's being. This was the possibility before Christ, since that absorption is not really a ``seeing'' of God. Hence, there is no love and no praise.

The \textbf{Beatific Vision}, on the other hand, retains the personality. Tomberg explains how:

\begin{quotex}
For this ``vision'' takes place in the domain of essence transcending all substance; it is not a fusion, but an \emph{encounter} in the domain of essence, in which the human personality (the consciousness of self) remains not only intact and without impediment, but also becomes ``that which it is'', i.e. becomes truly itself---such as the Thought of God has conceived it for all eternity.
\end{quotex}

The Beatific Vision is the union of essences, yet the two substances are separate. The individual personality persists.


\flrightit{Posted on 2020-04-19 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Samuel Fiebig on 2020-04-20 at 09:04 said:}

Absolutely beautiful. Thank you.

\hfill

\texttt{Matt on 2020-05-15 at 15:25 said:}

Are you using the term ``personality'' interchangeably with ``essence''. Maurice Nicoll, in his \textit{Psychological Commentaries}, distinguishes the two by stating that personality is the mask we put on in order to make our way through the contingencies of life and society. It is our external manifestation, and forms the meat around the seed of our essence. This allows some (maybe not many) to grow their essence later in life by feeding off the personality. So instead of dying as a seed, they become fully realized. How one's essence ``feeds'' and grows from one's personality is the part I'm not clear about in Nicoll's writing. In any case, I'm assuming this is not quite the meaning of personality that you're using. The more fully developed essence is that which persists, I would assume, which could be thought of as a union of one's inner essence with one's external personality.

\hfill

\texttt{Cologero on 2020-05-15 at 22:54 said:}

Yes, more like your final sentence. I was using it in Tomberg's sense as the last section of the post shows. Unfortunately, there is not a consistent philosophical language in English, and on top of that, we have to deal with translations from other languages like French or Russian; hence, context is always important. Your point, I believe, illustrates what we have been doing in our meetings. We are taking those negative aspects of the personality, which steal our psychic energies, and converting that energy into the development of essence. This is not different from using the awareness of sinfulness as motivation to a higher purpose. Tomberg's point is that nothing in our life is lost; this is unlike some eastern and similar teachings that strive for depersonalisation. Augustine addresses this issue in the City of God when he writes about the general resurrection: how does our earthly life correspond to the resurrection life?

\end{sffamily}
\end{footnotesize}

