\section{God and Life's Participation Trophy}

\begin{quotex}
Conversations about matters other than religion or art are always trivial and vain. \flright{\textsc{Joris-Karl Huysmans}}

\end{quotex}
\paragraph{Life's Participation Trophy}
\textbf{Aaron Rodgers}, who is a major star in American football, was raised in an evangelical family. Lately, he has caused some consternation through his public admission that he cannot believe in a God who would send people to everlasting hell. Of course, no one should consult an athlete on theological issues. So a theologian like David Bentley Hart came to a different conclusion: he believes in God but not in hell, if that sounds better to you.

It solves the problem of hell, but brings in a more subtle issue. If heaven is everyone's ultimate destination, doesn't that trivialize life? In American youth sports, saturated with egalitarianism, in lieu of giving out trophies to the best teams, every player gets a ``participation trophy". That seems to be the mindset of Rodgers and Hart\footnote{\url{https://www.youtube.com/watch?v=5QbsvsxFyDo}}. They can only believe in a God who gives out participation trophies for having lived.

\paragraph{Sudden Conversion}
Someone recently asked me if I had a conversion or a deeper understanding of life after my recent surgery. As though I'm going to appear on a TV show with tales of discovering God or finally understanding the meaning of life. No, I don't trust anyone who has such a conversion, although, I suppose, one conversion would be acceptable.

I've lived the life I want; I have no bucket list. I've gone on my adventures when I was young. I've met a lot of people, some rich and powerful, so I understand how they think and act. Not at all like you. Also many on different spiritual paths. In my work, I've learned about many commercial sectors: telecommunications, electronics, finance, manufacturing, banking, and so on.

Most readers here are young. Do what you want now, don't put it off.

\paragraph{True Esoterism}
Before you decide to embark on an esoteric way, consider seriously\textbf{ Boris Mouravieff}'s warning;

\begin{quotex}
The more man progresses on the Way, the more his feeling of being a stranger is intensified. Soon he will become boring: later still he will become unbearable, and finally, odious. That is why `the prophet is despised in his country, among his close relations and in his own house.' That indication is precise and leaves no room for doubt. He who wants to start esoteric studies is invited to think it over twice, and weigh it all, before he rushes to cross the moat-threshold. We repeat that it will not be possible for him to return to exterior life and to find his place, his pleasures and his satisfaction there as in the past.

\end{quotex}
On the other hand, there are benefits and you will make new connections:

\begin{quotex}
However, as well as the difficulties which are the first results of his evolution, such a man will receive comforting impressions, especially in his human relations. He will be surprised to perceive one day that certain faces which only yesterday appeared to him ordinary, today shine in his eyes with a bright beauty. It is because his sight, sharpened by esoteric work, has acquired the faculty of penetrating beyond the external crust. It is amongst these brighter beings that he will find his new friends. Their society will welcome him as one of their own. He will be understood among them, and their community of common aims and interests will be a stimulus and a help for all.

\end{quotex}
Read this short poem \textit{The Threshold}\footnote{\url{https://novaonline.nvcc.edu/eli/evans/his241/Notes/Turgenev_Threshold.html}} by Turgenev.

\paragraph{Toxic Esoterism}
There are dangers in pursuing esoterism, especially without a guide. Valentin Tomberg warns against specific dangers, some of which are described below. The first danger of misunderstanding is relying solely on intellectual arguments without the mystical experience to validate it. The second danger is to desire wonderful experiences. The third is to confuse one's own stream of consciousness with higher knowledge. The fourth is to presume to have reached the highest possible state of being.

\textbf{Misunderstanding}: There is no authentic and sincere religious life without faith, hope and love; but there is no faith, hope and love without mystical experience or, what is the same thing, without grace. No intellectual argument can awaken faith; what it can do, at best, is to eliminate obstacles, misunderstandings and prejudices, and thus help to establish the state of interior silence necessary for the experience of the divine breath. But faith itself is the divine breath whose origin is found neither in logical reasoning, nor in aesthetic impression, nor in human moral action.

\textbf{Spiritual drunkard}: Passing on to mysticism which has not given birth to gnosis, magic and Hermetic philosophy —such a mysticism must, sooner or later, necessarily degenerate into spiritual enjoyment" or ``intoxication". The mystic who wants only the experience of mystical states without understanding them, without drawing practical conclusions from them for life, and without wanting to be useful to others, who forgets everyone and everything in order to enjoy the mystical experience, can be compared to a spiritual drunkard.

\textbf{Charlatanism}: This is why anyone who confuses lack of concentration with concentration without effort, and streams of simple mental associations with the vision without effort of correspondences by analogy, will necessarily become a charlatan.

\textbf{Megalomania}: It is not thought as such which allows the desire for personal greatness or the tendency towards megalomania, but rather the will which makes use of the head and which can take hold of thought and reduce it to the role of its instrument. Mystical megalomania, where one deifies the regulating centre of one's own being, one's ego, and where one sees the divine only within oneself and becomes blind to the divine above and outside of oneself. The ``higher Self is then experienced as the supreme and unique Self of t h e world, although it is only higher in relation to the ordinary, empirical self, and it is fat from the supreme and unique being… far from being God.

\paragraph{Esoteric Writing}
There are a few different understandings of what esoteric writing means.

\textbf{Theosophical}: New Age systems based on Theosophy assume that esoteric writing is the ``real" teaching, so that the exoteric teaching can be ignored. The effect of this, in practice, is that the esoteric teaching just becomes a new dogma, replacing the exoteric dogma. That is the error of misunderstangind.

\textbf{Leo Strauss}: Strauss made the idea of esoteric writing acceptable in some academic circles. He understands it in a political sense. His premise is that philosophers faced dangers if they wrote too plainly about their views. Hence, they disguised their real intentions by writing in various ways (we can't go into those ways right now.) For example, Thomas Aquinas did not explicitly deny any exoteric teachings, but he wrote esoterically. His real views, seemingly quite different from those of simple believers, are in his philosophy.

Strauss subjected the writings of several philosophers to his theories. There are some Straussian tricks even on Gornahoor, such as obscure allusions. They will probably die with me.

\textbf{Rene Guenon}: Guenon claimed that the exoteric teachings include rites, rituals, dogmas, and moral teaching. However, a valid exoteric teaching must be fertilised with a living esoteric tradition. Unlike the multiplicity of exoteric teachings, there is but one esoteric teaching common to them all. Nevertheless, the esoteric teaching does not exist in a vacuum so that the esoterist must also follow the exoteric teaching.

\textbf{Valentin Tomberg}: Tomberg regarded the esoteric teaching as a deeper understanding of the exoteric teaching. Hence, he does not reject the exoteric teaching, but expresses them in deeper, more subtle, ways.

\paragraph{Negative Critiques}
There is no end of the gnashing of teeth and the wringing of hands. There is criticism of everything, if you are impressed by such things. Few of those professional criticisers, however, are able to provide a convincing positive alternative.

\paragraph{Atheism}
Atheism is not a defensible intellectual position, rather it is a moral flaw.



\flrightit{Posted on 2020-01-23 by Cologero }

\begin{center}* * *\end{center}

\begin{footnotesize}\begin{sffamily}



\texttt{Matt on 2020-01-25 at 19:27 said: }

``The more man progresses on the Way, the more his feeling of being a stranger is intensified."

With the small progress I have made, my experience has confirmed this statement by Mouravieff. Yet, I can't turn back from the path, not only because I can't just forget/ignore what I have read – and experienced – but the benefits of the online group are clear.

Thank you for pointing to that poem by Turgenev.

As for Hart and universal salvation, I take what he has written and said on it seriously – but I'm not exactly convinced. He is essentially giving what most in the modern West want to hear, just packaging it in theologically sophisticated language. There is a video on youtube of him speaking his case to an audience in an Episcopal church, rather than say, you know, a church within his claimed tradition (Eastern Orthodoxy); that is not a trivial fact. Of course, there are many in Orthodoxy – and the Catholic Church – that want to hear their universalist desires affirmed too, but it's telling that his platform is in an organization that ignores the teachings of the fathers, doctors, saints, and mystics of the Church and eats up whatever some contemporary theologian says on this or that. Hart, unlike his hosts, at least recognizes the importance of what the Fathers wrote, but the witnesses among them he has tried to marshal to his defense seem to be rather few. He will always assert there are ``numerous", and sure, you can pose further questions and point at evidence clearly to the contrary, but you'll just get the (in)famous Hart response: sighs and face palms. With all that said, I'll still read his work and listen to him speak.

On the topic of an ``eternal hell", it would be of value for both the universalists AND the super correct guardians of orthodoxy to read what Tomberg wrote on this matter in MOT. A David Bentley Hart would probably be open to reading the book; a Taylor Marshall would probably burn a copy if presented to him.


\hfill

\texttt{Michael on 2020-01-26 at 11:48 said: }

Perhaps when Hart thinks on Hell there are problems from where his thoughts arise from in regards with Time, Eternity, and the Human Being that bring him to such a conclusion.

Thirteen years ago an excerpt from Strauss' What is Political Philosophy began a journey in a way that was not unlike the Parable of the Coach. The effect it had turned slowly into a path continues today, revisiting this will be interesting.


\hfill

\texttt{Nicodemus on 2021-01-23 at 15:40 said: }

For what it's worth, Hart didn't choose the venue for the universalism lecture. It was part of a trip during which he also spoke on consciousness at the adjacent university. Both lectures were open to the public and attracted a wide array of audience members (academics, students, Christians of various stripes, a few esoterists, etc.)


\hfill

\texttt{Max on 2021-01-23 at 20:04 said: }

It is said here that there is one common esoteric teaching. The exoteric teaching tends to change with time and place though. In the ideal case they are compatible, which brings situations like those described by Guenon and Tomberg, with clear channels of communication open between them. Public religious and political controversies compete about getting to define what the exoteric teaching should be. When it changes quickly or suddenly, the connection is more difficult to maintain and becomes irregular. It is easy to see how ``also following the exoteric teaching", as Guénon says, might not really be a choice left to the individual at all. Having the choice would be an anomaly, perhaps afforded in periods of collapse. The question naturally presents itself, especially in turbulent times, exactly what the exoteric teaching says, or even if it will stay the same from one day to the next. If not, it is not as if the esoterist would necessarily be eager to madly pursue every new opinion of the day, in an effort to ``express it in a deeper subtler way". The returns are diminishing. Tomberg seems to write for a social situation which does not presently exist, even if that does not stop us from learning what we can from him. If a once exoteric teaching becomes unlawful or unsupported in the worldly envinronment, it can still be reabsorbed into the esoteric current.

Matt points out that the theologian Hart ``is essentially giving what most in the modern West want to hear, just packaging it in theologically sophisticated language." This phenomenon is of course nothing new, but has repeated itself many times before. The issue resembles that faced by the Amish, basically arresting progress at a certain point, which necessitates a measure of worldly power to do the same.


\end{sffamily}\end{footnotesize}
