\section{The True Horoscope}

\begin{quotationx}
Man alone is capable of communion with God. For to man alone among the living creatures does God speak: at night through
dreams, by day through the intellect. And He uses every means to foretell and prefigure the future blessings that will
be given to those worthy of Him. \flright{\textsc{Saint Antony the Great}}

The True is the good of the intellect. \flright{\textsc{Dante}}

The first stage of [esoterism] is attained when one enters conscious interaction with the beings of the angelic
hierarchy. [It] does not involve knowledge of universal laws, but entering conscious interaction with the beings who
know the mysteries of birth. The true horoscope will not be reached by a path of calculation but through a path of
interaction with suprasensory beings. What angels have imparted to humankind is the horoscope in the true sense. \flright{
\textsc{Valentin Tomberg}, \emph{Meditations on the Old Testament}}

\end{quotationx}

\paragraph{Free Will is the Greatest Gift}
\begin{quotex}
The greatest gift that in his largess God\newline
Creating made, and unto his own goodness\newline
Nearest conformed, and that which he doth prize\newline
Most highly, is the freedom of the will,\newline
Wherewith the creatures of intelligence\newline
Both all and only were and are endowed. \flright{\textsc{Dante}, \emph{Paradiso}}

\end{quotex}

Dante learns an interesting lesson on Mercury: Humanity used free will to fall into sin by Adam's rejection
of the Supreme Good. He took it on himself to know good and evil. and endured the just consequences of what was done
freely. He thereby cut humanity off from higher influences, which was then left to its own devices. The two obvious
solutions are unworkable.

\begin{enumerate}
\item One option is for humanity to save itself. That hardly seems possible since no political program, medical
procedure, pharmaceutical, eugenic practice, educational technique, etc., has proved efficacious. 
\item Or else, God could simply pardon Adam, but that would negate the whole point of free will. 
\end{enumerate}
A crime demands restitution. In this context, the crucifixion was paradoxically both a just penalty against
humanity's sin and also a crime against Christ.

\paragraph{Providence}
\begin{quotex}
A generated nature its own way\newline
Would always make like its progenitors,\newline
If Providence divine were not triumphant. 

\end{quotex}
On Venus, the discussion turned to Providence. Francesco da Buti, who wrote a commentary on the \emph{Divine Comedy} in
the 14th century, in which his lists both the negative and positive astrological influences of the planets. Regarding
them, he concludes:

\begin{quotex}
All are not found in one man, but a part in one, and a part in another, according to Divine Providence; and the wise man
adheres to the good, and overcomes the others. 
\end{quotex}

In other words, the planetary effects do not compel on their own, but all ultimately are subject to Divine Providence. A
fortiori, Providence even supersedes genetic and family influences. Otherwise, without being open to higher influences,
nothing could change in the world. Moreover, the wise man can pick and choose which influences to follow.

\paragraph{Vows}
Since free will is the greatest gift to humanity, so a vow, which is the voluntary suspension of free will, is a
sacrifice of great worth. The three traditional vows are obedience, poverty, and chastity. Valentin Tom\-berg describes
their esoteric meaning:

\begin{itemize}
\item The vow of \textbf{obedience} is the silencing of personal desires, emotions and imagination in the face of reason
and conscience; it is the primacy of the ideal as opposed to the apparent, the nation as opposed to the personal,
humanity as opposed to the nation, and God as opposed to humanity. It is the life of cosmic and human hierarchical
ordering. That which is above serves that which is below and that which is below obeys that which is above. 
\item The vow of \textbf{poverty} is the practice of inner emptiness, which is established as a consequence of the
silencing of personal desires, emotions and imagination so that the soul is capable of receiving from above the
revelation of the word, the life and the light. Poverty is perpetual active vigil and expectation before the eternal
sources of creativity; it is the soul awaiting what is new and unexpected; it is the aptitude for learning always and
everywhere. 
\item The vow of \textbf{chastity} means the resolution to live according to solar law. Although virtue is boring and
vice is disgusting, that which lives at the foundation of the heart is neither boring nor disgusting. The foundation of
the heart is love. The heart lives only when it loves. It is then like the sun. And chastity is the state of the human
being in which the heart, having become solar, is the centre of gravity. 
\end{itemize}

\paragraph{The True Horoscope}
\begin{quotex}
The power and motion of the holy spheres,\newline
As from the artisan the hammer's craft,\newline
Forth from the blessed motors must proceed.\newline
The heaven, which lights so manifold make fair,\newline
From the Intelligence profound, which turns it. 

\end{quotex}
The heavenly planets do not move on their own according to some fixed immutable physical law, but rather are moved by
the angelic hierarchy. Henry Wadsworth Longfellow, in his commentary, points out this correspondence:

\begin{quotationx}
The Intelligences, ruling and guiding the several heavens (receiving power from above, and distributing it downward,
taking their impression from God and stamping it like a seal upon the spheres below), according to Dionysius the
Areopagite are as follows:

\begin{itemize}[nosep]
\item
The Seraphim – Primum Mobile 

\item
The Cherubim – The Fixed Stars 

\item
The Thrones – Saturn 

\item
The Dominions – Jupiter 

\item
The Virtues – Mars 

\item
The Powers – The Sun 

\item
The Principalities – Venus 

\item
The Archangels – Mercury 

\item
The Angels – The Moon 
\end{itemize}
\end{quotationx}

Valentin Tomberg makes this perfectly clear:

\begin{quotex}
The spiritual world is not a world of laws, principles and ideas; it is a world of spiritual beings — human
souls, Angels, Archangels, Principalities, Powers, Virtues, Dominions, Thrones, Cherubim, Seraphim and the Holy
Trinity: the Holy Spirit, the Son and the Father. 

\end{quotex}

\paragraph{Takeaway}
In order to understand your horoscope, i.e., your place in the world, your nature and destiny, it is necessary to know
the spiritual beings active in your life. For example,

\begin{itemize}
\item Your Guardian Angel, who is an all too often neglected aspect of your life 
\item The archangels who influence your space: family, nation, etc. 
\item The principalities, the spirits of time (zeitgeist), who are guides through history 
\end{itemize}
This is a partial list. Moreover, there are malignant spirits simultaneously influencing your life. Generational spirits
may even attach themselves to families for years.

\textbf{Anne Catherine Emmerich} saw how these angelic influences are assigned at birth:

\begin{quotex}
I have seen that everyone receives two spirits at birth, one good, the other evil. The good one is heavenly by nature
and belongs to the lowest hierarchy; the evil one is not a devil, not yet in torments, though deprived of the vision of
God. 

\end{quotex}
Her vision leaves open the possibility that the evil one could even be redeemed.

\textbf{Saint Diadochos of Photiki} in the \emph{Philokalia} shows how they can act in two ways:

\begin{quotationx}
Spiritual knowledge teaches us that there are two kinds of evil spirits: some are more subtle, others more material in
nature. The more subtle demons attack the soul, while the others hold the flesh captive through their lascivious
enticements. Thus there is a complete contrast between the demons that attack the soul and those that attack the body,
even though they have the same propensity to inflict harm on mankind. When grace does not dwell in a man, they lurk
like serpents in the depths of the heart, never allowing the soul to aspire towards God. But when grace is hidden in
the intellect, they then move like dark clouds through the different parts of the heart, taking the form of sinful
passions or of all kinds of day-dreams, thus distracting the intellect from the remembrance of God and cutting it off
from grace. 

\end{quotationx}
\textbf{Ultimate control is always under Divine Providence.}

\begin{center}\scshape Deus vult\end{center}

\flright{\itshape Posted on 2020-11-11 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Patricia Kay on 2022-11-11 at 16:40 said: }

Would you give a brief description of these intelligences? I'm not sure what “principalities” or “thrones”
or “dominions” are?

\hfill

\texttt{Cologero on 2022-11-11 at 16:51 said: }

Tomberg recognizes that few people are able to experience the Divine Light in its fullness. The task of the angelic
hierarchy, therefore, is to attenuate the Light in a manner appropriate to each one as it descends through the
hierarchy. The esoteric path then is to reverse course and ascend through the hierarchy. To quote:

Each lower rank of hierarchy is a “moon” in relationship to the “sun” of the rank above it.

The angels transmit the tumultuous, strong impulses of the archangels in a bearable form, suited to human individuals,
i.e., in the form of the soft light of moral clarity.

The archangels adopt the radical, valid-to-all-mankind commandments and prohibitions of the principalities (archai) to
suit the special characters and capacities of the various peoples, thereby protecting them from becoming
over-pressured.

And something similar is effected by the principalities in relation to the powers (exusiai), the powers toward the
virtues (dynamis), the virtues toward the dominions (kryriotetes), the dominions toward the thrones, the thrones toward
the cherubim, the cherubim toward the seraphim, and the seraphim toward the eternal Trinity of God.

See also:\newline
Angelic States of Being\newline
Death by a Kiss

\hfill
\end{sffamily}
\end{footnotesize}
