\section{Solemnity of Mary}

The Feminine is the esoteric and the esoteric is Feminine. It is the part of the Self that is most hidden. Conversely, the Masculine is exoteric, the Law. The Virginal Conception means that Christianity originates in the purely esoteric. Without a male father, then, Jesus does not bring a new Law, which would be exoteric, but rather fulfills the Law.

Nevertheless, the understanding of salvation has over-emphasized the legalistic aspect. Something is done ``for the soul'', from the outside as it were. The point is then lost. This is not the place to rehash the theological conundrums associated with this view.

The esoteric purpose is to restore the Primordial State, that is, a return to the consciousness of Adam before the Fall. Amazingly, this undoes the consequences of one's own acts, it is like getting a Mulligan\footnote{\url{https://en.wikipedia.org/wiki/Mulligan\_(games)}}. The curses of toil, suffering and death are reversed, and replaced by even greater things.

The virginal conception and birth are not just physical events, they are spiritual events in human history. This is not an allegorical or metaphoric truth, but rather the literal truth.

Hail Holy Queen, Mother of Mercy, our Life our Sweetness and our Hope.


\flrightit{Posted on 2021-01-01 by Cologero}

\begin{center}* * *\end{center}

\begin{footnotesize}
\begin{sffamily}
\texttt{Cologero on 2021-01-01 at 13:26 said:}

This was intended as part of a more comprehensive upcoming post, but I thought it appropriate to pull this part out today. Sources and further reflections are coming.

\end{sffamily}
\end{footnotesize}

