\section{Axioms of the Will}

\begin{enumerate}[label=\Roman*.]

\item Nothing resists the will of man when he knows the truth and wills the good.

\textit{Commentary:}~The Truth is the Idea identical with Being. Will means altering Being in conformance with the Idea. The True Idea is. To will the good is to will the Idea.

\item To will evil is to will death. A perverse will is the beginning of suicide.

\textit{Commentary:}~Evil is disorder, chaos, formlessness, multiplicity. To will evil, then, is to will dissolution.

\item To will good with violence is to will evil, for violence produces disorder and disorder produces evil.

\textit{Commentary:}~The Will of God is actionless action. Violence is its opposite and is destructive of order.

\item One can, and one should, accept evil as the means of good; but one must never will it or do it, otherwise one would destroy with one hand what one builds with the other. Good faith never justifies bad means; it corrects them when one undergoes them, and condemns them when one takes them.

\textit{Commentary:}~Evil increases the virtue of the wise but corrupts the weak.

\item To have the right to possess always, one must will patiently and long.

\textit{Commentary:}~Man must know how to wait.

\item To pass one's life in willing what it is impossible to possess always, is to abdicate life and accept the eternity of death.

\textit{Commentary:}~To will the transient, the ephemeral, the impermanent is to will death.

\item The more obstacles the will surmounts, the stronger it is. It is for this reason that Christ glorified poverty and sorrow.

\textit{Commentary:}~A soft life makes man weak. The Will needs opposition and friction.

\item When the will is vowed to the absurd, it is reproved by eternal reason.

\textit{Commentary:}~The absurd cannot be willed for long, because Reality will punish it.

\item The will of the just man is the will of God himself, and the law of Nature.

\textit{Commentary:}~The will of God is Providence and is free. The law of Nature is Destiny and is determined. The Will of man is the mediating term. For the just man --- or True Man --- the three terms coincide.

\item It is by the will that the intelligence sees. If the will is healthy, the sight is just. God said: ``Let there be light!'' and light is; the will says: ``Let the world be as I will to see it!'' and the intelligence sees it as the will has willed. This is the meaning of the word, ``So be it,'' which confirms acts of faith.

\textit{Commentary:}~The world as Will and Representation ... only by seeing that I have willed my life and my world, can I then understand it.

\item When one creates phantoms for oneself, one puts vampires into the world, and one must nourish these children of a voluntary nightmare with one's blood, one's life, one's intelligence, and one's reason, without ever satisfying them.

\textit{Commentary:}~It is man's idiosyncrasy to create problems for himself in order to resolve them.

\item To affirm and to will what ought to be is to create; to affirm and will what ought not to be, is to destroy.

\item Light is an electric fire put by Nature at the service of the will; it lights those who know how to use it, it burns those who abuse it.

\item The empire of the world is the empire of the light.

\item Great intellects whose wills are badly balanced are like comets which are aborted suns.

\item To do nothing is as fatal as to do evil, but it is more cowardly. The most unpardonable of mortal sins is inertia.

\item To suffer is to work. A great sorrow suffered is a progress accomplished. Those who suffer much live more than those who do not suffer.

\item Voluntary death from devotion is not suicide; it is the apotheosis of the will.

\item Fear is nothing but idleness of the will, and for that reason public opinion scourges cowards.

\item Succeed in not fearing the lion, and the lion will fear you. Say to sorrow: ``I will that you be a pleasure, more even than a pleasure, a happiness.''

\item A chain of iron is easier to break than a chain of flowers.

\item Before saying that a man is happy or unhappy, find out what the direction of his will has made of him: Tiberius died every day at Capri, while Jesus proved his immortality and even his divinity on Calvary and upon the Cross.
\end{enumerate}

From \textit{The Key of the Mysteries} by Eliphas Levi.

Commentary added by editor.

Originally published on medtarot, 5 May 2011.


\flrightit{Posted on 2023-02-21 by Cologero}

